### 2017-11-06 analysis/paper/sim_survey_paper.Rmd
Add abstract and introduction crafted for a previous itteration of this analysis

---
title: Using simulation to optimize the sampling design of fisheries-independent trawl surveys
author: "Paul M. Regular, Fran Mowbray"
date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
affiliation: "This is a test"
output: 
  word_document:
    reference_docx: template.docx
bibliography: references.bib
csl: ices-journal-of-marine-science.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

A central question to fish stock assessments is whether enough samples were collected to adequately represent the population being monitored. For assessments based on trawl surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling. We therefore need to assess the impact of haul design and length and age sampling on survey accuracy to determine optimal sampling effort. In this study we used simulations to try to answer this question. The method draws repeated stratified-random samples from a spatially correlated field of simulated fish; these samples were then analyzed and estimates of abundance at age were compared against a known truth. By varying the sampling protocol, we demonstrate that increasing the number of sets conducted per stratum provides the greatest gains in precision when sampling a clustered population. Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent. These results indicate that length and age sampling efforts for cod (*Gadus morhua*) in the trawl survey conducted by Fisheries and Oceans Canada, Newfoundland region, can be scaled back without significant losses to precision. Time saved by reducing said sampling may afford more time to conduct more sets and/or increase sampling efforts of other species. While this simulation is tailored to a specific case, the method may be adapted to different systems and survey procedures.

**Keywords** - length and age distribution; simulation; spatial correlation; stratified analysis; statistical assessment models; survey design

# Introduction

Fisheries-independent trawl surveys have become an increasingly important research tool for the management of dynamic fish stocks. These surveys provide reliable indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998]. While costly to obtain, this information forms the basis of of many stock assessments models used throughout the world. Effective and informed management decisions therefore require surveys that maximize information while minimizing the expense of data collection. Determining the number of samples required to adequately characterize a fish population is particularly challenging since data tend to be collected across one or more stages [@Aanes2015]. Optimal survey design is further complicated by the fact that trawls sample clusters of fish that tend to have similar characteristics, such as length, age and stomach contents. This phenomenon results in samples with positive intra-haul correlation, which can markedly reduce the effective sample size for estimating length and age frequencies of the target population [@Pennington1994; @Pennington2002; @Stewart2014]. The complexities of multi-stage sampling and intra-haul correlation make it difficult to answer the question *"how much sampling is enough?"*

Two main approaches have been applied to try and answer this question: one involves statistical estimators of effective sample size [@Pennington2002; @Stewart2014], and the other employs resampling methods to test alternate sampling protocol [@Cervino2006; @Zhang2013]. Both approaches have shown that lengths are usually oversampled and a reduction in said sampling can often be done without significant loss of accuracy [@Pennington2002]. The same result holds for other positively correlated characteristics, such as age [@Coggins2013] and stomach contents [@Bogstad1995]. Pennington and Vølstad [-@Pennington1994] concluded that the best way to improve the accuracy of surveys with strong intra-haul correlation is to take fewer samples per sampling unit and increase the total number of sampling locations. Few studies, however, have tested these conclusions using simulations. A key advantage of a pure simulation approach is that "true" population size and all stages of sampling can be controlled. The relative effects of each stage of sampling on our perception of the "truth" can then be tested. This level of testing is of praticular importance for age composition estimates, and the contemporary age-based assessment models that rely on these estimates, as they are affected by all stages of sampling. A key challenge, of course, is building a simulation that emulates reality.

The objective of this paper is to use a realistic simulation to evaluate the efficacy of various sampling protocol. First we simulated a spatially autocorrelated target population, and then we simulated surveys over this field. Set density was varied across surveys, as was length and age sampling efforts. Estimates of abundance at age were then obtained using a standard stratified approach [@Smith1981] and deviation from the truth was assessed. Given previous findings, we predict that the greatest gains in precision will come from increasing the number of hauls; increased sampling of correlated variables (length and age) on a set by set basis will be relatively ineffective. The bottom trawl survey of cod (*Gadus morhua*) in NAFO division 3Ps, Newfoundland, was used as a case study for this analysis.

# References

### 2017-12-21 inst/rmd/vis_sim.Rmd
Add function for visualizing simulation results

---
title: "Simulation Explorer"
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r global, include=FALSE}

library(plotly)
library(viridis)
library(shiny)
e <- new.env() # make a new environment to keep the global clean
e$sim <- temp  # temp object (in global) created by vis_sim function
rm(temp)       # clean-up
e$ages <- e$sim$ages
e$years <- e$sim$years
e$vis_distribution <- "sp_N" %in% names(e$sim)

```

Abundance
================================================================================

Row
--------------------------------------------------------------------------------

### Total abundance

```{r}
e$tot <- data.frame(Year = e$years, N = colSums(e$sim$N))
e$tot$text <- paste0("Year: ", e$tot$year, "\nN: ", round(e$tot$N))
plot_ly(x = ~Year, y = ~N, text = ~text, hoverinfo = "text",
        type = "scatter", mode = "lines", data = e$tot) %>%
  layout(xaxis = list(rangeslider = list(type = "date")))
```

Row
--------------------------------------------------------------------------------

### Abundance at age

```{r}
e$d <- as.data.frame.table(e$sim$N, responseName = "N")
names(e$d) <- c("Age", "Year", "N")
e$d$text <- paste0("Age: ", e$d$Age, "\nYear: ", e$d$Year, "\nN: ", round(e$d$N))
plot_ly(x = ~Age, y = ~Year, z = ~N, color = ~Age,
        text = ~text, hoverinfo = "text",
        colors = viridis(length(e$ages)),
        type = "scatter3d", mode = "lines",
        line = list(width = 8),
        data = e$d)
```

### Total mortality

```{r}
plot_ly(x = e$ages, y = e$years, z = e$sim$Z, type = "surface") %>% 
  layout(
    scene = list(
      xaxis = list(title = "Age"),
      yaxis = list(title = "Year"),
      zaxis = list(title = "Z")
    ))
```

Distribution { `r if(!e$vis_distribution) ".hidden"`}
================================================================================

Column {.sidebar}
--------------------------------------------------------------------------------

```{r}

sliderInput("age", "Age", min = min(e$ages), max = max(e$ages), 
            value = 1, step = 1)
sliderInput("year", "Year", min = min(e$years), max = max(e$years), 
            value = 1, step = 1)
selectInput("type", "Type", choices = c("surface", "heatmap", "contour"), 
            selected = "surface")
checkboxInput("freez", label = "Free z limits?", value = FALSE)

```

Column
--------------------------------------------------------------------------------

### Distribution

```{r, eval=e$vis_distribution}

e$d <- e$sim$sp_N
e$x <- sort(unique(e$d$x))
e$y <- sort(unique(e$d$y))
e$p <- plot_ly(x = e$x, y = e$y)

renderPlotly({
  e$z <- t(xtabs(N ~ x + y, subset = age == input$age & year == input$year, data = e$d))
  if (input$freez) {
    e$r <- range(e$z)
  } else {
    e$r <- range(e$d$N)
  }
  add_trace(e$p, z = e$z, type = input$type, zauto = input$freez, zmin = min(e$d$N), zmax = max(e$d$N)) %>%
    layout(autosize = TRUE,
           scene = list(zaxis = list(title = "N", range = e$r)))
})

```

### 2017-12-21 R/vis_sim.R

@@ -1,0 +1,25 @@
+#' Make a flexdashboard for visualizing population simulation
+#'
+#' Assumes the working directory is the project directory
+#'
+#' @param sim            Object produced by \code{\link{sim_abundance}} or \code{\link{sim_distribution}}
+#' @param ...            Additional arguments to send to \link[rmarkdown]{run}
+#'
+#' @export
+#'
+
+vis_sim <- function(sim, ...) {
+
+  pkg <- c("rmarkdown", "shiny", "flexdashboard", "plotly", "viridis")
+  for (p in pkg) {
+    if (!requireNamespace(p, quietly = TRUE)) {
+      stop(paste(p, "is needed for vis_sim to work. Please install it."), call. = FALSE)
+    }
+  }
+
+  ## add a temp object to the global environment, otherwise rmarkdown
+  ## can't see these data
+  assign("temp", sim, envir = globalenv())
+  rmarkdown::run(file = "inst/rmd/vis_sim.Rmd", ...)
+
+}

### 2017-12-21 tests/informal_tests.R

@@ -1,75 +1,17 @@
 
-library(ggplot2)
-library(viridis)
 library(SimSurvey)
-library(plotly)
-library(shiny)
 
+sim <- sim_abundance(years = 1:100)
+vis_sim(sim)
 
-sp_pop <- sim_distribution(pop = sim_abundance(),
-                           grid = sim_grid(res = c(3.5, 3.5)),
-                           space_covar = sim_sp_covar(range = 50, sd = 0.1),
-                           ay_covar = sim_ay_covar(sd = 10,
-                                                   phi_age = 0.5,
-                                                   phi_year = 0.5),
-                           depth_par = sim_parabola(alpha = 0, sigma = 50))
-# sp_pop <- sim_distribution()
-sp_pop$N
-d <- sp_pop$sp_N
-ages <- sp_pop$ages
-years <- sp_pop$years
-head(d)
-
-# p <- ggplot(d, aes(x = x, y = y)) +
-#   facet_grid(as.numeric(age) ~ as.numeric(year)) +
-#   scale_fill_viridis() + theme_void() +
-#   theme(panel.spacing = unit(-0.1, "lines"))
-# p <- p + geom_tile(aes(fill = prob))
-# # p + geom_tile(aes(fill = N))
-#
-# ggplot(d[sample(seq(nrow(d)), 10000), ]) + geom_point(aes(x = depth, y = N))
-
-
-x <- sort(unique(d$x))
-y <- sort(unique(d$y))
-p <- plot_ly(x = x, y = y)
-
-ui <- fluidPage(
-  headerPanel("Distribution Explorer"),
-  sidebarPanel(
-    sliderInput("age", "Age", min = min(ages), max = max(ages),
-                value = 1, step = 1),
-    sliderInput("year", "Year", min = min(years), max = max(years),
-                value = 1, step = 1),
-    selectInput("type", "Type", choices = c("surface", "heatmap", "contour"), selected = "surface"),
-    checkboxInput("freez", label = "Free z limits?", value = FALSE)
-  ),
-  mainPanel(
-    plotlyOutput("surface", height = "900px", width = "900px")
-  )
-)
-
-server <- function(input, output) {
-  output$surface <- renderPlotly({
-    z <- t(xtabs(N ~ x + y, subset = age == input$age & year == input$year, data = d))
-    if (input$freez) {
-      r <- range(z)
-    } else {
-      r <- range(d$N)
-    }
[…]

### 2017-12-21 inst/rmd/vis_sim.Rmd
Fix hover text label for total abundance plot in vis_sim.Rmd

@@
 library(shiny)
-e <- new.env() # make a new environment to keep the global clean
+e <- new.env() # make a new environment to keep the global clean (rmarkdown::run wants to work in the global)
 e$sim <- temp  # temp object (in global) created by vis_sim function
@@
 e$tot <- data.frame(Year = e$years, N = colSums(e$sim$N))
-e$tot$text <- paste0("Year: ", e$tot$year, "\nN: ", round(e$tot$N))
+e$tot$text <- paste0("Year: ", e$tot$Year, "\nN: ", round(e$tot$N))
 plot_ly(x = ~Year, y = ~N, text = ~text, hoverinfo = "text",

### 2018-05-17 inst/rmd/vis_sim.Rmd
Minor clean-up during review

@@
 ```{r}
-e$d <- as.data.frame.table(e$sim$N, responseName = "N")
-names(e$d) <- c("Age", "Year", "N")
-e$d$text <- paste0("Age: ", e$d$Age, "\nYear: ", e$d$Year, "\nN: ", round(e$d$N))
-plot_ly(x = ~Age, y = ~Year, z = ~N, color = ~Age,
-        text = ~text, hoverinfo = "text",
-        colors = viridis(length(e$ages)),
-        type = "scatter3d", mode = "lines",
-        line = list(width = 8),
-        data = e$d)
+
+plot_ly(x = e$ages, y = e$years, z = e$sim$N, type = "surface") %>% 
+  layout(
+    scene = list(
+      xaxis = list(title = "Age"),
+      yaxis = list(title = "Year"),
+      zaxis = list(title = "N")
+    ))
+
 ```

### 2018-05-17 R/sim_dist.R

@@ -127,7 +127,7 @@ sim_parabola <- function(alpha = 0, mu = 250, sigma = 50, plot = FALSE) {
 #' applies correlated space, age and year error to simulate the distribution
 #' of the population.
 #'
-#' @param pop         An abundance at age matrix like one produced by \code{\link{sim_abundance}})
+#' @param pop         An abundance at age matrix like one produced by \code{\link{sim_abundance}}
 #' @param grid        A raster object defining the survey grid, like \code{\link{survey_grid}}
 #'                    or one produced by \code{\link{sim_grid}}
 #' @param space_covar Closure for simulating spatial covariance

### 2018-06-01 inst/rmd/vis_sim.Rmd
Add some plotting functions and fix up vis_sim flexdashboard

@@
 output: flexdashboard::flex_dashboard
-runtime: shiny
+always_allow_html: yes
 ---
 
-```{r global, include=FALSE}
-
+```{r setup, echo=FALSE}
 library(plotly)
 library(viridis)
-library(shiny)
-e <- new.env() # make a new environment to keep the global clean (rmarkdown::run wants to work in the global)
-e$sim <- temp  # temp object (in global) created by vis_sim function
-rm(temp)       # clean-up
-e$ages <- e$sim$ages
-e$years <- e$sim$years
-e$vis_distribution <- "sp_N" %in% names(e$sim)
+knitr::opts_knit$set(root.dir = tmp$output_dir)
+knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8)
+```
 
+```{r data, include=FALSE}
+sim <- tmp$sim
 ```
@@
 ```{r}
-e$tot <- data.frame(Year = e$years, N = colSums(e$sim$N))
-e$tot$text <- paste0("Year: ", e$tot$Year, "\nN: ", round(e$tot$N))
-plot_ly(x = ~Year, y = ~N, text = ~text, hoverinfo = "text",
-        type = "scatter", mode = "lines", data = e$tot) %>%
-  layout(xaxis = list(rangeslider = list(type = "date")))
+plot_trend(sim)
 ```
@@
 ```{r}
-
-plot_ly(x = e$ages, y = e$years, z = e$sim$N, type = "surface") %>% 
-  layout(
-    scene = list(
-      xaxis = list(title = "Age"),
-      yaxis = list(title = "Year"),
-      zaxis = list(title = "N")
-    ))
-
+plot_surface(sim, mat = "N")
 ```
@@
 ```{r}
-plot_ly(x = e$ages, y = e$years, z = e$sim$Z, type = "surface") %>% 
-  layout(
-    scene = list(
-      xaxis = list(title = "Age"),
-      yaxis = list(title = "Year"),
-      zaxis = list(title = "Z")
-    ))
+plot_surface(sim, mat = "Z")
 ```
 
-Distribution { `r if(!e$vis_distribution) ".hidden"`}
+Distribution { `r if(!"sp_N" %in% names(sim)) ".hidden"`}
 ================================================================================
@@
 ```{r}
-
-sliderInput("age", "Age", min = min(e$ages), max = max(e$ages), 
+sliderInput("age", "Age", min = min(sim$ages), max = max(sim$ages), 
             value = 1, step = 1)
-sliderInput("year", "Year", min = min(e$years), max = max(e$years), 
+sliderInput("year", "Year", min = min(sim$years), max = max(sim$years), 
             value = 1, step = 1)
-selectInput("type", "Type", choices = c("surface", "heatmap", "contour"), 
-            selected = "surface")
-checkboxInput("freez", label = "Free z limits?", value = FALSE)
-
+selectInput("type", "Type", choices = c("heatmap", "contour"), 
+            selected = "contour")
+selectInput("scale", "Scale", choices = c("natural", "log"), 
+            selected = "natural")
 ```
@@
 
-```{r, eval=e$vis_distribution}
-
-e$d <- e$sim$sp_N
-e$x <- sort(unique(e$d$x))
-e$y <- sort(unique(e$d$y))
-e$p <- plot_ly(x = e$x, y = e$y)
-
+```{r, eval = "sp_N" %in% names(sim)}
 renderPlotly({
-  e$z <- t(xtabs(N ~ x + y, subset = age == input$age & year == input$year, data = e$d))
-  if (input$freez) {
-    e$r <- range(e$z)
-  } else {
-    e$r <- range(e$d$N)
-  }
-  add_trace(e$p, z = e$z, type = input$type, zauto = input$freez, zmin = min(e$d$N), zmax = max(e$d$N)) %>%
-    layout(autosize = TRUE,
-           scene = list(zaxis = list(title = "N", range = e$r)))
+  plot_distribution(sim, ages = input$age, year = input$year, type = input$type,
+                    scale = input$scale)
 })
-
 ```

### 2018-06-01 R/plot_sim.R

@@ -1,19 +1,71 @@
-plot_sim <- function(gird, zcol = NULL,
-                     ramp = colorRampPalette(c("#FFFFFF", "#2166AC"))) {
-  op <- par()
-  layout(t(1:2), widths = c(0.9, 0.1))
-  b <- seq(min(grid[[zcol]], na.rm = TRUE),
-           max(grid[[zcol]], na.rm = TRUE),
-           length.out = 225)
-  col_ramp <- ramp(length(b))
-  cols <- col_ramp[findInterval(grid[[zcol]], b)]
-  par(mar = c(1, 1, 3, 0))
-  plot(grid, col = cols, border = NA)
-  box()
-  par(mar = c(1, 0, 3, 3))
-  image(1, b, t(seq_along(b)), col = col_ramp, axes = FALSE,
-        main = zcol)
-  axis(4, las = 2)
-  box()
-  on.exit(par(mar = op$mar, mfcol = c(1, 1)))
+
+#' Simple plotting functions
+#'
+#' @description These functions are simple plotting helpers to get some quick
+#' visuals of values produced by \code{\link{sim_abundance}},
+#' \code{\link{sim_distribution}}, etc.
+#'
+#' @param sim    Object returned by \code{\link{sim_abundance}},
+#'               \code{\link{sim_distribution}}, etc.
+#' @param mat    Name of matrix in \code{sim} list to plot.
+#' @param ages   Subset data to one or more ages.
+#' @param years  Subset data to one or more years.
+#' @param scale  Plot response on "natural" or "log" scale?
+#' @param ...    Additional arguments to pass to \code{\link{plotly::plot_ly}}.
+#'
+#' @import plotly
+#'
+#' @export
+#' @rdname plot_trend
+plot_trend <- function(sim, ...) {
+  tot <- data.frame(Year = sim$years, N = colSums(sim$N))
+  tot$text <- paste0("Year: ", tot$Year, "\nN: ", round(tot$N))
+  plot_ly(x = ~Year, y = ~N, text = ~text, hoverinfo = "text",
+          type = "scatter", mode = "lines", data = tot, ...)
 }
+
+#' @export
+#' @rdname plot_trend
+plot_surface <- function(sim, mat = "N", ...) {
+  plot_ly(x = sim$ages, y = sim$years, z = sim[[mat]], type = "surface",
+          ...) %>%
+    layout(
+      scene = list(
[…]

### 2018-06-01 R/vis_sim.R

@@ -1,8 +1,9 @@
-#' Make a flexdashboard for visualizing population simulation
+#' Make a flexdashboard for visualizing the simulation
 #'
 #' Assumes the working directory is the project directory
 #'
-#' @param sim            Object produced by \code{\link{sim_abundance}} or \code{\link{sim_distribution}}
+#' @param sim            Object produced by \code{\link{sim_abundance}}, \code{\link{sim_distribution}},
+#'                       \code{\link{sim_survey}} or \code{\link{test_surveys}}.
 #' @param ...            Additional arguments to send to \link[rmarkdown]{run}
 #'
 #' @export
@@ -13,13 +14,19 @@ vis_sim <- function(sim, ...) {
   pkg <- c("rmarkdown", "shiny", "flexdashboard", "plotly", "viridis")
   for (p in pkg) {
     if (!requireNamespace(p, quietly = TRUE)) {
-      stop(paste(p, "is needed for vis_sim to work. Please install it."), call. = FALSE)
+      stop(paste(p, "is needed for vis_fit to work. Please install it."), call. = FALSE)
     }
   }
 
-  ## add a temp object to the global environment, otherwise rmarkdown
-  ## can't see these data
-  assign("temp", sim, envir = globalenv())
-  rmarkdown::run(file = "inst/rmd/vis_sim.Rmd", ...)
+  ## make a tmp object in the global environment for use by rmarkdown
+  ## (rmarkdown::run likes to operate on objects in the global environment)
+  tmp <- as.list(environment())
+  assign("tmp", tmp, globalenv())
+
+  ## keep most objects created in the rmd file local (i.e. not in the global)
+  local({
+    rmarkdown::run(file = "inst/rmd/vis_sim.Rmd")
+  })
 
 }
+

### 2018-06-03 inst/rmd/vis_sim.Rmd
Allow years and ages to have equal spatial distribution and revise cod simulation parameters

@@
 library(viridis)
-knitr::opts_knit$set(root.dir = tmp$output_dir)
 knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8)
@@
 })
+
 ```

### 2018-06-03 R/sim_dist.R

@@ -12,14 +12,16 @@
 #'
 #' @description These functions return a function to use inside \code{\link{sim_distribution}}.
 #'
-#' @param range    Decorrelation range
-#' @param lambda   Controls the degree of smoothness of Matern covariance process
-#' @param sd       Variance
-#' @param phi_age  Defines autocorrelation through ages. Can be one value or a vector of the same
-#'                 length as ages
-#' @param phi_year Defines autocorrelation through years. Can be one value or a vector of the same
-#'                 length as years
-#' @param model    String indicating either "exponential" or "matern" as the correlation function
+#' @param range       Decorrelation range
+#' @param lambda      Controls the degree of smoothness of Matern covariance process
+#' @param sd          Variance (can be age specific in \code{sim_ay_covar}).
+#' @param phi_age     Defines autocorrelation through ages. Can be one value or a vector of the same
+#'                    length as ages
+#' @param phi_year    Defines autocorrelation through years. Can be one value or a vector of the same
+#'                    length as years
+#' @param group_ages  Make space-age-year variance equal across these ages
+#' @param group_years Make space-age-year variance equal across these years
+#' @param model       String indicating either "exponential" or "matern" as the correlation function
 #'
 #' @rdname sim_sp_covar
 #' @export
@@ -45,7 +47,8 @@ sim_sp_covar <- function(range = 40, lambda = 1, sd = 0.1, model = "matern") {
 
 #' @rdname sim_sp_covar
 #' @export
-sim_ay_covar <- function(sd = 10, phi_age = 0.5, phi_year = 0.5) {
+sim_ay_covar <- function(sd = 10, phi_age = 0.5, phi_year = 0.5,
+                         group_ages = NULL, group_years = NULL) {
   function(ages = NULL, years = NULL, cells = NULL, w = NULL) {
 
     # Simple description of covariance:
@@ -56,6 +59,13 @@ sim_ay_covar <- function(sd = 10, phi_age = 0.5, phi_year = 0.5) {
     na <- length(ages)
     ny <- length(years)
     nc <- length(cells)
+    if (length(sd) == 1) {
+      sd <- rep(sd, na)
[…]

### 2018-06-03 analysis/cod_sim.R

@@ -26,8 +26,9 @@ pop <- sim_abundance(ages = 1:20,
                    space_covar = sim_sp_covar(range = 40,
                                               sd = 0.1),
                    ay_covar = sim_ay_covar(sd = 10,
-                                           phi_age = c(rep(0.1, 3), rep(0.9, 17)),
-                                           phi_year = 0.7),
+                                           phi_age = 0.1,
+                                           phi_year = 0.8,
+                                           group_ages = 5:20),
                    depth_par = sim_parabola(mu = 250,
                                             sigma = 50))
 

### 2018-06-03 analysis/imitate_cod_data.R

@@ -88,22 +88,52 @@ af$age <- as.integer(gsub("af", "", af$age))
 ## - Abundance parameters and catchability curve roughly based on NCAM estimates
 ## - Distribution parameters manually tweaked until results roughly corresponded to
 ##   observations from 3Ps cod
-abundance <- sim_abundance(ages = 1:20, years = 1:20,
-                           R = sim_R(mean = 100000000, log_sd = 0.5,
-                                     random_walk = TRUE),
-                           Z = sim_Z(mean = 0.5, log_sd = 0.2,
-                                     phi_age = 0.9, phi_year = 0.5))
-grid <- sim_grid(x_range = c(-140, 140), y_range = c(-140, 140), res = c(3.5, 3.5),
-                 shelf_depth = 200, shelf_width = 100, depth_range = c(0, 1000),
-                 n_div = 1, strat_breaks = seq(0, 1000, by = 20), strat_splits = 2)
-distribution <- sim_distribution(abundance, grid = grid,
-                                 space_covar = sim_sp_covar(range = 40, sd = 0.1),
-                                 ay_covar = sim_ay_covar(sd = 10,
-                                                         phi_age = c(rep(0.1, 3), rep(0.9, 17)),
-                                                         phi_year = 0.7),
-                                 depth_par = sim_parabola(mu = 250, sigma = 50))
-survey <- sim_survey(distribution, n_sims = 1, light = FALSE,
-                     set_den = 3 / 1000, lengths_cap = 400, ages_cap = 10,
+set.seed(438)
+pop <- sim_abundance(ages = 1:20,
+                     years = 1:20,
+                     R = sim_R(mean = 100000000,
+                               log_sd = 0.5,
+                               random_walk = TRUE),
+                     Z = sim_Z(mean = 0.5,
+                               log_sd = 0.2,
+                               phi_age = 0.9,
+                               phi_year = 0.5)) %>%
+  sim_distribution(grid = sim_grid(x_range = c(-140, 140),
+                                   y_range = c(-140, 140),
+                                   res = c(3.5, 3.5),
+                                   shelf_depth = 200,
+                                   shelf_width = 100,
[…]

### 2018-06-28 inst/rmd/vis_sim.Rmd
Add survey distribution, estimates and error plots to flexdashboard

@@
 output: flexdashboard::flex_dashboard
-always_allow_html: yes
+runtime: shiny
 ---
@@
 library(viridis)
-knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8)
+knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8, cache = FALSE)
 ```
@@
   plot_distribution(sim, ages = input$age, year = input$year, type = input$type,
-                    scale = input$scale)
+                    scale = input$scale, axes = FALSE)
 })
@@
 ```
+
+Survey { `r if(!"sp_I" %in% names(sim)) ".hidden"`}
+================================================================================
+
+Column {.sidebar}
+--------------------------------------------------------------------------------
+
+```{r}
+sliderInput("samp_year", "Year", 
+            min = min(sim$years), max = max(sim$years), 
+            value = 1, step = 1)
+sliderInput("samp_sim", "Simulation", 
+            min = min(sim$setdet$sim), max = max(sim$setdet$sim), 
+            value = 1, step = 1)
+```
+
+Column
+--------------------------------------------------------------------------------
+
+### Survey sampling
+
+```{r, eval = "sp_I" %in% names(sim)}
+renderPlotly({
+  plot_samp_dist(sim, which_year = input$samp_year, which_sim = input$samp_sim)
+})
+```
+
+Estimates { `r if(!"age_strat_error" %in% names(sim)) ".hidden"`}
+================================================================================
+
+Column {.sidebar}
+--------------------------------------------------------------------------------
+
+```{r}
+selectInput("set_den", "Set density", 
+            choices = sort(unique(sim$surveys$set_den)))
+selectInput("lengths_cap", "max(lengths)", 
+            choices = sort(unique(sim$surveys$lengths_cap)))
+selectInput("ages_cap", "max(ages)", 
+            choices = sort(unique(sim$surveys$ages_cap)))
+selectInput("facet_by", "Facet by", 
+            choices = c("age", "year"))
+```
+
+Column
+--------------------------------------------------------------------------------
+
+### True population size vs. stratified estimates
+
+```{r, eval = "age_strat_error" %in% names(sim)}
+renderPlotly({
+  surveys <- sim$surveys
+  i <- surveys$survey[surveys$set_den == input$set_den & 
+                      surveys$lengths_cap == input$lengths_cap &
+                      surveys$ages_cap == input$ages_cap]
+  plot_true_vs_est(sim, which_survey = i, facet_by = input$facet_by,
+                   max_sims = 100)
+})
+```
+
+Error { `r if(!"surveys" %in% names(sim)) ".hidden"`}
+================================================================================
+
+Column  {.tabset}
+--------------------------------------------------------------------------------
+
+### Error surface
+
+```{r, eval = "surveys" %in% names(sim)}
+plot_error_surface(sim)
+```
+
+### Error cross-sections
+
+```{r, eval = "surveys" %in% names(sim)}
+plot_error_cross_sections(sim)
+```

### 2018-06-28 R/plot_sim.R

@@ -197,18 +197,19 @@ plot_true_vs_est <- function(sim, which_survey = 1, max_sims = 50,  facet_by = "
       geom_line(aes(y = I_hat), color = I(cols[2]), alpha = 0.2, size = 0.1) +
       geom_line(aes(y = I), color = I(cols[1]), size = 0.1) +
       xlab("Year") + ylab("Index") +
-      facet_wrap(~ age, scales = "free_y") +
-      theme_minimal() + theme(axis.text.y = element_blank(),
-                              axis.ticks.y = element_blank())
+      facet_wrap(~ age, scales = "free_y")
   } else {
     p <- ggplot(data = sub_d, aes(x = age, group = sim)) +
       geom_line(aes(y = I_hat), color = I(cols[2]), alpha = 0.2, size = 0.1) +
       geom_line(aes(y = I), color = I(cols[1]), size = 0.1) +
       xlab("Age") + ylab("Index") +
-      facet_wrap(~ year, scales = "free_y") +
-      theme_minimal() + theme(axis.text.y = element_blank(),
-                              axis.ticks.y = element_blank())
+      facet_wrap(~ year, scales = "free_y")
+
   }
+
+  p <- p + theme_minimal() + theme(axis.text.y = element_blank(),
+                                   axis.ticks.y = element_blank(),
+                                   plot.margin = unit(c(0.1, 0.1, 0.5, 0.5), "cm"))
 
   ggplotly(p)
 
@@ -249,7 +250,7 @@ plot_samp_dist <- function(sim, which_year = 1, which_sim = 1,
                 showlegend = FALSE) %>%
     add_markers(data = setdet[setdet$n == 0, ], color = I(col),
                 x = ~x, y = ~y, text = ~n, size = I(5), symbol = I(4),
-                name = "zero", alpha = 0.5,
+                name = "zero", alpha = 0.2,
                 showlegend = FALSE) %>%
     add_paths(data = df_strat, x = ~long, y = ~lat, color = I("black"),
               hoverinfo = "none", size = I(0.5), showlegend = FALSE,

### 2018-06-28 analysis/cod_sim.R

@@ -57,3 +57,10 @@ res <- test_surveys(pop,
 # res <- resume_test(dir = "analysis/cod_sim_exports/2018-06-27_test")
 setMKLthreads() # turn hyperthreading on again
 
+
+## visualize results
+load("analysis/cod_sim_exports/2018-06-27_test/test_output.RData")
+vis_sim(sim)
+
+
+

### 2018-06-28 analysis/cod_sim_plots.R

@@ -1,22 +1,0 @@
-
-library(SimSurvey)
-library(plotly)
-library(data.table)
-
-load("analysis/cod_sim_exports/2018-06-27_test/test_output.RData")
-
-vis_sim(sim)
-
-
-
-
-plot_error_surface(sim)
-
-plot_true_vs_est(sim, which_survey = 542, max_sims = 100, facet_by = "age")
-
-plot_true_vs_est(sim, which_survey = 542, max_sims = 100, facet_by = "year")
-
-plot_error_cross_sections(sim)
-
-plot_samp_dist(sim, which_year = 14, which_sim = 4)
-

### 2018-07-03 analysis/paper/sim_survey_paper.Rmd
Edit current draft of the paper

@@
 title: Using simulation to optimize the sampling design of fisheries-independent trawl surveys
-author: "Paul M. Regular, Fran Mowbray"
+author: "Paul M. Regular, Fran Mowbray, et al.?"
 date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
-affiliation: "This is a test"
 output: 
@@
 
-A central question to fish stock assessments is whether enough samples were collected to adequately represent the population being monitored.
-For assessments based on trawl surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling.
+A central question to fish stock assessments is: how many samples are needed to adequately represent the population being monitored?
+For assessments based on trawl surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling (i.e.
+the number of sets, lengths and ages sampled in a survey).
 We therefore need to assess the impact of haul design and length and age sampling on survey accuracy to determine optimal sampling effort.
-In this study we used simulations to try to answer this question.
-The method draws repeated stratified-random samples from a spatially correlated field of simulated fish; these samples were then analyzed and estimates of abundance at age were compared against a known truth.
+In this study we simulated repeated stratified-random samples from a spatially correlated field of simulated fish.
+These samples were then analyzed and estimates of abundance at age were compared against a known truth.
 By varying the sampling protocol, we demonstrate that increasing the number of sets conducted per stratum provides the greatest gains in precision when sampling a clustered population.
 Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent.
-These results indicate that length and age sampling efforts for cod (*Gadus morhua*) in the trawl survey conducted by Fisheries and Oceans Canada, Newfoundland region, can be scaled back without significant losses to precision.
-Time saved by reducing said sampling may afford more time to conduct more sets and/or increase sampling efforts of other species.
+These results indicate that length and age sampling efforts for some species(e.g.
+cod *Gadus morhua*) in the trawl survey conducted by regional fisheries management organizations can be scaled back without significant losses to precision.
+Time saved by reducing said sampling may afford more time to conduct additional sets and/or increase sampling efforts of other species.
 While this simulation is tailored to a specific case, the method may be adapted to different systems and survey procedures.
@@
 Few studies, however, have tested these conclusions using simulations.
-A key advantage of a pure simulation approach is that "true" population size and all stages of sampling can be controlled.
+A key advantage of a pure simulation approach is that "true" population size and all stages of sampling can be controlled (set, length and age sampling protocol).
 The relative effects of each stage of sampling on our perception of the "truth" can then be tested.
-This level of testing is of praticular importance for age composition estimates, and the contemporary age-based assessment models that rely on these estimates, as they are affected by all stages of sampling.
+This level of testing is of particular importance for age composition estimates, and the contemporary age-based assessment models that rely on these estimates, as they are affected by all stages of sampling.
 A key challenge, of course, is building a simulation that emulates reality.
@@
 
+# Acknowledgements
+
+- Feedback and advice: Alejandro Buren, Mariano Koen-Alonso, Karen Dwyer, Joanne Morgan, Geoff Evans, Don Power, Brian Healey, Martha Robertson, Danny Ings, Don Stansbury, Mark Simpson, Derek Osborne, Peter Upward, Brad Squires, Dwayne Pittman, Paul Higdon
+- Dave Cote for a review of a pervious version of this manuscript
+
 # References

### 2018-07-04 analysis/paper/sim_survey_paper.Rmd
Describe cohort model

@@
 
+# Methods
+
+The simulation presented here starts with a common cohort model where the abundance at age $a$ in year $y$ ($N_{a,y}$) equals the abundance of that cohort in the previous year multiplied by the survival rate, which is expressed in terms of total mortality ($Z_{a,y}$):
+$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
+Here, numbers-at-age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e.
+recruits) are filled using a random walk, $log(N_{1,y}) = log(N_{1,y-1}) + \epsilon_{y-1}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(z) + \delta_{a,y}$.
+The error around the recruitment process was simulated using the normal distribution, $\epsilon_{y} \sim N(0, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016]
+$$\begin{eqnarray}
+Cov(\delta_{a,y}, \delta_{a-i,y-j}) &=& \frac{\sigma_{\delta}^2 \varphi_{\delta, age}^i \varphi_{\delta, year}^j}{(1 - \varphi_{\delta, age}^2)(1 - \varphi_{\delta, year}^2)} \\
+Corr(\delta_{a,y}, \delta_{a-i,y-j}) &=& \varphi_{\delta, age}^i \varphi_{\delta, year}^j
+\end{eqnarray}
+$$
+
+This structure allows for autocorrelation in process errors over ages and years (i.e.
+total mortality can be made to be more similar for fish that are closer together in age and time).
+Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.
+
 # Acknowledgements

### 2018-07-04 analysis/paper/sim_survey_paper.Rmd
Describe spatial covariance

@@
 
+The next step in the simulation is to distribute the abundance-at-age matrix simulated using the cohort model throughout a spatial field.
+Here, a grid of $s$ locations was generated with a resolution of $r$, a depth gratient, $S$ depth-based strata and $D$ divisions.
+The depth gradient was set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$.
+The probability of fish inhabiting each cell in the grid was then simulated by defining covariance between space, time and age.
+Matern covariance was used to control the level of spatial aggregation:
+
+$$Cov(s_i, s_j) = \sigma_s^2 \frac{2^{1-\lambda}}{\Gamma(\lambda)} (\kappa \left \|s_i - s_j \right\|) K_{\lambda}(\kappa \left \|s_i - s_j \right\|)$$
+where $\left \|s_i - s_j \right\|$ is the Euclidean distance between two locations, $\sigma_s^2$ is the marginal variance, $\Gamma$ is the gamma function, $K_{\lambda}$ denotes the modified Bessel function of the second kind, and $\lambda$ and $\kappa$ are parameters of the covariance [@blangiardo2015].
+
 # Acknowledgements

### 2018-07-06 analysis/paper/sim_survey_paper.Rmd
Discovered and fixed issue with age-year-space covariance

--- 2018-07-03 analysis/paper/sim_survey_paper.Rmd
@@
 The bottom trawl survey of cod (*Gadus morhua*) in NAFO division 3Ps, Newfoundland, was used as a case study for this analysis.
+While our case study is specific, we also hope that the approach presented here will provide a framework for optimizing sampling protocol for an array of species with different life-histories and distributions.
+To facilitate future work and collaborations, we have developed an R package for conducting the tests presented in this paper (https://github.com/PaulRegular/SimSurvey).
+
+# Methods
+
+The simulation presented here starts with a common cohort model where the abundance at age $a$ in year $y$ ($N_{a,y}$) equals the abundance of that cohort in the previous year multiplied by the survival rate, which is expressed in terms of total mortality ($Z_{a,y}$):
+
+$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
+Here, numbers-at-age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e.
+recruits) are filled using a random walk, $log(N_{1,y}) = log(N_{1,y-1}) + \epsilon_{y-1}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(z) + \delta_{a,y}$.
+The error around the recruitment process was simulated using the normal distribution, $\epsilon_{y} \sim N(0, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$.
+The covariance across ages and years is controled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, age}$ and $\varphi_{\delta, year}$, respectively).
+This structure allows for autocorrelation in process errors across ages and years (i.e.
+total mortality can be made to be more similar for fish that are closer together in age and time).
+Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.
+
+The next step in the simulation is to distribute the abundance-at-age matrix simulated using the cohort model throughout a spatial field.
+Here, a grid of $s$ locations was generated with a resolution of $r$, a depth gradient, $S$ depth-based strata and $D$ divisions.
+The depth gradient was set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$.
+The simulated population is distributed through the grid by simulating spatial-temporal noise controled by a parabolic relationship with depth and covariance between ages, years and space
+
+$$\begin{aligned}log(N_{a,y,s}) &= log(N_{a,y}) + \eta_{a,y,s} \\ \eta_{a,y,s} &= \frac{(d_s - \mu_{d}) ^ 2}{2 \sigma_d^2} + \xi_{a,y,s} \end{aligned}$$
+
+Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or disperson of abundance around the mean.
+This helps simulate depth preferences of praticular fish.
+Residual noise $\xi_{a,y,s}$, which is assumed to follow $\xi_{a,y,s} \sim N(0, \Sigma_{a,y,s})$, is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
+Matérn covariance was controled by a variance ($\sigma_s^2$), smoothing ($\lambda$) and scaling parameter ($\kappa$) [@Blangiardo2015].
+Like the process error, age and year covariance was controled by a variance parameter ($\sigma_{\xi}^2$) and two correlation parameters ($\varphi_{\xi, age}$ and $\varphi_{\xi, year}$).
+In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering.
+Note that $\eta_{a,y,s}$ is forced to sum to one in each year to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model.
+A more detailed description of the space-age-year covariance is included in Appendix 1.
+
+The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated using formula described above.
 
@@
 - Feedback and advice: Alejandro Buren, Mariano Koen-Alonso, Karen Dwyer, Joanne Morgan, Geoff Evans, Don Power, Brian Healey, Martha Robertson, Danny Ings, Don Stansbury, Mark Simpson, Derek Osborne, Peter Upward, Brad Squires, Dwayne Pittman, Paul Higdon
-- Dave Cote for a review of a pervious version of this manuscript
+- Dave Cote for a review of a previous version of this manuscript
+
+# Appendix 1: age-year-space covariance
+
+The simulation applied in this paper was set-up to control covariance across ages, years and space. Matern covariance was used to control the level of spatial aggregation:
+
+$$\Sigma_{s} = Cov(s_i, s_j) = \sigma_s^2 \frac{2^{1-\lambda}}{\Gamma(\lambda)} (\kappa \left |s_i - s_j \right|) K_{\lambda}(\kappa \left |s_i - s_j \right|)$$
+
+where $\left |s_i - s_j \right|$ is the Euclidean distance between two locations, $\sigma_s^2$ is the marginal variance, $\Gamma$ is the gamma function, $K_{\lambda}$ denotes the modified Bessel function of the second kind, and $\lambda$ and $\kappa$ are parameters of the covariance [@Blangiardo2015].
+This spatial covariance was used in conjunction with covariance described in Cadigan [-@Cadigan2016] to simulate error across ages, years and space:
+
+$$\xi_{a,y,s} = \left\{\begin{matrix}
+ 0 + v_{a,y} + s_{s} & i = 1, j = 1 & \\ 
+ &  & 
+\end{matrix}\right.$$
 

### 2018-07-06 R/sim_dist.R

@@ -8,13 +8,32 @@
   }
 }
 
-#' Simulate exponential or Matern covariance
+## Helper function for spatial covariance
+.sp_covar <- function(x = NULL, range = 40, lambda = 1, sd = 1, model = "matern") {
+  d <- .dist(x)
+  cormat <- switch(model,
+                   exponential = {
+                     exp(-d / range)
+                   },
+                   matern = {
+                     kappa <- sqrt(8 * lambda) / range # approximate kappa from range as per R-INLA book
+                     2 ^ (1 - lambda) / gamma(lambda) * (kappa * d) ^ lambda *
+                       besselK(x = d * kappa, nu = lambda)
+                   },
+                   stop("wrong or no specification of covariance model"))
+
+  diag(cormat) <- 1
+  covar <- (sd ^ 2) * cormat
+  covar
+}
+
+#' Simulate age-year-space covariance
 #'
 #' @description These functions return a function to use inside \code{\link{sim_distribution}}.
 #'
 #' @param range       Decorrelation range
 #' @param lambda      Controls the degree of smoothness of Matern covariance process
-#' @param sd          Variance (can be age specific in \code{sim_ay_covar}).
+#' @param sd          Variance (can be age specific).
 #' @param phi_age     Defines autocorrelation through ages. Can be one value or a vector of the same
 #'                    length as ages
 #' @param phi_year    Defines autocorrelation through years. Can be one value or a vector of the same
@@ -23,38 +42,20 @@
 #' @param group_years Make space-age-year variance equal across these years
 #' @param model       String indicating either "exponential" or "matern" as the correlation function
 #'
-#' @rdname sim_sp_covar
 #' @export
-sim_sp_covar <- function(range = 40, lambda = 1, sd = 0.1, model = "matern") {
-  function(x = NULL) {
-    d <- .dist(x)
-    cormat <- switch(model,
-                     exponential = {
-                       exp(-d / range)
-                     },
-                     matern = {
-                       kappa <- sqrt(8 * lambda) / range # approximate kappa from range as per R-INLA book
-                       2 ^ (1 - lambda) / gamma(lambda) * (kappa * d) ^ lambda *
[…]

### 2018-07-06 R/sim_pop.R

@@ -68,7 +68,7 @@ sim_Z <- function(mean = 0.7, log_sd = 0.2, phi_age = 0.9, phi_year = 0.5, plot 
       for (i in seq_along(ages)) {
         if ((i == 1) & (j == 1)) {
           m <- 0
-          s <- log_sd / pc_age
+          s <- log_sd / (pc_age * pc_year)
           Z[i, j] <- rnorm(1, m, s)
         }
         if ((i > 1) & (j == 1)) {

### 2018-07-06 analysis/imitate_cod_data.R

@@ -101,7 +101,7 @@ af$age <- as.integer(gsub("af", "", af$age))
 set.seed(438)
 pop <- sim_abundance(ages = 1:20,
                      years = 1:20,
-                     R = sim_R(mean = 100000000,
+                     R = sim_R(mean = 30000000,
                                log_sd = 0.5,
                                random_walk = TRUE),
                      Z = sim_Z(mean = 0.5,
@@ -117,23 +117,22 @@ pop <- sim_abundance(ages = 1:20,
                                    n_div = 1,
                                    strat_breaks = seq(0, 1000, by = 20),
                                    strat_splits = 2),
-                   space_covar = sim_sp_covar(range = 40,
-                                              sd = 0.1),
-                   ay_covar = sim_ay_covar(sd = 10,
-                                           phi_age = 0.1,
-                                           phi_year = 0.8,
-                                           group_ages = 5:20),
-                   depth_par = sim_parabola(mu = 250,
-                                            sigma = 50))
+                   ays_covar = sim_ays_covar(sd = 2.8,
+                                             range = 300,
+                                             phi_age = 0.1,
+                                             phi_year = 0.9,
+                                             group_ages = 5:20),
+                   depth_par = sim_parabola(mu = 200,
+                                            sigma = 70))
 
 ## Quick look at distribution
 sp_N <- data.frame(merge(pop$sp_N, pop$grid_xy, by = "cell"))
 for (j in rev(pop$ages)) {
-  z <- xtabs(N ~ x + y, subset = age == j & year == 20, data = sp_N)
+  z <- xtabs(N ~ x + y, subset = age == j & year == 1, data = sp_N)
   image(z = z, axes = FALSE, col = viridis::viridis(100), main = paste("age", j))
 }
 for (i in rev(pop$years)) {
-  z <- xtabs(N ~ x + y, subset = age == 10 & year == i, data = sp_N)
+  z <- xtabs(N ~ x + y, subset = age == 1 & year == i, data = sp_N)
   image(z = z, axes = FALSE, col = viridis::viridis(100), main = paste("year", i))
 }
 
@@ -207,13 +206,13 @@ symbols(survey$setdet$x, survey$setdet$y,
 
 
[…]

### 2018-07-07 analysis/paper/sim_survey_paper.Rmd
Describe age-year-space covariance

@@
 This helps simulate depth preferences of praticular fish.
-Residual noise $\xi_{a,y,s}$, which is assumed to follow $\xi_{a,y,s} \sim N(0, \Sigma_{a,y,s})$, is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
-Matérn covariance was controled by a variance ($\sigma_s^2$), smoothing ($\lambda$) and scaling parameter ($\kappa$) [@Blangiardo2015].
-Like the process error, age and year covariance was controled by a variance parameter ($\sigma_{\xi}^2$) and two correlation parameters ($\varphi_{\xi, age}$ and $\varphi_{\xi, year}$).
+Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
+Spatial correlations was controled by a smoothing ($\lambda$) and scaling parameter ($\kappa$) [@Blangiardo2015] and correlation across ages and years was controled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively.
+The overall variance of the process was controled by $\sigma_{\xi}^2$.
 In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering.
@@
 
-The simulation applied in this paper was set-up to control covariance across ages, years and space. Matern covariance was used to control the level of spatial aggregation:
+The simulation applied in this paper was set-up to control covariance across ages, years and space.
+To do this we used a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
+As described in Appendix A in Cadigan [-@Cadigan2016], the age-year covariance can be broken down into a series of AR(1) processes.
+We integrate Matérn covariance into this series of equations:
 
-$$\Sigma_{s} = Cov(s_i, s_j) = \sigma_s^2 \frac{2^{1-\lambda}}{\Gamma(\lambda)} (\kappa \left |s_i - s_j \right|) K_{\lambda}(\kappa \left |s_i - s_j \right|)$$
+$$\xi_{a,y,s} \sim \left\{\begin{matrix} MVN\left (0, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a = 1, y = 1 & \\  MVN\left (\varphi_{year} \xi_{1, y - 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)} \boldsymbol{R}_s \right ) & a = 1, y > 1 & \\ MVN\left (\varphi_{age} \xi_{a - 1, 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a > 1, y = 1 & \\ MVN\left (\varphi_{year} \xi_{a, y - 1, s} + \varphi_{age} (\xi_{a - 1, y, s} - \varphi_{year} \xi_{a - 1, y - 1, s}), \sigma_\xi^2 \boldsymbol{R}_s \right ) & a > 1, y > 1 & \end{matrix}\right.$$
 
-where $\left |s_i - s_j \right|$ is the Euclidean distance between two locations, $\sigma_s^2$ is the marginal variance, $\Gamma$ is the gamma function, $K_{\lambda}$ denotes the modified Bessel function of the second kind, and $\lambda$ and $\kappa$ are parameters of the covariance [@Blangiardo2015].
-This spatial covariance was used in conjunction with covariance described in Cadigan [-@Cadigan2016] to simulate error across ages, years and space:
+Where $MVN$ indicates the multivariate normal distribution, $\sigma_\xi^2$ controls the variance of the process, $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$ control correlation in the age and year dimension and $\boldsymbol{R}_s$ is defined by Matérn correlation:
 
-$$\xi_{a,y,s} = \left\{\begin{matrix}
- 0 + v_{a,y} + s_{s} & i = 1, j = 1 & \\ 
- &  & 
-\end{matrix}\right.$$
+$$\boldsymbol{R}_s = \frac{2^{1-\lambda}}{\Gamma(\lambda)} (\kappa \left |s_i - s_j \right|) K_{\lambda}(\kappa \left |s_i - s_j \right|)$$
+where $\left |s_i - s_j \right|$ is the Euclidean distance between two locations, $\Gamma$ is the gamma function, $K_{\lambda}$ denotes the modified Bessel function of the second kind, and $\lambda$ and $\kappa$ control the smoothness and scale of the spatial process [@Blangiardo2015].
+With this structure, simulated error is correlated across ages, years and space.
 

### 2018-07-07 analysis/cod_sim.R

@@ -44,7 +44,7 @@ res <- test_surveys(pop,
                     surveys = surveys,
                     keep_details = 98,
                     n_sims = 5,
-                    n_loops = 100,
+                    n_loops = 200,
                     cores = 6,
                     q = sim_logistic(k = 2, x0 = 3),
                     growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),

### 2018-07-09 analysis/paper/sim_survey_paper.Rmd
Complete initial description of simulation

@@
 
+## Simulate abundance
+
 The simulation presented here starts with a common cohort model where the abundance at age $a$ in year $y$ ($N_{a,y}$) equals the abundance of that cohort in the previous year multiplied by the survival rate, which is expressed in terms of total mortality ($Z_{a,y}$):
@@
 
+## Simulate spatial distribution
+
 The next step in the simulation is to distribute the abundance-at-age matrix simulated using the cohort model throughout a spatial field.
-Here, a grid of $s$ locations was generated with a resolution of $r$, a depth gradient, $S$ depth-based strata and $D$ divisions.
+Here, a grid of $s$ locations was generated with a resolution of $r$ and a depth gradient.
 The depth gradient was set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$.
@@
 Note that $\eta_{a,y,s}$ is forced to sum to one in each year to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model.
+Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid.
 A more detailed description of the space-age-year covariance is included in Appendix 1.
 
+## Simulate survey
+
 The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated using formula described above.
+First, the grid is divided into $N_{div}$ divisions and $N_{strat}$ depth-based strata.
+The area of each strata is calculated and this is used to define the sets to allocate to each strata under a praticular set density, $D_{sets}$.
+The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:
+
+$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A}{r^2} q_a )$$
+Where $n_{a,y,s}$ is the number of fish sampled by a set, $A$ indicates the area covered by the trawl, $r$ is the resolution of the grid cells, and $q_a$ is the catchability coefficient of each age.
+Here, catchabilities were defined using a logistic curve controled by a steepness, $k$, and midpoint parameter, $x_0$.
+The lengths of the fish sampled by the set are then simulated using the original von Bertalanffy growth curve [@Cailliet2006]:
+
+$$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$
+Where $l_{\infty}$ is the mean asympotic lengths, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$.
+Depending on the number of fish caught, sub-sampling is then conducted.
+Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division.
+
+## Simulation settings
 

### 2018-07-09 analysis/cod_sim.R

@@ -49,12 +49,12 @@ res <- test_surveys(pop,
                     q = sim_logistic(k = 2, x0 = 3),
                     growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
                     export = "analysis/cod_sim_exports")
-# res <- resume_test(dir = "analysis/cod_sim_exports/2018-06-27_test")
+# res <- resume_test(dir = "analysis/cod_sim_exports/2018-07-07_test")
 setMKLthreads() # turn hyperthreading on again
 
 
 ## visualize results
-load("analysis/cod_sim_exports/2018-06-27_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-07-07_test/test_output.RData")
 vis_sim(sim)
 
 

### 2018-07-09 analysis/paper/sim_survey_paper.Rmd
Start parameter setting table

@@
 output: 
-  word_document:
+  bookdown::word_document2:
     reference_docx: template.docx
@@
 Here, numbers-at-age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e.
-recruits) are filled using a random walk, $log(N_{1,y}) = log(N_{1,y-1}) + \epsilon_{y-1}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(z) + \delta_{a,y}$.
-The error around the recruitment process was simulated using the normal distribution, $\epsilon_{y} \sim N(0, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$.
-The covariance across ages and years is controled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, age}$ and $\varphi_{\delta, year}$, respectively).
+recruits) vary around a baseline value, $log(N_{1,y}) = log(\mu_r) + \epsilon_{y}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(\mu_Z) + \delta_{a,y}$.
+The error around the recruitment process was set to follow a random walk, $\epsilon_{y} \sim N(\epsilon_{y - 1}, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$.
+The covariance across ages and years is controlled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, age}$ and $\varphi_{\delta, year}$, respectively).
 This structure allows for autocorrelation in process errors across ages and years (i.e.
@@
 The depth gradient was set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$.
-The simulated population is distributed through the grid by simulating spatial-temporal noise controled by a parabolic relationship with depth and covariance between ages, years and space
+The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space
 
@@
 
-Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or disperson of abundance around the mean.
-This helps simulate depth preferences of praticular fish.
+Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean.
+This helps simulate depth preferences of particular fish.
 Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
-Spatial correlations was controled by a smoothing ($\lambda$) and scaling parameter ($\kappa$) [@Blangiardo2015] and correlation across ages and years was controled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively.
-The overall variance of the process was controled by $\sigma_{\xi}^2$.
+Spatial correlations was controlled by a smoothing ($\lambda$) and scaling parameter ($\kappa$) [@Blangiardo2015] and correlation across ages and years was controlled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively.
+The overall variance of the process was controlled by $\sigma_{\xi}^2$.
 In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering.
@@
 First, the grid is divided into $N_{div}$ divisions and $N_{strat}$ depth-based strata.
-The area of each strata is calculated and this is used to define the sets to allocate to each strata under a praticular set density, $D_{sets}$.
+The area of each strata is calculated and this is used to define the sets to allocate to each strata under a particular set density, $D_{sets}$.
 The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:
@@
 Where $n_{a,y,s}$ is the number of fish sampled by a set, $A$ indicates the area covered by the trawl, $r$ is the resolution of the grid cells, and $q_a$ is the catchability coefficient of each age.
-Here, catchabilities were defined using a logistic curve controled by a steepness, $k$, and midpoint parameter, $x_0$.
+Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
 The lengths of the fish sampled by the set are then simulated using the original von Bertalanffy growth curve [@Cailliet2006]:
@@
 $$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$
-Where $l_{\infty}$ is the mean asympotic lengths, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$.
+Where $l_{\infty}$ is the mean asymptotic lengths, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$.
 Depending on the number of fish caught, sub-sampling is then conducted.
@@
 ## Simulation settings
+
+Ideally the parameter settings used here would be based on model fits to real data, however, conditioning the operating model described above was outside the scope of this project.
+Moreover, many of the parameters described above may not be identifiable or it may be difficult to obtain un-biased estimates [e.g.
+bias introduced by length-stratified sampling complicate the fitting of growth models; @Morgan1997].
+As such, parameters were informally selected to roughly correspond to observations typically obtained by the Fisheries and Oceans Canada survey of cod in NAFO division 3Ps, Newfoundland.
+Though this approach is far from exact, the goal was to approximate the life-history and distribution of an actual population to simulate an arena for testing the consequences of surveys with different sampling protocol.
+Parameter settings and descriptions used in this case study are presented in Table \@ref(tab:test).
+
+# Tables
+
+Table: (\#tab:test) Parameter descriptions, symbols and settings used in the cod case study.
+
+**Parameter name**   **Symbol**   **Setting**
+------------------   ----------   -----------
+              Ages          $a$          1:20
+             Years          $y$          1:20
+  Mean recruitment      $\mu_r$           30M
 

### 2018-07-10 analysis/paper/sim_survey_paper.Rmd
Work table up to survey settings

@@
 The next step in the simulation is to distribute the abundance-at-age matrix simulated using the cohort model throughout a spatial field.
-Here, a grid of $s$ locations was generated with a resolution of $r$ and a depth gradient.
+Here, a grid of $s$ locations was generated with a resolution of $R$ and a depth gradient.
 The depth gradient was set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$.
@@
 Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
-Spatial correlations was controlled by a smoothing ($\lambda$) and scaling parameter ($\kappa$) [@Blangiardo2015] and correlation across ages and years was controlled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively.
+Spatial correlations was controlled by a smoothing ($\lambda$) and scaling parameter ($\kappa$) [here we approximate $\kappa$ from a range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years was controlled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively.
 The overall variance of the process was controlled by $\sigma_{\xi}^2$.
@@
 
-$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A}{r^2} q_a )$$
-Where $n_{a,y,s}$ is the number of fish sampled by a set, $A$ indicates the area covered by the trawl, $r$ is the resolution of the grid cells, and $q_a$ is the catchability coefficient of each age.
+$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{trawl}}{A_{cell}} q_a )$$
+Where $n_{a,y,s}$ is the number of fish sampled by a set, $A_{trawl}$ indicates the area covered by the trawl, $A_{cell}$ is the area of a grid cell (i.e.
+$R^2$), and $q_a$ is the catchability coefficient of each age.
 Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
@@
                                        **Parameter name**                   **Symbol**      **Setting**
-------------------   ----------   -----------
+---------------------------------------------------------   --------------------------   --------------
+                                              *Abundance*
                                                      Ages                          $a$             1:20
@@
                                          Mean recruitment                      $\mu_r$              30M
+                    Standard deviation of log-recruitment                   $\sigma_r$              0.5
+                                     Mean total mortality                      $\mu_Z$              0.7
+                    Standard deviation of total mortality                   $\sigma_Z$              0.2
+  Correlation across ages in error around total mortality      $\varphi_{\delta, age}$              0.9
+ Correlation across years in error around total mortality     $\varphi_{\delta, year}$              0.5
+                                                   *Grid*
+                                               Grid cells                          $s$           1:6400
+                                               Resolution                          $r$           3.5 km
+                                              Shelf depth                  $d_{shelf}$            200 m
+                                              Shelf width                  $w_{shelf}$           100 km
+                                              Depth range         $[d_{min}, d_{max}]$    [0 m, 1000 m]
+                                      Number of divisions                    $N_{div}$                1
+                                         Number of strata                  $N_{strat}$               42
+                                           *Distribution*
+        Standard deviation of age-year-space distribution               $\sigma_{\xi}$              2.8
+                        Smoothness of spatial correlation                    $\lambda$                1
+                             Range of spatial correlation                          $r$            50 km
+          Correlation across ages in spatial distribution         $\varphi_{\xi, age}$             0.1*
+         Correlation across years in spatial distribution        $\varphi_{\xi, year}$              0.9
+            Depth at which abundance is typically highest                      $\mu_d$            200 m
+                Dispersion around depth of peak abundance                   $\sigma_d$             70 m
+
+\* Distribution set to be equal for ages 5+
 

### 2018-07-10 analysis/paper/sim_survey_paper.Rmd
Complete draft of tables

@@
 $$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$
-Where $l_{\infty}$ is the mean asymptotic lengths, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$.
+Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$.
 Depending on the number of fish caught, sub-sampling is then conducted.
@@
 Though this approach is far from exact, the goal was to approximate the life-history and distribution of an actual population to simulate an arena for testing the consequences of surveys with different sampling protocol.
-Parameter settings and descriptions used in this case study are presented in Table \@ref(tab:test).
+Parameter settings and descriptions used in this case study are presented in Table \@ref(tab:par).
+A series of ### surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests).
 
@@
 
-Table: (\#tab:test) Parameter descriptions, symbols and settings used in the cod case study.
+Table: (\#tab:par) Parameter descriptions, symbols and global settings used in the cod case study.
 
-                                       **Parameter name**                   **Symbol**      **Setting**
----------------------------------------------------------   --------------------------   --------------
-                                              *Abundance*
-                                                     Ages                          $a$             1:20
-                                                    Years                          $y$             1:20
-                                         Mean recruitment                      $\mu_r$              30M
-                    Standard deviation of log-recruitment                   $\sigma_r$              0.5
-                                     Mean total mortality                      $\mu_Z$              0.7
-                    Standard deviation of total mortality                   $\sigma_Z$              0.2
-  Correlation across ages in error around total mortality      $\varphi_{\delta, age}$              0.9
- Correlation across years in error around total mortality     $\varphi_{\delta, year}$              0.5
-                                                   *Grid*
-                                               Grid cells                          $s$           1:6400
-                                               Resolution                          $r$           3.5 km
-                                              Shelf depth                  $d_{shelf}$            200 m
-                                              Shelf width                  $w_{shelf}$           100 km
-                                              Depth range         $[d_{min}, d_{max}]$    [0 m, 1000 m]
-                                      Number of divisions                    $N_{div}$                1
-                                         Number of strata                  $N_{strat}$               42
-                                           *Distribution*
-        Standard deviation of age-year-space distribution               $\sigma_{\xi}$              2.8
-                        Smoothness of spatial correlation                    $\lambda$                1
-                             Range of spatial correlation                          $r$            50 km
-          Correlation across ages in spatial distribution         $\varphi_{\xi, age}$             0.1*
-         Correlation across years in spatial distribution        $\varphi_{\xi, year}$              0.9
-            Depth at which abundance is typically highest                      $\mu_d$            200 m
-                Dispersion around depth of peak abundance                   $\sigma_d$             70 m
+|                                       **Parameter name** |                 **Symbol** |    **Setting** |
+|:-------------------------------------------------------- | :------------------------- | :------------- |
+|                                              *Abundance* |                            |                |  
+|                                                     Ages |                        $a$ |           1:20 |
+|                                                    Years |                        $y$ |           1:20 |
+|                                         Mean recruitment |                    $\mu_r$ |            30M |
+|                    Standard deviation of log-recruitment |                 $\sigma_r$ |            0.5 |
+|                                     Mean total mortality |                    $\mu_Z$ |            0.7 |
+|                    Standard deviation of total mortality |                 $\sigma_Z$ |            0.2 |
+|  Correlation across ages in error around total mortality |    $\varphi_{\delta, age}$ |            0.9 |
+| Correlation across years in error around total mortality |   $\varphi_{\delta, year}$ |            0.5 |
+|                                                          |                            |                |  
+|                                                   *Grid* |                            |                | 
+|                                               Grid cells |                        $s$ |         1:6400 |
+|                                               Resolution |                        $R$ |         3.5 km |
+|                                              Shelf depth |                $d_{shelf}$ |          200 m |
+|                                              Shelf width |                $w_{shelf}$ |         100 km |
+|                                              Depth range |       $[d_{min}, d_{max}]$ |  [0 m, 1000 m] |
+|                                      Number of divisions |                  $N_{div}$ |              1 |
+|                                         Number of strata |                $N_{strat}$ |             42 |
+|                                                          |                            |                |  
+|                                           *Distribution* |                            |                | 
+|        Standard deviation of age-year-space distribution |             $\sigma_{\xi}$ |            2.8 |
+|                        Smoothness of spatial correlation |                  $\lambda$ |              1 |
+|                             Range of spatial correlation |                        $r$ |          50 km |
+|          Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |           0.1* |
+|         Correlation across years in spatial distribution |      $\varphi_{\xi, year}$ |            0.9 |
+|            Depth at which abundance is typically highest |                    $\mu_d$ |          200 m |
+|                Dispersion around depth of peak abundance |                 $\sigma_d$ |           70 m |
+|                                                          |                            |                |  
+|                                                 *Survey* |                            |                |
+|                                  Area covered by a trawl |                $A_{trawl}$ |     0.03 km^2^ |
+|              Steepness of logistic curve of catchability |                        $k$ |              2 |
+|               Midpoint of logistic curve of catchability |                      $x_0$ |              3 |
+|                                   Mean asymptotic length |               $l_{\infty}$ |            120 |
+|                                          Length at birth |                      $l_0$ |              5 |
+|                                              Growth rate |                        $K$ |            0.1 |
+|                                             Length group |                $l_{group}$ |              1 |  
+\* Distribution set to be equal for ages 5+
 
-\* Distribution set to be equal for ages 5+
+Table: (\#tab:tests) Descriptions and settings of different survey sampling protocol tested in the cod case study. All possible combinations of the options below were tested.
+
+|      **Parameter name** |      **Symbol** |                                       **Setting** |
+|:----------------------- | :-------------- | :------------------------------------------------ |
+|             Set density |      $D_{sets}$ |    0.0005, 0.001, 0.002, 0.005, 0.01 sets / km^2^ |
+|  Length sampling effort |   $M_{lengths}$ |        5, 10, 20, 50, 100, 500, 100 lengths / set |
+|     Age sampling effort |      $M_{ages}$ |   2, 5, 10, 20, 50 ages / length group / division |
 

### 2018-07-11 analysis/paper/sim_survey_paper.Rmd
Complete draft of methods

@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_knit$set(root.dir = "../..")
+```
+
+```{r data, include = FALSE}
+load("analysis/cod_sim_exports/2018-07-07_test/test_output.RData")
 ```
@@
 
-## Simulation settings
+## Simulation settings and analysis
 
@@
 Parameter settings and descriptions used in this case study are presented in Table \@ref(tab:par).
-A series of ### surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests).
+A series of `r nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests).
+Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain an estimate of abundance at age for each year for each simulation, $\hat{I}_{a,y,k}$.
+We used root-mean-squared error (RMSE) as a measure of the percision and bias of the abundance estimates from each survey:
+
+$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})}{N_a N_y N_k}}$$
+Where $N_a$, $N_y$, and $N_k$ are the number of years, ages and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e.
+catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$).
+All simulations and analyses were conducted using `r substr(version$version.string, 1, 15)` and the code has been developed into a package called `SimSurvey` to facilitate future work and collaboration (https://github.com/PaulRegular/SimSurvey).
+
+# Results
+
+# Discussion
 

### 2018-07-12 R/plot_sim.R
Work up two plot functions for a plot for the paper

@@ -18,6 +18,7 @@
 #' @param facet_by       Facet plot by "age" or "year"?
 #' @param col            Plot color
 #' @param cols           Plot colors
+#' @param main           Plot title
 #' @param ...            Additional arguments to pass to \code{\link{plotly::plot_ly}}.
 #'
 #' @import plotly
@@ -284,5 +285,108 @@ plot_samp_dist <- function(sim, which_year = 1, which_sim = 1,
                   margin = 0.1),
           nrows = 1, margin = 0.02, widths = c(0.6, 0.4),
           titleX = TRUE, titleY = TRUE)
+
+}
+
+
+#' @export
+#' @rdname plot_trend
+plot_total_abundance <- function(sim, which_year = 1, main = "") {
+
+  m <- list(
+    l = 0,
+    r = 0,
+    b = 0,
+    t = 0,
+    pad = 0
+  )
+
+  xax <- list(
+    title = "",
+    zeroline = FALSE,
+    showline = FALSE,
+    showticklabels = FALSE,
+    showgrid = FALSE,
+    tickcolor = toRGB("white")
+  )
+  yax <- c(scaleanchor = "x", xax)
+
+  sp_N_tot <- sim$sp_N[year == which_year, list(N = sum(N)), by = c("year", "cell")]
+  xyz <- merge(sim$grid_xy, sp_N_tot, by = "cell")
+  size <- unique(diff(pretty(xyz$N, 25)))
+
+  sp_strat <- raster::rasterToPolygons(sim$grid$strat, dissolve = TRUE)
+  df_strat <- suppressWarnings(fortify(sp_strat) %>% group_by(group))
+
+  abun_plot <- plot_ly() %>%
+    add_trace(x = ~x, y = ~y, z = ~N, data = xyz, type = "contour",
+              contours = list(coloring = "lines", start = 0, end = max(xyz$N),
+                              size = size)) %>%
+    add_paths(data = df_strat, x = ~long, y = ~lat, color = I("black"),
+              hoverinfo = "none", size = I(0.5), showlegend = FALSE,
+              alpha = 0.1) %>%
+    colorbar(ticklen = 0, ypad = 20, xpad = 0, outlinewidth = 0, thickness = 10) %>%
+    layout(xaxis = xax, yaxis = yax, margin = m, legend = list(x = 0.5, y = 0.5),
+           annotations = list(text = main, xref = "paper", yref = "paper",
[…]

### 2018-07-12 analysis/cod_sim.R

@@ -57,5 +57,13 @@ setMKLthreads() # turn hyperthreading on again
 load("analysis/cod_sim_exports/2018-07-07_test/test_output.RData")
 vis_sim(sim)
 
+abun_plot <- plot_total_abundance(sim)
+set_plot <- plot_set_catch(sim)
+export(abun_plot, file = "analysis/paper/figures/total_abundance.png",
+       vwidth = 1000, vheight = 800, zoom = 20)
+
+# plotly_IMAGE(abun_plot, out_file = "analysis/paper/figures/total_abundance.png")
+
+
 
 

### 2018-07-12 analysis/paper/sim_survey_paper.Rmd
Work up two plot functions for a plot for the paper
Resort to static plots and add first figure to paper

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE,
+                      message = FALSE,
+                      warning = FALSE,
+                      dev = "png",
+                      dpi = 300)
 knitr::opts_knit$set(root.dir = "../..")
@@
 ```{r data, include = FALSE}
+library(SimSurvey)
+library(plotly)
+library(raster)
+library(lattice)
+library(viridis)
 load("analysis/cod_sim_exports/2018-07-07_test/test_output.RData")
@@
 
+# Acknowledgements
+
+- Feedback and advice: Alejandro Buren, Mariano Koen-Alonso, Karen Dwyer, Joanne Morgan, Geoff Evans, Don Power, Brian Healey, Martha Robertson, Danny Ings, Don Stansbury, Mark Simpson, Derek Osborne, Peter Upward, Brad Squires, Dwayne Pittman, Paul Higdon
+- Dave Cote for a review of a previous version of this manuscript
+
+# Appendix 1: age-year-space covariance
+
+The simulation applied in this paper was set-up to control covariance across ages, years and space.
+To do this we used a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
+As described in Appendix A in Cadigan [-@Cadigan2016], the age-year covariance can be broken down into a series of AR(1) processes.
+We integrate Matérn covariance into this series of equations:
+
+$$\xi_{a,y,s} \sim \left\{\begin{matrix} MVN\left (0, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a = 1, y = 1 & \\  MVN\left (\varphi_{year} \xi_{1, y - 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)} \boldsymbol{R}_s \right ) & a = 1, y > 1 & \\ MVN\left (\varphi_{age} \xi_{a - 1, 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a > 1, y = 1 & \\ MVN\left (\varphi_{year} \xi_{a, y - 1, s} + \varphi_{age} (\xi_{a - 1, y, s} - \varphi_{year} \xi_{a - 1, y - 1, s}), \sigma_\xi^2 \boldsymbol{R}_s \right ) & a > 1, y > 1 & \end{matrix}\right.$$
+
+Where $MVN$ indicates the multivariate normal distribution, $\sigma_\xi^2$ controls the variance of the process, $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$ control correlation in the age and year dimension and $\boldsymbol{R}_s$ is defined by Matérn correlation:
+
+$$\boldsymbol{R}_s = \frac{2^{1-\lambda}}{\Gamma(\lambda)} (\kappa \left |s_i - s_j \right|) K_{\lambda}(\kappa \left |s_i - s_j \right|)$$
+where $\left |s_i - s_j \right|$ is the Euclidean distance between two locations, $\Gamma$ is the gamma function, $K_{\lambda}$ denotes the modified Bessel function of the second kind, and $\lambda$ and $\kappa$ control the smoothness and scale of the spatial process [@Blangiardo2015].
+With this structure, simulated error is correlated across ages, years and space.
+
 # Tables
@@
 
-# Acknowledgements
+# Figures
 
-- Feedback and advice: Alejandro Buren, Mariano Koen-Alonso, Karen Dwyer, Joanne Morgan, Geoff Evans, Don Power, Brian Healey, Martha Robertson, Danny Ings, Don Stansbury, Mark Simpson, Derek Osborne, Peter Upward, Brad Squires, Dwayne Pittman, Paul Higdon
-- Dave Cote for a review of a previous version of this manuscript
+```{r ex_fig, fig.height = 3.8, fig.width = 8.5}
 
-# Appendix 1: age-year-space covariance
+which_year <- 7
+which_sim <- 4
 
-The simulation applied in this paper was set-up to control covariance across ages, years and space.
-To do this we used a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
-As described in Appendix A in Cadigan [-@Cadigan2016], the age-year covariance can be broken down into a series of AR(1) processes.
-We integrate Matérn covariance into this series of equations:
+sp_N_tot <- sim$sp_I[year == which_year & sim == which_sim, list(N = sum(I)), by = c("year", "cell")]
+xyz <- merge(sim$grid_xy, sp_N_tot, by = "cell")
+size <- unique(diff(pretty(xyz$N, 25)))
 
-$$\xi_{a,y,s} \sim \left\{\begin{matrix} MVN\left (0, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a = 1, y = 1 & \\  MVN\left (\varphi_{year} \xi_{1, y - 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)} \boldsymbol{R}_s \right ) & a = 1, y > 1 & \\ MVN\left (\varphi_{age} \xi_{a - 1, 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a > 1, y = 1 & \\ MVN\left (\varphi_{year} \xi_{a, y - 1, s} + \varphi_{age} (\xi_{a - 1, y, s} - \varphi_{year} \xi_{a - 1, y - 1, s}), \sigma_\xi^2 \boldsymbol{R}_s \right ) & a > 1, y > 1 & \end{matrix}\right.$$
+sp_strat <- raster::rasterToPolygons(sim$grid$strat, dissolve = TRUE)
+df_strat <- suppressWarnings(fortify(sp_strat) %>% group_by(group))
 
-Where $MVN$ indicates the multivariate normal distribution, $\sigma_\xi^2$ controls the variance of the process, $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$ control correlation in the age and year dimension and $\boldsymbol{R}_s$ is defined by Matérn correlation:
+r <- raster::rasterFromXYZ(xyz[, c("x", "y", "N")])
+l <- rasterToContour(r, levels = pretty(xyz$N, 20))
+df_lines <- suppressWarnings(fortify(l))
+levels <- as.numeric(as.character(l$level))
+levels <- paste0(levels / 1000, "k")
+labs <- levels
+labs[seq_along(labs) %% 2 == 0] <- ""
+levels <- factor(levels, levels = levels)
+names(levels) <- rownames(as.data.frame(l))
+df_lines$level <- levels[df_lines$id]
 
-$$\boldsymbol{R}_s = \frac{2^{1-\lambda}}{\Gamma(\lambda)} (\kappa \left |s_i - s_j \right|) K_{\lambda}(\kappa \left |s_i - s_j \right|)$$
-where $\left |s_i - s_j \right|$ is the Euclidean distance between two locations, $\Gamma$ is the gamma function, $K_{\lambda}$ denotes the modified Bessel function of the second kind, and $\lambda$ and $\kappa$ control the smoothness and scale of the spatial process [@Blangiardo2015].
-With this structure, simulated error is correlated across ages, years and space.
+abun_plot <- ggplot() +
+  geom_path(data = df_strat, aes(x = long, y = lat, group = group), 
+            color = "grey", size = 0.05) +
+  geom_path(data = df_lines, aes(x = long, y = lat, group = group, color = level),
+            size = 0.3) +
+  scale_color_viridis(labels = labs, discrete = TRUE, name = "N") + 
+  scale_x_continuous(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
+  coord_equal() +
+  theme_void() + guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
+                                             default.unit = "cm", ncol = 1)) +
+  theme(legend.justification = c(0, 1),
+        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
+
+setdet <- sim$setdet
+setdet <- setdet[setdet$year == which_year & setdet$sim == which_sim, ]
+
+set_plot <- ggplot() +
+  geom_path(data = df_strat, aes(x = long, y = lat, group = group), 
+            color = "grey", size = 0.05) +
+  geom_point(data = setdet, aes(x = x, y = y, size = n, color = n), alpha = 0.8) +
+  scale_size_area(name = "   n", max_size = 10, breaks = pretty(setdet$n, 8)) +
+  scale_color_viridis(name = "   n", breaks = pretty(setdet$n, 8)) + 
+  scale_x_continuous(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
+  coord_equal() +
+  theme_void() + guides(color = guide_legend(keywidth = 0.4, keyheight = 0.2, default.unit = "cm")) +
+  theme(legend.justification = c(0, 1),
+        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
+
+p <- cowplot::plot_grid(abun_plot, set_plot, labels = c("a)", "b)"), align = "hv",
+                   label_fontface = "plain")
+cowplot::save_plot(plot = p, filename = "analysis/paper/figures/abun_and_sets_example.pdf",
+                   base_height = 3.8, base_width = 8.5)
+p
+
+```
 

### 2018-07-12 R/plot_sim.R

@@ -379,10 +379,10 @@ plot_set_catch <- function(sim, which_year = 1, which_sim = 1, main = "") {
     add_paths(data = df_strat, x = ~long, y = ~lat, color = I("black"),
               hoverinfo = "none", size = I(0.5), showlegend = FALSE,
               alpha = 0.1) %>%
-    layout(xaxis = xax, yaxis = yax, legend = list(x = 1, y = 0.9),
+    layout(xaxis = xax, yaxis = yax, legend = list(x = 1, y = 0.85),
            annotations = list(
              list(text = "n", xref = "paper", yref = "paper",
-                              x = 1.06, y = 0.95, showarrow = FALSE),
+                              x = 1.09, y = 0.9, showarrow = FALSE),
              list(text = main, xref = "paper", yref = "paper",
                   x = 0.08, y = 1, showarrow = FALSE)
            ),

### 2018-07-12 analysis/cod_sim.R

@@ -57,13 +57,5 @@ setMKLthreads() # turn hyperthreading on again
 load("analysis/cod_sim_exports/2018-07-07_test/test_output.RData")
 vis_sim(sim)
 
-abun_plot <- plot_total_abundance(sim)
-set_plot <- plot_set_catch(sim)
-export(abun_plot, file = "analysis/paper/figures/total_abundance.png",
-       vwidth = 1000, vheight = 800, zoom = 20)
-
-# plotly_IMAGE(abun_plot, out_file = "analysis/paper/figures/total_abundance.png")
-
-
 
 

### 2018-07-12 analysis/paper/sim_survey_paper.Rmd
Add true vs estimated plot to the paper

@@
 library(viridis)
+library(data.table)
 load("analysis/cod_sim_exports/2018-07-07_test/test_output.RData")
@@
 
+- add population trend plot?
+
 # Discussion
@@
 
+```{r fig_fun}
+
+## http://www.moeding.net/archives/32-Metric-prefixes-for-ggplot2-scales.html
+format_si <- function(...) {
+  # Format a vector of numeric values according
+  # to the International System of Units.
+  # http://en.wikipedia.org/wiki/SI_prefix
+  #
+  # Based on code by Ben Tupper
+  # https://stat.ethz.ch/pipermail/r-help/2012-January/299804.html
+  # Args:
+  #   ...: Args passed to format()
+  #
+  # Returns:
+  #   A function to format a vector of strings using
+  #   SI prefix notation
+  #
+
+  function(x) {
+    limits <- c(1e-24, 1e-21, 1e-18, 1e-15, 1e-12,
+                1e-9,  1e-6,  1e-3,  1e0,   1e3,
+                1e6,   1e9,   1e12,  1e15,  1e18,
+                1e21,  1e24)
+    prefix <- c("y",   "z",   "a",   "f",   "p",
+                "n",   "µ",   "m",   " ",   "k",
+                "M",   "G",   "T",   "P",   "E",
+                "Z",   "Y")
+
+    # Vector with array indices according to position in intervals
+    i <- findInterval(abs(x), limits)
+
+    # Set prefix to " " for very small values < 1e-24
+    i <- ifelse(i == 0, which(limits == 1e0), i)
+
+    paste0(format(round(x/limits[i], 1),
+                 trim = TRUE, scientific = FALSE, ...),
+          prefix[i])
+  }
+}
+si <- format_si()
+
+```
+
 ```{r ex_fig, fig.height = 3.8, fig.width = 8.5}
@@
 levels <- as.numeric(as.character(l$level))
-levels <- paste0(levels / 1000, "k")
+levels <- si(levels)
 labs <- levels
@@
 
+```{r est_fig, fig.height = 9, fig.width = 8.5}
+
+which_survey <- 98
+
+d <- sim$age_strat_error
+d <- d[d$survey == which_survey & d$sim %in% 1:10, ]
+d$age[d$age > 15] <- 15 # aggregate beyond 15
+d$age <- as.character(d$age)
+d$age[d$age == "15"] <- "15+"
+d$age <- factor(d$age, levels = c(as.character(1:14), "15+"))
+d <- d[, list(I_hat = sum(I_hat), I = sum(I)), by = c("year", "age", "sim")]
+
+p <- ggplot(data = d) + 
+  geom_line(aes(x = year, y = I_hat, color = sim, group = sim), size = 0.3) +
+  geom_line(aes(x = year, y = I), color = viridis::viridis(1), size = 0.7) +
+  scale_color_viridis(begin = 0.5, guide = FALSE) +
+  facet_wrap(~age, scales = "free_y", dir = "v", nrow = 5) +
+  xlab("Year") + ylab("Abundance index") +
+  theme_minimal() +
+  scale_y_continuous(labels = format_si()) +
+  scale_x_continuous(expand = c(0.01, 0)) +
+  theme(legend.justification = c(0, 1),
+        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))
+
+cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est.pdf",
+                   base_height = 9, base_width = 8.5)
+p
+
+```
+
 # References

### 2018-07-12 analysis/paper/sim_survey_paper.Rmd
Add 3D wireframe figure

@@
 |             Set density |      $D_{sets}$ |    0.0005, 0.001, 0.002, 0.005, 0.01 sets / km^2^ |
-|  Length sampling effort |   $M_{lengths}$ |        5, 10, 20, 50, 100, 500, 100 lengths / set |
+|  Length sampling effort |   $M_{lengths}$ |       5, 10, 20, 50, 100, 500, 1000 lengths / set |
 |     Age sampling effort |      $M_{ages}$ |   2, 5, 10, 20, 50 ages / length group / division |
@@
 
+```{r error_fig, fig.height = 3, fig.width = 8.5, results = "hide"}
+
+trellis_theme <- list(
+  box.3d = list(col = "darkgrey"), 
+  axis.line = list(col = "transparent"),
+  strip.background = list(col = "transparent"), 
+  strip.border = list(col = "transparent"),
+   layout.heights = list(
+        top.padding = 0,
+        main.key.padding = 0,
+        key.axis.padding = 0,
+        axis.xlab.padding = 0,
+        xlab.key.padding = 0,
+        key.sub.padding = 0,
+        bottom.padding = -5
+    ),
+    layout.widths = list(
+        left.padding = 0,
+        key.ylab.padding = 0,
+        ylab.axis.padding = 0,
+        axis.key.padding = 0,
+        right.padding = 1
+    )
+)
+
+zscale_format <- function(lim, logsc = FALSE, ...) {
+    ans <- zscale.components.default(lim, logsc = logsc, ...)
+    ans$left$labels$labels <- si(ans$left$labels$at)
+    ans
+}
+
+d <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
+# d <- d[d$set_den != 0.0005, ]
+d <- d[d$set_den %in% c(0.0005, 0.002, 0.01)]
+set_den <- sort(unique(d$set_den))
+d$set_den <- factor(d$set_den)
+options(scipen = 999)
+labs <- sapply(seq_along(set_den), function(i) as.expression(bquote(D[sets] == .(set_den[i]))), 
+               USE.NAMES = FALSE)
+zlabs <- list(at = pretty(d$RMSE), labels = si(pretty(d$RMSE)))
+
+pdf("analysis/paper/figures/error_wireframe.pdf", height = 3, width = 8.5)
+
+p <- wireframe(RMSE ~  lengths_cap * ages_cap | set_den, data = d,
+          as.table = TRUE, drape = TRUE, col = "white", lwd = 0.5,
+          strip = strip.custom(factor.levels = labs),
+          scales = list(arrows = FALSE, cex = 0.7, z = zlabs),
+          col.regions = viridis::viridis(100),
+          screen = list(z = 230, x = 300), 
+          xlab = list(expression(M[lengths]), rot = 25), 
+          ylab = list(expression(M[ages]), rot = -15), 
+          zlab = list("RMSE", rot = 95),
+          colorkey = list(labels = zlabs, height = 0.6,
+                          axis.line = list(col = "darkgrey")),
+          par.settings = trellis_theme)
+p <- update(p, layout = c(3, 1))
+p
+
+dev.off()
+
+p
+
+```
+
 # References

### 2018-07-13 analysis/paper/sim_survey_paper.Rmd
Add figure captions

@@
 Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain an estimate of abundance at age for each year for each simulation, $\hat{I}_{a,y,k}$.
-We used root-mean-squared error (RMSE) as a measure of the percision and bias of the abundance estimates from each survey:
+We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance estimates from each survey:
 
@@
 - add population trend plot?
+
+\@ref(fig:sp-fig)
 
@@
 
-Table: (\#tab:par) Parameter descriptions, symbols and global settings used in the cod case study.
+Table: (\#tab:par) - Parameter descriptions, symbols and global settings used in the cod case study.
 
@@
 
-Table: (\#tab:tests) Descriptions and settings of different survey sampling protocol tested in the cod case study. All possible combinations of the options below were tested.
+Table: (\#tab:tests) - Descriptions and settings of different survey sampling protocol tested in the cod case study. All possible combinations of the options below were tested.
 
@@
 
-```{r ex_fig, fig.height = 3.8, fig.width = 8.5}
+```{r sp-fig, cache = TRUE, fig.height = 3.8, fig.width = 8.5, fig.cap = "- An example from year seven of the simulation of a) the distribution of total abundance available to the survey, and b) locations of the sets conducted and total numbers caught in a stratified random survey with a set density of 0.002 sets / km^2^."}
 
@@
 
-```{r est_fig, fig.height = 9, fig.width = 8.5}
+```{r est-fit, cache = TRUE, fig.height = 9, fig.width = 8.5, fig.cap = "- Stratified estimates of abundance at age across years from 10 simulations of a survey with a set density of 0.002 sets / km^2^ and sampling caps of 500 lengths per set and 10 ages per length group.
+The heavy dark line indicates the true tred in the population available to the survey."}
 
@@
 
-```{r error_fig, fig.height = 3, fig.width = 8.5, results = "hide"}
+```{r error-fig, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
+Panels represent surveys with different set densities ($D_{sets}$), x-axes represnet the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$)"}
 

### 2018-07-14 analysis/paper/sim_survey_paper.Rmd
Exploring alternate figures for paper

@@
 
-```{r sp-fig, cache = TRUE, fig.height = 3.8, fig.width = 8.5, fig.cap = "- An example from year seven of the simulation of a) the distribution of total abundance available to the survey, and b) locations of the sets conducted and total numbers caught in a stratified random survey with a set density of 0.002 sets / km^2^."}
+```{r sp-fig, cache = TRUE, fig.height = 3.8, fig.width = 8.5, fig.cap = "- An example from year seven of the simulation of a) the distribution of total abundance available to the survey, and b) locations of the sets conducted and total numbers caught in a stratified random survey with a set density of 0.002 sets / km^2^.
+Grey lines indicate depth-based survey strata."}
 
@@
 
-```{r est-fit, cache = TRUE, fig.height = 9, fig.width = 8.5, fig.cap = "- Stratified estimates of abundance at age across years from 10 simulations of a survey with a set density of 0.002 sets / km^2^ and sampling caps of 500 lengths per set and 10 ages per length group.
+```{r est-fit, cache = TRUE, fig.height = 8, fig.width = 7, fig.cap = "- Stratified estimates of abundance at age across years from 10 simulations of a survey with a set density of 0.002 sets / km^2^ and sampling caps of 500 lengths per set and 10 ages per length group.
 The heavy dark line indicates the true tred in the population available to the survey."}
@@
 d <- sim$age_strat_error
-d <- d[d$survey == which_survey & d$sim %in% 1:10, ]
-d$age[d$age > 15] <- 15 # aggregate beyond 15
+d <- d[d$survey == which_survey & d$sim %in% 1:1000, ]
+d$age[d$age > 8] <- 8 # aggregate beyond 10
 d$age <- as.character(d$age)
-d$age[d$age == "15"] <- "15+"
-d$age <- factor(d$age, levels = c(as.character(1:14), "15+"))
+d$age[d$age == "8"] <- "8+"
+d$age <- factor(d$age, levels = c(as.character(1:7), "8+"))
 d <- d[, list(I_hat = sum(I_hat), I = sum(I)), by = c("year", "age", "sim")]
@@
 p <- ggplot(data = d) + 
-  geom_line(aes(x = year, y = I_hat, color = sim, group = sim), size = 0.3) +
+  geom_line(aes(x = year, y = I_hat, color = sim, group = sim), size = 0.1) +
   geom_line(aes(x = year, y = I), color = viridis::viridis(1), size = 0.7) +
   scale_color_viridis(begin = 0.5, guide = FALSE) +
-  facet_wrap(~age, scales = "free_y", dir = "v", nrow = 5) +
+  facet_wrap(~age, scales = "free_y", dir = "v", nrow = 4) +
   xlab("Year") + ylab("Abundance index") +
@@
 cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est.pdf",
-                   base_height = 9, base_width = 8.5)
+                   base_height = 8, base_width = 7)
 p
@@
 
+```{r age-fig, fig.width = 7, fig.height = 6}
+
+which_set_dens <- c(0.0005, 0.002, 0.01)
+which_len_caps <- c(10, 100, 1000)
+which_year <- 5
+
+d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
+d <- d[d$sim %in% 1:1000 & d$year == which_year & d$set_den %in% which_set_dens &
+         d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
+d$age[d$age > 8] <- 8 # aggregate beyond 8
+d <- d[, list(I_hat = sum(I_hat), I = sum(I)), 
+       by = c("year", "age", "sim", "survey", "set_den", "lengths_cap")]
+
+p <- ggplot(data = d) + 
+  geom_line(aes(x = age, y = I_hat, color = sim, group = sim), size = 0.1) +
+  geom_line(aes(x = age, y = I), color = viridis::viridis(1), size = 0.5) +
+  scale_color_viridis(begin = 0.5, guide = FALSE) +
+  facet_grid(set_den ~ lengths_cap) +
+  xlab("Age") + ylab("Abundance index") +
+  theme_minimal() +
+  scale_y_continuous(labels = format_si()) +
+  scale_x_continuous(expand = c(0.01, 0), 
+                     labels = c("2", "4", "6", "8+"), 
+                     breaks = seq(2, 8 , 2)) +
+  theme(legend.justification = c(0, 1),
+        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))
+
+p
+
+```
+
+```{r year-fig, fig.width = 7, fig.height = 6}
+
+which_set_dens <- c(0.0005, 0.002, 0.01)
+which_len_caps <- c(10, 100, 1000)
+which_age <- 4
+
+d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
+d <- d[d$sim %in% 1:1000 & d$age == which_age & d$set_den %in% which_set_dens &
+         d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
+
+p <- ggplot(data = d) + 
+  geom_line(aes(x = year, y = I_hat, color = sim, group = sim), size = 0.1) +
+  geom_line(aes(x = year, y = I), color = viridis::viridis(1), size = 0.5) +
+  scale_color_viridis(begin = 0.5, guide = FALSE) +
+  facet_grid(set_den ~ lengths_cap) +
+  xlab("Year") + ylab("Abundance index") +
+  theme_minimal() +
+  scale_y_continuous(labels = format_si()) +
+  scale_x_continuous(expand = c(0.01, 0)) +
+  theme(legend.justification = c(0, 1),
+        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))
+
+p
+
+```
+
 # References

### 2018-07-15 analysis/paper/sim_survey_paper.Rmd
Draft "fan plots" for paper

@@
 |                             Range of spatial correlation |                        $r$ |          50 km |
-|          Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |           0.1* |
+|          Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |           0.5* |
 |         Correlation across years in spatial distribution |      $\varphi_{\xi, year}$ |            0.9 |
@@
 d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
-d <- d[d$sim %in% 1:1000 & d$year == which_year & d$set_den %in% which_set_dens &
+d <- d[d$year == which_year & d$set_den %in% which_set_dens &
          d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
@@
 
-p <- ggplot(data = d) + 
-  geom_line(aes(x = age, y = I_hat, color = sim, group = sim), size = 0.1) +
-  geom_line(aes(x = age, y = I), color = viridis::viridis(1), size = 0.5) +
-  scale_color_viridis(begin = 0.5, guide = FALSE) +
+## quantiles for fan plot
+quants <- seq(90, 10, by = -10)
+ints <- lapply(quants, function(q) {
+  d[, list(prob = paste0(q, "%"),
+           lower = quantile(I_hat, prob = (1 - q / 100) / 2),
+           upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
+    by = c("year", "age", "set_den", "lengths_cap")]
+})
+ints <- rbindlist(ints)
+ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
+
+true <- d[d$sim == 1, ]
+
+cols <- viridis::viridis(3)
+ggplot(data = ints) + 
+  geom_ribbon(aes(x = age, ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
+  geom_line(aes(x = age, y = I), data = d, color = cols[1], size = 1) +
+  scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
+  scale_alpha_discrete(name = "Pr") + 
   facet_grid(set_den ~ lengths_cap) +
@@
                      breaks = seq(2, 8 , 2)) +
+  guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
+                              default.unit = "cm", ncol = 1)) +
   theme(legend.justification = c(0, 1),
         plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))
-
-p
 
@@
 which_len_caps <- c(10, 100, 1000)
-which_age <- 4
+which_age <- 6
 
 d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
-d <- d[d$sim %in% 1:1000 & d$age == which_age & d$set_den %in% which_set_dens &
+d <- d[d$age == which_age & d$set_den %in% which_set_dens &
        d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
 
-p <- ggplot(data = d) + 
-  geom_line(aes(x = year, y = I_hat, color = sim, group = sim), size = 0.1) +
-  geom_line(aes(x = year, y = I), color = viridis::viridis(1), size = 0.5) +
-  scale_color_viridis(begin = 0.5, guide = FALSE) +
+## quantiles for fan plot
+quants <- seq(90, 10, by = -10)
+ints <- lapply(quants, function(q) {
+  d[, list(prob = paste0(q, "%"),
+           lower = quantile(I_hat, prob = (1 - q / 100) / 2),
+           upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
+    by = c("year", "age", "set_den", "lengths_cap")]
+})
+ints <- rbindlist(ints)
+ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
+
+true <- d[d$sim == 1, ]
+
+cols <- viridis::viridis(3)
+ggplot(data = ints) + 
+  geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
+  geom_line(aes(x = year, y = I), data = d, color = cols[1], size = 1) +
+  scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
+  scale_alpha_discrete(name = "Pr") + 
   facet_grid(set_den ~ lengths_cap) +
@@
   scale_x_continuous(expand = c(0.01, 0)) +
+  guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
+                              default.unit = "cm", ncol = 1)) +
   theme(legend.justification = c(0, 1),
@@
 
-p
-
 ```
 
+## Figure to-dos
+
+- Fix facet labels (make them look like the wireframe plot)
+- Add RMSE label to colorkey in wireframe plot
+- Change vis_sim estimates plot to fan plots like used here
+
 # References

### 2018-07-15 analysis/paper/sim_survey_paper.Rmd
Minor modification of the correlation in simulation

@@
 library(data.table)
-load("analysis/cod_sim_exports/2018-07-07_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
 ```
@@
 which_year <- 7
-which_sim <- 4
+which_sim <- 1
 
@@
 d <- sim$age_strat_error
-d <- d[d$survey == which_survey & d$sim %in% 1:1000, ]
+d <- d[d$survey == which_survey & d$sim %in% 1:10, ]
 d$age[d$age > 8] <- 8 # aggregate beyond 10
@@
 which_len_caps <- c(10, 100, 1000)
-which_year <- 5
+which_year <- 7
 
@@
 which_len_caps <- c(10, 100, 1000)
-which_age <- 6
+which_age <- 4
 
@@
 
+```{r tot-fig, fig.width = 7, fig.height = 3}
+
+which_set_dens <- c(0.0005, 0.002, 0.01)
+which_len_caps <- c(10, 100, 1000)
+
+d <- merge(sim$surveys, sim$total_strat_error, by = "survey")
+d <- d[d$set_den %in% which_set_dens & 
+         d$lengths_cap == 100 & d$ages_cap == 10, ] ## length and age sampling not a factor here
+
+## quantiles for fan plot
+quants <- seq(90, 10, by = -10)
+ints <- lapply(quants, function(q) {
+  d[, list(prob = paste0(q, "%"),
+           lower = quantile(I_hat, prob = (1 - q / 100) / 2),
+           upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
+    by = c("year", "set_den")]
+})
+ints <- rbindlist(ints)
+ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
+
+true <- d[d$sim == 1, ]
+
+cols <- viridis::viridis(3)
+ggplot(data = ints) + 
+  geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
+  geom_line(aes(x = year, y = I), data = d, color = cols[1], size = 1) +
+  scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
+  scale_alpha_discrete(name = "Pr") + 
+  facet_wrap(~ set_den, nrow = 1) +
+  xlab("Year") + ylab("Abundance index") +
+  theme_minimal() +
+  scale_y_continuous(labels = format_si()) +
+  scale_x_continuous(expand = c(0.01, 0)) +
+  guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
+                              default.unit = "cm", ncol = 1)) +
+  theme(legend.justification = c(0, 1),
+        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))
+
+```
+
 ## Figure to-dos

### 2018-07-15 analysis/cod_sim.R

@@ -25,7 +25,7 @@ pop <- sim_abundance(ages = 1:20,
                                    strat_splits = 2),
                    ays_covar = sim_ays_covar(sd = 2.8,
                                              range = 300,
-                                             phi_age = 0.1,
+                                             phi_age = 0.5,
                                              phi_year = 0.9,
                                              group_ages = 5:20),
                    depth_par = sim_parabola(mu = 200,
@@ -49,12 +49,12 @@ res <- test_surveys(pop,
                     q = sim_logistic(k = 2, x0 = 3),
                     growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
                     export = "analysis/cod_sim_exports")
-# res <- resume_test(dir = "analysis/cod_sim_exports/2018-07-07_test")
+# res <- resume_test(dir = "analysis/cod_sim_exports/2018-07-13_test")
 setMKLthreads() # turn hyperthreading on again
 
 
 ## visualize results
-load("analysis/cod_sim_exports/2018-07-07_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
 vis_sim(sim)
 
 

### 2018-07-15 analysis/imitate_cod_data.R

@@ -119,7 +119,7 @@ pop <- sim_abundance(ages = 1:20,
                                    strat_splits = 2),
                    ays_covar = sim_ays_covar(sd = 2.8,
                                              range = 300,
-                                             phi_age = 0.1,
+                                             phi_age = 0.5,
                                              phi_year = 0.9,
                                              group_ages = 5:20),
                    depth_par = sim_parabola(mu = 200,
@@ -232,7 +232,7 @@ symbols(survey$setdet$x, survey$setdet$y,
 
 
 ## Real data (hold age or year and animate the other)
-plot_ly(data = af[af$age == 3, ]) %>%
+plot_ly(data = af[af$age == 5, ]) %>%
   add_markers(x = ~easting, y = ~northing, size = ~freq, frame = ~survey.year,
               sizes = c(5, 1000), showlegend = FALSE) %>%
   animation_opts(frame = 5)
@@ -261,14 +261,14 @@ plot_ly(data = af[af$survey.year == 2013, ]) %>%
 
 sim_af <- data.frame(survey$full_setdet)
 sim_af %>%
-  filter(age == 3) %>%
+  filter(age == 5) %>%
   group_by(year) %>%
   plot_ly(x = ~x, y = ~y, size = ~n, frame = ~year,
           sizes = c(5, 1000), showlegend = FALSE) %>%
   add_markers() %>%
   animation_opts(frame = 5)
 sim_af %>%
-  filter(year == 20) %>%
+  filter(year == 5) %>%
   group_by(age) %>%
   plot_ly(x = ~x, y = ~y, size = ~n, frame = ~age,
           sizes = c(5, 1000), showlegend = FALSE) %>%

### 2018-07-17 analysis/paper/sim_survey_paper.Rmd
Add fancharts to paper (replace "worm plots")

@@
 
+## little helper function for the fan plot
+fan_plot <- function(d, x = "Age", total_strat = FALSE) {
+
+  if (total_strat) {
+    ints_by = c("year", "set_den")
+  } else {
+    ints_by = c("year", "age", "set_den", "lengths_cap")
+  }
+
+  ## quantiles for fan plot
+  quants <- seq(90, 10, by = -10)
+  ints <- lapply(quants, function(q) {
+    d[, list(prob = paste0(q, "%"),
+             lower = quantile(I_hat, prob = (1 - q / 100) / 2),
+             upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
+      by = ints_by]
+  })
+  ints <- rbindlist(ints)
+  ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
+
+  true <- d[d$sim %in% 1:3, ]
+
+  if (x == "Age") {
+    p <- ggplot(data = ints, aes(x = age))
+    xlabs <- c("2", "4", "6", "8+")
+    xbreaks <- seq(2, 8 , 2)
+  }
+  if (x == "Year") {
+    p <- ggplot(data = ints, aes(x = year))
+    xlabs <- waiver()
+    xbreaks <- waiver()
+  }
+
+  p <-  p + 
+    geom_ribbon(aes(ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
+    geom_line(aes(y = I), data = true, color = viridis::viridis(1), size = 0.7) +
+    scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
+    scale_alpha_discrete(name = "Pr") + 
+    xlab(x) + ylab("Abundance index") +
+    scale_y_continuous(labels = format_si()) +
+    scale_x_continuous(expand = c(0.01, 0), 
+                       labels = xlabs, 
+                       breaks = xbreaks) +
+    guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
+                                default.unit = "cm", ncol = 1))
+
+  if (total_strat) {
+    p <- p + facet_wrap(~ set_den, nrow = 1,
+                        labeller = label_bquote(cols = D[sets]==.(set_den)))
+  } else {
+    p <- p + facet_grid(lengths_cap ~ set_den, 
+                        labeller = label_bquote(cols = D[sets]==.(set_den),
+                                                rows = M[lengths]==.(lengths_cap)))
+  }
+
+  p + theme_bw() +
+    theme(legend.justification = c(0, 1),
+          plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"),
+          panel.grid.major = element_blank(), 
+          panel.grid.minor = element_blank(),
+          strip.background = element_blank(),
+          panel.border = element_rect(fill = NA, color = "darkgrey", size = 0.4),
+          axis.ticks = element_line(color = "darkgrey", size = 0.4))
+
+}
+
 ```
@@
 
-```{r est-fit, cache = TRUE, fig.height = 8, fig.width = 7, fig.cap = "- Stratified estimates of abundance at age across years from 10 simulations of a survey with a set density of 0.002 sets / km^2^ and sampling caps of 500 lengths per set and 10 ages per length group.
-The heavy dark line indicates the true tred in the population available to the survey."}
+\pagebreak
 
-which_survey <- 98
+```{r tot-fig, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
+The heavy dark line indicates the true tred in the total population available to the survey."}
 
-d <- sim$age_strat_error
-d <- d[d$survey == which_survey & d$sim %in% 1:10, ]
-d$age[d$age > 8] <- 8 # aggregate beyond 10
-d$age <- as.character(d$age)
-d$age[d$age == "8"] <- "8+"
-d$age <- factor(d$age, levels = c(as.character(1:7), "8+"))
-d <- d[, list(I_hat = sum(I_hat), I = sum(I)), by = c("year", "age", "sim")]
+which_set_dens <- c(0.0005, 0.002, 0.01)
+which_len_caps <- c(10, 100, 1000)
 
-p <- ggplot(data = d) + 
-  geom_line(aes(x = year, y = I_hat, color = sim, group = sim), size = 0.1) +
-  geom_line(aes(x = year, y = I), color = viridis::viridis(1), size = 0.7) +
-  scale_color_viridis(begin = 0.5, guide = FALSE) +
-  facet_wrap(~age, scales = "free_y", dir = "v", nrow = 4) +
-  xlab("Year") + ylab("Abundance index") +
-  theme_minimal() +
-  scale_y_continuous(labels = format_si()) +
-  scale_x_continuous(expand = c(0.01, 0)) +
-  theme(legend.justification = c(0, 1),
-        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))
+d <- merge(sim$surveys, sim$total_strat_error, by = "survey")
+d <- d[d$set_den %in% which_set_dens & 
+         d$lengths_cap == 100 & d$ages_cap == 10, ] ## length and age sampling not a factor here
 
-cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est.pdf",
-                   base_height = 8, base_width = 7)
+p <- fan_plot(d, x = "Year", total_strat = TRUE)
+cowplot::save_plot(plot = p, filename = "analysis/paper/figures/total_strat_true_vs_est.pdf",
+                   base_height = 3, base_width = 8)
 p
@@
 ```
+
+\pagebreak
+
+```{r age-fig, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
+The heavy dark line indicates the true population available to the survey."}
+
+which_set_dens <- c(0.0005, 0.002, 0.01)
+which_len_caps <- c(10, 100, 1000)
+which_year <- 7
+
+d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
+d <- d[d$year == which_year & d$set_den %in% which_set_dens &
+         d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
+d$age[d$age > 8] <- 8 # aggregate beyond 8
+d <- d[, list(I_hat = sum(I_hat), I = sum(I)), 
+       by = c("year", "age", "sim", "survey", "set_den", "lengths_cap")]
+
+p <- fan_plot(d, x = "Age")
+cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_age.pdf",
+                   base_height = 6, base_width = 8)
+p
+
+```
+
+\pagebreak
+
+```{r year-fig, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
+The heavy dark line indicates the true tred in the population available to the survey."}
+
+which_set_dens <- c(0.0005, 0.002, 0.01)
+which_len_caps <- c(10, 100, 1000)
+which_age <- 4
+
+d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
+d <- d[d$age == which_age & d$set_den %in% which_set_dens &
+       d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
+
+p <- fan_plot(d, x = "Year")
+cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_year.pdf",
+                   base_height = 6, base_width = 8)
+p
+
+```
+
+\pagebreak
 
@@
 )
+
+# wireframe_legend <- packGrob(
+#   draw.colorkey(key = list(col = viridis::viridis, at = do.breaks(range(z),100))), 
+#   textGrob("RMSE", x = 0, y = 0.5, just = c("left", "centre")), 
+#   height = unit(2, "lines"), side = "top", dynamic = TRUE)
 
@@
 p
+grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc"))
 
-dev.off()
+invisible(dev.off())
 
 p
+grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc"))
 
@@
 
-```{r age-fig, fig.width = 7, fig.height = 6}
-
-which_set_dens <- c(0.0005, 0.002, 0.01)
-which_len_caps <- c(10, 100, 1000)
-which_year <- 7
-
-d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
-d <- d[d$year == which_year & d$set_den %in% which_set_dens &
-         d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
-d$age[d$age > 8] <- 8 # aggregate beyond 8
-d <- d[, list(I_hat = sum(I_hat), I = sum(I)), 
-       by = c("year", "age", "sim", "survey", "set_den", "lengths_cap")]
-
-## quantiles for fan plot
-quants <- seq(90, 10, by = -10)
-ints <- lapply(quants, function(q) {
-  d[, list(prob = paste0(q, "%"),
-           lower = quantile(I_hat, prob = (1 - q / 100) / 2),
-           upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
-    by = c("year", "age", "set_den", "lengths_cap")]
-})
-ints <- rbindlist(ints)
-ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
-
-true <- d[d$sim == 1, ]
-
-cols <- viridis::viridis(3)
-ggplot(data = ints) + 
-  geom_ribbon(aes(x = age, ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
-  geom_line(aes(x = age, y = I), data = d, color = cols[1], size = 1) +
-  scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
-  scale_alpha_discrete(name = "Pr") + 
-  facet_grid(set_den ~ lengths_cap) +
-  xlab("Age") + ylab("Abundance index") +
-  theme_minimal() +
-  scale_y_continuous(labels = format_si()) +
-  scale_x_continuous(expand = c(0.01, 0), 
-                     labels = c("2", "4", "6", "8+"), 
-                     breaks = seq(2, 8 , 2)) +
-  guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
-                              default.unit = "cm", ncol = 1)) +
-  theme(legend.justification = c(0, 1),
-        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))
-
-```
-
-```{r year-fig, fig.width = 7, fig.height = 6}
-
-which_set_dens <- c(0.0005, 0.002, 0.01)
-which_len_caps <- c(10, 100, 1000)
-which_age <- 4
-
-d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
-d <- d[d$age == which_age & d$set_den %in% which_set_dens &
-       d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
-
-## quantiles for fan plot
-quants <- seq(90, 10, by = -10)
-ints <- lapply(quants, function(q) {
-  d[, list(prob = paste0(q, "%"),
-           lower = quantile(I_hat, prob = (1 - q / 100) / 2),
-           upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
-    by = c("year", "age", "set_den", "lengths_cap")]
-})
-ints <- rbindlist(ints)
-ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
-
-true <- d[d$sim == 1, ]
-
-cols <- viridis::viridis(3)
-ggplot(data = ints) + 
-  geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
-  geom_line(aes(x = year, y = I), data = d, color = cols[1], size = 1) +
-  scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
-  scale_alpha_discrete(name = "Pr") + 
-  facet_grid(set_den ~ lengths_cap) +
-  xlab("Year") + ylab("Abundance index") +
-  theme_minimal() +
-  scale_y_continuous(labels = format_si()) +
-  scale_x_continuous(expand = c(0.01, 0)) +
-  guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
-                              default.unit = "cm", ncol = 1)) +
-  theme(legend.justification = c(0, 1),
-        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))
-
-```
-
-```{r tot-fig, fig.width = 7, fig.height = 3}
-
-which_set_dens <- c(0.0005, 0.002, 0.01)
-which_len_caps <- c(10, 100, 1000)
-
-d <- merge(sim$surveys, sim$total_strat_error, by = "survey")
-d <- d[d$set_den %in% which_set_dens & 
-         d$lengths_cap == 100 & d$ages_cap == 10, ] ## length and age sampling not a factor here
-
-## quantiles for fan plot
-quants <- seq(90, 10, by = -10)
-ints <- lapply(quants, function(q) {
-  d[, list(prob = paste0(q, "%"),
-           lower = quantile(I_hat, prob = (1 - q / 100) / 2),
-           upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
-    by = c("year", "set_den")]
-})
-ints <- rbindlist(ints)
-ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
-
-true <- d[d$sim == 1, ]
-
-cols <- viridis::viridis(3)
-ggplot(data = ints) + 
-  geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
-  geom_line(aes(x = year, y = I), data = d, color = cols[1], size = 1) +
-  scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
-  scale_alpha_discrete(name = "Pr") + 
-  facet_wrap(~ set_den, nrow = 1) +
-  xlab("Year") + ylab("Abundance index") +
-  theme_minimal() +
-  scale_y_continuous(labels = format_si()) +
-  scale_x_continuous(expand = c(0.01, 0)) +
-  guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
-                              default.unit = "cm", ncol = 1)) +
-  theme(legend.justification = c(0, 1),
-        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"))
-
-```
-
-## Figure to-dos
-
-- Fix facet labels (make them look like the wireframe plot)
-- Add RMSE label to colorkey in wireframe plot
-- Change vis_sim estimates plot to fan plots like used here
+\pagebreak
 

### 2018-07-17 analysis/cod_sim.R

@@ -94,22 +94,22 @@ pop <- sim_abundance(ages = 1:20,
 ## Test a series of surveys
 ## Simulate surveys and compare stratified estimates to the true index
 setMKLthreads(1) # turn off MKL hyperthreading
-surveys <- expand_surveys(set_den = c(0.5, 2, 10) / 1000,
-                          lengths_cap = c(10, 100, 500, 1000),
-                          ages_cap = c(2, 10, 50))
+surveys <- expand_surveys(set_den = c(0.5, 1, 2, 5, 10) / 1000,
+                          lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
+                          ages_cap = c(2, 5, 10, 20, 50))
 surveys[surveys$set_den == 0.002 &
           surveys$lengths_cap == 500 &
-          surveys$ages_cap == 10, ]    ## survey 20 ~ roughly current protocol
+          surveys$ages_cap == 10, ]    ## survey 98 ~ roughly current protocol
 sim <- test_surveys(pop,
                     surveys = surveys,
-                    keep_details = 20,
+                    keep_details = 98,
                     n_sims = 5,
                     n_loops = 200,
                     cores = 7,
                     q = sim_logistic(k = 2, x0 = 3),
                     growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
                     export = "analysis/cod_sim_exports")
-# sim <- resume_test(dir = "analysis/cod_sim_exports/2018-07-15_test")
+# sim <- resume_test(dir = "analysis/cod_sim_exports/2018-07-16_test")
 setMKLthreads() # turn hyperthreading on again
 
 

### 2018-07-19 analysis/paper/sim_survey_paper.Rmd
Add a second appendix to the paper describing the results of the simulation with no intra-haul correlation

@@
 - Dave Cote for a review of a previous version of this manuscript
-
-# Appendix 1: age-year-space covariance
-
-The simulation applied in this paper was set-up to control covariance across ages, years and space.
-To do this we used a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
-As described in Appendix A in Cadigan [-@Cadigan2016], the age-year covariance can be broken down into a series of AR(1) processes.
-We integrate Matérn covariance into this series of equations:
-
-$$\xi_{a,y,s} \sim \left\{\begin{matrix} MVN\left (0, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a = 1, y = 1 & \\  MVN\left (\varphi_{year} \xi_{1, y - 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)} \boldsymbol{R}_s \right ) & a = 1, y > 1 & \\ MVN\left (\varphi_{age} \xi_{a - 1, 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a > 1, y = 1 & \\ MVN\left (\varphi_{year} \xi_{a, y - 1, s} + \varphi_{age} (\xi_{a - 1, y, s} - \varphi_{year} \xi_{a - 1, y - 1, s}), \sigma_\xi^2 \boldsymbol{R}_s \right ) & a > 1, y > 1 & \end{matrix}\right.$$
-
-Where $MVN$ indicates the multivariate normal distribution, $\sigma_\xi^2$ controls the variance of the process, $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$ control correlation in the age and year dimension and $\boldsymbol{R}_s$ is defined by Matérn correlation:
-
-$$\boldsymbol{R}_s = \frac{2^{1-\lambda}}{\Gamma(\lambda)} (\kappa \left |s_i - s_j \right|) K_{\lambda}(\kappa \left |s_i - s_j \right|)$$
-where $\left |s_i - s_j \right|$ is the Euclidean distance between two locations, $\Gamma$ is the gamma function, $K_{\lambda}$ denotes the modified Bessel function of the second kind, and $\lambda$ and $\kappa$ control the smoothness and scale of the spatial process [@Blangiardo2015].
-With this structure, simulated error is correlated across ages, years and space.
 
@@
 
-\pagebreak
+<br>
 
@@
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
-The heavy dark line indicates the true tred in the population available to the survey."}
+The heavy dark line indicates the true trend in the population available to the survey."}
 
@@
 
-\pagebreak
+<br>
 
 ```{r error-fig, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
-Panels represent surveys with different set densities ($D_{sets}$), x-axes represnet the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$)"}
+Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$)."}
 
@@
 )
-
-# wireframe_legend <- packGrob(
-#   draw.colorkey(key = list(col = viridis::viridis, at = do.breaks(range(z),100))), 
-#   textGrob("RMSE", x = 0, y = 0.5, just = c("left", "centre")), 
-#   height = unit(2, "lines"), side = "top", dynamic = TRUE)
 
@@
 
-\pagebreak
+<br>
+
+# Appendix 1: age-year-space covariance
+
+The simulation applied in this paper was set-up to control covariance across ages, years and space.
+To do this we used a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
+As described in Appendix A in Cadigan [-@Cadigan2016], the age-year covariance can be broken down into a series of AR(1) processes.
+We integrate Matérn covariance into this series of equations:
+
+$$\xi_{a,y,s} \sim \left\{\begin{matrix} MVN\left (0, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a = 1, y = 1 & \\  MVN\left (\varphi_{year} \xi_{1, y - 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)} \boldsymbol{R}_s \right ) & a = 1, y > 1 & \\ MVN\left (\varphi_{age} \xi_{a - 1, 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a > 1, y = 1 & \\ MVN\left (\varphi_{year} \xi_{a, y - 1, s} + \varphi_{age} (\xi_{a - 1, y, s} - \varphi_{year} \xi_{a - 1, y - 1, s}), \sigma_\xi^2 \boldsymbol{R}_s \right ) & a > 1, y > 1 & \end{matrix}\right.$$
+
+Where $MVN$ indicates the multivariate normal distribution, $\sigma_\xi^2$ controls the variance of the process, $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$ control correlation in the age and year dimension and $\boldsymbol{R}_s$ is defined by Matérn correlation:
+
+$$\boldsymbol{R}_s = \frac{2^{1-\lambda}}{\Gamma(\lambda)} (\kappa \left |s_i - s_j \right|) K_{\lambda}(\kappa \left |s_i - s_j \right|)$$
+where $\left |s_i - s_j \right|$ is the Euclidean distance between two locations, $\Gamma$ is the gamma function, $K_{\lambda}$ denotes the modified Bessel function of the second kind, and $\lambda$ and $\kappa$ control the smoothness and scale of the spatial process [@Blangiardo2015].
+With this structure, simulated error is correlated across ages, years and space.
+
+# Appendix 2: results in the absence of intra-haul correlation
+
+In addition to the simulation run presented in body of this paper, which simulates surveys over a population that forms age-specific clusters, we also simulated surveys over a population that assumes the same spatial clustering across all age groups.
+There is therefore no intra-haul correlation in this series of surveys.
+Aside from this change to the spatial distribution settings, all other settings were kept the same.
+Results are presented below in Figures A1, A2 and A3.
+Aside from systematic underestimation in the low set density scenarios, there are little signs of clear bias in the stratified estimates.
+The bias present under the low set density scenario is likely related to the disproportionate sampling of the narrow shallow and deep strata in the survey grid.
+The lack of bias from the other scenarios help verify the issues introduced by age-specific clustering to stratified estimates of abundance at age.
+These results also indicate that little precision is gained by increasing length or age sampling effort beyond 100 lengths per set and 10 ages per length group, respectively.
+Given this result, it is probably not unreasonable to speculate that such sampling effort helps define the shape of the age distribution, but extra effort does little to improve estimates of abundance at age; such gains are only obtained by increasing set density.
+
+```{r appen-data}
+## this export holds the same distribution test
+load("analysis/cod_sim_exports/2018-07-16_test/test_output.RData")
+```
+
+```{r appen-age-fig, cache = TRUE, fig.width = 8, fig.height = 6}
+
+which_set_dens <- c(0.0005, 0.002, 0.01)
+which_len_caps <- c(10, 100, 1000)
+which_year <- 7
+
+d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
+d <- d[d$year == which_year & d$set_den %in% which_set_dens &
+         d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
+d$age[d$age > 8] <- 8 # aggregate beyond 8
+d <- d[, list(I_hat = sum(I_hat), I = sum(I)), 
+       by = c("year", "age", "sim", "survey", "set_den", "lengths_cap")]
+
+p <- fan_plot(d, x = "Age")
+cowplot::save_plot(plot = p, filename = "analysis/paper/figures/appen_true_vs_est_by_age.pdf",
+                   base_height = 6, base_width = 8)
+p
+
+```
+Figure A1 - Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
+The heavy dark line indicates the true population available to the survey.
+
+<br>
+
+```{r appen-year-fig, cache = TRUE, fig.width = 8, fig.height = 6}
+
+which_set_dens <- c(0.0005, 0.002, 0.01)
+which_len_caps <- c(10, 100, 1000)
+which_age <- 4
+
+d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
+d <- d[d$age == which_age & d$set_den %in% which_set_dens &
+       d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
+
+p <- fan_plot(d, x = "Year")
+cowplot::save_plot(plot = p, filename = "analysis/paper/figures/appen_true_vs_est_by_year.pdf",
+                   base_height = 6, base_width = 8)
+p
+
+```
+Figure A2 - Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
+The heavy dark line indicates the true trend in the population available to the survey.
+
+<br>
+
+```{r appen-error-fig, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide"}
+
+d <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
+# d <- d[d$set_den != 0.0005, ]
+d <- d[d$set_den %in% c(0.0005, 0.002, 0.01)]
+set_den <- sort(unique(d$set_den))
+d$set_den <- factor(d$set_den)
+options(scipen = 999)
+labs <- sapply(seq_along(set_den), function(i) as.expression(bquote(D[sets] == .(set_den[i]))), 
+               USE.NAMES = FALSE)
+zlabs <- list(at = pretty(d$RMSE), labels = si(pretty(d$RMSE)))
+
+pdf("analysis/paper/figures/appen_error_wireframe.pdf", height = 3, width = 8.5)
+
+p <- wireframe(RMSE ~  lengths_cap * ages_cap | set_den, data = d,
+          as.table = TRUE, drape = TRUE, col = "white", lwd = 0.5,
+          strip = strip.custom(factor.levels = labs),
+          scales = list(arrows = FALSE, cex = 0.7, z = zlabs),
+          col.regions = viridis::viridis(100),
+          screen = list(z = 230, x = 300), 
+          xlab = list(expression(M[lengths]), rot = 25), 
+          ylab = list(expression(M[ages]), rot = -15), 
+          zlab = list("RMSE", rot = 95),
+          colorkey = list(labels = zlabs, height = 0.6,
+                          axis.line = list(col = "darkgrey")),
+          par.settings = trellis_theme)
+p <- update(p, layout = c(3, 1))
+p
+grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc"))
+
+invisible(dev.off())
+
+p
+grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc"))
+
+```
+Figure A3 - Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
+Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$).
+
+<br>
 

### 2018-07-19 analysis/cod_sim.R

@@ -114,7 +114,7 @@ setMKLthreads() # turn hyperthreading on again
 
 
 ## visualize results
-load("analysis/cod_sim_exports/2018-07-15_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-07-16_test/test_output.RData")
 vis_sim(sim)
 
 

### 2018-07-20 analysis/paper/sim_survey_paper.Rmd
Start results section (add first paragraph)

@@
 Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain an estimate of abundance at age for each year for each simulation, $\hat{I}_{a,y,k}$.
+Intra-haul correlation (i.e.
+intraclass correlation) of the set catch was calculated by fitting a mixed-model to the lengths data with set as a random effect [fit using the lme4 package; @Bates2015]; variance estimates from this model were then plugged into the ICC(1) formula presented in McGraw and Wong [-@Mcgraw1996].
 We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance estimates from each survey:
@@
 catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$).
-All simulations and analyses were conducted using `r substr(version$version.string, 1, 15)` and the code has been developed into a package called `SimSurvey` to facilitate future work and collaboration (https://github.com/PaulRegular/SimSurvey).
+All simulations and analyses were conducted using `r substr(version$version.string, 1, 15)` [@R2017] and the code has been developed into a package called `SimSurvey` to facilitate future work and collaboration (https://github.com/PaulRegular/SimSurvey).
 
@@
 
-- add population trend plot?
+```{r icc, cache = TRUE}
+len_samp <- merge(sim$setdet, sim$samp, by = "set")
+len_samp <- len_samp[len_samp$sim == 1 & len_samp$measured, ]
+sim_icc <- round(icc(len_samp$length, len_samp$set), 1)
+sim_icc <- sprintf("%.2f", sim_icc)
+real_icc <- "0.70" # calculated in the file "imatate_cod_data.R"
+```
 
-\@ref(fig:sp-fig)
+A dynamic population that varies across ages, time and space is simulated with the parameter settings used here (Table \@ref(tab:par)).
+Figure \@ref(fig:sp-fig) depicts the aggregate distribution of the population in one year of the simulation and the corresponding catches of one stratified random survey over this population.
+Layered within this, of course, is the distributions of each age, as well as age and length sampling details.
+Though age-specific details are not depicted for simplicity, the general pattern of spatial aggregation apparent in Figure \@ref(fig:sp-fig) is similar across all ages.
+That is, all ages tend to aggregate in fairly dense clusters and low-density zones are fairly wide ranging.
+The survey, therefore, tends to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:sp-fig)).
+The samples within each successful set are also dynamic and rarely include all age groups.
+This is because only moderate correlation was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations.
+This results in an intra-haul correlation of approximately `r sim_icc`, which, in this case, is of equal magnitude to the intra-haul correlation observed across more than 20 years of survey data of 3Ps cod.
+
+Next discuss trend...
 
@@
 
-In addition to the simulation run presented in body of this paper, which simulates surveys over a population that forms age-specific clusters, we also simulated surveys over a population that assumes the same spatial clustering across all age groups.
+In addition to the run presented in body of this paper, which simulates surveys over a population that forms age-specific clusters, we also simulated surveys over a population that assumes the same spatial clustering across all age groups.
 There is therefore no intra-haul correlation in this series of surveys.

### 2018-08-22 analysis/paper/sim_survey_paper.Rmd
Start presentation

@@
 ```{r setup, include=FALSE}
+
+devtools::install_github("rstudio/rmarkdown@v1.10")
+
 knitr::opts_chunk$set(echo = FALSE,
@@
 These results indicate that length and age sampling efforts for some species(e.g.
-cod *Gadus morhua*) in the trawl survey conducted by regional fisheries management organizations can be scaled back without significant losses to precision.
+cod *Gadus morhua*) in trawl surveys conducted by regional fisheries management organizations can be scaled back without significant losses to precision.
 Time saved by reducing said sampling may afford more time to conduct additional sets and/or increase sampling efforts of other species.

### 2018-08-22 analysis/presentation/sim_survey_presentation.Rmd
Start presentation

---
title: "Using simulation to optimize the sampling design of fisheries-independent trawl surveys"
author: "Paul M. Regular & Fran Mowbray"
date: "September XX, 2018"
output:
  ioslides_presentation:
    widescreen: true
    incremental: true
    logo: dfo_logo.png
    css: dfo_style.css
---

<style>

@import url('https://fonts.googleapis.com/css?family=Montserrat');

.forceBreak { -webkit-column-break-after: always; break-after: column; }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      dev = "png",
                      dpi = 300)
knitr::opts_knit$set(root.dir = "../..")
```

```{r data, include = FALSE}
library(SimSurvey)
library(plotly)
library(raster)
library(lattice)
library(viridis)
library(data.table)
# load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
```

## Fisheries-independent trawl surveys

<div class="columns-2">

- Conducted by many RFMOs
- Becoming increasingly important in stock assessment
- Very costly 💰
- General goal:
    - Maximize information & minimize expenses

<p class="forceBreak"></p>

<br>
<img src="graphics/teleost_twitter_crop.jpg" width="400"/>
<font size="1"> 
&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">@CoastGuardCAN</a>
</font>

</div>

## Fisheries-independent trawl surveys

<div class="columns-2">

How much survey effort is enough?

1. Sub-sampling
    - Truth unknown
    - Small sample size effect
2. Simulation
    - Truth controlled
    - Reality?

<p class="forceBreak"></p>

<br>
<img src="graphics/teleost_twitter_crop.jpg" width="400"/>
<font size="1"> 
&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">@CoastGuardCAN</a>
</font>

</div>

## **Reality is complicated** {data-background=graphics/school_light.png data-background-size=cover}

- **Spatial aggregation**
- **Habitat preferences**
- **Survey design**

### 2018-08-23 analysis/presentation/sim_survey_presentation.Rmd
Clue up draft of intro and start drafting the methods in the presentation

@@
 author: "Paul M. Regular & Fran Mowbray"
-date: "September XX, 2018"
+date: "September 6^th^, 2018"
 output:
@@
     css: dfo_style.css
+bibliography: references.bib
 ---
@@
 # load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
+
+boat_txt <- '<br>
+<img src="graphics/teleost_twitter_crop.jpg" width="400"/>
+<font size="1"> 
+<br>
+&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">Canadian Coast Guard | Twitter</a>
+</font>'
+
 ```
 
-## Fisheries-independent trawl surveys
-
-<div class="columns-2">
+## Fisheries-independent trawl surveys {.columns-2}
 
@@
 
+```{r results = "asis"}
+cat(boat_txt)
+```
+
+## Fisheries-independent trawl surveys {.columns-2}
+
+- How much survey effort is enough?
+    - Difficult to answer
+        - Multi-stage sampling
+        - Intra-haul correlation
+
+<p class="forceBreak"></p>
+
+```{r results = "asis"}
+cat(boat_txt)
+```
+
+## Multi-stage sampling
+
+Surveys are generally designed to:
+
+<font size="8"> 1 </font> | <font size="8"> 2 </font> | <font size="8"> 3 </font>
+:----: | :----: | :----:
+Conduct sets at multiple locations | Sub-sample catch for length measurements | Sub-sampling length groups for age determination
+<img src="graphics/trawl_survey_crop.jpg" height="180"/> | <img src="graphics/cod_length.jpg" height="120"/> | <img src="graphics/Otoliths2_crop.jpg" height="140"/>
+
+## Intra-haul correlation
+
+Sampling clusters of fish that have similar characteristics (e.g. length)
+
+- Results in samples with positive intra-haul correlation
+    - Reduces effective sample size
+
+<center> <img src="graphics/plos_figure_crop.png" height="300"/> 
 <br>
-<img src="graphics/teleost_twitter_crop.jpg" width="400"/>
 <font size="1"> 
-&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">@CoastGuardCAN</a>
+Graphic from <a href="http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002915">Tunstrøm et al (2013) PLoS Comput Biol 9(2): e1002915.</a> 
 </font>
+</center>
 
-</div>
+## Fisheries-independent trawl surveys {.columns-2}
 
-## Fisheries-independent trawl surveys
-
-<div class="columns-2">
-
-How much survey effort is enough?
-
-1. Sub-sampling
+- How much survey effort is enough?
+    - Attempt to answer using:
+        - Sub-sampling
+            - Based in reality
             - Truth unknown
-    - Small sample size effect
-2. Simulation
-    - Truth controlled
+        - Simulation
+            - Truth known
             - Reality?
@@
 
-<br>
-<img src="graphics/teleost_twitter_crop.jpg" width="400"/>
-<font size="1"> 
-&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">@CoastGuardCAN</a>
-</font>
-
-</div>
+```{r results = "asis"}
+cat(boat_txt)
+```
 
@@
 
+- **Dynamic abundance**
 - **Spatial aggregation**
-- **Habitat preferences**
 - **Survey design**
+- **_Objective: build a simulation that emulates reality_**
+
+## Step 1: Simulate abundance
+
+- Common cohort model
+    - $N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$
+- First year filled via exponential decay
+    -  $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$
+- Random-walk in recruitment
+    - $log(N_{1,y}) = log(\mu_r) + \epsilon_{y}$, where $\epsilon_{y} \sim N(\epsilon_{y - 1}, \sigma_{r}^{2})$
+- Total mortality correlated across ages and years
+    - $log(Z_{a,y}) = log(\mu_Z) + \delta_{a,y}$, where $\delta_{a,y} \sim N(0, \Sigma_{a,y})$
+        - Using covariance, $\Sigma_{a,y}$, described in Cadigan [-@Cadigan2016]
+
+## References

### 2018-08-24 analysis/presentation/sim_survey_presentation.Rmd
Start to describe simulation of spatial distribution

@@
                       warning = FALSE, 
-                      dev = "png",
-                      dpi = 300)
+                      fig.height = 4.5, 
+                      fig.width = 10)
 knitr::opts_knit$set(root.dir = "../..")
@@
 library(data.table)
-# load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
 
@@
 - **_Objective: build a simulation that emulates reality_**
+- **_Roughly based simulation on 3Ps Cod_**
 
-## Step 1: Simulate abundance
+## Simulate abundance
 
@@
 
+## Simulate abundance
+
+Trend in total abundance
+
+<dev class = "centered"> 
+<!-- There is a bug somewhere...had to contain plotly plots in a dev to get it working -->
+```{r}
+plot_trend(sim) %>% 
+  layout(plot_bgcolor = "transparent",
+         paper_bgcolor = "transparent",
+         margin = list(l = 70, r = 0, t = 0, b = 40, pad = 0),
+         font = list(size = 14))
+```
+</dev>
+
+## Simulate abundance
+
+Trend in abundance at age
+
+<dev class = "centered">
+```{r}
+plot_surface(sim) %>% 
+  layout(plot_bgcolor = "transparent",
+         paper_bgcolor = "transparent",
+         margin = list(l = 0, r = 0, t = 0, b = 0, pad = 0),
+         font = list(size = 14))
+```
+</dev>
+
+## Simulate spatial aggregation
+
+- Generate a square grid of $s$ locations, resolution $R$ and a depth gradient
+- Distribute population through space and add spatial-temporal noise
+    - $log(N_{a,y,s}) = log(N_{a,y}) + \eta_{a,y,s}$
+    - $\eta_{a,y,s}$ includes parabolic relationship with depth and covariance between ages, years and space
+        - $\eta_{a,y,s} = \frac{(d_s - \mu_{d}) ^ 2}{2 \sigma_d^2} + \xi_{a,y,s}$
+        - $\xi_{a,y,s}$ is a combination of Matérn covariance and the age-year covariance described in Cadigan [-@Cadigan2016]
+
+## Todo
+
+- Try and figure out the bug that prevents the plotly plots from working in this part of the slideshow (alternate fix to containing it in a device)
+- Figure out how to use the Google Font with plotly
+- Set-up global layout arguents somehow
+
 ## References

### 2018-08-24 R/plot_sim.R

@@ -35,7 +35,7 @@ plot_trend <- function(sim, col = viridis::viridis(1), ...) {
 #' @export
 #' @rdname plot_trend
 plot_surface <- function(sim, mat = "N", ...) {
-  plot_ly(x = sim$ages, y = sim$years, z = sim[[mat]], type = "surface",
+  plot_ly(x = sim$ages, y = sim$years, z = t(sim[[mat]]), type = "surface",
           ...) %>%
     layout(
       scene = list(

### 2018-08-28 analysis/presentation/sim_survey_presentation.Rmd
Add a plot for describing the spatial simulation to the presentation

@@
                       warning = FALSE, 
-                      fig.height = 4.5, 
-                      fig.width = 10)
+                      fig.width = 10, 
+                      fig.height = 5)
 knitr::opts_knit$set(root.dir = "../..")
@@
 library(plotly)
+library(crosstalk)
 library(raster)
@@
 </font>'
+
+## wrapper for layout with some different defaults for this slideshow
+tight_layout <- function(p, 
+                         plot_bgcolor = "transparent",
+                         paper_bgcolor = "transparent",
+                         margin = list(l = 0, r = 0, t = 0, b = 0, pad = 0),
+                         font = list(size = 14), ..., data = NULL) {
+  layout(p, plot_bgcolor = plot_bgcolor, paper_bgcolor = paper_bgcolor,
+         margin = margin, font = font, ..., data = data)
+}
 
@@
 
-## Simulate abundance
+## Simulate abundance {.tighter}
 
-Trend in total abundance
+### Trend in total abundance
 
-<dev class = "centered"> 
+<dev class = "left">
 <!-- There is a bug somewhere...had to contain plotly plots in a dev to get it working -->
@@
 plot_trend(sim) %>% 
-  layout(plot_bgcolor = "transparent",
-         paper_bgcolor = "transparent",
-         margin = list(l = 70, r = 0, t = 0, b = 40, pad = 0),
-         font = list(size = 14))
+  tight_layout(margin = list(l = 70, r = 0, t = 0, b = 40, pad = 0))
 ```
@@
 
-## Simulate abundance
+## Simulate abundance {.tighter}
 
-Trend in abundance at age
+### Trend in abundance at age
 
-<dev class = "centered">
+<dev class = "left">
 ```{r}
-plot_surface(sim) %>% 
-  layout(plot_bgcolor = "transparent",
-         paper_bgcolor = "transparent",
-         margin = list(l = 0, r = 0, t = 0, b = 0, pad = 0),
-         font = list(size = 14))
+plot_surface(sim) %>% tight_layout()
 ```
@@
 
+## Simulate spatial aggregation  {.tighter}
+
+### Square grid with depth gradient
+
+<dev class = "left">
+```{r}
+plot_ly(data = sim$grid_xy, x = ~x, y = ~y, z = ~depth, type = "heatmap") %>%
+  colorbar(title = "Depth") %>%
+  tight_layout(yaxis = list(title = "", scaleanchor = "x", zeroline = FALSE, 
+                            showgrid = FALSE, showticklabels = FALSE,
+                            tickcolor = "transparent"),
+               xaxis = list(title = "", zeroline = FALSE, 
+                            showgrid = FALSE, showticklabels = FALSE,
+                            tickcolor = "transparent"))
+```
+</dev>
+
+## Simulate spatial aggregation {.tighter}
+
+### Distribution for ages 1-6 in years 1-6
+
+<dev class = "left">
+```{r}
+plot_distribution_slider(sim, ages = 1:6, years = 1:6) %>% 
+  tight_layout(margin = list(l = 0, r = 0, t = 40, b = 0, pad = 0))
+```
+</dev>
+
 ## Todo
@@
 - Set-up global layout arguents somehow
+- Figure out how to make a sub-title for a slide below main title
+- Reverse color ramp in grid figure
+- Frame by year for spatial abundance plot (consider fixing one year, animate age / fix age and animate year)
 

### 2018-08-28 R/plot_sim.R

@@ -57,7 +57,7 @@ plot_distribution <- function(sim, ages = 1, years = 1,
     showline = FALSE,
     showticklabels = FALSE,
     showgrid = FALSE,
-    tickcolor = toRGB("white")
+    tickcolor = "transparent"
   )
   yax <- c(scaleanchor = "x", xax)
 
@@ -76,6 +76,60 @@ plot_distribution <- function(sim, ages = 1, years = 1,
     }
   }
   subplot(p, nrows = length(ages), shareX = TRUE, shareY = TRUE)
+
+}
+
+
+#' @export
+#' @rdname plot_trend
+plot_distribution_slider <- function(sim, ages = 1, years = 1) {
+
+  xax <- list(
+    title = "",
+    zeroline = FALSE,
+    showline = FALSE,
+    showticklabels = FALSE,
+    showgrid = FALSE,
+    tickcolor = "transparent"
+  )
+  yax <- c(scaleanchor = "x", xax)
+
+  d <- merge(sim$grid_xy, sim$sp_N[age %in% ages & year %in% years, ], by = "cell")
+  d$ay <- paste(d$age, d$year, sep = "-")
+  split_d <- split(d, d$ay)
+  split_d <- lapply(split_d, function(.) {
+    xtabs(N ~ x + y, data = ., subset = NULL)
+  })
+  x <- sort(unique(d$x))
+  y <- sort(unique(d$y))
+
+  p <- plot_ly(x = ~x, y = ~y)
+  visible <- rep(TRUE, length(split_d))
+  showscale <- rep(FALSE, length(split_d))
+  showscale[1] <- TRUE
+  steps <- list()
+  for (i in seq_along(split_d)) {
+    vis <- rep(FALSE, length(split_d))
+    vis[i] <- TRUE
+    p <- p %>% add_trace(type = "contour",
+                         z = split_d[[i]],
+                         visible = i == 1,
+                         showscale = vis,
+                         name = names(split_d)[i],
+                         colorbar = list(title = "N")) %>%
+      layout(xaxis = xax, yaxis = yax)
+
+    steps[[i]] <- list(args = list(list(visible = vis,
+                                        showscale = vis)),
+                       method = "update",
+                       label = names(split_d)[i])
+  }
+
+  p %>%
+    layout(sliders = list(list(
+      currentvalue = list(prefix = "Age-Year: "),
+      steps = steps
+    )))
 
 }
 

### 2018-08-28 analysis/presentation/sim_survey_presentation.Rmd
Describe survey portion of the simulation in the presentation

@@
                          margin = list(l = 0, r = 0, t = 0, b = 0, pad = 0),
-                         font = list(size = 14), ..., data = NULL) {
+                         font = list(size = 14, family = "'Montserrat', sans-serif"), 
+                         ..., data = NULL) {
   layout(p, plot_bgcolor = plot_bgcolor, paper_bgcolor = paper_bgcolor,
@@
 
+## Simulate survey
+
+- Split grid into $N_{div}$ divisions and $N_{strat}$ depth-based strata
+- Set allocation defined by specifying set density, $D_{sets}$
+- Binomial sampling was used to simulate catch:
+    - $n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{trawl}}{A_{cell}} q_a )$
+    - Accounts for trawl size relative to cell size, $\frac{A_{trawl}}{A_{cell}}$, and catchability, $q_a$ (controled using a logistic curve)
+
+## Simulate survey
+
+- Lengths simulated using the original von Bertalanffy growth curve [@Cailliet2006]:
+    - $log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$, where $\varepsilon \sim N(0, \sigma_{l}^2)$
+- Sub-sampling is then conducted
+    - A maximum number of lengths are measured per set, $M_{lengths}$
+    - A maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division
+
+## Simulate survey {.tighter}
+
+### Catch in year 1 survey simulation 1
+
+<dev class = "left">
+```{r}
+plot_samp_dist(sim) %>% 
+  tight_layout(margin = list(l = 0, r = 0, t = 0, b = 40, pad = 0))
+```
+</dev>
+
+## Acknowledgements
+
+- Todo
+
 ## Todo
@@
 - Try and figure out the bug that prevents the plotly plots from working in this part of the slideshow (alternate fix to containing it in a device)
-- Figure out how to use the Google Font with plotly
-- Set-up global layout arguents somehow
-- Figure out how to make a sub-title for a slide below main title
-- Reverse color ramp in grid figure
-- Frame by year for spatial abundance plot (consider fixing one year, animate age / fix age and animate year)
 
-## References
+## References {.smaller}

### 2018-08-28 R/plot_sim.R

@@ -281,7 +281,7 @@ plot_samp_dist <- function(sim, which_year = 1, which_sim = 1,
     showline = FALSE,
     showticklabels = FALSE,
     showgrid = FALSE,
-    tickcolor = toRGB("white")
+    tickcolor = "transparent"
   )
   yax <- c(scaleanchor = "x", xax)
 
@@ -361,7 +361,7 @@ plot_total_abundance <- function(sim, which_year = 1, main = "") {
     showline = FALSE,
     showticklabels = FALSE,
     showgrid = FALSE,
-    tickcolor = toRGB("white")
+    tickcolor = "transparent"
   )
   yax <- c(scaleanchor = "x", xax)
 
@@ -406,7 +406,7 @@ plot_set_catch <- function(sim, which_year = 1, which_sim = 1, main = "") {
     showline = FALSE,
     showticklabels = FALSE,
     showgrid = FALSE,
-    tickcolor = toRGB("white")
+    tickcolor = "transparent"
   )
   yax <- c(scaleanchor = "x", xax)
 

### 2018-08-28 analysis/presentation/sim_survey_presentation.Rmd
Finish drafting methods portion of the presentation

@@
 - **_Objective: build a simulation that emulates reality_**
-- **_Roughly based simulation on 3Ps Cod_**
 
@@
 ```{r}
-plot_ly(data = sim$grid_xy, x = ~x, y = ~y, z = ~depth, type = "heatmap") %>%
+plot_ly(data = sim$grid_xy, x = ~x, y = ~y, color = ~depth, 
+        symbol = I(15), size = I(4.5), name = "grid", type = "scatter", 
+        text = ~depth) %>%
   colorbar(title = "Depth") %>%
@@
 
+## Settings and analysis {.build}
+
+- Roughly based simulation on 3Ps Cod
+- A series of 175 surveys were tested
+
+|      **Parameter name** |      **Symbol** |                                       **Setting** |
+|:----------------------- | :-------------- | :------------------------------------------------ |
+|             Set density |      $D_{sets}$ |    0.0005, 0.001, 0.002, 0.005, 0.01 sets / km^2^ |
+|  Length sampling effort |   $M_{lengths}$ |       5, 10, 20, 50, 100, 500, 1000 lengths / set |
+|     Age sampling effort |      $M_{ages}$ |   2, 5, 10, 20, 50 ages / length group / division |
+
+## Settings and analysis
+
+- Each survey was run 1000 times over the same population
+- Abundance estimates, $\hat{I}_{a,y,k}$, were obtained using a stratified analysis [@Smith1981]
+- Used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance estimates from each survey:
+    - $RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})}{N_a N_y N_k}}$
+    - Where $I_{a,y}$ is the true abundance available to the survey
+- Powered by R and worked into a package (https://github.com/PaulRegular/SimSurvey)
+
 ## Acknowledgements

### 2018-08-30 analysis/presentation/sim_survey_presentation.Rmd
Add plot_ly based fan plot for true vs est comps

@@
 plot_trend(sim) %>% 
-  tight_layout(margin = list(l = 70, r = 0, t = 0, b = 40, pad = 0))
+  tight_layout(yaxis = list(rangemode = "tozero"),
+               margin = list(l = 70, r = 0, t = 0, b = 40, pad = 0))
 ```

### 2018-08-30 R/plot_sim.R

@@ -16,6 +16,9 @@
 #' @param which_sim      Subset to specific sim
 #' @param max_sims       Maximum number of sims to plot
 #' @param facet_by       Facet plot by "age" or "year"?
+#' @param survey         Subset data to one or more surveys.
+#' @param frame          Frame plots by "age" or "year"?
+#' @param quants         Quantile intervals to display on fan plot
 #' @param col            Plot color
 #' @param cols           Plot colors
 #' @param main           Plot title
@@ -436,7 +439,7 @@ plot_set_catch <- function(sim, which_year = 1, which_sim = 1, main = "") {
     layout(xaxis = xax, yaxis = yax, legend = list(x = 1, y = 0.85),
            annotations = list(
              list(text = "n", xref = "paper", yref = "paper",
-                              x = 1.09, y = 0.9, showarrow = FALSE),
+                  x = 1.09, y = 0.9, showarrow = FALSE),
              list(text = main, xref = "paper", yref = "paper",
                   x = 0.08, y = 1, showarrow = FALSE)
            ),
@@ -444,3 +447,99 @@ plot_set_catch <- function(sim, which_year = 1, which_sim = 1, main = "") {
   set_plot
 
 }
+
+
+## helper function for making a fan plot
+## column named prob is expected in d
+.add_fan <- function(p, x = NULL, ymin = NULL, ymax = NULL, ..., data = NULL) {
+  if (is.null(data)) data <- plotly_data(p)
+  cols <- toRGB(viridis::viridis(nlevels(data$prob) - 1),
+                alpha = seq(0.1, 1, length = nlevels(data$prob) - 1))
+  cols[length(cols) + 1] <- toRGB(viridis::viridis(1), alpha = 1)
+  names(cols) <- levels(data$prob)
+  for (q in levels(data$prob)) {
+    p <- p %>% add_ribbons(data = data[data$prob == q, ],
+                           ymin = ymin, ymax = ymax, name = q,
+                           fillcolor = cols[q],
+                           line = list(color = cols[q],
+                                       width = ifelse(q == "True", 1.6, 0)),
+                           ...)
+  }
+  p
+}
+
+#' @export
+#' @rdname plot_trend
+plot_true_vs_est_fan <- function(sim, surveys = 1, frame = "year",
+                                 quants = seq(90, 10, by = -10)) {
+
+  if (frame == "year") {
+    f <- ~year
+    x <- ~age
[…]

### 2018-08-30 analysis/presentation/sim_survey_presentation.Rmd
Add another true vs est plotting option using crosstalk::filter_select

@@
                       fig.width = 10, 
-                      fig.height = 5)
+                      fig.height = 5.2)
 knitr::opts_knit$set(root.dir = "../..")
@@
 
+## Age strat
+
+### Some results
+
+<dev class = "left">
+```{r}
+plot_age_strat_fan(sim, surveys = 1:5, 
+                   plot_bgcolor = "transparent",
+                   paper_bgcolor = "transparent",
+                   font = list(size = 14, family = "'Montserrat', sans-serif"))
+```
+</dev>
+
 ## Acknowledgements

### 2018-08-30 R/plot_sim.R

@@ -16,8 +16,8 @@
 #' @param which_sim      Subset to specific sim
 #' @param max_sims       Maximum number of sims to plot
 #' @param facet_by       Facet plot by "age" or "year"?
+#' @param select_by      Select plot by "age" or "year"?
 #' @param survey         Subset data to one or more surveys.
-#' @param frame          Frame plots by "age" or "year"?
 #' @param quants         Quantile intervals to display on fan plot
 #' @param col            Plot color
 #' @param cols           Plot colors
@@ -449,44 +449,15 @@ plot_set_catch <- function(sim, which_year = 1, which_sim = 1, main = "") {
 }
 
 
-## helper function for making a fan plot
-## column named prob is expected in d
-.add_fan <- function(p, x = NULL, ymin = NULL, ymax = NULL, ..., data = NULL) {
-  if (is.null(data)) data <- plotly_data(p)
-  cols <- toRGB(viridis::viridis(nlevels(data$prob) - 1),
-                alpha = seq(0.1, 1, length = nlevels(data$prob) - 1))
-  cols[length(cols) + 1] <- toRGB(viridis::viridis(1), alpha = 1)
-  names(cols) <- levels(data$prob)
-  for (q in levels(data$prob)) {
-    p <- p %>% add_ribbons(data = data[data$prob == q, ],
-                           ymin = ymin, ymax = ymax, name = q,
-                           fillcolor = cols[q],
-                           line = list(color = cols[q],
-                                       width = ifelse(q == "True", 1.6, 0)),
-                           ...)
-  }
-  p
-}
-
 #' @export
 #' @rdname plot_trend
-plot_true_vs_est_fan <- function(sim, surveys = 1, frame = "year",
-                                 quants = seq(90, 10, by = -10)) {
-
-  if (frame == "year") {
-    f <- ~year
-    x <- ~age
-    flab <- "Year: "
-    xlab <- "Age"
-  } else {
-    f <- ~age
-    x <- ~year
-    flab <- "Age: "
-    xlab <- "Year"
-  }
+plot_age_strat_fan <- function(sim, surveys = 1, select_by = "year",
+                               quants = seq(90, 10, by = -10),
+                               ...) {
 
   d <- sim$age_strat_error
   sub_d <- d[survey %in% surveys, ]
+  true <- sub_d[sim == 1, list(year, age, survey, I)]
 
   ## Calculate a series of quantiles
   ints <- lapply(quants, function(q) {
[…]

### 2018-08-30 R/plot_sim.R
Add fan plot function for total strat estimates

@@ -488,6 +488,7 @@ plot_age_strat_fan <- function(sim, surveys = 1, select_by = "year",
 
   crosstalk::bscols(
     list(
+      htmltools::HTML("<br>"),
       f,
       filter_select("set_den", "Set density", shared_ints, ~set_den, multiple = FALSE),
       filter_select("lengths_cap", "Lengths cap", shared_ints, ~lengths_cap, multiple = FALSE),
@@ -502,7 +503,54 @@ plot_age_strat_fan <- function(sim, surveys = 1, select_by = "year",
                 showlegend = FALSE) %>%
       layout(yaxis = list(title = "Abundance index",
                           range = c(min(ints$lower), max(ints$upper))),
-             xaxis = list(title = xlab)),
+             xaxis = list(title = xlab), ...),
+    widths = c(3, NA)
+  )
+
+}
+
+
+#' @export
+#' @rdname plot_trend
+plot_total_strat_fan <- function(sim, surveys = 1,
+                                 quants = seq(90, 10, by = -10),
+                                 ...) {
+
+  d <- sim$total_strat_error
+  sub_d <- d[survey %in% surveys, ]
+  true <- sub_d[sim == 1, list(year, survey, I)]
+
+  ## Calculate a series of quantiles
+  ints <- lapply(quants, function(q) {
+    sub_d[, list(prob = paste0(q, "%"),
+                 lower = quantile(I_hat, prob = (1 - q / 100) / 2),
+                 upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)),
+          by = c("year", "survey")]
+  })
+  ints <- rbindlist(ints)
+  ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
+
+  ints <- merge(sim$surveys, ints, by = "survey", all.y = TRUE)
+  ints$lab <- paste(formatC(ints$set_den, format = "fg"),
+                    ints$lengths_cap, ints$ages_cap, sep = "-")
+  ints <- merge(ints, true, by = c("year", "survey"))
+
+  shared_ints <- crosstalk::SharedData$new(ints)
+
+  crosstalk::bscols(
+    list(
+      htmltools::HTML("<br>"),
+      filter_select("set_den", "Set density", shared_ints, ~set_den, multiple = FALSE)
+    ),
+    plot_ly(data = shared_ints, x = ~year, height = "100%") %>%
+      add_ribbons(ymin = ~lower, ymax = ~upper, line = list(width = 0),
[…]

### 2018-08-30 analysis/presentation/sim_survey_presentation.Rmd
Add new truth vs estimate plots to presentation

@@
 
-## Age strat
+## Truth vs. estimate
 
-### Some results
+### Total abundance
 
@@
 ```{r}
-plot_age_strat_fan(sim, surveys = 1:5, 
-                   plot_bgcolor = "transparent",
-                   paper_bgcolor = "transparent",
+plot_total_strat_fan(sim, surveys = 1:5, 
+                     plot_bgcolor = "transparent", paper_bgcolor = "transparent",
+                     font = list(size = 14, family = "'Montserrat', sans-serif"))
+```
+</dev>
+
+## Truth vs. estimate
+
+### Abundance at age
+
+<dev class = "left">
+```{r}
+surveys <- sim$surveys[set_den %in% c(0.0005, 0.001, 0.002, 0.01) &
+                         lengths_cap %in% c(10, 100, 500, 1000) &
+                         ages_cap %in% c(5, 10, 50), ]
+plot_age_strat_fan(sim, surveys = surveys$survey, 
+                   years = 1:5, ages = sim$ages, select_by = "year",
+                   plot_bgcolor = "transparent", paper_bgcolor = "transparent",
+                   font = list(size = 14, family = "'Montserrat', sans-serif"))
+```
+</dev>
+
+## Truth vs. estimate
+
+### Abundance at age
+
+<dev class = "left">
+```{r}
+plot_age_strat_fan(sim, surveys = surveys$survey, 
+                   ages = 2:6, years = sim$years, select_by = "age",
+                   plot_bgcolor = "transparent", paper_bgcolor = "transparent",
                    font = list(size = 14, family = "'Montserrat', sans-serif"))

### 2018-08-30 R/plot_sim.R

@@ -451,12 +451,13 @@ plot_set_catch <- function(sim, which_year = 1, which_sim = 1, main = "") {
 
 #' @export
 #' @rdname plot_trend
-plot_age_strat_fan <- function(sim, surveys = 1, select_by = "year",
+plot_age_strat_fan <- function(sim, surveys = 1:5, years = 1:10,
+                               ages = 1:10, select_by = "year",
                                quants = seq(90, 10, by = -10),
                                ...) {
 
   d <- sim$age_strat_error
-  sub_d <- d[survey %in% surveys, ]
+  sub_d <- d[survey %in% surveys & year %in% years & age %in% ages, ]
   true <- sub_d[sim == 1, list(year, age, survey, I)]
 
   ## Calculate a series of quantiles
@@ -488,7 +489,7 @@ plot_age_strat_fan <- function(sim, surveys = 1, select_by = "year",
 
   crosstalk::bscols(
     list(
-      htmltools::HTML("<br>"),
+      htmltools::div(style = htmltools::css(height = "10px")), # small margin
       f,
       filter_select("set_den", "Set density", shared_ints, ~set_den, multiple = FALSE),
       filter_select("lengths_cap", "Lengths cap", shared_ints, ~lengths_cap, multiple = FALSE),
@@ -512,7 +513,7 @@ plot_age_strat_fan <- function(sim, surveys = 1, select_by = "year",
 
 #' @export
 #' @rdname plot_trend
-plot_total_strat_fan <- function(sim, surveys = 1,
+plot_total_strat_fan <- function(sim, surveys = 1:5,
                                  quants = seq(90, 10, by = -10),
                                  ...) {
 
@@ -539,7 +540,7 @@ plot_total_strat_fan <- function(sim, surveys = 1,
 
   crosstalk::bscols(
     list(
-      htmltools::HTML("<br>"),
+      htmltools::div(style = htmltools::css(height = "10px")), # small margin
       filter_select("set_den", "Set density", shared_ints, ~set_den, multiple = FALSE)
     ),
     plot_ly(data = shared_ints, x = ~year, height = "100%") %>%

### 2018-08-31 analysis/presentation/sim_survey_presentation.Rmd
Add plots for bias test run to presentation (plus minor tweaks)

@@
 
-<font size="8"> 1 </font> | <font size="8"> 2 </font> | <font size="8"> 3 </font>
+<font size="8"> &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; 1 </font> | <font size="8"> &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; 2 </font> | <font size="8"> &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; 3 </font>
 :----: | :----: | :----:
@@
 <dev class = "left">
-<!-- There is a bug somewhere...had to contain plotly plots in a dev to get it working -->
 ```{r}
@@
 
+## Truth vs. estimate {.tighter}
+
+### RMSE
+
+<dev class = "left">
+```{r}
+plot_error_surface(sim) %>% tight_layout()
+```
+</dev>
+
+## Truth vs. estimate {.smaller}
+
+### Recap
+
+- **Sets = primary sampling unit**
+    - Only way to improve stratified estimates of total abundance
+    - Best way to improve stratified estimates of abundance at age
+- **Increasing length sampling does not always improve estimates**
+    - At low set densities, more is worse
+    - At high set densities, more is better
+- **Sampling effort affects both bias and percision**
+    - Stratified estimates are supposed to be unbiased?!
+    - Likely related to age-specific clustering
+
+## Check bias
+
+### Turned off age-specific clustering
+
+<dev class = "left">
+```{r}
+load("analysis/cod_sim_exports/2018-07-16_test/test_output.RData")
+plot_age_strat_fan(sim, surveys = surveys$survey,
+                   ages = 2:6, years = sim$years, select_by = "age",
+                   plot_bgcolor = "transparent", paper_bgcolor = "transparent",
+                   font = list(size = 14, family = "'Montserrat', sans-serif"))
+```
+</dev>
+
+## Check bias {.tighter}
+
+### Turned off age-specific clustering
+
+<dev class = "left">
+```{r}
+plot_error_surface(sim) %>% tight_layout()
+```
+</dev>
+
 ## Acknowledgements
@@
 
-## Todo
-
-- Try and figure out the bug that prevents the plotly plots from working in this part of the slideshow (alternate fix to containing it in a device)
+<!-- There is a conflict somewhere...had to contain plotly plots in a dev to get it working and I had to add a bunch of spaces to fudge centering in one of the tables -->
 

### 2018-08-31 analysis/presentation/sim_survey_presentation.Rmd
Add interpretations of the results to the presentation and start conclusions

@@
   ioslides_presentation:
+    mathjax: local
+    self_contained: false
     widescreen: true
@@
     - $n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{trawl}}{A_{cell}} q_a )$
-    - Accounts for trawl size relative to cell size, $\frac{A_{trawl}}{A_{cell}}$, and catchability, $q_a$ (controled using a logistic curve)
+    - Accounts for trawl size relative to cell size, $\frac{A_{trawl}}{A_{cell}}$, and catchability, $q_a$ (controlled using a logistic curve)
 
@@
 ```{r}
-plot_error_surface(sim) %>% tight_layout()
+slide_font <- list(size = 14, color = rgb(121, 121, 121, maxColorValue = 255))
+surface_font <- list(titlefont = slide_font, 
+                     tickfont = slide_font)
+surface_scene <- list(
+  xaxis = surface_font,
+  yaxis = surface_font,
+  zaxis = surface_font
+)
+plot_error_surface(sim) %>% 
+  tight_layout(scene = surface_scene,
+               margin = list(l = 0, r = 0, t = 40, b = 0, pad = 0))
 ```
@@
 ```{r}
+main_sim <- sim
 load("analysis/cod_sim_exports/2018-07-16_test/test_output.RData")
@@
 ```{r}
-plot_error_surface(sim) %>% tight_layout()
+plot_error_surface(sim) %>% 
+  tight_layout(scene = surface_scene,
+               margin = list(l = 0, r = 0, t = 40, b = 0, pad = 0))
 ```
 </dev>
+
+## Check bias
+
+### Turned off age-specific clustering
+
+<br>
+
+>- Increases in secondary sampling does little to improve estimates
+>- Bias is negligible in this scenario
+
+## Interpretation
+
+Why is there bias? Why is extra sub-sampling ineffective or detrimental?
+
+- With age-specific clustering, fish of different ages are not encountered with equal probability
+    - A survey with few sets will be less likely to encounter groups in a tight cluster
+        - Survey-level length / age distribution ∴ more likely to be incomplete 
+            - Increasing length / age sampling cannot improve perception of the population-level distribution
+
+## Interpretation {.tighter .smaller}
+
+Ages caught and sampled in year 18 from one survey scenario:
+
+>- $D_{sets} = 0.002, M_{lengths} = 500, M_{ages} = 10$
+
+<dev class = "left">
+```{r, fig.height = 4.8}
+
+samp <- merge(main_sim$setdet, main_sim$samp, by = "set")
+samp <- samp[sim == 1 & year == 18, ]
+caught <- as.data.frame(table(samp$age))
+caught <- data.frame(group = "Caught", age = caught$Var1, freq = caught$Freq,
+                     prop = caught$Freq / sum(caught$Freq))
+measured <- as.data.frame(table(samp$age[samp$measured]))
+measured <- data.frame(group = "Sampled", age = measured$Var1, freq = measured$Freq,
+                       prop = measured$Freq / sum(measured$Freq))
+true <- data.frame(group = "True", age = sim$ages, freq = sim$I[, 1],
+                   prop = sim$I[, 1] / sum(sim$I[, 1]))
+d <- rbind(caught, measured, true)
+
+plot_ly(data = d) %>% 
+  add_lines(x = ~age, y = ~prop, color = ~group,
+            colors = rev(viridis::viridis(3))) %>% 
+  tight_layout(xaxis = list(title = "Age"),
+               yaxis = list(title = "Proportion"),
+               margin = list(r = 60))
+
+```
+</dev>
+
+## Conclusions
+
+- Increasing sub-sampling effort may be an inefficient means of improving population estimates
+    - Especially problimatic if sub-sampling affects primary sampling effort
+
+## Potential future directions
+
+- Taylor to other species and assess sampling protocol
+- Add a time component to assess trade-offs / efficiency
+- Assess the effects of
+    - Missing strata
+    - Variable trawl area
+- Use this as a platform to test spatial models
+- Use as an operating model for an MSE
+- Others?
 

### 2018-09-04 analysis/paper/sim_survey_paper.Rmd
Finish drafting presentation

@@
 
-- Feedback and advice: Alejandro Buren, Mariano Koen-Alonso, Karen Dwyer, Joanne Morgan, Geoff Evans, Don Power, Brian Healey, Martha Robertson, Danny Ings, Don Stansbury, Mark Simpson, Derek Osborne, Peter Upward, Brad Squires, Dwayne Pittman, Paul Higdon
+- Feedback and advice: Alejandro Buren, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Dwayne Pittman, Don Power, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
 - Dave Cote for a review of a previous version of this manuscript

### 2018-09-04 analysis/presentation/sim_survey_presentation.Rmd
Finish drafting presentation

@@
 
-- Increasing sub-sampling effort may be an inefficient means of improving population estimates
-    - Especially problimatic if sub-sampling affects primary sampling effort
+Increasing sub-sampling effort may be an inefficient means of improving population estimates
+
+- Large catches are relatively infrequent
+    - Samples in large sets are not independent
+        - If the goal is to maximize information, it is likely better to stop collecting psudoreplicates and, instead, focus efforts on other species or the next set
 
@@
 
-- Todo
+>- **Feedback and advice:** Alejandro Buren, Dave Cote, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Dwayne Pittman, Don Power, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward, ...
+>- **Support:** Fisheries and Oceans Canada and NSERC
 

### 2018-09-04 R/sim_survey.R
Add a sample size summary table (samp_totals) to the output from test_surveys function

@@ -254,12 +254,18 @@ sim_survey <- function(sim, n_sims = 1, q = sim_logistic(), growth = sim_vonB(),
   setdet$n_measured[is.na(setdet$n_measured)] <- 0
   setdet$n_aged[is.na(setdet$n_aged)] <- 0
 
+  ## Further summarize samples
+  samp_totals <- setdet[, list(sets = .N, n_caught = sum(n),
+                               n_measured = sum(n_measured),
+                               n_aged = sum(n_aged)), by = c("sim", "year")]
+
   ## Add new stuff to main object
   sim$I <- I
   if (!light) sim$sp_I <- sp_I
   if (!light) sim$full_setdet <- full_setdet
   sim$setdet <- setdet
   sim$samp <- samp
+  sim$samp_totals <- samp_totals
   sim
 
 }

### 2018-09-04 R/test_surveys.R

@@ -50,22 +50,27 @@ expand_surveys <- function(set_den = c(0.5, 1, 2, 5, 10) / 1000,
                                               ages_cap = surveys$ages_cap[i],
                                               ...) %>%
                               run_strat() %>% strat_error()
+                            samp_totals <- res$samp_totals
+                            samp_totals$sim <- samp_totals$sim + (j * n_sims - n_sims)
                             total_strat_error <- res$total_strat_error
                             total_strat_error$sim <- total_strat_error$sim + (j * n_sims - n_sims)
                             age_strat_error <- res$age_strat_error
                             age_strat_error$sim <- age_strat_error$sim + (j * n_sims - n_sims)
-                            list(total_strat_error = total_strat_error, age_strat_error = age_strat_error)
+                            list(samp_totals = samp_totals,
+                                 total_strat_error = total_strat_error,
+                                 age_strat_error = age_strat_error)
                           }
     stopCluster(cl) # stop parallel process
 
+    samp_totals <- data.table::rbindlist(lapply(loop_error, `[[`, "samp_totals"))
     total_strat_error <- data.table::rbindlist(lapply(loop_error, `[[`, "total_strat_error"))
     age_strat_error <- data.table::rbindlist(lapply(loop_error, `[[`, "age_strat_error"))
+    samp_totals$survey <- i
     total_strat_error$survey <- i
     age_strat_error$survey <- i
+    survey_error[[i]]$samp_totals <- samp_totals
     survey_error[[i]]$total_strat_error <- total_strat_error
     survey_error[[i]]$age_strat_error <- age_strat_error
-    survey_error[[i]]$total_strat_error_stats <- c(survey = i, error_stats(total_strat_error$error))
-    survey_error[[i]]$age_strat_error_stats <- c(survey = i, error_stats(age_strat_error$error))
 
     if (!is.null(export)) {
       save(survey_error, file = file.path(export_dir, "survey_error.RData"))
@@ -82,18 +87,18 @@ expand_surveys <- function(set_den = c(0.5, 1, 2, 5, 10) / 1000,
 .test_fin <- function(sim = NULL,  n_sims = NULL, surveys = NULL, keep_details = NULL,
                       survey_error = NULL, export = NULL, export_dir = NULL, ...) {
 
-  total_strat_error <- data.table::rbindlist(lapply(survey_error, `[[`, "total_strat_error"))
[…]

### 2018-09-04 analysis/cod_sim.R

@@ -49,7 +49,7 @@ sim <- test_surveys(pop,
                     q = sim_logistic(k = 2, x0 = 3),
                     growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
                     export = "analysis/cod_sim_exports")
-# sim <- resume_test(dir = "analysis/cod_sim_exports/2018-07-13_test")
+# sim <- resume_test(dir = "analysis/cod_sim_exports/2018-09-04_test")
 setMKLthreads() # turn hyperthreading on again
 
 

### 2018-09-06 analysis/presentation/sim_survey_presentation.Rmd
Refine presentation

@@
 :----: | :----: | :----:
-Conduct sets at multiple locations | Sub-sample catch for length measurements | Sub-sampling length groups for age determination
+Conduct sets at multiple locations | Sub-sample catch for length measurements | Sub-sample length groups for age determination
 <img src="graphics/trawl_survey_crop.jpg" height="180"/> | <img src="graphics/cod_length.jpg" height="120"/> | <img src="graphics/Otoliths2_crop.jpg" height="140"/>
@@
 - **_Objective: build a simulation that emulates reality_**
+
+## Simulation steps {.smaller}
+
+1. **Simulate abundance**
+    - Common cohort model
+2. **Simulate spatial aggregation**
+    - Includes depth associations and noise correlated across space, years and ages
+3. **Simulate surveys**
+    - Stratified random surveys with different sampling protocol
+5. **Analyze**
+    - Conduct stratified analyses and compare <a href="javascript:slidedeck.loadSlide(20)">truth vs. estimate</a>
 
@@
 
+>- Bias is negligible in this scenario
 >- Increases in secondary sampling does little to improve estimates
->- Bias is negligible in this scenario
 
@@
 
-- With age-specific clustering, fish of different ages are not encountered with equal probability
-    - A survey with few sets will be less likely to encounter groups in a tight cluster
-        - Survey-level length / age distribution ∴ more likely to be incomplete 
-            - Increasing length / age sampling cannot improve perception of the population-level distribution
-
-## Interpretation {.tighter .smaller}
-
-Ages caught and sampled in year 18 from one survey scenario:
-
->- $D_{sets} = 0.002, M_{lengths} = 500, M_{ages} = 10$
-
-<dev class = "left">
-```{r, fig.height = 4.8}
-
-samp <- merge(main_sim$setdet, main_sim$samp, by = "set")
-samp <- samp[sim == 1 & year == 18, ]
-caught <- as.data.frame(table(samp$age))
-caught <- data.frame(group = "Caught", age = caught$Var1, freq = caught$Freq,
-                     prop = caught$Freq / sum(caught$Freq))
-measured <- as.data.frame(table(samp$age[samp$measured]))
-measured <- data.frame(group = "Sampled", age = measured$Var1, freq = measured$Freq,
-                       prop = measured$Freq / sum(measured$Freq))
-true <- data.frame(group = "True", age = sim$ages, freq = sim$I[, 1],
-                   prop = sim$I[, 1] / sum(sim$I[, 1]))
-d <- rbind(caught, measured, true)
-
-plot_ly(data = d) %>% 
-  add_lines(x = ~age, y = ~prop, color = ~group,
-            colors = rev(viridis::viridis(3))) %>% 
-  tight_layout(xaxis = list(title = "Age"),
-               yaxis = list(title = "Proportion"),
-               margin = list(r = 60))
-
-```
-</dev>
+- Likely related to age-specific clustering
+- Alternate analyses (e.g. ogmap, geostatistical models) may be able to account for such clustering and, therefore, provide estimates with less bias
+- Requires further investigation
 
@@
 
-- Large catches are relatively infrequent
-    - Samples in large sets are not independent
+- Large catches are relatively infrequent and samples in these sets are not independent
 - If the goal is to maximize information, it is likely better to stop collecting psudoreplicates and, instead, focus efforts on other species or the next set

### 2018-09-06 analysis/presentation/sim_survey_presentation.Rmd
Add an assumptions slide and revise conclusions

@@
 library(data.table)
-load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-09-04_test/test_output.RData")
 
@@
 
+## Assumptions
+
+>- Population is uniformly distributed within a cell
+>- The survey is an instantaneous snapshot of the population
+>- Fish are aged at random throughout the division within length bins
+>- Ages are estimated without error
+>- Trawl dimensions are perfectly standard
+
 ## Truth vs. estimate
@@
     - Stratified estimates are supposed to be unbiased?!
-    - Likely related to age-specific clustering
+    - May be related to age-specific clustering
 
@@
 
-Why is there bias? Why is extra sub-sampling ineffective or detrimental?
+Why is there bias? Why is extra sub-sampling sometimes useful and sometimes detrimental?
 
@@
 
-Increasing sub-sampling effort may be an inefficient means of improving population estimates
+Results suggest that
 
-- Large catches are relatively infrequent and samples in these sets are not independent
+- Stratified estimates are not always unbiased
+- Sets are the most effective way to improve estimates
+- Excessive sub-sampling may be detrimental if set density is low
+
+## Conclusions
+
+**Caution:** this simple simulation is far from perfect, focuses on one case study, and lacks a cost component
+
+- Makes it difficult to be prescriptive
+- Nonetheless, results echo the growing body of literature that concludes
+    - Extra sub-sampling is an ineffective means of improving estimates because within-set samples are correlated
     - If the goal is to maximize information, it is likely better to stop collecting psudoreplicates and, instead, focus efforts on other species or the next set
@@
     - Variable trawl area
+    - Random vs. stratified age sampling
 - Use this as a platform to test spatial models
@@
 
+## Code {.smaller}
+
+### Population
+
+```{r, echo = TRUE, eval = FALSE}
+
+      set.seed(438)
+      pop <- sim_abundance(ages = 1:20, years = 1:20,
+                           R = sim_R(mean = 30000000, log_sd = 0.5,
+                                     random_walk = TRUE),
+                           Z = sim_Z(mean = 0.5, log_sd = 0.2,
+                                     phi_age = 0.9, phi_year = 0.5)) %>%
+        sim_distribution(grid = sim_grid(x_range = c(-140, 140), y_range = c(-140, 140),
+                                         res = c(3.5, 3.5), shelf_depth = 200,
+                                         shelf_width = 100, depth_range = c(0, 1000),
+                                         n_div = 1, strat_breaks = seq(0, 1000, by = 20),
+                                         strat_splits = 2),
+                         ays_covar = sim_ays_covar(sd = 2.8, range = 300,
+                                                   phi_age = 0.5, phi_year = 0.9,
+                                                   group_ages = 5:20),
+                         depth_par = sim_parabola(mu = 200, sigma = 70))
+
+```
+
+## Code {.smaller}
+
+### Surveys
+
+```{r, echo = TRUE, eval = FALSE}
+
+      surveys <- expand_surveys(set_den = c(0.5, 1, 2, 5, 10) / 1000,
+                                lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
+                                ages_cap = c(2, 5, 10, 20, 50))
+      sim <- test_surveys(pop,
+                          surveys = surveys, n_sims = 5,
+                          n_loops = 200, cores = 6,
+                          q = sim_logistic(k = 2, x0 = 3),
+                          growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
+                          export = "analysis/cod_sim_exports")
+
+```
+
 <!-- There is a conflict somewhere...had to contain plotly plots in a dev to get it working and I had to add a bunch of spaces to fudge centering in one of the tables -->

### 2018-09-07 R/sim_survey.R
Add a function for simulating surveys in parallel

@@ -255,7 +255,7 @@ sim_survey <- function(sim, n_sims = 1, q = sim_logistic(), growth = sim_vonB(),
   setdet$n_aged[is.na(setdet$n_aged)] <- 0
 
   ## Further summarize samples
-  samp_totals <- setdet[, list(sets = .N, n_caught = sum(n),
+  samp_totals <- setdet[, list(n_sets = .N, n_caught = sum(n),
                                n_measured = sum(n_measured),
                                n_aged = sum(n_aged)), by = c("sim", "year")]
 
@@ -271,5 +271,85 @@ sim_survey <- function(sim, n_sims = 1, q = sim_logistic(), growth = sim_vonB(),
 }
 
 
+#' Simulate stratified random surveys using parallel computation
+#'
+#' This function is a wrapper for \code{\link{sim_survey}} except it allows for
+#' many more total itterations to be run than \code{\link{sim_survey}} before running
+#' into RAM limitations.
+#'
+#' @param sim               Simulation from \code{\link{sim_distribution}}
+#' @param n_sims            Number of times to simulate a survey over the simulated population.
+#'                          Requesting a large number of simulations here may max out your RAM.
+#' @param n_loops           Number of times to run the \code{\link{sim_survey}} function. Total
+#'                          simulations run will be the product of \code{n_sims} and \code{n_loops}
+#'                          arguments. Low numbers of \code{n_sims} and high numbers of \code{n_loops}
+#'                          will be easier on RAM, but may be slower.
+#' @param cores             Number of cores to use in parallel. More cores should speed up the process.
+#' @param quiet             Print message on what to expect for duration?
+#' @inheritDotParams sim_survey
+#'
+#' @details \code{\link{sim_survey}} is hard-wired here to be "light" to minimize object size.
+#'
+#' @export
+#'
 
+sim_survey_parallel <- function(sim, n_sims = 1, n_loops = 100,
+                                cores = 6, quiet = FALSE, ...) {
 
+  start <- Sys.time()
+  one_res <- sim_survey(sim, n_sims = n_sims, light = TRUE, ...)
+  end <- Sys.time()
+  elapsed <- end - start
+  max_dur <- end + (elapsed * n_loops) - start
+
+  if (!quiet) {
+    message(paste("One run of sim_survey took ~",
[…]

### 2018-09-07 analysis/cod_sim.R

@@ -31,6 +31,7 @@ pop <- sim_abundance(ages = 1:20,
                    depth_par = sim_parabola(mu = 200,
                                             sigma = 70))
 
+
 ## Test a series of surveys
 ## Simulate surveys and compare stratified estimates to the true index
 setMKLthreads(1) # turn off MKL hyperthreading

### 2018-09-07 R/plot_sim.R
Draft a plot for visualizing relationship between sample size and error

@@ -558,3 +558,56 @@ plot_total_strat_fan <- function(sim, surveys = 1:5,
 }
 
 
+#' @export
+#' @rdname plot_trend
+plot_effort_error_surface <- function(sim, ...) {
+
+  totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught),
+                                   n_measured = mean(n_measured), n_aged = mean(n_aged)),
+                            by = "survey"]
+  errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
+  d <- merge(errors, totals, by = "survey")
+
+  d$text <- with(d, paste("<br>set density:", set_den,
+                          "<br>lengths cap:", lengths_cap,
+                          "<br>ages cap:", ages_cap,
+                          "<br><br>N sets:", n_sets,
+                          "<br>N lengths:", round(n_measured),
+                          "<br>N ages:", round(n_aged)))
+
+  shared_d <- crosstalk::SharedData$new(d)
+
+  crosstalk::bscols(
+    widths = c(3, NA),
+    list(
+      htmltools::div(style = htmltools::css(height = "10px")), # small margin
+      filter_select("n_sets", "N set", shared_d, ~n_sets, multiple = FALSE)
+    ),
+    plot_ly(data = shared_d, x = ~n_aged, y = ~n_measured, z = ~RMSE) %>%
+      add_trace(type = "mesh3d", flatshading = TRUE, intensity = ~RMSE, name = "RMSE",
+                hoverinfo = "skip", contour = list(show = TRUE, width = 15, color = toRGB("white"))) %>%
+      add_markers(z = ~RMSE, color = I("lightgrey"), size = I(1), name = "RMSE",
+                  text = ~text, hoverinfo = "text+z",
+                  showlegend = FALSE) %>%
+      add_markers(z = ~RMSE, color = I("lightgrey"), name = "RMSE",
+                  opacity = 0, text = ~text, hoverinfo = "text+z",
+                  showlegend = FALSE) %>%
+      layout(
+        scene = list(
+          xaxis = list(title = "N ages"),
+          yaxis = list(title = "N lengths"),
+          zaxis = list(title = "RMSE", range = range(d$RMSE)),
+          camera = list(eye = list(x = 2, y = 1.5, z = 1.5))
+        )) %>%
+      highlight(on = NULL)
+  )
+
+}
+
+
+
+
+
+
+
+

### 2018-09-10 analysis/paper/sim_survey_paper.Rmd
Add an option to plot_error_surface function for viewing results against sample size and revise presentation accordingly

@@
 
-$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})}{N_a N_y N_k}}$$
+$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})^2}{N_a N_y N_k}}$$
 Where $N_a$, $N_y$, and $N_k$ are the number of years, ages and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e.

### 2018-09-10 analysis/presentation/sim_survey_presentation.Rmd
Draft a plot for visualizing relationship between sample size and error
Add an option to plot_error_surface function for viewing results against sample size and revise presentation accordingly

@@
 author: "Paul M. Regular & Fran Mowbray"
-date: "September 6^th^, 2018"
+date: "September 12^th^, 2018"
 output:
@@
 - Used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance estimates from each survey:
-    - $RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})}{N_a N_y N_k}}$
+    - $RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})^2}{N_a N_y N_k}}$
     - Where $I_{a,y}$ is the true abundance available to the survey
@@
 
-### RMSE
+### RMSE ~ sampling protocol
 
@@
 )
-plot_error_surface(sim) %>% 
+plot_error_surface(sim, plot_by = "rule") %>% 
+  tight_layout(scene = surface_scene,
+               margin = list(l = 0, r = 0, t = 40, b = 0, pad = 0))
+```
+</dev>
+
+## Truth vs. estimate {.tighter}
+
+### RMSE ~ sample size
+
+<dev class = "left">
+```{r}
+plot_error_surface(sim, plot_by = "samples") %>% 
   tight_layout(scene = surface_scene,
@@
 main_sim <- sim
-load("analysis/cod_sim_exports/2018-07-16_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-09-07_test/test_output.RData")
 plot_age_strat_fan(sim, surveys = surveys$survey,
@@
 ```{r}
-plot_error_surface(sim) %>% 
+plot_error_surface(sim, plot_by = "samples") %>% 
   tight_layout(scene = surface_scene,
@@
 
->- Bias is negligible in this scenario
->- Increases in secondary sampling does little to improve estimates
+In this scenario,
+
+- Bias is negligible
+- Increasing sub-sampling is always benifical
+- Sampling fewer fish at more locations appears better than many fish at fewer locations
 

### 2018-09-10 R/plot_sim.R

@@ -17,12 +17,12 @@
 #' @param max_sims       Maximum number of sims to plot
 #' @param facet_by       Facet plot by "age" or "year"?
 #' @param select_by      Select plot by "age" or "year"?
+#' @param plot_by        Plot error surface by "rule" or "samples"?
 #' @param survey         Subset data to one or more surveys.
 #' @param quants         Quantile intervals to display on fan plot
 #' @param col            Plot color
 #' @param cols           Plot colors
 #' @param main           Plot title
-#' @param autoscale      Should axes automatically adjust or be fixed to the range of the data?
 #' @param ...            Additional arguments to pass to \code{\link{plotly::plot_ly}}.
 #'
 #' @import plotly
@@ -140,54 +140,143 @@ plot_distribution_slider <- function(sim, ages = 1, years = 1) {
 
 #' @export
 #' @rdname plot_trend
-plot_error_surface <- function(sim) {
+plot_error_surface <- function(sim, plot_by = "rule") {
 
-  d <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
-  split_d <- split(d, d$set_den)
-  split_d <- lapply(split_d, function(.) {
-    xtabs(RMSE ~ ages_cap + lengths_cap, data = ., subset = NULL)
-  })
-  x <- sort(unique(d$lengths_cap))
-  y <- sort(unique(d$ages_cap))
+  totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught),
+                                   n_measured = mean(n_measured), n_aged = mean(n_aged)),
+                            by = "survey"]
+  errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
+  d <- merge(errors, totals, by = "survey")
 
-  p <- plot_ly(x = ~x, y = ~y, cmin = min(d$RMSE), cmax = max(d$RMSE))
-  visible <- showscale <- rep(TRUE, length(split_d))
-  steps <- list()
-  showscale <- c(TRUE, rep(FALSE, length(split_d) - 1))
-  for (i in seq_along(split_d)) {
-    vis <- rep(FALSE, length(split_d))
-    vis[i] <- TRUE
-    p <- p %>% add_surface(z = split_d[[i]],
-                           visible = i == 1,
-                           showscale = vis,
-                           name = names(split_d)[i],
-                           colorbar = list(title = "RMSE"))
-    steps[[i]] <- list(args = list(list(visible = vis,
-                                        showscale = vis)),
[…]

### 2018-09-10 analysis/presentation/sim_survey_presentation.Rmd
Tweak interpretations in presentation and clean-up test_survey documentation

@@
 
-<br>
+- In this scenario, 
+    - Bias is negligible
+    - Increasing sub-sampling is always beneficial
+        - No age-specific clustering means that one fish is as "effective" as the next
+    - Sub-sampling fewer fish at more locations appears better than many fish at fewer locations
+        - Though more sub-samples improve perception of the age-distribution, it does little to inform magnitude
 
-In this scenario,
+## Check bias
 
-- Bias is negligible
-- Increasing sub-sampling is always benifical
-- Sampling fewer fish at more locations appears better than many fish at fewer locations
-
-## Interpretation
-
-Why is there bias? Why is extra sub-sampling sometimes useful and sometimes detrimental?
+Why is there bias in the main scenario? Why is extra sub-sampling sometimes useful / sometimes detrimental?
 

### 2018-09-10 R/test_surveys.R

@@ -129,8 +129,7 @@ expand_surveys <- function(set_den = c(0.5, 1, 2, 5, 10) / 1000,
 #' Test sampling design of multiple surveys
 #'
 #' @description   This function allows a series of sampling design settings to
-#' be set and tested on the simulated population. All combinations of the
-#' supplied settings (\code{set_den}, \code{lengths_cap}, \code{ages_cap}) are tested.
+#' be set and tested on the simulated population.
 #'
 #' @param sim               Simulation from \code{\link{sim_distribution}}.
 #' @param surveys           A data.frame or data.table with a sequence of surveys and their settings
@@ -190,6 +189,8 @@ test_surveys <- function(sim, surveys = expand_surveys(), keep_details = 1,
 #'
 #' @param dir  Export directory specified when \code{\link{test_surveys}} was run.
 #'
+#' @details Progress bar time estimates will be biased here by previous completions
+#'
 #' @export
 #'
 
@@ -203,4 +204,5 @@ resume_test <- function(dir = NULL) {
             n_sims = n_sims, keep_details = keep_details,
             export = export, export_dir = export_dir, ...)
 }
+
 

### 2018-09-10 analysis/cod_sim.R
Export data for Geoff (for analysis using ogmap)

@@ -59,6 +59,42 @@ load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
 vis_sim(sim)
 
 
+## Exports for Geoff -----------------------------------------------------------
+
+library(data.table)
+
+## ~ current sampling protocol
+base_case <- sim_survey_parallel(pop, n_sims = 5, n_loops = 200, cores = 6,
+                                 set_den = 2e-03, lengths_cap = 500, ages_cap = 10,
+                                 q = sim_logistic(k = 2, x0 = 3),
+                                 growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
+                                 quiet = FALSE)
+setdet <- base_case$setdet[, list(year, x, y, depth, strat, n, sim, set)]
+samp <- base_case$samp[base_case$samp$measured, ]
+samp$age[!samp$aged] <- NA
+samp$measured <- samp$aged <- NULL
+samp <- merge(setdet[, list(year, sim, set)], samp, by = "set")
+fwrite(setdet, file = "analysis/cod_sim_exports/2018-09-10_for_Geoff/01_setdet.csv")
+fwrite(samp, file = "analysis/cod_sim_exports/2018-09-10_for_Geoff/01_samp.csv")
+rm(base_case)
+gc()
+
+## One example of a protocol that results in bias
+bias_case <- sim_survey_parallel(pop, n_sims = 1, n_loops = 1000, cores = 6,
+                                 set_den = 1e-02, lengths_cap = 10, ages_cap = 10,
+                                 q = sim_logistic(k = 2, x0 = 3),
+                                 growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
+                                 quiet = FALSE)
+setdet <- bias_case$setdet[, list(year, x, y, depth, strat, n, sim, set)]
+samp <- bias_case$samp[bias_case$samp$measured, ]
+samp$age[!samp$aged] <- NA
+samp$measured <- samp$aged <- NULL
+samp <- merge(setdet[, list(year, sim, set)], samp, by = "set")
+fwrite(setdet, file = "analysis/cod_sim_exports/2018-09-10_for_Geoff/02_setdet.csv")
+fwrite(samp, file = "analysis/cod_sim_exports/2018-09-10_for_Geoff/02_samp.csv")
+rm(bias_case)
+gc()
+
 
 ## Simulate same distribution across ages --------------------------------------
 

### 2018-09-11 analysis/paper/sim_survey_paper.Rmd
Use new simulation run in paper and tweak figures

@@
 
-devtools::install_github("rstudio/rmarkdown@v1.10")
+# devtools::install_github("rstudio/rmarkdown@v1.10")
 
@@
 library(data.table)
-load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-09-04_test/test_output.RData")
 ```
@@
     geom_ribbon(aes(ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
-    geom_line(aes(y = I), data = true, color = viridis::viridis(1), size = 0.7) +
+    geom_line(aes(y = I), data = true, color = "black", size = 0.6) +
     scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
-    scale_alpha_discrete(name = "Pr") + 
+    scale_alpha_discrete(name = "Pr", range = c(0.5, 1)) + 
     xlab(x) + ylab("Abundance index") +
@@
 ## this export holds the same distribution test
-load("analysis/cod_sim_exports/2018-07-16_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-09-07_test/test_output.RData")
 ```

### 2018-09-11 analysis/paper/sim_survey_paper.Rmd
Add a RMSE comparison table to the paper

@@
 load("analysis/cod_sim_exports/2018-09-04_test/test_output.RData")
+```
+
+```{r fun}
+
+## http://www.moeding.net/archives/32-Metric-prefixes-for-ggplot2-scales.html
+format_si <- function(...) {
+  # Format a vector of numeric values according
+  # to the International System of Units.
+  # http://en.wikipedia.org/wiki/SI_prefix
+  #
+  # Based on code by Ben Tupper
+  # https://stat.ethz.ch/pipermail/r-help/2012-January/299804.html
+  # Args:
+  #   ...: Args passed to format()
+  #
+  # Returns:
+  #   A function to format a vector of strings using
+  #   SI prefix notation
+  #
+
+  function(x) {
+    limits <- c(1e-24, 1e-21, 1e-18, 1e-15, 1e-12,
+                1e-9,  1e-6,  1e-3,  1e0,   1e3,
+                1e6,   1e9,   1e12,  1e15,  1e18,
+                1e21,  1e24)
+    prefix <- c("y",   "z",   "a",   "f",   "p",
+                "n",   "µ",   "m",   " ",   "k",
+                "M",   "G",   "T",   "P",   "E",
+                "Z",   "Y")
+
+    # Vector with array indices according to position in intervals
+    i <- findInterval(abs(x), limits)
+
+    # Set prefix to " " for very small values < 1e-24
+    i <- ifelse(i == 0, which(limits == 1e0), i)
+
+    paste0(format(round(x/limits[i], 1),
+                 trim = TRUE, scientific = FALSE, ...),
+          prefix[i])
+  }
+}
+si <- format_si()
+
+## little helper function for the fan plot
+fan_plot <- function(d, x = "Age", total_strat = FALSE) {
+
+  if (total_strat) {
+    ints_by = c("year", "set_den")
+  } else {
+    ints_by = c("year", "age", "set_den", "lengths_cap")
+  }
+
+  ## quantiles for fan plot
+  quants <- seq(90, 10, by = -10)
+  ints <- lapply(quants, function(q) {
+    d[, list(prob = paste0(q, "%"),
+             lower = quantile(I_hat, prob = (1 - q / 100) / 2),
+             upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
+      by = ints_by]
+  })
+  ints <- rbindlist(ints)
+  ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
+
+  true <- d[d$sim %in% 1:3, ]
+
+  if (x == "Age") {
+    p <- ggplot(data = ints, aes(x = age))
+    xlabs <- c("2", "4", "6", "8+")
+    xbreaks <- seq(2, 8 , 2)
+  }
+  if (x == "Year") {
+    p <- ggplot(data = ints, aes(x = year))
+    xlabs <- waiver()
+    xbreaks <- waiver()
+  }
+
+  p <-  p + 
+    geom_ribbon(aes(ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
+    geom_line(aes(y = I), data = true, color = "black", size = 0.6) +
+    scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
+    scale_alpha_discrete(name = "Pr", range = c(0.5, 1)) + 
+    xlab(x) + ylab("Abundance index") +
+    scale_y_continuous(labels = format_si()) +
+    scale_x_continuous(expand = c(0.01, 0), 
+                       labels = xlabs, 
+                       breaks = xbreaks) +
+    guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
+                                default.unit = "cm", ncol = 1))
+
+  if (total_strat) {
+    p <- p + facet_wrap(~ set_den, nrow = 1,
+                        labeller = label_bquote(cols = D[sets]==.(set_den)))
+  } else {
+    p <- p + facet_grid(lengths_cap ~ set_den, 
+                        labeller = label_bquote(cols = D[sets]==.(set_den),
+                                                rows = M[lengths]==.(lengths_cap)))
+  }
+
+  p + theme_bw() +
+    theme(legend.justification = c(0, 1),
+          plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"),
+          panel.grid.major = element_blank(), 
+          panel.grid.minor = element_blank(),
+          strip.background = element_blank(),
+          panel.border = element_rect(fill = NA, color = "darkgrey", size = 0.4),
+          axis.ticks = element_line(color = "darkgrey", size = 0.4))
+
+}
+
 ```
@@
 A dynamic population that varies across ages, time and space is simulated with the parameter settings used here (Table \@ref(tab:par)).
-Figure \@ref(fig:sp-fig) depicts the aggregate distribution of the population in one year of the simulation and the corresponding catches of one stratified random survey over this population.
+Figure \@ref(fig:sp) depicts the aggregate distribution of the population in one year of the simulation and the corresponding catches of one stratified random survey over this population.
 Layered within this, of course, is the distributions of each age, as well as age and length sampling details.
-Though age-specific details are not depicted for simplicity, the general pattern of spatial aggregation apparent in Figure \@ref(fig:sp-fig) is similar across all ages.
+Though age-specific details are not depicted for simplicity, the general pattern of spatial aggregation apparent in Figure \@ref(fig:sp) is similar across all ages.
 That is, all ages tend to aggregate in fairly dense clusters and low-density zones are fairly wide ranging.
-The survey, therefore, tends to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:sp-fig)).
+The survey, therefore, tends to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:sp)).
 The samples within each successful set are also dynamic and rarely include all age groups.
@@
 
+<br>
+
 Table: (\#tab:tests) - Descriptions and settings of different survey sampling protocol tested in the cod case study. All possible combinations of the options below were tested.
@@
 |             Set density |      $D_{sets}$ |    0.0005, 0.001, 0.002, 0.005, 0.01 sets / km^2^ |
-|  Length sampling effort |   $M_{lengths}$ |       5, 10, 20, 50, 100, 500, 1000 lengths / set |
-|     Age sampling effort |      $M_{ages}$ |   2, 5, 10, 20, 50 ages / length group / division |
+|    Length sampling rule |   $M_{lengths}$ |       5, 10, 20, 50, 100, 500, 1000 lengths / set |
+|       Age sampling rule |      $M_{ages}$ |   2, 5, 10, 20, 50 ages / length group / division |
+
+<br>
+
+Table: (\#tab:comp) - Root-mean-squared errors (RMSE) and number of samples [sets ($N_{sets}$), lengths ($N_{lengths}$) and ages ($N_{ages}$)] obtained from an array of surveys with different sampling protocol [set densities ($D_{sets}$), maximum sampling effort of lengths per set ($M_{lengths}$), and maximum number of ages to collect per length group ($M_{ages}$)].
+Differences from the minimum RMSE value ($\Delta ~ RMSE$) within each set density option are presented to aid comparisons.
+
+```{r comp, results = "asis"}
+
+totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught),
+                                 n_measured = mean(n_measured), n_aged = mean(n_aged)),
+                          by = "survey"]
+errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
+d <- merge(errors, totals, by = "survey")
+
+which_set_dens <- c(0.0005, 0.002, 0.01)
+which_len_caps <- c(10, 100, 1000)
+which_age_caps <- c(5, 10, 50)
+
+tab <- d[d$set_den %in% which_set_dens & d$lengths_cap %in% which_len_caps & 
+           d$ages_cap %in% which_age_caps, 
+         list(lengths_cap, ages_cap,
+              n_sets = round(n_sets), n_measured = round(n_measured), n_aged = round(n_aged),
+              RMSE = RMSE, delta_RMSE = RMSE - min(RMSE)), 
+         by = "set_den"]
+tab <- tab[order(set_den, delta_RMSE), ]
+tab$set_den <- formatC(tab$set_den)
+tab$set_den[tab$delta_RMSE != 0] <- NA
+tab$RMSE <- si(tab$RMSE)
+tab$delta_RMSE <- si(tab$delta_RMSE)
+
+split_tab <- split(tab, tab$n_sets)
+for (i in seq_along(split_tab)[-length(split_tab)]) {
+  split_tab[[i]] <- rbind(split_tab[[i]], data.table(RMSE = NA), fill = TRUE)
+}
+tab <- rbindlist(split_tab)
+
+names(tab) <- c("$D_{sets}$", "$M_{lengths}$", "$M_{ages}$", 
+                "$N_{sets}$", "$N_{lengths}$", "$N_{ages}$",
+                "$RMSE$", "$\\Delta ~ RMSE$")
+
+options(knitr.kable.NA = "")
+knitr::kable(tab, align = "r", format = "markdown")
+
+```
 
@@
 
-```{r fig_fun}
-
-## http://www.moeding.net/archives/32-Metric-prefixes-for-ggplot2-scales.html
-format_si <- function(...) {
-  # Format a vector of numeric values according
-  # to the International System of Units.
-  # http://en.wikipedia.org/wiki/SI_prefix
-  #
-  # Based on code by Ben Tupper
-  # https://stat.ethz.ch/pipermail/r-help/2012-January/299804.html
-  # Args:
-  #   ...: Args passed to format()
-  #
-  # Returns:
-  #   A function to format a vector of strings using
-  #   SI prefix notation
-  #
-
-  function(x) {
-    limits <- c(1e-24, 1e-21, 1e-18, 1e-15, 1e-12,
-                1e-9,  1e-6,  1e-3,  1e0,   1e3,
-                1e6,   1e9,   1e12,  1e15,  1e18,
-                1e21,  1e24)
-    prefix <- c("y",   "z",   "a",   "f",   "p",
-                "n",   "µ",   "m",   " ",   "k",
-                "M",   "G",   "T",   "P",   "E",
-                "Z",   "Y")
-
-    # Vector with array indices according to position in intervals
-    i <- findInterval(abs(x), limits)
-
-    # Set prefix to " " for very small values < 1e-24
-    i <- ifelse(i == 0, which(limits == 1e0), i)
-
-    paste0(format(round(x/limits[i], 1),
-                 trim = TRUE, scientific = FALSE, ...),
-          prefix[i])
-  }
-}
-si <- format_si()
-
-## little helper function for the fan plot
-fan_plot <- function(d, x = "Age", total_strat = FALSE) {
-
-  if (total_strat) {
-    ints_by = c("year", "set_den")
-  } else {
-    ints_by = c("year", "age", "set_den", "lengths_cap")
-  }
-
-  ## quantiles for fan plot
-  quants <- seq(90, 10, by = -10)
-  ints <- lapply(quants, function(q) {
-    d[, list(prob = paste0(q, "%"),
-             lower = quantile(I_hat, prob = (1 - q / 100) / 2),
-             upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
-      by = ints_by]
-  })
-  ints <- rbindlist(ints)
-  ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
-
-  true <- d[d$sim %in% 1:3, ]
-
-  if (x == "Age") {
-    p <- ggplot(data = ints, aes(x = age))
-    xlabs <- c("2", "4", "6", "8+")
-    xbreaks <- seq(2, 8 , 2)
-  }
-  if (x == "Year") {
-    p <- ggplot(data = ints, aes(x = year))
-    xlabs <- waiver()
-    xbreaks <- waiver()
-  }
-
-  p <-  p + 
-    geom_ribbon(aes(ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
-    geom_line(aes(y = I), data = true, color = "black", size = 0.6) +
-    scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
-    scale_alpha_discrete(name = "Pr", range = c(0.5, 1)) + 
-    xlab(x) + ylab("Abundance index") +
-    scale_y_continuous(labels = format_si()) +
-    scale_x_continuous(expand = c(0.01, 0), 
-                       labels = xlabs, 
-                       breaks = xbreaks) +
-    guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
-                                default.unit = "cm", ncol = 1))
-
-  if (total_strat) {
-    p <- p + facet_wrap(~ set_den, nrow = 1,
-                        labeller = label_bquote(cols = D[sets]==.(set_den)))
-  } else {
-    p <- p + facet_grid(lengths_cap ~ set_den, 
-                        labeller = label_bquote(cols = D[sets]==.(set_den),
-                                                rows = M[lengths]==.(lengths_cap)))
-  }
-
-  p + theme_bw() +
-    theme(legend.justification = c(0, 1),
-          plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"),
-          panel.grid.major = element_blank(), 
-          panel.grid.minor = element_blank(),
-          strip.background = element_blank(),
-          panel.border = element_rect(fill = NA, color = "darkgrey", size = 0.4),
-          axis.ticks = element_line(color = "darkgrey", size = 0.4))
-
-}
-
-```
-
-```{r sp-fig, cache = TRUE, fig.height = 3.8, fig.width = 8.5, fig.cap = "- An example from year seven of the simulation of a) the distribution of total abundance available to the survey, and b) locations of the sets conducted and total numbers caught in a stratified random survey with a set density of 0.002 sets / km^2^.
+```{r sp, cache = TRUE, fig.height = 3.8, fig.width = 8.5, fig.cap = "- An example from year seven of the simulation of a) the distribution of total abundance available to the survey, and b) locations of the sets conducted and total numbers caught in a stratified random survey with a set density of 0.002 sets / km^2^.
 Grey lines indicate depth-based survey strata."}
@@
 
-```{r tot-fig, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
+```{r tot, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
 The heavy dark line indicates the true tred in the total population available to the survey."}
@@
 
-```{r age-fig, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+```{r age, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
@@
 
-```{r year-fig, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+```{r year, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
@@
 
-```{r error-fig, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
+```{r error, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
 Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$)."}
@@
 
-```{r appen-age-fig, cache = TRUE, fig.width = 8, fig.height = 6}
+```{r appen-age, cache = TRUE, fig.width = 8, fig.height = 6}
 
@@
 
-```{r appen-year-fig, cache = TRUE, fig.width = 8, fig.height = 6}
+```{r appen-year, cache = TRUE, fig.width = 8, fig.height = 6}
 
@@
 
-```{r appen-error-fig, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide"}
+```{r appen-error, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide"}
 

### 2018-09-12 analysis/paper/sim_survey_paper.Rmd
Minor tweaks of presentation, paper and data export for Geoff

@@
 
-Next discuss trend...
+### TODO
+
+- Next discuss trend...
+- Look into the ictopelagic sampling design literature (suggestion from Pierre)
+- Mention that simulation is an approach widely applied to inform sampling design (program MARK, distance, RandomForest, etc.); i.e. cite the gods
+- Look into work done on occupancy modeling and sampling design for such studies (i.e. book suggestion by Aaron)
 
@@
   coord_equal() +
-  theme_void() + guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
+  theme_void() + guides(color = guide_legend(keywidth = 0.4, keyheight = 0, 
                                              default.unit = "cm", ncol = 1)) +
@@
 setdet <- setdet[setdet$year == which_year & setdet$sim == which_sim, ]
+breaks <- pretty(setdet$n, 7)
 
@@
   geom_point(data = setdet, aes(x = x, y = y, size = n, color = n), alpha = 0.8) +
-  scale_size_area(name = "   n", max_size = 10, breaks = pretty(setdet$n, 8)) +
-  scale_color_viridis(name = "   n", breaks = pretty(setdet$n, 8)) + 
+  scale_size_area(name = "   n", max_size = 10, limits = range(breaks), breaks = breaks) +
+  scale_color_viridis(name = "   n", limits = range(breaks), breaks = breaks) + 
   scale_x_continuous(expand = c(0,0)) +

### 2018-09-12 analysis/presentation/sim_survey_presentation.Rmd
Minor tweaks of presentation, paper and data export for Geoff

@@
   ioslides_presentation:
-    mathjax: local
-    self_contained: false
     widescreen: true
@@
 
-- Taylor to other species and assess sampling protocol
+- Tailor to other species and assess sampling protocol
 - Add a time component to assess trade-offs / efficiency

### 2018-09-12 analysis/cod_sim.R

@@ -64,7 +64,7 @@ vis_sim(sim)
 library(data.table)
 
 ## ~ current sampling protocol
-base_case <- sim_survey_parallel(pop, n_sims = 5, n_loops = 200, cores = 6,
+base_case <- sim_survey_parallel(pop, n_sims = 5, n_loops = 5, cores = 6,
                                  set_den = 2e-03, lengths_cap = 500, ages_cap = 10,
                                  q = sim_logistic(k = 2, x0 = 3),
                                  growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
@@ -74,13 +74,13 @@ samp <- base_case$samp[base_case$samp$measured, ]
 samp$age[!samp$aged] <- NA
 samp$measured <- samp$aged <- NULL
 samp <- merge(setdet[, list(year, sim, set)], samp, by = "set")
-fwrite(setdet, file = "analysis/cod_sim_exports/2018-09-10_for_Geoff/01_setdet.csv")
-fwrite(samp, file = "analysis/cod_sim_exports/2018-09-10_for_Geoff/01_samp.csv")
+fwrite(setdet, file = "analysis/cod_sim_exports/2018-09-12_for_Geoff/01_setdet.csv")
+fwrite(samp, file = "analysis/cod_sim_exports/2018-09-12_for_Geoff/01_samp.csv")
 rm(base_case)
 gc()
 
 ## One example of a protocol that results in bias
-bias_case <- sim_survey_parallel(pop, n_sims = 1, n_loops = 1000, cores = 6,
+bias_case <- sim_survey_parallel(pop, n_sims = 1, n_loops = 25, cores = 6,
                                  set_den = 1e-02, lengths_cap = 10, ages_cap = 10,
                                  q = sim_logistic(k = 2, x0 = 3),
                                  growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
@@ -90,8 +90,8 @@ samp <- bias_case$samp[bias_case$samp$measured, ]
 samp$age[!samp$aged] <- NA
 samp$measured <- samp$aged <- NULL
 samp <- merge(setdet[, list(year, sim, set)], samp, by = "set")
-fwrite(setdet, file = "analysis/cod_sim_exports/2018-09-10_for_Geoff/02_setdet.csv")
[…]

### 2018-09-14 analysis/paper/sim_survey_paper.Rmd
Add abundance-at-length calculations

@@
 total mortality can be made to be more similar for fish that are closer together in age and time).
+Abundance-at-age is then converted to abundance-at-length using the original von Bertalanffy growth curve [@Cailliet2006]:
+
+$$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$
+
+Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$.
+Abundance in discrete length groups, $l_{group}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance-at-age to abundance-at-length.
 Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.
@@
 A more detailed description of the space-age-year covariance is included in Appendix 1.
+Note that the the distributions of specific length groups were not simulated because of the greater computational cost of storing spatially dis-aggregated abundance across all length groups.
 
@@
 Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
-The lengths of the fish sampled by the set are then simulated using the original von Bertalanffy growth curve [@Cailliet2006]:
-
-$$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$
-Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$.
+The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the [Simulate abundance] section.
 Depending on the number of fish caught, sub-sampling is then conducted.
@@
 A series of `r nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests).
-Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain an estimate of abundance at age for each year for each simulation, $\hat{I}_{a,y,k}$.
-Intra-haul correlation (i.e.
-intraclass correlation) of the set catch was calculated by fitting a mixed-model to the lengths data with set as a random effect [fit using the lme4 package; @Bates2015]; variance estimates from this model were then plugged into the ICC(1) formula presented in McGraw and Wong [-@Mcgraw1996].
-We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance estimates from each survey:
+Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain annual estimates of total abundance ($\hat{I}_{y,k}$), abundance-at-length ($\hat{I}_{l,y,k}$) and abundance-at-age ($\hat{I}_{a,y,k}$).
+We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance-at-age estimates from each survey:
 
@@
 catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$).
+Intra-haul correlation (i.e.
+intraclass correlation) of set catch was also calculated by fitting a mixed-model to the lengths data with set as a random effect [fit using the lme4 package; @Bates2015]; variance estimates from this model were then plugged into the ICC(1) formula presented in McGraw and Wong [-@Mcgraw1996].
 All simulations and analyses were conducted using `r substr(version$version.string, 1, 15)` [@R2017] and the code has been developed into a package called `SimSurvey` to facilitate future work and collaboration (https://github.com/PaulRegular/SimSurvey).
@@
 
-- Feedback and advice: Alejandro Buren, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Dwayne Pittman, Don Power, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
+- Feedback and advice: Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Dwayne Pittman, Don Power, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
 - Dave Cote for a review of a previous version of this manuscript
@@
 | Correlation across years in error around total mortality |   $\varphi_{\delta, year}$ |            0.5 |
+|                                   Mean asymptotic length |               $l_{\infty}$ |         120 cm |
+|                                          Length at birth |                      $l_0$ |           5 cm |
+|                                              Growth rate |                        $K$ |            0.1 |
+|                     Length group for abundance at length |            $l_{group}^{N}$ |           3 cm |
 |                                                          |                            |                |  
@@
 |               Midpoint of logistic curve of catchability |                      $x_0$ |              3 |
-|                                   Mean asymptotic length |               $l_{\infty}$ |            120 |
-|                                          Length at birth |                      $l_0$ |              5 |
-|                                              Growth rate |                        $K$ |            0.1 |
-|                                             Length group |                $l_{group}$ |              1 |  
+|                            Length group for age sampling |                $l_{group}$ |           1 cm |  
 \* Distribution set to be equal for ages 5+

### 2018-09-14 R/plot_sim.R

@@ -9,6 +9,7 @@
 #'                       \code{\link{sim_distribution}}, etc.
 #' @param mat            Name of matrix in \code{sim} list to plot.
 #' @param ages           Subset data to one or more ages.
+#' @param lengths        Subset data to one or more length groups.
 #' @param years          Subset data to one or more years.
 #' @param scale          Plot response on "natural" or "log" scale?
 #' @param which_survey   Subset to specific survey
@@ -16,7 +17,7 @@
 #' @param which_sim      Subset to specific sim
 #' @param max_sims       Maximum number of sims to plot
 #' @param facet_by       Facet plot by "age" or "year"?
-#' @param select_by      Select plot by "age" or "year"?
+#' @param select_by      Select plot by "age", "length" or "year"?
 #' @param plot_by        Plot error surface by "rule" or "samples"?
 #' @param survey         Subset data to one or more surveys.
 #' @param quants         Quantile intervals to display on fan plot
@@ -603,6 +604,68 @@ plot_age_strat_fan <- function(sim, surveys = 1:5, years = 1:10,
 
 #' @export
 #' @rdname plot_trend
+plot_length_strat_fan <- function(sim, surveys = 1:5, years = 1:10,
+                                  lengths = 1:50, select_by = "year",
+                                  quants = seq(90, 10, by = -10),
+                                  ...) {
+
+  d <- sim$length_strat_error
+  sub_d <- d[survey %in% surveys & year %in% years & length %in% lengths, ]
+  true <- sub_d[sim == 1, list(year, length, survey, I)]
+
+  ## Calculate a series of quantiles
+  ints <- lapply(quants, function(q) {
+    sub_d[, list(prob = paste0(q, "%"),
+                 lower = quantile(I_hat, prob = (1 - q / 100) / 2),
+                 upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)),
+          by = c("length", "year", "survey")]
+  })
+  ints <- rbindlist(ints)
+  ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
+
+  ints <- merge(sim$surveys, ints, by = "survey", all.y = TRUE)
+  ints$lab <- paste(formatC(ints$set_den, format = "fg"),
+                    ints$lengths_cap, sep = "-")
+  ints <- merge(ints, true, by = c("length", "year", "survey"))
+
[…]

### 2018-09-14 R/run_strat.R

@@ -219,6 +219,22 @@ strat_error <- function(sim) {
   sim$total_strat_error <- comp
   sim$total_strat_error_stats <- means
 
+  ## length_strat
+  I_hat <- sim$length_strat[, list(sim, year, length, total)]
+  names(I_hat) <- c("sim", "year", "length", "I_hat")
+  sly <- expand.grid(sim = seq(max(sim$total_strat$sim)),
+                     year = sim$years, length = sim$lengths)
+  I_hat <- merge(sly, I_hat, by = c("sim", "year", "length"), all = TRUE) # expand to all lengths
+  I_hat$I_hat[is.na(I_hat$I_hat)] <- 0                                    # fill missing lengths with zero
+  I <- as.data.frame.table(sim$I_at_length, responseName = "I")
+  I$year <- as.numeric(as.character(I$year))
+  I$length <- as.numeric(as.character(I$length))
+  comp <- merge(data.table(I_hat), data.table(I), by = c("year", "length"))
+  comp$error <- comp$I_hat - comp$I
+  means <- error_stats(comp$error)
+  sim$length_strat_error <- comp
+  sim$length_strat_error_stats <- means
+
   ## age_strat
   I_hat <- sim$age_strat[, list(sim, year, age, total)]
   names(I_hat) <- c("sim", "year", "age", "I_hat")
@@ -227,8 +243,8 @@ strat_error <- function(sim) {
   I_hat <- merge(say, I_hat, by = c("sim", "year", "age"), all = TRUE) # expand to all ages
   I_hat$I_hat[is.na(I_hat$I_hat)] <- 0                                 # fill missing ages with zero
   I <- as.data.frame.table(sim$I, responseName = "I")
-  I$year <- as.numeric(I$year)
-  I$age <- as.numeric(I$age)
+  I$year <- as.numeric(as.character(I$year))
+  I$age <- as.numeric(as.character(I$age))
   comp <- merge(data.table(I_hat), data.table(I), by = c("year", "age"))
   comp$error <- comp$I_hat - comp$I
   means <- error_stats(comp$error)

### 2018-09-14 R/sim_dist.R

@@ -182,7 +182,7 @@ sim_parabola <- function(alpha = 0, mu = 250, sigma = 50, plot = FALSE) {
 #' @description Provided an abundance at age matrix (like one provided by \code{\link{sim_abundance}})
 #' and a survey grid (like \code{\link{survey_grid}}) to populate, this function
 #' applies correlated space, age and year error to simulate the distribution
-#' of the population.
+#' of the population. The ability to simulate distributions by length is yet to be implemented.
 #'
 #' @param sim         An abundance at age matrix like one produced by \code{\link{sim_abundance}}
 #' @param grid        A raster object defining the survey grid, like \code{\link{survey_grid}}

### 2018-09-14 R/sim_pop.R

@@ -115,28 +115,146 @@ sim_N0 <- function(N0 = "exp", plot = FALSE) {
 
 
 
+#' Convert length to length group
+#'
+#' @description Helper function for converting lengths to length groups
+#' (Note: this isn't a general function; the output midpoints defining the
+#' groups alligns with DFO specific method/labeling)
+#'
+#' @param length       Interval from \code{\link{base::findInterval}}
+#' @param group        Length group used to cut the length data
+#'
+#' @export
+#'
+
+group_lengths <- function(length, group) {
+  breaks <- seq(0, max(length, na.rm = TRUE) * 2, group)
+  interval <- findInterval(length, breaks)
+  l <- breaks[interval]
+  if (group == 0.5 | group == 1) { m <- l }
+  if (group > 1) { m <- l + (0.5 * (group - 1)) }
+  m
+}
+
+
+#' Closure for simulating length given age using von Bertalanffy notation
+#'
+#' This function outputs a function which holds the parameter values supplied and
+#' the function either simulates lengths given ages or generates a length age key
+#' give a sequence of ages.
+#'
+#' @param Linf          Mean asymptotic length
+#' @param L0            Length at birth
+#' @param K             Growth rate parameter
+#' @param log_sd        Standard deviation of the relationship in log scale
+#' @param length_group  Length group for length age key (note that labels on the matrix produced are
+#'                      midpoints using the DFO conventions; see \code{\link{group_lengths}})
+#' @param digits        Integer indicating the number of decimal places to round the values to
+#' @param plot          Produce a simple plot of the simulated values?
+#'
+#' @export
+#'
+
+sim_vonB <- function(Linf = 120, L0 = 5, K = 0.1, log_sd = 0.1,
+                     length_group = 3, digits = 0, plot = FALSE) {
+
+  function(age = NULL, length_age_key = FALSE) {
+
+    pred_length <- Linf - (Linf - L0) * exp(-K * age)
+
+    if (length_age_key) {
+      breaks <- seq(0, max(pred_length) * 10, length_group)
+      breaks[1] <- 0.0000001
+      lak <- matrix(NA, ncol = length(pred_length), nrow = length(breaks) - 1,
+                    dimnames = list(length = group_lengths(breaks, length_group)[-length(breaks)],
[…]

### 2018-09-14 R/sim_survey.R

@@ -1,28 +1,4 @@
 
-
-#' Convert length to length group
-#'
-#' @description Helper function for converting lengths to length groups
-#' (Note: this isn't a general function; the output midpoints defining the
-#' groups alligns with DFO specific method/labeling)
-#'
-#' @param length       Interval from \code{\link{base::findInterval}}
-#' @param group        Length group used to cut the length data
-#'
-#' @export
-#'
-
-group_lengths <- function(length, group) {
-  breaks <- seq(0, max(length, na.rm = TRUE) * 2, group)
-  interval <- findInterval(length, breaks)
-  l <- breaks[interval]
-  if (group == 0.5 | group == 1) { m <- l }
-  if (group > 1) { m <- l + (0.5 * (group - 1)) }
-  m
-}
-
-
-
 #' Closure for simulating logistic curve
 #'
 #' @description This closure is useful for simulating q inside the
@@ -44,30 +20,6 @@ sim_logistic <- function(k = 2, x0 = 3, plot = FALSE) {
 }
 
 
-#' Closure for simulating length given age using von Bertalanffy notation
-#'
-#' @param Linf     Mean asymptotic length
-#' @param L0       Length at birth
-#' @param K        Growth rate parameter
-#' @param log_sd   Standard deviation of the relationship in log scale
-#' @param digits   Integer indicating the number of decimal places to round the values to
-#' @param plot     Produce a simple plot of the simulated values?
-#'
-#' @export
-#'
-
-sim_vonB <- function(Linf = 120, L0 = 5, K = 0.1, log_sd = 0.1,
-                     digits = 0, plot = FALSE) {
-  function(age = NULL) {
-    pred_length <- Linf - (Linf - L0) * exp(-K * age)
-    log_length <- rnorm(length(age), log(pred_length), sd = log_sd)
-    length <- round(exp(log_length), digits)
-    if (plot) plot(age, length)
-    length
-  }
-}
-
-
 #' Round simulated population
 #'
 #' @param sim Simulation from \code{\link{sim_distribution}}
@@ -80,6 +32,8 @@ round_sim <- function(sim) {
   N <- tapply(sim$sp_N$N, list(sim$sp_N$age, sim$sp_N$year), sum)
   N <- N[rownames(sim$N), colnames(sim$N)]
   dimnames(N) <- dimnames(sim$N)
+  sim$N_at_length <- convert_N(N_at_age = N,
[…]

### 2018-09-14 R/test_surveys.R

@@ -54,22 +54,28 @@ expand_surveys <- function(set_den = c(0.5, 1, 2, 5, 10) / 1000,
                             samp_totals$sim <- samp_totals$sim + (j * n_sims - n_sims)
                             total_strat_error <- res$total_strat_error
                             total_strat_error$sim <- total_strat_error$sim + (j * n_sims - n_sims)
+                            length_strat_error <- res$length_strat_error
+                            length_strat_error$sim <- length_strat_error$sim + (j * n_sims - n_sims)
                             age_strat_error <- res$age_strat_error
                             age_strat_error$sim <- age_strat_error$sim + (j * n_sims - n_sims)
                             list(samp_totals = samp_totals,
                                  total_strat_error = total_strat_error,
+                                 length_strat_error = length_strat_error,
                                  age_strat_error = age_strat_error)
                           }
     stopCluster(cl) # stop parallel process
 
     samp_totals <- data.table::rbindlist(lapply(loop_error, `[[`, "samp_totals"))
     total_strat_error <- data.table::rbindlist(lapply(loop_error, `[[`, "total_strat_error"))
+    length_strat_error <- data.table::rbindlist(lapply(loop_error, `[[`, "length_strat_error"))
     age_strat_error <- data.table::rbindlist(lapply(loop_error, `[[`, "age_strat_error"))
     samp_totals$survey <- i
     total_strat_error$survey <- i
+    length_strat_error$survey <- i
     age_strat_error$survey <- i
     survey_error[[i]]$samp_totals <- samp_totals
     survey_error[[i]]$total_strat_error <- total_strat_error
+    survey_error[[i]]$length_strat_error <- length_strat_error
     survey_error[[i]]$age_strat_error <- age_strat_error
 
     if (!is.null(export)) {
@@ -90,11 +96,16 @@ expand_surveys <- function(set_den = c(0.5, 1, 2, 5, 10) / 1000,
   sim$surveys <- data.table::data.table(surveys)
   sim$samp_totals <- do.call(rbind, lapply(survey_error, `[[`, "samp_totals"))
   sim$total_strat_error <- data.table::rbindlist(lapply(survey_error, `[[`, "total_strat_error"))
+  sim$length_strat_error <- data.table::rbindlist(lapply(survey_error, `[[`, "length_strat_error"))
   sim$age_strat_error <- data.table::rbindlist(lapply(survey_error, `[[`, "age_strat_error"))
   sim$total_strat_error_stats <- sim$total_strat_error[, list(MAE = mean(abs(error)),
[…]

### 2018-09-14 analysis/cod_sim.R

@@ -13,7 +13,10 @@ pop <- sim_abundance(ages = 1:20,
                      Z = sim_Z(mean = 0.5,
                                log_sd = 0.2,
                                phi_age = 0.9,
-                               phi_year = 0.5)) %>%
+                               phi_year = 0.5),
+                     growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1,
+                                       length_group = 3,
+                                       digits = 0)) %>%
   sim_distribution(grid = sim_grid(x_range = c(-140, 140),
                                    y_range = c(-140, 140),
                                    res = c(3.5, 3.5),
@@ -48,7 +51,6 @@ sim <- test_surveys(pop,
                     n_loops = 200,
                     cores = 6,
                     q = sim_logistic(k = 2, x0 = 3),
-                    growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
                     export = "analysis/cod_sim_exports")
 # sim <- resume_test(dir = "analysis/cod_sim_exports/2018-09-04_test")
 setMKLthreads() # turn hyperthreading on again

### 2018-09-14 tests/informal_tests.R

@@ -1,29 +1,60 @@
 
+
 library(SimSurvey)
 
+## Simulate cod-like population
+## See "imitate_cod_data.R" file for details on parameter choices
+
 set.seed(438)
-pop <- sim_abundance(years = 1:10, R = sim_R(mean = 1e+09)) %>%
-  sim_distribution(grid = sim_grid(res = c(3.5, 3.5)),
-                   space_covar = sim_sp_covar(range = 50, sd = 0.1),
-                   ay_covar = sim_ay_covar(sd = 10, phi_age = 0.5, phi_year = 0.5),
-                   depth_par = sim_parabola(alpha = 0, sigma = 50))
-# vis_sim(pop)
+pop <- sim_abundance(ages = 1:20,
+                     years = 1:20,
+                     R = sim_R(mean = 30000000,
+                               log_sd = 0.5,
+                               random_walk = TRUE),
+                     Z = sim_Z(mean = 0.5,
+                               log_sd = 0.2,
+                               phi_age = 0.9,
+                               phi_year = 0.5),
+                     growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1,
+                                       length_group = 3,
+                                       digits = 0)) %>%
+  sim_distribution(grid = sim_grid(x_range = c(-140, 140),
+                                   y_range = c(-140, 140),
+                                   res = c(3.5, 3.5),
+                                   shelf_depth = 200,
+                                   shelf_width = 100,
+                                   depth_range = c(0, 1000),
+                                   n_div = 1,
+                                   strat_breaks = seq(0, 1000, by = 20),
+                                   strat_splits = 2),
+                   ays_covar = sim_ays_covar(sd = 2.8,
+                                             range = 300,
+                                             phi_age = 0.5,
+                                             phi_year = 0.9,
+                                             group_ages = 5:20),
+                   depth_par = sim_parabola(mu = 200,
+                                            sigma = 70))
 
-survey <- sim_survey(pop, n_sims = 10) %>%
-  run_strat() %>% strat_error()
+error <- pop %>%
+  sim_survey(n_sims = 10) %>%
+  run_strat() %>%
+  strat_error()
+rm(error)
 
+## Test a series of surveys
+## Simulate surveys and compare stratified estimates to the true index
 setMKLthreads(1) # turn off MKL hyperthreading
[…]

### 2018-09-18 analysis/paper/sim_survey_paper.Rmd
Work to speed up the test_surveys loop and improve RAM management; use data.table::fwrite for export

@@
 library(data.table)
-load("analysis/cod_sim_exports/2018-09-04_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-09-14_test/test_output.RData")
 ```

### 2018-09-18 R/test_surveys.R

@@ -25,92 +25,117 @@ expand_surveys <- function(set_den = c(0.5, 1, 2, 5, 10) / 1000,
 
 ## helper functions for test_surveys and resume_test
 .test_loop <- function(sim = NULL, surveys = NULL, n_sims = NULL,
-                       n_loops = NULL, cores = NULL, export = NULL,
-                       export_dir = NULL, survey_error = NULL, ...) {
+                       n_loops = NULL, cores = NULL, export_dir = NULL,
+                       complete = NULL, keep_details = NULL, ...) {
 
-  if (is.null(survey_error)) {
-    survey_error <- vector("list", nrow(surveys))
+  ## Containers
+  samp_totals <- total_strat_error <- length_strat_error <- age_strat_error <- vector("list", nrow(surveys))
+  if (is.null(complete) && !is.null(export_dir)) {
+    dir.create(file.path(export_dir, "samp_totals"), showWarnings = FALSE)
+    dir.create(file.path(export_dir, "total_strat_error"), showWarnings = FALSE)
+    dir.create(file.path(export_dir, "length_strat_error"), showWarnings = FALSE)
+    dir.create(file.path(export_dir, "age_strat_error"), showWarnings = FALSE)
   }
-  incomplete <- which(sapply(survey_error, is.null))
+  if (is.null(complete)) {
+    complete <- rep(FALSE, nrow(surveys))
+  }
 
-  pb <- progress::progress_bar$new(
-    format = "Survey :current of :total [:bar] :percent in :elapsed (eta: :eta)",
-    total = nrow(surveys), clear = FALSE, show_after = 0, width = 100)
-  invisible(pb$tick(min(incomplete) - 1))
+  if (any(!complete)) {
 
-  for (i in incomplete) {
+    incomplete <- which(!complete)
 
-    cl <- makeCluster(cores) # use parallel computation
-    registerDoParallel(cl)
-    loop_error <- foreach(j = seq(n_loops),
-                          .packages = "SimSurvey") %dopar% {
-                            res <- sim_survey(sim, n_sims = n_sims,
-                                              set_den = surveys$set_den[i],
-                                              lengths_cap = surveys$lengths_cap[i],
-                                              ages_cap = surveys$ages_cap[i],
-                                              ...) %>%
-                              run_strat() %>% strat_error()
-                            samp_totals <- res$samp_totals
-                            samp_totals$sim <- samp_totals$sim + (j * n_sims - n_sims)
-                            total_strat_error <- res$total_strat_error
[…]

### 2018-09-18 analysis/cod_sim.R

@@ -49,59 +49,62 @@ sim <- test_surveys(pop,
                     keep_details = 98,
                     n_sims = 5,
                     n_loops = 200,
-                    cores = 6,
+                    cores = 7,
                     q = sim_logistic(k = 2, x0 = 3),
-                    export = "analysis/cod_sim_exports")
-# sim <- resume_test(dir = "analysis/cod_sim_exports/2018-09-04_test")
+                    export_dir = "analysis/cod_sim_exports/2018-09-17_age_clust_test")
+# sim <- resume_test(export_dir = "analysis/cod_sim_exports/2018-09-17_age_clust_test")
 setMKLthreads() # turn hyperthreading on again
 
 
-## visualize results
-load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
-vis_sim(sim)
+# ## visualize results
+# load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
+# vis_sim(sim)
 
 
-## Exports for Geoff -----------------------------------------------------------
-
-library(data.table)
-
-## ~ current sampling protocol
-base_case <- sim_survey_parallel(pop, n_sims = 5, n_loops = 5, cores = 6,
-                                 set_den = 2e-03, lengths_cap = 500, ages_cap = 10,
-                                 q = sim_logistic(k = 2, x0 = 3),
-                                 growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
-                                 quiet = FALSE)
-setdet <- base_case$setdet[, list(year, x, y, depth, strat, n, sim, set)]
-samp <- base_case$samp[base_case$samp$measured, ]
-samp$age[!samp$aged] <- NA
-samp$measured <- samp$aged <- NULL
-samp <- merge(setdet[, list(year, sim, set)], samp, by = "set")
-fwrite(setdet, file = "analysis/cod_sim_exports/2018-09-12_for_Geoff/01_setdet.csv")
-fwrite(samp, file = "analysis/cod_sim_exports/2018-09-12_for_Geoff/01_samp.csv")
-rm(base_case)
-gc()
-
-## One example of a protocol that results in bias
-bias_case <- sim_survey_parallel(pop, n_sims = 1, n_loops = 25, cores = 6,
-                                 set_den = 1e-02, lengths_cap = 10, ages_cap = 10,
-                                 q = sim_logistic(k = 2, x0 = 3),
[…]

### 2018-09-18 tests/informal_tests.R

@@ -44,15 +44,17 @@ rm(error)
 ## Test a series of surveys
 ## Simulate surveys and compare stratified estimates to the true index
 setMKLthreads(1) # turn off MKL hyperthreading
-surveys <- expand_surveys(set_den = c(10) / 1000,
+surveys <- expand_surveys(set_den = c(5, 10) / 1000,
                           lengths_cap = c(10, 1000),
                           ages_cap = c(10))
 sim <- test_surveys(pop,
                     surveys = surveys,
-                    n_sims = 5,
-                    n_loops = 100,
-                    cores = 7,
-                    q = sim_logistic(k = 2, x0 = 3))
+                    n_sims = 1,
+                    n_loops = 1,
+                    cores = 1,
+                    q = sim_logistic(k = 2, x0 = 3),
+                    export_dir = "tests/exports")
+# sim <- resume_test(dir = "tests/exports")
 setMKLthreads() # turn hyperthreading on again
 
 

### 2018-09-20 analysis/paper/sim_survey_paper.Rmd
Add two new figures to the paper given new length-based output

@@
 library(data.table)
-load("analysis/cod_sim_exports/2018-09-14_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-09-17_age_clust_test/test_output.RData")
 ```
@@
 ## little helper function for the fan plot
-fan_plot <- function(d, x = "Age", total_strat = FALSE) {
+fan_plot <- function(d, x = "Age", which_strat = "Age") {
 
-  if (total_strat) {
+  if (which_strat == "Total") {
     ints_by = c("year", "set_den")
-  } else {
+  }
+  if (which_strat == "Age") {
     ints_by = c("year", "age", "set_den", "lengths_cap")
+  }
+  if (which_strat == "Length") {
+    ints_by = c("year", "length", "set_den", "lengths_cap")
   }
@@
     p <- ggplot(data = ints, aes(x = age))
-    xlabs <- c("2", "4", "6", "8+")
-    xbreaks <- seq(2, 8 , 2)
+    xlabs <- c("2", "4", "6", "8", "10+")
+    xbreaks <- seq(2, 10, 2)
+  }
+  if (x == "Length") {
+    p <- ggplot(data = ints, aes(x = length))
+    xbreaks <- xlabs <- seq(0, 100, 20)
+    xlabs[xlabs == 100] <- "100+"
   }
@@
 
-  if (total_strat) {
+  if (which_strat == "Total") {
     p <- p + facet_wrap(~ set_den, nrow = 1,
@@
 $$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
-Here, numbers-at-age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e.
+Here, numbers at age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e.
 recruits) vary around a baseline value, $log(N_{1,y}) = log(\mu_r) + \epsilon_{y}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(\mu_Z) + \delta_{a,y}$.
@@
 total mortality can be made to be more similar for fish that are closer together in age and time).
-Abundance-at-age is then converted to abundance-at-length using the original von Bertalanffy growth curve [@Cailliet2006]:
+Abundance at age is then converted to abundance at length using the original von Bertalanffy growth curve [@Cailliet2006]:
 
@@
 Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$.
-Abundance in discrete length groups, $l_{group}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance-at-age to abundance-at-length.
+Abundance in discrete length groups, $l_{group}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance at age to abundance at length.
 Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.
@@
 
-The next step in the simulation is to distribute the abundance-at-age matrix simulated using the cohort model throughout a spatial field.
+The next step in the simulation is to distribute the abundance at age matrix simulated using the cohort model throughout a spatial field.
 Here, a grid of $s$ locations was generated with a resolution of $R$ and a depth gradient.
@@
 A series of `r nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests).
-Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain annual estimates of total abundance ($\hat{I}_{y,k}$), abundance-at-length ($\hat{I}_{l,y,k}$) and abundance-at-age ($\hat{I}_{a,y,k}$).
-We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance-at-age estimates from each survey:
+Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain annual estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
+We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:
 
@@
 catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$).
+RMSE was also calculated for abundance at length estimates.
 Intra-haul correlation (i.e.
@@
 
-- Feedback and advice: Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Dwayne Pittman, Don Power, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
+- Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Dwayne Pittman, Don Power, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
 - Dave Cote for a review of a previous version of this manuscript
@@
 
-p <- fan_plot(d, x = "Year", total_strat = TRUE)
+p <- fan_plot(d, x = "Year", which_strat = "Total")
 cowplot::save_plot(plot = p, filename = "analysis/paper/figures/total_strat_true_vs_est.pdf",
@@
 <br>
+
+```{r length, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+The heavy dark line indicates the true population available to the survey."}
+
+which_set_dens <- c(0.0005, 0.002, 0.01)
+which_len_caps <- c(10, 100, 1000)
+which_year <- 7
+
+sub_surveys <- sim$surveys[sim$survey$set_den %in% which_set_dens &
+                             sim$survey$lengths_cap %in% which_len_caps & 
+                             sim$survey$ages_cap == 10, ]
+sub_strat <- sim$length_strat_error[eval(sim$length_strat_error$survey %in% sub_surveys$survey &
+                                           sim$length_strat_error$year == which_year), ]
+d <- merge(sub_surveys, sub_strat, by = "survey")
+d$length[d$length > 100] <- 100 # aggregate beyond 100 cm
+d <- d[, list(I_hat = sum(I_hat), I = sum(I)), 
+        by = c("year", "length", "sim", "survey", "set_den", "lengths_cap")]
+
+p <- fan_plot(d, x = "Length", which_strat = "Length")
+cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_length.pdf",
+                   base_height = 6, base_width = 8)
+p
+
+```
 
@@
 
-d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
+sub_surveys <- sim$surveys[sim$survey$set_den %in% which_set_dens &
+                             sim$survey$lengths_cap %in% which_len_caps & 
+                             sim$survey$ages_cap == 10, ]
+sub_strat <- sim$age_strat_error[eval(sim$age_strat_error$survey %in% sub_surveys$survey &
+                                        sim$age_strat_error$year == which_year), ]
+d <- merge(sub_surveys, sub_strat, by = "survey")
 d <- d[d$year == which_year & d$set_den %in% which_set_dens &
          d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
-d$age[d$age > 8] <- 8 # aggregate beyond 8
+d$age[d$age > 10] <- 10 # aggregate beyond 10
 d <- d[, list(I_hat = sum(I_hat), I = sum(I)), 
@@
 
-```{r error, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
+```{r error-lines, cache = TRUE, fig.height = 3.8, fig.width = 8.5, results = "hide", fig.cap = "- Root-mean-squared error (RMSE) of abundance at length and abundance at age estimates as a function of the number of fish measured ($N_{lengths}$) from an array of surveys with different sampling protocol.
+The labels represent surveys with different set densities ($D_{sets}$), where number of sets are shown in brackets, and the lines are color coded by length sampling protocol ($M_{lengths}$).
+Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios."}
+
+totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught),
+                                 n_measured = mean(n_measured), n_aged = mean(n_aged)),
+                          by = "survey"]
+
+errors <- merge(sim$surveys, sim$length_strat_error_stats, by = "survey")
+length_d <- merge(errors, totals, by = "survey")
+length_d <- length_d[length_d$ages_cap == 10, ]
+length_d$strat <- "Abundance at length"
+
+errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
+age_d <- merge(errors, totals, by = "survey")
+age_d <- age_d[age_d$ages_cap == 10, ]
+age_d$strat <- "Abundance at age"
+
+d <- rbind(length_d, age_d)
+d$cap <- factor(d$lengths_cap)
+d$strat <- factor(d$strat, levels = c("Abundance at length", "Abundance at age"))
+
+lab_d <- d[lengths_cap == max(lengths_cap),
+           list(lengths_cap, n_measured, n_sets, cap, RMSE),
+           by = c("set_den", "strat")]
+lab_d$lab <- paste0(formatC(lab_d$set_den), " (", lab_d$n_sets, ")")
+
+p <- ggplot(data = d, aes(x = n_measured, y = RMSE, group = set_den)) +
+  geom_line(aes(color = cap), size = 2, lineend = "round") +
+  geom_text(data = lab_d, aes(label = lab), nudge_x = 500, hjust = 0) +
+  geom_point(data = lab_d, aes(color = cap)) +
+  xlab(bquote(N[lengths])) +
+  scale_y_continuous(labels = format_si()) +
+  scale_x_continuous(labels = format_si(), limits = c(0, max(d$n_measured) * 1.3),
+                     breaks = pretty(d$n_measured, 6)) +
+  scale_color_viridis(name = bquote(M[lengths]), discrete = TRUE) +
+  facet_wrap(~ strat, scales = "free", dir = "h") +
+  theme_bw() +
+  theme(legend.justification = c(0, 1),
+        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+        panel.border = element_rect(fill = NA, color = "darkgrey", size = 0.4),
+        axis.ticks = element_line(color = "darkgrey", size = 0.4))
+
+cowplot::save_plot(plot = p, filename = "analysis/paper/figures/error_lines.pdf",
+                   base_height = 3.8, base_width = 8.5)
+p
+
+```
+
+<br>
+
+```{r error-wireframe, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
 Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$)."}
@@
 ## this export holds the same distribution test
-load("analysis/cod_sim_exports/2018-09-07_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-09-18_no_age_clust_test/test_output.RData")
 ```

### 2018-09-24 analysis/paper/sim_survey_paper.Rmd
Draft results section of paper

@@
 
-A dynamic population that varies across ages, time and space is simulated with the parameter settings used here (Table \@ref(tab:par)).
+A dynamic population that varies across ages, time and space was simulated with the parameter settings used here (Table \@ref(tab:par)).
 Figure \@ref(fig:sp) depicts the aggregate distribution of the population in one year of the simulation and the corresponding catches of one stratified random survey over this population.
@@
 
+The stratified analyses run on these simulated data provided estimates of total abundance, abundance at length and abundance at age, and different stages of sampling affect these estimates.
+Specifically, set density affects all estimates, length sampling affects estimates of abundance at length and abundance at age, and age sampling affects abundance at age estimates.
+Figure \@ref(fig:tot) depicts the consequences of different set densities on the estimates of total abundance available to the survey through the simulated series.
+It is clear from this figure that the probability envelopes are tighter (i.e.
+estimates are more precise) and bias is negligible under the higher set density scenarios ($D_{sets}$ = 0.002, and $D_{sets}$ = 0.01).
+Trends in total abundance appear to be slightly underestimated under the low set density scenario ($D_{sets}$ = 0.0005; Figure \@ref(fig:tot)).
+Likewise, abundance at length and abundance at age estimates were generally biased downward in all low set density scenarios ($D_{sets}$ = 0.0005; Figures \@ref(fig:length), \@ref(fig:age), \@ref(fig:year)).
+At higher set densities, abundance at length estimates are more precise and bias is less apparent (Figure \@ref(fig:length)).
+With the exception of a length sampling protocol of 10 measurements per set ($M_{lengths}$ = 10), there appears to be little difference in the estimates of abundance at length when the number of measurements are capped at 100 ($M_{lengths}$ = 100) or 1000 ($M_{lengths}$ = 1000; Figure \@ref(fig:length)).
+Some downward bias is apparent in the abundance at length estimates, especially at larger lengths, under the medium set density scenario with low length sampling effort ($D_{sets}$ = 0.002 and $M_{lengths}$ = 10; Figure \@ref(fig:length)).
+Upward or downward bias in abundance at age estimates is apparent in all cases except the high set density and high length sampling scenario ($D_{sets}$ = 0.01 and $M_{lengths}$ = 1000; Figure \@ref(fig:age), \@ref(fig:year)).
+Under both low and medium set densities ($D_{sets}$ = 0.0005 and 0.02, respectively), increasing length sampling effort from 10 or 100 measurements per set to 1000 appears to increase bias or reduce precision slightly (Figure \@ref(fig:age), \@ref(fig:year)).
+
+For a less visual approach to assessing precision and bias, RMSE results are presented in Figures \@ref(fig:error-lines) and \@ref(fig:error-wireframe) and Table \@ref(tab:comp).
+Figure \@ref(fig:error-lines) depicts the effects of set density and length sampling effort on abundance at length and abundance at age estimates.
+In all cases, increases in set density results in large declines in RMSE (Figure \@ref(fig:error-lines)).
+For abundance at length estimates, RMSE declines as the number of lengths measured increases, however, gains plateau beyond a cretin point.
+Likewise, RMSE of abundance at age estimates decline with increased length sampling effort, but RMSE does not plateau in all set density scenarios (Figure \@ref(fig:error-lines)).
+While RMSE continues to decline with increased length sampling effort under the high set density scenarios ($D_{sets}$ = 0.005 to 0.01 sets / km^2^), RMSE reaches a plateau and then increases with increased length sampling effort under the medium to low set density scenarios ($D_{sets}$ = 0.002 to 0.0005 sets / km^2^; Figure \@ref(fig:error-lines)).
+In many cases, RMSE for both metrics is lower for scenarios with higher set density but lower length sampling effort (Figure \@ref(fig:error-lines)).
+Similar patterns are apparent in Figure \@ref(fig:error-wireframe) and Table \@ref(tab:comp), which focus on the consequences of set density and length and age sampling protocol on the estimates of abundance at age.
+Again, RMSE clearly declines with increases to set density and sometimes declines with increased length sampling effort (Figure \@ref(fig:error-wireframe)).
+Under both low and medium set density scenarios ($D_{sets}$ = 0.005 to 0.01 sets / km^2^), measuring a maximum of 100 fish per set appears to be better than measuring a maximum of 1000 fish per set (Figure \@ref(fig:error-wireframe); Table \@ref(tab:comp)).
+Decreasing the age sampling protocol below 10 ages sampled per set per length group appears to result in relatively small increases to RMSE (Figure \@ref(fig:error-wireframe); Table \@ref(tab:comp)).
+
+# Discussion
+
 ### TODO
 
-- Next discuss trend...
 - Look into the ictopelagic sampling design literature (suggestion from Pierre)
@@
 
-# Discussion
-
 # Acknowledgements
 
-- Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Dwayne Pittman, Don Power, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
+- Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
 - Dave Cote for a review of a previous version of this manuscript

### 2018-09-25 analysis/paper/sim_survey_paper.Rmd
Tweak intro, add assumptions section and start discussion

@@
 Few studies, however, have tested these conclusions using simulations.
+Simulations are a useful approach for informing the design of surveys [@sutherland2006] and they have been used widely for exploring the number of samples needed to obtain estimates with reasonable precision [e.g.
+distance sampling surveys; @thomas2010].
 A key advantage of a pure simulation approach is that "true" population size and all stages of sampling can be controlled (set, length and age sampling protocol).
@@
 
+## Assumptions
+
+Like any model, this simulation is a simplification of a much more complex reality.
+For instance, the population is assumed to be uniformly distributed within a cell when, in the real-world, there are likely finer-scale aggregations.
+The survey is also an instantaneous snapshot of the population, meaning that the population is assumed to be in the same location from the beginning to the end of the survey.
+Also, fish are aged at random within length bins and ages are estimated without error.
+Finally, area trawled is assumed to be perfectly standard.
+This is a short, non-inclusive, list of additional variability that likely occurs in the real-world but was not implemented here for simplicity.
+
 ## Simulation settings and analysis
@@
 
+The simple simulation presented here revealed some expected and unexpected patterns.
+First, there were clear improvements to precision in all population estimates as the number of sets were increased.
+This result was expected as it aligns with a basic principle of survey design that the precision and representativeness of a survey is increased by increasing the number of sampling units [@sutherland2006].
+
 ### TODO
@@
 - Look into the ictopelagic sampling design literature (suggestion from Pierre)
-- Mention that simulation is an approach widely applied to inform sampling design (program MARK, distance, RandomForest, etc.); i.e. cite the gods
 - Look into work done on occupancy modeling and sampling design for such studies (i.e. book suggestion by Aaron)

### 2018-09-26 analysis/paper/sim_survey_paper.Rmd
Tweak results section, continue to work on discussion and drop Appendix 2

@@
 It is clear from this figure that the probability envelopes are tighter (i.e.
-estimates are more precise) and bias is negligible under the higher set density scenarios ($D_{sets}$ = 0.002, and $D_{sets}$ = 0.01).
-Trends in total abundance appear to be slightly underestimated under the low set density scenario ($D_{sets}$ = 0.0005; Figure \@ref(fig:tot)).
-Likewise, abundance at length and abundance at age estimates were generally biased downward in all low set density scenarios ($D_{sets}$ = 0.0005; Figures \@ref(fig:length), \@ref(fig:age), \@ref(fig:year)).
-At higher set densities, abundance at length estimates are more precise and bias is less apparent (Figure \@ref(fig:length)).
+estimates are more precise) and bias is negligible.
+Likewise, abundance at length and abundance at age estimates are more precise as set density is increased (Figure \@ref(fig:length), \@ref(fig:age), \@ref(fig:year)).
 With the exception of a length sampling protocol of 10 measurements per set ($M_{lengths}$ = 10), there appears to be little difference in the estimates of abundance at length when the number of measurements are capped at 100 ($M_{lengths}$ = 100) or 1000 ($M_{lengths}$ = 1000; Figure \@ref(fig:length)).
-Some downward bias is apparent in the abundance at length estimates, especially at larger lengths, under the medium set density scenario with low length sampling effort ($D_{sets}$ = 0.002 and $M_{lengths}$ = 10; Figure \@ref(fig:length)).
+Some downward bias is apparent in the abundance at length estimates, especially at larger lengths, under the low to medium set density scenarios with low length sampling effort ($D_{sets}$ = 0.0005 or 0.002 and $M_{lengths}$ = 10; Figure \@ref(fig:length)).
 Upward or downward bias in abundance at age estimates is apparent in all cases except the high set density and high length sampling scenario ($D_{sets}$ = 0.01 and $M_{lengths}$ = 1000; Figure \@ref(fig:age), \@ref(fig:year)).
@@
 First, there were clear improvements to precision in all population estimates as the number of sets were increased.
-This result was expected as it aligns with a basic principle of survey design that the precision and representativeness of a survey is increased by increasing the number of sampling units [@sutherland2006].
+This result was expected as it aligns with a basic principle of survey design that the precision and representativeness of a survey increases as the number of sampling units are increased [@sutherland2006].
+Second, stratified estimates of abundance at age were often biased and, in some cases, estimates were poorer when sub-sampling effort was increased.
+This result was unexpected because design-based estimators, such as the stratified analysis applied here, are expected to be unbiased [@smith1990] and increases to sub-sampling effort was expected to be relatively ineffective, not detrimental.
+Results from stratified estimates of abundance at length, in contrast, better align with these expectations.
+By deduction, these results indicate that the issue stems from the intervening age-length key and not the design-based estimator.
+
+Using an age-length key in conjunction with the length distribution to estimate the age distribution is standard procedure in the analysis of fisheries-independent survey data as only a small fraction of the catch is typically aged.
+This is because the aging procedure is costly and time-consuming whereas length measurements are relatively easy to obtain.
+Ages are generally determined from length-stratified sub-samples of the catch and raw proportions of age-at-length are used to assign ages to fish in specific length groups.
+Age-length keys are rarely stratified due to a small number of samples and, as such, they are usually constructed at larger spatial scales.
+Here we construct one age-length key for the division, as is done for the analysis of 3Ps cod.
+There is, however, a potential cost to the spatial scale of the key.
+Namely, it is unlikely that one age-length key is representative for the whole region because the probability of being a specific age given length varies in space [@Berg2012].
+This would not be an issue if there was no age-specific clustering, but this seems unlikely given the common observation that fish caught together tend to be more similar than those in the general population [@Pennington1994; @Pennington2002; @Stewart2014].
+Because different age groups sometimes occur in different places, the translation of lengths to ages may be biased by the samples used to generate the age-length key.
+This bias is perhaps compounded by the length-stratified sampling of ages such that the ages sampled may be skewed towards sets with the most catch, especially under scenarios where length sampling effort is high.
+If this is the case, then the representativeness of the age-length key to the whole population could be deminished by excessive length sampling.
+
+Requires further research. Solutions to the problem... (e.g. berge and kristensen).
 
@@
 
-# Appendix 2: results in the absence of intra-haul correlation
-
-In addition to the run presented in body of this paper, which simulates surveys over a population that forms age-specific clusters, we also simulated surveys over a population that assumes the same spatial clustering across all age groups.
-There is therefore no intra-haul correlation in this series of surveys.
-Aside from this change to the spatial distribution settings, all other settings were kept the same.
-Results are presented below in Figures A1, A2 and A3.
-Aside from systematic underestimation in the low set density scenarios, there are little signs of clear bias in the stratified estimates.
-The bias present under the low set density scenario is likely related to the disproportionate sampling of the narrow shallow and deep strata in the survey grid.
-The lack of bias from the other scenarios help verify the issues introduced by age-specific clustering to stratified estimates of abundance at age.
-These results also indicate that little precision is gained by increasing length or age sampling effort beyond 100 lengths per set and 10 ages per length group, respectively.
-Given this result, it is probably not unreasonable to speculate that such sampling effort helps define the shape of the age distribution, but extra effort does little to improve estimates of abundance at age; such gains are only obtained by increasing set density.
-
-```{r appen-data}
-## this export holds the same distribution test
-load("analysis/cod_sim_exports/2018-09-18_no_age_clust_test/test_output.RData")
-```
-
-```{r appen-age, cache = TRUE, fig.width = 8, fig.height = 6}
-
-which_set_dens <- c(0.0005, 0.002, 0.01)
-which_len_caps <- c(10, 100, 1000)
-which_year <- 7
-
-d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
-d <- d[d$year == which_year & d$set_den %in% which_set_dens &
-         d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
-d$age[d$age > 8] <- 8 # aggregate beyond 8
-d <- d[, list(I_hat = sum(I_hat), I = sum(I)), 
-       by = c("year", "age", "sim", "survey", "set_den", "lengths_cap")]
-
-p <- fan_plot(d, x = "Age")
-cowplot::save_plot(plot = p, filename = "analysis/paper/figures/appen_true_vs_est_by_age.pdf",
-                   base_height = 6, base_width = 8)
-p
-
-```
-Figure A1 - Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
-Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
-The heavy dark line indicates the true population available to the survey.
-
-<br>
-
-```{r appen-year, cache = TRUE, fig.width = 8, fig.height = 6}
-
-which_set_dens <- c(0.0005, 0.002, 0.01)
-which_len_caps <- c(10, 100, 1000)
-which_age <- 4
-
-d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
-d <- d[d$age == which_age & d$set_den %in% which_set_dens &
-       d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
-
-p <- fan_plot(d, x = "Year")
-cowplot::save_plot(plot = p, filename = "analysis/paper/figures/appen_true_vs_est_by_year.pdf",
-                   base_height = 6, base_width = 8)
-p
-
-```
-Figure A2 - Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
-Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
-The heavy dark line indicates the true trend in the population available to the survey.
-
-<br>
-
-```{r appen-error, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide"}
-
-d <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
-# d <- d[d$set_den != 0.0005, ]
-d <- d[d$set_den %in% c(0.0005, 0.002, 0.01)]
-set_den <- sort(unique(d$set_den))
-d$set_den <- factor(d$set_den)
-options(scipen = 999)
-labs <- sapply(seq_along(set_den), function(i) as.expression(bquote(D[sets] == .(set_den[i]))), 
-               USE.NAMES = FALSE)
-zlabs <- list(at = pretty(d$RMSE), labels = si(pretty(d$RMSE)))
-
-pdf("analysis/paper/figures/appen_error_wireframe.pdf", height = 3, width = 8.5)
-
-p <- wireframe(RMSE ~  lengths_cap * ages_cap | set_den, data = d,
-          as.table = TRUE, drape = TRUE, col = "white", lwd = 0.5,
-          strip = strip.custom(factor.levels = labs),
-          scales = list(arrows = FALSE, cex = 0.7, z = zlabs),
-          col.regions = viridis::viridis(100),
-          screen = list(z = 230, x = 300), 
-          xlab = list(expression(M[lengths]), rot = 25), 
-          ylab = list(expression(M[ages]), rot = -15), 
-          zlab = list("RMSE", rot = 95),
-          colorkey = list(labels = zlabs, height = 0.6,
-                          axis.line = list(col = "darkgrey")),
-          par.settings = trellis_theme)
-p <- update(p, layout = c(3, 1))
-p
-grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc"))
-
-invisible(dev.off())
-
-p
-grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc"))
-
-```
-Figure A3 - Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
-Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$).
-
 <br>

### 2018-09-27 analysis/paper/sim_survey_paper.Rmd
Finish drafting paper!

@@
 Like any model, this simulation is a simplification of a much more complex reality.
-For instance, the population is assumed to be uniformly distributed within a cell when, in the real-world, there are likely finer-scale aggregations.
+For instance, the population is assumed to aggregate by age-class and be uniformly distributed within a cell when, in the real-world, they may aggregate by length and form finer-scale clusters.
 The survey is also an instantaneous snapshot of the population, meaning that the population is assumed to be in the same location from the beginning to the end of the survey.
@@
 
-Using an age-length key in conjunction with the length distribution to estimate the age distribution is standard procedure in the analysis of fisheries-independent survey data as only a small fraction of the catch is typically aged.
+Using an age-length key in conjunction with the length distribution to estimate abundance at age is standard procedure in the analysis of fisheries-independent survey data as only a small fraction of the catch is typically aged.
 This is because the aging procedure is costly and time-consuming whereas length measurements are relatively easy to obtain.
 Ages are generally determined from length-stratified sub-samples of the catch and raw proportions of age-at-length are used to assign ages to fish in specific length groups.
-Age-length keys are rarely stratified due to a small number of samples and, as such, they are usually constructed at larger spatial scales.
+Age-length keys are usually constructed at larger spatial scales because there are rarely enough samples to stratify these data.
 Here we construct one age-length key for the division, as is done for the analysis of 3Ps cod.
@@
 Namely, it is unlikely that one age-length key is representative for the whole region because the probability of being a specific age given length varies in space [@Berg2012].
-This would not be an issue if there was no age-specific clustering, but this seems unlikely given the common observation that fish caught together tend to be more similar than those in the general population [@Pennington1994; @Pennington2002; @Stewart2014].
+This would not be an issue if there was no size (age) specific clustering, however, this may not always be the case because different size groups often occur in different places because of ontogenetic habitat shifts [@dahlgren2000; @galaiduk2017].
+For instance, it is not uncommon for populations to form distinct nursery and spawning areas [@marteinsdottir2000; @booth2000].
 Because different age groups sometimes occur in different places, the translation of lengths to ages may be biased by the samples used to generate the age-length key.
 This bias is perhaps compounded by the length-stratified sampling of ages such that the ages sampled may be skewed towards sets with the most catch, especially under scenarios where length sampling effort is high.
-If this is the case, then the representativeness of the age-length key to the whole population could be deminished by excessive length sampling.
+If this is the case, then the representativeness of the age-length key to the whole population could be diminished by excessive length sampling.
 
-Requires further research. Solutions to the problem... (e.g. berge and kristensen).
+The negative consequences of excessive length sampling on abundance at age estimates are most apparent under the low set density scenarios.
+The precision of the estimates are also the poorest when fewer sets are conducted.
+Sets, however, are the most costly sample to obtain and, as such, the set densities of fisheries-independent surveys tend to be on the lower end of the scenarios presented in this paper.
+The set density of the multi-species survey conducted by Fisheries and Oceans Canada in the Newfoundland region, for instance, ranges between 0.001 and 0.002 sets / km^2^.
+Length sampling effort, in contrast, tends to be on the higher end of the scenarios presented here because lengths are relatively cheap and easy to obtain.
+Again, using Fisheries and Oceans Canada sampling protocol from the Newfoundland region as an example, it is not uncommon for length sampling effort to be capped at 500 measurements per set.
+The results presented in this paper paradoxically suggest that better abundance at age estimates are obtained by lowering the length sampling cap.
+Results also indicate that reductions to length sampling effort would not significantly impact abundance at length estimates.
+Nonetheless, it is hard to be prescriptive because this simulation is far from a perfect reflection of reality and it focuses on one case study.
+Further research is required on the consequences of some of the assumptions of this simulation as well as the interaction between age-specific clustering, length-stratified sampling and the age-length key.
+Ideally, the simulation would also include multiple species with different life-histories and distributions, and also integrate a cost component to assess the trade-offs between information and cost.
+Finally, it would be interesting to test alternate analyses of these data that may account for the spatial structure of the age-length key [e.g.
+@Berg2012].
 
-### TODO
-
-- Look into the ictopelagic sampling design literature (suggestion from Pierre)
-- Look into work done on occupancy modeling and sampling design for such studies (i.e. book suggestion by Aaron)
+While much work has yet to be done, results from the simple simulation echo the growing body of literature which concludes that extra sub-sampling is an ineffective means of improving estimates relative to sampling more locations [@Pennington1994; @Pennington2002; @Stewart2014; @Bogstad1995; @Coggins2013; @Zhang2013].
+In general, the most significant source of variation in fisheries-independent surveys stems from set-to-set variation, and not variability from individual sub-samples.
+This is largely because fish caught together tend to be more similar than those in the general population [@Pennington1994].
+Therefore, if the goal of a trawl survey is to maximize information, it is likely better to stop collecting psudoreplicates and, instead, focus efforts on other species or the next set.
 
@@
 - Dave Cote for a review of a previous version of this manuscript
+- NSERC visiting-fellow program
 

### 2018-09-28 analysis/paper/sim_survey_paper.Rmd
Tweak abstract

@@
 By varying the sampling protocol, we demonstrate that increasing the number of sets conducted per stratum provides the greatest gains in precision when sampling a clustered population.
-Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent.
-These results indicate that length and age sampling efforts for some species(e.g.
-cod *Gadus morhua*) in trawl surveys conducted by regional fisheries management organizations can be scaled back without significant losses to precision.
+Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent and improvements were typically marginal, and sometimes squandered, when sub-sampling effort was increased.
+These results indicate that high levels of length and age sampling effort can be scaled back without significant losses to precision and, in some cases, estimates of abundance at age may be improved by reducing sub-sampling effort.
 Time saved by reducing said sampling may afford more time to conduct additional sets and/or increase sampling efforts of other species.
@@
 
-**Keywords** - length and age distribution; simulation; spatial correlation; stratified analysis; statistical assessment models; survey design
+**Keywords** - length and age distribution; simulation; spatial correlation; stratified analysis; survey design
 
@@
 
-- Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
+- Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
 - Dave Cote for a review of a previous version of this manuscript

### 2018-10-25 inst/rmd/vis_sim.Rmd
Update vis_sim function and add example to help file

--- 2018-06-03 inst/rmd/vis_sim.Rmd
@@
 
-Row
+Row {.tabset}
 --------------------------------------------------------------------------------
@@
 
-Row
---------------------------------------------------------------------------------
+### Recruitment
+
+```{r}
+plot_trend(sim, sum_ages = min(sim$ages))
+```
+
+### Total mortality
+
+```{r}
+plot_surface(sim, mat = "Z")
+```
 
@@
 
-### Total mortality
+### Abundance at length
 
 ```{r}
-plot_surface(sim, mat = "Z")
+plot_surface(sim, mat = "N_at_length", xlab = "Length", zlab = "N")
 ```
@@
 ```
+
+Survey { `r if(!"setdet" %in% names(sim)) ".hidden"`}
+================================================================================
+
+Column {.sidebar}
+--------------------------------------------------------------------------------
+
+```{r}
+sliderInput("samp_year", "Year", 
+            min = min(sim$years), max = max(sim$years), 
+            value = 1, step = 1)
+sliderInput("samp_sim", "Simulation", 
+            min = min(sim$setdet$sim), max = max(sim$setdet$sim), 
+            value = 1, step = 1)
+```
+
+Column
+--------------------------------------------------------------------------------
+
+### Survey sampling
+
+```{r, eval = "setdet" %in% names(sim)}
+renderPlotly({
+  plot_survey(sim, which_year = input$samp_year, which_sim = input$samp_sim)
+})
+```
+
+Estimates { `r if(!"age_strat_error" %in% names(sim)) ".hidden"`}
+================================================================================
+
+Column {.sidebar}
+--------------------------------------------------------------------------------
+
+```{r}
+selectInput("set_den", "Set density", 
+            choices = sort(unique(sim$surveys$set_den)))
+selectInput("lengths_cap", "Lengths cap", 
+            choices = sort(unique(sim$surveys$lengths_cap)))
+selectInput("ages_cap", "Ages cap", 
+            choices = sort(unique(sim$surveys$ages_cap)))
+selectInput("fan_length", "Length", 
+            choices = sim$lengths)
+selectInput("fan_age", "Age", 
+            choices = sim$ages)
+```
+
+Column {.tabset}
+--------------------------------------------------------------------------------
+
+### Total abundance
+
+```{r, eval = "age_strat_error" %in% names(sim)}
+renderPlotly({
+  s <- sim$surveys
+  i <- s$survey[s$set_den == input$set_den & 
+                        s$lengths_cap == max(s$lengths_cap) &
+                        s$ages_cap == max(s$ages_cap)]
+  plot_total_strat_fan(sim, surveys = i)
+})
+```
+
+### Abundance at length
+
+```{r, eval = "age_strat_error" %in% names(sim)}
+renderPlotly({
+  s <- sim$surveys
+  i <- s$survey[s$set_den == input$set_den & 
+                        s$lengths_cap == input$lengths_cap &
+                        s$ages_cap == max(s$ages_cap)]
+  plot_length_strat_fan(sim, surveys = i, years = sim$years,
+                        lengths = input$fan_length, select_by = "length")
+})
+```
+
+### Abundance at age
+
+```{r, eval = "age_strat_error" %in% names(sim)}
+renderPlotly({
+  s <- sim$surveys
+  i <- s$survey[s$set_den == input$set_den & 
+                        s$lengths_cap == input$lengths_cap &
+                        s$ages_cap == input$ages_cap]
+  plot_age_strat_fan(sim, surveys = i, years = sim$years,
+                        ages = input$fan_age, select_by = "age")
+})
+```
+
+Error { `r if(!"surveys" %in% names(sim)) ".hidden"`}
+================================================================================
+
+Column  {.tabset}
+--------------------------------------------------------------------------------
+
+### Error surface by rule
+
+```{r, eval = "surveys" %in% names(sim)}
+plot_error_surface(sim, plot_by = "rule")
+```
+
+### Error surface by sample size
+
+```{r, eval = "surveys" %in% names(sim)}
+plot_error_surface(sim, plot_by = "samples")
+```

### 2018-10-25 R/plot_sim.R

@@ -9,6 +9,7 @@
 #'                       \code{\link{sim_distribution}}, etc.
 #' @param mat            Name of matrix in \code{sim} list to plot.
 #' @param xlab,ylab,zlab Axes labels.
+#' @param sum_ages       Sum across these ages
 #' @param ages           Subset data to one or more ages.
 #' @param lengths        Subset data to one or more length groups.
 #' @param years          Subset data to one or more years.
@@ -31,11 +32,16 @@
 #'
 #' @export
 #' @rdname plot_trend
-plot_trend <- function(sim, col = viridis::viridis(1), ...) {
-  tot <- data.frame(Year = sim$years, N = colSums(sim$N))
+plot_trend <- function(sim, sum_ages = sim$ages, col = viridis::viridis(1), ...) {
+  if (length(sum_ages) == 1) {
+    n <- sim$N[sum_ages, ]
+  } else {
+    n <- colSums(sim$N[sum_ages, ])
+  }
+  tot <- data.frame(Year = sim$years, N = n)
   tot$text <- paste0("Year: ", tot$Year, "\nN: ", round(tot$N))
-  p1 <- plot_ly(x = ~Year, y = ~N, text = ~text, hoverinfo = "text",
-                type = "scatter", mode = "lines", color = I(col), data = tot, ...)
+  plot_ly(x = ~Year, y = ~N, text = ~text, hoverinfo = "text",
+          type = "scatter", mode = "lines", color = I(col), data = tot, ...)
 }
 
 #' @export

### 2018-10-25 R/vis_sim.R

@@ -6,6 +6,27 @@
 #'                       \code{\link{sim_survey}} or \code{\link{test_surveys}}.
 #' @param ...            Additional arguments to send to \link[rmarkdown]{run}
 #'
+#' @examples
+#'
+#' pop <- sim_abundance(ages = 1:20, years = 1:20)
+#' vis_sim(pop)
+#'
+#' dist <- sim_distribution(pop, grid = make_grid(res = c(10, 10)))
+#' vis_sim(dist)
+#'
+#' ## Run one survey design
+#' survey <- sim_survey(dist, n_sims = 5)
+#' vis_sim(survey)
+#'
+#' ## Run several survey designs and assess stratified estimates
+#' surveys <- expand_surveys(set_den = c(1, 2) / 1000,
+#'                           lengths_cap = c(100, 500),
+#'                           ages_cap = c(5, 20))
+#' tests <- test_surveys(dist, surveys = surveys, keep_details = 1,
+#'                       n_sims = 10, n_loops = 50, cores = 2)
+#' vis_sim(tests)
+#'
+#'
 #' @export
 #'
 

### 2018-10-25 analysis/cod_sim.R
Revisit cod_sim following package edits

@@ -51,13 +51,13 @@ sim <- test_surveys(pop,
                     n_loops = 200,
                     cores = 7,
                     q = sim_logistic(k = 2, x0 = 3),
-                    export_dir = "analysis/cod_sim_exports/2018-09-17_age_clust_test")
-# sim <- resume_test(export_dir = "analysis/cod_sim_exports/2018-09-17_age_clust_test")
+                    export_dir = "analysis/cod_sim_exports/2018-10-26_age_clust_test")
+# sim <- resume_test(export_dir = "analysis/cod_sim_exports/2018-10-26_age_clust_test")
 setMKLthreads() # turn hyperthreading on again
 
 
 # ## visualize results
-# load("analysis/cod_sim_exports/2018-07-13_test/test_output.RData")
+# load("analysis/cod_sim_exports/2018-10-26_age_clust_test/test_output.RData")
 # vis_sim(sim)
 
 
@@ -152,13 +152,13 @@ sim <- test_surveys(pop,
                     n_loops = 200,
                     cores = 7,
                     q = sim_logistic(k = 2, x0 = 3),
-                    export = "analysis/cod_sim_exports/2018-09-18_no_age_clust_test")
-# sim <- resume_test(dir = "analysis/cod_sim_exports/2018-09-18_no_age_clust_test")
+                    export = "analysis/cod_sim_exports/2018-10-27_no_age_clust_test")
+# sim <- resume_test(dir = "analysis/cod_sim_exports/2018-10-27_no_age_clust_test")
 setMKLthreads() # turn hyperthreading on again
 
 
 # ## visualize results
-# load("analysis/cod_sim_exports/2018-07-16_test/test_output.RData")
+# load("analysis/cod_sim_exports/2018-10-27_no_age_clust_test/test_output.RData")
 # vis_sim(sim)
 
 

### 2018-10-25 analysis/imitate_cod_data.R

@@ -109,7 +109,7 @@ pop <- sim_abundance(ages = 1:20,
                                phi_age = 0.9,
                                phi_year = 0.5),
                      growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0)) %>%
-  sim_distribution(grid = sim_grid(x_range = c(-140, 140),
+  sim_distribution(grid = make_grid(x_range = c(-140, 140),
                                    y_range = c(-140, 140),
                                    res = c(3.5, 3.5),
                                    shelf_depth = 200,
@@ -188,7 +188,7 @@ plot(length ~ age, data = sim_I, main = "simulated data",
 ## (could use to improve the accuracy of depth in the real data, but the pattern is clear)
 data_I <- setdet
 sim_I <- survey$setdet
-plot(as.numeric(data_I$max.depth), data_I$number, xlab = "depth",
+plot(as.numeric(data_I$set.depth.min), data_I$number, xlab = "depth",
      ylab = "number", main = "real data", xlim = c(0, 1000))
 plot(sim_I$depth, sim_I$n, xlab = "depth",
      ylab = "number", main = "simulated data", xlim = c(0, 1000))

### 2018-11-05 analysis/paper/sim_survey_paper.Rmd
Draft a new table 1 that includes function calls, descriptions and parameter symbols

@@
 ```{r setup, include=FALSE}
-
 # devtools::install_github("rstudio/rmarkdown@v1.10")
-
 knitr::opts_chunk$set(echo = FALSE,
@@
 library(data.table)
-load("analysis/cod_sim_exports/2018-09-17_age_clust_test/test_output.RData")
-```
-
-```{r fun}
-
-## http://www.moeding.net/archives/32-Metric-prefixes-for-ggplot2-scales.html
-format_si <- function(...) {
-  # Format a vector of numeric values according
-  # to the International System of Units.
-  # http://en.wikipedia.org/wiki/SI_prefix
-  #
-  # Based on code by Ben Tupper
-  # https://stat.ethz.ch/pipermail/r-help/2012-January/299804.html
-  # Args:
-  #   ...: Args passed to format()
-  #
-  # Returns:
-  #   A function to format a vector of strings using
-  #   SI prefix notation
-  #
-
-  function(x) {
-    limits <- c(1e-24, 1e-21, 1e-18, 1e-15, 1e-12,
-                1e-9,  1e-6,  1e-3,  1e0,   1e3,
-                1e6,   1e9,   1e12,  1e15,  1e18,
-                1e21,  1e24)
-    prefix <- c("y",   "z",   "a",   "f",   "p",
-                "n",   "µ",   "m",   " ",   "k",
-                "M",   "G",   "T",   "P",   "E",
-                "Z",   "Y")
-
-    # Vector with array indices according to position in intervals
-    i <- findInterval(abs(x), limits)
-
-    # Set prefix to " " for very small values < 1e-24
-    i <- ifelse(i == 0, which(limits == 1e0), i)
-
-    paste0(format(round(x/limits[i], 1),
-                 trim = TRUE, scientific = FALSE, ...),
-          prefix[i])
-  }
-}
-si <- format_si()
-
-## little helper function for the fan plot
-fan_plot <- function(d, x = "Age", which_strat = "Age") {
-
-  if (which_strat == "Total") {
-    ints_by = c("year", "set_den")
-  }
-  if (which_strat == "Age") {
-    ints_by = c("year", "age", "set_den", "lengths_cap")
-  }
-  if (which_strat == "Length") {
-    ints_by = c("year", "length", "set_den", "lengths_cap")
-  }
-
-  ## quantiles for fan plot
-  quants <- seq(90, 10, by = -10)
-  ints <- lapply(quants, function(q) {
-    d[, list(prob = paste0(q, "%"),
-             lower = quantile(I_hat, prob = (1 - q / 100) / 2),
-             upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)), 
-      by = ints_by]
-  })
-  ints <- rbindlist(ints)
-  ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
-
-  true <- d[d$sim %in% 1:3, ]
-
-  if (x == "Age") {
-    p <- ggplot(data = ints, aes(x = age))
-    xlabs <- c("2", "4", "6", "8", "10+")
-    xbreaks <- seq(2, 10, 2)
-  }
-  if (x == "Length") {
-    p <- ggplot(data = ints, aes(x = length))
-    xbreaks <- xlabs <- seq(0, 100, 20)
-    xlabs[xlabs == 100] <- "100+"
-  }
-  if (x == "Year") {
-    p <- ggplot(data = ints, aes(x = year))
-    xlabs <- waiver()
-    xbreaks <- waiver()
-  }
-
-  p <-  p + 
-    geom_ribbon(aes(ymin = lower, ymax = upper, fill = prob, alpha = prob)) +
-    geom_line(aes(y = I), data = true, color = "black", size = 0.6) +
-    scale_fill_viridis(name = "Pr", discrete = TRUE, begin = 0.1) +
-    scale_alpha_discrete(name = "Pr", range = c(0.5, 1)) + 
-    xlab(x) + ylab("Abundance index") +
-    scale_y_continuous(labels = format_si()) +
-    scale_x_continuous(expand = c(0.01, 0), 
-                       labels = xlabs, 
-                       breaks = xbreaks) +
-    guides(color = guide_legend(keywidth = 0.4, keyheight = 0.25, 
-                                default.unit = "cm", ncol = 1))
-
-  if (which_strat == "Total") {
-    p <- p + facet_wrap(~ set_den, nrow = 1,
-                        labeller = label_bquote(cols = D[sets]==.(set_den)))
-  } else {
-    p <- p + facet_grid(lengths_cap ~ set_den, 
-                        labeller = label_bquote(cols = D[sets]==.(set_den),
-                                                rows = M[lengths]==.(lengths_cap)))
-  }
-
-  p + theme_bw() +
-    theme(legend.justification = c(0, 1),
-          plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"),
-          panel.grid.major = element_blank(), 
-          panel.grid.minor = element_blank(),
-          strip.background = element_blank(),
-          panel.border = element_rect(fill = NA, color = "darkgrey", size = 0.4),
-          axis.ticks = element_line(color = "darkgrey", size = 0.4))
-
-}
-
+load("analysis/cod_sim_exports/2018-10-26_age_clust_test/test_output.RData")
+source("analysis/paper/helper_functions.R")
 ```
@@
 
+The SimSurvey package was written in the programming language R [@R] and, in short, the package allows for the simulation of stratified-random surveys of an age-structured population that varies in space and time.
+SimSurvey relies heavily on functions from the data.table [@dowle2017], raster [@hijmans2016] and plotly [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively.
+All the source R code behind SimSurvey is available on GitHub (https://github.com/PaulRegular/SimSurvey) and the 0.1.0 release of the package can be installed via GitHub:
+
+```{r, echo = TRUE, eval = FALSE}
+install.packages("devtools")
+devtools::install_github("PaulRegular/SimSurvey@v0.1.0") 
+```
+
+## Case study
+
+The construction of the SimSurvey package was motivated by a regional need for assessing the sampling design of trawl surveys conducted along the Newfoundland and Labrador shelf by Fisheries and Oceans Canada.
+As a tangible first step, we focused our efforts on emulating data from the survey of cod in NAFO division 3Ps since it is one of the most dynamic and heavily sampled stocks in the region.
+Below we outline the equations and SimSurvey functions used to simulate survey data with similar features to data obtained by the survey of 3Ps cod.
+The settings were iteratively modified until it was difficult to distinguish the simulated data from the real survey data (see Table \@ref(tab:par) for parameter settings and descriptions).
+Here we evaluate the efficacy of alternate sampling protocol by assessing deviation between stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
+While assessing model-based analyses was an option, we focused on the design-based stratified analysis for simplicity and because of its widespread use.
+
 ## Simulate abundance
 
-The simulation presented here starts with a common cohort model where the abundance at age $a$ in year $y$ ($N_{a,y}$) equals the abundance of that cohort in the previous year multiplied by the survival rate, which is expressed in terms of total mortality ($Z_{a,y}$):
+The simulation starts with a common cohort model where the abundance at age $a$ in year $y$ ($N_{a,y}$) equals the abundance of that cohort in the previous year multiplied by the survival rate, which is expressed in terms of total mortality ($Z_{a,y}$):
 
@@
 
+Abundance is simulated using the `sim_abundance` function and the below code was used to simulate the case study population:
+
+```{r, echo = TRUE, eval = FALSE}
+set.seed(438) # set seed for reproducible results
+n <- sim_abundance(ages = 1:20,
+                   years = 1:20,
+                   R = sim_R(mean = 30000000,
+                             log_sd = 0.5,
+                             random_walk = TRUE),
+                   Z = sim_Z(mean = 0.5,
+                             log_sd = 0.2,
+                             phi_age = 0.9,
+                             phi_year = 0.5),
+                   growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1,
+                                     log_sd = 0.1, length_group = 3,
+                                     digits = 0))
+```
+
+The resultant object, `n`, is a list with data such as the abundance at age matrix (named `N`).
+Note that the nested functions `sim_R`, `sim_Z` and `sim_vonB` are actually "closures" (i.e.
+functions that contain data and return functions).
+These closures avoid the repeated specifications of ages and years and also allows users to use their own closures with a similar structure but different underlying formula.
+
 ## Simulate spatial distribution
@@
 The depth gradient was set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$.
+The grid is divided into $N_{div}$ divisions and $N_{strat}$ depth-based strata.
 The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space
@@
 
+The `make_grid` function is used to generate the grid used in the case study, and the `sim_distribution` function distributes the population, `n`, simulated above using `sim_abundance` through the grid:
+
+```{r, echo = TRUE, eval = FALSE}
+g <- make_grid(x_range = c(-140, 140),
+               y_range = c(-140, 140),
+               res = c(3.5, 3.5),
+               shelf_depth = 200,
+               shelf_width = 100,
+               depth_range = c(0, 1000),
+               n_div = 1,
+               strat_breaks = seq(0, 1000, by = 40),
+               strat_splits = 2)
+
+d <- sim_distribution(n, grid = g,
+                      ays_covar = sim_ays_covar(sd = 2.8,
+                                                range = 300,
+                                                phi_age = 0.5,
+                                                phi_year = 0.9,
+                                                group_ages = 5:20),
+                      depth_par = sim_parabola(mu = 200,
+                                               sigma = 70))
+```
+
+This function retains all the data simulated by `sim_abundance` and adds a `data.table` with abundance split by age, year and cell (named `sp_N`).
+Again, note the use of the closures `sim_ays_covar` and `sim_parabola` to impose the age-year-space covariance and the relationship with depth, respectively.
+
 ## Simulate survey
@@
 The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated using formula described above.
-First, the grid is divided into $N_{div}$ divisions and $N_{strat}$ depth-based strata.
 The area of each strata is calculated and this is used to define the sets to allocate to each strata under a particular set density, $D_{sets}$.
@@
 
-## Assumptions
+The function `sim_survey` can be used to simulate data from one survey over a population simulated using `sim_distribution`:
 
-Like any model, this simulation is a simplification of a much more complex reality.
-For instance, the population is assumed to aggregate by age-class and be uniformly distributed within a cell when, in the real-world, they may aggregate by length and form finer-scale clusters.
-The survey is also an instantaneous snapshot of the population, meaning that the population is assumed to be in the same location from the beginning to the end of the survey.
-Also, fish are aged at random within length bins and ages are estimated without error.
-Finally, area trawled is assumed to be perfectly standard.
-This is a short, non-inclusive, list of additional variability that likely occurs in the real-world but was not implemented here for simplicity.
+```{r, echo = TRUE, eval = FALSE}
+s <- sim_survey(d, n_sims = 1,
+                q = sim_logistic(k = 2, x0 = 3),
+                trawl_dim = c(1.5, 0.02),
+                set_den = 2 / 1000,
+                lengths_cap = 500,
+                length_group = 1,
+                ages_cap = 10)
+```
 
-## Simulation settings and analysis
+Again, this function retains all the objects listed in `d` and adds objects that detail the set locations (`setdet`) and sampling details (`samp`).
 
-Ideally the parameter settings used here would be based on model fits to real data, however, conditioning the operating model described above was outside the scope of this project.
-Moreover, many of the parameters described above may not be identifiable or it may be difficult to obtain un-biased estimates [e.g.
-bias introduced by length-stratified sampling complicate the fitting of growth models; @Morgan1997].
-As such, parameters were informally selected to roughly correspond to observations typically obtained by the Fisheries and Oceans Canada survey of cod in NAFO division 3Ps, Newfoundland.
-Though this approach is far from exact, the goal was to approximate the life-history and distribution of an actual population to simulate an arena for testing the consequences of surveys with different sampling protocol.
-Parameter settings and descriptions used in this case study are presented in Table \@ref(tab:par).
+## Stratified analysis
+
 A series of `r nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests).
@@
 All simulations and analyses were conducted using `r substr(version$version.string, 1, 15)` [@R2017] and the code has been developed into a package called `SimSurvey` to facilitate future work and collaboration (https://github.com/PaulRegular/SimSurvey).
+
+## Assumptions
+
+Like any model, this simulation is a simplification of a much more complex reality.
+For instance, the population is assumed to aggregate by age-class and be uniformly distributed within a cell when, in the real-world, they may aggregate by length and form finer-scale clusters.
+The survey is also an instantaneous snapshot of the population, meaning that the population is assumed to be in the same location from the beginning to the end of the survey.
+Also, fish are aged at random within length bins and ages are estimated without error.
+Finally, area trawled is assumed to be perfectly standard.
+This is a short, non-inclusive, list of additional variability that likely occurs in the real-world but was not implemented here for simplicity.
 
@@
 
-Table: (\#tab:par) - Parameter descriptions, symbols and global settings used in the cod case study.
+Table: (\#tab:par) - Main function calls, with argument descriptions and associated parameter symbols, used to simulate the cod-like population used in the case study presented here.
 
-|                                       **Parameter name** |                 **Symbol** |    **Setting** |
-|:-------------------------------------------------------- | :------------------------- | :------------- |
-|                                              *Abundance* |                            |                |  
-|                                                     Ages |                        $a$ |           1:20 |
-|                                                    Years |                        $y$ |           1:20 |
-|                                         Mean recruitment |                    $\mu_r$ |            30M |
-|                    Standard deviation of log-recruitment |                 $\sigma_r$ |            0.5 |
-|                                     Mean total mortality |                    $\mu_Z$ |            0.7 |
-|                    Standard deviation of total mortality |                 $\sigma_Z$ |            0.2 |
-|  Correlation across ages in error around total mortality |    $\varphi_{\delta, age}$ |            0.9 |
-| Correlation across years in error around total mortality |   $\varphi_{\delta, year}$ |            0.5 |
-|                                   Mean asymptotic length |               $l_{\infty}$ |         120 cm |
-|                                          Length at birth |                      $l_0$ |           5 cm |
-|                                              Growth rate |                        $K$ |            0.1 |
-|                     Length group for abundance at length |            $l_{group}^{N}$ |           3 cm |
-|                                                          |                            |                |  
-|                                                   *Grid* |                            |                | 
-|                                               Grid cells |                        $s$ |         1:6400 |
-|                                               Resolution |                        $R$ |         3.5 km |
-|                                              Shelf depth |                $d_{shelf}$ |          200 m |
-|                                              Shelf width |                $w_{shelf}$ |         100 km |
-|                                              Depth range |       $[d_{min}, d_{max}]$ |  [0 m, 1000 m] |
-|                                      Number of divisions |                  $N_{div}$ |              1 |
-|                                         Number of strata |                $N_{strat}$ |             42 |
-|                                                          |                            |                |  
-|                                           *Distribution* |                            |                | 
-|        Standard deviation of age-year-space distribution |             $\sigma_{\xi}$ |            2.8 |
-|                        Smoothness of spatial correlation |                  $\lambda$ |              1 |
-|                             Range of spatial correlation |                        $r$ |          50 km |
-|          Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |           0.5* |
-|         Correlation across years in spatial distribution |      $\varphi_{\xi, year}$ |            0.9 |
-|            Depth at which abundance is typically highest |                    $\mu_d$ |          200 m |
-|                Dispersion around depth of peak abundance |                 $\sigma_d$ |           70 m |
-|                                                          |                            |                |  
-|                                                 *Survey* |                            |                |
-|                                  Area covered by a trawl |                $A_{trawl}$ |     0.03 km^2^ |
-|              Steepness of logistic curve of catchability |                        $k$ |              2 |
-|               Midpoint of logistic curve of catchability |                      $x_0$ |              3 |
-|                            Length group for age sampling |                $l_{group}$ |           1 cm |  
-\* Distribution set to be equal for ages 5+
+|                                       **Function call** |                                            **Description** |                 **Symbol** |
+|:------------------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
+| `> set.seed(438)`                                       |      Set random number generator seed to replicate results |                            |
+| `>`                                                     |                                                            |                            |
+| `> n <- sim_abundance(`                                 |                                       *Simulate abundance* |                            |         
+| `+   ages = 1:20,`                                      |                                                       Ages |                        $a$ |
+| `+   years = 1:20,`                                     |                                                      Years |                        $y$ |
+| `+   R = sim_R(mean = 30000000,`                        |                                           Mean recruitment |                    $\mu_r$ |
+| `+             log_sd = 0.5),`                          |                      Standard deviation of log-recruitment |                 $\sigma_r$ |
+| `+   Z = sim_Z(mean = 0.5,`                             |                                       Mean total mortality |                    $\mu_Z$ |
+| `+             log_sd = 0.2,`                           |                      Standard deviation of total mortality |                 $\sigma_Z$ |
+| `+             phi_age = 0.9,`                          |    Correlation across ages in error around total mortality |    $\varphi_{\delta, age}$ |
+| `+             phi_year = 0.5),`                        |   Correlation across years in error around total mortality |   $\varphi_{\delta, year}$ |
+| `+   growth = sim_vonB(Linf = 120,`                     |                                Mean asymptotic length (cm) |               $l_{\infty}$ |
+| `+                     L0 = 5,`                         |                                       Length at birth (cm) |                      $l_0$ |
+| `+                     K = 0.1,`                        |                                                Growth rate |                        $K$ |
+| `+                     log_sd = 0.1,`                   |      Standard devation of the von Bertalanffy growth curve |                 $\sigma_l$ |
+| `+                     length_group = 3))`              |                  Length group for abundance at length (cm) |            $l_{group}^{N}$ |      
+| `>`                                                     |                                                            |                            |
+| `> g <- make_grid(`                                     |                                              *Make a grid* |                            |
+| `+   x_range = c(-140, 140),`                           |                           Range of grid in the x dimension |                            |                   
+| `+   y_range = c(-140, 140),`                           |                           Range of grid in the y dimension |                            | 
+| `+   res = c(3.5, 3.5),`                                |                                       Grid resolution (km) |                        $R$ |
+| `+   shelf_depth = 200,`                                |                                           Shelf depth (km) |                $d_{shelf}$ |
+| `+   shelf_width = 100,`                                |                                           Shelf width (km) |                $w_{shelf}$ |
+| `+   depth_range = c(0, 1000),`                         |                                            Depth range (m) |       $[d_{min}, d_{max}]$ |
+| `+   n_div = 1,`                                        |                                        Number of divisions |                  $N_{div}$ |
+| `+   strat_breaks = seq(0, 1000, by = 40))`             |                 Series of depth breaks for defining strata |                            |
+| `>`                                                     |                                                            |                            |
+| `> d <- sim_distribution(`                              |                                    *Simulate distribution* |                            |     
+| `+   sim = n,`                                          |                Object produced by `sim_abundance` function |                            |     
+| `+   grid = g,`                                         |                    Object produced by `make_grid` function |                            |
+| `+   ays_covar = sim_ays_covar(sd = 2.8,`               |          Standard deviation of age-year-space distribution |             $\sigma_{\xi}$ |
+| `+                             range = 300,`            |                          Range of spatial correlation (km) |                        $r$ |
+| `+                             phi_age = 0.5,`          |            Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |
+| `+                             phi_year = 0.9,`         |           Correlation across years in spatial distribution |      $\varphi_{\xi, year}$ |
+| `+                             group_ages = 5:20),`     |       Make space-age-year variance equal across these ages |                            |
+| `+   depth_par = sim_parabola(mu = 200,`                |          Depth at which abundance is typically highest (m) |                    $\mu_d$ |
+| `+                            sigma = 70))`             |              Dispersion around depth of peak abundance (m) |                 $\sigma_d$ |
+| `>`                                                     |                                                            |                            |
+| `> e <- expand_surveys(`                                |                       *Define a series of surveys to test* |                            |
+| `+   set_den = c(0.0005, 0.001, 0.002, 0.005, 0.01),`   |                                       Set density (km^-2^) |                 $D_{sets}$ |
+| `+   lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),`   |                 Maximum number of lengths to collect / set |              $M_{lengths}$ |
+| `+   ages_cap = c(2, 5, 10, 20, 50))`                   | Maximum number of ages to sample / length group / division |                 $M_{ages}$ |
+| `>`                                                     |                                                            |                            |
+| `> s <- test_surveys(`                                  |                             *Simulate and analyze surveys* |                            |
+| `+   sim = d,`                                          |             Object produced by `sim_distribution` function |                            |
+| `+   surveys = e,`                                      |               Object produced by `expand_surveys` function |                            |
+| `+   n_loops = 1000,`                                   |                    Number of times to simulate each survey |                            | 
+| `+   q = sim_logistic(k = 2,`                           |                Steepness of logistic curve of catchability |                        $k$ |
+| `+                   x0 = 3),`                          |                 Midpoint of logistic curve of catchability |                      $x_0$ |
+| `+   trawl_dim = c(1.5, 0.02),`                         |       Trawl dimensions (km) - i.e. area covered by a trawl |                $A_{trawl}$ |
+| `+   length_group = 1)`                                 |                         Length group for age sampling (cm) |                $l_{group}$ |
+
+*Caution:* as written, the `test_surveys` function call may take days or weeks to run.
+Run duration can be decreased by running multiple simulations per loop (`n_sims` argument) and by running the loops in parallel (`cores` argument).
+See help file for more details (`?test_surveys`).
 
 <br>
-
-Table: (\#tab:tests) - Descriptions and settings of different survey sampling protocol tested in the cod case study. All possible combinations of the options below were tested.
-
-|      **Parameter name** |      **Symbol** |                                       **Setting** |
-|:----------------------- | :-------------- | :------------------------------------------------ |
-|             Set density |      $D_{sets}$ |    0.0005, 0.001, 0.002, 0.005, 0.01 sets / km^2^ |
-|    Length sampling rule |   $M_{lengths}$ |       5, 10, 20, 50, 100, 500, 1000 lengths / set |
-|       Age sampling rule |      $M_{ages}$ |   2, 5, 10, 20, 50 ages / length group / division |
-
-<br>
-
-Table: (\#tab:comp) - Root-mean-squared errors (RMSE) and number of samples [sets ($N_{sets}$), lengths ($N_{lengths}$) and ages ($N_{ages}$)] obtained from an array of surveys with different sampling protocol [set densities ($D_{sets}$), maximum sampling effort of lengths per set ($M_{lengths}$), and maximum number of ages to collect per length group ($M_{ages}$)].
-Differences from the minimum RMSE value ($\Delta ~ RMSE$) within each set density option are presented to aid comparisons.
-
-```{r comp, results = "asis"}
-
-totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught),
-                                 n_measured = mean(n_measured), n_aged = mean(n_aged)),
-                          by = "survey"]
-errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
-d <- merge(errors, totals, by = "survey")
-
-which_set_dens <- c(0.0005, 0.002, 0.01)
-which_len_caps <- c(10, 100, 1000)
-which_age_caps <- c(5, 10, 50)
-
-tab <- d[d$set_den %in% which_set_dens & d$lengths_cap %in% which_len_caps & 
-           d$ages_cap %in% which_age_caps, 
-         list(lengths_cap, ages_cap,
-              n_sets = round(n_sets), n_measured = round(n_measured), n_aged = round(n_aged),
-              RMSE = RMSE, delta_RMSE = RMSE - min(RMSE)), 
-         by = "set_den"]
-tab <- tab[order(set_den, delta_RMSE), ]
-tab$set_den <- formatC(tab$set_den)
-tab$set_den[tab$delta_RMSE != 0] <- NA
-tab$RMSE <- si(tab$RMSE)
-tab$delta_RMSE <- si(tab$delta_RMSE)
-
-split_tab <- split(tab, tab$n_sets)
-for (i in seq_along(split_tab)[-length(split_tab)]) {
-  split_tab[[i]] <- rbind(split_tab[[i]], data.table(RMSE = NA), fill = TRUE)
-}
-tab <- rbindlist(split_tab)
-
-names(tab) <- c("$D_{sets}$", "$M_{lengths}$", "$M_{ages}$", 
-                "$N_{sets}$", "$N_{lengths}$", "$N_{ages}$",
-                "$RMSE$", "$\\Delta ~ RMSE$")
-
-options(knitr.kable.NA = "")
-knitr::kable(tab, align = "r", format = "markdown")
-
-```
 

### 2018-11-05 analysis/cod_sim.R

@@ -152,13 +152,13 @@ sim <- test_surveys(pop,
                     n_loops = 200,
                     cores = 7,
                     q = sim_logistic(k = 2, x0 = 3),
-                    export = "analysis/cod_sim_exports/2018-10-27_no_age_clust_test")
-# sim <- resume_test(dir = "analysis/cod_sim_exports/2018-10-27_no_age_clust_test")
+                    export = "analysis/cod_sim_exports/2018-10-28_no_age_clust_test")
+# sim <- resume_test(export_dir = "analysis/cod_sim_exports/2018-10-28_no_age_clust_test")
 setMKLthreads() # turn hyperthreading on again
 
 
 # ## visualize results
-# load("analysis/cod_sim_exports/2018-10-27_no_age_clust_test/test_output.RData")
+# load("analysis/cod_sim_exports/2018-10-28_no_age_clust_test/test_output.RData")
 # vis_sim(sim)
 
 

### 2018-11-05 analysis/paper/helper_functions.R

@@ -1,0 +1,116 @@
+
+## http://www.moeding.net/archives/32-Metric-prefixes-for-ggplot2-scales.html
+format_si <- function(...) {
+  # Format a vector of numeric values according
+  # to the International System of Units.
+  # http://en.wikipedia.org/wiki/SI_prefix
+  #
+  # Based on code by Ben Tupper
+  # https://stat.ethz.ch/pipermail/r-help/2012-January/299804.html
+  # Args:
+  #   ...: Args passed to format()
+  #
+  # Returns:
+  #   A function to format a vector of strings using
+  #   SI prefix notation
+  #
+
+  function(x) {
+    limits <- c(1e-24, 1e-21, 1e-18, 1e-15, 1e-12,
+                1e-9,  1e-6,  1e-3,  1e0,   1e3,
+                1e6,   1e9,   1e12,  1e15,  1e18,
+                1e21,  1e24)
+    prefix <- c("y",   "z",   "a",   "f",   "p",
+                "n",   "µ",   "m",   " ",   "k",
+                "M",   "G",   "T",   "P",   "E",
+                "Z",   "Y")
+
+    # Vector with array indices according to position in intervals
+    i <- findInterval(abs(x), limits)
+
+    # Set prefix to " " for very small values < 1e-24
+    i <- ifelse(i == 0, which(limits == 1e0), i)
+
+    paste0(format(round(x/limits[i], 1),
+                  trim = TRUE, scientific = FALSE, ...),
+           prefix[i])
+  }
+}
+si <- format_si()
+
+## little helper function for the fan plot
+fan_plot <- function(d, x = "Age", which_strat = "Age") {
+
+  if (which_strat == "Total") {
+    ints_by = c("year", "set_den")
+  }
+  if (which_strat == "Age") {
+    ints_by = c("year", "age", "set_den", "lengths_cap")
+  }
+  if (which_strat == "Length") {
+    ints_by = c("year", "length", "set_den", "lengths_cap")
+  }
+
+  ## quantiles for fan plot
+  quants <- seq(90, 10, by = -10)
+  ints <- lapply(quants, function(q) {
+    d[, list(prob = paste0(q, "%"),
+             lower = quantile(I_hat, prob = (1 - q / 100) / 2),
+             upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)),
+      by = ints_by]
+  })
+  ints <- rbindlist(ints)
[…]

### 2018-11-15 analysis/paper/sim_survey_paper.Rmd
Demonstrate use of sim_abundance and plot_surface

@@
 library(data.table)
-load("analysis/cod_sim_exports/2018-10-26_age_clust_test/test_output.RData")
+Sys.setenv("MAPBOX_TOKEN" = "pk.eyJ1IjoicGF1bHJlZ3VsYXIiLCJhIjoiY2pueGIxejMxMGM0MjNqbjR1aXdwYm92bSJ9.R-t1SCa0VHKf4N-PQmYI5Q")
+# load("analysis/cod_sim_exports/2018-10-26_age_clust_test/test_output.RData")
 source("analysis/paper/helper_functions.R")
@@
 As a tangible first step, we focused our efforts on emulating data from the survey of cod in NAFO division 3Ps since it is one of the most dynamic and heavily sampled stocks in the region.
-Below we outline the equations and SimSurvey functions used to simulate survey data with similar features to data obtained by the survey of 3Ps cod.
-The settings were iteratively modified until it was difficult to distinguish the simulated data from the real survey data (see Table \@ref(tab:par) for parameter settings and descriptions).
+Below we outline the equations behind a series of SimSurvey functions that were used to simulate survey data with similar features to data obtained by the survey of 3Ps cod.
+The settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data.
 Here we evaluate the efficacy of alternate sampling protocol by assessing deviation between stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
 While assessing model-based analyses was an option, we focused on the design-based stratified analysis for simplicity and because of its widespread use.
+See Table \@ref(tab:par) for a description of the code used to generate the case study results.
 
@@
 
-Abundance is simulated using the `sim_abundance` function and the below code was used to simulate the case study population:
+Abundance is simulated using the `sim_abundance` function and the settings used for our case study are shown in Table \@ref(tab:par).
+These settings are also the defaults of the function.
+This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`).
+The package includes the functions `sim_R`, `sim_Z` and `sim_vonB` for simulating recruitment, total mortality and growth, respectively.
+These functions are "closures" [i.e.
+functions that contain data and return functions; @wickham2014] and this structure was chosen to avoid the repeated specifications of ages and years.
+The use of closures also allows users to use their own closures with a similar structure but different underlying formula.
+Overall, the function provides a simple tool for simulating a range dynamic age-structured populations.
+For instance, below we simulate a relatively long and short lived species:
+
+```{r, echo = TRUE, eval = TRUE}
+set.seed(438)
+a <- sim_abundance(ages = 1:20,
+                   R = sim_R(mean = 3e+07),
+                   Z = sim_Z(mean = 0.5))
+b <- sim_abundance(ages = 1:6,
+                   R = sim_R(mean = 1e+10),
+                   Z = sim_Z(mean = 0.8))
+```
+
+The `sim_abundance` function returns a list with the sequence of ages (`ages`), sequence of years (`years`), sequence of lengths (`lengths`), numbers of recruits across all years (`R`), numbers at age in the first year (`N0`), total mortality matrix (`Z`), abundance at age matrix (`N`), abundance at length matrix (`N_at_length`) and the function supplied to the `growth` argument (`sim_length`).
+The growth function is retained for later use in `sim_survey` to simulate lengths given simulated catch at age in a simulated survey.
+
+The package also includes several ploting function for making quick plotly-based [@sievert2018] interactive visuals of the simulated population.
+The plots output from the below code are combined into Figure \@ref(fig:surface).
 
 ```{r, echo = TRUE, eval = FALSE}
-set.seed(438) # set seed for reproducible results
-n <- sim_abundance(ages = 1:20,
-                   years = 1:20,
-                   R = sim_R(mean = 30000000,
-                             log_sd = 0.5,
-                             random_walk = TRUE),
-                   Z = sim_Z(mean = 0.5,
-                             log_sd = 0.2,
-                             phi_age = 0.9,
-                             phi_year = 0.5),
-                   growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1,
-                                     log_sd = 0.1, length_group = 3,
-                                     digits = 0))
+plot_surface(a, mat = "N")
+plot_surface(b, mat = "N")
 ```
 
-The resultant object, `n`, is a list with data such as the abundance at age matrix (named `N`).
-Note that the nested functions `sim_R`, `sim_Z` and `sim_vonB` are actually "closures" (i.e.
-functions that contain data and return functions).
-These closures avoid the repeated specifications of ages and years and also allows users to use their own closures with a similar structure but different underlying formula.
+```{r surface, fig.cap = "- Plots produced by `plot_surface` when supplied an list produced by a) `sim_abundance(ages = 1:20, R = sim_R(mean = 3e+07), Z = sim_Z(mean = 0.5))` and b) `sim_abundance(ages = 1:6, R = sim_R(mean = 1e+10), Z = sim_Z(mean = 0.8))`.
+The abundance at age matrix (`N`) is displayed by default; the `mat` argument can be used to specify another matrix contained within the supplied list, such as `Z`."}
+
+set.seed(438)
+a <- sim_abundance(ages = 1:20,
+                   R = sim_R(mean = 3e+07),
+                   Z = sim_Z(mean = 0.5))
+b <- sim_abundance(ages = 1:6,
+                   R = sim_R(mean = 1e+10),
+                   Z = sim_Z(mean = 0.8))
+
+pa <- plot_surface(a, mat = "N", scene = "scene1", showscale = FALSE) %>% 
+  add_annotations(x = 0.05, y = 0.9, xref = "paper", yref = "paper", showarrow = FALSE,
+                  text = "a)")
+pb <- plot_surface(b, mat = "N", scene = "scene2", showscale = FALSE) %>% 
+  add_annotations(x = 0.05, y = 0.9, xref = "paper", yref = "paper", showarrow = FALSE,
+                  text = "b)")
+
+p <- subplot(pa, pb) %>%
+  layout(scene = list(domain = list(x = c(0, 0.5), y = c(0, 1))),
+         scene2 = list(domain = list(x = c(0.5, 1), y = c(0, 1)),
+                       xaxis = list(title = "Age"),
+                       yaxis = list(title = "Year"),
+                       zaxis = list(title = "N")))
+
+export_plot(p, file = "analysis/paper/figures/plot_surface.png", width = 800, height = 370)
+knitr::include_graphics("figures/plot_surface.png")
+```
 
@@
 
-A series of `r nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests).
-Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain annual estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
+A series of `r # nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests).
+Each survey was simulated `r # max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain annual estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
 We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:
@@
 
-# Figures
+<!-- # Figures -->
 
-```{r sp, cache = TRUE, fig.height = 3.8, fig.width = 8.5, fig.cap = "- An example from year seven of the simulation of a) the distribution of total abundance available to the survey, and b) locations of the sets conducted and total numbers caught in a stratified random survey with a set density of 0.002 sets / km^2^.
-Grey lines indicate depth-based survey strata."}
+<!-- ```{r sp, cache = TRUE, fig.height = 3.8, fig.width = 8.5, fig.cap = "- An example from year seven of the simulation of a) the distribution of total abundance available to the survey, and b) locations of the sets conducted and total numbers caught in a stratified random survey with a set density of 0.002 sets / km^2^.
+Grey lines indicate depth-based survey strata."} -->
 
-which_year <- 7
-which_sim <- 1
+<!-- which_year <- 7 -->
+<!-- which_sim <- 1 -->
 
-sp_N_tot <- sim$sp_I[year == which_year & sim == which_sim, list(N = sum(I)), by = c("year", "cell")]
-xyz <- merge(sim$grid_xy, sp_N_tot, by = "cell")
-size <- unique(diff(pretty(xyz$N, 25)))
+<!-- sp_N_tot <- sim$sp_I[year == which_year & sim == which_sim, list(N = sum(I)), by = c("year", "cell")] -->
+<!-- xyz <- merge(sim$grid_xy, sp_N_tot, by = "cell") -->
+<!-- size <- unique(diff(pretty(xyz$N, 25))) -->
 
-sp_strat <- raster::rasterToPolygons(sim$grid$strat, dissolve = TRUE)
-df_strat <- suppressWarnings(fortify(sp_strat) %>% group_by(group))
+<!-- sp_strat <- raster::rasterToPolygons(sim$grid$strat, dissolve = TRUE) -->
+<!-- df_strat <- suppressWarnings(fortify(sp_strat) %>% group_by(group)) -->
 
-r <- raster::rasterFromXYZ(xyz[, c("x", "y", "N")])
-l <- rasterToContour(r, levels = pretty(xyz$N, 20))
-df_lines <- suppressWarnings(fortify(l))
-levels <- as.numeric(as.character(l$level))
-levels <- si(levels)
-labs <- levels
-labs[seq_along(labs) %% 2 == 0] <- ""
-levels <- factor(levels, levels = levels)
-names(levels) <- rownames(as.data.frame(l))
-df_lines$level <- levels[df_lines$id]
+<!-- r <- raster::rasterFromXYZ(xyz[, c("x", "y", "N")]) -->
+<!-- l <- rasterToContour(r, levels = pretty(xyz$N, 20)) -->
+<!-- df_lines <- suppressWarnings(fortify(l)) -->
+<!-- levels <- as.numeric(as.character(l$level)) -->
+<!-- levels <- si(levels) -->
+<!-- labs <- levels -->
+<!-- labs[seq_along(labs) %% 2 == 0] <- "" -->
+<!-- levels <- factor(levels, levels = levels) -->
+<!-- names(levels) <- rownames(as.data.frame(l)) -->
+<!-- df_lines$level <- levels[df_lines$id] -->
 
-abun_plot <- ggplot() +
-  geom_path(data = df_strat, aes(x = long, y = lat, group = group), 
-            color = "grey", size = 0.05) +
-  geom_path(data = df_lines, aes(x = long, y = lat, group = group, color = level),
-            size = 0.3) +
-  scale_color_viridis(labels = labs, discrete = TRUE, name = "N") + 
-  scale_x_continuous(expand = c(0,0)) +
-  scale_y_continuous(expand = c(0,0)) +
-  coord_equal() +
-  theme_void() + guides(color = guide_legend(keywidth = 0.4, keyheight = 0, 
-                                             default.unit = "cm", ncol = 1)) +
-  theme(legend.justification = c(0, 1),
-        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
+<!-- abun_plot <- ggplot() + -->
+<!--   geom_path(data = df_strat, aes(x = long, y = lat, group = group),  -->
+<!--             color = "grey", size = 0.05) + -->
+<!--   geom_path(data = df_lines, aes(x = long, y = lat, group = group, color = level), -->
+<!--             size = 0.3) + -->
+<!--   scale_color_viridis(labels = labs, discrete = TRUE, name = "N") +  -->
+<!--   scale_x_continuous(expand = c(0,0)) + -->
+<!--   scale_y_continuous(expand = c(0,0)) + -->
+<!--   coord_equal() + -->
+<!--   theme_void() + guides(color = guide_legend(keywidth = 0.4, keyheight = 0,  -->
+<!--                                              default.unit = "cm", ncol = 1)) + -->
+<!--   theme(legend.justification = c(0, 1), -->
+<!--         plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) -->
 
-setdet <- sim$setdet
-setdet <- setdet[setdet$year == which_year & setdet$sim == which_sim, ]
-breaks <- pretty(setdet$n, 7)
+<!-- setdet <- sim$setdet -->
+<!-- setdet <- setdet[setdet$year == which_year & setdet$sim == which_sim, ] -->
+<!-- breaks <- pretty(setdet$n, 7) -->
 
-set_plot <- ggplot() +
-  geom_path(data = df_strat, aes(x = long, y = lat, group = group), 
-            color = "grey", size = 0.05) +
-  geom_point(data = setdet, aes(x = x, y = y, size = n, color = n), alpha = 0.8) +
-  scale_size_area(name = "   n", max_size = 10, limits = range(breaks), breaks = breaks) +
-  scale_color_viridis(name = "   n", limits = range(breaks), breaks = breaks) + 
-  scale_x_continuous(expand = c(0,0)) +
-  scale_y_continuous(expand = c(0,0)) +
-  coord_equal() +
-  theme_void() + guides(color = guide_legend(keywidth = 0.4, keyheight = 0.2, default.unit = "cm")) +
-  theme(legend.justification = c(0, 1),
-        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
+<!-- set_plot <- ggplot() + -->
+<!--   geom_path(data = df_strat, aes(x = long, y = lat, group = group),  -->
+<!--             color = "grey", size = 0.05) + -->
+<!--   geom_point(data = setdet, aes(x = x, y = y, size = n, color = n), alpha = 0.8) + -->
+<!--   scale_size_area(name = "   n", max_size = 10, limits = range(breaks), breaks = breaks) + -->
+<!--   scale_color_viridis(name = "   n", limits = range(breaks), breaks = breaks) +  -->
+<!--   scale_x_continuous(expand = c(0,0)) + -->
+<!--   scale_y_continuous(expand = c(0,0)) + -->
+<!--   coord_equal() + -->
+<!--   theme_void() + guides(color = guide_legend(keywidth = 0.4, keyheight = 0.2, default.unit = "cm")) + -->
+<!--   theme(legend.justification = c(0, 1), -->
+<!--         plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) -->
 
-p <- cowplot::plot_grid(abun_plot, set_plot, labels = c("a)", "b)"), align = "hv",
-                   label_fontface = "plain")
-cowplot::save_plot(plot = p, filename = "analysis/paper/figures/abun_and_sets_example.pdf",
-                   base_height = 3.8, base_width = 8.5)
-p
+<!-- p <- cowplot::plot_grid(abun_plot, set_plot, labels = c("a)", "b)"), align = "hv", -->
+<!--                    label_fontface = "plain") -->
+<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/abun_and_sets_example.pdf", -->
+<!--                    base_height = 3.8, base_width = 8.5) -->
+<!-- p -->
 
-```
+<!-- ``` -->
 
-<br>
+<!-- <br> -->
 
-```{r tot, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
-The heavy dark line indicates the true tred in the total population available to the survey."}
+<!-- ```{r tot, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
+The heavy dark line indicates the true tred in the total population available to the survey."} -->
 
-which_set_dens <- c(0.0005, 0.002, 0.01)
-which_len_caps <- c(10, 100, 1000)
+<!-- which_set_dens <- c(0.0005, 0.002, 0.01) -->
+<!-- which_len_caps <- c(10, 100, 1000) -->
 
-d <- merge(sim$surveys, sim$total_strat_error, by = "survey")
-d <- d[d$set_den %in% which_set_dens & 
-         d$lengths_cap == 100 & d$ages_cap == 10, ] ## length and age sampling not a factor here
+<!-- d <- merge(sim$surveys, sim$total_strat_error, by = "survey") -->
+<!-- d <- d[d$set_den %in% which_set_dens &  -->
+<!--          d$lengths_cap == 100 & d$ages_cap == 10, ] ## length and age sampling not a factor here -->
 
-p <- fan_plot(d, x = "Year", which_strat = "Total")
-cowplot::save_plot(plot = p, filename = "analysis/paper/figures/total_strat_true_vs_est.pdf",
-                   base_height = 3, base_width = 8)
-p
+<!-- p <- fan_plot(d, x = "Year", which_strat = "Total") -->
+<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/total_strat_true_vs_est.pdf", -->
+<!--                    base_height = 3, base_width = 8) -->
+<!-- p -->
 
-```
+<!-- ``` -->
 
-<br>
+<!-- <br> -->
 
-```{r length, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
-The heavy dark line indicates the true population available to the survey."}
+<!-- ```{r length, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+The heavy dark line indicates the true population available to the survey."} -->
 
-which_set_dens <- c(0.0005, 0.002, 0.01)
-which_len_caps <- c(10, 100, 1000)
-which_year <- 7
+<!-- which_set_dens <- c(0.0005, 0.002, 0.01) -->
+<!-- which_len_caps <- c(10, 100, 1000) -->
+<!-- which_year <- 7 -->
 
-sub_surveys <- sim$surveys[sim$survey$set_den %in% which_set_dens &
-                             sim$survey$lengths_cap %in% which_len_caps & 
-                             sim$survey$ages_cap == 10, ]
-sub_strat <- sim$length_strat_error[eval(sim$length_strat_error$survey %in% sub_surveys$survey &
-                                           sim$length_strat_error$year == which_year), ]
-d <- merge(sub_surveys, sub_strat, by = "survey")
-d$length[d$length > 100] <- 100 # aggregate beyond 100 cm
-d <- d[, list(I_hat = sum(I_hat), I = sum(I)), 
-        by = c("year", "length", "sim", "survey", "set_den", "lengths_cap")]
+<!-- sub_surveys <- sim$surveys[sim$survey$set_den %in% which_set_dens & -->
+<!--                              sim$survey$lengths_cap %in% which_len_caps &  -->
+<!--                              sim$survey$ages_cap == 10, ] -->
+<!-- sub_strat <- sim$length_strat_error[eval(sim$length_strat_error$survey %in% sub_surveys$survey & -->
+<!--                                            sim$length_strat_error$year == which_year), ] -->
+<!-- d <- merge(sub_surveys, sub_strat, by = "survey") -->
+<!-- d$length[d$length > 100] <- 100 # aggregate beyond 100 cm -->
+<!-- d <- d[, list(I_hat = sum(I_hat), I = sum(I)),  -->
+<!--         by = c("year", "length", "sim", "survey", "set_den", "lengths_cap")] -->
 
-p <- fan_plot(d, x = "Length", which_strat = "Length")
-cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_length.pdf",
-                   base_height = 6, base_width = 8)
-p
+<!-- p <- fan_plot(d, x = "Length", which_strat = "Length") -->
+<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_length.pdf", -->
+<!--                    base_height = 6, base_width = 8) -->
+<!-- p -->
 
-```
+<!-- ``` -->
 
-```{r age, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+<!-- ```{r age, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
-The heavy dark line indicates the true population available to the survey."}
+The heavy dark line indicates the true population available to the survey."} -->
 
-which_set_dens <- c(0.0005, 0.002, 0.01)
-which_len_caps <- c(10, 100, 1000)
-which_year <- 7
+<!-- which_set_dens <- c(0.0005, 0.002, 0.01) -->
+<!-- which_len_caps <- c(10, 100, 1000) -->
+<!-- which_year <- 7 -->
 
-sub_surveys <- sim$surveys[sim$survey$set_den %in% which_set_dens &
-                             sim$survey$lengths_cap %in% which_len_caps & 
-                             sim$survey$ages_cap == 10, ]
-sub_strat <- sim$age_strat_error[eval(sim$age_strat_error$survey %in% sub_surveys$survey &
-                                        sim$age_strat_error$year == which_year), ]
-d <- merge(sub_surveys, sub_strat, by = "survey")
-d <- d[d$year == which_year & d$set_den %in% which_set_dens &
-         d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
-d$age[d$age > 10] <- 10 # aggregate beyond 10
-d <- d[, list(I_hat = sum(I_hat), I = sum(I)), 
-       by = c("year", "age", "sim", "survey", "set_den", "lengths_cap")]
+<!-- sub_surveys <- sim$surveys[sim$survey$set_den %in% which_set_dens & -->
+<!--                              sim$survey$lengths_cap %in% which_len_caps &  -->
+<!--                              sim$survey$ages_cap == 10, ] -->
+<!-- sub_strat <- sim$age_strat_error[eval(sim$age_strat_error$survey %in% sub_surveys$survey & -->
+<!--                                         sim$age_strat_error$year == which_year), ] -->
+<!-- d <- merge(sub_surveys, sub_strat, by = "survey") -->
+<!-- d <- d[d$year == which_year & d$set_den %in% which_set_dens & -->
+<!--          d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ] -->
+<!-- d$age[d$age > 10] <- 10 # aggregate beyond 10 -->
+<!-- d <- d[, list(I_hat = sum(I_hat), I = sum(I)),  -->
+<!--        by = c("year", "age", "sim", "survey", "set_den", "lengths_cap")] -->
 
-p <- fan_plot(d, x = "Age")
-cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_age.pdf",
-                   base_height = 6, base_width = 8)
-p
+<!-- p <- fan_plot(d, x = "Age") -->
+<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_age.pdf", -->
+<!--                    base_height = 6, base_width = 8) -->
+<!-- p -->
 
-```
+<!-- ``` -->
 
-<br>
+<!-- <br> -->
 
-```{r year, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+<!-- ```{r year, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
-The heavy dark line indicates the true trend in the population available to the survey."}
+The heavy dark line indicates the true trend in the population available to the survey."} -->
 
-which_set_dens <- c(0.0005, 0.002, 0.01)
-which_len_caps <- c(10, 100, 1000)
-which_age <- 4
+<!-- which_set_dens <- c(0.0005, 0.002, 0.01) -->
+<!-- which_len_caps <- c(10, 100, 1000) -->
+<!-- which_age <- 4 -->
 
-d <- merge(sim$surveys, sim$age_strat_error, by = "survey")
-d <- d[d$age == which_age & d$set_den %in% which_set_dens &
-       d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ]
+<!-- d <- merge(sim$surveys, sim$age_strat_error, by = "survey") -->
+<!-- d <- d[d$age == which_age & d$set_den %in% which_set_dens & -->
+<!--        d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ] -->
 
-p <- fan_plot(d, x = "Year")
-cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_year.pdf",
-                   base_height = 6, base_width = 8)
-p
+<!-- p <- fan_plot(d, x = "Year") -->
+<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_year.pdf", -->
+<!--                    base_height = 6, base_width = 8) -->
+<!-- p -->
 
-```
+<!-- ``` -->
 
-<br>
+<!-- <br> -->
 
-```{r error-lines, cache = TRUE, fig.height = 3.8, fig.width = 8.5, results = "hide", fig.cap = "- Root-mean-squared error (RMSE) of abundance at length and abundance at age estimates as a function of the number of fish measured ($N_{lengths}$) from an array of surveys with different sampling protocol.
+<!-- ```{r error-lines, cache = TRUE, fig.height = 3.8, fig.width = 8.5, results = "hide", fig.cap = "- Root-mean-squared error (RMSE) of abundance at length and abundance at age estimates as a function of the number of fish measured ($N_{lengths}$) from an array of surveys with different sampling protocol.
 The labels represent surveys with different set densities ($D_{sets}$), where number of sets are shown in brackets, and the lines are color coded by length sampling protocol ($M_{lengths}$).
-Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios."}
+Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios."} -->
 
-totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught),
-                                 n_measured = mean(n_measured), n_aged = mean(n_aged)),
-                          by = "survey"]
+<!-- totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught), -->
+<!--                                  n_measured = mean(n_measured), n_aged = mean(n_aged)), -->
+<!--                           by = "survey"] -->
 
-errors <- merge(sim$surveys, sim$length_strat_error_stats, by = "survey")
-length_d <- merge(errors, totals, by = "survey")
-length_d <- length_d[length_d$ages_cap == 10, ]
-length_d$strat <- "Abundance at length"
+<!-- errors <- merge(sim$surveys, sim$length_strat_error_stats, by = "survey") -->
+<!-- length_d <- merge(errors, totals, by = "survey") -->
+<!-- length_d <- length_d[length_d$ages_cap == 10, ] -->
+<!-- length_d$strat <- "Abundance at length" -->
 
-errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
-age_d <- merge(errors, totals, by = "survey")
-age_d <- age_d[age_d$ages_cap == 10, ]
-age_d$strat <- "Abundance at age"
+<!-- errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey") -->
+<!-- age_d <- merge(errors, totals, by = "survey") -->
+<!-- age_d <- age_d[age_d$ages_cap == 10, ] -->
+<!-- age_d$strat <- "Abundance at age" -->
 
-d <- rbind(length_d, age_d)
-d$cap <- factor(d$lengths_cap)
-d$strat <- factor(d$strat, levels = c("Abundance at length", "Abundance at age"))
+<!-- d <- rbind(length_d, age_d) -->
+<!-- d$cap <- factor(d$lengths_cap) -->
+<!-- d$strat <- factor(d$strat, levels = c("Abundance at length", "Abundance at age")) -->
 
-lab_d <- d[lengths_cap == max(lengths_cap),
-           list(lengths_cap, n_measured, n_sets, cap, RMSE),
-           by = c("set_den", "strat")]
-lab_d$lab <- paste0(formatC(lab_d$set_den), " (", lab_d$n_sets, ")")
+<!-- lab_d <- d[lengths_cap == max(lengths_cap), -->
+<!--            list(lengths_cap, n_measured, n_sets, cap, RMSE), -->
+<!--            by = c("set_den", "strat")] -->
+<!-- lab_d$lab <- paste0(formatC(lab_d$set_den), " (", lab_d$n_sets, ")") -->
 
-p <- ggplot(data = d, aes(x = n_measured, y = RMSE, group = set_den)) +
-  geom_line(aes(color = cap), size = 2, lineend = "round") +
-  geom_text(data = lab_d, aes(label = lab), nudge_x = 500, hjust = 0) +
-  geom_point(data = lab_d, aes(color = cap)) +
-  xlab(bquote(N[lengths])) +
-  scale_y_continuous(labels = format_si()) +
-  scale_x_continuous(labels = format_si(), limits = c(0, max(d$n_measured) * 1.3),
-                     breaks = pretty(d$n_measured, 6)) +
-  scale_color_viridis(name = bquote(M[lengths]), discrete = TRUE) +
-  facet_wrap(~ strat, scales = "free", dir = "h") +
-  theme_bw() +
-  theme(legend.justification = c(0, 1),
-        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        strip.background = element_blank(),
-        panel.border = element_rect(fill = NA, color = "darkgrey", size = 0.4),
-        axis.ticks = element_line(color = "darkgrey", size = 0.4))
+<!-- p <- ggplot(data = d, aes(x = n_measured, y = RMSE, group = set_den)) + -->
+<!--   geom_line(aes(color = cap), size = 2, lineend = "round") + -->
+<!--   geom_text(data = lab_d, aes(label = lab), nudge_x = 500, hjust = 0) + -->
+<!--   geom_point(data = lab_d, aes(color = cap)) + -->
+<!--   xlab(bquote(N[lengths])) + -->
+<!--   scale_y_continuous(labels = format_si()) + -->
+<!--   scale_x_continuous(labels = format_si(), limits = c(0, max(d$n_measured) * 1.3), -->
+<!--                      breaks = pretty(d$n_measured, 6)) + -->
+<!--   scale_color_viridis(name = bquote(M[lengths]), discrete = TRUE) + -->
+<!--   facet_wrap(~ strat, scales = "free", dir = "h") + -->
+<!--   theme_bw() + -->
+<!--   theme(legend.justification = c(0, 1), -->
+<!--         plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"), -->
+<!--         panel.grid.major = element_blank(), -->
+<!--         panel.grid.minor = element_blank(), -->
+<!--         strip.background = element_blank(), -->
+<!--         panel.border = element_rect(fill = NA, color = "darkgrey", size = 0.4), -->
+<!--         axis.ticks = element_line(color = "darkgrey", size = 0.4)) -->
 
-cowplot::save_plot(plot = p, filename = "analysis/paper/figures/error_lines.pdf",
-                   base_height = 3.8, base_width = 8.5)
-p
+<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/error_lines.pdf", -->
+<!--                    base_height = 3.8, base_width = 8.5) -->
+<!-- p -->
 
-```
+<!-- ``` -->
 
-<br>
+<!-- <br> -->
 
-```{r error-wireframe, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
-Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$)."}
+<!-- ```{r error-wireframe, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
+Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$)."} -->
 
-trellis_theme <- list(
-  box.3d = list(col = "darkgrey"), 
-  axis.line = list(col = "transparent"),
-  strip.background = list(col = "transparent"), 
-  strip.border = list(col = "transparent"),
-   layout.heights = list(
-        top.padding = 0,
-        main.key.padding = 0,
-        key.axis.padding = 0,
-        axis.xlab.padding = 0,
-        xlab.key.padding = 0,
-        key.sub.padding = 0,
-        bottom.padding = -5
-    ),
-    layout.widths = list(
-        left.padding = 0,
-        key.ylab.padding = 0,
-        ylab.axis.padding = 0,
-        axis.key.padding = 0,
-        right.padding = 1
-    )
-)
+<!-- trellis_theme <- list( -->
+<!--   box.3d = list(col = "darkgrey"),  -->
+<!--   axis.line = list(col = "transparent"), -->
+<!--   strip.background = list(col = "transparent"),  -->
+<!--   strip.border = list(col = "transparent"), -->
+<!--    layout.heights = list( -->
+<!--         top.padding = 0, -->
+<!--         main.key.padding = 0, -->
+<!--         key.axis.padding = 0, -->
+<!--         axis.xlab.padding = 0, -->
+<!--         xlab.key.padding = 0, -->
+<!--         key.sub.padding = 0, -->
+<!--         bottom.padding = -5 -->
+<!--     ), -->
+<!--     layout.widths = list( -->
+<!--         left.padding = 0, -->
+<!--         key.ylab.padding = 0, -->
+<!--         ylab.axis.padding = 0, -->
+<!--         axis.key.padding = 0, -->
+<!--         right.padding = 1 -->
+<!--     ) -->
+<!-- ) -->
 
-zscale_format <- function(lim, logsc = FALSE, ...) {
-    ans <- zscale.components.default(lim, logsc = logsc, ...)
-    ans$left$labels$labels <- si(ans$left$labels$at)
-    ans
-}
+<!-- zscale_format <- function(lim, logsc = FALSE, ...) { -->
+<!--     ans <- zscale.components.default(lim, logsc = logsc, ...) -->
+<!--     ans$left$labels$labels <- si(ans$left$labels$at) -->
+<!--     ans -->
+<!-- } -->
 
-d <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
-# d <- d[d$set_den != 0.0005, ]
-d <- d[d$set_den %in% c(0.0005, 0.002, 0.01)]
-set_den <- sort(unique(d$set_den))
-d$set_den <- factor(d$set_den)
-options(scipen = 999)
-labs <- sapply(seq_along(set_den), function(i) as.expression(bquote(D[sets] == .(set_den[i]))), 
-               USE.NAMES = FALSE)
-zlabs <- list(at = pretty(d$RMSE), labels = si(pretty(d$RMSE)))
+<!-- d <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey") -->
+<!-- # d <- d[d$set_den != 0.0005, ] -->
+<!-- d <- d[d$set_den %in% c(0.0005, 0.002, 0.01)] -->
+<!-- set_den <- sort(unique(d$set_den)) -->
+<!-- d$set_den <- factor(d$set_den) -->
+<!-- options(scipen = 999) -->
+<!-- labs <- sapply(seq_along(set_den), function(i) as.expression(bquote(D[sets] == .(set_den[i]))),  -->
+<!--                USE.NAMES = FALSE) -->
+<!-- zlabs <- list(at = pretty(d$RMSE), labels = si(pretty(d$RMSE))) -->
 
-pdf("analysis/paper/figures/error_wireframe.pdf", height = 3, width = 8.5)
+<!-- pdf("analysis/paper/figures/error_wireframe.pdf", height = 3, width = 8.5) -->
 
-p <- wireframe(RMSE ~  lengths_cap * ages_cap | set_den, data = d,
-          as.table = TRUE, drape = TRUE, col = "white", lwd = 0.5,
-          strip = strip.custom(factor.levels = labs),
-          scales = list(arrows = FALSE, cex = 0.7, z = zlabs),
-          col.regions = viridis::viridis(100),
-          screen = list(z = 230, x = 300), 
-          xlab = list(expression(M[lengths]), rot = 25), 
-          ylab = list(expression(M[ages]), rot = -15), 
-          zlab = list("RMSE", rot = 95),
-          colorkey = list(labels = zlabs, height = 0.6,
-                          axis.line = list(col = "darkgrey")),
-          par.settings = trellis_theme)
-p <- update(p, layout = c(3, 1))
-p
-grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc"))
+<!-- p <- wireframe(RMSE ~  lengths_cap * ages_cap | set_den, data = d, -->
+<!--           as.table = TRUE, drape = TRUE, col = "white", lwd = 0.5, -->
+<!--           strip = strip.custom(factor.levels = labs), -->
+<!--           scales = list(arrows = FALSE, cex = 0.7, z = zlabs), -->
+<!--           col.regions = viridis::viridis(100), -->
+<!--           screen = list(z = 230, x = 300),  -->
+<!--           xlab = list(expression(M[lengths]), rot = 25),  -->
+<!--           ylab = list(expression(M[ages]), rot = -15),  -->
+<!--           zlab = list("RMSE", rot = 95), -->
+<!--           colorkey = list(labels = zlabs, height = 0.6, -->
+<!--                           axis.line = list(col = "darkgrey")), -->
+<!--           par.settings = trellis_theme) -->
+<!-- p <- update(p, layout = c(3, 1)) -->
+<!-- p -->
+<!-- grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc")) -->
 
-invisible(dev.off())
+<!-- invisible(dev.off()) -->
 
-p
-grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc"))
+<!-- p -->
+<!-- grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc")) -->
 
-```
+<!-- ``` -->
 
-<br>
+<!-- <br> -->
 

### 2018-11-15 analysis/paper/helper_functions.R

@@ -114,3 +114,9 @@ fan_plot <- function(d, x = "Age", which_strat = "Age") {
 }
 
 
+## wrapper for orca that puts file somewhere other than the working dir
+export_plot <- function(p, file = NULL, width = 700, height = 500, scale = 4, ...) {
+  ext <- tools::file_ext(file)
+  orca(p, paste0("tmp.", ext), width = width, height = height, scale = scale, ...)
+  invisible(file.rename(paste0("tmp.", ext), file))
+}

### 2018-11-15 analysis/paper/sim_survey_paper.Rmd
Add plot_grid function

@@
 
-The `make_grid` function is used to generate the grid used in the case study, and the `sim_distribution` function distributes the population, `n`, simulated above using `sim_abundance` through the grid:
+The `make_grid` function is used to generate a survey grid for use in the `sim_distribution` function.
+The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`.
+If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied to `sim_distribution` if the same structure is used.
+The settings used to make the grid used in our case study are shown in Table \@ref(tab:par) and these settings are also the defaults.
+Below we generate and plot (combined in Figure \@ref(fig:grid)) a default grid and another grid with a more linear gradient in depth.
+
+```{r, echo = TRUE, eval = FALSE}
+a <- make_grid()
+b <- make_grid(shelf_depth = 500, shelf_width = 0, depth_range = c(0, 1000))
+plot_grid(a)
+plot_grid(b)
+```
+
+```{r grid, fig.cap = "- Plots produced by `plot_grid` when supplied an raster object produced by a) `make_grid()` and b) `make_grid(shelf_depth = 500, shelf_width = 0, depth_range = c(0, 1000))`."}
+
+a <- make_grid() %>% 
+  plot_grid(showscale = FALSE) %>% 
+  add_annotations(x = 0, y = 1, xref = "paper", yref = "paper", showarrow = FALSE,
+                  text = "a)")
+b <- make_grid(shelf_depth = 500, shelf_width = 0, depth_range = c(0, 1000)) %>% 
+  plot_grid(showscale = FALSE) %>% 
+  add_annotations(x = 0.01, y = 1, xref = "paper", yref = "paper", showarrow = FALSE,
+                  text = "b)")
+p <- subplot(a, b)
+
+export_plot(p, file = "analysis/paper/figures/plot_grid.png", width = 800, height = 460)
+knitr::include_graphics("figures/plot_grid.png")
+
+```
+
+With a population simulated
+
+, and the `sim_distribution` function distributes the population simulated above using `sim_abundance` through the grid:
 

### 2018-11-15 R/plot_sim.R

@@ -8,6 +8,7 @@
 #' @param sim            Object returned by \code{\link{sim_abundance}},
 #'                       \code{\link{sim_distribution}}, etc.
 #' @param mat            Name of matrix in \code{sim} list to plot.
+#' @param grid           Grid produced by \code{\link{make_grid}}.
 #' @param xlab,ylab,zlab Axes labels.
 #' @param sum_ages       Sum across these ages
 #' @param ages           Subset data to one or more ages.
@@ -53,6 +54,39 @@ plot_surface <- function(sim, mat = "N", xlab = "Age", ylab = "Year", zlab = mat
       ))
 }
 
+
+#' @export
+#' @rdname plot_trend
+plot_grid <- function(grid, ...) {
+
+  xax <- list(
+    title = "",
+    zeroline = FALSE,
+    showline = FALSE,
+    showticklabels = FALSE,
+    showgrid = FALSE,
+    tickcolor = "transparent"
+  )
+  yax <- c(scaleanchor = "x", xax)
+
+  sp_div <- raster::rasterToPolygons(grid$division, dissolve = TRUE)
+  df_div <- suppressMessages(ggplot2::fortify(sp_div) %>% group_by(group))
+  sp_strat <- raster::rasterToPolygons(grid$strat, dissolve = TRUE)
+  df_strat <- suppressMessages(ggplot2::fortify(sp_strat) %>% group_by(group))
+  xyz <- data.frame(rasterToPoints(grid))
+
+  plot_ly(...) %>%
+    add_trace(data = xyz, x = ~x, y = ~y, z = ~depth,
+              text = ~paste("x:", x, "<br>y:", y, "<br>depth:", depth, "<br>cell:", cell,
+                            "<br>division:", division, "<br>strat:", strat),
+              hoverinfo = "text", type = "heatmap") %>%
+    add_paths(data = df_strat, x = ~long, y = ~lat, color = I("white"),
+              hoverinfo = "none", size = I(0.5), showlegend = FALSE) %>%
+    add_paths(data = df_div, x = ~long, y = ~lat, color = I("darkgrey"),
+              hoverinfo = "none", size = I(2), showlegend = FALSE) %>%
+    layout(xaxis = xax, yaxis = yax)
+
+}
 
 #' @export
 #' @rdname plot_trend
@@ -135,7 +169,7 @@ plot_survey <- function(sim, which_year = 1, which_sim = 1,
[…]

### 2018-11-16 analysis/paper/sim_survey_paper.Rmd
Draft demo of sim_distribution and plot_distribution

--- 2018-11-15 analysis/paper/sim_survey_paper.Rmd
@@
 
-The package also includes several ploting function for making quick plotly-based [@sievert2018] interactive visuals of the simulated population.
+The package also includes several plotting function for making quick plotly-based [@sievert2018] interactive visuals of the simulated population.
 The plots output from the below code are combined into Figure \@ref(fig:surface).
@@
 
-```{r surface, fig.cap = "- Plots produced by `plot_surface` when supplied an list produced by a) `sim_abundance(ages = 1:20, R = sim_R(mean = 3e+07), Z = sim_Z(mean = 0.5))` and b) `sim_abundance(ages = 1:6, R = sim_R(mean = 1e+10), Z = sim_Z(mean = 0.8))`.
+```{r surface, fig.cap = "- Plots produced by `plot_surface` when supplied an list produced by `sim_abundance` with settings that simulate a relatively a) long and b) short lived species.
 The abundance at age matrix (`N`) is displayed by default; the `mat` argument can be used to specify another matrix contained within the supplied list, such as `Z`."}
@@
 
-The `make_grid` function is used to generate the grid used in the case study, and the `sim_distribution` function distributes the population, `n`, simulated above using `sim_abundance` through the grid:
+The `make_grid` function is used to generate a survey grid for use in the `sim_distribution` function.
+The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`.
+If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied to `sim_distribution` if the same structure is used.
+The settings used to make the grid used in our case study are shown in Table \@ref(tab:par) and these settings are also the defaults.
+Below we generate and plot (combined in Figure \@ref(fig:grid)) a default grid and another grid with four divisions and a more linear gradient in depth.
 
 ```{r, echo = TRUE, eval = FALSE}
-g <- make_grid(x_range = c(-140, 140),
-               y_range = c(-140, 140),
-               res = c(3.5, 3.5),
-               shelf_depth = 200,
-               shelf_width = 100,
-               depth_range = c(0, 1000),
-               n_div = 1,
-               strat_breaks = seq(0, 1000, by = 40),
-               strat_splits = 2)
-
-d <- sim_distribution(n, grid = g,
-                      ays_covar = sim_ays_covar(sd = 2.8,
-                                                range = 300,
-                                                phi_age = 0.5,
-                                                phi_year = 0.9,
-                                                group_ages = 5:20),
-                      depth_par = sim_parabola(mu = 200,
-                                               sigma = 70))
+a <- make_grid(n_div = 1, shelf_depth = 200,
+               shelf_width = 100, depth_range = c(0, 1000))
+b <- make_grid(n_div = 4, shelf_depth = 500, 
+               shelf_width = 0, depth_range = c(0, 1000))
+plot_grid(a)
+plot_grid(b)
 ```
 
-This function retains all the data simulated by `sim_abundance` and adds a `data.table` with abundance split by age, year and cell (named `sp_N`).
-Again, note the use of the closures `sim_ays_covar` and `sim_parabola` to impose the age-year-space covariance and the relationship with depth, respectively.
+```{r grid, fig.cap = "- Plots produced by `plot_grid` when supplied an raster object produced by `make_grid` a) using default settings and b) settings that generate a more linear depth gradient.
+In these plots, the color gradient represents depth, the grey lines deliniate divisions and white lines deliniate strata."}
+
+a <- make_grid() %>% 
+  plot_grid(showscale = TRUE) %>% 
+  add_annotations(x = 0, y = 1, xref = "paper", yref = "paper", showarrow = FALSE,
+                  text = "a)")
+b <- make_grid(n_div = 4, shelf_depth = 500, shelf_width = 0, depth_range = c(0, 1000)) %>% 
+  plot_grid(showscale = FALSE) %>% 
+  add_annotations(x = 0.01, y = 1, xref = "paper", yref = "paper", showarrow = FALSE,
+                  text = "b)")
+p <- subplot(a, b)
+
+export_plot(p, file = "analysis/paper/figures/plot_grid.png", width = 800, height = 430)
+knitr::include_graphics("figures/plot_grid.png")
+
+```
+
+The `sim_distribution` function distributes the population simulated using `sim_abundance` throughout the grid generated using `make_grid`.
+Table \@ref(tab:par) outlines the default settings used for the function and the case study.
+In addition to supplying objects produced by `sim_abundance` and `make_grid`, this function requires two closures that describe the age-year-space covariance and the relationship with depth.
+Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by tweaking a few parameters in these closures.
+Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e.
+wider range) and exhibits stronger correlation across ages than years.
+Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years`, respectively, arguments in the `sim_ays_covar` closure.
+In other words, these parameters can be modified to control the degree of age-specific clustering and inter-annual site-fedelity exhibited by the simulated population.
+Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete.
+Also note that the key functions in the SimSurvey package have been set-up to be pipe [`%>%`; @bache2014] friendly.
+
+```{r, echo = TRUE, eval = TRUE, cache = TRUE}
+
+set.seed(438)
+a <- sim_abundance() %>% 
+  sim_distribution(ays_covar = sim_ays_covar(range = 300,
+                                             phi_year = 0.9,
+                                             phi_age = 0.5))
+b <- sim_abundance() %>% 
+  sim_distribution(ays_covar = sim_ays_covar(range = 2000,
+                                             phi_year = 0.2,
+                                             phi_age = 0.9))
+
+```
+
+This function retains all the data simulated by `sim_abundance` and adds a data.table [@dowle2017], named `sp_N`, with abundance (`N`) split by `age`, `year` and `cell`.
+The function also retains the grid object and converts these data into a data.table, named `grid_xy`, with headers `x`, `y`, `depth`, `cell`, `division` and `strat`.
+The `sp_N` object can be merge with the `grid_xy` data by `cell` to associate abundance with specific locations, depth, divisions or strata.
+The `plot_distribution` function can be used to provide a quick visual of the distribution across ages and years.
+The code below will generate interactive plots with an Age-Year slider, however, for this paper we present a facet plot of the simulated data in Figure \@ref(fig:distribution).
+
+```{r, echo = TRUE, eval = FALSE}
+plot_distribution(a, ages = 1:3, years = 1:3, type = "heatmap")
+plot_distribution(b, ages = 1:3, years = 1:3, type = "heatmap")
+```
+
+```{r distribution, fig.cap = "- Distribution plots of a population simulated using `sim_distribution` that forms a) tight clusters with stronger correlation through years than ages, and b) relatively diffuse clusters with stronger correlation through ages than years."}
+
+facet_distribution <- function(data) {
+  ay <- expand.grid(age = 1:3, year = 1:3)
+  plot_list <- vector("list", nrow(ay))
+  for (i in 1:3) {
+    for (j in 1:3) {
+      p <- plot_distribution(data, ages = i, years = j, type = "heatmap") %>% 
+        hide_colorbar() %>% 
+        layout(autosize = FALSE, width = 100, height = 100,
+               margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0))
+      plot_list[[which(ay$age == i & ay$year == j)]] <- p
+    }
+  }
+  subplot(plot_list, nrows = 3, shareX = TRUE, shareY = TRUE, margin = 0.01) %>% 
+    layout(autosize = FALSE, width = 300, height = 300,
+           margin = list(t = 20, l = 20, b = 10, r = 10))
+}
+
+a_plots <- facet_distribution(a)
+b_plots <- facet_distribution(b)
+
+p <- subplot(a_plots, b_plots) %>% 
+  add_labels(text = rep(c("Age 1", "Age 2", "Age 3"), 2), x = c(0.04, 0.20, 0.4, 0.6, 0.8, 0.95), y = rep(0.995, 6), yshift = 20) %>% 
+  add_labels(text = rep(c("Year 1", "Year 2", "Year 3"), 2), x = c(rep(0.005, 3), rep(0.54, 3)), y = c(0.08, 0.5, 0.92), xshift = -20, textangle = -90) %>% 
+  add_labels(text = "a)", x = 0, y = 1, yshift = 20, xshift = -20) %>% 
+  add_labels(text = "b)", x = 0.54, y = 1, yshift = 20, xshift = -20)
+
+export_plot(p, file = "analysis/paper/figures/plot_distribution.png", width = 600, height = 300)
+knitr::include_graphics("figures/plot_distribution.png")
+
+```
 
@@
 | `+                     K = 0.1,`                        |                                                Growth rate |                        $K$ |
-| `+                     log_sd = 0.1,`                   |      Standard devation of the von Bertalanffy growth curve |                 $\sigma_l$ |
+| `+                     log_sd = 0.1,`                   |      Standard deviation of the von Bertalanffy growth curve |                 $\sigma_l$ |
 | `+                     length_group = 3))`              |                  Length group for abundance at length (cm) |            $l_{group}^{N}$ |      

### 2018-11-16 R/plot_sim.R

@@ -65,7 +65,7 @@ plot_grid <- function(grid, ...) {
     showline = FALSE,
     showticklabels = FALSE,
     showgrid = FALSE,
-    tickcolor = "transparent"
+    ticks = ""
   )
   yax <- c(scaleanchor = "x", xax)
 
@@ -99,7 +99,7 @@ plot_distribution <- function(sim, ages = 1:10, years = 1:10,
     showline = FALSE,
     showticklabels = FALSE,
     showgrid = FALSE,
-    tickcolor = "transparent"
+    ticks = ""
   )
   yax <- c(scaleanchor = "x", xax)
 
@@ -108,7 +108,7 @@ plot_distribution <- function(sim, ages = 1:10, years = 1:10,
   d$ay <- paste(d$age, d$year, sep = "-")
   split_d <- split(d, d$ay)
   split_d <- lapply(split_d, function(.) {
-    stats::xtabs(N ~ x + y, data = ., subset = NULL)
+    stats::xtabs(N ~ y + x, data = ., subset = NULL)
   })
   x <- sort(unique(d$x))
   y <- sort(unique(d$y))
@@ -164,7 +164,7 @@ plot_survey <- function(sim, which_year = 1, which_sim = 1,
     showline = FALSE,
     showticklabels = FALSE,
     showgrid = FALSE,
-    tickcolor = "transparent"
+    ticks = ""
   )
   yax <- c(scaleanchor = "x", xax)
 

### 2018-11-16 analysis/paper/helper_functions.R

@@ -120,3 +120,10 @@ export_plot <- function(p, file = NULL, width = 700, height = 500, scale = 4, ..
   orca(p, paste0("tmp.", ext), width = width, height = height, scale = scale, ...)
   invisible(file.rename(paste0("tmp.", ext), file))
 }
+
+## wrapper of add_annotations with defaults set-up for plot labeling
+add_labels <- function(p, text = NULL, x = 0, y = 1, xref = "paper", yref = "paper",
+                      showarrow = FALSE, ..., data = NULL) {
+  add_annotations(p, text = text, x = x, y = y, xref = xref, yref = yref,
+                  showarrow = showarrow, ..., data = data)
+}

### 2018-11-18 analysis/paper/sim_survey_paper.Rmd
Improve the plot_survey function (include true, caught and measured length and age distributions)

@@
 
+Table: (\#tab:par) - Main function calls, with argument descriptions and associated parameter symbols, used to simulate the cod-like population used in the case study presented here.
+
+|                                       **Function call** |                                            **Description** |                 **Symbol** |
+|:------------------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
+| `> set.seed(438)`                                       |      Set random number generator seed to replicate results |                            |
+| `>`                                                     |                                                            |                            |
+| `> n <- sim_abundance(`                                 |                                       *Simulate abundance* |                            |         
+| `+   ages = 1:20,`                                      |                                                       Ages |                        $a$ |
+| `+   years = 1:20,`                                     |                                                      Years |                        $y$ |
+| `+   R = sim_R(mean = 30000000,`                        |                                           Mean recruitment |                    $\mu_r$ |
+| `+             log_sd = 0.5),`                          |                      Standard deviation of log-recruitment |                 $\sigma_r$ |
+| `+   Z = sim_Z(mean = 0.5,`                             |                                       Mean total mortality |                    $\mu_Z$ |
+| `+             log_sd = 0.2,`                           |                      Standard deviation of total mortality |                 $\sigma_Z$ |
+| `+             phi_age = 0.9,`                          |    Correlation across ages in error around total mortality |    $\varphi_{\delta, age}$ |
+| `+             phi_year = 0.5),`                        |   Correlation across years in error around total mortality |   $\varphi_{\delta, year}$ |
+| `+   growth = sim_vonB(Linf = 120,`                     |                                Mean asymptotic length (cm) |               $l_{\infty}$ |
+| `+                     L0 = 5,`                         |                                       Length at birth (cm) |                      $l_0$ |
+| `+                     K = 0.1,`                        |                                                Growth rate |                        $K$ |
+| `+                     log_sd = 0.1,`                   |     Standard deviation of the von Bertalanffy growth curve |                 $\sigma_l$ |
+| `+                     length_group = 3))`              |                  Length group for abundance at length (cm) |            $l_{group}^{N}$ |      
+| `>`                                                     |                                                            |                            |
+| `> g <- make_grid(`                                     |                                              *Make a grid* |                            |
+| `+   x_range = c(-140, 140),`                           |                           Range of grid in the x dimension |                            |                   
+| `+   y_range = c(-140, 140),`                           |                           Range of grid in the y dimension |                            | 
+| `+   res = c(3.5, 3.5),`                                |                                       Grid resolution (km) |                        $R$ |
+| `+   shelf_depth = 200,`                                |                                           Shelf depth (km) |                $d_{shelf}$ |
+| `+   shelf_width = 100,`                                |                                           Shelf width (km) |                $w_{shelf}$ |
+| `+   depth_range = c(0, 1000),`                         |                                            Depth range (m) |       $[d_{min}, d_{max}]$ |
+| `+   n_div = 1,`                                        |                                        Number of divisions |                  $N_{div}$ |
+| `+   strat_breaks = seq(0, 1000, by = 40))`             |                 Series of depth breaks for defining strata |                            |
+| `>`                                                     |                                                            |                            |
+| `> d <- sim_distribution(`                              |                                    *Simulate distribution* |                            |     
+| `+   sim = n,`                                          |                Object produced by `sim_abundance` function |                            |     
+| `+   grid = g,`                                         |                    Object produced by `make_grid` function |                            |
+| `+   ays_covar = sim_ays_covar(sd = 2.8,`               |          Standard deviation of age-year-space distribution |             $\sigma_{\xi}$ |
+| `+                             range = 300,`            |                          Range of spatial correlation (km) |                        $r$ |
+| `+                             phi_age = 0.5,`          |            Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |
+| `+                             phi_year = 0.9,`         |           Correlation across years in spatial distribution |      $\varphi_{\xi, year}$ |
+| `+                             group_ages = 5:20),`     |       Make space-age-year variance equal across these ages |                            |
+| `+   depth_par = sim_parabola(mu = 200,`                |          Depth at which abundance is typically highest (m) |                    $\mu_d$ |
+| `+                            sigma = 70))`             |              Dispersion around depth of peak abundance (m) |                 $\sigma_d$ |
+| `>`                                                     |                                                            |                            |
+| `> e <- expand_surveys(`                                |                       *Define a series of surveys to test* |                            |
+| `+   set_den = c(0.0005, 0.001, 0.002, 0.005, 0.01),`   |                                       Set density (km^-2^) |                 $D_{sets}$ |
+| `+   lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),`   |                 Maximum number of lengths to collect / set |              $M_{lengths}$ |
+| `+   ages_cap = c(2, 5, 10, 20, 50))`                   | Maximum number of ages to sample / length group / division |                 $M_{ages}$ |
+| `>`                                                     |                                                            |                            |
+| `> s <- test_surveys(`                                  |                             *Simulate and analyze surveys* |                            |
+| `+   sim = d,`                                          |             Object produced by `sim_distribution` function |                            |
+| `+   surveys = e,`                                      |               Object produced by `expand_surveys` function |                            |
+| `+   n_loops = 1000,`                                   |                    Number of times to simulate each survey |                            | 
+| `+   q = sim_logistic(k = 2,`                           |                Steepness of logistic curve of catchability |                        $k$ |
+| `+                   x0 = 3),`                          |                 Midpoint of logistic curve of catchability |                      $x_0$ |
+| `+   trawl_dim = c(1.5, 0.02),`                         |       Trawl dimensions (km) - i.e. area covered by a trawl |                $A_{trawl}$ |
+| `+   length_group = 1)`                                 |                         Length group for age sampling (cm) |                $l_{group}$ |
+
+*Caution:* as written, the `test_surveys` function call may take days or weeks to run.
+Run duration can be decreased by running multiple simulations per loop (`n_sims` argument) and by running the loops in parallel (`cores` argument).
+See help file for more details (`?test_surveys`).
+
+<br>
+
 ## Simulate abundance
@@
 pa <- plot_surface(a, mat = "N", scene = "scene1", showscale = FALSE) %>% 
-  add_annotations(x = 0.05, y = 0.9, xref = "paper", yref = "paper", showarrow = FALSE,
-                  text = "a)")
+  add_labels(x = 0.05, y = 0.9, text = "a)")
 pb <- plot_surface(b, mat = "N", scene = "scene2", showscale = FALSE) %>% 
-  add_annotations(x = 0.05, y = 0.9, xref = "paper", yref = "paper", showarrow = FALSE,
-                  text = "b)")
+  add_labels(x = 0.05, y = 0.9, text = "b)")
 
@@
 knitr::include_graphics("figures/plot_surface.png")
+
 ```
@@
 ```{r, echo = TRUE, eval = FALSE}
+
 a <- make_grid(n_div = 1, shelf_depth = 200,
@@
 plot_grid(b)
+
 ```
@@
   plot_grid(showscale = TRUE) %>% 
-  add_annotations(x = 0, y = 1, xref = "paper", yref = "paper", showarrow = FALSE,
-                  text = "a)")
+  add_labels(x = 0, y = 1, text = "a)")
 b <- make_grid(n_div = 4, shelf_depth = 500, shelf_width = 0, depth_range = c(0, 1000)) %>% 
   plot_grid(showscale = FALSE) %>% 
-  add_annotations(x = 0.01, y = 1, xref = "paper", yref = "paper", showarrow = FALSE,
-                  text = "b)")
+  add_labels(x = 0.01, y = 1, text = "b)")
 p <- subplot(a, b)
@@
 Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years`, respectively, arguments in the `sim_ays_covar` closure.
-In other words, these parameters can be modified to control the degree of age-specific clustering and inter-annual site-fedelity exhibited by the simulated population.
+In other words, these parameters can be modified to control the degree of age-specific clustering and inter-annual site-fidelity exhibited by the simulated population.
 Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete.
@@
 ```{r, echo = TRUE, eval = FALSE}
+
 plot_distribution(a, ages = 1:3, years = 1:3, type = "heatmap")
 plot_distribution(b, ages = 1:3, years = 1:3, type = "heatmap")
+
 ```
@@
 
-The function `sim_survey` can be used to simulate data from one survey over a population simulated using `sim_distribution`:
+The function `sim_survey` can be used to simulate data from one survey over a population simulated using `sim_distribution`.
+This function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and definitions of the design of the survey.
+Specifically, the `q` argument requires a closure, such as `sim_logistic`, for defining the probability of catching specific age groups, trawl dimensions are defined in the `trawl_dim` argument, and set, length and age sampling effort are defined using the `set_den`, `lengths_cap` and `ages_cap` arguments, respectively.
+Multiple simulations of the same survey can be run using the `n_sims` argument, however, it is important not to request large numbers of simulations because this function can quickly deplete a computers RAM.
+Defaults used in the function and for the case study roughly correspond to the assumed catchability in NAFO division 3Ps as well as the sampling protocol used in the survey of this stock (Table \@ref(tab:par)).
+Below we use `sim_survey` to simulate two surveys over the same population, of which one is set-up to have higher sampling effort than the other.
+
+```{r, echo = TRUE, eval = TRUE, cache = TRUE}
+
+set.seed(438)
+pop <- sim_abundance() %>% 
+  sim_distribution()
+a <- pop %>% 
+  sim_survey(n_sims = 5,
+             set_den = 1 / 1000,
+             lengths_cap = 100,
+             ages_cap = 5)
+b <- pop %>% 
+  sim_survey(n_sims = 5,
+             set_den = 5 / 1000,
+             lengths_cap = 500,
+             ages_cap = 25)
+
+```
+
+Again, this function retains all the objects listed in the output of `sim_distribution` and adds data.tables that detail the set locations (`setdet`) and sampling details (`samp`).
+Catchability corrected abundance matrices, named `I` and `I_at_length`, are also produced and added to the output; these matrices are useful for comparing the true abundance available to the survey to abundance estimates obtained using survey-based or model-based analyses of the simulated survey data.
+Specific surveys can be explored using the `plot_survey` function, which uses plotly [@sievert2018] and crosstalk [@cheng2016] in the background to link the bubble plot of aggregate set catch to the histogram of lengths and ages sampled to facilitate explorations of set-specific catches.
+The plots generated below have been combined here into a static plot for this paper (Figure \@ref(fig:survey)).
 
 ```{r, echo = TRUE, eval = FALSE}
-s <- sim_survey(d, n_sims = 1,
-                q = sim_logistic(k = 2, x0 = 3),
-                trawl_dim = c(1.5, 0.02),
-                set_den = 2 / 1000,
-                lengths_cap = 500,
-                length_group = 1,
-                ages_cap = 10)
+
+plot_survey(a, which_sim = 1, which_year = 20)
+plot_survey(b, which_sim = 1, which_year = 20)
+
 ```
 
-Again, this function retains all the objects listed in `d` and adds objects that detail the set locations (`setdet`) and sampling details (`samp`).
+```{r survey, fig.cap = "- Bubble plots and histograms of set catches produced by `plot_survey` when supplied survey data simulated using `sim_survey` under relatively a) high and b) low sampling effort.
+Note that the first simulation of the survey in the first year is depicted here."}
+
+plot_a <- plot_survey(a, which_sim = 1, which_year = 1)
+plot_b <- plot_survey(b, which_sim = 1, which_year = 1)
+
+p <- subplot(plot_a, plot_b, nrows = 2) %>% 
+  add_labels(text = "a)", x = 0, y = 1, yshift = 20, xshift = -20) %>% 
+  add_labels(text = "b)", x = 0, y = 0.5, yshift = 20, xshift = -20)
+
+export_plot(p, file = "analysis/paper/figures/plot_survey.png", width = 500, height = 600)
+knitr::include_graphics("figures/plot_survey.png")
+
+```
+
+- You need to fix up the aspect ratios and margins in the plot_survey subplot
+- Mention the `n_sims` argument, mention RAM limitations, and then move to `sim_survey_parallel`
 
@@
 # Tables
-
-Table: (\#tab:par) - Main function calls, with argument descriptions and associated parameter symbols, used to simulate the cod-like population used in the case study presented here.
-
-|                                       **Function call** |                                            **Description** |                 **Symbol** |
-|:------------------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
-| `> set.seed(438)`                                       |      Set random number generator seed to replicate results |                            |
-| `>`                                                     |                                                            |                            |
-| `> n <- sim_abundance(`                                 |                                       *Simulate abundance* |                            |         
-| `+   ages = 1:20,`                                      |                                                       Ages |                        $a$ |
-| `+   years = 1:20,`                                     |                                                      Years |                        $y$ |
-| `+   R = sim_R(mean = 30000000,`                        |                                           Mean recruitment |                    $\mu_r$ |
-| `+             log_sd = 0.5),`                          |                      Standard deviation of log-recruitment |                 $\sigma_r$ |
-| `+   Z = sim_Z(mean = 0.5,`                             |                                       Mean total mortality |                    $\mu_Z$ |
-| `+             log_sd = 0.2,`                           |                      Standard deviation of total mortality |                 $\sigma_Z$ |
-| `+             phi_age = 0.9,`                          |    Correlation across ages in error around total mortality |    $\varphi_{\delta, age}$ |
-| `+             phi_year = 0.5),`                        |   Correlation across years in error around total mortality |   $\varphi_{\delta, year}$ |
-| `+   growth = sim_vonB(Linf = 120,`                     |                                Mean asymptotic length (cm) |               $l_{\infty}$ |
-| `+                     L0 = 5,`                         |                                       Length at birth (cm) |                      $l_0$ |
-| `+                     K = 0.1,`                        |                                                Growth rate |                        $K$ |
-| `+                     log_sd = 0.1,`                   |      Standard deviation of the von Bertalanffy growth curve |                 $\sigma_l$ |
-| `+                     length_group = 3))`              |                  Length group for abundance at length (cm) |            $l_{group}^{N}$ |      
-| `>`                                                     |                                                            |                            |
-| `> g <- make_grid(`                                     |                                              *Make a grid* |                            |
-| `+   x_range = c(-140, 140),`                           |                           Range of grid in the x dimension |                            |                   
-| `+   y_range = c(-140, 140),`                           |                           Range of grid in the y dimension |                            | 
-| `+   res = c(3.5, 3.5),`                                |                                       Grid resolution (km) |                        $R$ |
-| `+   shelf_depth = 200,`                                |                                           Shelf depth (km) |                $d_{shelf}$ |
-| `+   shelf_width = 100,`                                |                                           Shelf width (km) |                $w_{shelf}$ |
-| `+   depth_range = c(0, 1000),`                         |                                            Depth range (m) |       $[d_{min}, d_{max}]$ |
-| `+   n_div = 1,`                                        |                                        Number of divisions |                  $N_{div}$ |
-| `+   strat_breaks = seq(0, 1000, by = 40))`             |                 Series of depth breaks for defining strata |                            |
-| `>`                                                     |                                                            |                            |
-| `> d <- sim_distribution(`                              |                                    *Simulate distribution* |                            |     
-| `+   sim = n,`                                          |                Object produced by `sim_abundance` function |                            |     
-| `+   grid = g,`                                         |                    Object produced by `make_grid` function |                            |
-| `+   ays_covar = sim_ays_covar(sd = 2.8,`               |          Standard deviation of age-year-space distribution |             $\sigma_{\xi}$ |
-| `+                             range = 300,`            |                          Range of spatial correlation (km) |                        $r$ |
-| `+                             phi_age = 0.5,`          |            Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |
-| `+                             phi_year = 0.9,`         |           Correlation across years in spatial distribution |      $\varphi_{\xi, year}$ |
-| `+                             group_ages = 5:20),`     |       Make space-age-year variance equal across these ages |                            |
-| `+   depth_par = sim_parabola(mu = 200,`                |          Depth at which abundance is typically highest (m) |                    $\mu_d$ |
-| `+                            sigma = 70))`             |              Dispersion around depth of peak abundance (m) |                 $\sigma_d$ |
-| `>`                                                     |                                                            |                            |
-| `> e <- expand_surveys(`                                |                       *Define a series of surveys to test* |                            |
-| `+   set_den = c(0.0005, 0.001, 0.002, 0.005, 0.01),`   |                                       Set density (km^-2^) |                 $D_{sets}$ |
-| `+   lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),`   |                 Maximum number of lengths to collect / set |              $M_{lengths}$ |
-| `+   ages_cap = c(2, 5, 10, 20, 50))`                   | Maximum number of ages to sample / length group / division |                 $M_{ages}$ |
-| `>`                                                     |                                                            |                            |
-| `> s <- test_surveys(`                                  |                             *Simulate and analyze surveys* |                            |
-| `+   sim = d,`                                          |             Object produced by `sim_distribution` function |                            |
-| `+   surveys = e,`                                      |               Object produced by `expand_surveys` function |                            |
-| `+   n_loops = 1000,`                                   |                    Number of times to simulate each survey |                            | 
-| `+   q = sim_logistic(k = 2,`                           |                Steepness of logistic curve of catchability |                        $k$ |
-| `+                   x0 = 3),`                          |                 Midpoint of logistic curve of catchability |                      $x_0$ |
-| `+   trawl_dim = c(1.5, 0.02),`                         |       Trawl dimensions (km) - i.e. area covered by a trawl |                $A_{trawl}$ |
-| `+   length_group = 1)`                                 |                         Length group for age sampling (cm) |                $l_{group}$ |
-
-*Caution:* as written, the `test_surveys` function call may take days or weeks to run.
-Run duration can be decreased by running multiple simulations per loop (`n_sims` argument) and by running the loops in parallel (`cores` argument).
-See help file for more details (`?test_surveys`).
-
-<br>
 

### 2018-11-18 R/plot_sim.R

@@ -155,8 +155,7 @@ plot_distribution <- function(sim, ages = 1:10, years = 1:10,
 
 #' @export
 #' @rdname plot_trend
-plot_survey <- function(sim, which_year = 1, which_sim = 1,
-                        col = viridis::viridis(1)) {
+plot_survey <- function(sim, which_year = 1, which_sim = 1) {
 
   xax <- list(
     title = "",
@@ -173,8 +172,21 @@ plot_survey <- function(sim, which_year = 1, which_sim = 1,
 
   setdet <- sim$setdet
   setdet <- setdet[setdet$year == which_year & setdet$sim == which_sim, ]
+  samp <- sim$samp
+  samp <- samp[samp$set %in% setdet$set, ]
   samp <- merge(setdet[, c("year", "sim", "set", "x", "y")],
-                sim$samp, by = "set")
+                samp, by = "set", all = TRUE)
+
+  l <- sim$I_at_length[, as.character(which_year)]
+  l <- (l / sum(l)) * 100
+  true_length <- data.frame(length = as.numeric(names(l)), percent = l)
+  a <- sim$I[, as.character(which_year)]
+  a <- (a / sum(a)) * 100
+  true_age <- data.frame(age = as.numeric(names(a)), percent = a)
+
+  ## impose the same length grouping on the data as chosen for the abundance
+  ## at length grouping
+  length_group <- get("length_group", envir = environment(sim$sim_length))
 
   d <- crosstalk::SharedData$new(samp, ~set)
 
@@ -182,46 +194,54 @@ plot_survey <- function(sim, which_year = 1, which_sim = 1,
 
   sp_p <- base %>%
     group_by(set) %>%
-    summarise(x = unique(x), y = unique(y), n = n()) %>%
+    summarise(x = unique(x), y = unique(y), n = sum(!is.na(measured))) %>%
     add_markers(x = ~x, y = ~y, size = ~n, text = ~n,
-                color = I(col), sizes = c(5, 500), name = "n",
-                showlegend = FALSE) %>%
-    add_markers(data = setdet[setdet$n == 0, ], color = I(col),
-                x = ~x, y = ~y, text = ~n, size = I(5), symbol = I(4),
-                name = "zero", alpha = 0.2,
+                color = ~n, sizes = c(2, 600), name = "n",
                 showlegend = FALSE) %>%
     add_paths(data = df_strat, x = ~long, y = ~lat, color = I("black"),
               hoverinfo = "none", size = I(0.5), showlegend = FALSE,
[…]

### 2018-11-19 analysis/paper/sim_survey_paper.Rmd
Draft demo of sim_survey and plot_survey

@@
   add_labels(text = rep(c("Age 1", "Age 2", "Age 3"), 2), x = c(0.04, 0.20, 0.4, 0.6, 0.8, 0.95), y = rep(0.995, 6), yshift = 20) %>% 
-  add_labels(text = rep(c("Year 1", "Year 2", "Year 3"), 2), x = c(rep(0.005, 3), rep(0.54, 3)), y = c(0.08, 0.5, 0.92), xshift = -20, textangle = -90) %>% 
+  add_labels(text = rep(c("Year 3", "Year 2", "Year 1"), 2), x = c(rep(0.005, 3), rep(0.54, 3)), y = c(0.08, 0.5, 0.92), xshift = -20, textangle = -90) %>% 
   add_labels(text = "a)", x = 0, y = 1, yshift = 20, xshift = -20) %>% 
@@
 ```{r survey, fig.cap = "- Bubble plots and histograms of set catches produced by `plot_survey` when supplied survey data simulated using `sim_survey` under relatively a) high and b) low sampling effort.
-Note that the first simulation of the survey in the first year is depicted here."}
+Histograms of length and age composition include the distribution caught and sampled overlaid with a line of the true distribution of length and ages available to the survey.
+Note that the first simulation of the survey in year 20 is depicted here."}
 
-plot_a <- plot_survey(a, which_sim = 1, which_year = 1)
-plot_b <- plot_survey(b, which_sim = 1, which_year = 1)
+plot_a <- plot_survey(a, which_sim = 1, which_year = 20) %>% 
+  add_labels(text = "a)", x = 0, y = 1, yshift = 5, xshift = -5) %>% 
+  layout(margin = list(l = 10, r = 10, t = 10, b = 10))
+plot_b <- plot_survey(b, which_sim = 1, which_year = 20) %>% 
+  add_labels(text = "b)", x = 0, y = 1, yshift = 5, xshift = -5) %>% 
+  layout(margin = list(l = 10, r = 10, t = 10, b = 10))
 
-p <- subplot(plot_a, plot_b, nrows = 2) %>% 
-  add_labels(text = "a)", x = 0, y = 1, yshift = 20, xshift = -20) %>% 
-  add_labels(text = "b)", x = 0, y = 0.5, yshift = 20, xshift = -20)
+## plotly::subplot was not amenable to combining these two plots, I therefore used ImageMagick to combine a and b
 
-export_plot(p, file = "analysis/paper/figures/plot_survey.png", width = 500, height = 600)
+paths <- c(a = "analysis/paper/figures/plot_survey_a.png", b = "analysis/paper/figures/plot_survey_b.png")
+export_plot(plot_a, file = paths["a"], width = 800, height = 400)
+export_plot(plot_b, file = paths["b"], width = 800, height = 400)
+
+image_a <- magick::image_read(paths["a"])
+image_b <- magick::image_read(paths["b"])
+p <- magick::image_append(c(image_a, image_b), stack = TRUE)
+magick::image_write(p, path = "analysis/paper/figures/plot_survey.png")
+
 knitr::include_graphics("figures/plot_survey.png")
@@
 
-- You need to fix up the aspect ratios and margins in the plot_survey subplot
-- Mention the `n_sims` argument, mention RAM limitations, and then move to `sim_survey_parallel`
+As noted above, the amount of RAM a user's computer has may limit the utility of the `sim_survey` function for running thousands of simulations of the same survey.
+The `sim_survey_parallel` function was therefore constructed to facilitate the running of thousands of simulations without running into RAM limitations.
+This function is set-up to run multiple `sim_survey` calls in parallel using the doParallel package [@weston2015] and, as such, multiple loops can be run using the `n_loops` argument and within each loop multiple simulations can be run (controled using the `n_sims` argument).
+Total simulations will be the product of `n_loops` and `n_sims` arguments.
+If more than one core (`cores`) is specified, then the simulations will be run in parallel to speed up the process.
+Low numbers of `n_sims` and high numbers of `n_loops` will be easier on RAM, but may be slower.
+The optimum ratio of `n_sims` to `n_loops` will depend on the amount of RAM and number of cores the user's computer has.
+In any case, this function simplifies the process of running thousands of simulations of the same survey and the simulated data can then be supplied to survey-based or model-based analyses that require simulation testing.
+Because the SimSurvey package was constructed in a context where survey-based analyses are commonplace, specifically stratified analyses, the package facilities include functions for calculating and assessing bias and precision of stratified estimates of abundance.
 

### 2018-11-19 R/sim_survey.R

@@ -278,7 +278,7 @@ sim_survey <- function(sim, n_sims = 1, q = sim_logistic(), trawl_dim = c(1.5, 0
 #'
 
 sim_survey_parallel <- function(sim, n_sims = 1, n_loops = 100,
-                                cores = 6, quiet = FALSE, ...) {
+                                cores = 1, quiet = FALSE, ...) {
 
   start <- Sys.time()
   one_res <- sim_survey(sim, n_sims = n_sims, light = TRUE, ...)

### 2018-11-19 analysis/paper/sim_survey_paper.Rmd
Draft stratified analysis section

@@
 In any case, this function simplifies the process of running thousands of simulations of the same survey and the simulated data can then be supplied to survey-based or model-based analyses that require simulation testing.
-Because the SimSurvey package was constructed in a context where survey-based analyses are commonplace, specifically stratified analyses, the package facilities include functions for calculating and assessing bias and precision of stratified estimates of abundance.
 
@@
 
-A series of `r # nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests).
-Each survey was simulated `r # max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain annual estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
+Once the survey data are simulated, they then need to be analyzed to obtain an index of abundance.
+While there are many model-based options for obtaining an index (**TODO: add references**), survey-based approaches, such as stratified analyses, are often used.
+Here we apply formula presented in Smith and Somerton [-@Smith1981] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
+Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be bumped up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key.
 We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:
@@
 $$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})^2}{N_a N_y N_k}}$$
+
 Where $N_a$, $N_y$, and $N_k$ are the number of years, ages and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e.
 catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$).
-RMSE was also calculated for abundance at length estimates.
-Intra-haul correlation (i.e.
-intraclass correlation) of set catch was also calculated by fitting a mixed-model to the lengths data with set as a random effect [fit using the lme4 package; @Bates2015]; variance estimates from this model were then plugged into the ICC(1) formula presented in McGraw and Wong [-@Mcgraw1996].
-All simulations and analyses were conducted using `r substr(version$version.string, 1, 15)` [@R2017] and the code has been developed into a package called `SimSurvey` to facilitate future work and collaboration (https://github.com/PaulRegular/SimSurvey).
+RMSE was also calculated for abundance at length estimates, where the above formula is indexed by length groups $l$, and total abundance, which lacks a group index of $a$ or $l$.
+
+Stratified estimates of abundance obtained are by supplying the output from `sim_survey` to the `run_strat` function.
+RMSE of the stratified estimates can then be calculated using the `strat_error` function.
+Results and error of a stratified analysis of one default survey over a default population are obtained using the following code:
+
+```{r, echo = TRUE, eval = FALSE}
+
+set.seed(438)
+sim <- sim_abundance() %>% 
+  sim_distribution() %>% 
+  sim_survey() %>% 
+  run_strat() %>% 
+  strat_error()
+
+```
+
+The returned object will include all the objects accumulated through the `sim_abundance` to `strat_error`.
+The `run_strat` function adds three data.tables called `total_strat`, `length_strat` and `age_strat` that include stratified estimates of total abundance, abundance at length, and abundance at age, respectively.
+To this, `strat_error` adds data.tables ending with `_strat_error` or `_strat_error_stats`.
+The `_strat_error` objects simply contain stratified estimates of abundance (column named `I_hat`) with corresponding true values of abundance available to the survey (column named `I`) and the `strat_error_stats` data.frame includes metrics of mean absolute error (`MAE`), mean-squared error (`MSE`) and root-mean-squared error (`RMSE`).
+
+## Testing protocol
+
+...
 

### 2018-11-19 R/make_grid.R

@@ -84,7 +84,7 @@ make_grid <- function(x_range = c(-140, 140), y_range = c(-140, 140),
   rstrat <- r$strat
   max_id <- 0
   for (s in sort(unique(rstrat))) {
-    rc <- clump(rstrat == s)
+    rc <- suppressMessages(raster::clump(rstrat == s))
     new_ids <- rc[!is.na(rc[])] + max_id
     r$strat[!is.na(rc)] <- new_ids
     max_id <- max(new_ids)

### 2018-11-20 analysis/paper/sim_survey_paper.Rmd
Finish drafting rejigged methods section

@@
 
-```{r data, include = FALSE}
+```{r packages, include = FALSE}
 library(SimSurvey)
@@
 Sys.setenv("MAPBOX_TOKEN" = "pk.eyJ1IjoicGF1bHJlZ3VsYXIiLCJhIjoiY2pueGIxejMxMGM0MjNqbjR1aXdwYm92bSJ9.R-t1SCa0VHKf4N-PQmYI5Q")
-# load("analysis/cod_sim_exports/2018-10-26_age_clust_test/test_output.RData")
 source("analysis/paper/helper_functions.R")
@@
 | `>`                                                     |                                                            |                            |
-| `> n <- sim_abundance(`                                 |                                       *Simulate abundance* |                            |         
+| `> abundance <- sim_abundance(`                         |                                       *Simulate abundance* |                            |         
 | `+   ages = 1:20,`                                      |                                                       Ages |                        $a$ |
@@
 | `>`                                                     |                                                            |                            |
-| `> g <- make_grid(`                                     |                                              *Make a grid* |                            |
+| `> grid <- make_grid(`                                  |                                              *Make a grid* |                            |
 | `+   x_range = c(-140, 140),`                           |                           Range of grid in the x dimension |                            |                   
@@
 | `>`                                                     |                                                            |                            |
-| `> d <- sim_distribution(`                              |                                    *Simulate distribution* |                            |     
-| `+   sim = n,`                                          |                Object produced by `sim_abundance` function |                            |     
-| `+   grid = g,`                                         |                    Object produced by `make_grid` function |                            |
+| `> distribution <- sim_distribution(`                   |                                    *Simulate distribution* |                            |     
+| `+   sim = abundance,`                                  |                Object produced by `sim_abundance` function |                            |     
+| `+   grid = grid,`                                      |                    Object produced by `make_grid` function |                            |
 | `+   ays_covar = sim_ays_covar(sd = 2.8,`               |          Standard deviation of age-year-space distribution |             $\sigma_{\xi}$ |
@@
 | `>`                                                     |                                                            |                            |
-| `> e <- expand_surveys(`                                |                       *Define a series of surveys to test* |                            |
+| `> surveys <- expand_surveys(`                          |                       *Define a series of surveys to test* |                            |
 | `+   set_den = c(0.0005, 0.001, 0.002, 0.005, 0.01),`   |                                       Set density (km^-2^) |                 $D_{sets}$ |
@@
 | `>`                                                     |                                                            |                            |
-| `> s <- test_surveys(`                                  |                             *Simulate and analyze surveys* |                            |
-| `+   sim = d,`                                          |             Object produced by `sim_distribution` function |                            |
-| `+   surveys = e,`                                      |               Object produced by `expand_surveys` function |                            |
+| `> tests <- test_surveys(`                              |                             *Simulate and analyze surveys* |                            |
+| `+   sim = distribution,`                               |             Object produced by `sim_distribution` function |                            |
+| `+   surveys = surveys,`                                |               Object produced by `expand_surveys` function |                            |
 | `+   n_loops = 1000,`                                   |                    Number of times to simulate each survey |                            | 
@@
 The `sim_survey_parallel` function was therefore constructed to facilitate the running of thousands of simulations without running into RAM limitations.
-This function is set-up to run multiple `sim_survey` calls in parallel using the doParallel package [@weston2015] and, as such, multiple loops can be run using the `n_loops` argument and within each loop multiple simulations can be run (controled using the `n_sims` argument).
+This function is set-up to run multiple `sim_survey` calls in parallel using the doParallel package [@weston2015] and, as such, multiple loops can be run using the `n_loops` argument and within each loop multiple simulations can be run (controlled using the `n_sims` argument).
 Total simulations will be the product of `n_loops` and `n_sims` arguments.
@@
 
-## Testing protocol
+## Testing survey protocol
 
-...
+Assuming a stratified analysis as the default method for obtaining an index of abundance, a series of survey protocol can be tested using the `test_surveys` function.
+Provided a simulated population from `sim_distribution` and a series of survey protocol from `expand_surveys`, this function will simulate and analyze data from each survey using the `sim_survey`, `run_strat` and `strat_error` functions.
+Like `sim_survey_parallel`, this function operates in parallel and allows the specification of `n_sims` and `n_loops`, and the product of these two arguments equals the number of times each survey is simulated.
+Keep in mind that low numbers of `n_sims` and high numbers of `n_loops` will be easier on RAM, but may be slower, especially if the work is spread across few `cores`.
+The below code will run a small subset of the simulation tests run for the case study presented in this paper (see Table \@ref(tab:par) for code to replicate the case study results).
+Specifically, `expand_surveys` sets up a series of eight surveys to test (i.e.
+all possible combinations of `set_den`, `lengths_cap` and `ages_cap` protocol) and the `test_surveys` function will run 250 simulations of each survey and compare stratified estimates of abundance to the true abundance available to the survey.
+Though a limited test, this may take a while to run.
+
+```{r, echo = TRUE, eval = FALSE}
+
+set.seed(438)
+pop <- sim_abundance() %>%
+  sim_distribution()
+
+surveys <- expand_surveys(set_den = c(1, 2) / 1000,
+                          lengths_cap = c(100, 500),
+                          ages_cap = c(5, 20))
+
+tests <- test_surveys(pop, surveys = surveys,
+                      n_sims = 5, n_loops = 50, cores = 2)
+
+```
+
+Processing time will system (i.e.
+amount of RAM and number of cores) and setting (i.e.
+`n_loops` and `n_sims` ratio) dependent.
+The `test_survey` function will print a progress bar, generated using the progress package [@csardi2016], which details percent completion and will also include an eta after the first step of the loop completes.
+Thousands of simulations of hundreds of survey designs, like run here for our case study, may take days to run.
+The `test_surveys` function therefore includes an option for exporting intermediate results to a local directory, via the `export_dir` argument, and the `resume_test` function can be used resume a `test_surveys` run that had to be stopped part way through the process.
+The final object produced will be a list that includes all objects from `sim_abundance` and `sim_distribution` with the table of survey designs tested (named `surveys`) and tables produced by `strat_error` that end with the names `_strat_error` and `_strat_error_stats`.
+These tables include a `survey` column to allow merging of the survey protocol table with the error tables.
+Objects produced by `sim_survey` (set and sampling details) and `run_strat` (full stratified analysis results) are not retained to minimize object size.
+Like other core functions, some convenience functions are included in SimSurvey for creating interactive plots of the results from `test_surveys`.
+These will be demonstrated in the [Results] section using the results from the case study.
 
@@
 This is a short, non-inclusive, list of additional variability that likely occurs in the real-world but was not implemented here for simplicity.
+Nevertheless, the SimSurvey package provides the most realistic operating model that we are aware of for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
 
@@
 
-```{r icc, cache = TRUE}
-len_samp <- merge(sim$setdet, sim$samp, by = "set")
-len_samp <- len_samp[len_samp$sim == 1 & len_samp$measured, ]
-sim_icc <- round(icc(len_samp$length, len_samp$set), 1)
-sim_icc <- sprintf("%.2f", sim_icc)
-real_icc <- "0.70" # calculated in the file "imatate_cod_data.R"
+```{r data, cache = TRUE}
+load("analysis/cod_sim_exports/2018-10-26_age_clust_test/test_output.RData")
+tests <- sim
 ```

### 2018-11-21 analysis/paper/sim_survey_paper.Rmd
Revisit methods. Start adding plots for test_surveys output. Also set-up a manual cache for the plots.

@@
 ```{r setup, include=FALSE}
+
 # devtools::install_github("rstudio/rmarkdown@v1.10")
@@
 knitr::opts_knit$set(root.dir = "../..")
+options(scipen = 999) # turn off scientific notation to simplify printing
+
 ```
@@
 ```{r packages, include = FALSE}
+
 library(SimSurvey)
@@
 source("analysis/paper/helper_functions.R")
+
+replot <- FALSE # plot the figures again? (acts as a manual cache)
+
 ```
@@
 Below we outline the equations behind a series of SimSurvey functions that were used to simulate survey data with similar features to data obtained by the survey of 3Ps cod.
-The settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data.
+The settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data; these values were set as the defaults for the main functions in the package.
 Here we evaluate the efficacy of alternate sampling protocol by assessing deviation between stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
@@
 Table: (\#tab:par) - Main function calls, with argument descriptions and associated parameter symbols, used to simulate the cod-like population used in the case study presented here.
+Note that the settings below largely correspond with the defaults used in the functions.
 
@@
 Overall, the function provides a simple tool for simulating a range dynamic age-structured populations.
-For instance, below we simulate a relatively long and short lived species:
+For instance, below we simulate a relatively long (default case study settings) and short lived species.
 
-```{r, echo = TRUE, eval = TRUE}
+```{r, echo = TRUE, eval = replot}
 set.seed(438)
@@
 The package also includes several plotting function for making quick plotly-based [@sievert2018] interactive visuals of the simulated population.
+For instance, the `plot_surface` function can be used to make quick visuals of matricies contained within the list returned by `sim_abundance`.
 The plots output from the below code are combined into Figure \@ref(fig:surface).
+Here we display the abundance at age matrix (object named `N` in the list produceb by `sim_abundance`); other names can be supplied to the `mat` argument to visualize a different matrix from the `sim_abundance` list, such as `Z`.
 
@@
 
-```{r surface, fig.cap = "- Plots produced by `plot_surface` when supplied an list produced by `sim_abundance` with settings that simulate a relatively a) long and b) short lived species.
-The abundance at age matrix (`N`) is displayed by default; the `mat` argument can be used to specify another matrix contained within the supplied list, such as `Z`."}
-
+```{r, eval = replot}
 set.seed(438)
@@
 export_plot(p, file = "analysis/paper/figures/plot_surface.png", width = 800, height = 370)
+```
+
+```{r surface, fig.cap = "- Surface plots of simulated abundance at age of a relatively a) long lived (case study settings) and b) short lived species.
+These plots were produced by `plot_surface` when supplied a list produced by `sim_abundance`."}
+
 knitr::include_graphics("figures/plot_surface.png")
@@
 If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied to `sim_distribution` if the same structure is used.
-The settings used to make the grid used in our case study are shown in Table \@ref(tab:par) and these settings are also the defaults.
-Below we generate and plot (combined in Figure \@ref(fig:grid)) a default grid and another grid with four divisions and a more linear gradient in depth.
+The package includes a manually constructed survey grid of NAFO division 3Ps (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid.
+However, for simplicity, we use `make_grid` to construct a square grid for our case study and the settings used are shown in Table \@ref(tab:par).
+Below we generate and plot (combined in Figure \@ref(fig:grid)) a default case study grid and another grid with four divisions and a more linear gradient in depth.
 
@@
 
-```{r grid, fig.cap = "- Plots produced by `plot_grid` when supplied an raster object produced by `make_grid` a) using default settings and b) settings that generate a more linear depth gradient.
-In these plots, the color gradient represents depth, the grey lines deliniate divisions and white lines deliniate strata."}
+```{r eval = replot}
 
@@
 export_plot(p, file = "analysis/paper/figures/plot_grid.png", width = 800, height = 430)
+
+```
+
+```{r grid, fig.cap = "- Plots produced by `plot_grid` when supplied an raster object produced by `make_grid` a) using default case study settings and b) settings generates a grid with four divisions and a more linear depth gradient.
+In these plots, the color gradient represents depth, the grey lines deliniate divisions and white lines deliniate strata."}
+
 knitr::include_graphics("figures/plot_grid.png")
@@
 
-```{r, echo = TRUE, eval = TRUE, cache = TRUE}
+```{r, echo = TRUE, eval = replot}
 
@@
 
-```{r distribution, fig.cap = "- Distribution plots of a population simulated using `sim_distribution` that forms a) tight clusters with stronger correlation through years than ages, and b) relatively diffuse clusters with stronger correlation through ages than years."}
+```{r, eval = replot}
 
@@
 export_plot(p, file = "analysis/paper/figures/plot_distribution.png", width = 600, height = 300)
+
+```
+
+```{r distribution, fig.cap = "- Distribution plots of simulated populations that form a) tight clusters with stronger correlation through years than ages (case study settings), and b) relatively diffuse clusters with stronger correlation through ages than years.
+This plot is a facet of plots produced by `plot_distribution` when supplied simulations from `sim_distribution`."}
+
 knitr::include_graphics("figures/plot_distribution.png")
@@
 Defaults used in the function and for the case study roughly correspond to the assumed catchability in NAFO division 3Ps as well as the sampling protocol used in the survey of this stock (Table \@ref(tab:par)).
-Below we use `sim_survey` to simulate two surveys over the same population, of which one is set-up to have higher sampling effort than the other.
+Below we use `sim_survey` to simulate two surveys over our case study population, of which one is set-up to have higher sampling effort than the other.
 
-```{r, echo = TRUE, eval = TRUE, cache = TRUE}
+```{r, echo = TRUE, eval = replot}
 
@@
 
-```{r survey, fig.cap = "- Bubble plots and histograms of set catches produced by `plot_survey` when supplied survey data simulated using `sim_survey` under relatively a) high and b) low sampling effort.
-Histograms of length and age composition include the distribution caught and sampled overlaid with a line of the true distribution of length and ages available to the survey.
-Note that the first simulation of the survey in year 20 is depicted here."}
+```{r eval = replot}
 
@@
 magick::image_write(p, path = "analysis/paper/figures/plot_survey.png")
+
+```
+
+```{r survey, fig.cap = "- Bubble plots and histograms of set catches from a simulated stratified-random survey of the case study population under relatively a) high and b) low sampling effort.
+Histograms of length and age composition include the distribution caught and sampled overlaid with a line of the true distribution of length and ages available to the survey.
+Note that the first simulation of the survey in year 20 is depicted here.
+These plots are produced by `plot_survey` when supplied survey data simulated using `sim_survey`."}
 
@@
 Keep in mind that low numbers of `n_sims` and high numbers of `n_loops` will be easier on RAM, but may be slower, especially if the work is spread across few `cores`.
-The below code will run a small subset of the simulation tests run for the case study presented in this paper (see Table \@ref(tab:par) for code to replicate the case study results).
-Specifically, `expand_surveys` sets up a series of eight surveys to test (i.e.
-all possible combinations of `set_den`, `lengths_cap` and `ages_cap` protocol) and the `test_surveys` function will run 250 simulations of each survey and compare stratified estimates of abundance to the true abundance available to the survey.
-Though a limited test, this may take a while to run.
+Because most of the default settings of the functions match the case study settings, the below code will replicate the results from our case study (see Table \@ref(tab:par) for a more detail).
+The `expand_surveys` function sets up a series of `r nrow(expand_surveys())` surveys to test (i.e.
+all possible combinations of `set_den`, `lengths_cap` and `ages_cap` protocol) and the `test_surveys` function will run 1000 simulations of each survey and compare stratified estimates of abundance to the true abundance available to the survey.
 
@@
 
-surveys <- expand_surveys(set_den = c(1, 2) / 1000,
-                          lengths_cap = c(100, 500),
-                          ages_cap = c(5, 20))
+surveys <- expand_surveys(set_den = c(0.0005, 0.001, 0.002, 0.005, 0.01),
+                          lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
+                          ages_cap = c(2, 5, 10, 20, 50))
 
 tests <- test_surveys(pop, surveys = surveys,
-                      n_sims = 5, n_loops = 50, cores = 2)
+                      n_sims = 5, n_loops = 200, cores = 2)
+
+```
+
+```{r data, eval = replot}
+
+load("analysis/cod_sim_exports/2018-10-26_age_clust_test/test_output.RData")
+tests <- sim
 
@@
 The `test_survey` function will print a progress bar, generated using the progress package [@csardi2016], which details percent completion and will also include an eta after the first step of the loop completes.
-Thousands of simulations of hundreds of survey designs, like run here for our case study, may take days to run.
+Keep in mind that thousands of simulations of hundreds of survey designs may take days to run.
 The `test_surveys` function therefore includes an option for exporting intermediate results to a local directory, via the `export_dir` argument, and the `resume_test` function can be used resume a `test_surveys` run that had to be stopped part way through the process.
@@
 Like other core functions, some convenience functions are included in SimSurvey for creating interactive plots of the results from `test_surveys`.
-These will be demonstrated in the [Results] section using the results from the case study.
+For instance a series of plotting functions ending in `_fan` produces fancharts where stratified estimates of abundance from each simulated survey are converted into a series of quantiles to depict the probability that estimates fall within a praticular range.
+True values of abundance available to the survey are overlaid on the series of probability envelopes.
+These plots help visually assess the level of percision and bias from a specific set of survey protocol.
+The three lines of code below will produce interactive fancharts for stratified estimates of total abundance, abundance at length and abundance at age, respectively.
+A subset of these plots have been combined into static plots (Figure \@ref(fig:fan1), \@ref(fig:fan2), \@ref(fig:fan3)).
+
+```{r, echo = TRUE, eval = FALSE}
+
+plot_total_strat_fan(tests)
+plot_length_strat_fan(tests, years = 1:20, lengths = 1:100)
+plot_age_strat_fan(tests, years = 1:20, ages = 1:10)
+
+```
+
+```{r eval = replot}
+
+surveys <- tests$surveys
+set_dens <- c(0.0005, 0.002, 0.01)
+i <- surveys$survey[surveys$set_den == set_dens[1] & surveys$lengths_cap == 10 & surveys$ages_cap == 10] # lengths_cap and ages_cap does not matter here
+a <- plot_total_strat_fan(tests, surveys = i)
+i <- surveys$survey[surveys$set_den == set_dens[2] & surveys$lengths_cap == 10 & surveys$ages_cap == 10]
+b <- plot_total_strat_fan(tests, surveys = i)
+i <- surveys$survey[surveys$set_den == set_dens[3] & surveys$lengths_cap == 10 & surveys$ages_cap == 10]
+c <- plot_total_strat_fan(tests, surveys = i)
+
+p <- subplot(a, b, c, nrows = 1, shareY = TRUE, titleX = FALSE) %>% 
+  add_labels(x = 0.5, y = 0, text = "Year", yshift = -40) %>% 
+  add_labels(x = c(0.1, 0.5, 0.9), y = rep(1, 3), yshift = 20, 
+             text = paste("D<sub>sets</sub> =", set_dens))
+
+export_plot(p, file = "analysis/paper/figures/plot_total_strat_fan.png", width = 700, height = 250)
+
+```
+
+```{r fan1, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
+The heavy dark line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
+
+knitr::include_graphics("figures/plot_total_strat_fan.png")
+
+```
+
+```{r eval = replot}
+
+facet_fan <- function(data, ages = NULL, lengths = NULL, ylim = NULL, set_dens = c(0.0005, 0.002, 0.01), lengths_caps = c(10, 100, 1000)) {
+  surveys <- data$surveys
+  s <- expand.grid(set_den = set_dens, lengths_cap = lengths_caps)
+  plot_list <- vector("list", nrow(s))
+  for (i in seq_along(set_dens)) {
+    for (j in seq_along(lengths_caps)) {
+      s_i <- surveys$survey[surveys$set_den == set_dens[i] & surveys$lengths_cap == lengths_caps[j] & surveys$ages_cap == 10]
+      if (is.null(ages)) {
+        p <- plot_length_strat_fan(data, surveys = s_i, years = 7, lengths = lengths) %>% 
+         layout(yaxis = list(range = ylim))
+      } else {
+        p <- plot_age_strat_fan(data, surveys = s_i, years = 7, ages = ages) %>% 
+          layout(yaxis = list(range = ylim))
+      }
+      plot_list[[which(s$set_den == set_dens[i] & s$lengths_cap == lengths_caps[j])]] <- p
+    }
+  }
+  subplot(plot_list, nrows = length(set_dens), shareX = TRUE, shareY = TRUE, titleX = FALSE, titleY = FALSE)
+}
+
+p <- facet_fan(tests, lengths = 1:100, ylim = c(0, 21000000)) %>% 
+  layout(margin = list(r = 30)) %>% 
+  add_labels(x = c(0.1, 0.5, 0.9), y = rep(1, 3), yshift = 20, 
+             text = paste("D<sub>sets</sub> =", c(0.0005, 0.002, 0.01))) %>% 
+  add_labels(x = rep(1, 3), y = c(0.05, 0.5, 0.95), xshift = 30, textangle = 90,
+             text = paste("M<sub>lengths</sub> =", rev(c(10, 100, 1000)))) %>% 
+  add_labels(x = 0.5, y = 0, text = "Year", yshift = -40) %>% 
+  add_labels(x = 0, y = 0.5, text = "Abundance index", xshift = -60, textangle = -90)
+
+export_plot(p, file = "analysis/paper/figures/plot_length_strat_fan.png", width = 700, height = 600)
+
+```
+
+```{r fan2, fig.cap = "- Fanchart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+The heavy dark line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
+
+knitr::include_graphics("figures/plot_length_strat_fan.png")
+
+```
+
+```{r, eval = replot}
+
+p <- facet_fan(tests, ages = 1:10, ylim = c(0, 51000000)) %>% 
+  layout(margin = list(r = 30)) %>% 
+  add_labels(x = c(0.1, 0.5, 0.9), y = rep(1, 3), yshift = 20, 
+             text = paste("D<sub>sets</sub> =", c(0.0005, 0.002, 0.01))) %>% 
+  add_labels(x = rep(1, 3), y = c(0.05, 0.5, 0.95), xshift = 30, textangle = 90,
+             text = paste("M<sub>lengths</sub> =", rev(c(10, 100, 1000)))) %>% 
+  add_labels(x = 0.5, y = 0, text = "Year", yshift = -40) %>% 
+  add_labels(x = 0, y = 0.5, text = "Abundance index", xshift = -60, textangle = -90)
+
+export_plot(p, file = "analysis/paper/figures/plot_age_strat_fan.png", width = 700, height = 600)
+
+```
+
+```{r fan3, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
+The heavy dark line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
+
+knitr::include_graphics("figures/plot_age_strat_fan.png")
+
+```
 
@@
 
-```{r data, cache = TRUE}
-load("analysis/cod_sim_exports/2018-10-26_age_clust_test/test_output.RData")
-tests <- sim
-```
+- Make the results more general, almost about the flexibility of the package, then hone in on the case study??
 
 A dynamic population that varies across ages, time and space was simulated with the parameter settings used here (Table \@ref(tab:par)).
-Figure \@ref(fig:sp) depicts the aggregate distribution of the population in one year of the simulation and the corresponding catches of one stratified random survey over this population.
-Layered within this, of course, is the distributions of each age, as well as age and length sampling details.
-Though age-specific details are not depicted for simplicity, the general pattern of spatial aggregation apparent in Figure \@ref(fig:sp) is similar across all ages.
-That is, all ages tend to aggregate in fairly dense clusters and low-density zones are fairly wide ranging.
-The survey, therefore, tends to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:sp)).
+Because the default settings of the functions match the case study settings, Figure \@ref(fig:surface)a depicts abundance at age through time and Figure \@ref(fig:distribution)a depicts the spatial distribution of ages 1-3 through years 1-3.
+In general, the parameter settings used dictate that all ages tend to aggregate in fairly dense clusters and that low-density zones are relatively wide-ranging.
+Surveys, therefore, tend to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:survey)).
 The samples within each successful set are also dynamic and rarely include all age groups.
-This is because only moderate correlation was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations.
-This results in an intra-haul correlation of approximately `r sim_icc`, which, in this case, is of equal magnitude to the intra-haul correlation observed across more than 20 years of survey data of 3Ps cod.
+This is because only moderate correlation ($\varphi_{\xi, age}$ of 0.5) was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations.
+These settings were chosen to simulated data that roughly correspond to actual survey data of cod from NAFO division 3Ps, where cod between the ages of 1-5 tend to be caught at different locations.
 
@@
 Specifically, set density affects all estimates, length sampling affects estimates of abundance at length and abundance at age, and age sampling affects abundance at age estimates.
+This is because stratified estimates of total abundance is based on aggregate catch while abundance at length estimates requires the length frequency data and abundance at age estimates requires both length and age data to construct an age-length key.
+
 Figure \@ref(fig:tot) depicts the consequences of different set densities on the estimates of total abundance available to the survey through the simulated series.

### 2018-11-21 R/test_surveys.R

@@ -70,7 +70,7 @@ expand_surveys <- function(set_den = c(0.5, 1, 2, 5, 10) / 1000,
 #'                           lengths_cap = c(100, 500),
 #'                           ages_cap = c(5, 20))
 #'
-#' ## This call runs 200 simulations of 8 different surveys over the same
+#' ## This call runs 500 simulations of 8 different surveys over the same
 #' ## population, and then runs a stratified analysis and compares true vs
 #' ## estimated values. It may take a while to run.
 #' tests <- test_surveys(pop, surveys = surveys, keep_details = 1,

### 2018-11-21 analysis/paper/helper_functions.R

@@ -1,119 +1,4 @@
 
-## http://www.moeding.net/archives/32-Metric-prefixes-for-ggplot2-scales.html
-format_si <- function(...) {
-  # Format a vector of numeric values according
-  # to the International System of Units.
-  # http://en.wikipedia.org/wiki/SI_prefix
-  #
-  # Based on code by Ben Tupper
-  # https://stat.ethz.ch/pipermail/r-help/2012-January/299804.html
-  # Args:
-  #   ...: Args passed to format()
-  #
-  # Returns:
-  #   A function to format a vector of strings using
-  #   SI prefix notation
-  #
-
-  function(x) {
-    limits <- c(1e-24, 1e-21, 1e-18, 1e-15, 1e-12,
-                1e-9,  1e-6,  1e-3,  1e0,   1e3,
-                1e6,   1e9,   1e12,  1e15,  1e18,
-                1e21,  1e24)
-    prefix <- c("y",   "z",   "a",   "f",   "p",
-                "n",   "µ",   "m",   " ",   "k",
-                "M",   "G",   "T",   "P",   "E",
-                "Z",   "Y")
-
-    # Vector with array indices according to position in intervals
-    i <- findInterval(abs(x), limits)
-
-    # Set prefix to " " for very small values < 1e-24
-    i <- ifelse(i == 0, which(limits == 1e0), i)
-
-    paste0(format(round(x/limits[i], 1),
-                  trim = TRUE, scientific = FALSE, ...),
-           prefix[i])
-  }
-}
-si <- format_si()
-
-## little helper function for the fan plot
-fan_plot <- function(d, x = "Age", which_strat = "Age") {
-
-  if (which_strat == "Total") {
-    ints_by = c("year", "set_den")
-  }
-  if (which_strat == "Age") {
-    ints_by = c("year", "age", "set_den", "lengths_cap")
-  }
-  if (which_strat == "Length") {
-    ints_by = c("year", "length", "set_den", "lengths_cap")
-  }
-
-  ## quantiles for fan plot
-  quants <- seq(90, 10, by = -10)
-  ints <- lapply(quants, function(q) {
-    d[, list(prob = paste0(q, "%"),
-             lower = quantile(I_hat, prob = (1 - q / 100) / 2),
-             upper = quantile(I_hat, prob = 1 - (1 - q / 100) / 2)),
-      by = ints_by]
-  })
-  ints <- rbindlist(ints)
-  ints$prob <- factor(ints$prob, levels = paste0(quants, "%"))
-
[…]

### 2019-02-06 analysis/paper/sim_survey_paper.Rmd
Add divergent dot plot and error surface plot to paper (next step: fix docx template for code block and increase font of surface plot)

@@
 
+The relative performance of the surveys tested can be compared using `plot_survey_rank` and `plot_error_surface`.
+The `plot_survey_rank` function produces a divergent dot plot of the results which ranks the surveys by RMSE.
+Using the `wich_strat` argument, the plot can be focused on total, length or age based stratified results.
+The `plot_error_surface` focuses solely on the age based stratified results by plotting a surface of RMSE (z-axis) by set (drop down selection), length (y-axis) and age (z-axis) sampling effort.
+The sampling effort axes can be rule or sample size based (`plot_by = "rule"` or `plot_by = "samples"`, respectively).
+See figures \@ref(fig:rank) and \@ref(fig:rmse-surface) for examples of the output from the code below.
+
+```{r, echo = TRUE, eval = FALSE}
+
+plot_survey_rank(tests, which_strat = "length")
+plot_error_surface(tests, plot_by = "rule")
+
+```
+
+```{r, eval = replot}
+
+sim <- tests
+sim$surveys <- sim$surveys[sim$surveys$set_den %in% c("0.01", "0.002", "0.0005")]
+
+p <- plot_survey_rank(sim, which_strat = "length")
+export_plot(p, file = "analysis/paper/figures/plot_survey_rank.png", width = 450, height = 400)
+
+totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught),
+                                 n_measured = mean(n_measured), n_aged = mean(n_aged)),
+                          by = "survey"]
+errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
+d <- merge(errors, totals, by = "survey")
+
+split_d <- split(d, d$set_den)
+x <- sort(unique(d$ages_cap))
+y <- sort(unique(d$lengths_cap))
+
+scene <- list(
+  xaxis = list(title = "Ages cap"),
+  yaxis = list(title = "Lengths cap"),
+  zaxis = list(title = "RMSE"),
+  camera = list(eye = list(x = 1.5, y = 1.5, z = 1.5))
+)
+
+xshift <- c(-30, 0, 30)
+p_list <- lapply(seq_along(split_d), function(i) {
+  scene <- paste0("scene", i)
+  z <- stats::xtabs(RMSE ~ ages_cap + lengths_cap, data = split_d[[i]], subset = NULL)
+  plot_ly(x = x, y = y, scene = scene) %>% add_surface(z = t(z), showscale = FALSE) %>% 
+    add_labels(x = 0.5, y = 1, text = paste("D<sub>sets</sub> =", names(split_d[i])), 
+               yshift = -20, xshift = xshift[i])
+})
+
+p <- subplot(p_list) %>%  
+  layout(scene = c(list(domain = list(x = c(0, 1/3), y = c(0, 1))), scene),
+         scene2 = c(list(domain = list(x = c(1/3, 2/3), y = c(0, 1))), scene),
+         scene3 = c(list(domain = list(x = c(2/3, 1), y = c(0, 1))), scene),
+         autosize = FALSE, margin = list(t = 0, l = 10, r = 10, b = 10))
+
+export_plot(p, file = "analysis/paper/figures/plot_error_surface.png", width = 1250, height = 400)
+
+```
+
+```{r rank, fig.cap = "- Divergent dot plot of the percision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{sets}$] and length measurements [$N_{measured}$]), under various sampling protocol (set density [$D_{sets}$] and maximum number of lengths measured per set [$M_{lengths}$])."}
+
+knitr::include_graphics("figures/plot_survey_rank.png")
+
+```
+
+```{r rmse-surface, fig.cap = "- Surface plots of RMSE from an array of surveys with different sampling protocol.
+Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group (M_{ages}).
+This plot is a facet of plots produced by `plot_error_surface` when supplied results from `test_surveys`.
+Note that the z-axes scales are different across the facets."}
+
+knitr::include_graphics("figures/plot_error_surface.png")
+
+```
+
 ## Assumptions

### 2019-02-07 analysis/paper/sim_survey_paper.Rmd
Improve font size of error_surface figure and improve docx template

@@
 
-export_plot(p, file = "analysis/paper/figures/plot_error_surface.png", width = 1250, height = 400)
+export_plot(p, file = "analysis/paper/figures/plot_error_surface.png", width = 990, height = 270)
 
@@
 ```{r rmse-surface, fig.cap = "- Surface plots of RMSE from an array of surveys with different sampling protocol.
-Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group (M_{ages}).
+Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$).
 This plot is a facet of plots produced by `plot_error_surface` when supplied results from `test_surveys`.

### 2019-02-08 analysis/presentation/sim_survey_presentation.Rmd
Tweak presentation for MUN Biology seminar

@@
 author: "Paul M. Regular & Fran Mowbray"
-date: "September 12^th^, 2018"
+date: "2019-02-08"
 output:
@@
 knitr::opts_knit$set(root.dir = "../..")
+
 ```
@@
 ```{r data, include = FALSE}
+
+# xaringan::inf_mr()
+
 library(SimSurvey)
@@
 library(data.table)
-load("analysis/cod_sim_exports/2018-09-04_test/test_output.RData")
+load("analysis/cod_sim_exports/2018-10-26_age_clust_test/test_output.RData")
 
@@
 
+# N?
+
 ## Fisheries-independent trawl surveys {.columns-2}
@@
 - Becoming increasingly important in stock assessment
-- Very costly 💰
+- Very costly!
 - General goal:
@@
 
-<dev class = "left">
 ```{r}
@@
 ```
-</dev>
 
@@
 
-<dev class = "left">
 ```{r}
-plot_surface(sim) %>% tight_layout()
+plot_surface(sim) %>% tight_layout(scene = list(xaxis = list(range = c(1, 10))))
 ```
-</dev>
 
@@
 
-### Square grid with depth gradient
-
-<dev class = "left">
-```{r}
-plot_ly(data = sim$grid_xy, x = ~x, y = ~y, color = ~depth, 
-        symbol = I(15), size = I(4.5), name = "grid", type = "scatter", 
-        text = ~depth) %>%
-  colorbar(title = "Depth") %>%
-  tight_layout(yaxis = list(title = "", scaleanchor = "x", zeroline = FALSE, 
-                            showgrid = FALSE, showticklabels = FALSE,
-                            tickcolor = "transparent"),
-               xaxis = list(title = "", zeroline = FALSE, 
-                            showgrid = FALSE, showticklabels = FALSE,
-                            tickcolor = "transparent"))
-```
-</dev>
-
-## Simulate spatial aggregation {.tighter}
-
 ### Distribution for ages 1-6 in years 1-6
 
-<dev class = "left">
 ```{r}
-plot_distribution_slider(sim, ages = 1:6, years = 1:6) %>% 
+plot_distribution(sim, ages = 1:6, years = 1:6, type = "heatmap") %>% 
   tight_layout(margin = list(l = 0, r = 0, t = 40, b = 0, pad = 0))
 ```
-</dev>
 
@@
 
-<dev class = "left">
 ```{r}
-plot_samp_dist(sim) %>% 
+plot_survey(sim) %>% 
   tight_layout(margin = list(l = 0, r = 0, t = 0, b = 40, pad = 0))
 ```
-</dev>
 
@@
 
-## Code {.smaller}
-
-### Population
-
-```{r, echo = TRUE, eval = FALSE}
-
-      set.seed(438)
-      pop <- sim_abundance(ages = 1:20, years = 1:20,
-                           R = sim_R(mean = 30000000, log_sd = 0.5,
-                                     random_walk = TRUE),
-                           Z = sim_Z(mean = 0.5, log_sd = 0.2,
-                                     phi_age = 0.9, phi_year = 0.5)) %>%
-        sim_distribution(grid = sim_grid(x_range = c(-140, 140), y_range = c(-140, 140),
-                                         res = c(3.5, 3.5), shelf_depth = 200,
-                                         shelf_width = 100, depth_range = c(0, 1000),
-                                         n_div = 1, strat_breaks = seq(0, 1000, by = 20),
-                                         strat_splits = 2),
-                         ays_covar = sim_ays_covar(sd = 2.8, range = 300,
-                                                   phi_age = 0.5, phi_year = 0.9,
-                                                   group_ages = 5:20),
-                         depth_par = sim_parabola(mu = 200, sigma = 70))
-
-```
-
-## Code {.smaller}
-
-### Surveys
-
-```{r, echo = TRUE, eval = FALSE}
-
-      surveys <- expand_surveys(set_den = c(0.5, 1, 2, 5, 10) / 1000,
-                                lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
-                                ages_cap = c(2, 5, 10, 20, 50))
-      sim <- test_surveys(pop,
-                          surveys = surveys, n_sims = 5,
-                          n_loops = 200, cores = 6,
-                          q = sim_logistic(k = 2, x0 = 3),
-                          growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1, digits = 0),
-                          export = "analysis/cod_sim_exports")
-
-```
-
-<!-- There is a conflict somewhere...had to contain plotly plots in a dev to get it working and I had to add a bunch of spaces to fudge centering in one of the tables -->
-
 ## References {.smaller}

### 2019-02-12 analysis/paper/sim_survey_paper.Rmd
Draft new first paragraph of results section

@@
 The package also includes several plotting function for making quick plotly-based [@sievert2018] interactive visuals of the simulated population.
-For instance, the `plot_surface` function can be used to make quick visuals of matricies contained within the list returned by `sim_abundance`.
+For instance, the `plot_surface` function can be used to make quick visuals of matrices contained within the list returned by `sim_abundance`.
 The plots output from the below code are combined into Figure \@ref(fig:surface).
-Here we display the abundance at age matrix (object named `N` in the list produceb by `sim_abundance`); other names can be supplied to the `mat` argument to visualize a different matrix from the `sim_abundance` list, such as `Z`.
+Here we display the abundance at age matrix (object named `N` in the list produced by `sim_abundance`); other names can be supplied to the `mat` argument to visualize a different matrix from the `sim_abundance` list, such as `Z`.
 
@@
 Like other core functions, some convenience functions are included in SimSurvey for creating interactive plots of the results from `test_surveys`.
-For instance a series of plotting functions ending in `_fan` produces fancharts where stratified estimates of abundance from each simulated survey are converted into a series of quantiles to depict the probability that estimates fall within a praticular range.
+For instance a series of plotting functions ending in `_fan` produces fan charts where stratified estimates of abundance from each simulated survey are converted into a series of quantiles to depict the probability that estimates fall within a particular range.
 True values of abundance available to the survey are overlaid on the series of probability envelopes.
-These plots help visually assess the level of percision and bias from a specific set of survey protocol.
-The three lines of code below will produce interactive fancharts for stratified estimates of total abundance, abundance at length and abundance at age, respectively.
+These plots help visually assess the level of precision and bias from a specific set of survey protocol.
+The three lines of code below will produce interactive fan charts for stratified estimates of total abundance, abundance at length and abundance at age, respectively.
 A subset of these plots have been combined into static plots (Figure \@ref(fig:fan1), \@ref(fig:fan2), \@ref(fig:fan3)).
@@
 
-```{r fan1, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
+```{r fan1, fig.cap = "- Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
 The heavy dark line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
@@
 
-```{r fan2, fig.cap = "- Fanchart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+```{r fan2, fig.cap = "- Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 The heavy dark line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
@@
 
-```{r fan3, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+```{r fan3, fig.cap = "- Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
@@
 # Results
+
+The SimSurvey package serves as a tool for simulating surveys of dynamic populations that vary across ages, time and space.
+The core of the simulation is based on the widely used cohort equation and, even though the processes that define recruitment and total mortality are simple, a wide range of stock dynamics can be simulated by changing a few parameters (Figure \@ref(fig:surface)).
+This base population can then be distributed through a grid (Figure \@ref(fig:grid)) and relationships with depth can be defined as can the nature of the correlation across ages, years and space (Figure \@ref(fig:distribution)).
+Together, two functions (`sim_abundance` and `sim_distribution`) are capable of simulating a wide range of populations with different life histories and spatial properties.
+The next necessary step to generating data similar to actual observations is to conduct a survey.
+In this package we implement a function, `sim_survey`, that conducts a stratified random survey of the population.
+The sampling process is governed by the area covered by the trawl as well as age-specific catchability.
+Sub-sampling protocol (length and age sampling) can also be varied.
+As such, data from a wide range of surveys can be simulated (Figure \@ref(fig:survey)).
 
@@
 
-<!-- ```{r tot, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
+<!-- ```{r tot, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
 The heavy dark line indicates the true tred in the total population available to the survey."} -->
@@
 
-<!-- ```{r length, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+<!-- ```{r length, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 The heavy dark line indicates the true population available to the survey."} -->
@@
 
-<!-- ```{r age, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+<!-- ```{r age, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fan chart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
@@
 
-<!-- ```{r year, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+<!-- ```{r year, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.

### 2019-02-12 R/plot_sim.R

@@ -681,7 +681,8 @@ plot_survey_rank <- function(sim, which_strat = "age") {
 
   suppressWarnings({
     subplot(a, b, titleX = TRUE, shareY = TRUE, nrows = 1, widths = c(0.3, 0.7),
-            margin = 0.01)
+            margin = 0.01) %>%
+      layout(margin = list(b = 10, l = 10))
   })
 
 }

### 2019-03-07 analysis/paper/sim_survey_paper.Rmd
Complete draft of new results section

@@
 
-The SimSurvey package serves as a tool for simulating surveys of dynamic populations that vary across ages, time and space.
+The SimSurvey package serves as a tool for simulating stratified random surveys of dynamic populations that vary across ages, time and space.
 The core of the simulation is based on the widely used cohort equation and, even though the processes that define recruitment and total mortality are simple, a wide range of stock dynamics can be simulated by changing a few parameters (Figure \@ref(fig:surface)).
 This base population can then be distributed through a grid (Figure \@ref(fig:grid)) and relationships with depth can be defined as can the nature of the correlation across ages, years and space (Figure \@ref(fig:distribution)).
-Together, two functions (`sim_abundance` and `sim_distribution`) are capable of simulating a wide range of populations with different life histories and spatial properties.
+Together, two functions (`sim_abundance` and `sim_distribution`) are capable of simulating a wide range of populations with different life histories, depth associations and spatial properties.
 The next necessary step to generating data similar to actual observations is to conduct a survey.
@@
 As such, data from a wide range of surveys can be simulated (Figure \@ref(fig:survey)).
-
-- Make the results more general, almost about the flexibility of the package, then hone in on the case study??
-
-A dynamic population that varies across ages, time and space was simulated with the parameter settings used here (Table \@ref(tab:par)).
-Because the default settings of the functions match the case study settings, Figure \@ref(fig:surface)a depicts abundance at age through time and Figure \@ref(fig:distribution)a depicts the spatial distribution of ages 1-3 through years 1-3.
-In general, the parameter settings used dictate that all ages tend to aggregate in fairly dense clusters and that low-density zones are relatively wide-ranging.
+There are a plethora of statistical models that may be tested using these simulated data, however, implementing a variety of analylitical approaches was outside the scope of this work.
+Insted we focus on analyzing simulated stratified-random survey data using a design-based stratified analysis.
+A stratified analyis is facilitated using the `run_strat` function and the precision and accuracy of the results (e.g.
+RMSE) can be calculated using `strat_error`.
+This is a simple and widly-used analysis, and the speed at which it runs allows for a wide range of survey designs to be tested, via the `test_surveys` function, in a reasonable timeframe.
+
+Results from the case study presented provide one example of how this package can be used to explore the effacacy of various survey designs.
+A dynamic population with a patchy and age-clustered distribution was simulated for this case study (Figure \@ref(fig:surface)a, \ref(fig:distribution)a).
+In general, the parameter settings used (function defaults) dictate that all ages tend to aggregate in fairly dense clusters and that low-density zones are relatively wide-ranging.
 Surveys, therefore, tend to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:survey)).
@@
 
-The stratified analyses run on these simulated data provided estimates of total abundance, abundance at length and abundance at age, and different stages of sampling affect these estimates.
+A series of surveys were run over the simulated population (see `expand_surveys` call in Table \@ref(tab:par)) and stratified analyses were run on the data obtained.
+The stratified analyses provided estimates of total abundance, abundance at length and abundance at age, and different stages of sampling affect these estimates.
 Specifically, set density affects all estimates, length sampling affects estimates of abundance at length and abundance at age, and age sampling affects abundance at age estimates.
-This is because stratified estimates of total abundance is based on aggregate catch while abundance at length estimates requires the length frequency data and abundance at age estimates requires both length and age data to construct an age-length key.
-
-Figure \@ref(fig:tot) depicts the consequences of different set densities on the estimates of total abundance available to the survey through the simulated series.
-It is clear from this figure that the probability envelopes are tighter (i.e.
-estimates are more precise) and bias is negligible.
-Likewise, abundance at length and abundance at age estimates are more precise as set density is increased (Figure \@ref(fig:length), \@ref(fig:age), \@ref(fig:year)).
-With the exception of a length sampling protocol of 10 measurements per set ($M_{lengths}$ = 10), there appears to be little difference in the estimates of abundance at length when the number of measurements are capped at 100 ($M_{lengths}$ = 100) or 1000 ($M_{lengths}$ = 1000; Figure \@ref(fig:length)).
-Some downward bias is apparent in the abundance at length estimates, especially at larger lengths, under the low to medium set density scenarios with low length sampling effort ($D_{sets}$ = 0.0005 or 0.002 and $M_{lengths}$ = 10; Figure \@ref(fig:length)).
-Upward or downward bias in abundance at age estimates is apparent in all cases except the high set density and high length sampling scenario ($D_{sets}$ = 0.01 and $M_{lengths}$ = 1000; Figure \@ref(fig:age), \@ref(fig:year)).
-Under both low and medium set densities ($D_{sets}$ = 0.0005 and 0.02, respectively), increasing length sampling effort from 10 or 100 measurements per set to 1000 appears to increase bias or reduce precision slightly (Figure \@ref(fig:age), \@ref(fig:year)).
-
-For a less visual approach to assessing precision and bias, RMSE results are presented in Figures \@ref(fig:error-lines) and \@ref(fig:error-wireframe) and Table \@ref(tab:comp).
-Figure \@ref(fig:error-lines) depicts the effects of set density and length sampling effort on abundance at length and abundance at age estimates.
-In all cases, increases in set density results in large declines in RMSE (Figure \@ref(fig:error-lines)).
-For abundance at length estimates, RMSE declines as the number of lengths measured increases, however, gains plateau beyond a cretin point.
-Likewise, RMSE of abundance at age estimates decline with increased length sampling effort, but RMSE does not plateau in all set density scenarios (Figure \@ref(fig:error-lines)).
-While RMSE continues to decline with increased length sampling effort under the high set density scenarios ($D_{sets}$ = 0.005 to 0.01 sets / km^2^), RMSE reaches a plateau and then increases with increased length sampling effort under the medium to low set density scenarios ($D_{sets}$ = 0.002 to 0.0005 sets / km^2^; Figure \@ref(fig:error-lines)).
-In many cases, RMSE for both metrics is lower for scenarios with higher set density but lower length sampling effort (Figure \@ref(fig:error-lines)).
-Similar patterns are apparent in Figure \@ref(fig:error-wireframe) and Table \@ref(tab:comp), which focus on the consequences of set density and length and age sampling protocol on the estimates of abundance at age.
-Again, RMSE clearly declines with increases to set density and sometimes declines with increased length sampling effort (Figure \@ref(fig:error-wireframe)).
-Under both low and medium set density scenarios ($D_{sets}$ = 0.005 to 0.01 sets / km^2^), measuring a maximum of 100 fish per set appears to be better than measuring a maximum of 1000 fish per set (Figure \@ref(fig:error-wireframe); Table \@ref(tab:comp)).
-Decreasing the age sampling protocol below 10 ages sampled per set per length group appears to result in relatively small increases to RMSE (Figure \@ref(fig:error-wireframe); Table \@ref(tab:comp)).
+This is because stratified estimates of total abundance is based on aggregate catch while abundance at length estimates requires both total and length frequency data; finally, abundance at age estimates requires set totals, and length and age data to construct an age-length key.
+
+The effect of set density on estimates of total abundance, abundance at length and abundance at age are clear across all figures generated from the `test_surveys` results (Figures \@ref(fig:fan1), \@ref(fig:fan2), \@ref(fig:fan3), \@ref(fig:rank), \@ref(fig:rmse-surface)).
+Across all fan plots, it is clear that the probability envelopes tighten (i.e.
+estimates are more precise) as set density is increased (Figures \@ref(fig:fan1), \@ref(fig:fan2), \@ref(fig:fan3)).
+Clear declines in RMSE are also apparant in estimates of abundance at length as set density is increased (Figure \@ref(fig:rank)) and there are notable differences in the scale of RMSE in estimates of abundance at age as set density is increased (Figure \@ref(fig:rmse-surface)).
+
+Compared to increases in set density, the effects of increased length sampling effort on abundance at length estimates are less clear.
+Regarding estimates of abundance at length, RMSE appears to reach a platau at when length sampling effort is increased 100 measurements per set (Figure \@ref(fig:rank)).
+Increasing the sampling rule above 100 measurements per set results in many more fish being measured, however, the increase in sampling effort is not matched with substantive declines in RMSE (Figure \@ref(fig:rank)).
+Interestingly, it appears that measuring fewer total fish at more locations appears to be more benificial than many fish at fewer locations (Figure \@ref(fig:rank)).
+
+All levels of sampling affect the abundance at age estimates as set catches define the maginitude of the population, length sampling define the length distribution and data from the length-stratified age sampling is used to construct an age-length-key to convert length frequenceis to age frequencies.
+Like the abundance at length estimates, the greatest improvements to RMSE come from increasing set density rather than sub-sampling effort (Figure \@ref(fig:rmse-surface)).
+Specifically, decreasing the age sampling protocol below 10 ages sampled per length group per division appears to result in relatively small increases to RMSE (Figure \@ref(fig:rmse-surface)).
+Lenght sampling effort, in contrast, has an uneven impact depending on the set density scenario.
+At low set densities ($D_{sets}$ = 0.0005 sets / km^2^), RMSE declines when length sampling effort is increased from around 5 to around 100 measurements per set, and RMSE starts to increase as length sampling effort is increased beyond ~100 measurements per set; in fact, RMSE values appear higher at the highest length sampling scenario (1000 measurements / set) than the lowest (5 measurements / set; Figure \@ref(fig:rmse-surface)).
+A similar pattern is apparant under the medium set density scenario ($D_{sets}$ = 0.002 sets / km^2^), however, RMSE values under the lowest and highest length sampling scenarios are of similar magnitude.
+Finally, RMSE continues to decline with increased length sampling effort under the high set density scenarios ($D_{sets}$ = 0.01 sets / km^2^; Figure \@ref(fig:rmse-surface)).
 

### 2019-03-07 R/plot_sim.R
Use base::length insted of dplyr::n inside the plot_survey function

@@ -210,7 +210,7 @@ plot_survey <- function(sim, which_year = 1, which_sim = 1) {
     group_by(set) %>%
     filter(!is.na(measured)) %>%
     slice(rep(1:n(), each = 2)) %>%
-    mutate(lab = rep(c("caught", "sampled"), times = n() / 2))
+    mutate(lab = rep(c("caught", "sampled"), times = length(measured) / 2))
 
   lf_p <- hist_base %>%
     filter(measured & lab == "sampled" | lab == "caught") %>%

### 2019-03-07 R/plot_sim.R
Missed one use of dplyr::n; replaced with base::length

@@ -209,7 +209,7 @@ plot_survey <- function(sim, which_year = 1, which_sim = 1) {
     mutate(length = group_lengths(length, length_group)) %>%
     group_by(set) %>%
     filter(!is.na(measured)) %>%
-    slice(rep(1:n(), each = 2)) %>%
+    slice(rep(1:length(measured), each = 2)) %>%
     mutate(lab = rep(c("caught", "sampled"), times = length(measured) / 2))
 
   lf_p <- hist_base %>%

### 2019-05-21 analysis/paper/sim_survey_paper.Rmd
Start revising the discussion

@@
 Once the survey data are simulated, they then need to be analyzed to obtain an index of abundance.
-While there are many model-based options for obtaining an index (**TODO: add references**), survey-based approaches, such as stratified analyses, are often used.
+While there are many model-based options for obtaining an index [e.g.
+@Thorson2015], survey-based approaches, such as stratified analyses, are often used.
 Here we apply formula presented in Smith and Somerton [-@Smith1981] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
@@
 # Discussion
+
+While there are several simulation frameworks for testing the design of fisheries-independent trawl surveys [e.g.
+@puerta2019; @schnute2003], there are few tools that facilitate such work.
+Here we document an R package with a series of functions for testing the multi-layred sampling design of spatial, age-structured populations.
+The package can also serve as a medium for the simulation testing of methods for analyzing such data.
+"Good survey design is a crucial prerequisite for obtaining reliable results" [@thomas2010]...
 

### 2019-05-24 analysis/paper/sim_survey_paper.Rmd
Structural changes to the paper

@@
 ---
-title: Using simulation to optimize the sampling design of fisheries-independent trawl surveys
+title: "`SimSurvey`: an `R` package for simulation testing the design and analysis of fisheries-independent trawl surveys"
 author: "Paul M. Regular, Fran Mowbray, et al.?"
@@
 
+- As with any population monitoring program, accurate and reliable results from a fisheries-independent trawl survey depends on good survey design.
+Ideally, the sampling design of large and costly surveys should be simulation tested before embarking on an expensive field program.
+
+- While there are several simulation frameworks for testing the design of fisheries-independent trawl surveys [e.g.
+@puerta2019; @schnute2003], there are few tools that facilitate such work.
+Here we document an R package with a series of functions for testing the multi-layered sampling design of spatial, age-structured populations.
+The package can also serve as a medium for the simulation testing of methods for analyzing such data.
+
 Fisheries-independent trawl surveys have become an increasingly important research tool for the management of dynamic fish stocks.
@@
 
-# Methods
-
-The SimSurvey package was written in the programming language R [@R] and, in short, the package allows for the simulation of stratified-random surveys of an age-structured population that varies in space and time.
-SimSurvey relies heavily on functions from the data.table [@dowle2017], raster [@hijmans2016] and plotly [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively.
-All the source R code behind SimSurvey is available on GitHub (https://github.com/PaulRegular/SimSurvey) and the 0.1.0 release of the package can be installed via GitHub:
+# The **`SimSurvey`** package
+
+The **`SimSurvey`** package was written in the programming language R [@R] and, in short, the package allows for the simulation of stratified-random surveys of an age-structured population that varies in space and time.
+**`SimSurvey`** relies heavily on functions from the data.table [@dowle2017], raster [@hijmans2016] and plotly [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively.
+All the source R code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey) and the 0.1.0 release of the package can be installed via GitHub:
 
@@
 
-The construction of the SimSurvey package was motivated by a regional need for assessing the sampling design of trawl surveys conducted along the Newfoundland and Labrador shelf by Fisheries and Oceans Canada.
+The construction of the **`SimSurvey`** package was motivated by a regional need for assessing the sampling design of trawl surveys conducted along the Newfoundland and Labrador shelf by Fisheries and Oceans Canada.
 As a tangible first step, we focused our efforts on emulating data from the survey of cod in NAFO division 3Ps since it is one of the most dynamic and heavily sampled stocks in the region.
-Below we outline the equations behind a series of SimSurvey functions that were used to simulate survey data with similar features to data obtained by the survey of 3Ps cod.
+Below we outline the equations behind a series of **`SimSurvey`** functions that were used to simulate survey data with similar features to data obtained by the survey of 3Ps cod.
 The settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data; these values were set as the defaults for the main functions in the package.
@@
 See Table \@ref(tab:par) for a description of the code used to generate the case study results.
+Text below provides more detail on each function and also demonstrates a series of functions for visualizing the results.
+See Appendix 1 for an interpretation and discussion of the case study results.
 
@@
 Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid.
-A more detailed description of the space-age-year covariance is included in Appendix 1.
+A more detailed description of the space-age-year covariance is included in Appendix 2.
 Note that the the distributions of specific length groups were not simulated because of the greater computational cost of storing spatially dis-aggregated abundance across all length groups.
@@
 Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete.
-Also note that the key functions in the SimSurvey package have been set-up to be pipe [`%>%`; @bache2014] friendly.
+Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe [`%>%`; @bache2014] friendly.
 
@@
 Objects produced by `sim_survey` (set and sampling details) and `run_strat` (full stratified analysis results) are not retained to minimize object size.
-Like other core functions, some convenience functions are included in SimSurvey for creating interactive plots of the results from `test_surveys`.
+Like other core functions, some convenience functions are included in **`SimSurvey`** for creating interactive plots of the results from `test_surveys`.
 For instance a series of plotting functions ending in `_fan` produces fan charts where stratified estimates of abundance from each simulated survey are converted into a series of quantiles to depict the probability that estimates fall within a particular range.
@@
 This is a short, non-inclusive, list of additional variability that likely occurs in the real-world but was not implemented here for simplicity.
-Nevertheless, the SimSurvey package provides the most realistic operating model that we are aware of for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
-
-# Results
-
-The SimSurvey package serves as a tool for simulating stratified random surveys of dynamic populations that vary across ages, time and space.
+Nevertheless, the **`SimSurvey`** package provides the most realistic operating model that we are aware of for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
+
+# Research opportunities
+
+- Catchability
+- Stratified vs. random age sampling
+- Model testing
+
+# Future directions
+
+- Cost-benefit
+
+# Summary
+
+The **`SimSurvey`** package serves as a tool for simulating stratified random surveys of dynamic populations that vary across ages, time and space.
 The core of the simulation is based on the widely used cohort equation and, even though the processes that define recruitment and total mortality are simple, a wide range of stock dynamics can be simulated by changing a few parameters (Figure \@ref(fig:surface)).
@@
 As such, data from a wide range of surveys can be simulated (Figure \@ref(fig:survey)).
-There are a plethora of statistical models that may be tested using these simulated data, however, implementing a variety of analylitical approaches was outside the scope of this work.
-Insted we focus on analyzing simulated stratified-random survey data using a design-based stratified analysis.
-A stratified analyis is facilitated using the `run_strat` function and the precision and accuracy of the results (e.g.
+There are a plethora of statistical models that may be tested using these simulated data, however, implementing a variety of analytical approaches was outside the scope of this work.
+Instead we focus on analyzing simulated stratified-random survey data using a design-based stratified analysis.
+A stratified analysis is facilitated using the `run_strat` function and the precision and accuracy of the results (e.g.
 RMSE) can be calculated using `strat_error`.
-This is a simple and widly-used analysis, and the speed at which it runs allows for a wide range of survey designs to be tested, via the `test_surveys` function, in a reasonable timeframe.
-
-Results from the case study presented provide one example of how this package can be used to explore the effacacy of various survey designs.
-A dynamic population with a patchy and age-clustered distribution was simulated for this case study (Figure \@ref(fig:surface)a, \ref(fig:distribution)a).
+This is a simple and widely-used analysis, and the speed at which it runs allows for a wide range of survey designs to be tested, via the `test_surveys` function, in a reasonable time-frame.
+
+# Acknowledgements
+
+- Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
+- Dave Cote for a review of a previous version of this manuscript
+- NSERC visiting-fellow program
+
+# Appendix 1: case study results and discussion
+
+## Results
+
+Results from the case study presented provide one example of how this package can be used to explore the efficacy of various survey designs.
+A dynamic population with a patchy and age-clustered distribution was simulated for this case study (Figure \@ref(fig:surface)a, \@ref(fig:distribution)a).
 In general, the parameter settings used (function defaults) dictate that all ages tend to aggregate in fairly dense clusters and that low-density zones are relatively wide-ranging.
@@
 estimates are more precise) as set density is increased (Figures \@ref(fig:fan1), \@ref(fig:fan2), \@ref(fig:fan3)).
-Clear declines in RMSE are also apparant in estimates of abundance at length as set density is increased (Figure \@ref(fig:rank)) and there are notable differences in the scale of RMSE in estimates of abundance at age as set density is increased (Figure \@ref(fig:rmse-surface)).
+Clear declines in RMSE are also apparent in estimates of abundance at length as set density is increased (Figure \@ref(fig:rank)) and there are notable differences in the scale of RMSE in estimates of abundance at age as set density is increased (Figure \@ref(fig:rmse-surface)).
 
 Compared to increases in set density, the effects of increased length sampling effort on abundance at length estimates are less clear.
-Regarding estimates of abundance at length, RMSE appears to reach a platau at when length sampling effort is increased 100 measurements per set (Figure \@ref(fig:rank)).
+Regarding estimates of abundance at length, RMSE appears to reach a plateau at when length sampling effort is increased 100 measurements per set (Figure \@ref(fig:rank)).
 Increasing the sampling rule above 100 measurements per set results in many more fish being measured, however, the increase in sampling effort is not matched with substantive declines in RMSE (Figure \@ref(fig:rank)).
-Interestingly, it appears that measuring fewer total fish at more locations appears to be more benificial than many fish at fewer locations (Figure \@ref(fig:rank)).
-
-All levels of sampling affect the abundance at age estimates as set catches define the maginitude of the population, length sampling define the length distribution and data from the length-stratified age sampling is used to construct an age-length-key to convert length frequenceis to age frequencies.
+Interestingly, it appears that measuring fewer total fish at more locations appears to be more beneficial than many fish at fewer locations (Figure \@ref(fig:rank)).
+
+All levels of sampling affect the abundance at age estimates as set catches define the magnitude of the population, length sampling define the length distribution and data from the length-stratified age sampling is used to construct an age-length-key to convert length frequencies to age frequencies.
 Like the abundance at length estimates, the greatest improvements to RMSE come from increasing set density rather than sub-sampling effort (Figure \@ref(fig:rmse-surface)).
 Specifically, decreasing the age sampling protocol below 10 ages sampled per length group per division appears to result in relatively small increases to RMSE (Figure \@ref(fig:rmse-surface)).
-Lenght sampling effort, in contrast, has an uneven impact depending on the set density scenario.
+Length sampling effort, in contrast, has an uneven impact depending on the set density scenario.
 At low set densities ($D_{sets}$ = 0.0005 sets / km^2^), RMSE declines when length sampling effort is increased from around 5 to around 100 measurements per set, and RMSE starts to increase as length sampling effort is increased beyond ~100 measurements per set; in fact, RMSE values appear higher at the highest length sampling scenario (1000 measurements / set) than the lowest (5 measurements / set; Figure \@ref(fig:rmse-surface)).
-A similar pattern is apparant under the medium set density scenario ($D_{sets}$ = 0.002 sets / km^2^), however, RMSE values under the lowest and highest length sampling scenarios are of similar magnitude.
+A similar pattern is apparent under the medium set density scenario ($D_{sets}$ = 0.002 sets / km^2^), however, RMSE values under the lowest and highest length sampling scenarios are of similar magnitude.
 Finally, RMSE continues to decline with increased length sampling effort under the high set density scenarios ($D_{sets}$ = 0.01 sets / km^2^; Figure \@ref(fig:rmse-surface)).
 
-# Discussion
-
-While there are several simulation frameworks for testing the design of fisheries-independent trawl surveys [e.g.
-@puerta2019; @schnute2003], there are few tools that facilitate such work.
-Here we document an R package with a series of functions for testing the multi-layred sampling design of spatial, age-structured populations.
-The package can also serve as a medium for the simulation testing of methods for analyzing such data.
-"Good survey design is a crucial prerequisite for obtaining reliable results" [@thomas2010]...
-
-The simple simulation presented here revealed some expected and unexpected patterns.
+## Discussion
+
+The simple case study presented here revealed some expected and unexpected patterns.
 First, there were clear improvements to precision in all population estimates as the number of sets were increased.
@@
 
-# Acknowledgements
-
-- Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
-- Dave Cote for a review of a previous version of this manuscript
-- NSERC visiting-fellow program
-
-# Tables
-
-<!-- # Figures -->
-
-<!-- ```{r sp, cache = TRUE, fig.height = 3.8, fig.width = 8.5, fig.cap = "- An example from year seven of the simulation of a) the distribution of total abundance available to the survey, and b) locations of the sets conducted and total numbers caught in a stratified random survey with a set density of 0.002 sets / km^2^.
-Grey lines indicate depth-based survey strata."} -->
-
-<!-- which_year <- 7 -->
-<!-- which_sim <- 1 -->
-
-<!-- sp_N_tot <- sim$sp_I[year == which_year & sim == which_sim, list(N = sum(I)), by = c("year", "cell")] -->
-<!-- xyz <- merge(sim$grid_xy, sp_N_tot, by = "cell") -->
-<!-- size <- unique(diff(pretty(xyz$N, 25))) -->
-
-<!-- sp_strat <- raster::rasterToPolygons(sim$grid$strat, dissolve = TRUE) -->
-<!-- df_strat <- suppressWarnings(fortify(sp_strat) %>% group_by(group)) -->
-
-<!-- r <- raster::rasterFromXYZ(xyz[, c("x", "y", "N")]) -->
-<!-- l <- rasterToContour(r, levels = pretty(xyz$N, 20)) -->
-<!-- df_lines <- suppressWarnings(fortify(l)) -->
-<!-- levels <- as.numeric(as.character(l$level)) -->
-<!-- levels <- si(levels) -->
-<!-- labs <- levels -->
-<!-- labs[seq_along(labs) %% 2 == 0] <- "" -->
-<!-- levels <- factor(levels, levels = levels) -->
-<!-- names(levels) <- rownames(as.data.frame(l)) -->
-<!-- df_lines$level <- levels[df_lines$id] -->
-
-<!-- abun_plot <- ggplot() + -->
-<!--   geom_path(data = df_strat, aes(x = long, y = lat, group = group),  -->
-<!--             color = "grey", size = 0.05) + -->
-<!--   geom_path(data = df_lines, aes(x = long, y = lat, group = group, color = level), -->
-<!--             size = 0.3) + -->
-<!--   scale_color_viridis(labels = labs, discrete = TRUE, name = "N") +  -->
-<!--   scale_x_continuous(expand = c(0,0)) + -->
-<!--   scale_y_continuous(expand = c(0,0)) + -->
-<!--   coord_equal() + -->
-<!--   theme_void() + guides(color = guide_legend(keywidth = 0.4, keyheight = 0,  -->
-<!--                                              default.unit = "cm", ncol = 1)) + -->
-<!--   theme(legend.justification = c(0, 1), -->
-<!--         plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) -->
-
-<!-- setdet <- sim$setdet -->
-<!-- setdet <- setdet[setdet$year == which_year & setdet$sim == which_sim, ] -->
-<!-- breaks <- pretty(setdet$n, 7) -->
-
-<!-- set_plot <- ggplot() + -->
-<!--   geom_path(data = df_strat, aes(x = long, y = lat, group = group),  -->
-<!--             color = "grey", size = 0.05) + -->
-<!--   geom_point(data = setdet, aes(x = x, y = y, size = n, color = n), alpha = 0.8) + -->
-<!--   scale_size_area(name = "   n", max_size = 10, limits = range(breaks), breaks = breaks) + -->
-<!--   scale_color_viridis(name = "   n", limits = range(breaks), breaks = breaks) +  -->
-<!--   scale_x_continuous(expand = c(0,0)) + -->
-<!--   scale_y_continuous(expand = c(0,0)) + -->
-<!--   coord_equal() + -->
-<!--   theme_void() + guides(color = guide_legend(keywidth = 0.4, keyheight = 0.2, default.unit = "cm")) + -->
-<!--   theme(legend.justification = c(0, 1), -->
-<!--         plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm")) -->
-
-<!-- p <- cowplot::plot_grid(abun_plot, set_plot, labels = c("a)", "b)"), align = "hv", -->
-<!--                    label_fontface = "plain") -->
-<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/abun_and_sets_example.pdf", -->
-<!--                    base_height = 3.8, base_width = 8.5) -->
-<!-- p -->
-
-<!-- ``` -->
-
-<!-- <br> -->
-
-<!-- ```{r tot, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
-The heavy dark line indicates the true tred in the total population available to the survey."} -->
-
-<!-- which_set_dens <- c(0.0005, 0.002, 0.01) -->
-<!-- which_len_caps <- c(10, 100, 1000) -->
-
-<!-- d <- merge(sim$surveys, sim$total_strat_error, by = "survey") -->
-<!-- d <- d[d$set_den %in% which_set_dens &  -->
-<!--          d$lengths_cap == 100 & d$ages_cap == 10, ] ## length and age sampling not a factor here -->
-
-<!-- p <- fan_plot(d, x = "Year", which_strat = "Total") -->
-<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/total_strat_true_vs_est.pdf", -->
-<!--                    base_height = 3, base_width = 8) -->
-<!-- p -->
-
-<!-- ``` -->
-
-<!-- <br> -->
-
-<!-- ```{r length, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
-The heavy dark line indicates the true population available to the survey."} -->
-
-<!-- which_set_dens <- c(0.0005, 0.002, 0.01) -->
-<!-- which_len_caps <- c(10, 100, 1000) -->
-<!-- which_year <- 7 -->
-
-<!-- sub_surveys <- sim$surveys[sim$survey$set_den %in% which_set_dens & -->
-<!--                              sim$survey$lengths_cap %in% which_len_caps &  -->
-<!--                              sim$survey$ages_cap == 10, ] -->
-<!-- sub_strat <- sim$length_strat_error[eval(sim$length_strat_error$survey %in% sub_surveys$survey & -->
-<!--                                            sim$length_strat_error$year == which_year), ] -->
-<!-- d <- merge(sub_surveys, sub_strat, by = "survey") -->
-<!-- d$length[d$length > 100] <- 100 # aggregate beyond 100 cm -->
-<!-- d <- d[, list(I_hat = sum(I_hat), I = sum(I)),  -->
-<!--         by = c("year", "length", "sim", "survey", "set_den", "lengths_cap")] -->
-
-<!-- p <- fan_plot(d, x = "Length", which_strat = "Length") -->
-<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_length.pdf", -->
-<!--                    base_height = 6, base_width = 8) -->
-<!-- p -->
-
-<!-- ``` -->
-
-<!-- ```{r age, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fan chart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
-Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
-The heavy dark line indicates the true population available to the survey."} -->
-
-<!-- which_set_dens <- c(0.0005, 0.002, 0.01) -->
-<!-- which_len_caps <- c(10, 100, 1000) -->
-<!-- which_year <- 7 -->
-
-<!-- sub_surveys <- sim$surveys[sim$survey$set_den %in% which_set_dens & -->
-<!--                              sim$survey$lengths_cap %in% which_len_caps &  -->
-<!--                              sim$survey$ages_cap == 10, ] -->
-<!-- sub_strat <- sim$age_strat_error[eval(sim$age_strat_error$survey %in% sub_surveys$survey & -->
-<!--                                         sim$age_strat_error$year == which_year), ] -->
-<!-- d <- merge(sub_surveys, sub_strat, by = "survey") -->
-<!-- d <- d[d$year == which_year & d$set_den %in% which_set_dens & -->
-<!--          d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ] -->
-<!-- d$age[d$age > 10] <- 10 # aggregate beyond 10 -->
-<!-- d <- d[, list(I_hat = sum(I_hat), I = sum(I)),  -->
-<!--        by = c("year", "age", "sim", "survey", "set_den", "lengths_cap")] -->
-
-<!-- p <- fan_plot(d, x = "Age") -->
-<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_age.pdf", -->
-<!--                    base_height = 6, base_width = 8) -->
-<!-- p -->
-
-<!-- ``` -->
-
-<!-- <br> -->
-
-<!-- ```{r year, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
-Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
-The heavy dark line indicates the true trend in the population available to the survey."} -->
-
-<!-- which_set_dens <- c(0.0005, 0.002, 0.01) -->
-<!-- which_len_caps <- c(10, 100, 1000) -->
-<!-- which_age <- 4 -->
-
-<!-- d <- merge(sim$surveys, sim$age_strat_error, by = "survey") -->
-<!-- d <- d[d$age == which_age & d$set_den %in% which_set_dens & -->
-<!--        d$lengths_cap %in% which_len_caps & d$ages_cap == 10, ] -->
-
-<!-- p <- fan_plot(d, x = "Year") -->
-<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/true_vs_est_by_year.pdf", -->
-<!--                    base_height = 6, base_width = 8) -->
-<!-- p -->
-
-<!-- ``` -->
-
-<!-- <br> -->
-
-<!-- ```{r error-lines, cache = TRUE, fig.height = 3.8, fig.width = 8.5, results = "hide", fig.cap = "- Root-mean-squared error (RMSE) of abundance at length and abundance at age estimates as a function of the number of fish measured ($N_{lengths}$) from an array of surveys with different sampling protocol.
-The labels represent surveys with different set densities ($D_{sets}$), where number of sets are shown in brackets, and the lines are color coded by length sampling protocol ($M_{lengths}$).
-Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios."} -->
-
-<!-- totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught), -->
-<!--                                  n_measured = mean(n_measured), n_aged = mean(n_aged)), -->
-<!--                           by = "survey"] -->
-
-<!-- errors <- merge(sim$surveys, sim$length_strat_error_stats, by = "survey") -->
-<!-- length_d <- merge(errors, totals, by = "survey") -->
-<!-- length_d <- length_d[length_d$ages_cap == 10, ] -->
-<!-- length_d$strat <- "Abundance at length" -->
-
-<!-- errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey") -->
-<!-- age_d <- merge(errors, totals, by = "survey") -->
-<!-- age_d <- age_d[age_d$ages_cap == 10, ] -->
-<!-- age_d$strat <- "Abundance at age" -->
-
-<!-- d <- rbind(length_d, age_d) -->
-<!-- d$cap <- factor(d$lengths_cap) -->
-<!-- d$strat <- factor(d$strat, levels = c("Abundance at length", "Abundance at age")) -->
-
-<!-- lab_d <- d[lengths_cap == max(lengths_cap), -->
-<!--            list(lengths_cap, n_measured, n_sets, cap, RMSE), -->
-<!--            by = c("set_den", "strat")] -->
-<!-- lab_d$lab <- paste0(formatC(lab_d$set_den), " (", lab_d$n_sets, ")") -->
-
-<!-- p <- ggplot(data = d, aes(x = n_measured, y = RMSE, group = set_den)) + -->
-<!--   geom_line(aes(color = cap), size = 2, lineend = "round") + -->
-<!--   geom_text(data = lab_d, aes(label = lab), nudge_x = 500, hjust = 0) + -->
-<!--   geom_point(data = lab_d, aes(color = cap)) + -->
-<!--   xlab(bquote(N[lengths])) + -->
-<!--   scale_y_continuous(labels = format_si()) + -->
-<!--   scale_x_continuous(labels = format_si(), limits = c(0, max(d$n_measured) * 1.3), -->
-<!--                      breaks = pretty(d$n_measured, 6)) + -->
-<!--   scale_color_viridis(name = bquote(M[lengths]), discrete = TRUE) + -->
-<!--   facet_wrap(~ strat, scales = "free", dir = "h") + -->
-<!--   theme_bw() + -->
-<!--   theme(legend.justification = c(0, 1), -->
-<!--         plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm"), -->
-<!--         panel.grid.major = element_blank(), -->
-<!--         panel.grid.minor = element_blank(), -->
-<!--         strip.background = element_blank(), -->
-<!--         panel.border = element_rect(fill = NA, color = "darkgrey", size = 0.4), -->
-<!--         axis.ticks = element_line(color = "darkgrey", size = 0.4)) -->
-
-<!-- cowplot::save_plot(plot = p, filename = "analysis/paper/figures/error_lines.pdf", -->
-<!--                    base_height = 3.8, base_width = 8.5) -->
-<!-- p -->
-
-<!-- ``` -->
-
-<!-- <br> -->
-
-<!-- ```{r error-wireframe, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
-Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$)."} -->
-
-<!-- trellis_theme <- list( -->
-<!--   box.3d = list(col = "darkgrey"),  -->
-<!--   axis.line = list(col = "transparent"), -->
-<!--   strip.background = list(col = "transparent"),  -->
-<!--   strip.border = list(col = "transparent"), -->
-<!--    layout.heights = list( -->
-<!--         top.padding = 0, -->
-<!--         main.key.padding = 0, -->
-<!--         key.axis.padding = 0, -->
-<!--         axis.xlab.padding = 0, -->
-<!--         xlab.key.padding = 0, -->
-<!--         key.sub.padding = 0, -->
-<!--         bottom.padding = -5 -->
-<!--     ), -->
-<!--     layout.widths = list( -->
-<!--         left.padding = 0, -->
-<!--         key.ylab.padding = 0, -->
-<!--         ylab.axis.padding = 0, -->
-<!--         axis.key.padding = 0, -->
-<!--         right.padding = 1 -->
-<!--     ) -->
-<!-- ) -->
-
-<!-- zscale_format <- function(lim, logsc = FALSE, ...) { -->
-<!--     ans <- zscale.components.default(lim, logsc = logsc, ...) -->
-<!--     ans$left$labels$labels <- si(ans$left$labels$at) -->
-<!--     ans -->
-<!-- } -->
-
-<!-- d <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey") -->
-<!-- # d <- d[d$set_den != 0.0005, ] -->
-<!-- d <- d[d$set_den %in% c(0.0005, 0.002, 0.01)] -->
-<!-- set_den <- sort(unique(d$set_den)) -->
-<!-- d$set_den <- factor(d$set_den) -->
-<!-- options(scipen = 999) -->
-<!-- labs <- sapply(seq_along(set_den), function(i) as.expression(bquote(D[sets] == .(set_den[i]))),  -->
-<!--                USE.NAMES = FALSE) -->
-<!-- zlabs <- list(at = pretty(d$RMSE), labels = si(pretty(d$RMSE))) -->
-
-<!-- pdf("analysis/paper/figures/error_wireframe.pdf", height = 3, width = 8.5) -->
-
-<!-- p <- wireframe(RMSE ~  lengths_cap * ages_cap | set_den, data = d, -->
-<!--           as.table = TRUE, drape = TRUE, col = "white", lwd = 0.5, -->
-<!--           strip = strip.custom(factor.levels = labs), -->
-<!--           scales = list(arrows = FALSE, cex = 0.7, z = zlabs), -->
-<!--           col.regions = viridis::viridis(100), -->
-<!--           screen = list(z = 230, x = 300),  -->
-<!--           xlab = list(expression(M[lengths]), rot = 25),  -->
-<!--           ylab = list(expression(M[ages]), rot = -15),  -->
-<!--           zlab = list("RMSE", rot = 95), -->
-<!--           colorkey = list(labels = zlabs, height = 0.6, -->
-<!--                           axis.line = list(col = "darkgrey")), -->
-<!--           par.settings = trellis_theme) -->
-<!-- p <- update(p, layout = c(3, 1)) -->
-<!-- p -->
-<!-- grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc")) -->
-
-<!-- invisible(dev.off()) -->
-
-<!-- p -->
-<!-- grid::grid.text("RMSE", x = unit(0.952, "npc"), y = unit(0.77, "npc")) -->
-
-<!-- ``` -->
-
-<!-- <br> -->
-
-# Appendix 1: age-year-space covariance
+# Appendix 2: age-year-space covariance
 

### 2019-05-28 analysis/paper/sim_survey_paper.Rmd
Add a research oppertunity section

@@
 
-- Catchability
-- Stratified vs. random age sampling
-- Model testing
+The case study used here provides one example of ways in which the **`SimSurvey`** package can be used to simulation test the design of fisheries-independent trawl surveys.
+There are multiple layers to the sampling design of such surveys and the end results are inextricably linked with the way such data are analyzed.
+Below we outline some examples where the **`SimSurvey`** package may aid future research efforts.
+
+## Design or model-based approach
+
+The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formula for stratified-random designs [e.g.
+@Cochran1977].
+Nevertheless, there has long been interest in using model-based approaches to improve abundance estimates [e.g.
+@smith1990; @Berg2014; @Thorson2015].
+The **`SimSurvey`** can serve as a convenient tool for simulation testing the mean and variance estimates provided by a range of different approaches (design-based analyses, bootstrap estimates, generalized additive models, geostatistical models, etc.).
+Moreover, the full analytical pathway for obtaining age-disaggregated estimates of abundance has rarely been simulation tested, perhaps because there are few alternate options to a design-based approach [but see @Berg2014].
+Existing and future approaches to the integrated use of age-length keys or von Bertalanffy growth and catch-at-length data to obtain age-based indices of abundance can be simulation tested using *`SimSurvey`*.
+
+## Growth analyses
+
+Assessing ages for a large number of fish is very time-consuming and, as such, length-stratified sampling is often used to estimate age frequencies of fish populations.
+The resultant sub-sample is used to construct an age-length key (i.e.
+the probability a fish is a specific age given length) and age frequencies are obtained by applying this key to length-frequencies obtained via more expansive random sampling.
+One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in the abundance-at-age estimates [@Berg2012].
+Results from the case study presented here (Appendix 1) reiterate this message and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
+@Berg2012].
+
+## Random or stratified sampling
+
+**`SimSurvey`** can be used to compare the precision and bias of population estimates obtained using random or stratified sampling.
+Simple random sampling can be implemented using a grid with one strata (e.g.
+`make_grid(depth_range = c(0, 1000), strat_breaks = c(0, 1000), strat_split = 0)`).
+Sub-sampling of ages can also be random rather than length-stratified by setting the `age_sampling` argument in the `sim_survey` function to `"random"` rather than `"stratified"`.
+This can facilitate research similar to work presented in Puerta et al.
+[-@puerta2019].
 
@@
 
+- Multi-species, multiple goals
 - Cost-benefit
+- Survey coverage
+
+Up to now, the package has focused on testing...
 

### 2019-05-29 analysis/paper/sim_survey_paper.Rmd
Draft a future directions section

@@
 
-The case study used here provides one example of ways in which the **`SimSurvey`** package can be used to simulation test the design of fisheries-independent trawl surveys.
+The case study used here provides one example of ways in which **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
 There are multiple layers to the sampling design of such surveys and the end results are inextricably linked with the way such data are analyzed.
@@
 @smith1990; @Berg2014; @Thorson2015].
-The **`SimSurvey`** can serve as a convenient tool for simulation testing the mean and variance estimates provided by a range of different approaches (design-based analyses, bootstrap estimates, generalized additive models, geostatistical models, etc.).
-Moreover, the full analytical pathway for obtaining age-disaggregated estimates of abundance has rarely been simulation tested, perhaps because there are few alternate options to a design-based approach [but see @Berg2014].
-Existing and future approaches to the integrated use of age-length keys or von Bertalanffy growth and catch-at-length data to obtain age-based indices of abundance can be simulation tested using *`SimSurvey`*.
+**`SimSurvey`** can serve as a convenient tool for simulation testing mean and variance estimates provided by a range of different approaches (design-based analyses, bootstrap estimates, generalized additive models, geostatistical models, etc.).
+Moreover, the full analytical pathway for obtaining age-disaggregated estimates of abundance has rarely been simulation tested.
+Existing and future approaches for calculating age-based indices of abundance can be simulation tested using *`SimSurvey`*.
 
@@
 the probability a fish is a specific age given length) and age frequencies are obtained by applying this key to length-frequencies obtained via more expansive random sampling.
-One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in the abundance-at-age estimates [@Berg2012].
+One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012].
 Results from the case study presented here (Appendix 1) reiterate this message and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
@@
 
-- Multi-species, multiple goals
-- Cost-benefit
-- Survey coverage
-
-Up to now, the package has focused on testing...
+Up to now, the package has focused on the effects of sampling design on the precision and bias of population estimates obtained from fisheries-independent surveys, however, the costs associated with sampling has yet to be considered.
+In future iterations of **`SimSurvey`**, we hope to add options for integrating data on the time and monetary costs associated with each level of sampling (sets, length measurements, age determination) to facilitate cost-benefit analyses.
+We also realize that a single fisheries-independent survey may have multiple goals as data obtained are often used to assess multiple species or to conduct community analyses.
+We therefore endeavor to add functions for simulating multi-species surveys.
+Finally, it would be useful to add an option for testing the consequences of surveys with partial coverage of a population as survey coverage is a frequent concern in stock assessment.
 
@@
 
-- Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
+- Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
 - Dave Cote for a review of a previous version of this manuscript

### 2019-05-29 analysis/paper/sim_survey_paper.Rmd
Modify and augment summary section

@@
 The **`SimSurvey`** package serves as a tool for simulating stratified random surveys of dynamic populations that vary across ages, time and space.
-The core of the simulation is based on the widely used cohort equation and, even though the processes that define recruitment and total mortality are simple, a wide range of stock dynamics can be simulated by changing a few parameters (Figure \@ref(fig:surface)).
-This base population can then be distributed through a grid (Figure \@ref(fig:grid)) and relationships with depth can be defined as can the nature of the correlation across ages, years and space (Figure \@ref(fig:distribution)).
+The core of the simulation is based on the widely used cohort equation and, even though the processes that define recruitment and total mortality are simple, a wide range of stock dynamics can be simulated by changing a few parameters.
+This base population can then be distributed through a grid and relationships with depth can be defined as can the nature of the correlation across ages, years and space.
 Together, two functions (`sim_abundance` and `sim_distribution`) are capable of simulating a wide range of populations with different life histories, depth associations and spatial properties.
@@
 Sub-sampling protocol (length and age sampling) can also be varied.
-As such, data from a wide range of surveys can be simulated (Figure \@ref(fig:survey)).
-There are a plethora of statistical models that may be tested using these simulated data, however, implementing a variety of analytical approaches was outside the scope of this work.
+As such, data from a wide range of surveys can be simulated.
+
+There are a plethora of statistical models that may be tested using these simulated data, however, implementing a variety of analytical approaches was outside the scope of this package.
 Instead we focus on analyzing simulated stratified-random survey data using a design-based stratified analysis.
@@
 
+Simulation testing is an important tool in the field of fisheries science as the inferred status of fish stocks hing on the data and models used to assess fish populations.
+Simulations provide an opportunity to explore survey and model performance, and such explorations are becoming increasingly important as model complexity increases.
+It is also important to continually assess the efficacy and efficiency of sampling programs given their costs and the constant scrutiny of the value added by such surveys.
+These are some of the reasons multiple simulation frameworks, including **`SimSurvey`**, have been developed to test the design and analyses of complex surveys.
+We have made **`SimSurvey`** as open and accessible as possible to allow the broader community to validate, reuse and improve this package.
+We hope that open-source sharing will extend the value of such simulation frameworks and we encourage users to extend the package for their own needs and contribute to future versions.
+
 # Acknowledgements
@@
 - Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
-- Dave Cote for a review of a previous version of this manuscript
+- Dave Cote, Keith Lewis and Greg Robertson for a review of a previous version of this manuscript
 - NSERC visiting-fellow program

### 2019-05-30 analysis/paper/sim_survey_paper.Rmd
Start to re-work the introduction

@@
 
+It is well known that the accuracy and reliability of any population monitoring program hinges on good survey design and analysis.
+Ideally, the design and analysis of any large and costly survey should be simulation tested before embarking on expensive field programs.
+This is especially true for the large-scale and costly fisheries-independent trawl surveys conducted by numurous regional fisheries management organizations.
+These surveys have become an increasingly important research tool for the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
+While costly to obtain, this information forms the basis of of many stock assessments used throughout the world.
+Efficient, effective and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection.
+Solving this optimization problem is not a trivial task given the complexity of the population processes we aim to represent and the multi-stage nature of the typical sampling program.
+
+The complexity of the system and the surveys may be why simulation tests of the full sampling and analylicical pathway are rare.
+Simulations provide a playground of sorts...multiple frameworks...few accessible...
+
 Fisheries-independent trawl surveys have become an increasingly important research tool for the management of dynamic fish stocks.
-These surveys provide reliable indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
+These surveys provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
 While costly to obtain, this information forms the basis of of many stock assessments models used throughout the world.
@@
 Determining the number of samples required to adequately characterize a fish population is particularly challenging since data tend to be collected across one or more stages [@Aanes2015].
-Optimal survey design is further complicated by the fact that trawls sample clusters of fish that tend to have similar characteristics, such as length, age and stomach contents.
+Optimal survey design is further complicated by the fact that trawls sample clusters of fish that tend to have similar characteristics, such as length and age.
 This phenomenon results in samples with positive intra-haul correlation, which can markedly reduce the effective sample size for estimating length and age frequencies of the target population [@Pennington1994; @Pennington2002; @Stewart2014].
-The complexities of multi-stage sampling and intra-haul correlation make it difficult to answer the question *"how much sampling is enough?"*
+The complexities of multi-stage sampling and intra-haul correlation make it difficult to answer the question *"how much sampling is enough?"* Moreover, the answer to this question is inextricably linked with the way in witch these data are analyzed.
+
+While analyses of fisheries-independent survey data has typically been confined to design-based approaches [e.g.
+stratified-random survey data tend to be analyzed using standard stratified-analysis formulae;  @Cochran1977], there are a growing number of model-based approaches for analyzing such data [e.g.
+@smith1990; @Berg2014; @Thorson2015].
+The merits of both approaches have long been debated, however, simulation tests of the full sampling and analylitical pathway are rarely conducted.
+This is especially true for design- or model-based estimates of abundance-at-age, which is a criticial input to any age-based stock assessment model.
+Ideally, the design and analysis of any large and costly survey should be simulation tested before embarking on an expensive field program.
 
@@
 Again, this function retains all the objects listed in the output of `sim_distribution` and adds data.tables that detail the set locations (`setdet`) and sampling details (`samp`).
-Catchability corrected abundance matrices, named `I` and `I_at_length`, are also produced and added to the output; these matrices are useful for comparing the true abundance available to the survey to abundance estimates obtained using survey-based or model-based analyses of the simulated survey data.
+Catchability corrected abundance matrices, named `I` and `I_at_length`, are also produced and added to the output; these matrices are useful for comparing the true abundance available to the survey to abundance estimates obtained using design-based or model-based analyses of the simulated survey data.
 Specific surveys can be explored using the `plot_survey` function, which uses plotly [@sievert2018] and crosstalk [@cheng2016] in the background to link the bubble plot of aggregate set catch to the histogram of lengths and ages sampled to facilitate explorations of set-specific catches.
@@
 The case study used here provides one example of ways in which **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
-There are multiple layers to the sampling design of such surveys and the end results are inextricably linked with the way such data are analyzed.
+There are multiple layers to the sampling design of such surveys and the end results are linked with the way such data are analyzed.
 Below we outline some examples where the **`SimSurvey`** package may aid future research efforts.
@@
 
-- Feedback and advice: Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
+- Feedback and advice: Aaron Adamack, Jonathan Babyn, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
+- Package contributions: Jonathan Babyn, Greg Robertson
 - Dave Cote, Keith Lewis and Greg Robertson for a review of a previous version of this manuscript

### 2019-05-31 analysis/paper/sim_survey_paper.Rmd
Simplify the introduction

@@
 
-- As with any population monitoring program, accurate and reliable results from a fisheries-independent trawl survey depends on good survey design.
-Ideally, the sampling design of large and costly surveys should be simulation tested before embarking on an expensive field program.
+It is well known that the accuracy and reliability of results from any population monitoring program hinges on good survey design and analysis [@thomas2010].
+Ideally, the design and analysis of large and costly survey should be simulation tested before embarking on expensive field programs.
+This is especially true for the large-scale and costly fisheries-independent trawl surveys conducted by numerous regional fisheries management organizations.
+These surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
+This information forms the basis of of many stock assessments throughout the world.
+Efficient, effective and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection.
+While simulations provide a platform for exploring solutions to this optimization problem, building the necessary simulation framework is not a trivial task given the multi-stage nature of the typical sampling program and the complexity of the population processes we aim to represent.
+The complexity of this problem is perhaps why simulations that test the full sampling and analytical pathway of fisheries-independent surveys are rare [but see @puerta2019; @schnute2003].
 
-- While there are several simulation frameworks for testing the design of fisheries-independent trawl surveys [e.g.
-@puerta2019; @schnute2003], there are few tools that facilitate such work.
-Here we document an R package with a series of functions for testing the multi-layered sampling design of spatial, age-structured populations.
-The package can also serve as a medium for the simulation testing of methods for analyzing such data.
-
-It is well known that the accuracy and reliability of any population monitoring program hinges on good survey design and analysis.
-Ideally, the design and analysis of any large and costly survey should be simulation tested before embarking on expensive field programs.
-This is especially true for the large-scale and costly fisheries-independent trawl surveys conducted by numurous regional fisheries management organizations.
-These surveys have become an increasingly important research tool for the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
-While costly to obtain, this information forms the basis of of many stock assessments used throughout the world.
-Efficient, effective and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection.
-Solving this optimization problem is not a trivial task given the complexity of the population processes we aim to represent and the multi-stage nature of the typical sampling program.
-
-The complexity of the system and the surveys may be why simulation tests of the full sampling and analylicical pathway are rare.
-Simulations provide a playground of sorts...multiple frameworks...few accessible...
-
-Fisheries-independent trawl surveys have become an increasingly important research tool for the management of dynamic fish stocks.
-These surveys provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
-While costly to obtain, this information forms the basis of of many stock assessments models used throughout the world.
-Effective and informed management decisions therefore require surveys that maximize information while minimizing the expense of data collection.
-Determining the number of samples required to adequately characterize a fish population is particularly challenging since data tend to be collected across one or more stages [@Aanes2015].
-Optimal survey design is further complicated by the fact that trawls sample clusters of fish that tend to have similar characteristics, such as length and age.
-This phenomenon results in samples with positive intra-haul correlation, which can markedly reduce the effective sample size for estimating length and age frequencies of the target population [@Pennington1994; @Pennington2002; @Stewart2014].
-The complexities of multi-stage sampling and intra-haul correlation make it difficult to answer the question *"how much sampling is enough?"* Moreover, the answer to this question is inextricably linked with the way in witch these data are analyzed.
-
-While analyses of fisheries-independent survey data has typically been confined to design-based approaches [e.g.
-stratified-random survey data tend to be analyzed using standard stratified-analysis formulae;  @Cochran1977], there are a growing number of model-based approaches for analyzing such data [e.g.
-@smith1990; @Berg2014; @Thorson2015].
-The merits of both approaches have long been debated, however, simulation tests of the full sampling and analylitical pathway are rarely conducted.
-This is especially true for design- or model-based estimates of abundance-at-age, which is a criticial input to any age-based stock assessment model.
-Ideally, the design and analysis of any large and costly survey should be simulation tested before embarking on an expensive field program.
-
-Two main approaches have been applied to try and answer this question: one involves statistical estimators of effective sample size [@Pennington2002; @Stewart2014], and the other employs resampling methods to test alternate sampling protocol [@Cervino2006; @Zhang2013].
-Both approaches have shown that lengths are usually oversampled and a reduction in said sampling can often be done without significant loss of accuracy [@Pennington2002].
-The same result holds for other positively correlated characteristics, such as age [@Coggins2013] and stomach contents [@Bogstad1995].
-Pennington and Vølstad [-@Pennington1994] concluded that the best way to improve the accuracy of surveys with strong intra-haul correlation is to take fewer samples per sampling unit and increase the total number of sampling locations.
-Few studies, however, have tested these conclusions using simulations.
-Simulations are a useful approach for informing the design of surveys [@sutherland2006] and they have been used widely for exploring the number of samples needed to obtain estimates with reasonable precision [e.g.
-distance sampling surveys; @thomas2010].
-A key advantage of a pure simulation approach is that "true" population size and all stages of sampling can be controlled (set, length and age sampling protocol).
-The relative effects of each stage of sampling on our perception of the "truth" can then be tested.
-This level of testing is of particular importance for age composition estimates, and the contemporary age-based assessment models that rely on these estimates, as they are affected by all stages of sampling.
-A key challenge, of course, is building a simulation that emulates reality.
-
-The objective of this paper is to use a realistic simulation to evaluate the efficacy of various sampling protocol.
-First we simulated a spatially autocorrelated target population, and then we simulated surveys over this field.
-Set density was varied across surveys, as was length and age sampling efforts.
-Estimates of abundance at age were then obtained using a standard stratified approach [@Smith1981] and deviation from the truth was assessed.
-Given previous findings, we predict that the greatest gains in precision will come from increasing the number of hauls; increased sampling of correlated variables (length and age) on a set by set basis will be relatively ineffective.
-The bottom trawl survey of cod (*Gadus morhua*) in NAFO division 3Ps, Newfoundland, was used as a case study for this analysis.
-While our case study is specific, we also hope that the approach presented here will provide a framework for optimizing sampling protocol for an array of species with different life-histories and distributions.
-To facilitate future work and collaborations, we have developed an R package for conducting the tests presented in this paper (https://github.com/PaulRegular/SimSurvey).
+Here we document **`SimSurvey`**, an **`R`** package designed to facilitate realistic simulations of fisheries-independent trawl surveys.
+This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al.
+[-@puerta2019], however, efforts were focused here on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways.
+The steps taken to simulate surveys of spatial, age-structure populations are outlined below, with specific emphasis of a case study that motivated this work.
 
@@
 
-The **`SimSurvey`** package was written in the programming language R [@R] and, in short, the package allows for the simulation of stratified-random surveys of an age-structured population that varies in space and time.
-**`SimSurvey`** relies heavily on functions from the data.table [@dowle2017], raster [@hijmans2016] and plotly [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively.
-All the source R code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey) and the 0.1.0 release of the package can be installed via GitHub:
+The **`SimSurvey`** package was written in the programming language **`R`** [@R] and, in short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time.
+**`SimSurvey`** relies heavily on functions from the **`data.table`** [@dowle2017], **`raster`** [@hijmans2016] and **`plotly`** [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively.
+All the source **`R`** code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey) and the package can be installed via GitHub:
 
@@
 install.packages("devtools")
-devtools::install_github("PaulRegular/SimSurvey@v0.1.0") 
+devtools::install_github("PaulRegular/SimSurvey")
 ```
@@
 Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division.
+Such sub-sampling is common in fisheries-independent surveys as it is costly to sample every fish sampled.
+Age determination is especially time-consuming, which is why age sampling tends to be length-stratified.
 
@@
 Moreover, the full analytical pathway for obtaining age-disaggregated estimates of abundance has rarely been simulation tested.
-Existing and future approaches for calculating age-based indices of abundance can be simulation tested using *`SimSurvey`*.
+Existing and future approaches for calculating age-based indices of abundance can be simulation tested using **`SimSurvey`**.
 

### 2019-05-31 analysis/paper/sim_survey_paper.Rmd
Revise abstrat, complete draft

@@
 
-# Abstract
+# Summary
 
-A central question to fish stock assessments is: how many samples are needed to adequately represent the population being monitored?
-For assessments based on trawl surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling (i.e.
-the number of sets, lengths and ages sampled in a survey).
-We therefore need to assess the impact of haul design and length and age sampling on survey accuracy to determine optimal sampling effort.
-In this study we simulated repeated stratified-random samples from a spatially correlated field of simulated fish.
-These samples were then analyzed and estimates of abundance at age were compared against a known truth.
-By varying the sampling protocol, we demonstrate that increasing the number of sets conducted per stratum provides the greatest gains in precision when sampling a clustered population.
-Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent and improvements were typically marginal, and sometimes squandered, when sub-sampling effort was increased.
-These results indicate that high levels of length and age sampling effort can be scaled back without significant losses to precision and, in some cases, estimates of abundance at age may be improved by reducing sub-sampling effort.
-Time saved by reducing said sampling may afford more time to conduct additional sets and/or increase sampling efforts of other species.
-While this simulation is tailored to a specific case, the method may be adapted to different systems and survey procedures.
+**1\.** Fisheries scientists are often faced with two critical questions: (i) how many samples are needed to adequately represent a population?
+And, (ii) how should samples from multi-stage sampling be analyzed?
 
-**Keywords** - length and age distribution; simulation; spatial correlation; stratified analysis; survey design
+**2\.** For assessments based on fisheries-independent surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling (i.e.
+the number of sets, lengths and ages sampled in a survey) as well as the methods used to analyze the data.
+
+**3\.** Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, system and survey complexity have limited the number of frameworks that simulation test the full sampling and analytical pathway of fisheries-independent survey data.
+
+**4\.** Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of spatial fish populations.
+The packages includes methods for simulating age-structured populations that varies in space and time as well as multi-stage surveys that sample the virtual population.
+A series of parameters can be used to control the characteristics of the population and the sampling intensity of the survey.
+**`SimSurvey`** also include a function for running a stratified-random analysis of the simulated data.
+
+**5\.** **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies of a range of species.
+Various design- or model-based approaches can be applied to the results to test the efficacy of different approaches.
+
+**Key-words:** length and age distribution; simulation; spatial correlation; stratified analysis; survey design
 
@@
 These surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
-This information forms the basis of of many stock assessments throughout the world.
+This information forms the basis of many stock assessments throughout the world.
 Efficient, effective and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection.
@@
 
-Here we document **`SimSurvey`**, an **`R`** package designed to facilitate realistic simulations of fisheries-independent trawl surveys.
+Here we document **`SimSurvey`**, an `R` package designed to facilitate realistic simulations of fisheries-independent trawl surveys.
 This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al.
-[-@puerta2019], however, efforts were focused here on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways.
+[-@puerta2019], however, efforts were focused on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways.
 The steps taken to simulate surveys of spatial, age-structure populations are outlined below, with specific emphasis of a case study that motivated this work.
@@
 
-The **`SimSurvey`** package was written in the programming language **`R`** [@R] and, in short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time.
+The **`SimSurvey`** package was written in the programming language `R` [@R] and, in short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time.
 **`SimSurvey`** relies heavily on functions from the **`data.table`** [@dowle2017], **`raster`** [@hijmans2016] and **`plotly`** [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively.
-All the source **`R`** code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey) and the package can be installed via GitHub:
+All the source `R` code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey) and the package can be installed via GitHub:
 
@@
 ```{r grid, fig.cap = "- Plots produced by `plot_grid` when supplied an raster object produced by `make_grid` a) using default case study settings and b) settings generates a grid with four divisions and a more linear depth gradient.
-In these plots, the color gradient represents depth, the grey lines deliniate divisions and white lines deliniate strata."}
+In these plots, the color gradient represents depth, the grey lines delineate divisions and white lines delineate strata."}
 
@@
 The `sim_survey_parallel` function was therefore constructed to facilitate the running of thousands of simulations without running into RAM limitations.
-This function is set-up to run multiple `sim_survey` calls in parallel using the doParallel package [@weston2015] and, as such, multiple loops can be run using the `n_loops` argument and within each loop multiple simulations can be run (controlled using the `n_sims` argument).
+This function is set-up to run multiple `sim_survey` calls in parallel using the **`doParallel`** package [@weston2015] and, as such, multiple loops can be run using the `n_loops` argument and within each loop multiple simulations can be run (controlled using the `n_sims` argument).
 Total simulations will be the product of `n_loops` and `n_sims` arguments.
@@
 `n_loops` and `n_sims` ratio) dependent.
-The `test_survey` function will print a progress bar, generated using the progress package [@csardi2016], which details percent completion and will also include an eta after the first step of the loop completes.
+The `test_survey` function will print a progress bar, generated using the **`progress`** package [@csardi2016], which details percent completion and will also include an eta after the first step of the loop completes.
 Keep in mind that thousands of simulations of hundreds of survey designs may take days to run.
@@
 
-```{r rank, fig.cap = "- Divergent dot plot of the percision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{sets}$] and length measurements [$N_{measured}$]), under various sampling protocol (set density [$D_{sets}$] and maximum number of lengths measured per set [$M_{lengths}$])."}
+```{r rank, fig.cap = "- Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{sets}$] and length measurements [$N_{measured}$]), under various sampling protocol (set density [$D_{sets}$] and maximum number of lengths measured per set [$M_{lengths}$])."}
 
@@
 
-- Feedback and advice: Aaron Adamack, Jonathan Babyn, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
-- Package contributions: Jonathan Babyn, Greg Robertson
-- Dave Cote, Keith Lewis and Greg Robertson for a review of a previous version of this manuscript
-- NSERC visiting-fellow program
+**Feedback and advice:** Aaron Adamack, Jonathan Babyn, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
+
+**Package contributions:** Jonathan Babyn, Greg Robertson
+
+**Review of previous draft:** Dave Cote, Keith Lewis, Greg Robertson
+
+**Funding:** NSERC visiting-fellow program
 
@@
 # References
+
+<!-- # Quotes -->
+
+<!-- "To effectively manage rare populations, accurate monitoring data are critical.
+Yet many monitoring programs are initiated without careful consideration of whether chosen sampling designs will provide accurate estimates of population parameters.
+Obtaining accurate estimates is especially difficult when natural variability is high, or limited budgets determine that only a small fraction of the population can be sampled." - Morrison et al.
+(2008) Popul Ecol 50:417–425 -->
+
+<!-- "Fishery independent surveys provide valuable information on fish population characteristics and play important roles in fisheries stock assessment (Gunderson, 1993; Rago, 2005).
+The importance of fishery independent surveys has been well recognized, but the important role of sampling design for these surveys is less appreciated.
+Often, traditional survey methods are used without considering survey designs that maximize the precision and accuracy of the estimator (Mier and Picquelle, 2008).
+Fishery independent surveys are usually expensive and time-consuming, and an efficient sampling design is key to the success of these surveys.
+Recently, investigators have begun to pay more attention to the importance of survey sampling and are exploring a variety of aspects of designs (Bez, 2002; Brown, 2003; Kimura and Somerton, 2006).
+Based on the previous studies, we compared the performance of traditional and adaptive sampling designs with emphasis on simulation application for performance comparison." - Yu et al.
+(2012) Fisheries Research 113: 173–181 -->
+
+<!-- # Examples -->
+
+<!-- Good examples of papers on packages: -->
+<!-- - https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0092725#s4 -->
+<!-- - file:///C:/Users/RegularP/Downloads/Mildenberger_et_al-2017-Methods_in_Ecology_and_Evolution.pdf -->
+
+<!-- # READ -->
+
+<!-- - https://onlinelibrary.wiley.com/doi/pdf/10.1046/j.1467-2960.2001.00047.x -->

### 2019-06-19 analysis/paper/sim_survey_paper.Rmd
Modify abstract given Greg and Keith's suggestions

@@
 title: "`SimSurvey`: an `R` package for simulation testing the design and analysis of fisheries-independent trawl surveys"
+subtitle: "ALTERNATE TITLE - `SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
 author: "Paul M. Regular, Fran Mowbray, et al.?"
@@
 
-**1\.** Fisheries scientists are often faced with two critical questions: (i) how many samples are needed to adequately represent a population?
-And, (ii) how should samples from multi-stage sampling be analyzed?
+**1\.** Age-structured fish stocks often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective population surveys.
+Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision and increasing the risk of bias) and over-sampling (through the extensive sampling of correlated metrics).
 
-**2\.** For assessments based on fisheries-independent surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling (i.e.
-the number of sets, lengths and ages sampled in a survey) as well as the methods used to analyze the data.
+**2\.** For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of fishing sets and sub-sampling the fish captured to measure biological characteristics (e.g.
+lengths or ages).
+Multiple analysis approaches are also now available.
 
-**3\.** Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, system and survey complexity have limited the number of frameworks that simulation test the full sampling and analytical pathway of fisheries-independent survey data.
+**3\.** Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, complex fish population structure has limited the number of tools that allow for simulation testing of a range of sampling and analytical pathways for fisheries-independent survey data.
 
-**4\.** Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of spatial fish populations.
-The packages includes methods for simulating age-structured populations that varies in space and time as well as multi-stage surveys that sample the virtual population.
-A series of parameters can be used to control the characteristics of the population and the sampling intensity of the survey.
+**4\.** Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations.
+The package allows the user to simulate age-structured populations that vary in space and time and then explore the efficacy of a range of sampling protocols to reproduce the population parameters of the known population.
 **`SimSurvey`** also include a function for running a stratified-random analysis of the simulated data.
 
-**5\.** **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies of a range of species.
-Various design- or model-based approaches can be applied to the results to test the efficacy of different approaches.
+**5\.** **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring.
+Various design- or model-based approaches can then be applied to the results to test the efficacy of different analytical approaches.
 
@@
 
-Table: (\#tab:par) - Main function calls, with argument descriptions and associated parameter symbols, used to simulate the cod-like population used in the case study presented here.
+Table: (\#tab:par)Main function calls, with argument descriptions and associated parameter symbols, used to simulate the cod-like population used in the case study presented here.
 Note that the settings below largely correspond with the defaults used in the functions.
@@
 
-```{r surface, fig.cap = "- Surface plots of simulated abundance at age of a relatively a) long lived (case study settings) and b) short lived species.
+```{r surface, fig.cap = "Surface plots of simulated abundance at age of a relatively a) long lived (case study settings) and b) short lived species.
 These plots were produced by `plot_surface` when supplied a list produced by `sim_abundance`."}
@@
 
-```{r grid, fig.cap = "- Plots produced by `plot_grid` when supplied an raster object produced by `make_grid` a) using default case study settings and b) settings generates a grid with four divisions and a more linear depth gradient.
+```{r grid, fig.cap = "Plots produced by `plot_grid` when supplied an raster object produced by `make_grid` a) using default case study settings and b) settings generates a grid with four divisions and a more linear depth gradient.
 In these plots, the color gradient represents depth, the grey lines delineate divisions and white lines delineate strata."}
@@
 
-```{r distribution, fig.cap = "- Distribution plots of simulated populations that form a) tight clusters with stronger correlation through years than ages (case study settings), and b) relatively diffuse clusters with stronger correlation through ages than years.
+```{r distribution, fig.cap = "Distribution plots of simulated populations that form a) tight clusters with stronger correlation through years than ages (case study settings), and b) relatively diffuse clusters with stronger correlation through ages than years.
 This plot is a facet of plots produced by `plot_distribution` when supplied simulations from `sim_distribution`."}
@@
 
-```{r survey, fig.cap = "- Bubble plots and histograms of set catches from a simulated stratified-random survey of the case study population under relatively a) high and b) low sampling effort.
+```{r survey, fig.cap = "Bubble plots and histograms of set catches from a simulated stratified-random survey of the case study population under relatively a) high and b) low sampling effort.
 Histograms of length and age composition include the distribution caught and sampled overlaid with a line of the true distribution of length and ages available to the survey.
@@
 
-```{r fan1, fig.cap = "- Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
+```{r fan1, fig.cap = "Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
 The heavy dark line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
@@
 
-```{r fan2, fig.cap = "- Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+```{r fan2, fig.cap = "Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 The heavy dark line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
@@
 
-```{r fan3, fig.cap = "- Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+```{r fan3, fig.cap = "Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
@@
 
-```{r rank, fig.cap = "- Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{sets}$] and length measurements [$N_{measured}$]), under various sampling protocol (set density [$D_{sets}$] and maximum number of lengths measured per set [$M_{lengths}$])."}
+```{r rank, fig.cap = "Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{sets}$] and length measurements [$N_{measured}$]), under various sampling protocol (set density [$D_{sets}$] and maximum number of lengths measured per set [$M_{lengths}$])."}
 
@@
 
-```{r rmse-surface, fig.cap = "- Surface plots of RMSE from an array of surveys with different sampling protocol.
+```{r rmse-surface, fig.cap = "Surface plots of RMSE from an array of surveys with different sampling protocol.
 Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$).
@@
 [-@puerta2019].
+
+<!-- ESTIMATION OF VARIANCE. What does the stratified estimates of variance look like with the simulations you already have? Do negative values disapear above a praticular sampling intensity? -->
 

### 2019-06-20 analysis/paper/sim_survey_paper.Rmd
Modify introduction | Greg, Keith eidts

@@
 
-It is well known that the accuracy and reliability of results from any population monitoring program hinges on good survey design and analysis [@thomas2010].
-Ideally, the design and analysis of large and costly survey should be simulation tested before embarking on expensive field programs.
-This is especially true for the large-scale and costly fisheries-independent trawl surveys conducted by numerous regional fisheries management organizations.
-These surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
-This information forms the basis of many stock assessments throughout the world.
-Efficient, effective and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection.
+Fisheries-independent trawl surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
+While costly to obtain, this information forms the basis of many stock assessments throughout the world.
+Efficient and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection.
 While simulations provide a platform for exploring solutions to this optimization problem, building the necessary simulation framework is not a trivial task given the multi-stage nature of the typical sampling program and the complexity of the population processes we aim to represent.
-The complexity of this problem is perhaps why simulations that test the full sampling and analytical pathway of fisheries-independent surveys are rare [but see @puerta2019; @schnute2003].
+Notably, many fish stocks show size-specific patterns of aggregation, making it difficult to collect biological sub-samples (e.g.
+lengths and ages) that are representative of the entire target population.
+This undermines the assumptions of many analyses as sub-samples from such populations tend to exhibit strong positive intracluster correlation [@Aanes2015].
+These complexities may explain why simulations that test the full sampling and analytical pathway of fisheries-independent surveys are rare [but see @puerta2019; @schnute2003].
 
-Here we document **`SimSurvey`**, an `R` package designed to facilitate realistic simulations of fisheries-independent trawl surveys.
+Here we document **`SimSurvey`**, an `R` package designed to simplify and facilitate realistic simulations of fisheries-independent trawl surveys.
+The package has two main components, the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated fish stock distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.
 This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al.
 [-@puerta2019], however, efforts were focused on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways.
-The steps taken to simulate surveys of spatial, age-structure populations are outlined below, with specific emphasis of a case study that motivated this work.
+The steps taken to simulate surveys of spatial, age-structure populations are outlined below.
 

### 2019-06-27 analysis/paper/sim_survey_paper.Rmd
Start moving the case study details to the appendix and add a table that describes the key functions of the package

@@
 library(data.table)
+library(magrittr)
 Sys.setenv("MAPBOX_TOKEN" = "pk.eyJ1IjoicGF1bHJlZ3VsYXIiLCJhIjoiY2pueGIxejMxMGM0MjNqbjR1aXdwYm92bSJ9.R-t1SCa0VHKf4N-PQmYI5Q")
@@
 Here we document **`SimSurvey`**, an `R` package designed to simplify and facilitate realistic simulations of fisheries-independent trawl surveys.
+In short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time
+
 The package has two main components, the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated fish stock distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.
+
 This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al.
@@
 
-The **`SimSurvey`** package was written in the programming language `R` [@R] and, in short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time.
+The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for simulating the abundance and distribution of virtural fish populations with correlation across space, time and age dimensions, as well as functions for simulating surveys with a range of sampling stratigies and intensities (Table \@ref(tab:fun)).
+The equations behind these functions are detailed in the sections below.
 **`SimSurvey`** relies heavily on functions from the **`data.table`** [@dowle2017], **`raster`** [@hijmans2016] and **`plotly`** [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively.
-All the source `R` code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey) and the package can be installed via GitHub:
+Package documentation has been published online using **`pkgdown`** (https://paulregular.github.io/SimSurvey/) and all the source `R` code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey).
+**`SimSurvey`** can be installed via GitHub using the **`remotes`** package:
 
 ```{r, echo = TRUE, eval = FALSE}
-install.packages("devtools")
-devtools::install_github("PaulRegular/SimSurvey")
+install.packages("remotes")
+remotes::install_github("PaulRegular/SimSurvey")
 ```
 
-## Case study
+Table: (\#tab:fun) Names and descritpions of the key functions of **`SimSurvey`**.
+Functions in bold font are core functions and those in medium font are designed for use inside the core functions.
+The latter are typicaly closures, which are functions that contain data and return functions [@wickham2014]; here they are used to store parameter values and return functions that require dimensions, such as ages or years, to be supplied.
 
-The construction of the **`SimSurvey`** package was motivated by a regional need for assessing the sampling design of trawl surveys conducted along the Newfoundland and Labrador shelf by Fisheries and Oceans Canada.
-As a tangible first step, we focused our efforts on emulating data from the survey of cod in NAFO division 3Ps since it is one of the most dynamic and heavily sampled stocks in the region.
-Below we outline the equations behind a series of **`SimSurvey`** functions that were used to simulate survey data with similar features to data obtained by the survey of 3Ps cod.
-The settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data; these values were set as the defaults for the main functions in the package.
-Here we evaluate the efficacy of alternate sampling protocol by assessing deviation between stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
-While assessing model-based analyses was an option, we focused on the design-based stratified analysis for simplicity and because of its widespread use.
-See Table \@ref(tab:par) for a description of the code used to generate the case study results.
-Text below provides more detail on each function and also demonstrates a series of functions for visualizing the results.
-See Appendix 1 for an interpretation and discussion of the case study results.
-
-Table: (\#tab:par)Main function calls, with argument descriptions and associated parameter symbols, used to simulate the cod-like population used in the case study presented here.
-Note that the settings below largely correspond with the defaults used in the functions.
-
-|                                       **Function call** |                                            **Description** |                 **Symbol** |
-|:------------------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
-| `> set.seed(438)`                                       |      Set random number generator seed to replicate results |                            |
-| `>`                                                     |                                                            |                            |
-| `> abundance <- sim_abundance(`                         |                                       *Simulate abundance* |                            |
-| `+   ages = 1:20,`                                      |                                                       Ages |                        $a$ |
-| `+   years = 1:20,`                                     |                                                      Years |                        $y$ |
-| `+   R = sim_R(mean = 30000000,`                        |                                           Mean recruitment |                    $\mu_r$ |
-| `+             log_sd = 0.5),`                          |                      Standard deviation of log-recruitment |                 $\sigma_r$ |
-| `+   Z = sim_Z(mean = 0.5,`                             |                                       Mean total mortality |                    $\mu_Z$ |
-| `+             log_sd = 0.2,`                           |                      Standard deviation of total mortality |                 $\sigma_Z$ |
-| `+             phi_age = 0.9,`                          |    Correlation across ages in error around total mortality |    $\varphi_{\delta, age}$ |
-| `+             phi_year = 0.5),`                        |   Correlation across years in error around total mortality |   $\varphi_{\delta, year}$ |
-| `+   growth = sim_vonB(Linf = 120,`                     |                                Mean asymptotic length (cm) |               $l_{\infty}$ |
-| `+                     L0 = 5,`                         |                                       Length at birth (cm) |                      $l_0$ |
-| `+                     K = 0.1,`                        |                                                Growth rate |                        $K$ |
-| `+                     log_sd = 0.1,`                   |     Standard deviation of the von Bertalanffy growth curve |                 $\sigma_l$ |
-| `+                     length_group = 3))`              |                  Length group for abundance at length (cm) |            $l_{group}^{N}$ |      
-| `>`                                                     |                                                            |                            |
-| `> grid <- make_grid(`                                  |                                              *Make a grid* |                            |
-| `+   x_range = c(-140, 140),`                           |                           Range of grid in the x dimension |                            |
-| `+   y_range = c(-140, 140),`                           |                           Range of grid in the y dimension |                            | 
-| `+   res = c(3.5, 3.5),`                                |                                       Grid resolution (km) |                        $R$ |
-| `+   shelf_depth = 200,`                                |                                           Shelf depth (km) |                $d_{shelf}$ |
-| `+   shelf_width = 100,`                                |                                           Shelf width (km) |                $w_{shelf}$ |
-| `+   depth_range = c(0, 1000),`                         |                                            Depth range (m) |       $[d_{min}, d_{max}]$ |
-| `+   n_div = 1,`                                        |                                        Number of divisions |                  $N_{div}$ |
-| `+   strat_breaks = seq(0, 1000, by = 40))`             |                 Series of depth breaks for defining strata |                            |
-| `>`                                                     |                                                            |                            |
-| `> distribution <- sim_distribution(`                   |                                    *Simulate distribution* |                            |     
-| `+   sim = abundance,`                                  |                Object produced by `sim_abundance` function |                            |     
-| `+   grid = grid,`                                      |                    Object produced by `make_grid` function |                            |
-| `+   ays_covar = sim_ays_covar(sd = 2.8,`               |          Standard deviation of age-year-space distribution |             $\sigma_{\xi}$ |
-| `+                             range = 300,`            |                          Range of spatial correlation (km) |                        $r$ |
-| `+                             phi_age = 0.5,`          |            Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |
-| `+                             phi_year = 0.9,`         |           Correlation across years in spatial distribution |      $\varphi_{\xi, year}$ |
-| `+                             group_ages = 5:20),`     |       Make space-age-year variance equal across these ages |                            |
-| `+   depth_par = sim_parabola(mu = 200,`                |          Depth at which abundance is typically highest (m) |                    $\mu_d$ |
-| `+                            sigma = 70))`             |              Dispersion around depth of peak abundance (m) |                 $\sigma_d$ |
-| `>`                                                     |                                                            |                            |
-| `> surveys <- expand_surveys(`                          |                       *Define a series of surveys to test* |                            |
-| `+   set_den = c(0.0005, 0.001, 0.002, 0.005, 0.01),`   |                                       Set density (km^-2^) |                 $D_{sets}$ |
-| `+   lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),`   |                 Maximum number of lengths to collect / set |              $M_{lengths}$ |
-| `+   ages_cap = c(2, 5, 10, 20, 50))`                   | Maximum number of ages to sample / length group / division |                 $M_{ages}$ |
-| `>`                                                     |                                                            |                            |
-| `> tests <- test_surveys(`                              |                             *Simulate and analyze surveys* |                            |
-| `+   sim = distribution,`                               |             Object produced by `sim_distribution` function |                            |
-| `+   surveys = surveys,`                                |               Object produced by `expand_surveys` function |                            |
-| `+   n_loops = 1000,`                                   |                    Number of times to simulate each survey |                            | 
-| `+   q = sim_logistic(k = 2,`                           |                Steepness of logistic curve of catchability |                        $k$ |
-| `+                   x0 = 3),`                          |                 Midpoint of logistic curve of catchability |                      $x_0$ |
-| `+   trawl_dim = c(1.5, 0.02),`                         |       Trawl dimensions (km) - i.e. area covered by a trawl |                $A_{trawl}$ |
-| `+   length_group = 1)`                                 |                         Length group for age sampling (cm) |                $l_{group}$ |
-
-*Caution:* as written, the `test_surveys` function call may take days or weeks to run.
-Run duration can be decreased by running multiple simulations per loop (`n_sims` argument) and by running the loops in parallel (`cores` argument).
-See help file for more details (`?test_surveys`).
+| Function | Description                   |
+|:---------|:------------------------------|
+| **`sim_abundance`**  | Simulate a basic age-structured population dynamics model |
+| `sim_R`, `sim_Z`, `sim_N0`, `sim_vonB` | Closures, to use inside `sim_abundance`, for simulating recruitment, total mortality, initial abundance and growth, respectively | 
+| **`sim_distribution`** | Simulate spatial and temporal distribution of an age-structured population |
+| `sim_ays_covar`, `sim_parabola` | Closures, to use inside `sim_distribution`, for simulating age-year-space covariance and parabolic relationships with covariates (e.g. depth), respectively |
+| `make_grid` | Make a basic depth stratified square grid to use inside `sim_distribution` |
+| **`sim_survey`** | Simulate a survey of a spatial, age-structured population |
+| `sim_logistic` | Closure, to use inside `sim_survey`, for simulating age-specific catchability as a logistic curve |
+| **`run_strat`** | Run a stratified analysis on simulated survey data |
+| **`strat_error`** | Calculate the error of stratified estimates (e.g. root mean squared error of stratified estimates from true values) |
+| **`test_surveys`** | Test the sampling design of multiple surveys using a stratified analysis (internally loops over `sim_survey`, `run_strat` and `strat_error`) |
+| `expand_surveys` | Create a data frame, for use in `test_surveys`,  with all combinations of supplied survey settings |
 
 <br>
+
+TODO:
+> increase the margins on the table in the template
+> re-read the initial SimSurvey section and expand as necessary to better allign with the table
+> in each section, expand upon the details of the core functions and the associated equation symbols (i.e. break down case-study table into several tables)
+> in the case study section, copy and paste (and clean-up) the code used to generate the case study results
 
@@
 
-# Appendix 1: case study results and discussion
+# Appendix 1: case study
+
+The construction of the **`SimSurvey`** package was motivated by a regional need to assess the sampling design of trawl surveys conducted along the Newfoundland and Labrador shelf by Fisheries and Oceans Canada.
+As a tangible first step, we focused our efforts on emulating data from the survey of cod in NAFO division 3Ps since it is one of the most dynamic and heavily sampled stocks in the region.
+Function settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data; these values were set as the defaults for the main functions in the package.
+Here we evaluate the efficacy of alternate sampling protocol by assessing deviation between stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
+Although assessing model-based analyses was an option, we focused on the design-based stratified analysis for simplicity and because of its widespread use.
+See Table \@ref(tab:par) for a description of the code used to generate the case study results.
+
+Table: (\#tab:par) Main function calls, with argument descriptions and associated parameter symbols, used to simulate the cod-like population used in the case study presented here.
+Note that the settings below largely correspond with the defaults used in the functions.
+
+|                                       **Function call** |                                            **Description** |                 **Symbol** |
+|:------------------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
+| `> set.seed(438)`                                       |      Set random number generator seed to replicate results |                            |
+| `>`                                                     |                                                            |                            |
+| `> abundance <- sim_abundance(`                         |                                       *Simulate abundance* |                            |
+| `+   ages = 1:20,`                                      |                                                       Ages |                        $a$ |
+| `+   years = 1:20,`                                     |                                                      Years |                        $y$ |
+| `+   R = sim_R(mean = 30000000,`                        |                                           Mean recruitment |                    $\mu_r$ |
+| `+             log_sd = 0.5),`                          |                      Standard deviation of log-recruitment |                 $\sigma_r$ |
+| `+   Z = sim_Z(mean = 0.5,`                             |                                       Mean total mortality |                    $\mu_Z$ |
+| `+             log_sd = 0.2,`                           |                      Standard deviation of total mortality |                 $\sigma_Z$ |
+| `+             phi_age = 0.9,`                          |    Correlation across ages in error around total mortality |    $\varphi_{\delta, age}$ |
+| `+             phi_year = 0.5),`                        |   Correlation across years in error around total mortality |   $\varphi_{\delta, year}$ |
+| `+   growth = sim_vonB(Linf = 120,`                     |                                Mean asymptotic length (cm) |               $l_{\infty}$ |
+| `+                     L0 = 5,`                         |                                       Length at birth (cm) |                      $l_0$ |
+| `+                     K = 0.1,`                        |                                                Growth rate |                        $K$ |
+| `+                     log_sd = 0.1,`                   |     Standard deviation of the von Bertalanffy growth curve |                 $\sigma_l$ |
+| `+                     length_group = 3))`              |                  Length group for abundance at length (cm) |            $l_{group}^{N}$ |      
+| `>`                                                     |                                                            |                            |
+| `> grid <- make_grid(`                                  |                                              *Make a grid* |                            |
+| `+   x_range = c(-140, 140),`                           |                           Range of grid in the x dimension |                            |
+| `+   y_range = c(-140, 140),`                           |                           Range of grid in the y dimension |                            | 
+| `+   res = c(3.5, 3.5),`                                |                                       Grid resolution (km) |                        $R$ |
+| `+   shelf_depth = 200,`                                |                                           Shelf depth (km) |                $d_{shelf}$ |
+| `+   shelf_width = 100,`                                |                                           Shelf width (km) |                $w_{shelf}$ |
+| `+   depth_range = c(0, 1000),`                         |                                            Depth range (m) |       $[d_{min}, d_{max}]$ |
+| `+   n_div = 1,`                                        |                                        Number of divisions |                  $N_{div}$ |
+| `+   strat_breaks = seq(0, 1000, by = 40))`             |                 Series of depth breaks for defining strata |                            |
+| `>`                                                     |                                                            |                            |
+| `> distribution <- sim_distribution(`                   |                                    *Simulate distribution* |                            |     
+| `+   sim = abundance,`                                  |                Object produced by `sim_abundance` function |                            |     
+| `+   grid = grid,`                                      |                    Object produced by `make_grid` function |                            |
+| `+   ays_covar = sim_ays_covar(sd = 2.8,`               |          Standard deviation of age-year-space distribution |             $\sigma_{\xi}$ |
+| `+                             range = 300,`            |                          Range of spatial correlation (km) |                        $r$ |
+| `+                             phi_age = 0.5,`          |            Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |
+| `+                             phi_year = 0.9,`         |           Correlation across years in spatial distribution |      $\varphi_{\xi, year}$ |
+| `+                             group_ages = 5:20),`     |       Make space-age-year variance equal across these ages |                            |
+| `+   depth_par = sim_parabola(mu = 200,`                |          Depth at which abundance is typically highest (m) |                    $\mu_d$ |
+| `+                            sigma = 70))`             |              Dispersion around depth of peak abundance (m) |                 $\sigma_d$ |
+| `>`                                                     |                                                            |                            |
+| `> surveys <- expand_surveys(`                          |                       *Define a series of surveys to test* |                            |
+| `+   set_den = c(0.0005, 0.001, 0.002, 0.005, 0.01),`   |                                       Set density (km^-2^) |                 $D_{sets}$ |
+| `+   lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),`   |                 Maximum number of lengths to collect / set |              $M_{lengths}$ |
+| `+   ages_cap = c(2, 5, 10, 20, 50))`                   | Maximum number of ages to sample / length group / division |                 $M_{ages}$ |
+| `>`                                                     |                                                            |                            |
+| `> tests <- test_surveys(`                              |                             *Simulate and analyze surveys* |                            |
+| `+   sim = distribution,`                               |             Object produced by `sim_distribution` function |                            |
+| `+   surveys = surveys,`                                |               Object produced by `expand_surveys` function |                            |
+| `+   n_loops = 1000,`                                   |                    Number of times to simulate each survey |                            | 
+| `+   q = sim_logistic(k = 2,`                           |                Steepness of logistic curve of catchability |                        $k$ |
+| `+                   x0 = 3),`                          |                 Midpoint of logistic curve of catchability |                      $x_0$ |
+| `+   trawl_dim = c(1.5, 0.02),`                         |       Trawl dimensions (km) - i.e. area covered by a trawl |                $A_{trawl}$ |
+| `+   length_group = 1)`                                 |                         Length group for age sampling (cm) |                $l_{group}$ |
+
+*Caution:* as written, the `test_surveys` function call may take days or weeks to run.
+Run duration can be decreased by running multiple simulations per loop (`n_sims` argument) and by running the loops in parallel (`cores` argument).
+See help file for more details (`?test_surveys`).
+
+<br>
 

### 2019-07-03 analysis/paper/sim_survey_paper.Rmd
Split the case study table into tables on the core functions

@@
 
-The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for simulating the abundance and distribution of virtural fish populations with correlation across space, time and age dimensions, as well as functions for simulating surveys with a range of sampling stratigies and intensities (Table \@ref(tab:fun)).
+The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtural fish populations with correlation across space, time and age dimensions, 2) simulating surveys with a range of sampling stratigies and intensities, and 3) running stratified analyses of simulated survey data (Table \@ref(tab:fun)).
 The equations behind these functions are detailed in the sections below.
@@
 TODO:
-> increase the margins on the table in the template
-> re-read the initial SimSurvey section and expand as necessary to better allign with the table
 > in each section, expand upon the details of the core functions and the associated equation symbols (i.e. break down case-study table into several tables)
@@
 
-Abundance is simulated using the `sim_abundance` function and the settings used for our case study are shown in Table \@ref(tab:par).
-These settings are also the defaults of the function.
+Table: (\#tab:abundance) Default `sim_abundance` function call, with argument descriptions and associated parameter symbols.
+
+| **Function call**                           | **Description**                                            | **Symbol**                 |
+|:------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
+| `sim_abundance(`                            | &nbsp;                                                     | &nbsp;                     |
+| `+  ages = 1:20,`                           | Ages                                                       | $a$                        |
+| `+  years = 1:20,`                          | Years                                                      | $y$                        |
+| `+  R = sim_R(mean = 30000000,`             | Mean recruitment                                           | $\mu_r$                    |
+| `+            log_sd = 0.5),`               | Standard deviation of log-recruitment                      | $\sigma_r$                 |
+| `+  Z = sim_Z(mean = 0.5,`                  | Mean total mortality                                       | $\mu_Z$                    |
+| `+            log_sd = 0.2,`                | Standard deviation of total mortality                      | $\sigma_Z$                 |
+| `+            phi_age = 0.9,`               | Correlation across ages in error around total mortality    | $\varphi_{\delta, age}$    |
+| `+            phi_year = 0.5),`             | Correlation across years in error around total mortality   | $\varphi_{\delta, year}$   |
+| `+  growth = sim_vonB(Linf = 120,`          | Mean asymptotic length (cm)                                | $l_{\infty}$               |
+| `+                    L0 = 5,`              | Length at birth (cm)                                       | $l_0$                      |
+| `+                    K = 0.1,`             | Growth rate                                                | $K$                        |
+| `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_l$                 |
+| `+                    length_group = 3))`   | Length group for abundance at length (cm)                  | $l_{group}^{N}$            |
+
+<br>
+
+Abundance is simulated using the `sim_abundance` function and a default function call is described in Table \@ref(tab:abundance).
 This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`).
@@
 
+Table: (\#tab:distribution) Default `sim_distribution` function call, with argument descriptions and associated parameter symbols.
+
+| **Function call**                                       | **Description**                                       | **Symbol**            |
+|:------------------------------------------------------- |:----------------------------------------------------- | :-------------------- |
+| `sim_distribution(`                                     | &nbsp;                                                | &nbsp;                |
+| `+  sim,`                                               | Simulated population from `sim_abundance`             | &nbsp;                |
+| `+  grid = make_grid(x_range = c(-140, 140),`           | Range of grid in the x dimension                      | &nbsp;                |
+| `+                   y_range = c(-140, 140),`           | Range of grid in the y dimension                      | &nbsp;                |
+| `+                   res = c(3.5, 3.5),`                | Grid resolution (km)                                  | $R$                   |
+| `+                   shelf_depth = 200,`                | Shelf depth (km)                                      | $d_{shelf}$           |
+| `+                   shelf_width = 100,`                | Shelf width (km)                                      | $w_{shelf}$           |
+| `+                   depth_range = c(0, 1000),`         | Depth range (m)                                       | $[d_{min}, d_{max}]$  |
+| `+                   n_div = 1,`                        | Number of divisions                                   | $N_{div}$             |
+| `+                   strat_breaks = seq(0, 1000, 40)),` | Series of depth breaks for defining strata            | &nbsp;                |
+| `+  ays_covar = sim_ays_covar(sd = 2.8,`                | Standard deviation of age-year-space distribution     | $\sigma_{\xi}$        |
+| `+                            range = 300,`             | Range of spatial correlation (km)                     | $r$                   |
+| `+                            phi_age = 0.5,`           | Correlation across ages in spatial distribution       | $\varphi_{\xi, age}$  |
+| `+                            phi_year = 0.9,`          | Correlation across years in spatial distribution      | $\varphi_{\xi, year}$ |
+| `+                            group_ages = 5:20),`      | Make space-age-year variance equal across these ages  | &nbsp;                |
+| `+  depth_par = sim_parabola(mu = 200,`                 | Depth at which abundance is typically highest (m)     | $\mu_d$               |
+| `+                           sigma = 70))`              | Dispersion around depth of peak abundance (m)         | $\sigma_d$            |
+
+<br>
+
 The `make_grid` function is used to generate a survey grid for use in the `sim_distribution` function.
@@
 The `sim_distribution` function distributes the population simulated using `sim_abundance` throughout the grid generated using `make_grid`.
-Table \@ref(tab:par) outlines the default settings used for the function and the case study.
+Table \@ref(tab:distribution) outlines the default settings used for the function.
 In addition to supplying objects produced by `sim_abundance` and `make_grid`, this function requires two closures that describe the age-year-space covariance and the relationship with depth.
@@
 
+Table: (\#tab:survey) Default `sim_survey` function call, with descriptions and associated parameter symbols of key arguments.
+
+| **Function call**                   | **Description**                                                         | **Symbol**     |
+|:----------------------------------- |:----------------------------------------------------------------------- | :------------- |
+| `sim_survey(`                       | &nbsp;                                                                  | &nbsp;         |
+| `+  sim,`                           | Simulated spatial population from `sim_distribution`                    | &nbsp;         |
+| `+  n_sims = 1`                     | Number of times to repeat the survey                                    | &nbsp;         |
+| `+  q = sim_logistic(k = 2,`        | Steepness of logistic curve of catchability                             | $k$            |
+| `+                   x0 = 3),`      | Midpoint of logistic curve of catchability                              | $x_0$          |
+| `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (length, width; km) - i.e. area covered by a trawl     | $A_{trawl}$    |
+| `+  min_sets = 2`                   | Minimum number of sets to conduct per strata                            | &nbsp;         |
+| `+  set_den = 2/1000,`              | Set density (km^-2^)                                                    | $D_{sets}$     |
+| `+  lengths_cap = 500,`             | Maximum number of lengths to collect / set                              | $M_{lengths}$  |
+| `+  length_group = 1,`              | Length group for age sampling (cm)                                      | $l_{group}$    |
+| `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / division              | $M_{ages}$     |
+| `+  age_sammpling = "stratified")`  | Controls whether age sampling is length `"stratified"` or `"random"`    | &nbsp;         |
+
+<br>
+
 The function `sim_survey` can be used to simulate data from one survey over a population simulated using `sim_distribution`.
+A default function call is described in Table \@ref(tab:survey)).
 This function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and definitions of the design of the survey.
@@
 Multiple simulations of the same survey can be run using the `n_sims` argument, however, it is important not to request large numbers of simulations because this function can quickly deplete a computers RAM.
-Defaults used in the function and for the case study roughly correspond to the assumed catchability in NAFO division 3Ps as well as the sampling protocol used in the survey of this stock (Table \@ref(tab:par)).
-Below we use `sim_survey` to simulate two surveys over our case study population, of which one is set-up to have higher sampling effort than the other.
+Below we use `sim_survey` to simulate two surveys over a default population, of which one is set-up to have higher sampling effort than the other.
 
@@
 
-# Appendix 1: case study
+# Appendix 1: Case study
 
@@
 Although assessing model-based analyses was an option, we focused on the design-based stratified analysis for simplicity and because of its widespread use.
-See Table \@ref(tab:par) for a description of the code used to generate the case study results.
+The code below will replicate the results of this caset study:
 
-Table: (\#tab:par) Main function calls, with argument descriptions and associated parameter symbols, used to simulate the cod-like population used in the case study presented here.
-Note that the settings below largely correspond with the defaults used in the functions.
+```{r, echo = TRUE, eval = FALSE}
 
-|                                       **Function call** |                                            **Description** |                 **Symbol** |
-|:------------------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
-| `> set.seed(438)`                                       |      Set random number generator seed to replicate results |                            |
-| `>`                                                     |                                                            |                            |
-| `> abundance <- sim_abundance(`                         |                                       *Simulate abundance* |                            |
-| `+   ages = 1:20,`                                      |                                                       Ages |                        $a$ |
-| `+   years = 1:20,`                                     |                                                      Years |                        $y$ |
-| `+   R = sim_R(mean = 30000000,`                        |                                           Mean recruitment |                    $\mu_r$ |
-| `+             log_sd = 0.5),`                          |                      Standard deviation of log-recruitment |                 $\sigma_r$ |
-| `+   Z = sim_Z(mean = 0.5,`                             |                                       Mean total mortality |                    $\mu_Z$ |
-| `+             log_sd = 0.2,`                           |                      Standard deviation of total mortality |                 $\sigma_Z$ |
-| `+             phi_age = 0.9,`                          |    Correlation across ages in error around total mortality |    $\varphi_{\delta, age}$ |
-| `+             phi_year = 0.5),`                        |   Correlation across years in error around total mortality |   $\varphi_{\delta, year}$ |
-| `+   growth = sim_vonB(Linf = 120,`                     |                                Mean asymptotic length (cm) |               $l_{\infty}$ |
-| `+                     L0 = 5,`                         |                                       Length at birth (cm) |                      $l_0$ |
-| `+                     K = 0.1,`                        |                                                Growth rate |                        $K$ |
-| `+                     log_sd = 0.1,`                   |     Standard deviation of the von Bertalanffy growth curve |                 $\sigma_l$ |
-| `+                     length_group = 3))`              |                  Length group for abundance at length (cm) |            $l_{group}^{N}$ |      
-| `>`                                                     |                                                            |                            |
-| `> grid <- make_grid(`                                  |                                              *Make a grid* |                            |
-| `+   x_range = c(-140, 140),`                           |                           Range of grid in the x dimension |                            |
-| `+   y_range = c(-140, 140),`                           |                           Range of grid in the y dimension |                            | 
-| `+   res = c(3.5, 3.5),`                                |                                       Grid resolution (km) |                        $R$ |
-| `+   shelf_depth = 200,`                                |                                           Shelf depth (km) |                $d_{shelf}$ |
-| `+   shelf_width = 100,`                                |                                           Shelf width (km) |                $w_{shelf}$ |
-| `+   depth_range = c(0, 1000),`                         |                                            Depth range (m) |       $[d_{min}, d_{max}]$ |
-| `+   n_div = 1,`                                        |                                        Number of divisions |                  $N_{div}$ |
-| `+   strat_breaks = seq(0, 1000, by = 40))`             |                 Series of depth breaks for defining strata |                            |
-| `>`                                                     |                                                            |                            |
-| `> distribution <- sim_distribution(`                   |                                    *Simulate distribution* |                            |     
-| `+   sim = abundance,`                                  |                Object produced by `sim_abundance` function |                            |     
-| `+   grid = grid,`                                      |                    Object produced by `make_grid` function |                            |
-| `+   ays_covar = sim_ays_covar(sd = 2.8,`               |          Standard deviation of age-year-space distribution |             $\sigma_{\xi}$ |
-| `+                             range = 300,`            |                          Range of spatial correlation (km) |                        $r$ |
-| `+                             phi_age = 0.5,`          |            Correlation across ages in spatial distribution |       $\varphi_{\xi, age}$ |
-| `+                             phi_year = 0.9,`         |           Correlation across years in spatial distribution |      $\varphi_{\xi, year}$ |
-| `+                             group_ages = 5:20),`     |       Make space-age-year variance equal across these ages |                            |
-| `+   depth_par = sim_parabola(mu = 200,`                |          Depth at which abundance is typically highest (m) |                    $\mu_d$ |
-| `+                            sigma = 70))`             |              Dispersion around depth of peak abundance (m) |                 $\sigma_d$ |
-| `>`                                                     |                                                            |                            |
-| `> surveys <- expand_surveys(`                          |                       *Define a series of surveys to test* |                            |
-| `+   set_den = c(0.0005, 0.001, 0.002, 0.005, 0.01),`   |                                       Set density (km^-2^) |                 $D_{sets}$ |
-| `+   lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),`   |                 Maximum number of lengths to collect / set |              $M_{lengths}$ |
-| `+   ages_cap = c(2, 5, 10, 20, 50))`                   | Maximum number of ages to sample / length group / division |                 $M_{ages}$ |
-| `>`                                                     |                                                            |                            |
-| `> tests <- test_surveys(`                              |                             *Simulate and analyze surveys* |                            |
-| `+   sim = distribution,`                               |             Object produced by `sim_distribution` function |                            |
-| `+   surveys = surveys,`                                |               Object produced by `expand_surveys` function |                            |
-| `+   n_loops = 1000,`                                   |                    Number of times to simulate each survey |                            | 
-| `+   q = sim_logistic(k = 2,`                           |                Steepness of logistic curve of catchability |                        $k$ |
-| `+                   x0 = 3),`                          |                 Midpoint of logistic curve of catchability |                      $x_0$ |
-| `+   trawl_dim = c(1.5, 0.02),`                         |       Trawl dimensions (km) - i.e. area covered by a trawl |                $A_{trawl}$ |
-| `+   length_group = 1)`                                 |                         Length group for age sampling (cm) |                $l_{group}$ |
+set.seed(438)
+
+abundance <- sim_abundance(
+  ages = 1:20,
+  years = 1:20,
+  R = sim_R(mean = 30000000,
+            log_sd = 0.5),
+  Z = sim_Z(mean = 0.5,
+            log_sd = 0.2,
+            phi_age = 0.9,
+            phi_year = 0.5),
+  growth = sim_vonB(Linf = 120, 
+                    L0 = 5, 
+                    K = 0.1,
+                    log_sd = 0.1, 
+                    length_group = 3)
+)
+
+distribution <- sim_distribution(
+  abundance,
+  grid = make_grid(x_range = c(-140, 140),
+                   y_range = c(-140, 140),
+                   res = c(3.5, 3.5),
+                   shelf_depth = 200,
+                   shelf_width = 100,
+                   depth_range = c(0, 1000),
+                   n_div = 1,
+                   strat_breaks = seq(0, 1000, by = 40),
+                   strat_splits = 2),
+  ays_covar = sim_ays_covar(sd = 2.8,
+                            range = 300,
+                            phi_age = 0.5,
+                            phi_year = 0.9,
+                            group_ages = 5:20),
+  depth_par = sim_parabola(mu = 200,
+                           sigma = 70)
+)
+
+surveys <- test_surveys(
+  distribution,
+  surveys = expand_surveys(set_den = c(0.5, 1, 2, 5, 10) / 1000,
+                           lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
+                           ages_cap = c(2, 5, 10, 20, 50)),
+  n_loops = 1000,
+  q = sim_logistic(k = 2, x0 = 3)
+)
+
+```
 
@@
 See help file for more details (`?test_surveys`).
-
-<br>
 
@@
 
-# Appendix 2: age-year-space covariance
+# Appendix 2: Age-year-space covariance
 

### 2019-07-04 analysis/paper/sim_survey_paper.Rmd
Address the remainder of Greg's comments (except suggestion for new plot that compares real and simulated survey data; work on that next)

@@
 
-TODO:
-> in each section, expand upon the details of the core functions and the associated equation symbols (i.e. break down case-study table into several tables)
-> in the case study section, copy and paste (and clean-up) the code used to generate the case study results
-
 ## Simulate abundance
@@
 
-Table: (\#tab:abundance) Default `sim_abundance` function call, with argument descriptions and associated parameter symbols.
+Table: (\#tab:abundance) Default `sim_abundance` function call, with descriptions and associated parameter symbols of key arguments.
 
@@
 | `+  Z = sim_Z(mean = 0.5,`                  | Mean total mortality                                       | $\mu_Z$                    |
-| `+            log_sd = 0.2,`                | Standard deviation of total mortality                      | $\sigma_Z$                 |
+| `+            log_sd = 0.2,`                | Standard deviation of total mortality (log)                | $\sigma_Z$                 |
 | `+            phi_age = 0.9,`               | Correlation across ages in error around total mortality    | $\varphi_{\delta, age}$    |
@@
 | `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_l$                 |
-| `+                    length_group = 3))`   | Length group for abundance at length (cm)                  | $l_{group}^{N}$            |
+| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{group}^{N}$            |
 
@@
 Overall, the function provides a simple tool for simulating a range dynamic age-structured populations.
-For instance, below we simulate a relatively long (default case study settings) and short lived species.
+For instance, below we simulate a relatively long and short lived species (with default variance settings).
 
@@
 For instance, the `plot_surface` function can be used to make quick visuals of matrices contained within the list returned by `sim_abundance`.
-The plots output from the below code are combined into Figure \@ref(fig:surface).
-Here we display the abundance at age matrix (object named `N` in the list produced by `sim_abundance`); other names can be supplied to the `mat` argument to visualize a different matrix from the `sim_abundance` list, such as `Z`.
+As an example, we display the abundance at age matrix (object named `N` in the list produced by `sim_abundance`; Figure \@ref(fig:surface)); other names can be supplied to the `mat` argument to visualize a different matrix from the `sim_abundance` list, such as `Z`.
 
@@
 
-```{r surface, fig.cap = "Surface plots of simulated abundance at age of a relatively a) long lived (case study settings) and b) short lived species.
+```{r surface, fig.cap = "Surface plots of simulated abundance at age of a relatively a) long lived and b) short lived species.
 These plots were produced by `plot_surface` when supplied a list produced by `sim_abundance`."}
@@
 The next step in the simulation is to distribute the abundance at age matrix simulated using the cohort model throughout a spatial field.
-Here, a grid of $s$ locations was generated with a resolution of $R$ and a depth gradient.
-The depth gradient was set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$.
+Here, a grid of $s$ locations is generated with a resolution of $R$ and a depth gradient.
+The depth gradient is set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$.
 The grid is divided into $N_{div}$ divisions and $N_{strat}$ depth-based strata.
@@
 
-Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean.
+Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean depth.
 This helps simulate depth preferences of particular fish.
-Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
-Spatial correlations was controlled by a smoothing ($\lambda$) and scaling parameter ($\kappa$) [here we approximate $\kappa$ from a range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years was controlled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively.
-The overall variance of the process was controlled by $\sigma_{\xi}^2$.
+Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and the two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
+Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively.
+The overall variance of the process is controlled by $\sigma_{\xi}^2$.
 In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering.
@@
 A more detailed description of the space-age-year covariance is included in Appendix 2.
-Note that the the distributions of specific length groups were not simulated because of the greater computational cost of storing spatially dis-aggregated abundance across all length groups.
-
-Table: (\#tab:distribution) Default `sim_distribution` function call, with argument descriptions and associated parameter symbols.
+
+Table: (\#tab:distribution) Default `sim_distribution` function call, with descriptions and associated parameter symbols of key arguments.
 
@@
 | `+                            range = 300,`             | Range of spatial correlation (km)                     | $r$                   |
+| `+                            lambda = 1,`              | Smoothness of spatial correlation                     | $\lambda$             |
 | `+                            phi_age = 0.5,`           | Correlation across ages in spatial distribution       | $\varphi_{\xi, age}$  |
@@
 The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`.
-If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied to `sim_distribution` if the same structure is used.
+If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure is used.
 The package includes a manually constructed survey grid of NAFO division 3Ps (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid.
-However, for simplicity, we use `make_grid` to construct a square grid for our case study and the settings used are shown in Table \@ref(tab:par).
-Below we generate and plot (combined in Figure \@ref(fig:grid)) a default case study grid and another grid with four divisions and a more linear gradient in depth.
+However, for simplicity, we use `make_grid` to construct a square grid for a default run of `sim_distribution`.
+Below we generate and plot (Figure \@ref(fig:grid)) a default grid, another grid with the number of strata increased by increasing the number of `strat_splits`, and another with four divisions and a more linear gradient in depth.
 
@@
 
-a <- make_grid(n_div = 1, shelf_depth = 200,
+a <- make_grid(n_div = 1, strat_splits = 2, shelf_depth = 200,
                shelf_width = 100, depth_range = c(0, 1000))
-b <- make_grid(n_div = 4, shelf_depth = 500, 
+b <- make_grid(n_div = 1, strat_splits = 3, shelf_depth = 200, 
+               shelf_width = 100, depth_range = c(0, 1000))
+c <- make_grid(n_div = 4, strat_splits = 1, shelf_depth = 500, 
                shelf_width = 0, depth_range = c(0, 1000))
@@
 plot_grid(b)
+plot_grid(c)
 
@@
   add_labels(x = 0, y = 1, text = "a)")
-b <- make_grid(n_div = 4, shelf_depth = 500, shelf_width = 0, depth_range = c(0, 1000)) %>% 
+b <- make_grid(strat_splits = 3) %>% 
   plot_grid(showscale = FALSE) %>% 
-  add_labels(x = 0.01, y = 1, text = "b)")
-p <- subplot(a, b)
-
-export_plot(p, file = "analysis/paper/figures/plot_grid.png", width = 800, height = 430)
-
-```
-
-```{r grid, fig.cap = "Plots produced by `plot_grid` when supplied an raster object produced by `make_grid` a) using default case study settings and b) settings generates a grid with four divisions and a more linear depth gradient.
-In these plots, the color gradient represents depth, the grey lines delineate divisions and white lines delineate strata."}
+  add_labels(x = 0.04, y = 1, text = "b)")
+c <- make_grid(n_div = 4, shelf_depth = 500, shelf_width = 0, depth_range = c(0, 1000)) %>% 
+  plot_grid(showscale = FALSE) %>% 
+  add_labels(x = 0.08, y = 1, text = "c)")
+p <- subplot(a, b, c)
+
+export_plot(p, file = "analysis/paper/figures/plot_grid.png", width = 800, height = 320)
+
+```
+
+```{r grid, fig.cap = "Plots produced by `plot_grid` when supplied a raster object produced by `make_grid` a) using default settings, b) settings that increase the number of times depth strata are horizontally split, and c) settings that produce a more linear depth gradient and increase the number of divisions.
+In these plots, the color gradient represents depth, the thick grey lines delineate divisions and thin white lines delineate strata."}
 
@@
 Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years`, respectively, arguments in the `sim_ays_covar` closure.
+Variance in the size of the clusters can also be modified by changing the `sd` argument in the `sim_ays_covar` function.
 In other words, these parameters can be modified to control the degree of age-specific clustering and inter-annual site-fidelity exhibited by the simulated population.
@@
 set.seed(438)
-a <- sim_abundance() %>% 
-  sim_distribution(ays_covar = sim_ays_covar(range = 300,
+a <- sim_distribution(sim = sim_abundance(), # nested approach
+                      ays_covar = sim_ays_covar(range = 300,
                                                 phi_year = 0.9,
                                                 phi_age = 0.5))
-b <- sim_abundance() %>% 
+b <- sim_abundance() %>%  # pipe approach
   sim_distribution(ays_covar = sim_ays_covar(range = 2000,
@@
 The `plot_distribution` function can be used to provide a quick visual of the distribution across ages and years.
-The code below will generate interactive plots with an Age-Year slider, however, for this paper we present a facet plot of the simulated data in Figure \@ref(fig:distribution).
+The code below will generate interactive plots with an Age-Year slider, however, for this paper we present a facet plot of the simulated data (Figure \@ref(fig:distribution)).
 
@@
 
-```{r distribution, fig.cap = "Distribution plots of simulated populations that form a) tight clusters with stronger correlation through years than ages (case study settings), and b) relatively diffuse clusters with stronger correlation through ages than years.
+```{r distribution, fig.cap = "Distribution plots of simulated populations that form a) tight clusters with stronger correlation through years than ages (default settings), and b) relatively diffuse clusters with stronger correlation through ages than years.
 This plot is a facet of plots produced by `plot_distribution` when supplied simulations from `sim_distribution`."}
@@
 
-The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated using formula described above.
-The area of each strata is calculated and this is used to define the sets to allocate to each strata under a particular set density, $D_{sets}$.
+The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated.
+The area of each strata is calculated and this is used to define the number of sets to allocate to each strata under a particular set density, $D_{sets}$.
 The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:
@@
 $$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{trawl}}{A_{cell}} q_a )$$
-Where $n_{a,y,s}$ is the number of fish sampled by a set, $A_{trawl}$ indicates the area covered by the trawl, $A_{cell}$ is the area of a grid cell (i.e.
-$R^2$), and $q_a$ is the catchability coefficient of each age.
+Where $n_{a,y,s}$ is the number of fish sampled by a set, $A_{trawl}$ indicates the area covered by the trawl, $A_{cell}$ is the area of a grid cell, and $q_a$ is the catchability coefficient of each age.
 Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
@@
 Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division.
-Such sub-sampling is common in fisheries-independent surveys as it is costly to sample every fish sampled.
+Such sub-sampling is common in fisheries-independent surveys as it is costly to sample every fish captured.
 Age determination is especially time-consuming, which is why age sampling tends to be length-stratified.
@@
 | `+  lengths_cap = 500,`             | Maximum number of lengths to collect / set                              | $M_{lengths}$  |
-| `+  length_group = 1,`              | Length group for age sampling (cm)                                      | $l_{group}$    |
+| `+  length_group = 1,`              | Length group bin size for age sampling (cm)                             | $l_{group}$    |
 | `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / division              | $M_{ages}$     |
@@
 
-The function `sim_survey` can be used to simulate data from one survey over a population simulated using `sim_distribution`.
+The function `sim_survey` can be used to simulate data from one survey over a population created using `sim_distribution`.
 A default function call is described in Table \@ref(tab:survey)).
-This function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and definitions of the design of the survey.
+The `sim_survey` function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and definitions of the design of the survey.
 Specifically, the `q` argument requires a closure, such as `sim_logistic`, for defining the probability of catching specific age groups, trawl dimensions are defined in the `trawl_dim` argument, and set, length and age sampling effort are defined using the `set_den`, `lengths_cap` and `ages_cap` arguments, respectively.
-Multiple simulations of the same survey can be run using the `n_sims` argument, however, it is important not to request large numbers of simulations because this function can quickly deplete a computers RAM.
+Multiple simulations of the same survey can be run using the `n_sims` argument, however, requesting large numbers of simulations can be computational demanding depending on the processing capacity available.
 Below we use `sim_survey` to simulate two surveys over a default population, of which one is set-up to have higher sampling effort than the other.
@@
 Catchability corrected abundance matrices, named `I` and `I_at_length`, are also produced and added to the output; these matrices are useful for comparing the true abundance available to the survey to abundance estimates obtained using design-based or model-based analyses of the simulated survey data.
-Specific surveys can be explored using the `plot_survey` function, which uses plotly [@sievert2018] and crosstalk [@cheng2016] in the background to link the bubble plot of aggregate set catch to the histogram of lengths and ages sampled to facilitate explorations of set-specific catches.
-The plots generated below have been combined here into a static plot for this paper (Figure \@ref(fig:survey)).
+Specific surveys can be explored using the `plot_survey` function, which uses plotly [@sievert2018] and crosstalk [@cheng2016] in the background to link the bubble plot of aggregate set catch to the histogram of lengths and ages sampled to facilitate explorations of set-specific catches (Figure \@ref(fig:survey)).
 
@@
 
-```{r survey, fig.cap = "Bubble plots and histograms of set catches from a simulated stratified-random survey of the case study population under relatively a) high and b) low sampling effort.
+```{r survey, fig.cap = "Bubble plots and histograms of set catches from a simulated stratified-random survey of a default population under relatively a) high and b) low sampling effort.
 Histograms of length and age composition include the distribution caught and sampled overlaid with a line of the true distribution of length and ages available to the survey.
@@
 
-As noted above, the amount of RAM a user's computer has may limit the utility of the `sim_survey` function for running thousands of simulations of the same survey.
+As noted above, available RAM may limit the utility of the `sim_survey` function for running thousands of simulations of the same survey.
 The `sim_survey_parallel` function was therefore constructed to facilitate the running of thousands of simulations without running into RAM limitations.
@@
 Here we apply formula presented in Smith and Somerton [-@Smith1981] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
-Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be bumped up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key.
+Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be scaled up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key.
 We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:
@@
 RMSE of the stratified estimates can then be calculated using the `strat_error` function.
-Results and error of a stratified analysis of one default survey over a default population are obtained using the following code:
+Results and error of a stratified analysis of one survey over a population are obtained using the following code (using default values):
 
@@
 Assuming a stratified analysis as the default method for obtaining an index of abundance, a series of survey protocol can be tested using the `test_surveys` function.
-Provided a simulated population from `sim_distribution` and a series of survey protocol from `expand_surveys`, this function will simulate and analyze data from each survey using the `sim_survey`, `run_strat` and `strat_error` functions.
+Provided a simulated population from `sim_distribution` and a series of survey protocols from `expand_surveys`, this function will simulate and analyze data from each survey using the `sim_survey`, `run_strat` and `strat_error` functions.
 Like `sim_survey_parallel`, this function operates in parallel and allows the specification of `n_sims` and `n_loops`, and the product of these two arguments equals the number of times each survey is simulated.
-Keep in mind that low numbers of `n_sims` and high numbers of `n_loops` will be easier on RAM, but may be slower, especially if the work is spread across few `cores`.
-Because most of the default settings of the functions match the case study settings, the below code will replicate the results from our case study (see Table \@ref(tab:par) for a more detail).
+Keep in mind that low numbers of `n_sims` and high numbers of `n_loops` will be less demanding on RAM, but may be slower, especially if the work is spread across few `cores`.
+Because most of the default settings of the functions match the case study settings, the code below will replicate the results from our case study (see Appendix 1 for more detail).
 The `expand_surveys` function sets up a series of `r nrow(expand_surveys())` surveys to test (i.e.
@@
 
-Processing time will system (i.e.
+Processing time will be system (i.e.
 amount of RAM and number of cores) and setting (i.e.
@@
 The `test_survey` function will print a progress bar, generated using the **`progress`** package [@csardi2016], which details percent completion and will also include an eta after the first step of the loop completes.
-Keep in mind that thousands of simulations of hundreds of survey designs may take days to run.
 The `test_surveys` function therefore includes an option for exporting intermediate results to a local directory, via the `export_dir` argument, and the `resume_test` function can be used resume a `test_surveys` run that had to be stopped part way through the process.
@@
 These plots help visually assess the level of precision and bias from a specific set of survey protocol.
-The three lines of code below will produce interactive fan charts for stratified estimates of total abundance, abundance at length and abundance at age, respectively.
-A subset of these plots have been combined into static plots (Figure \@ref(fig:fan1), \@ref(fig:fan2), \@ref(fig:fan3)).
+The three lines of code below will produce interactive fan charts for stratified estimates of total abundance, abundance at length and abundance at age, respectively (e.g.
+Figure \@ref(fig:fan1), \@ref(fig:fan2), \@ref(fig:fan3)).
 
@@
 ```{r fan1, fig.cap = "Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
-The heavy dark line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+The thick black line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
 This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
@@
 ```{r fan2, fig.cap = "Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
-The heavy dark line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+The thick black line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
 This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
@@
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
-The heavy dark line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+The thick black line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
 This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
@@
 The `plot_survey_rank` function produces a divergent dot plot of the results which ranks the surveys by RMSE.
-Using the `wich_strat` argument, the plot can be focused on total, length or age based stratified results.
+Using the `which_strat` argument, the plot can be focused on total, length or age based stratified results.
 The `plot_error_surface` focuses solely on the age based stratified results by plotting a surface of RMSE (z-axis) by set (drop down selection), length (y-axis) and age (z-axis) sampling effort.
@@
 Like any model, this simulation is a simplification of a much more complex reality.
-For instance, the population is assumed to aggregate by age-class and be uniformly distributed within a cell when, in the real-world, they may aggregate by length and form finer-scale clusters.
+For instance, the population is assumed to aggregate by age-class and be uniformly distributed within a cell, instead fish may aggregate by length and form finer-scale clusters.
 The survey is also an instantaneous snapshot of the population, meaning that the population is assumed to be in the same location from the beginning to the end of the survey.
@@
 Finally, area trawled is assumed to be perfectly standard.
-This is a short, non-inclusive, list of additional variability that likely occurs in the real-world but was not implemented here for simplicity.
+These assumptions, plus a range of others, will surely under-represent the natural variability of fish populations.
 Nevertheless, the **`SimSurvey`** package provides the most realistic operating model that we are aware of for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
@@
 One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012].
-Results from the case study presented here (Appendix 1) reiterate this message and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
+Results from the case study presented here (Appendix 1) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
 @Berg2012].
@@
 
-Simulation testing is an important tool in the field of fisheries science as the inferred status of fish stocks hing on the data and models used to assess fish populations.
+Simulation testing is an important tool in the field of fisheries science as the inferred status of fish stocks hinge on the data and models used to assess fish populations.
 Simulations provide an opportunity to explore survey and model performance, and such explorations are becoming increasingly important as model complexity increases.
@@
 
-Results from the case study presented provide one example of how this package can be used to explore the efficacy of various survey designs.
 A dynamic population with a patchy and age-clustered distribution was simulated for this case study (Figure \@ref(fig:surface)a, \@ref(fig:distribution)a).
@@
 The samples within each successful set are also dynamic and rarely include all age groups.
-This is because only moderate correlation ($\varphi_{\xi, age}$ of 0.5) was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations.
+This is because only moderate correlation ($\varphi_{\xi, age}$ of 0.5) was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations, as is typically observed in the actual survey.
 These settings were chosen to simulated data that roughly correspond to actual survey data of cod from NAFO division 3Ps, where cod between the ages of 1-5 tend to be caught at different locations.
@@
 A series of surveys were run over the simulated population (see `expand_surveys` call in Table \@ref(tab:par)) and stratified analyses were run on the data obtained.
-The stratified analyses provided estimates of total abundance, abundance at length and abundance at age, and different stages of sampling affect these estimates.
-Specifically, set density affects all estimates, length sampling affects estimates of abundance at length and abundance at age, and age sampling affects abundance at age estimates.
-This is because stratified estimates of total abundance is based on aggregate catch while abundance at length estimates requires both total and length frequency data; finally, abundance at age estimates requires set totals, and length and age data to construct an age-length key.
+The stratified analyses provided estimates of total abundance, abundance at length and abundance at age.
+Different stages of sampling affect these estimates, specifically, set density affects all estimates, length sampling affects estimates of abundance at length and abundance at age, and age sampling affects abundance at age estimates.
 
@@
 
-The simple case study presented here revealed some expected and unexpected patterns.
-First, there were clear improvements to precision in all population estimates as the number of sets were increased.
-This result was expected as it aligns with a basic principle of survey design that the precision and representativeness of a survey increases as the number of sampling units are increased [@sutherland2006].
+The simple case study presented here revealed some expected, but also a few unexpected, patterns.
+First, and not surprisingly, there were clear improvements to precision in all population estimates as the number of sets were increased [@sutherland2006].
 Second, stratified estimates of abundance at age were often biased and, in some cases, estimates were poorer when sub-sampling effort was increased.
-This result was unexpected because design-based estimators, such as the stratified analysis applied here, are expected to be unbiased [@smith1990] and increases to sub-sampling effort was expected to be relatively ineffective, not detrimental.
+This result was unexpected because design-based estimators, such as the stratified analysis applied here, are expected to be unbiased [@smith1990] and large increases to sub-sampling effort could be expected to be relatively ineffective, but not detrimental.
 Results from stratified estimates of abundance at length, in contrast, better align with these expectations.
@@
 Using an age-length key in conjunction with the length distribution to estimate abundance at age is standard procedure in the analysis of fisheries-independent survey data as only a small fraction of the catch is typically aged.
-This is because the aging procedure is costly and time-consuming whereas length measurements are relatively easy to obtain.
+The aging procedure is costly and time-consuming whereas length measurements are relatively easy to obtain.
 Ages are generally determined from length-stratified sub-samples of the catch and raw proportions of age-at-length are used to assign ages to fish in specific length groups.
@@
 The set density of the multi-species survey conducted by Fisheries and Oceans Canada in the Newfoundland region, for instance, ranges between 0.001 and 0.002 sets / km^2^.
-Length sampling effort, in contrast, tends to be on the higher end of the scenarios presented here because lengths are relatively cheap and easy to obtain.
+Length sampling effort, in contrast, tends to be on the higher end of the scenarios presented here because lengths are relatively easy to obtain.
 Again, using Fisheries and Oceans Canada sampling protocol from the Newfoundland region as an example, it is not uncommon for length sampling effort to be capped at 500 measurements per set.
@@
 This is largely because fish caught together tend to be more similar than those in the general population [@Pennington1994].
-Therefore, if the goal of a trawl survey is to maximize information, it is likely better to stop collecting psudoreplicates and, instead, focus efforts on other species or the next set.
+Therefore, if the goal of a trawl survey is to maximize information, it is likely better to stop collecting correlated samples and, instead, focus efforts on other species or the next set.
 

### 2019-07-04 R/plot_sim.R

@@ -84,7 +84,7 @@ plot_grid <- function(grid, ...) {
     add_paths(data = df_strat, x = ~long, y = ~lat, color = I("white"),
               hoverinfo = "none", size = I(0.5), showlegend = FALSE) %>%
     add_paths(data = df_div, x = ~long, y = ~lat, color = I("darkgrey"),
-              hoverinfo = "none", size = I(2), showlegend = FALSE) %>%
+              hoverinfo = "none", size = I(4), showlegend = FALSE) %>%
     layout(xaxis = xax, yaxis = yax)
 
 }

### 2019-07-06 analysis/paper/appendix_1_figure.R
Draft a figure for the appendix showing the similarity between real and simulated data

@@ -1,0 +1,253 @@
+
+library(raster)
+library(SimSurvey)
+library(Rstrap)
+library(plotly)
+library(dplyr)
+
+source("analysis/paper/helper_functions.R")
+
+## REAL DATA ----------
+
+## Load survey data compiled for Rstrap
+load("analysis/rv_data/converted_set_details_2018-02-26.Rdata")
+load("analysis/rv_data/converted_length-frequency_data_2018-02-26.Rdata")
+load("analysis/rv_data/age-growth_data_2018-02-26.Rdata")
+
+## Subset data to cod
+con.setdet <- con.setdet[(con.setdet$rec == 5 | (con.setdet$rec == 6 & con.setdet$spec == 438)), ]
+con.lf <- con.lf[con.lf$spec == 438, ]
+ag <- ag[ag$spec == 438, ]
+rv_data <- list(setdet = con.setdet, lf = con.lf, ag = ag)
+
+## Save Rstrap output
+index.strata <- c(293:300, 306:326, 705:708, 711:716, 779:783)
+out <- strat.fun(setdet = rv_data$setdet, lf = rv_data$lf, ag = rv_data$ag,
+                 data.series = "Campelen", program = "strat2 & strat1",
+                 species = 438, survey.year = 1995:2017, season = "spring",
+                 NAFOdiv = "3P", strat = c(293:300, 306:326, 705:708, 711:716, 779:783),
+                 sex = c("male","female","unsexed"), length.group = 3,
+                 group.by = "length & age",
+                 export = NULL, plot.results = FALSE)
+
+## Convert lat and lon to UTM
+setdet <- data.table(out$raw.data$set.details)
+xy <- cbind(-setdet$long.start, setdet$lat.start) %>%
+  rgdal::project(., proj = sp::proj4string(survey_grid)) %>%
+  data.frame(.)
+names(xy) <- c("easting", "northing")
+setdet <- cbind(setdet, xy)
+
+## Make a unique identifier
+setdet$id <- with(setdet, paste(survey.year, vessel, trip, set))
+ag <- out$raw.data$age.growth
+ag$id <- with(ag, paste(survey.year, vessel, trip, set))
+
+## Melt age frequency data
+af <- data.table::melt(setdet,
+                       id.vars = c("survey.year", "vessel", "trip", "set", "easting", "northing"),
+                       measure.vars = names(setdet)[grepl("^af", names(setdet))],
[…]

### 2019-07-08 analysis/paper/sim_survey_paper.Rmd
Draft a figure for the appendix showing the similarity between real and simulated data
Add new figure to the appendix

@@
 Overall, the function provides a simple tool for simulating a range dynamic age-structured populations.
-For instance, below we simulate a relatively long and short lived species (with default variance settings).
+For instance, below we simulate a relatively long and short lived species (with default variance and growth settings).
 
@@
 ```{r, eval = replot}
-set.seed(438)
-a <- sim_abundance(ages = 1:20,
-                   R = sim_R(mean = 3e+07),
-                   Z = sim_Z(mean = 0.5))
-b <- sim_abundance(ages = 1:6,
-                   R = sim_R(mean = 1e+10),
-                   Z = sim_Z(mean = 0.8))
 
@@
 Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid.
-A more detailed description of the space-age-year covariance is included in Appendix 2.
+A more detailed description of the space-age-year covariance is included in Appendix B.
 
@@
 Keep in mind that low numbers of `n_sims` and high numbers of `n_loops` will be less demanding on RAM, but may be slower, especially if the work is spread across few `cores`.
-Because most of the default settings of the functions match the case study settings, the code below will replicate the results from our case study (see Appendix 1 for more detail).
+Because most of the default settings of the functions match the case study settings, the code below will replicate the results from our case study (see Appendix A for more detail).
 The `expand_surveys` function sets up a series of `r nrow(expand_surveys())` surveys to test (i.e.
@@
 One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012].
-Results from the case study presented here (Appendix 1) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
+Results from the case study presented here (Appendix A) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
 @Berg2012].
@@
 
-# Appendix 1: Case study
+# Appendix A: Case study
 
@@
 Surveys, therefore, tend to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:survey)).
-The samples within each successful set are also dynamic and rarely include all age groups.
+The samples within each successful set are also dynamic and rarely include all length or age groups.
 This is because only moderate correlation ($\varphi_{\xi, age}$ of 0.5) was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations, as is typically observed in the actual survey.
 These settings were chosen to simulated data that roughly correspond to actual survey data of cod from NAFO division 3Ps, where cod between the ages of 1-5 tend to be caught at different locations.
-
-A series of surveys were run over the simulated population (see `expand_surveys` call in Table \@ref(tab:par)) and stratified analyses were run on the data obtained.
+In short, simulated data roughly emulate actual survey data (Figure \@ref(fig:real-sim-comp)).
+
+```{r real-sim-comp, fig.cap = "Comparison of a) real and b) simulated survey data where, from top to bottom, histograms of numbers caught across sets, distributions of set catches and length-frequency distributions from three random sets are shown from a randomly selected year from the real and simulated 3Ps cod survey data.
+The deeper shades in the length-frequency plots represent numbers of fish sampled for age determination."}
+
+knitr::include_graphics("figures/real_sim_comp.png")
+
+```
+
+A series of surveys were run over the simulated population and stratified analyses were run on the data obtained.
 The stratified analyses provided estimates of total abundance, abundance at length and abundance at age.
@@
 
-# Appendix 2: Age-year-space covariance
+# Appendix B: Age-year-space covariance
 

### 2019-07-08 analysis/paper/appendix_figures.R

@@ -274,5 +274,5 @@ p <- subplot(hist_row,
 
 p
 
-export_plot(p, file = "analysis/paper/figures/appendix_1_figure.png", width = 500, height = 650)
+export_plot(p, file = "analysis/paper/figures/real_sim_comp.png", width = 500, height = 650)
 

### 2019-07-08 analysis/paper/sim_survey_paper.Rmd
Address the remainder of Keith's suggestions

@@
 subtitle: "ALTERNATE TITLE - `SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
-author: "Paul M. Regular, Fran Mowbray, et al.?"
+author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, Jonathan Babyn, Fran Mowbray, etc.?"
 date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
@@
 ```
+
+<!-- Items to discuss -->
+<!-- - Title -->
+<!-- - Target journal -->
+<!-- - Actual figures in appendix (html article, with interactive figures, as supplement?) -->
 
@@
 
-Abundance is simulated using the `sim_abundance` function and a default function call is described in Table \@ref(tab:abundance).
+Abundance at age and length is simulated using the `sim_abundance` function and a default function call is described in Table \@ref(tab:abundance) along with associated symbols from the equations outlined above.
 This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`).
@@
 Overall, the function provides a simple tool for simulating a range dynamic age-structured populations.
-For instance, below we simulate a relatively long and short lived species (with default variance and growth settings).
+For instance, below we simulate a relatively long (object `a`) and short (object `b`) lived species, with default variance, starting abundance and growth settings.
 
@@
 
-The `make_grid` function is used to generate a survey grid for use in the `sim_distribution` function.
+The above equations are used in the `make_grid`, `sim_ays_covar` and `sim_parabola` functions, and these functions are used within `sim_distribution` to distribute a population simulated using `sim_abundance` throughout a grid (Table \@ref(tab:distribution)).
 The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`.
@@
 
-The `sim_distribution` function distributes the population simulated using `sim_abundance` throughout the grid generated using `make_grid`.
-Table \@ref(tab:distribution) outlines the default settings used for the function.
-In addition to supplying objects produced by `sim_abundance` and `make_grid`, this function requires two closures that describe the age-year-space covariance and the relationship with depth.
+In addition to supplying objects produced by `sim_abundance` and `make_grid`, the `sim_distribution` function requires two closures that describe the age-year-space covariance and the relationship with depth.
 Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by tweaking a few parameters in these closures.
@@
 wider range) and exhibits stronger correlation across ages than years.
-Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years`, respectively, arguments in the `sim_ays_covar` closure.
+Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years` arguments, respectively, in the `sim_ays_covar` closure.
 Variance in the size of the clusters can also be modified by changing the `sd` argument in the `sim_ays_covar` function.
@@
 Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
-The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the [Simulate abundance] section.
+The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the [Simulate Abundance] section.
 Depending on the number of fish caught, sub-sampling is then conducted.
@@
 
-Where $N_a$, $N_y$, and $N_k$ are the number of years, ages and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e.
+Where $N_a$, $N_y$, and $N_k$ are the number of ages, years and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e.
 catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$).
@@
 
-Assuming a stratified analysis as the default method for obtaining an index of abundance, a series of survey protocol can be tested using the `test_surveys` function.
+Assuming a stratified analysis as the default method for obtaining an index of abundance, a series of survey protocols can be tested using the `test_surveys` function.
 Provided a simulated population from `sim_distribution` and a series of survey protocols from `expand_surveys`, this function will simulate and analyze data from each survey using the `sim_survey`, `run_strat` and `strat_error` functions.
@@
 `n_loops` and `n_sims` ratio) dependent.
-The `test_survey` function will print a progress bar, generated using the **`progress`** package [@csardi2016], which details percent completion and will also include an eta after the first step of the loop completes.
+The `test_survey` function will print a progress bar, generated using the **`progress`** package [@csardi2016], which details percent completion and will also include an estimate time of arrival (eta) after the first step of the loop completes.
 The `test_surveys` function therefore includes an option for exporting intermediate results to a local directory, via the `export_dir` argument, and the `resume_test` function can be used resume a `test_surveys` run that had to be stopped part way through the process.
@@
 ```{r fan1, fig.cap = "Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
-The thick black line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
 This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
@@
 ```{r fan2, fig.cap = "Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
-The thick black line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
 This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
@@
   add_labels(x = rep(1, 3), y = c(0.05, 0.5, 0.95), xshift = 30, textangle = 90,
-             text = paste("M<sub>lengths</sub> =", rev(c(10, 100, 1000)))) %>% 
+             text = paste("M<sub>age</sub> =", rev(c(10, 100, 1000)))) %>% 
   add_labels(x = 0.5, y = 0, text = "Year", yshift = -40) %>% 
@@
 Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
-The thick black line indicates the true tred in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
 This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
@@
 The `plot_survey_rank` function produces a divergent dot plot of the results which ranks the surveys by RMSE.
-Using the `which_strat` argument, the plot can be focused on total, length or age based stratified results.
+Using the `which_strat` argument, the plot can be focused on total, length or age based stratified results (Figure \@ref(fig:rank)).
 The `plot_error_surface` focuses solely on the age based stratified results by plotting a surface of RMSE (z-axis) by set (drop down selection), length (y-axis) and age (z-axis) sampling effort.
-The sampling effort axes can be rule or sample size based (`plot_by = "rule"` or `plot_by = "samples"`, respectively).
-See figures \@ref(fig:rank) and \@ref(fig:rmse-surface) for examples of the output from the code below.
+The sampling effort axes can be rule or sample size based (`plot_by = "rule"` or `plot_by = "samples"`, respectively; Figure \@ref(fig:rmse-surface)).
 
@@
 
-```{r rank, fig.cap = "Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{sets}$] and length measurements [$N_{measured}$]), under various sampling protocol (set density [$D_{sets}$] and maximum number of lengths measured per set [$M_{lengths}$])."}
+```{r rank, fig.cap = "Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{sets}$] and length measurements [$N_{measured}$]), under various sampling protocol (set density [$D_{sets}$] and maximum number of lengths measured per set [$M_{lengths}$]).
+Records are ranked by lowest to highest RMSE score."}
 
@@
 
-There are a plethora of statistical models that may be tested using these simulated data, however, implementing a variety of analytical approaches was outside the scope of this package.
+A large number of statistical models that may be tested using these simulated data, but, implementing a variety of analytical approaches was outside the scope of this package.
 Instead we focus on analyzing simulated stratified-random survey data using a design-based stratified analysis.
@@
 
-**Feedback and advice:** Aaron Adamack, Jonathan Babyn, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Keith Lewis, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Greg Robertson, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
-
-**Package contributions:** Jonathan Babyn, Greg Robertson
-
-**Review of previous draft:** Dave Cote, Keith Lewis, Greg Robertson
+**Feedback and advice:** Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
+
+**Review of previous draft:** Dave Cote
 
@@
 Function settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data; these values were set as the defaults for the main functions in the package.
-Here we evaluate the efficacy of alternate sampling protocol by assessing deviation between stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
+Here we evaluate the efficacy of alternate sampling protocols by assessing deviation between stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
 Although assessing model-based analyses was an option, we focused on the design-based stratified analysis for simplicity and because of its widespread use.
@@
 Increasing the sampling rule above 100 measurements per set results in many more fish being measured, however, the increase in sampling effort is not matched with substantive declines in RMSE (Figure \@ref(fig:rank)).
-Interestingly, it appears that measuring fewer total fish at more locations appears to be more beneficial than many fish at fewer locations (Figure \@ref(fig:rank)).
+Further, it appears that measuring fewer total fish at more locations is more beneficial than measuring many fish at fewer locations (Figure \@ref(fig:rank)).
 
@@
 
-Using an age-length key in conjunction with the length distribution to estimate abundance at age is standard procedure in the analysis of fisheries-independent survey data as only a small fraction of the catch is typically aged.
+Using an age-length key in conjunction with the length distribution to estimate abundance at age is standard procedure in the analysis of fisheries-independent survey data as only a small fraction of the catch are typically aged.
 The aging procedure is costly and time-consuming whereas length measurements are relatively easy to obtain.

### 2019-07-30 analysis/paper/sim_survey_paper.Rmd
Address Joanne's suggestions

@@
 ---
-title: "`SimSurvey`: an `R` package for simulation testing the design and analysis of fisheries-independent trawl surveys"
-subtitle: "ALTERNATE TITLE - `SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
+title: "`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
 author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, Jonathan Babyn, Fran Mowbray, etc.?"
@@
 
-Fisheries-independent trawl surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
+Fisheries-independent surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
 While costly to obtain, this information forms the basis of many stock assessments throughout the world.
@@
 
-The package has two main components, the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated fish stock distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.
+The package has two main components, the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated population distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.
 
@@
 
-The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtural fish populations with correlation across space, time and age dimensions, 2) simulating surveys with a range of sampling stratigies and intensities, and 3) running stratified analyses of simulated survey data (Table \@ref(tab:fun)).
+The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtural fish populations with correlation across space, time and age, 2) simulating surveys with a range of sampling stratigies and intensities, and 3) running stratified analyses of simulated survey data (Table \@ref(tab:fun)).
 The equations behind these functions are detailed in the sections below.
@@
 The error around the recruitment process was set to follow a random walk, $\epsilon_{y} \sim N(\epsilon_{y - 1}, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$.
-The covariance across ages and years is controlled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, age}$ and $\varphi_{\delta, year}$, respectively).
+The covariance across ages and years is controlled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, \mathrm{age}}$ and $\varphi_{\delta, \mathrm{year}}$, respectively).
 This structure allows for autocorrelation in process errors across ages and years (i.e.
 total mortality can be made to be more similar for fish that are closer together in age and time).
-Abundance at age is then converted to abundance at length using the original von Bertalanffy growth curve [@Cailliet2006]:
+Abundance at age is then converted to abundance at length using the original von Bertalanffy growth curve [@von1938]:
 
@@
 Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$.
-Abundance in discrete length groups, $l_{group}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance at age to abundance at length.
+Abundance in discrete length groups, $l_{\mathrm{group}}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance at age to abundance at length.
 Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.
@@
 | **Function call**                           | **Description**                                            | **Symbol**                          |
-|:------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
+|:------------------------------------------- |:---------------------------------------------------------- | :---------------------------------- |
 | `sim_abundance(`                            | &nbsp;                                                     | &nbsp;                              |
@@
 | `+            log_sd = 0.2,`                | Standard deviation of total mortality (log)                | $\sigma_Z$                          |
-| `+            phi_age = 0.9,`               | Correlation across ages in error around total mortality    | $\varphi_{\delta, age}$    |
-| `+            phi_year = 0.5),`             | Correlation across years in error around total mortality   | $\varphi_{\delta, year}$   |
+| `+            phi_age = 0.9,`               | Correlation across ages in error around total mortality    | $\varphi_{\delta, \mathrm{age}}$    |
+| `+            phi_year = 0.5),`             | Correlation across years in error around total mortality   | $\varphi_{\delta, \mathrm{year}}$   |
 | `+  growth = sim_vonB(Linf = 120,`          | Mean asymptotic length (cm)                                | $l_{\infty}$                        |
@@
 | `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_l$                          |
-| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{group}^{N}$            |
+| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{\mathrm{group}}^{N}$            |
 
@@
 Abundance at age and length is simulated using the `sim_abundance` function and a default function call is described in Table \@ref(tab:abundance) along with associated symbols from the equations outlined above.
-This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`).
-The package includes the functions `sim_R`, `sim_Z` and `sim_vonB` for simulating recruitment, total mortality and growth, respectively.
-These functions are "closures" [i.e.
-functions that contain data and return functions; @wickham2014] and this structure was chosen to avoid the repeated specifications of ages and years.
-The use of closures also allows users to use their own closures with a similar structure but different underlying formula.
+This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions known as "closures", such as `sim_R`, `sim_Z` and `sim_vonB`, for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`), respectively.
+Closures are functions that contain data and return functions [@wickham2014] and this structure was chosen to avoid the repeated specifications of ages and years.
+The use of closures also allows users to construct and use their own closures with a similar structure but different underlying formula.
 Overall, the function provides a simple tool for simulating a range dynamic age-structured populations.
-For instance, below we simulate a relatively long (object `a`) and short (object `b`) lived species, with default variance, starting abundance and growth settings.
+For instance, below we provide examples where we simulate a relatively long and short lived species (note that default variance, starting abundance and growth settings were used in both simulations).
 
@@
 set.seed(438)
-a <- sim_abundance(ages = 1:20,
+long <- sim_abundance(ages = 1:20,
                       R = sim_R(mean = 3e+07),
-                   Z = sim_Z(mean = 0.5))
-b <- sim_abundance(ages = 1:6,
+                      Z = sim_Z(mean = 0.2))
+short <- sim_abundance(ages = 1:6,
                        R = sim_R(mean = 1e+10),
@@
 
-The package also includes several plotting function for making quick plotly-based [@sievert2018] interactive visuals of the simulated population.
+The package also includes several plotting functions for making quick plotly-based [@sievert2018] interactive visuals of the simulated population.
 For instance, the `plot_surface` function can be used to make quick visuals of matrices contained within the list returned by `sim_abundance`.
@@
 ```{r, echo = TRUE, eval = FALSE}
-plot_surface(a, mat = "N")
-plot_surface(b, mat = "N")
+plot_surface(long, mat = "N")
+plot_surface(short, mat = "N")
 ```
@@
 
-pa <- plot_surface(a, mat = "N", scene = "scene1", showscale = FALSE) %>% 
+pa <- plot_surface(long, mat = "N", scene = "scene1", showscale = FALSE) %>% 
   add_labels(x = 0.05, y = 0.9, text = "a)")
-pb <- plot_surface(b, mat = "N", scene = "scene2", showscale = FALSE) %>% 
+pb <- plot_surface(short, mat = "N", scene = "scene2", showscale = FALSE) %>% 
   add_labels(x = 0.05, y = 0.9, text = "b)")
@@
 The next step in the simulation is to distribute the abundance at age matrix simulated using the cohort model throughout a spatial field.
-Here, a grid of $s$ locations is generated with a resolution of $R$ and a depth gradient.
-The depth gradient is set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$.
-The grid is divided into $N_{div}$ divisions and $N_{strat}$ depth-based strata.
+Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
+The grid is divided into $N_{\mathrm{div}}$ divisions (e.g.
+NAFO or ICES divisions) and $N_{\mathrm{strat}}$ depth-based strata.
 The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space
@@
 Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and the two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
-Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively.
+Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
 The overall variance of the process is controlled by $\sigma_{\xi}^2$.
@@
 Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid.
-A more detailed description of the space-age-year covariance is included in Appendix B.
+A more detailed description of the space-age-year covariance is included in [*Appendix B*][Appendix B: Age-year-space covariance].
 
@@
 | **Function call**                                       | **Description**                                              | **Symbol**                              |
-|:------------------------------------------------------- |:----------------------------------------------------- | :-------------------- |
+|:------------------------------------------------------- |:------------------------------------------------------------ | :-------------------------------------- |
 | `sim_distribution(`                                     | &nbsp;                                                       | &nbsp;                                  |
@@
 | `+                   y_range = c(-140, 140),`           | Range of grid in the y dimension                             | &nbsp;                                  |
-| `+                   res = c(3.5, 3.5),`                | Grid resolution (km)                                  | $R$                   |
-| `+                   shelf_depth = 200,`                | Shelf depth (km)                                      | $d_{shelf}$           |
-| `+                   shelf_width = 100,`                | Shelf width (km)                                      | $w_{shelf}$           |
-| `+                   depth_range = c(0, 1000),`         | Depth range (m)                                       | $[d_{min}, d_{max}]$  |
-| `+                   n_div = 1,`                        | Number of divisions                                   | $N_{div}$             |
+| `+                   res = c(3.5, 3.5),`                | Grid resolution in x and y dimensions (km) - i.e. cell area  | $A_{\mathrm{cell}}$                     |
+| `+                   shelf_depth = 200,`                | Shelf depth (m)                                              | $d_{\mathrm{shelf}}$                    |
+| `+                   shelf_width = 100,`                | Shelf width (km)                                             | $w_{\mathrm{shelf}}$                    |
+| `+                   depth_range = c(0, 1000),`         | Depth range from coast to slope (m)                          | $[d_{\mathrm{min}}, d_{\mathrm{max}}]$  |
+| `+                   n_div = 1,`                        | Number of divisions                                          | $N_{\mathrm{div}}$                      |
 | `+                   strat_breaks = seq(0, 1000, 40)),` | Series of depth breaks for defining strata                   | &nbsp;                                  |
@@
 | `+                            lambda = 1,`              | Smoothness of spatial correlation                            | $\lambda$                               |
-| `+                            phi_age = 0.5,`           | Correlation across ages in spatial distribution       | $\varphi_{\xi, age}$  |
-| `+                            phi_year = 0.9,`          | Correlation across years in spatial distribution      | $\varphi_{\xi, year}$ |
+| `+                            phi_age = 0.5,`           | Correlation across ages in spatial distribution              | $\varphi_{\xi, \mathrm{age}}$           |
+| `+                            phi_year = 0.9,`          | Correlation across years in spatial distribution             | $\varphi_{\xi, \mathrm{year}}$          |
 | `+                            group_ages = 5:20),`      | Make space-age-year variance equal across these ages         | &nbsp;                                  |
@@
 If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure is used.
-The package includes a manually constructed survey grid of NAFO division 3Ps (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid.
+The package includes a manually constructed survey grid of NAFO Subdivision 3Ps (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid.
 However, for simplicity, we use `make_grid` to construct a square grid for a default run of `sim_distribution`.
-Below we generate and plot (Figure \@ref(fig:grid)) a default grid, another grid with the number of strata increased by increasing the number of `strat_splits`, and another with four divisions and a more linear gradient in depth.
+Below we generate and plot (Figure \@ref(fig:grid)) a default grid, another grid with the number of strata increased by increasing the number of `strat_splits`, and another with four divisions and a linear depth gradient (the sigmoid curve is forced to be linear when `shelf_width` is set to zero).
 
@@
 Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e.
-wider range) and exhibits stronger correlation across ages than years.
+wider `range`) and exhibits stronger correlation across ages than years (i.e.
+lower `phi_year` and higher `phi_age`).
 Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years` arguments, respectively, in the `sim_ays_covar` closure.
@@
 The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated.
-The area of each strata is calculated and this is used to define the number of sets to allocate to each strata under a particular set density, $D_{sets}$.
+The area of each strata is calculated and this is used to define the number of sets to allocate to each strata under a particular set density, $D_{\mathrm{sets}}$.
 The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:
 
-$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{trawl}}{A_{cell}} q_a )$$
-Where $n_{a,y,s}$ is the number of fish sampled by a set, $A_{trawl}$ indicates the area covered by the trawl, $A_{cell}$ is the area of a grid cell, and $q_a$ is the catchability coefficient of each age.
+$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{\mathrm{trawl}}}{A_{\mathrm{cell}}} q_a )$$
+
+Where $n_{a,y,s}$ is the number of fish of age `a` in year `y` sampled by a set at location `s`, $A_{\mathrm{trawl}}$ indicates the area covered by the trawl, $A_{\mathrm{cell}}$ is the area of a grid cell, and $q_a$ is the catchability coefficient of each age.
 Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
-The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the [Simulate Abundance] section.
+The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the *[Simulate abundance]* section.
 Depending on the number of fish caught, sub-sampling is then conducted.
-Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division.
+Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division.
 Such sub-sampling is common in fisheries-independent surveys as it is costly to sample every fish captured.
-Age determination is especially time-consuming, which is why age sampling tends to be length-stratified.
+Age determination is especially time-consuming, which is why otoliths for age-determination tend to be sub-sampled by length-bin to obtain a representative age sample across a wider range of lengths than would be obtained via random sampling.
 
@@
 | **Function call**                   | **Description**                                                         | **Symbol**              |
-|:----------------------------------- |:----------------------------------------------------------------------- | :------------- |
+|:----------------------------------- |:----------------------------------------------------------------------- | :---------------------- |
 | `sim_survey(`                       | &nbsp;                                                                  | &nbsp;                  |
@@
 | `+                   x0 = 3),`      | Midpoint of logistic curve of catchability                              | $x_0$                   |
-| `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (length, width; km) - i.e. area covered by a trawl     | $A_{trawl}$    |
+| `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (length, width; km) - i.e. area covered by a trawl     | $A_{\mathrm{trawl}}$    |
 | `+  min_sets = 2`                   | Minimum number of sets to conduct per strata                            | &nbsp;                  |
-| `+  set_den = 2/1000,`              | Set density (km^-2^)                                                    | $D_{sets}$     |
-| `+  lengths_cap = 500,`             | Maximum number of lengths to collect / set                              | $M_{lengths}$  |
-| `+  length_group = 1,`              | Length group bin size for age sampling (cm)                             | $l_{group}$    |
-| `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / division              | $M_{ages}$     |
+| `+  set_den = 2/1000,`              | Set density (km^-2^)                                                    | $D_{\mathrm{sets}}$     |
+| `+  lengths_cap = 500,`             | Maximum number of lengths to collect / set                              | $M_{\mathrm{lengths}}$  |
+| `+  length_group = 1,`              | Length group bin size for age sampling (cm)                             | $l_{\mathrm{group}}$    |
+| `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / division              | $M_{\mathrm{ages}}$     |
 | `+  age_sammpling = "stratified")`  | Controls whether age sampling is length `"stratified"` or `"random"`    | &nbsp;                  |
@@
 Multiple simulations of the same survey can be run using the `n_sims` argument, however, requesting large numbers of simulations can be computational demanding depending on the processing capacity available.
-Below we use `sim_survey` to simulate two surveys over a default population, of which one is set-up to have higher sampling effort than the other.
+Below we use `sim_survey` to simulate two surveys over a default population, of which one is set-up to have higher set density (`set_den`) than the other.
 
@@
 
-```{r survey, fig.cap = "Bubble plots and histograms of set catches from a simulated stratified-random survey of a default population under relatively a) high and b) low sampling effort.
-Histograms of length and age composition include the distribution caught and sampled overlaid with a line of the true distribution of length and ages available to the survey.
+```{r survey, fig.cap = "Bubble plots and histograms of set catches from a simulated stratified-random survey of a default population under relatively a) low and b) high sampling effort.
+Histograms of length and age composition include the distribution of all fish caught next to those sampled overlaid with a line of the true distribution of lengths and ages available to the survey.
 Note that the first simulation of the survey in year 20 is depicted here.
@@
 As noted above, available RAM may limit the utility of the `sim_survey` function for running thousands of simulations of the same survey.
-The `sim_survey_parallel` function was therefore constructed to facilitate the running of thousands of simulations without running into RAM limitations.
+The `sim_survey_parallel` function was therefore constructed to facilitate this process.
 This function is set-up to run multiple `sim_survey` calls in parallel using the **`doParallel`** package [@weston2015] and, as such, multiple loops can be run using the `n_loops` argument and within each loop multiple simulations can be run (controlled using the `n_sims` argument).
@@
 While there are many model-based options for obtaining an index [e.g.
-@Thorson2015], survey-based approaches, such as stratified analyses, are often used.
-Here we apply formula presented in Smith and Somerton [-@Smith1981] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
+@Thorson2015], design-based approaches, such as stratified analyses, are often used.
+Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in [*Appendix C*][Appendix C: Stratified analysis equations]] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
 Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be scaled up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key.
@@
 Keep in mind that low numbers of `n_sims` and high numbers of `n_loops` will be less demanding on RAM, but may be slower, especially if the work is spread across few `cores`.
-Because most of the default settings of the functions match the case study settings, the code below will replicate the results from our case study (see Appendix A for more detail).
+Because most of the default settings of the functions match the case study settings, the code below will replicate the results from our case study (see [*Appendix A*][Appendix A: Case study] for more detail).
 The `expand_surveys` function sets up a series of `r nrow(expand_surveys())` surveys to test (i.e.
-all possible combinations of `set_den`, `lengths_cap` and `ages_cap` protocol) and the `test_surveys` function will run 1000 simulations of each survey and compare stratified estimates of abundance to the true abundance available to the survey.
+all possible combinations of the `set_den`, `lengths_cap` and `ages_cap` vectors) and the `test_surveys` function will run 1000 simulations of each survey and compare stratified estimates of abundance to the true abundance available to the survey.
 
@@
 tests <- test_surveys(pop, surveys = surveys,
-                      n_sims = 5, n_loops = 200, cores = 2)
+                      n_sims = 5, n_loops = 200, cores = 3)
 
@@
 The `test_survey` function will print a progress bar, generated using the **`progress`** package [@csardi2016], which details percent completion and will also include an estimate time of arrival (eta) after the first step of the loop completes.
-The `test_surveys` function therefore includes an option for exporting intermediate results to a local directory, via the `export_dir` argument, and the `resume_test` function can be used resume a `test_surveys` run that had to be stopped part way through the process.
+The `test_surveys` function therefore includes an option for exporting intermediate results to a local directory, via the `export_dir` argument, and the `resume_test` function can be used to resume a `test_surveys` run that had to be stopped part way through the process.
 The final object produced will be a list that includes all objects from `sim_abundance` and `sim_distribution` with the table of survey designs tested (named `surveys`) and tables produced by `strat_error` that end with the names `_strat_error` and `_strat_error_stats`.
@@
 
-```{r fan1, fig.cap = "Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$.
+```{r fan1, fig.cap = "Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{\\mathrm{sets}}$.
 The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
@@
 
-```{r fan2, fig.cap = "Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
+```{r fan2, fig.cap = "Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{\\mathrm{sets}}$, and length sampling protocol, $M_{\\mathrm{lengths}}$.
 The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
@@
 
-```{r fan3, fig.cap = "Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$.
-Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios.
+```{r fan3, fig.cap = "Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{\\mathrm{sets}}$, and length sampling protocol, $M_{\\mathrm{lengths}}$.
+Number of ages sampled per length group, $M_{\\mathrm{ages}}$, was 10 in all scenarios.
 The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
@@
 
-```{r rank, fig.cap = "Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{sets}$] and length measurements [$N_{measured}$]), under various sampling protocol (set density [$D_{sets}$] and maximum number of lengths measured per set [$M_{lengths}$]).
+```{r rank, fig.cap = "Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{\\mathrm{sets}}$] and length measurements [$N_{\\mathrm{measured}}$]), under various sampling protocol (set density [$D_{\\mathrm{sets}}$] and maximum number of lengths measured per set [$M_{\\mathrm{lengths}}$]).
 Records are ranked by lowest to highest RMSE score."}
@@
 ```{r rmse-surface, fig.cap = "Surface plots of RMSE from an array of surveys with different sampling protocol.
-Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$).
+Panels represent surveys with different set densities ($D_{\\mathrm{sets}}$), x-axes represent the maximum sampling effort of lengths per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per length group ($M_{\\mathrm{ages}}$).
 This plot is a facet of plots produced by `plot_error_surface` when supplied results from `test_surveys`.
@@
 One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012].
-Results from the case study presented here (Appendix A) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
+Results from the case study presented here ([*Appendix A*][Appendix A: Case study]) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
 @Berg2012].
@@
 We also realize that a single fisheries-independent survey may have multiple goals as data obtained are often used to assess multiple species or to conduct community analyses.
-We therefore endeavor to add functions for simulating multi-species surveys.
+We will therefore endeavor to add functions for simulating multi-species surveys.
 Finally, it would be useful to add an option for testing the consequences of surveys with partial coverage of a population as survey coverage is a frequent concern in stock assessment.
@@
 
-**Review of previous draft:** Dave Cote
+**Review of previous draft:** Dave Cote, Joanne Morgan
 
@@
 
-The construction of the **`SimSurvey`** package was motivated by a regional need to assess the sampling design of trawl surveys conducted along the Newfoundland and Labrador shelf by Fisheries and Oceans Canada.
-As a tangible first step, we focused our efforts on emulating data from the survey of cod in NAFO division 3Ps since it is one of the most dynamic and heavily sampled stocks in the region.
+The construction of the **`SimSurvey`** package was motivated by a need to assess the sampling design of trawl surveys conducted along the Newfoundland and Labrador shelf by Fisheries and Oceans Canada.
+As a tangible first step, we focused our efforts on emulating data from the survey of cod in NAFO Subdivision 3Ps since it is one of the most dynamic and heavily sampled stocks in the region.
 Function settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data; these values were set as the defaults for the main functions in the package.
-Here we evaluate the efficacy of alternate sampling protocols by assessing deviation between stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
+Here we evaluate the efficacy of alternate sampling protocols by assessing deviation of stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
 Although assessing model-based analyses was an option, we focused on the design-based stratified analysis for simplicity and because of its widespread use.
-The code below will replicate the results of this caset study:
+The code below displays the core defaults will replicate the results of this case study.
+Results produced here are the same as those produced and displayed in a series of figures in the *[Testing survey protocol]* section of the paper.
 
@@
                            ages_cap = c(2, 5, 10, 20, 50)),
-  n_loops = 1000,
+  n_sims = 5, n_loops = 200, cores = 3,
   q = sim_logistic(k = 2, x0 = 3)
@@
 
-*Caution:* as written, the `test_surveys` function call may take days or weeks to run.
-Run duration can be decreased by running multiple simulations per loop (`n_sims` argument) and by running the loops in parallel (`cores` argument).
-See help file for more details (`?test_surveys`).
-
 ## Results
 
-A dynamic population with a patchy and age-clustered distribution was simulated for this case study (Figure \@ref(fig:surface)a, \@ref(fig:distribution)a).
-In general, the parameter settings used (function defaults) dictate that all ages tend to aggregate in fairly dense clusters and that low-density zones are relatively wide-ranging.
+The default settings of `sim_abundance` and `sim_distribution` produces a dynamic population with a patchy and age-clustered distribution (Figure \@ref(fig:surface)a, \@ref(fig:distribution)a).
+In general, the function defaults dictate that all ages tend to aggregate in fairly dense clusters and that low-density zones are relatively wide-spread.
 Surveys, therefore, tend to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:survey)).
-The samples within each successful set are also dynamic and rarely include all length or age groups.
-This is because only moderate correlation ($\varphi_{\xi, age}$ of 0.5) was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations, as is typically observed in the actual survey.
-These settings were chosen to simulated data that roughly correspond to actual survey data of cod from NAFO division 3Ps, where cod between the ages of 1-5 tend to be caught at different locations.
-In short, simulated data roughly emulate actual survey data (Figure \@ref(fig:real-sim-comp)).
+The samples within each successful set are also variable and rarely include all length or age groups.
+This is because only moderate correlation ($\varphi_{\xi, \mathrm{age}}$ of 0.5) was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations, as is typically observed in the actual survey.
+These settings were chosen to simulated data that roughly correspond to actual survey data of cod from NAFO Subdivision 3Ps, where each age group between 1-4 tend to be caught at different locations while cod older than 4 years tend to be found at similar locations.
+In short, simulated data roughly emulate actual survey data with clustered catches and length samples with clear intra-haul correlation (Figure \@ref(fig:real-sim-comp)).
 
 ```{r real-sim-comp, fig.cap = "Comparison of a) real and b) simulated survey data where, from top to bottom, histograms of numbers caught across sets, distributions of set catches and length-frequency distributions from three random sets are shown from a randomly selected year from the real and simulated 3Ps cod survey data.
+The colors of the length-frequency plots correspond to the colored locations.
 The deeper shades in the length-frequency plots represent numbers of fish sampled for age determination."}
@@
 
-A series of surveys were run over the simulated population and stratified analyses were run on the data obtained.
+Using the `test_surveys` function, a series of surveys were run over the simulated population and stratified analyses were run on the data obtained.
 The stratified analyses provided estimates of total abundance, abundance at length and abundance at age.
@@
 Length sampling effort, in contrast, has an uneven impact depending on the set density scenario.
-At low set densities ($D_{sets}$ = 0.0005 sets / km^2^), RMSE declines when length sampling effort is increased from around 5 to around 100 measurements per set, and RMSE starts to increase as length sampling effort is increased beyond ~100 measurements per set; in fact, RMSE values appear higher at the highest length sampling scenario (1000 measurements / set) than the lowest (5 measurements / set; Figure \@ref(fig:rmse-surface)).
-A similar pattern is apparent under the medium set density scenario ($D_{sets}$ = 0.002 sets / km^2^), however, RMSE values under the lowest and highest length sampling scenarios are of similar magnitude.
-Finally, RMSE continues to decline with increased length sampling effort under the high set density scenarios ($D_{sets}$ = 0.01 sets / km^2^; Figure \@ref(fig:rmse-surface)).
+At low set densities ($D_{\mathrm{sets}}$ = 0.0005 sets / km^2^), RMSE declines when length sampling effort is increased from around 5 to around 100 measurements per set, and RMSE starts to increase as length sampling effort is increased beyond ~100 measurements per set; in fact, RMSE values appear higher at the highest length sampling scenario (1000 measurements / set) than the lowest (5 measurements / set; Figure \@ref(fig:rmse-surface)).
+A similar pattern is apparent under the medium set density scenario ($D_{\mathrm{sets}}$ = 0.002 sets / km^2^), however, RMSE values under the lowest and highest length sampling scenarios are of similar magnitude.
+Finally, RMSE continues to decline with increased length sampling effort under the high set density scenarios ($D_{\mathrm{sets}}$ = 0.01 sets / km^2^; Figure \@ref(fig:rmse-surface)).
 
@@
 Ages are generally determined from length-stratified sub-samples of the catch and raw proportions of age-at-length are used to assign ages to fish in specific length groups.
-Age-length keys are usually constructed at larger spatial scales because there are rarely enough samples to stratify these data.
+Age-length keys are usually constructed at larger spatial scales because there are rarely enough samples to construct a key at a finer spatial scale.
 Here we construct one age-length key for the division, as is done for the analysis of 3Ps cod.
@@
 Therefore, if the goal of a trawl survey is to maximize information, it is likely better to stop collecting correlated samples and, instead, focus efforts on other species or the next set.
+This is simply one example of the questions that can be explored using **`SimSurvey`**.
 
@@
 
-$$\xi_{a,y,s} \sim \left\{\begin{matrix} MVN\left (0, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a = 1, y = 1 & \\  MVN\left (\varphi_{year} \xi_{1, y - 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{age}^2)} \boldsymbol{R}_s \right ) & a = 1, y > 1 & \\ MVN\left (\varphi_{age} \xi_{a - 1, 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{year}^2)} \boldsymbol{R}_s \right ) & a > 1, y = 1 & \\ MVN\left (\varphi_{year} \xi_{a, y - 1, s} + \varphi_{age} (\xi_{a - 1, y, s} - \varphi_{year} \xi_{a - 1, y - 1, s}), \sigma_\xi^2 \boldsymbol{R}_s \right ) & a > 1, y > 1 & \end{matrix}\right.$$
-
-Where $MVN$ indicates the multivariate normal distribution, $\sigma_\xi^2$ controls the variance of the process, $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$ control correlation in the age and year dimension and $\boldsymbol{R}_s$ is defined by Matérn correlation:
+$$\xi_{a,y,s} \sim \left\{\begin{matrix} MVN\left (0, \frac{\sigma_\xi^2}{(1 - \varphi_{\mathrm{age}}^2)(1 - \varphi_{\mathrm{year}}^2)} \boldsymbol{R}_s \right ) & a = 1, y = 1 & \\  MVN\left (\varphi_{\mathrm{year}} \xi_{1, y - 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{\mathrm{age}}^2)} \boldsymbol{R}_s \right ) & a = 1, y > 1 & \\ MVN\left (\varphi_{\mathrm{age}} \xi_{a - 1, 1, s}, \frac{\sigma_\xi^2}{(1 - \varphi_{\mathrm{year}}^2)} \boldsymbol{R}_s \right ) & a > 1, y = 1 & \\ MVN\left (\varphi_{\mathrm{year}} \xi_{a, y - 1, s} + \varphi_{\mathrm{age}} (\xi_{a - 1, y, s} - \varphi_{\mathrm{year}} \xi_{a - 1, y - 1, s}), \sigma_\xi^2 \boldsymbol{R}_s \right ) & a > 1, y > 1 & \end{matrix}\right.$$
+
+Where $MVN$ indicates the multivariate normal distribution, $\sigma_\xi^2$ controls the variance of the process, $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$ control correlation in the age and year dimension and $\boldsymbol{R}_s$ is defined by Matérn correlation:
 
@@
 <br>
+
+# Appendix C: Stratified analysis equations
+
+Standard notation for use in the analysis of stratified-random survey data [modified from @Smith1981]:
+
+| **Equation** | **Description** |
+|:---------- |:----- |
+| $L$ | Number of strata ($h = 1, 2, ....., L$) |
+| $A_h$ | Area of the $h^{th}$ stratum |
+| $A_{\mathrm{trawl}}$ | Area covered by a standard trawl |
+| $N_h = \frac{A_h}{A_{\mathrm{trawl}}}$ | Total number of sample units in the $h^{th}$ stratum |
+| $n_h$ | Total number of units sampled in the $h^{th}$ stratum ($i = 1, 2, ....., n_h$) |
+| $N = \sum^{L}_{h = 1} N_h$ | Total number of sample units in the survey |
+| $n = \sum^{L}_{h = 1} n_h$ | Total number of observations in the survey |
+| $W_h = \frac{N_h}{N}$ | Stratum weight |
+| $f_h = \frac{n_h}{N_h}$ | Sampling fraction in the $h^{th}$ stratum |
+| $\overline{I}_h = \sum^{n_h}_{i = 1} \frac{I_{h, i}}{n_h}$ | Sample mean in the $h^{th}$ stratum, where $I_{h, i}$ is the $i^{th}$ observation in the $h^{th}$ stratum |
+| $s^2_h = \sum^{n_h}_{i = 1} \frac{(I_{h, i} - \overline{I}_h)}{(n_h - 1)}$ | Sample variance in the $h^{th}$ stratum |
+| $\overline{I} = \sum^{L}_{h = 1} W_h \overline{I}_h$ | Estimate of the population mean per unit (i.e. stratified mean catch per tow) |
+| $s^2(\overline{I}) = \frac{1}{N^2} \sum^{L}_{h=1} N_h (N_h - n_h) \frac{s^2_h}{n_h}$ | Estimate of the variance of the stratified mean |
+| $\widehat{I} = N \overline{I}$ | Estimate of the population total over the survey area |
 

### 2019-07-31 analysis/paper/sim_survey_paper.Rmd
Address Keith's latest suggestions

@@
 **1\.** Age-structured fish stocks often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective population surveys.
-Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision and increasing the risk of bias) and over-sampling (through the extensive sampling of correlated metrics).
-
-**2\.** For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of fishing sets and sub-sampling the fish captured to measure biological characteristics (e.g.
+Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision and increasing the risk of bias) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics).
+
+**2\.** For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of fishing sets as well as the sub-sampling of fish captured to measure biological characteristics (e.g.
 lengths or ages).
-Multiple analysis approaches are also now available.
-
-**3\.** Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, complex fish population structure has limited the number of tools that allow for simulation testing of a range of sampling and analytical pathways for fisheries-independent survey data.
+Population estimates are also affected by the pathway taken to analyze such data.
+
+**3\.** Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for fisheries-independent survey data.
 
 **4\.** Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations.
-The package allows the user to simulate age-structured populations that vary in space and time and then explore the efficacy of a range of sampling protocols to reproduce the population parameters of the known population.
+The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of sampling protocols to reproduce the population parameters of the known population.
 **`SimSurvey`** also include a function for running a stratified-random analysis of the simulated data.
@@
 **5\.** **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring.
-Various design- or model-based approaches can then be applied to the results to test the efficacy of different analytical approaches.
+Various statistical approaches can then be applied to the results to test the efficacy of different analytical approaches.
 
@@
 
-Fisheries-independent surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998].
-While costly to obtain, this information forms the basis of many stock assessments throughout the world.
+Fisheries-independent surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies.
+While costly to obtain, this information forms the basis of many stock assessments throughout the world [@Pennington1998].
 Efficient and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection.
@@
 Here we document **`SimSurvey`**, an `R` package designed to simplify and facilitate realistic simulations of fisheries-independent trawl surveys.
-In short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time
-
-The package has two main components, the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated population distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.
+In short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time.
+The package has two main components: the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated population distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.
 
@@
 The steps taken to simulate surveys of spatial, age-structure populations are outlined below.
+While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study.
+Output from default function calls are therefore relavant to the case study and these results are described and discussed in [*Appendix A*][Appendix A: Case study].
 
@@
 Functions in bold font are core functions and those in medium font are designed for use inside the core functions.
-The latter are typicaly closures, which are functions that contain data and return functions [@wickham2014]; here they are used to store parameter values and return functions that require dimensions, such as ages or years, to be supplied.
+The latter are typically closures, which are functions that contain data and return functions [@wickham2014]; here they are used to store parameter values and return functions that require dimensions, such as ages or years, to be supplied.
 
@@
 Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
-The overall variance of the process is controlled by $\sigma_{\xi}^2$.
+The overall variance of the process is controlled by $\sigma_{\xi}^2$ (see [*Appendix B*][Appendix B: Age-year-space covariance] for a more detailed description of the space-age-year covariance structure).
 In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering.
@@
 Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid.
-A more detailed description of the space-age-year covariance is included in [*Appendix B*][Appendix B: Age-year-space covariance].
 
@@
 The function also retains the grid object and converts these data into a data.table, named `grid_xy`, with headers `x`, `y`, `depth`, `cell`, `division` and `strat`.
-The `sp_N` object can be merge with the `grid_xy` data by `cell` to associate abundance with specific locations, depth, divisions or strata.
+The `sp_N` object can be merged with the `grid_xy` data by `cell` to associate abundance with specific locations, depth, divisions or strata.
 The `plot_distribution` function can be used to provide a quick visual of the distribution across ages and years.
@@
 The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated.
-The area of each strata is calculated and this is used to define the number of sets to allocate to each strata under a particular set density, $D_{\mathrm{sets}}$.
+The area of each strata is calculated and this is used to define the number of sets allocated to each strata under a particular set density, $D_{\mathrm{sets}}$.
 The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:
@@
 The function `sim_survey` can be used to simulate data from one survey over a population created using `sim_distribution`.
-A default function call is described in Table \@ref(tab:survey)).
+A default function call is described in Table \@ref(tab:survey).
 The `sim_survey` function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and definitions of the design of the survey.
 Specifically, the `q` argument requires a closure, such as `sim_logistic`, for defining the probability of catching specific age groups, trawl dimensions are defined in the `trawl_dim` argument, and set, length and age sampling effort are defined using the `set_den`, `lengths_cap` and `ages_cap` arguments, respectively.
-Multiple simulations of the same survey can be run using the `n_sims` argument, however, requesting large numbers of simulations can be computational demanding depending on the processing capacity available.
+Like `sim_abundance` and `sim_distribution`, custom closures can be supplied to `sim_survey` to impose alternate parametric curves for catchability at age.
+Multiple simulations of the same survey can be run using the `n_sims` argument, however, requesting large numbers of simulations can be computationally demanding depending on the processing capacity available.
 Below we use `sim_survey` to simulate two surveys over a default population, of which one is set-up to have higher set density (`set_den`) than the other.
@@
 
-```{r survey, fig.cap = "Bubble plots and histograms of set catches from a simulated stratified-random survey of a default population under relatively a) low and b) high sampling effort.
+```{r survey, fig.cap = "Bubble plots of abundance and histograms of set catches from a simulated stratified-random survey of a default population under relatively a) low and b) high sampling effort.
+Point size and color are scaled by abundance in the bubble plots.
 Histograms of length and age composition include the distribution of all fish caught next to those sampled overlaid with a line of the true distribution of lengths and ages available to the survey.
@@
 The `sim_survey_parallel` function was therefore constructed to facilitate this process.
-This function is set-up to run multiple `sim_survey` calls in parallel using the **`doParallel`** package [@weston2015] and, as such, multiple loops can be run using the `n_loops` argument and within each loop multiple simulations can be run (controlled using the `n_sims` argument).
+This function is set-up to run multiple `sim_survey` calls in parallel using the **`doParallel`** package [@weston2015] and, as such, multiple loops can be run using the `n_loops` argument and, within each loop, multiple simulations can be run (controlled using the `n_sims` argument).
 Total simulations will be the product of `n_loops` and `n_sims` arguments.
@@
 Low numbers of `n_sims` and high numbers of `n_loops` will be easier on RAM, but may be slower.
-The optimum ratio of `n_sims` to `n_loops` will depend on the amount of RAM and number of cores the user's computer has.
+The optimum ratio of `n_sims` to `n_loops` will depend on the amount of RAM and number of cores in a given computer.
 In any case, this function simplifies the process of running thousands of simulations of the same survey and the simulated data can then be supplied to survey-based or model-based analyses that require simulation testing.
@@
 
-Once the survey data are simulated, they then need to be analyzed to obtain an index of abundance.
+Survey data need to be analyzed to obtain an index of abundance.
 While there are many model-based options for obtaining an index [e.g.
@@
 
-Stratified estimates of abundance obtained are by supplying the output from `sim_survey` to the `run_strat` function.
+Stratified estimates of abundance are obtained by supplying the output from `sim_survey` to the `run_strat` function.
 RMSE of the stratified estimates can then be calculated using the `strat_error` function.
@@
 Using the `which_strat` argument, the plot can be focused on total, length or age based stratified results (Figure \@ref(fig:rank)).
-The `plot_error_surface` focuses solely on the age based stratified results by plotting a surface of RMSE (z-axis) by set (drop down selection), length (y-axis) and age (z-axis) sampling effort.
+The `plot_error_surface` displays the age based stratified results by plotting a surface of RMSE (z-axis) by set (drop down selection), length (y-axis) and age (z-axis) sampling effort.
 The sampling effort axes can be rule or sample size based (`plot_by = "rule"` or `plot_by = "samples"`, respectively; Figure \@ref(fig:rmse-surface)).
@@
 
-```{r rank, fig.cap = "Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{\\mathrm{sets}}$] and length measurements [$N_{\\mathrm{measured}}$]), under various sampling protocol (set density [$D_{\\mathrm{sets}}$] and maximum number of lengths measured per set [$M_{\\mathrm{lengths}}$]).
-Records are ranked by lowest to highest RMSE score."}
+```{r rank, fig.cap = "Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{\\mathrm{sets}}$] and length measurements [$N_{\\mathrm{measured}}$]), under various sampling protocols (set density [$D_{\\mathrm{sets}}$] and maximum number of lengths measured per set [$M_{\\mathrm{lengths}}$]).
+Records are ranked by lowest to highest RMSE score.
+Within each plot, a color ramp is applied from lowest to highest value as an additional visual aid.
+Note that the exponent format of the axes defaults to SI unit symbols (e.g.
+M for million)."}
 
@@
 This plot is a facet of plots produced by `plot_error_surface` when supplied results from `test_surveys`.
-Note that the z-axes scales are different across the facets."}
+Note that RMSE scales are different across $D_{\\mathrm{sets}}$ facets.
+Note that the exponent format of the axes defaults to SI unit symbols (e.g.
+M for million)."}
 
@@
 Finally, area trawled is assumed to be perfectly standard.
-These assumptions, plus a range of others, will surely under-represent the natural variability of fish populations.
+These assumptions, plus a range of others, will surely under-represent the natural variability of fish populations and survey protocol.
 Nevertheless, the **`SimSurvey`** package provides the most realistic operating model that we are aware of for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
 
+<!-- # Should we include a summary of the case study - previous abstrat included below -->
+
+<!-- A central question to fish stock assessments is: how many samples are needed to adequately represent the population being monitored?
+For assessments based on trawl surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling (i.e.
+the number of sets, lengths and ages sampled in a survey).
+We therefore need to assess the impact of haul design and length and age sampling on survey accuracy to determine optimal sampling effort.
+In this study we simulated repeated stratified-random samples from a spatially correlated field of simulated fish.
+These samples were then analyzed and estimates of abundance at age were compared against a known truth.
+By varying the sampling protocol, we demonstrate that increasing the number of sets conducted per stratum provides the greatest gains in precision when sampling a clustered population.
+Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent and improvements were typically marginal, and sometimes squandered, when sub-sampling effort was increased.
+These results indicate that high levels of length and age sampling effort can be scaled back without significant losses to precision and, in some cases, estimates of abundance at age may be improved by reducing sub-sampling effort.
+Time saved by reducing said sampling may afford more time to conduct additional sets and/or increase sampling efforts of other species.
+While this simulation is tailored to a specific case, the method may be adapted to different systems and survey procedures.
+-->
+
 # Research opportunities
 
-The case study used here provides one example of ways in which **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
+The case study described in [*Appendix A*][Appendix A: Case study] provides one example of ways in which **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
 There are multiple layers to the sampling design of such surveys and the end results are linked with the way such data are analyzed.
@@
 One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012].
-Results from the case study presented here ([*Appendix A*][Appendix A: Case study]) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
+Results from the case study (see [*Appendix A*][Appendix A: Case study]) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
 @Berg2012].
@@
 
-Up to now, the package has focused on the effects of sampling design on the precision and bias of population estimates obtained from fisheries-independent surveys, however, the costs associated with sampling has yet to be considered.
+Up to now, the package has focused on the effects of sampling design on the precision and bias of population estimates obtained from fisheries-independent surveys; however, the costs associated with sampling has yet to be considered.
 In future iterations of **`SimSurvey`**, we hope to add options for integrating data on the time and monetary costs associated with each level of sampling (sets, length measurements, age determination) to facilitate cost-benefit analyses.
@@
 Together, two functions (`sim_abundance` and `sim_distribution`) are capable of simulating a wide range of populations with different life histories, depth associations and spatial properties.
-The next necessary step to generating data similar to actual observations is to conduct a survey.
+The next necessary steps to generating data similar to actual observations is to conduct a survey.
 In this package we implement a function, `sim_survey`, that conducts a stratified random survey of the population.
@@
 
-```{r real-sim-comp, fig.cap = "Comparison of a) real and b) simulated survey data where, from top to bottom, histograms of numbers caught across sets, distributions of set catches and length-frequency distributions from three random sets are shown from a randomly selected year from the real and simulated 3Ps cod survey data.
+```{r real-sim-comp, fig.cap = "Comparison of a) real and b) simulated survey data where, from top to bottom, histograms of numbers caught across sets, distributions of set catches, and length-frequency distributions from three random sets are shown from a randomly selected year from the real and simulated 3Ps cod survey data.
 The colors of the length-frequency plots correspond to the colored locations.
@@
 Compared to increases in set density, the effects of increased length sampling effort on abundance at length estimates are less clear.
-Regarding estimates of abundance at length, RMSE appears to reach a plateau at when length sampling effort is increased 100 measurements per set (Figure \@ref(fig:rank)).
+Regarding estimates of abundance at length, RMSE appears to reach a plateau at when length sampling effort is increased to 100 measurements per set (Figure \@ref(fig:rank)).
 Increasing the sampling rule above 100 measurements per set results in many more fish being measured, however, the increase in sampling effort is not matched with substantive declines in RMSE (Figure \@ref(fig:rank)).
@@
 This is largely because fish caught together tend to be more similar than those in the general population [@Pennington1994].
-Therefore, if the goal of a trawl survey is to maximize information, it is likely better to stop collecting correlated samples and, instead, focus efforts on other species or the next set.
+Therefore, if the goal of a trawl survey is to maximize information, it is likely better to stop collecting correlated samples and, instead, focus efforts the next set and/or other species.
 This is simply one example of the questions that can be explored using **`SimSurvey`**.

### 2019-09-16 analysis/paper/sim_survey_paper.Rmd
Address Greg's latest edits to the paper

@@
 title: "`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
-author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, Jonathan Babyn, Fran Mowbray, etc.?"
+author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, Jonathan Babyn, Fran Mowbray"
 date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
@@
 
-replot <- FALSE # plot the figures again? (acts as a manual cache)
+replot <- TRUE # plot the figures again? (acts as a manual cache)
 
@@
 <!-- - Actual figures in appendix (html article, with interactive figures, as supplement?) -->
+
+# Things to manually fix in word
+
+1. Adjust font size of code strings
+2. Decrease font size of tables
+3. Indent closures in Table \@ref(tab:fun)
+4. Replace pluses (`+`) with spaces in Tables \@ref(tab:abundance) to (\#tab:survey)
+1. Remove this section
 
@@
 While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study.
-Output from default function calls are therefore relavant to the case study and these results are described and discussed in [*Appendix A*][Appendix A: Case study].
+Output from default function calls are therefore relevant to the case study and these results are described and discussed in [*Appendix A*][Appendix A: Case study].
 
@@
 
-The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtural fish populations with correlation across space, time and age, 2) simulating surveys with a range of sampling stratigies and intensities, and 3) running stratified analyses of simulated survey data (Table \@ref(tab:fun)).
+The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtual fish populations with correlation across space, time and age, 2) simulating surveys with a range of sampling strategies and intensities, and 3) running stratified analyses of simulated survey data (Table \@ref(tab:fun)).
 The equations behind these functions are detailed in the sections below.
@@
 
-Table: (\#tab:fun) Names and descritpions of the key functions of **`SimSurvey`**.
+Table: (\#tab:fun) Names and descriptions of the key functions of **`SimSurvey`**.
 Functions in bold font are core functions and those in medium font are designed for use inside the core functions.
@@
 This structure allows for autocorrelation in process errors across ages and years (i.e.
-total mortality can be made to be more similar for fish that are closer together in age and time).
+total mortality can be made to be more similar for fish that are closer together in age and/or time).
 Abundance at age is then converted to abundance at length using the original von Bertalanffy growth curve [@von1938]:
@@
 Abundance in discrete length groups, $l_{\mathrm{group}}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance at age to abundance at length.
-Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.
-
-Table: (\#tab:abundance) Default `sim_abundance` function call, with descriptions and associated parameter symbols of key arguments.
+Though some typical relationships have yet to be implemented (e.g.
+stock-recruitment), this formulation facilitates the simulation of age-structured populations that are dynamic enough to test an array of survey designs.
+
+Table: (\#tab:abundance) Default `sim_abundance` function call, with descriptions, default values and associated parameter symbols of key arguments.
 
@@
 This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions known as "closures", such as `sim_R`, `sim_Z` and `sim_vonB`, for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`), respectively.
-Closures are functions that contain data and return functions [@wickham2014] and this structure was chosen to avoid the repeated specifications of ages and years.
+Closures are functions that contain data and return functions [@wickham2014].
+For example, `sim_R(mean = 500, log_sd = 0.5)` returns a function that holds the supplied parameter values and requires a sequence of years to be supplied.
+(Note that each of the above-mentioned closures include a plot argument such that quick visuals can be obtained using a line of code like this: `sim_R(mean = 500, log_sd = 0.5, plot = TRUE)(years = 1:100)`).
+This structure was chosen to avoid the repeated specifications of ages and years.
 The use of closures also allows users to construct and use their own closures with a similar structure but different underlying formula.
-Overall, the function provides a simple tool for simulating a range dynamic age-structured populations.
+Overall, the function provides a simple tool for simulating a range of dynamic age-structured populations.
 For instance, below we provide examples where we simulate a relatively long and short lived species (note that default variance, starting abundance and growth settings were used in both simulations).
@@
 Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
-The grid is divided into $N_{\mathrm{div}}$ divisions (e.g.
-NAFO or ICES divisions) and $N_{\mathrm{strat}}$ depth-based strata.
+The grid is divided into $L_{\mathrm{div}}$ divisions (e.g.
+NAFO or ICES divisions) and $L_{\mathrm{strat}}$ depth-based strata.
 The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space
@@
 Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean depth.
-This helps simulate depth preferences of particular fish.
-Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and the two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
+Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and a two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
 Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
-The overall variance of the process is controlled by $\sigma_{\xi}^2$ (see [*Appendix B*][Appendix B: Age-year-space covariance] for a more detailed description of the space-age-year covariance structure).
+The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see [*Appendix B*][Appendix B: Age-year-space covariance] for a more detailed description of the space-age-year covariance structure).
 In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering.
@@
 | `+                   depth_range = c(0, 1000),`         | Depth range from coast to slope (m)                          | $[d_{\mathrm{min}}, d_{\mathrm{max}}]$  |
-| `+                   n_div = 1,`                        | Number of divisions                                          | $N_{\mathrm{div}}$                      |
+| `+                   n_div = 1,`                        | Number of divisions                                          | $L_{\mathrm{div}}$                      |
+| `+                   strat_splits = 2,                  | Number of times to horizontally split strat                  | &nbsp;                                  |
 | `+                   strat_breaks = seq(0, 1000, 40)),` | Series of depth breaks for defining strata                   | &nbsp;                                  |
@@
 | `+  q = sim_logistic(k = 2,`        | Steepness of logistic curve of catchability                             | $k$                     |
-| `+                   x0 = 3),`      | Midpoint of logistic curve of catchability                              | $x_0$                   |
+| `+                   x0 = 3),`      | Midpoint of logistic curve of catchability (age)                        | $x_0$                   |
 | `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (length, width; km) - i.e. area covered by a trawl     | $A_{\mathrm{trawl}}$    |
@@
              text = paste("M<sub>lengths</sub> =", rev(c(10, 100, 1000)))) %>% 
-  add_labels(x = 0.5, y = 0, text = "Year", yshift = -40) %>% 
+  add_labels(x = 0.5, y = 0, text = "Length", yshift = -40) %>% 
   add_labels(x = 0, y = 0.5, text = "Abundance index", xshift = -60, textangle = -90)
@@
   add_labels(x = rep(1, 3), y = c(0.05, 0.5, 0.95), xshift = 30, textangle = 90,
-             text = paste("M<sub>age</sub> =", rev(c(10, 100, 1000)))) %>% 
-  add_labels(x = 0.5, y = 0, text = "Year", yshift = -40) %>% 
+             text = paste("M<sub>length</sub> =", rev(c(10, 100, 1000)))) %>% 
+  add_labels(x = 0.5, y = 0, text = "Age", yshift = -40) %>% 
   add_labels(x = 0, y = 0.5, text = "Abundance index", xshift = -60, textangle = -90)
@@
 These assumptions, plus a range of others, will surely under-represent the natural variability of fish populations and survey protocol.
-Nevertheless, the **`SimSurvey`** package provides the most realistic operating model that we are aware of for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
-
-<!-- # Should we include a summary of the case study - previous abstrat included below -->
-
-<!-- A central question to fish stock assessments is: how many samples are needed to adequately represent the population being monitored?
-For assessments based on trawl surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling (i.e.
-the number of sets, lengths and ages sampled in a survey).
-We therefore need to assess the impact of haul design and length and age sampling on survey accuracy to determine optimal sampling effort.
-In this study we simulated repeated stratified-random samples from a spatially correlated field of simulated fish.
-These samples were then analyzed and estimates of abundance at age were compared against a known truth.
-By varying the sampling protocol, we demonstrate that increasing the number of sets conducted per stratum provides the greatest gains in precision when sampling a clustered population.
-Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent and improvements were typically marginal, and sometimes squandered, when sub-sampling effort was increased.
-These results indicate that high levels of length and age sampling effort can be scaled back without significant losses to precision and, in some cases, estimates of abundance at age may be improved by reducing sub-sampling effort.
-Time saved by reducing said sampling may afford more time to conduct additional sets and/or increase sampling efforts of other species.
-While this simulation is tailored to a specific case, the method may be adapted to different systems and survey procedures.
--->
+Nevertheless, the **`SimSurvey`** package provides a relatively complex and flexible operating model for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
 
@@
 
-The case study described in [*Appendix A*][Appendix A: Case study] provides one example of ways in which **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
+The case study described in [*Appendix A*][Appendix A: Case study] provides one example of how **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
 There are multiple layers to the sampling design of such surveys and the end results are linked with the way such data are analyzed.
@@
 The core of the simulation is based on the widely used cohort equation and, even though the processes that define recruitment and total mortality are simple, a wide range of stock dynamics can be simulated by changing a few parameters.
-This base population can then be distributed through a grid and relationships with depth can be defined as can the nature of the correlation across ages, years and space.
+This base population can then be distributed through a grid and relationships with depth and correlation across ages, years and space can be defined.
 Together, two functions (`sim_abundance` and `sim_distribution`) are capable of simulating a wide range of populations with different life histories, depth associations and spatial properties.
@@
 ```{r, echo = TRUE, eval = FALSE}
+
+# install.packages("remotes")
+# remotes::install_github("PaulRegular/SimSurvey")
+# library(SimSurvey)
 

### 2019-09-17 analysis/paper/sim_survey_paper.Rmd
Address Brian's suggestions and produce next draft of the paper

@@
 
-replot <- TRUE # plot the figures again? (acts as a manual cache)
+replot <- FALSE # plot the figures again? (acts as a manual cache)
 
@@
 4. Replace pluses (`+`) with spaces in Tables \@ref(tab:abundance) to (\#tab:survey)
-1. Remove this section
+5. Adjust size of footnote below Table \@ref(tab:abundance)
+6. Adjust size of Figure \@ref(fig:real-sim-comp)
+7. Remove this section
 
@@
 
-**1\.** Age-structured fish stocks often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective population surveys.
+**1\.** Fish populations often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective surveys.
 Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision and increasing the risk of bias) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics).
 
-**2\.** For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of fishing sets as well as the sub-sampling of fish captured to measure biological characteristics (e.g.
+**2\.** For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of fish captured to measure biological characteristics (e.g.
 lengths or ages).
@@
 The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of sampling protocols to reproduce the population parameters of the known population.
-**`SimSurvey`** also include a function for running a stratified-random analysis of the simulated data.
+**`SimSurvey`** also includes a function for estimating the stratified mean and variance of the simulated data.
 
@@
 Fisheries-independent surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies.
-While costly to obtain, this information forms the basis of many stock assessments throughout the world [@Pennington1998].
-Efficient and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection.
+While costly to obtain, this information forms the basis of many stock assessments throughout the world [@Pennington1998] and the quality this information depends on surveys and analyses that maximize information while minimizing the expense of data collection.
 While simulations provide a platform for exploring solutions to this optimization problem, building the necessary simulation framework is not a trivial task given the multi-stage nature of the typical sampling program and the complexity of the population processes we aim to represent.
-Notably, many fish stocks show size-specific patterns of aggregation, making it difficult to collect biological sub-samples (e.g.
-lengths and ages) that are representative of the entire target population.
+Notably, many fish stocks show size-specific patterns of aggregation, making it difficult to collect sufficient biological sub-samples (e.g.
+lengths and ages) to represent the entire target population.
 This undermines the assumptions of many analyses as sub-samples from such populations tend to exhibit strong positive intracluster correlation [@Aanes2015].
@@
 
-The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtual fish populations with correlation across space, time and age, 2) simulating surveys with a range of sampling strategies and intensities, and 3) running stratified analyses of simulated survey data (Table \@ref(tab:fun)).
+The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtual fish populations with correlation across space, time and age, 2) simulating surveys with a range of sampling strategies and intensities, and 3) estimating the stratified mean and variance of simulated survey data (Table \@ref(tab:fun)).
 The equations behind these functions are detailed in the sections below.
@@
 
-The simulation starts with a common cohort model where the abundance at age $a$ in year $y$ ($N_{a,y}$) equals the abundance of that cohort in the previous year multiplied by the survival rate, which is expressed in terms of total mortality ($Z_{a,y}$):
+The simulation starts with an exponential decay  cohort model where the abundance at age $a$ in year $y$ ($N_{a,y}$) equals the abundance of that cohort in the previous year multiplied by the associated survival rate, which is expressed in terms of total mortality ($Z_{a,y}$):
 
@@
 | `+  years = 1:20,`                          | Years                                                      | $y$                                 |
-| `+  R = sim_R(mean = 30000000,`             | Mean recruitment                                           | $\mu_r$                             |
+| `+  R = sim_R(mean = 30000000,`             | Mean recruitment^1^                                        | $\mu_r$                             |
 | `+            log_sd = 0.5),`               | Standard deviation of log-recruitment                      | $\sigma_r$                          |
-| `+  Z = sim_Z(mean = 0.5,`                  | Mean total mortality                                       | $\mu_Z$                             |
+| `+  Z = sim_Z(mean = 0.5,`                  | Mean total mortality^2^                                    | $\mu_Z$                             |
 | `+            log_sd = 0.2,`                | Standard deviation of total mortality (log)                | $\sigma_Z$                          |
@@
 | `+  growth = sim_vonB(Linf = 120,`          | Mean asymptotic length (cm)                                | $l_{\infty}$                        |
-| `+                    L0 = 5,`              | Length at birth (cm)                                       | $l_0$                               |
+| `+                    L0 = 5,`              | Length in birth year (cm)                                  | $l_0$                               |
 | `+                    K = 0.1,`             | Growth rate                                                | $K$                                 |
@@
 | `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{\mathrm{group}}^{N}$            |
+
+^1^ Can be a vector of means with a length equal to the number of years in the simulation.  
+^2^ Can be a matrix of means with number of rows and columns equaling the number of ages and years in the simulation, respectively.
 
@@
 The next step in the simulation is to distribute the abundance at age matrix simulated using the cohort model throughout a spatial field.
-Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
+Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across longitude only, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
 The grid is divided into $L_{\mathrm{div}}$ divisions (e.g.
 NAFO or ICES divisions) and $L_{\mathrm{strat}}$ depth-based strata.
-The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space
+The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space via:
 
@@
 The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see [*Appendix B*][Appendix B: Age-year-space covariance] for a more detailed description of the space-age-year covariance structure).
-In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering.
-Note that $\eta_{a,y,s}$ is forced to sum to one in each year to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model.
-Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid.
+In short, this formulation allows control of depth preferences, the level of spatial aggregation and the degree of age and year specific clustering.
+Note that $\eta_{a,y,s}$ values are rounded and re-scaled to ensure that discrete numbers of fish reside in each grid cell and to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model.
 
@@
 | `+                   n_div = 1,`                        | Number of divisions                                          | $L_{\mathrm{div}}$                      |
-| `+                   strat_splits = 2,                  | Number of times to horizontally split strat                  | &nbsp;                                  |
+| `+                   strat_splits = 2,`                 | Number of times to horizontally split strat                  | &nbsp;                                  |
 | `+                   strat_breaks = seq(0, 1000, 40)),` | Series of depth breaks for defining strata                   | &nbsp;                                  |
@@
 If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure is used.
-The package includes a manually constructed survey grid of NAFO Subdivision 3Ps (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid.
+The package includes a manually constructed survey grid of NAFO Subdivision 3Ps off the southern coast of Newfoundland (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid.
 However, for simplicity, we use `make_grid` to construct a square grid for a default run of `sim_distribution`.
@@
 
-The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated.
-The area of each strata is calculated and this is used to define the number of sets allocated to each strata under a particular set density, $D_{\mathrm{sets}}$.
+The final step in the simulation is to sample the simulated population over the age-year-space array generated.
+The sampling is stratified random, emulating real-world surveys conducted by many research institutions around the world.
+The area of each strata is calculated and this is used to define the number of sampling stations, hereafter referred to as sets, allocated to each strata under a particular set density, $D_{\mathrm{sets}}$.
 The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:
@@
 
-Where $n_{a,y,s}$ is the number of fish of age `a` in year `y` sampled by a set at location `s`, $A_{\mathrm{trawl}}$ indicates the area covered by the trawl, $A_{\mathrm{cell}}$ is the area of a grid cell, and $q_a$ is the catchability coefficient of each age.
+Where $n_{a,y,s}$ is the number of fish of age $a$ in year $y$ sampled by a set at location $s$, $A_{\mathrm{trawl}}$ indicates the area covered by the trawl, $A_{\mathrm{cell}}$ is the area of a grid cell, and $q_a$ is the catchability coefficient of each age (i.e.
+the ability of the trawling gear to catch specific age groups).
 Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
@@
 Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division.
-Such sub-sampling is common in fisheries-independent surveys as it is costly to sample every fish captured.
+Such sub-sampling is common in fisheries-independent surveys as it is costly, impractical and unnecessary to sample every fish captured.
 Age determination is especially time-consuming, which is why otoliths for age-determination tend to be sub-sampled by length-bin to obtain a representative age sample across a wider range of lengths than would be obtained via random sampling.
@@
 
-Survey data need to be analyzed to obtain an index of abundance.
-While there are many model-based options for obtaining an index [e.g.
+While there are many model-based options for obtaining an abundance index from survey data [e.g.
 @Thorson2015], design-based approaches, such as stratified analyses, are often used.
@@
 
-<!-- ESTIMATION OF VARIANCE. What does the stratified estimates of variance look like with the simulations you already have? Do negative values disapear above a praticular sampling intensity? -->
+<!-- ESTIMATION OF VARIANCE. What does the stratified estimates of variance look like with the simulations you already have? Do negative values disapear above a particular sampling intensity? -->
 
@@
 
-**Review of previous draft:** Dave Cote, Joanne Morgan
+**Review of previous draft:** Dave Cote, Joanne Morgan, Brian Healey
 

### 2019-09-18 analysis/paper/sim_survey_paper.Rmd
Prep manuscript for plos one submission

@@
 title: "`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
-author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, Jonathan Babyn, Fran Mowbray"
-date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
 output: 
@@
 bibliography: references.bib
-csl: ices-journal-of-marine-science.csl
+csl: plos.csl
 ---
@@
 
-<!-- Items to discuss -->
-<!-- - Title -->
-<!-- - Target journal -->
-<!-- - Actual figures in appendix (html article, with interactive figures, as supplement?) -->
+<br>
+<br>
+
+Paul M. Regular^1^*, Gregory J. Robertson^1^, Keith P. Lewis^1^, Jonathan Babyn^1^, Brian Healey^1^, Fran Mowbray^1^
+
+<br>
+<br>
+
+^1^ Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada
+
+<br>
+<br>
+
+*Corresponding author  
+E-mail: Paul.Regular@dfo-mpo.gc.ca (PMR)
+
+<br>
+<br>
 
@@
 
-# Summary
-
-**1\.** Fish populations often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective surveys.
+# Abstract
+
+Fish populations often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective surveys.
 Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision and increasing the risk of bias) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics).
-
-**2\.** For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of fish captured to measure biological characteristics (e.g.
+For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of fish captured to measure biological characteristics (e.g.
 lengths or ages).
 Population estimates are also affected by the pathway taken to analyze such data.
-
-**3\.** Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for fisheries-independent survey data.
-
-**4\.** Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations.
+Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for fisheries-independent survey data.
+Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations.
 The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of sampling protocols to reproduce the population parameters of the known population.
 **`SimSurvey`** also includes a function for estimating the stratified mean and variance of the simulated data.
-
-**5\.** **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring.
+**`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring.
 Various statistical approaches can then be applied to the results to test the efficacy of different analytical approaches.
 
-**Key-words:** length and age distribution; simulation; spatial correlation; stratified analysis; survey design
+<!-- **Key-words:** length and age distribution; simulation; spatial correlation; stratified analysis; survey design -->
 
@@
 While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study.
-Output from default function calls are therefore relevant to the case study and these results are described and discussed in [*Appendix A*][Appendix A: Case study].
+Output from default function calls are therefore relevant to the case study and these results are described and discussed in [*S1 Appendix*][S1 Appendix: Case study].
 
@@
 Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
-The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see [*Appendix B*][Appendix B: Age-year-space covariance] for a more detailed description of the space-age-year covariance structure).
+The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see [*S2 Appendix*][S2 Appendix: Age-year-space covariance] for a more detailed description of the space-age-year covariance structure).
 In short, this formulation allows control of depth preferences, the level of spatial aggregation and the degree of age and year specific clustering.
@@
 @Thorson2015], design-based approaches, such as stratified analyses, are often used.
-Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in [*Appendix C*][Appendix C: Stratified analysis equations]] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
+Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in [*S3 Appendix*][S3 Appendix: Stratified analysis equations]] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
 Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be scaled up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key.
@@
 Keep in mind that low numbers of `n_sims` and high numbers of `n_loops` will be less demanding on RAM, but may be slower, especially if the work is spread across few `cores`.
-Because most of the default settings of the functions match the case study settings, the code below will replicate the results from our case study (see [*Appendix A*][Appendix A: Case study] for more detail).
+Because most of the default settings of the functions match the case study settings, the code below will replicate the results from our case study (see [*S1 Appendix*][S1 Appendix: Case study] for more detail).
 The `expand_surveys` function sets up a series of `r nrow(expand_surveys())` surveys to test (i.e.
@@
 
-The case study described in [*Appendix A*][Appendix A: Case study] provides one example of how **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
+The case study described in [*S1 Appendix*][S1 Appendix: Case study] provides one example of how **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
 There are multiple layers to the sampling design of such surveys and the end results are linked with the way such data are analyzed.
@@
 One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012].
-Results from the case study (see [*Appendix A*][Appendix A: Case study]) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
+Results from the case study (see [*S1 Appendix*][S1 Appendix: Case study]) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
 @Berg2012].
@@
 [-@puerta2019].
-
-<!-- ESTIMATION OF VARIANCE. What does the stratified estimates of variance look like with the simulations you already have? Do negative values disapear above a particular sampling intensity? -->
 
@@
 
-**Feedback and advice:** Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
-
-**Review of previous draft:** Dave Cote, Joanne Morgan, Brian Healey
-
-**Funding:** NSERC visiting-fellow program
-
-# Appendix A: Case study
+This work has benifited from valuable feedback from numerous colleagues, including Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury and Peter Upward.
+We also thank Dave Cote and Joanne Morgan for providing constructive comments on a previous version of this manuscript.
+This work was supporded by the NSERC visiting-fellow program.
+
+# S1 Appendix: Case study
 
@@
 
-# Appendix B: Age-year-space covariance
+# S2 Appendix: Age-year-space covariance
 
@@
 
-# Appendix C: Stratified analysis equations
+# S3 Appendix: Stratified analysis equations
 
@@
 # References
-
-<!-- # Quotes -->
-
-<!-- "To effectively manage rare populations, accurate monitoring data are critical.
-Yet many monitoring programs are initiated without careful consideration of whether chosen sampling designs will provide accurate estimates of population parameters.
-Obtaining accurate estimates is especially difficult when natural variability is high, or limited budgets determine that only a small fraction of the population can be sampled." - Morrison et al.
-(2008) Popul Ecol 50:417–425 -->
-
-<!-- "Fishery independent surveys provide valuable information on fish population characteristics and play important roles in fisheries stock assessment (Gunderson, 1993; Rago, 2005).
-The importance of fishery independent surveys has been well recognized, but the important role of sampling design for these surveys is less appreciated.
-Often, traditional survey methods are used without considering survey designs that maximize the precision and accuracy of the estimator (Mier and Picquelle, 2008).
-Fishery independent surveys are usually expensive and time-consuming, and an efficient sampling design is key to the success of these surveys.
-Recently, investigators have begun to pay more attention to the importance of survey sampling and are exploring a variety of aspects of designs (Bez, 2002; Brown, 2003; Kimura and Somerton, 2006).
-Based on the previous studies, we compared the performance of traditional and adaptive sampling designs with emphasis on simulation application for performance comparison." - Yu et al.
-(2012) Fisheries Research 113: 173–181 -->
-
-<!-- # Examples -->
-
-<!-- Good examples of papers on packages: -->
-<!-- - https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0092725#s4 -->
-<!-- - file:///C:/Users/RegularP/Downloads/Mildenberger_et_al-2017-Methods_in_Ecology_and_Evolution.pdf -->
-
-<!-- # READ -->
-
-<!-- - https://onlinelibrary.wiley.com/doi/pdf/10.1046/j.1467-2960.2001.00047.x -->

### 2019-09-19 analysis/paper/sim_survey_paper.Rmd
Render both html and word copies of the paper

@@
     reference_docx: template.docx
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: true
 bibliography: references.bib
@@
                       dev = "png",
-                      dpi = 300)
+                      dpi = 300,
+                      out.width = "100%")
 knitr::opts_knit$set(root.dir = "../..")
@@
 replot <- FALSE # plot the figures again? (acts as a manual cache)
+is_html <- knitr::is_html_output()
 
@@
 <br>
-<br>
 
@@
 
-<br>
-<br>
-
 ^1^ Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada
-
-<br>
-<br>
 
@@
 
-<br>
-<br>
-
-# Things to manually fix in word
-
-1. Adjust font size of code strings
-2. Decrease font size of tables
-3. Indent closures in Table \@ref(tab:fun)
-4. Replace pluses (`+`) with spaces in Tables \@ref(tab:abundance) to (\#tab:survey)
-5. Adjust size of footnote below Table \@ref(tab:abundance)
-6. Adjust size of Figure \@ref(fig:real-sim-comp)
-7. Remove this section
+```{r, eval = !is_html, results = "asis"}
+
+cat("# Things to manually fix in word  \n")
+
+cat("1. Adjust font size of code strings  \n")
+cat("2. Decrease font size of tables  \n")
+cat("3. Indent closures in Table \\@ref(tab:fun)  \n")
+cat("4. Replace pluses (`+`) with spaces in Tables \\@ref(tab:abundance) to \\@ref(tab:survey)  \n")
+cat("5. Adjust size of footnote below Table \\@ref(tab:abundance)  \n")
+cat("6. Adjust size of Figure \\@ref(fig:real-sim-comp)  \n")
+cat("7. Place supplements in separate documents  \n")
+cat("8. Remove this section  \n")
+
+```
 
@@
 
-```{r, echo = TRUE, eval = replot}
+```{r, echo = TRUE, eval = replot | is_html}
 set.seed(438)
@@
 
-```{r, eval = replot}
+```{r, eval = replot | is_html}
 
@@
 
-export_plot(p, file = "analysis/paper/figures/plot_surface.png", width = 800, height = 370)
-```
-
-```{r surface, fig.cap = "Surface plots of simulated abundance at age of a relatively a) long lived and b) short lived species.
+if (replot) export_plot(p, file = "analysis/paper/figures/plot_surface.png", width = 800, height = 370)
+
+```
+
+```{r surface, out.height = "370px", fig.cap = "Surface plots of simulated abundance at age of a relatively a) long lived and b) short lived species.
 These plots were produced by `plot_surface` when supplied a list produced by `sim_abundance`."}
 
+if (is_html) {
+  p
+} else {
   knitr::include_graphics("figures/plot_surface.png")
+}
 
@@
 
-```{r eval = replot}
+```{r eval = replot | is_html}
 
@@
 
-export_plot(p, file = "analysis/paper/figures/plot_grid.png", width = 800, height = 320)
-
-```
-
-```{r grid, fig.cap = "Plots produced by `plot_grid` when supplied a raster object produced by `make_grid` a) using default settings, b) settings that increase the number of times depth strata are horizontally split, and c) settings that produce a more linear depth gradient and increase the number of divisions.
+if (replot) export_plot(p, file = "analysis/paper/figures/plot_grid.png", width = 800, height = 320)
+
+```
+
+```{r grid, out.height = "320px", fig.cap = "Plots produced by `plot_grid` when supplied a raster object produced by `make_grid` a) using default settings, b) settings that increase the number of times depth strata are horizontally split, and c) settings that produce a more linear depth gradient and increase the number of divisions.
 In these plots, the color gradient represents depth, the thick grey lines delineate divisions and thin white lines delineate strata."}
 
+if (is_html) {
+  p
+} else {
   knitr::include_graphics("figures/plot_grid.png")
+}
 
@@
 
-```{r, echo = TRUE, eval = replot}
+```{r, echo = TRUE, eval = replot | is_html}
 
@@
 
-export_plot(p, file = "analysis/paper/figures/plot_distribution.png", width = 600, height = 300)
-
-```
-
-```{r distribution, fig.cap = "Distribution plots of simulated populations that form a) tight clusters with stronger correlation through years than ages (default settings), and b) relatively diffuse clusters with stronger correlation through ages than years.
-This plot is a facet of plots produced by `plot_distribution` when supplied simulations from `sim_distribution`."}
-
+if (replot) export_plot(p, file = "analysis/paper/figures/plot_distribution.png", width = 600, height = 300)
+
+a_plots <- plot_distribution(a, ages = 1:3, years = 1:3, type = "heatmap") %>% 
+  add_labels(x = 0.08, y = 1, text = "a)")
+b_plots <- plot_distribution(b, ages = 1:3, years = 1:3, type = "heatmap") %>% 
+  add_labels(x = 0.08, y = 1, text = "b)")
+
+```
+
+```{r}
+cap <- "Distribution plots of simulated populations that form a) tight clusters with stronger correlation through years than ages (default settings), and b) relatively diffuse clusters with stronger correlation through ages than years."
+if (is_html) {
+  cap <- paste(cap, "These plots were produced by `plot_distribution` when supplied simulations from `sim_distribution`.")
+} else {
+  cap <- paste(cap, "This plot is a facet of plots produced by `plot_distribution` when supplied simulations from `sim_distribution`.")
+}
+```
+
+```{r eval = is_html, out.height = "400px"}
+a_plots
+```
+
+```{r distribution, out.height = "400px", fig.cap = cap}
+
+if (is_html) {
+  b_plots
+} else {
   knitr::include_graphics("figures/plot_distribution.png")
+}
 
@@
 ## plotly::subplot was not amenable to combining these two plots, I therefore used ImageMagick to combine a and b
-
+if (replot) {
   paths <- c(a = "analysis/paper/figures/plot_survey_a.png", b = "analysis/paper/figures/plot_survey_b.png")
@@
   magick::image_write(p, path = "analysis/paper/figures/plot_survey.png")
-
-```
-
-```{r survey, fig.cap = "Bubble plots of abundance and histograms of set catches from a simulated stratified-random survey of a default population under relatively a) low and b) high sampling effort.
+}
+
+```
+
+```{r eval = is_html, out.height = "400px"}
+plot_a
+```
+
+```{r survey, out.height = "400px", fig.cap = "Bubble plots of abundance and histograms of set catches from a simulated stratified-random survey of a default population under relatively a) low and b) high sampling effort.
 Point size and color are scaled by abundance in the bubble plots.
@@
 
+if (is_html) {
+  plot_b
+} else {
   knitr::include_graphics("figures/plot_survey.png")
+}
 
@@
 
-```{r data, eval = replot}
+```{r data, eval = replot | is_html}
 
@@
 
-export_plot(p, file = "analysis/paper/figures/plot_total_strat_fan.png", width = 700, height = 250)
-
-```
-
-```{r fan1, fig.cap = "Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{\\mathrm{sets}}$.
+if (replot) export_plot(p, file = "analysis/paper/figures/plot_total_strat_fan.png", width = 700, height = 250)
+
+```
+
+```{r fan1, out.height = "250px", fig.cap = "Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{\\mathrm{sets}}$.
 The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
@@
 
+if (is_html) {
+  p
+} else {
   knitr::include_graphics("figures/plot_total_strat_fan.png")
-
-```
-
-```{r eval = replot}
+}
+
+```
+
+```{r eval = replot | is_html}
 
@@
   }
+  rm(data)
   subplot(plot_list, nrows = length(set_dens), shareX = TRUE, shareY = TRUE, titleX = FALSE, titleY = FALSE)
@@
 
-export_plot(p, file = "analysis/paper/figures/plot_length_strat_fan.png", width = 700, height = 600)
-
-```
-
-```{r fan2, fig.cap = "Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{\\mathrm{sets}}$, and length sampling protocol, $M_{\\mathrm{lengths}}$.
+if (replot) export_plot(p, file = "analysis/paper/figures/plot_length_strat_fan.png", width = 700, height = 600)
+
+```
+
+```{r fan2, out.height = "600px", fig.cap = "Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{\\mathrm{sets}}$, and length sampling protocol, $M_{\\mathrm{lengths}}$.
 The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
@@
 
+if (is_html) {
+  p
+} else {
   knitr::include_graphics("figures/plot_length_strat_fan.png")
-
-```
-
-```{r, eval = replot}
+}
+
+```
+
+```{r, eval = replot | is_html}
 
@@
 
-export_plot(p, file = "analysis/paper/figures/plot_age_strat_fan.png", width = 700, height = 600)
-
-```
-
-```{r fan3, fig.cap = "Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{\\mathrm{sets}}$, and length sampling protocol, $M_{\\mathrm{lengths}}$.
+if (replot) export_plot(p, file = "analysis/paper/figures/plot_age_strat_fan.png", width = 700, height = 600)
+
+```
+
+```{r fan3, out.height = "600px", fig.cap = "Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{\\mathrm{sets}}$, and length sampling protocol, $M_{\\mathrm{lengths}}$.
 Number of ages sampled per length group, $M_{\\mathrm{ages}}$, was 10 in all scenarios.
@@
 
+if (is_html) {
+  p
+} else {
   knitr::include_graphics("figures/plot_age_strat_fan.png")
+}
 
@@
 
-p <- plot_survey_rank(sim, which_strat = "length")
-export_plot(p, file = "analysis/paper/figures/plot_survey_rank.png", width = 450, height = 400)
+rank_p <- plot_survey_rank(sim, which_strat = "length")
+if (replot) export_plot(rank_p, file = "analysis/paper/figures/plot_survey_rank.png", width = 450, height = 400)
 
@@
 
-p <- subplot(p_list) %>%  
+surface_p <- subplot(p_list) %>% 
   layout(scene = c(list(domain = list(x = c(0, 1/3), y = c(0, 1))), scene),
@@
          scene3 = c(list(domain = list(x = c(2/3, 1), y = c(0, 1))), scene),
-         autosize = FALSE, margin = list(t = 0, l = 10, r = 10, b = 10))
-
-export_plot(p, file = "analysis/paper/figures/plot_error_surface.png", width = 990, height = 270)
-
-```
-
-```{r rank, fig.cap = "Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{\\mathrm{sets}}$] and length measurements [$N_{\\mathrm{measured}}$]), under various sampling protocols (set density [$D_{\\mathrm{sets}}$] and maximum number of lengths measured per set [$M_{\\mathrm{lengths}}$]).
+         autosize = is_html, margin = list(t = 0, l = 10, r = 10, b = 10))
+
+if (replot) export_plot(surface_p, file = "analysis/paper/figures/plot_error_surface.png", width = 1090, height = 320)
+
+surface_p <- plot_error_surface(tests, plot_by = "rule")
+
+```
+
+```{r rank, out.height = "400px", fig.cap = "Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{\\mathrm{sets}}$] and length measurements [$N_{\\mathrm{measured}}$]), under various sampling protocols (set density [$D_{\\mathrm{sets}}$] and maximum number of lengths measured per set [$M_{\\mathrm{lengths}}$]).
 Records are ranked by lowest to highest RMSE score.
@@
 
+if (is_html) {
+  rank_p
+} else {
   knitr::include_graphics("figures/plot_survey_rank.png")
-
-```
-
-```{r rmse-surface, fig.cap = "Surface plots of RMSE from an array of surveys with different sampling protocol.
+}
+
+```
+
+```{r}
+cap <- "Surface plots of RMSE from an array of surveys with different sampling protocol.
 Panels represent surveys with different set densities ($D_{\\mathrm{sets}}$), x-axes represent the maximum sampling effort of lengths per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per length group ($M_{\\mathrm{ages}}$).
 This plot is a facet of plots produced by `plot_error_surface` when supplied results from `test_surveys`.
-Note that RMSE scales are different across $D_{\\mathrm{sets}}$ facets.
+Note that RMSE scales are different across $D_{\\mathrm{sets}}$ panels.
 Note that the exponent format of the axes defaults to SI unit symbols (e.g.
-M for million)."}
-
+M for million)."
+if (is_html) {
+  cap <- gsub(" a facet of plots ", " ", cap)
+}
+
+```
+
+```{r rmse-surface, out.height = "500px", fig.cap = cap}
+
+if (is_html) {
+  surface_p
+} else {
   knitr::include_graphics("figures/plot_error_surface.png")
+}
 
@@
 
-```{r real-sim-comp, fig.cap = "Comparison of a) real and b) simulated survey data where, from top to bottom, histograms of numbers caught across sets, distributions of set catches, and length-frequency distributions from three random sets are shown from a randomly selected year from the real and simulated 3Ps cod survey data.
+```{r real-sim-comp, out.width = "90%", fig.cap = "Comparison of a) real and b) simulated survey data where, from top to bottom, histograms of numbers caught across sets, distributions of set catches, and length-frequency distributions from three random sets are shown from a randomly selected year from the real and simulated 3Ps cod survey data.
 The colors of the length-frequency plots correspond to the colored locations.

### 2019-09-19 analysis/paper/render_paper.R

@@ -1,0 +1,8 @@
+
+## Render both word and html documents
+rmarkdown::render("analysis/paper/sim_survey_paper.Rmd", output_format = "all")
+
+## Replace the +'s in the tables with empty space
+ms <- readLines("analysis/paper/sim_survey_paper.html")
+ms <- gsub("<code>[+]", "<code>&nbsp;", ms)
+writeLines(ms, "analysis/paper/sim_survey_paper.html")

### 2019-09-20 analysis/paper/cover_letter.Rmd
Draft a cover letter for PLOS ONE

---
output: word_document
---

Joerg Heber, Ph.D.  
Editor-in-Chief  
PLOS ONE  
1160 Battery Street  
Koshland Building East, Suite 225  
San Francisco, CA 94111  
United States

`r format(Sys.time(), '%Y-%m-%d')`

Dear Dr. Heber,

I am writing to submit our manuscript entitled, **"`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"**, for consideration as a **Research Article** in **PLOS ONE**. In our manuscript we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations. We believe our work to be a valuable contribution to the scientific literature as few tools have been built to facilitate simulation tests of the design and analysis of costly fishery independent surveys. Such surveys have become an integral part of the population monitoring programs led by many fisheries organizations and, as with any other monitoring program, good survey design and analysis is a critical prerequisite for obtaining reliable results. It has long been recognized that simulations are a useful tool for optimizing such surveys and, as such, several frameworks have been proposed and applied to try and answer the common question of "how much sampling is enough?"(e.g. Schnute and Haigh,  Canadian Journal of Fisheries and Aquatic Sciences, 2003). The **`SimSurvey`** package builds upon previous frameworks but extends the accessibility of the underlying simulation by wrapping it in an open-source `R` package.

This is our first interaction with PLOS regarding the submitted manuscript. We would like to suggest George Tserpes or Rui Coelho to act as the Academic Editor given their background in fisheries science.

We look forward to your decision.

Sincerely,

Paul Regular  
Fisheries and Oceans Canada  
Northwest Atlantic Fisheries Center  
80 East White Hills, St. John’s, NL  
A1C 5X1, Canada  
E-mail: Paul.Regular@dfo-mpo.gc.ca  
Phone: (709) 772-2067

### 2019-09-24 analysis/paper/cover_letter.Rmd
Edit cover letter given Greg's suggestions

@@
 I am writing to submit our manuscript entitled, **"`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"**, for consideration as a **Research Article** in **PLOS ONE**.
-In our manuscript we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations.
+In our manuscript we introduce the `R` package **`SimSurvey`**, which has been designed to allow the process of simulating surveys of complex age-structured and spatially-correlated fish populations.
 We believe our work to be a valuable contribution to the scientific literature as few tools have been built to facilitate simulation tests of the design and analysis of costly fishery independent surveys.
@@
 Schnute and Haigh,  Canadian Journal of Fisheries and Aquatic Sciences, 2003).
-The **`SimSurvey`** package builds upon previous frameworks but extends the accessibility of the underlying simulation by wrapping it in an open-source `R` package.
+The **`SimSurvey`** package builds upon previous frameworks by increasing the complexity of the simulated population to reflect the dynamic structure seen in fish populations and extends the accessibility of the underlying simulation by wrapping it in an open-source `R` package.
 

### 2019-11-26 analysis/exports_for_Geoff.R
New real 3Ps cod data export for Geoff

@@ -229,14 +229,14 @@ comp_fun(x_y_year = c(110.25, 68.25, 7),
 
 
 
-## 2018-12-17 - real 3Ps data --------------------------------------------------
+## 2019-11-26 - real 3Ps data --------------------------------------------------
 
 library(Rstrap)
 
 ## Load survey data compiled for Rstrap
-load("analysis/rv_data/converted_set_details_2018-02-26.Rdata")
-load("analysis/rv_data/converted_length-frequency_data_2018-02-26.Rdata")
-load("analysis/rv_data/age-growth_data_2018-02-26.Rdata")
+load("analysis/rv_data/converted_set_details_2019-10-03.Rdata")
+load("analysis/rv_data/converted_length-frequency_data_2019-10-03.Rdata")
+load("analysis/rv_data/age-growth_data_2019-10-03.Rdata")
 
 ## Subset data to cod
 con.setdet <- con.setdet[(con.setdet$rec == 5 |
@@ -256,7 +256,7 @@ out <- strat.fun(setdet = rv_data$setdet, lf = rv_data$lf, ag = rv_data$ag,
                  group.by = "length & age",
                  export = NULL, plot.results = FALSE)
 
-fwrite(out$raw.data$set.details, file = "analysis/cod_sim_exports/2018-09-12_for_Geoff/cod_3Ps_setdet.csv")
-fwrite(out$raw.data$age.growth, file = "analysis/cod_sim_exports/2018-09-12_for_Geoff/cod_3Ps_samp.csv")
+fwrite(out$raw.data$set.details, file = "analysis/cod_sim_exports/2019-11-26_for_Geoff/cod_3Ps_setdet.csv")
+fwrite(out$raw.data$age.growth, file = "analysis/cod_sim_exports/2019-11-26_for_Geoff/cod_3Ps_samp.csv")
 
 

### 2019-12-12 analysis/paper/rebuttal_letter.Rmd
Start working through ms and rebuttal letter, building upon Keith's efforts

---
output: word_document
---

Daniel E. Duplisea, PhD  
Fisheries and Oceans Canada  
850 Route de la Mer  
Mont‐Joli, Quebec  
G5H 3Z4, Canada

`r format(Sys.time(), '%Y-%m-%d')`

Dear Dr. Duplisea,

Thank you for considering a revision of the manuscript PONE-D-19-26904,  **"`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"** by Paul M. Regular, Gregory J. Robertson, Bob Rogers, and Keith P. Lewis. In accordance to one of the reviewers' suggestions, the paper has been renamed **"`SimSurvey`: an `R` package to compare the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"**.

We thank the Reviewers for their thoughtful and thorough comments which will clarify and improve the manuscript.  We agree with many of the comments and have made every effort to incorporate these into the manuscript.  Below are the reviewers' comments with a brief accounting of our response to each concern or suggestion.

Lines numbers refer to the revised manuscript.

We submit this revised manuscript for your consideration and look forward to your decision.

Sincerely,

Paul Regular  
Fisheries and Oceans Canada  
Northwest Atlantic Fisheries Center  
80 East White Hills, St. John’s, NL  
A1C 5X1, Canada  
E-mail: Paul.Regular@dfo-mpo.gc.ca  
Phone: (709) 772-2067

Reviewer #1:

This paper describes the R package “SimSurvey” which is a package to test the design and analysis of fisheries independent surveys.

I found the paper very interesting and the package of great potential value and interest to many fishery scientists. However, I have several comments that I would like the authors to address first before I recommend acceptance of this paper. My main comments relate to:

1. Verify some of the equations and model description used in the model. I think that some of them lack description to fully understand what was done and few others are misleading, even incorrect. The detail can be found below but in general, they are linked to the problem of bias correction for log-normal distribution, re-scaling”, moving from abundance at age to abundance at length.

*Please see our specific responses to lines X, X, and X.*

2. Add additional explanation to justify the choices. E.g. why population dynamics is modeled independently of the spatial distribution. Is this reasonable, limitation, etc.

*TODO*

3. Few suggestions (add-ons, renaming) to increase the generality of the paper (without requiring too much work)
Please see our specific responses to...

*TODO*

Detailed comments:

Abstract:

What about the flexibility to include new sampling strategies (user defined one) or including new estimation approaches

*We have noted that both built-in and user-defined strategies can be utilised.*

Introduction:

L43: quality of

*We have made the suggested change.*

It would be good to add few references here e.g. L50,

*We have added references as suggested.*

Method:

L69-73: I think it would be good to add the reference to the specific function name listed in Table2. This way, I would be more explicit and avoid possible confusion. For example, in my case, when I first read the section header “simulate abundance”, I was expecting to see the way you dealt with the spatial distribution of abundance.

*We have made the suggested change.*

L94&95&104: I just want to make sure that the bias correction has been applied to these two equations. If logN (or logZ) is normally distribution with mean mu and sigma, then the N (or Z) is lognormally distributed with mean exp(mu+sigma^2/2). Therefore, when you write “mean” in table 3 for R and Z, I hope that you modified in your actual equation for log(Z) to:
log(Z) ~ normal(log(mean)-sigma^2/2, sigma)
In other word, the mean(R) is not simply exp(meanR_log_scale). exp(meanR_log_scale) is the median of the distribution of R.

*Good catch. The answer to the question is no, bias correction was not applied. This is a matter of naming convention that we struggled with. The intent was to operate in log space for many of these equations and, as such, the simulations were set-up to receive log means. When developing the functions, however, we waffled on whether to use `log_mean` or `mean` as the argument name and we landed on `mean` as we figured this would be a more relatable value for users to supply. Kudos to the reviewer for noticing that we failed to consider bias-correction. To be more explicit with regards to what we "mean", we have renamed the arguments to `log_mean`.*

L107: there is not enough description to understand how you converted the abundance at age to abundance at length. What were the choice of length bin groups? Every 5 cm, 10cm? Such information is also missing in the tables. Please add that. But then you have to calculate some cumulative distribution (until length l) for the length at age distribution (using the VB growth equation) and do some subtractions to be able to allocate the abundance at age to each length group.
See equation A.1.14 in here for example. https://www.nwfsc.noaa.gov/news/events/program_reviews/documents/C.2_Methot_Wetzel_SSTechnicalDescription.pdf

*TODO*

“Simulate spatial distribution”:

I have a few comments on this section:

1. First of all, I think it is important to state that you assume that the population dynamic model and the spatial distribution is independent i.e. you do not have a spatially-explicit population dynamics model. And talk about what it means (is it realistic, limitation, etc).

*TODO*

2. L161. I think it would be better to rename “depth” as “main covariate influencing the species distribution”. And “depth” happens to be one of them for many species, but there are cases when this is not the case.

*See next comment*

3. L161. I think it would be good to say upfront that there are two ways of defining the spatial structure. Using the function in the package or user defined.

*In response to this suggestion and the previous one, we have modified the text (lines X-X)*

4. L161: I think another possibility is to generate a random field by using the package “Randomfields” for example. This way you can generate a map where “depth” is patchily distributed (maybe more like an island type case study)

*TODO*

5. L163-164: please add more description about the division or strata. How can we set it up and what do you specifically mean by division and strata? Is one nested within the other, or not necessarily? Your examples are based on Atlantic Ocean and people in other regions might not be familiar with how these divisions are created.

*TODO*

6. L163: Why only focus on “depth-based strata”? I think it would be good to allow the user to choose their own stratification approach. It could be depth based as you did (which happens most often in surveys) but it could technically be any other thing (user supplied). This allows more flexibility.

*TODO*

7. L167: the equation is misleading and I am not sure it is right based on what is written in the text. You mentioned later on L178: that you “re-scaled” so that the total number of fish in a specific year and age across space is equal to the single number from the population dynamics model. If so, the re-scaling should be done in the identify (natural) scale, not in log scale. In log scale, even if ,, sums to zero across space, the sum in the identity scale won’t match. This is often refer to as “bias correction” for the log normal models. And you should ideally show how the rescaling was done in terms of the equations too.

*TODO*

8. L167: this “depth preference” function is very simplistic and gives only very “smooth” symmetric distribution. More often, fish have a skewed depth preference: often right-skewed.

*TODO*

9. L173-174: I think it might be worth adding, in simpler terms, the meaning of the spatial smoothing and scaling parameters.

*TODO*

10. L178: it is another question of scaling. How did you exactly do the scaling? In the identity scale? By dividing my the sum of the effects? I am asking this because depending on how you did the rescaling, your correlation structure in space and age might have been affected and is not the same as the one specified in L172. Did you verify that?

*TODO*

Table 4. “group_ages”. Ok but how is the variance controlled for the other age classes?

*TODO*

L189: “user supplied”. This is a good feature. However, I think it is important to mention here that user have to make sure that they use the correct projection method to ensure that each grid is of the same dimension.

*We have made the suggested change.*

L216: reference to figure 3?

*We have made the suggested change.*

L221: there is not “group_years” argument in Table 4.

*TODO*

L237: “this function”. I think it would be better to replace “this function” with “sim_distribution” as you do not mention the word “sim_distribution” in the sentence above this.

*We have made the suggested change.*

L158-251: In general, I think re-organising this section using sub-section headers could be useful. Just to guide the readers

*This is an excellent suggestion and we have inserted headers where appropriate.*

L254: you say that sampling is stratified random but SRS is also an option based on Table5. Please correct.

*TODO*

L257: what does this mean? Does this control the number of set but how is this calculated?

*TODO*

L257: I do not see how you control for the total number of set in the survey? How do you control it?

*TODO*

L261: I think you should mention here that you can also force the sample size (as seen on table 5). Moreover, in table 5, it would be good to set-up a “ages_min” for the minimum number of ages to sample […]” so that it gives the ability to fix the sample size if needed by writing the same value for “ages_min” and “ages_cap”.

*TODO*

L261: How are you making sure that the number of sampled fish for that specific cell, age and year won’t be above the total number of fish in that cell, age, and year? The probability value could be close to one and if you fish in a few a time, then you are at risk. Especially because your population dynamics model is not spatially explicit and is completely independent of the distribution function itself i.e. you can technically fish out all the fish in an area but it will be populated back the year after the way you implemented in this study… Maybe you need to put a condition (or just a note) for general users to make sure that this probability value is much below 1?

*TODO*

L267: I recommend to clarify something here. 1. Depending on the number of fish caught? What do you mean? What is the rule you used? 2. The way you coded, sample by age is first decided, then the corresponding length is calculated, then age-subsample is determined. In reality, length sample is taken in the field, then age sub-samples are taken. While similar, I do not think it always equal. Especially, when you start including some correlation structure in the sampling. By the way, did you consider including some correlation structure in the sampling process to make more realistic?

*TODO*

L275 Table 1 on should be table 5

*We have changed the page numbers accordingly.*

L275: Table5: “age_sammpling” should be “age_sampling”

*We have made the suggested change.*

L275: “min_sets” you have not described it yet and what is it? You have sample from all cells? If no, this is not realistic.

*TODO*

L279: Table5 not Table1?

*We have changed the page numbers accordingly.*

L285-286: Could you be more specific on how custom closures can be supplied and where?

*TODO*

L306: how are these catchability corrected abundance matrices calculated? It is important to write this information somewhere (or write “please refer to the section “Stratified analysis” for further information on the calculation of abundance indices”) or something alike and Appendix S3.

*We have adopted the Reviewers suggestion and referred the reader to the section on Stratified analysis.*

L336: I think it would be good to say that other methods exist and people can use it in this package (maybe)?

*TODO*

L421: color gradient. Even though it is obvious it might be good to say green to purple gradient.

*We have made the suggested change.*

L427: instead of “sampling protocol”, I think it would be more meaning full to say the maximum number of length samples.

*We have made the suggested change.*

L452: say that the color ramps from yellow to purple

*We have made the suggested change.*

S1 appendix: missing figure in S1

*TODO*

Reviewer #2:

This manuscript describes an R package called SimSurvey. The package includes a set of functions for simulating point-based fisheries survey designs, e.g. bottom trawl surveys, for estimating abundance indices. It focuses on number of stations and number of fish sampled ignores other constraints such as distance between stations and day-time duration which impose strong constraints on real surveys. The functions included allow the user to first simulate age-and length structured population dynamics, distribute individuals randomly in space (assuming a certain correlation structure), carry out a survey and finally calculate abundance indices from the simulated data. The package will likely be of interest to researchers wanting to explore the precision achievable with different survey designs. My comments regarding the package and the presentation are summarised below.

1. Optimization

The title announces a package for optimizing survey designs. As far as I can see the package does not allow survey design optimization, neither in terms of defining survey strata nor in terms of number of stations per stratum. The strata are defined by the user. The only option available for the number of stations is proportional to stratum surface; the user sets the minimum number of stations taken in the smallest stratum. It would be useful to be able to specify the total number of stations and test different allocation schemes, such as proportional to surface area (implemented), equal number per stratum, Neyman allocation (accounting for surface area and abundance variability), etc.
Please consider revising the title (e.g. “compare” and instead of “optimize)” and spell out the available sampling design options.

*We agree with the Reviewer and retitled the ms "`SimSurvey`: an `R` package to compare the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks".*

2. Manuscript structure

The manuscript might be easier to follow if the manuscript was restructured: 1) Model description, 2) Using SimSurvey. The later section would then group all example code which could again be subdivided into running simulations and exploring results (plot functions).

*TODO*

3. Parameterisation
To use SimSurvey for a real world problem realistic parameter values are needed. The package comes with default values chosen for a particular case study. However, no mention is made in the manuscript how to choose appropriate values for the many model parameters to tailor the simulations to a population of interest. I suggest the authors add a section on parameterisation and a table summarizing all parameters with a column specifying how to parameterize. For example, parameters for population dynamics and growth could probably be taken from the literature (or a stock assessment report). However this is not possible for the parameters of the spatial distribution function sim_distribution() such as correlation between ages etc. Ideally the package would include a fitting function for estimating these parameters from actual data. These input data could come from a pilot survey and include location (lat, long) and numbers by length/age.

*TODO*

Minor issues
- line 93: I assume there is an age plus but this needs to be mentioned. Also, please specify how you set the initial numbers for plus group ().

*TODO*

- line 121: I don’t understand the explanation of a closure. What do you mean by “return functions”? Do you mean it returns an object with different attributes?

*TODO*

- line 125: the number of right and left brackets is unbalanced, please check

*We have made the suggested change.*

- line 126 “This structure was chosen to avoid the repeated specifications of ages and years”. As far as I can see the example code only specifies years, not ages.

*TODO*

- line 227 Please explain what a pipe is and how it is used. In the example I understand that the output of sim_abundance( ) is provided to (piped) into sim_distribution(). I am unclear what the object b contains. Is it the result of sim_distribution()?

*CHECK*

*We have added the following text: Also note that the key functions in the SimSurvey package have been set-up to be pipe [13] friendly, i.e. pipes are an operator to forward the values into a function or call expression.*

*Object b is the same as Object a but the variance in the size of clusters has been changed. We believe that this is clear from the text on L228-231 and the code on L235-243 but defer to the editor’s judgement on this matter.*

- There is no table 1, please revise table numbering.

*We have made the suggested change.*

### 2019-12-12 analysis/paper/sim_survey_paper.Rmd
Minor tweaks to paper and cover letter
Start working through ms and rebuttal letter, building upon Keith's efforts

@@
 ---
-title: "`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
+title: "`SimSurvey`: an `R` package to compare the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
 output: 
@@
 Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations.
-The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of sampling protocols to reproduce the population parameters of the known population.
+The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of built-in or user-defined sampling protocols to reproduce the population parameters of the known population.
 **`SimSurvey`** also includes a function for estimating the stratified mean and variance of the simulated data.
@@
 Fisheries-independent surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies.
-While costly to obtain, this information forms the basis of many stock assessments throughout the world [@Pennington1998] and the quality this information depends on surveys and analyses that maximize information while minimizing the expense of data collection.
+While costly to obtain, this information forms the basis of many stock assessments throughout the world [@Pennington1998] and the quality of this information depends on surveys and analyses that maximize information while minimizing the expense of data collection.
 While simulations provide a platform for exploring solutions to this optimization problem, building the necessary simulation framework is not a trivial task given the multi-stage nature of the typical sampling program and the complexity of the population processes we aim to represent.
 Notably, many fish stocks show size-specific patterns of aggregation, making it difficult to collect sufficient biological sub-samples (e.g.
-lengths and ages) to represent the entire target population.
+lengths and ages) to represent the entire target population [@Pennington1994; @Pennington2002; @Stewart2014].
 This undermines the assumptions of many analyses as sub-samples from such populations tend to exhibit strong positive intracluster correlation [@Aanes2015].
@@
 
-The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtual fish populations with correlation across space, time and age, 2) simulating surveys with a range of sampling strategies and intensities, and 3) estimating the stratified mean and variance of simulated survey data (Table \@ref(tab:fun)).
+The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtual fish populations with correlation across space, time and age (`sim_abundance`, `sim_distribution`), 2) simulating surveys with a range of sampling strategies and intensities (`sim_survey`), and 3) estimating the stratified mean and variance of simulated survey data (`test_surveys`; Table \@ref(tab:fun)).
 The equations behind these functions are detailed in the sections below.
@@
 We also thank Dave Cote and Joanne Morgan for providing constructive comments on a previous version of this manuscript.
-This work was supporded by the NSERC visiting-fellow program.
+This work was supported by the NSERC visiting-fellow program and Fisheries and Oceans Canada.
 

### 2019-12-16 analysis/paper/sim_survey_paper.Rmd
Rename argument from `mean` to `log_mean` inside the `sim_R` and `sim_Z` functions

@@
 | `+  years = 1:20,`                          | Years                                                      | $y$                                 |
-| `+  R = sim_R(mean = 30000000,`             | Mean recruitment^1^                                        | $\mu_r$                             |
+| `+  R = sim_R(log_mean = log(30000000),`    | Mean recruitment^1^                                        | $\mu_r$                             |
 | `+            log_sd = 0.5),`               | Standard deviation of log-recruitment                      | $\sigma_r$                          |
-| `+  Z = sim_Z(mean = 0.5,`                  | Mean total mortality^2^                                    | $\mu_Z$                             |
+| `+  Z = sim_Z(log_mean = log(0.5),`         | Mean total mortality^2^                                    | $\mu_Z$                             |
 | `+            log_sd = 0.2,`                | Standard deviation of total mortality (log)                | $\sigma_Z$                          |
@@
 Closures are functions that contain data and return functions [@wickham2014].
-For example, `sim_R(mean = 500, log_sd = 0.5)` returns a function that holds the supplied parameter values and requires a sequence of years to be supplied.
-(Note that each of the above-mentioned closures include a plot argument such that quick visuals can be obtained using a line of code like this: `sim_R(mean = 500, log_sd = 0.5, plot = TRUE)(years = 1:100)`).
+For example, `sim_R(log_mean = log(500), log_sd = 0.5)` returns a function that holds the supplied parameter values and requires a sequence of years to be supplied.
+(Note that each of the above-mentioned closures include a plot argument such that quick visuals can be obtained using a line of code like this: `sim_R(log_mean = log(500), log_sd = 0.5, plot = TRUE)(years = 1:100)`).
 This structure was chosen to avoid the repeated specifications of ages and years.
@@
 long <- sim_abundance(ages = 1:20,
-                      R = sim_R(mean = 3e+07),
-                      Z = sim_Z(mean = 0.2))
+                      R = sim_R(log_mean = log(3e+07)),
+                      Z = sim_Z(log_mean = log(0.2)))
 short <- sim_abundance(ages = 1:6,
-                       R = sim_R(mean = 1e+10),
-                       Z = sim_Z(mean = 0.8))
+                       R = sim_R(log_mean = log(1e+10)),
+                       Z = sim_Z(log_mean = log(0.8)))
 ```
@@
   years = 1:20,
-  R = sim_R(mean = 30000000,
+  R = sim_R(log_mean = log(30000000),
             log_sd = 0.5),
-  Z = sim_Z(mean = 0.5,
+  Z = sim_Z(log_mean = log(0.5),
             log_sd = 0.2,

### 2019-12-16 R/sim_pop.R

@@ -5,7 +5,7 @@
 #' @description These functions return a function to use inside \code{\link{sim_abundance}}.
 #' Given parameters, it generates N0, R and Z values.
 #'
-#' @param mean One mean value or a vector of means of length equal to years for \code{sim_R} or a matrix of means with
+#' @param log_mean One mean value or a vector of means, in log scale, of length equal to years for \code{sim_R} or a matrix of means with
 #' rows equaling the number of ages and colums equaling the number of years for \code{sim_Z}.
 #' @param random_walk Simulate recruitment as a random walk?
 #' @param log_sd Standard deviation of the variable in the log scale.
@@ -25,22 +25,22 @@
 #'
 #' @examples
 #'
-#' R_fun <- sim_R(mean = 100000, log_sd = 0.1, random_walk = TRUE, plot = TRUE)
+#' R_fun <- sim_R(log_mean = log(100000), log_sd = 0.1, random_walk = TRUE, plot = TRUE)
 #' R_fun(years = 1:100)
-#' sim_abundance(R = sim_R(mean = 100000, log_sd = 0.5))
+#' sim_abundance(R = sim_R(log_mean = log(100000), log_sd = 0.5))
 #' sim_abundance(years = 1:20,
-#'               R = sim_R(mean = c(rep(100000, 10), rep(10000, 10)), plot = TRUE))
+#'               R = sim_R(log_mean = log(c(rep(100000, 10), rep(10000, 10))), plot = TRUE))
 #'
-#' Z_fun <- sim_Z(mean = 0.5, log_sd = 0.1, phi_age = 0.9, phi_year = 0.9, plot = TRUE)
+#' Z_fun <- sim_Z(log_mean = log(0.5), log_sd = 0.1, phi_age = 0.9, phi_year = 0.9, plot = TRUE)
 #' Z_fun(years = 1:100, ages = 1:20)
-#' sim_abundance(Z = sim_Z(mean = 0.5, log_sd = 0.1, plot = TRUE))
+#' sim_abundance(Z = sim_Z(log_mean = log(0.5), log_sd = 0.1, plot = TRUE))
 #' Za_dev <- c(-0.2, -0.1, 0, 0.1, 0.2, 0.3, 0.3, 0.2, 0.1, 0)
[…]

### 2019-12-16 R/run_strat.R
Rename argument from `mean` to `log_mean` inside the `sim_R` call in the examples in the run_strat.R file

@@ -165,7 +165,8 @@ strat_means <- function(data = NULL, metric = NULL, strat_groups = NULL,
 #'
 #' @examples
 #'
-#' sim <- sim_abundance(ages = 1:10, years = 1:5, R = sim_R(mean = 1e+7),
+#' sim <- sim_abundance(ages = 1:10, years = 1:5,
+#'                      R = sim_R(log_mean = log(1e+7)),
 #'                      growth = sim_vonB(length_group = 1)) %>%
 #'            sim_distribution(grid = make_grid(res = c(10, 10)),
 #'                             ays_covar = sim_ays_covar(sd = 1)) %>%
@@ -224,7 +225,8 @@ run_strat <- function(sim,
 #'
 #' @examples
 #'
-#' sim <- sim_abundance(ages = 1:10, years = 1:5, R = sim_R(mean = 1e+7),
+#' sim <- sim_abundance(ages = 1:10, years = 1:5,
+#'                      R = sim_R(log_mean = log(1e+7)),
 #'                      growth = sim_vonB(length_group = 1)) %>%
 #'            sim_distribution(grid = make_grid(res = c(10, 10)),
 #'                             ays_covar = sim_ays_covar(sd = 1)) %>%

### 2019-12-16 analysis/paper/rebuttal_letter.Rmd
Rename argument from `mean` to `log_mean` inside the `sim_R` and `sim_Z` functions
Improve the description of the length-age-key

@@
 This is a matter of naming convention that we struggled with.
-The intent was to operate in log space for many of these equations and, as such, the simulations were set-up to receive log means.
+The intent was to operate in log space for these equations and, as such, the simulations were set-up to receive log means.
 When developing the functions, however, we waffled on whether to use `log_mean` or `mean` as the argument name and we landed on `mean` as we figured this would be a more relatable value for users to supply.
@@
 
-*TODO*
+*Another good catch as we did not provide enough information to replicate the approach.
+We have added more detail on the calculations behind the sim_vonB function; turns out our approach was similar to the one described in the exemplar the reviewer provided.*
 

### 2019-12-16 analysis/paper/sim_survey_paper.Rmd
Improve the description of the length-age-key

@@
 
-$$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$
-
-Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$.
-Abundance in discrete length groups, $l_{\mathrm{group}}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance at age to abundance at length.
+$$log(L) = log(L_{\infty} - (L_{\infty} - L_0) e^{-Ka}) + \varepsilon$$
+
+Where $L_{\infty}$ is the mean asymptotic length, $L_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{L}^2)$.
+Numbers at age are distributed across discrete length groups following a lognormal distribution by calculating the probability of being in a specific length group given age.
+These probabilities are calculated using the standard normal cumulative density function $\Phi$ for a sequence of length groups $l$ from length 0 to 10 times the maximum predicted length $L$ at an interval of $l_{group}^{N}$:
+
+$$ \phi_{a,l} = \Phi \left ( \frac{log(L_l)}{\sigma_{L}} \right ) - \Phi \left ( \frac{log(L_{l-1})}{\sigma_{L}} \right )$$
+
 Though some typical relationships have yet to be implemented (e.g.
@@
 | `+            phi_year = 0.5),`             | Correlation across years in error around total mortality   | $\varphi_{\delta, \mathrm{year}}$   |
-| `+  growth = sim_vonB(Linf = 120,`          | Mean asymptotic length (cm)                                | $l_{\infty}$                        |
-| `+                    L0 = 5,`              | Length in birth year (cm)                                  | $l_0$                               |
+| `+  growth = sim_vonB(Linf = 120,`          | Mean asymptotic length (cm)                                | $L_{\infty}$                        |
+| `+                    L0 = 5,`              | Length in birth year (cm)                                  | $L_0$                               |
 | `+                    K = 0.1,`             | Growth rate                                                | $K$                                 |
-| `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_l$                          |
-| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{\mathrm{group}}^{N}$            |
+| `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_L$                          |
+| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{group}^{N}$                     |
 
@@
 Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across longitude only, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
-The grid is divided into $L_{\mathrm{div}}$ divisions (e.g.
-NAFO or ICES divisions) and $L_{\mathrm{strat}}$ depth-based strata.
+The grid is divided into $H_{\mathrm{div}}$ divisions (e.g.
+NAFO or ICES divisions) and $H_{\mathrm{strat}}$ depth-based strata.
 The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space via:
@@
 | `+                   depth_range = c(0, 1000),`         | Depth range from coast to slope (m)                          | $[d_{\mathrm{min}}, d_{\mathrm{max}}]$  |
-| `+                   n_div = 1,`                        | Number of divisions                                          | $L_{\mathrm{div}}$                      |
+| `+                   n_div = 1,`                        | Number of divisions                                          | $H_{\mathrm{div}}$                      |
 | `+                   strat_splits = 2,`                 | Number of times to horizontally split strat                  | &nbsp;                                  |
@@
 
-This work has benifited from valuable feedback from numerous colleagues, including Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury and Peter Upward.
+This work has benefited from valuable feedback from numerous colleagues, including Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury and Peter Upward.
 We also thank Dave Cote and Joanne Morgan for providing constructive comments on a previous version of this manuscript.
@@
 |:---------- |:----- |
-| $L$ | Number of strata ($h = 1, 2, ....., L$) |
+| $H$ | Number of strata ($h = 1, 2, ....., H$) |
 | $A_h$ | Area of the $h^{th}$ stratum |
@@
 | $n_h$ | Total number of units sampled in the $h^{th}$ stratum ($i = 1, 2, ....., n_h$) |
-| $N = \sum^{L}_{h = 1} N_h$ | Total number of sample units in the survey |
-| $n = \sum^{L}_{h = 1} n_h$ | Total number of observations in the survey |
+| $N = \sum^{H}_{h = 1} N_h$ | Total number of sample units in the survey |
+| $n = \sum^{H}_{h = 1} n_h$ | Total number of observations in the survey |
 | $W_h = \frac{N_h}{N}$ | Stratum weight |
@@
 | $s^2_h = \sum^{n_h}_{i = 1} \frac{(I_{h, i} - \overline{I}_h)}{(n_h - 1)}$ | Sample variance in the $h^{th}$ stratum |
-| $\overline{I} = \sum^{L}_{h = 1} W_h \overline{I}_h$ | Estimate of the population mean per unit (i.e. stratified mean catch per tow) |
-| $s^2(\overline{I}) = \frac{1}{N^2} \sum^{L}_{h=1} N_h (N_h - n_h) \frac{s^2_h}{n_h}$ | Estimate of the variance of the stratified mean |
+| $\overline{I} = \sum^{H}_{h = 1} W_h \overline{I}_h$ | Estimate of the population mean per unit (i.e. stratified mean catch per tow) |
+| $s^2(\overline{I}) = \frac{1}{N^2} \sum^{H}_{h=1} N_h (N_h - n_h) \frac{s^2_h}{n_h}$ | Estimate of the variance of the stratified mean |
 | $\widehat{I} = N \overline{I}$ | Estimate of the population total over the survey area |

### 2019-12-16 R/sim_pop.R

@@ -185,8 +185,8 @@ sim_vonB <- function(Linf = 120, L0 = 5, K = 0.1, log_sd = 0.1,
     pred_length <- Linf - (Linf - L0) * exp(-K * age)
 
     if (length_age_key) {
-      breaks <- seq(0, max(pred_length) * 10, length_group)
-      breaks[1] <- 0.0000001
+
+      breaks <- seq(0, ceiling(max(pred_length)) * 10, length_group)
       lak <- matrix(NA, ncol = length(pred_length), nrow = length(breaks) - 1,
                     dimnames = list(length = group_lengths(breaks, length_group)[-length(breaks)],
                                     age = age))

### 2019-12-17 analysis/paper/rebuttal_letter.Rmd
Coarse re-structure of the paper based on a reviewers' suggestion

@@
 Lewis.
-In accordance to one of the reviewers' suggestions, the paper has been renamed **"`SimSurvey`: an `R` package to compare the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"**.
+In accordance to one of the reviewers' suggestions, the paper has been renamed **"`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"**.
 
@@
 
-*TODO*
+*We have prefaced this section with the caveat that population and spatial dynamics are modeled as independent processes and we note the pros and cons of this approach.
+In short: the con is that the approach is a simplification of reality as it does not explicitly account for dynamics such as larval dispersal, spatial differences in growth and meta-population connectivity; the pro is that this simplification limits the number of unknown parameters that need to be specified while facilitating the simulation of a sufficiently complex population for testing the efficacy of various survey designs.*
 
@@
 
-*See next comment*
+*Fair point, however, we focused on "depth" for two reasons: 1) it is an important covariate for many species (as the reviewer noted), and 2) most surveys that we are aware of are depth-stratified (also noted by the reviewer below).
+We therefore retained the "depth" naming as it ought to satisfy most use cases.
+Users can satisfy other, less common, cases by supplying user-specified grids and/or spatial populations.*
 
@@
 
-*In response to this suggestion and the previous one, we have modified the text (lines X-X)*
+*TODO*
 

### 2019-12-17 analysis/paper/sim_survey_paper.Rmd
Coarse re-structure of the paper based on a reviewers' suggestion

@@
 ---
-title: "`SimSurvey`: an `R` package to compare the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
+title: "`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
 output: 
@@
 The steps taken to simulate surveys of spatial, age-structure populations are outlined below.
+First we outline the equations underlying the package in the [*Model structure*][Model structure] section and then move into a vignette of the package under the [_Using **`SimSurvey`**_][Using **`SimSurvey`**] section.
 While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study.
@@
 
-# The **`SimSurvey`** package
+# Model structure
+
+In this section, we describe the framework currently implemented in the **`SimSurvey`** package.
+With this framework, we tried to strike a balance between realism, simplicity, generality and computational feasibility.
+The framework follows four general steps: 1) simulate a spatially-aggregated age-structured population; 2) distribute the population throughout a spatial grid, imposing correlation across space, time and age; 3) sample the population using random sampling; and, 4) obtain population estimates using a design-based analysis.
+Though there is a degree of flexibility in each of these steps, keep in mind that users can circumvent specific components by applying user defined equations, inputs and/or analyses.
+
+## Simulate abundance
+
+The simulation starts with an exponential decay cohort model where the abundance at age $a$ in year $y$ ($N_{a,y}$) equals the abundance of that cohort in the previous year multiplied by the associated survival rate, which is expressed in terms of total mortality ($Z_{a,y}$):
+
+$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
+
+Here, numbers at age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e.
+recruits) vary around a baseline value, $log(N_{1,y}) = log(\mu_r) + \epsilon_{y}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(\mu_Z) + \delta_{a,y}$.
+The error around the recruitment process was set to follow a random walk, $\epsilon_{y} \sim N(\epsilon_{y - 1}, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$.
+The covariance across ages and years is controlled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, \mathrm{age}}$ and $\varphi_{\delta, \mathrm{year}}$, respectively).
+This structure allows for autocorrelation in process errors across ages and years (i.e.
+total mortality can be made to be more similar for fish that are closer together in age and/or time).
+Abundance at age is then converted to abundance at length using the original von Bertalanffy growth curve [@von1938]:
+
+$$log(L) = log(L_{\infty} - (L_{\infty} - L_0) e^{-Ka}) + \varepsilon$$
+
+Where $L_{\infty}$ is the mean asymptotic length, $L_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{L}^2)$.
+Numbers at age are distributed across discrete length groups following a lognormal distribution by calculating the probability of being in a specific length group given age, $\phi_{a,l}$.
+These probabilities are calculated using the standard normal cumulative density function $\Phi$ for a sequence of length groups $l$ from length 0 to 10 times the maximum predicted length $L$ at an interval of $l_{group}^{N}$:
+
+$$ \phi_{a,l} = \Phi \left ( \frac{log(L_l)}{\sigma_{L}} \right ) - \Phi \left ( \frac{log(L_{l-1})}{\sigma_{L}} \right )$$
+
+Though some typical relationships have yet to be implemented (e.g.
+stock-recruitment), this formulation facilitates the simulation of age-structured populations that are dynamic enough to test an array of survey designs.
+
+## Simulate spatial distribution
+
+Rather than developing a full spatially-explicit model, population and spatial dynamics are modeled as independent processes for simplicity.
+The complexities of spatial population dynamics - such as larval dispersal, spatial differences in growth and meta-population connectivity - are therefor not accounted for and, as such, the model is a substantial simplification of reality.
+Despite this limitation, the approach taken facilitates the simulation of spatial, age-structured populations with sufficient complexity for testing the efficacy of various survey designs.
+The simplicity also limits the number of unknown parameters that need to be specified to simulate a population.
+Parameter estimates from  spatially-aggregated age-structured models, which are commonly used in stock assessments, can therefore be used to simulate a population using the cohort model and the resultant abundance at age values can be distributed across a spatial grid.
+Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across longitude only, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
+The grid is divided into $H_{\mathrm{div}}$ divisions (e.g.
+NAFO or ICES divisions) and $H_{\mathrm{strat}}$ depth-based strata.
+The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space via:
+
+$$\begin{aligned}log(N_{a,y,s}) &= log(N_{a,y}) + \eta_{a,y,s} \\ \eta_{a,y,s} &= \frac{(d_s - \mu_{d}) ^ 2}{2 \sigma_d^2} + \xi_{a,y,s} \end{aligned}$$
+
+Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean depth.
+Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and a two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
+Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
+The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see [*S2 Appendix*][S2 Appendix: Age-year-space covariance] for a more detailed description of the space-age-year covariance structure).
+In short, this formulation allows control of depth preferences, the level of spatial aggregation and the degree of age and year specific clustering.
+Note that $\eta_{a,y,s}$ values are rounded and re-scaled to ensure that discrete numbers of fish reside in each grid cell and to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model.
+
+## Simulate survey
+
+The final step in the simulation is to sample the simulated population over the age-year-space array generated.
+The sampling is stratified random, emulating real-world surveys conducted by many research institutions around the world.
+The area of each strata is calculated and this is used to define the number of sampling stations, hereafter referred to as sets, allocated to each strata under a particular set density, $D_{\mathrm{sets}}$.
+The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:
+
+$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{\mathrm{trawl}}}{A_{\mathrm{cell}}} q_a )$$
+
+Where $n_{a,y,s}$ is the number of fish of age $a$ in year $y$ sampled by a set at location $s$, $A_{\mathrm{trawl}}$ indicates the area covered by the trawl, $A_{\mathrm{cell}}$ is the area of a grid cell, and $q_a$ is the catchability coefficient of each age (i.e.
+the ability of the trawling gear to catch specific age groups).
+Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
+The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the *[Simulate abundance]* section.
+Depending on the number of fish caught, sub-sampling is then conducted.
+Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division.
+Such sub-sampling is common in fisheries-independent surveys as it is costly, impractical and unnecessary to sample every fish captured.
+Age determination is especially time-consuming, which is why otoliths for age-determination tend to be sub-sampled by length-bin to obtain a representative age sample across a wider range of lengths than would be obtained via random sampling.
+
+## Stratified analysis
+
+While there are many model-based options for obtaining an abundance index from survey data [e.g.
+@Thorson2015], design-based approaches, such as stratified analyses, are often used.
+Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in [*S3 Appendix*][S3 Appendix: Stratified analysis equations]] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
+Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be scaled up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key.
+We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:
+
+$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})^2}{N_a N_y N_k}}$$
+
+Where $N_a$, $N_y$, and $N_k$ are the number of ages, years and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e.
+catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$).
+RMSE was also calculated for abundance at length estimates, where the above formula is indexed by length groups $l$, and total abundance, which lacks a group index of $a$ or $l$.
+
+## Assumptions
+
+Like any model, this simulation is a simplification of a much more complex reality.
+For instance, the population is assumed to aggregate by age-class and be uniformly distributed within a cell, instead fish may aggregate by length and form finer-scale clusters.
+The survey is also an instantaneous snapshot of the population, meaning that the population is assumed to be in the same location from the beginning to the end of the survey.
+Also, fish are aged at random within length bins and ages are estimated without error.
+Finally, area trawled is assumed to be perfectly standard.
+These assumptions, plus a range of others, will surely under-represent the natural variability of fish populations and survey protocol.
+Nevertheless, the **`SimSurvey`** package provides a relatively complex and flexible operating model for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
+
+# Using **`SimSurvey`**
 
@@
 
-## Simulate abundance
-
-The simulation starts with an exponential decay  cohort model where the abundance at age $a$ in year $y$ ($N_{a,y}$) equals the abundance of that cohort in the previous year multiplied by the associated survival rate, which is expressed in terms of total mortality ($Z_{a,y}$):
-
-$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
-Here, numbers at age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e.
-recruits) vary around a baseline value, $log(N_{1,y}) = log(\mu_r) + \epsilon_{y}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(\mu_Z) + \delta_{a,y}$.
-The error around the recruitment process was set to follow a random walk, $\epsilon_{y} \sim N(\epsilon_{y - 1}, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$.
-The covariance across ages and years is controlled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, \mathrm{age}}$ and $\varphi_{\delta, \mathrm{year}}$, respectively).
-This structure allows for autocorrelation in process errors across ages and years (i.e.
-total mortality can be made to be more similar for fish that are closer together in age and/or time).
-Abundance at age is then converted to abundance at length using the original von Bertalanffy growth curve [@von1938]:
-
-$$log(L) = log(L_{\infty} - (L_{\infty} - L_0) e^{-Ka}) + \varepsilon$$
-
-Where $L_{\infty}$ is the mean asymptotic length, $L_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{L}^2)$.
-Numbers at age are distributed across discrete length groups following a lognormal distribution by calculating the probability of being in a specific length group given age.
-These probabilities are calculated using the standard normal cumulative density function $\Phi$ for a sequence of length groups $l$ from length 0 to 10 times the maximum predicted length $L$ at an interval of $l_{group}^{N}$:
-
-$$ \phi_{a,l} = \Phi \left ( \frac{log(L_l)}{\sigma_{L}} \right ) - \Phi \left ( \frac{log(L_{l-1})}{\sigma_{L}} \right )$$
-
-Though some typical relationships have yet to be implemented (e.g.
-stock-recruitment), this formulation facilitates the simulation of age-structured populations that are dynamic enough to test an array of survey designs.
+## `sim_abundance`
 
@@
 
-## Simulate spatial distribution
-
-The next step in the simulation is to distribute the abundance at age matrix simulated using the cohort model throughout a spatial field.
-Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across longitude only, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
-The grid is divided into $H_{\mathrm{div}}$ divisions (e.g.
-NAFO or ICES divisions) and $H_{\mathrm{strat}}$ depth-based strata.
-The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space via:
-
-$$\begin{aligned}log(N_{a,y,s}) &= log(N_{a,y}) + \eta_{a,y,s} \\ \eta_{a,y,s} &= \frac{(d_s - \mu_{d}) ^ 2}{2 \sigma_d^2} + \xi_{a,y,s} \end{aligned}$$
-
-Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean depth.
-Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and a two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
-Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
-The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see [*S2 Appendix*][S2 Appendix: Age-year-space covariance] for a more detailed description of the space-age-year covariance structure).
-In short, this formulation allows control of depth preferences, the level of spatial aggregation and the degree of age and year specific clustering.
-Note that $\eta_{a,y,s}$ values are rounded and re-scaled to ensure that discrete numbers of fish reside in each grid cell and to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model.
+## `sim_distribution`
 
@@
 
-## Simulate survey
-
-The final step in the simulation is to sample the simulated population over the age-year-space array generated.
-The sampling is stratified random, emulating real-world surveys conducted by many research institutions around the world.
-The area of each strata is calculated and this is used to define the number of sampling stations, hereafter referred to as sets, allocated to each strata under a particular set density, $D_{\mathrm{sets}}$.
-The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:
-
-$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{\mathrm{trawl}}}{A_{\mathrm{cell}}} q_a )$$
-
-Where $n_{a,y,s}$ is the number of fish of age $a$ in year $y$ sampled by a set at location $s$, $A_{\mathrm{trawl}}$ indicates the area covered by the trawl, $A_{\mathrm{cell}}$ is the area of a grid cell, and $q_a$ is the catchability coefficient of each age (i.e.
-the ability of the trawling gear to catch specific age groups).
-Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
-The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the *[Simulate abundance]* section.
-Depending on the number of fish caught, sub-sampling is then conducted.
-Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division.
-Such sub-sampling is common in fisheries-independent surveys as it is costly, impractical and unnecessary to sample every fish captured.
-Age determination is especially time-consuming, which is why otoliths for age-determination tend to be sub-sampled by length-bin to obtain a representative age sample across a wider range of lengths than would be obtained via random sampling.
+## `sim_survey`
 
@@
 
-## Stratified analysis
-
-While there are many model-based options for obtaining an abundance index from survey data [e.g.
-@Thorson2015], design-based approaches, such as stratified analyses, are often used.
-Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in [*S3 Appendix*][S3 Appendix: Stratified analysis equations]] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
-Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be scaled up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key.
-We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:
-
-$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})^2}{N_a N_y N_k}}$$
-
-Where $N_a$, $N_y$, and $N_k$ are the number of ages, years and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e.
-catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$).
-RMSE was also calculated for abundance at length estimates, where the above formula is indexed by length groups $l$, and total abundance, which lacks a group index of $a$ or $l$.
+## `run_strat`
 
@@
 
-## Testing survey protocol
+## `test_surveys`
 
@@
 
-## Assumptions
-
-Like any model, this simulation is a simplification of a much more complex reality.
-For instance, the population is assumed to aggregate by age-class and be uniformly distributed within a cell, instead fish may aggregate by length and form finer-scale clusters.
-The survey is also an instantaneous snapshot of the population, meaning that the population is assumed to be in the same location from the beginning to the end of the survey.
-Also, fish are aged at random within length bins and ages are estimated without error.
-Finally, area trawled is assumed to be perfectly standard.
-These assumptions, plus a range of others, will surely under-represent the natural variability of fish populations and survey protocol.
-Nevertheless, the **`SimSurvey`** package provides a relatively complex and flexible operating model for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
-
 # Research opportunities
@@
 We also thank Dave Cote and Joanne Morgan for providing constructive comments on a previous version of this manuscript.
+Finally, the package and the manuscript were greatly improved by feedback from two anonymous reviewers.
 This work was supported by the NSERC visiting-fellow program and Fisheries and Oceans Canada.

### 2019-12-17 analysis/paper/rebuttal_letter.Rmd
Complete draft of replies to Reviewer #1

@@
 
+----------------------------------------------------------------------------------------------------
+
 Reviewer #1:
@@
 
-*Please see our specific responses to lines X, X, and X.*
-
 2. Add additional explanation to justify the choices. E.g. why population dynamics is modeled independently of the spatial distribution. Is this reasonable, limitation, etc.
 
-*TODO*
+3. Few suggestions (add-ons, renaming) to increase the generality of the paper (without requiring too much work)
 
-3. Few suggestions (add-ons, renaming) to increase the generality of the paper (without requiring too much work)
-Please see our specific responses to...
-
-*TODO*
+*We hope we have addressed each of these main comments by 1) clarifying several equations, 2) adding more details and justifications, and 3) modifying naming conventions and clarifying extensibility.
+See replies below for more details.*
 
@@
 
-*TODO*
+*Given a suggestion from Reviewer #2, we have restructured the paper to have two core sections: "Model structure" and "Using SimSurvey".
+We have included a blanket statement under the "Model structure" section noting that users can circumvent specific components of the framework.*
 
@@
 
-*TODO*
+*Good idea!
+We actually pursued this idea in an earlier iteration of `make_grid`, however, we abandoned the option because it was difficult to automate; the random field created random problems with the depth-stratification (e.g.
+one cell strata).*
 
@@
 
-*TODO*
+*We have added more detail on the structure of the divisions and strata.*
 
@@
 
-*TODO*
+*See reply to point 2. above.*
 
@@
 
-*TODO*
+*Right, the equation presented was not an accurate reflection of the calculations. We have revised the equation to explicitly show how the values were normalized to sum to 1.*
 
@@
 
-*TODO*
+*True.
+Some users may find this parameterization insufficient for their species and we hope they will implement their own closure to use in the `sim_distribution` function to better simulate the effect.
+In addition to our blanket statement under the "Model structure" section, we have added a more specific statement under the "Using SimSurvey" section stating that alternate formulations can be used by supplying alternate closures to the core functions.*
 
@@
 
-*TODO*
+*We have prefaced that sentence with "The rate at which point-to-point spatial correlation decays with distance is controlled by...".*
 
@@
 
-*TODO*
+*See reply to point 7. above*
 
@@
 
-*TODO*
+*"Variance" was a poor word choice. We have replaced it with "noise" as it is the simulated noise that we fix across multiple age groups.*
 
@@
 
-*TODO*
+*We have added it to the table.*
 
@@
 
-*This is an excellent suggestion and we have inserted headers where appropriate.*
+*This is an excellent suggestion. By following a suggestion by Reviewer #2 to re-organize the paper into two core sections we have added more headers to help guide the readers.*
 
@@
 
-*TODO*
+*We have made the suggested change.*
 
@@
 
-*TODO*
+*We have clarified how number of sets per strata is calculated.*
 
@@
 
-*TODO*
+*We are not sure what the reviewer would like to have implemented here. Is the suggestion to impose a minimum number of ages to collect across all length groups?*
 
@@
 
-*TODO*
+*The sampling is implemented such that the number of fish sampled in a cell cannot exceed the number of fish in a cell because the population is split across sets in cases where more than one set is conducted in a cell.
+We have added this missing detail to our manuscript.
+We also added a note that the survey is assumed to have no impact on the population from one year to the next.*
 
@@
 
-*TODO*
+*Honestly, we do not recall what we mean by "depending on the number of fish caught".
+Perhaps we added those words to cover off cases where no fish are caught.
+Whatever the case, we have removed the statement to minimize confusion.
+We have also clarified the sub-sampling sequence.
+Finally, we have yet to consider including correlation structure in the sampling process as we went about imposing correlation via the spatial correlation of age groups (i.e.
+age-specific clustering tends to result in sets with high intraclass correlation).
+We are open to learning more about other processes that may contribute to correlated samples.*
 
@@
 
-*TODO*
+*We have clarified the meaning and utility of the `min_sets` argument (i.e.
+a small strata may be allocated only one set under a low set density scenario; this argument overrides the allocation and imposes the `min_sets` if it is greater than the allocation).*
 
@@
 
-*We have changed the page numbers accordingly.*
+*We have changed the numbers accordingly.*
 
@@
 
-*TODO*
+*We have included an example that ought to clarify how a custom closure can be supplied.*
 
@@
 
-*We have adopted the Reviewers suggestion and referred the reader to the section on Stratified analysis.*
+*We have clarified how this was calculated.*
 
@@
 
-*TODO*
+*Good point, however, we think this is covered by referencing a paper that describes a geostatistical R package and we also note that other options can be used under the "Research opportunities" section.*
 
@@
 
-*TODO*
+*We have included the figure*
+
+----------------------------------------------------------------------------------------------------
 
@@
 
-*We agree with the Reviewer and retitled the ms "`SimSurvey`: an `R` package to compare the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks".*
+*We agree with the Reviewer and retitled the ms "`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks".*
 

### 2019-12-17 analysis/paper/sim_survey_paper.Rmd
Complete draft of replies to Reviewer #1

@@
 The steps taken to simulate surveys of spatial, age-structure populations are outlined below.
-First we outline the equations underlying the package in the [*Model structure*][Model structure] section and then move into a vignette of the package under the [_Using **`SimSurvey`**_][Using **`SimSurvey`**] section.
+First we outline the equations underlying the package in the *[Model structure]* section and then move into a vignette of the package under the *[Using **`SimSurvey`**]* section.
 While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study.
@@
 Though there is a degree of flexibility in each of these steps, keep in mind that users can circumvent specific components by applying user defined equations, inputs and/or analyses.
+Details on how to use the package and circumvent some aspects of its structure are outlined in the *[Using **`SimSurvey`**]* section.
 
@@
 Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across longitude only, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
-The grid is divided into $H_{\mathrm{div}}$ divisions (e.g.
-NAFO or ICES divisions) and $H_{\mathrm{strat}}$ depth-based strata.
-The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space via:
-
-$$\begin{aligned}log(N_{a,y,s}) &= log(N_{a,y}) + \eta_{a,y,s} \\ \eta_{a,y,s} &= \frac{(d_s - \mu_{d}) ^ 2}{2 \sigma_d^2} + \xi_{a,y,s} \end{aligned}$$
+The grid can be divided into two hierarchical levels.
+For demonstration purposes, we envision these levels as part of a stratified-random survey within international fishery divisions, i.e., $H_{\mathrm{strat}}$ depth-based strata within $H_{\mathrm{div}}$ divisions (e.g.
+NAFO or ICES divisions).
+The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space.
+This noise term is normalized to sum to 1 to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model:
+
+$$\begin{aligned}N_{a,y,s} &= N_{a,y} \frac{\eta_{a,y,s}}{\sum_{s = 1}^{N_s} \eta_{a,y}} \\ \eta_{a,y,s} &= \frac{(d_s - \mu_{d}) ^ 2}{2 \sigma_d^2} + \xi_{a,y,s} \end{aligned}$$
 
@@
 Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and a two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
-Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
+The rate at which point-to-point spatial correlation decays with distance is controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
 The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see [*S2 Appendix*][S2 Appendix: Age-year-space covariance] for a more detailed description of the space-age-year covariance structure).
 In short, this formulation allows control of depth preferences, the level of spatial aggregation and the degree of age and year specific clustering.
-Note that $\eta_{a,y,s}$ values are rounded and re-scaled to ensure that discrete numbers of fish reside in each grid cell and to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model.
 
@@
 The final step in the simulation is to sample the simulated population over the age-year-space array generated.
-The sampling is stratified random, emulating real-world surveys conducted by many research institutions around the world.
-The area of each strata is calculated and this is used to define the number of sampling stations, hereafter referred to as sets, allocated to each strata under a particular set density, $D_{\mathrm{sets}}$.
+The sampling is random or stratified random, emulating real-world surveys conducted by many research institutions around the world.
+The area of each strata $A_{\mathrm{strat}}$ is calculated and this is used to define the number of sampling stations $H_{\mathrm{sets}}$, hereafter referred to as sets, allocated to one or more strata under a particular set density, $D_{\mathrm{sets}}$ (i.e.
+$H_{\mathrm{sets}} = A_{\mathrm{strat}}~D_{\mathrm{sets}}$).
 The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:
@@
 Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
-The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the *[Simulate abundance]* section.
-Depending on the number of fish caught, sub-sampling is then conducted.
+In cases where there are multiple sets in one cell, the population in that cell is split across the sets.
+While this means that numbers caught in an isolated simulation cannot exceed the numbers in the population, keep in mind that the survey, no matter how intense, is assumed to have no impact on the population from one year to the next.
+
+Once catches are simulated, lengths of the fish sampled by each set are simulated using the von Bertalanffy growth equation found above in the *[Simulate abundance]* section.
+Sub-sampling is then conducted whereby a subset of fish are sampled for length measurements and a subset of this subset are sampled for age determination.
 Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division.
@@
 The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtual fish populations with correlation across space, time and age (`sim_abundance`, `sim_distribution`), 2) simulating surveys with a range of sampling strategies and intensities (`sim_survey`), and 3) estimating the stratified mean and variance of simulated survey data (`test_surveys`; Table \@ref(tab:fun)).
-The equations behind these functions are detailed in the sections below.
 **`SimSurvey`** relies heavily on functions from the **`data.table`** [@dowle2017], **`raster`** [@hijmans2016] and **`plotly`** [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively.
@@
 <br>
+
+The equations behind the functions listed in Table \@ref(tab:fun) are detailed in the *[Model structure]* section.
+Note that several of the core equations are implemented using closures, which are functions that contain data and return functions [@wickham2014].
+This was done to avoid the repeated specifications of key arguments, such as ages and years, and also provide an option for advanced R users to supply custom closures with alternate equations.
 
@@
 (Note that each of the above-mentioned closures include a plot argument such that quick visuals can be obtained using a line of code like this: `sim_R(log_mean = log(500), log_sd = 0.5, plot = TRUE)(years = 1:100)`).
-This structure was chosen to avoid the repeated specifications of ages and years.
-The use of closures also allows users to construct and use their own closures with a similar structure but different underlying formula.
 Overall, the function provides a simple tool for simulating a range of dynamic age-structured populations.
@@
 | `+                            phi_year = 0.9,`          | Correlation across years in spatial distribution             | $\varphi_{\xi, \mathrm{year}}$          |
-| `+                            group_ages = 5:20),`      | Make space-age-year variance equal across these ages         | &nbsp;                                  |
+| `+                            group_ages = 5:20,`       | Make space-age-year noise equal across these ages^1^         | &nbsp;                                  |
+| `+                            group_years = NULL),`     | Make space-age-year noise equal across these years^1^        | &nbsp;                                  |
 | `+  depth_par = sim_parabola(mu = 200,`                 | Depth at which abundance is typically highest (m)            | $\mu_d$                                 |
@@
 
+^1^ All ages or years are independent if argument values is `NULL`.
+
 <br>
@@
 The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`.
-If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure is used.
+If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure structure and correct projection is used.
 The package includes a manually constructed survey grid of NAFO Subdivision 3Ps off the southern coast of Newfoundland (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid.
@@
 ```{r grid, out.height = "320px", fig.cap = "Plots produced by `plot_grid` when supplied a raster object produced by `make_grid` a) using default settings, b) settings that increase the number of times depth strata are horizontally split, and c) settings that produce a more linear depth gradient and increase the number of divisions.
-In these plots, the color gradient represents depth, the thick grey lines delineate divisions and thin white lines delineate strata."}
+In these plots, the yellow to purple color gradient represents depth, the thick grey lines delineate divisions and thin white lines delineate strata."}
 
@@
 In addition to supplying objects produced by `sim_abundance` and `make_grid`, the `sim_distribution` function requires two closures that describe the age-year-space covariance and the relationship with depth.
-Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by tweaking a few parameters in these closures.
+Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by tweaking a few parameters in these closures (Figure \@ref(fig:distribution)).
 Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e.
@@
 
-This function retains all the data simulated by `sim_abundance` and adds a data.table [@dowle2017], named `sp_N`, with abundance (`N`) split by `age`, `year` and `cell`.
+The `sim_distribution` function retains all the data simulated by `sim_abundance` and adds a data.table [@dowle2017], named `sp_N`, with abundance (`N`) split by `age`, `year` and `cell`.
 The function also retains the grid object and converts these data into a data.table, named `grid_xy`, with headers `x`, `y`, `depth`, `cell`, `division` and `strat`.
@@
 | `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / division              | $M_{\mathrm{ages}}$     |
-| `+  age_sammpling = "stratified")`  | Controls whether age sampling is length `"stratified"` or `"random"`    | &nbsp;                  |
+| `+  age_sampling = "stratified")`   | Controls whether age sampling is length `"stratified"` or `"random"`    | &nbsp;                  |
 
@@
 Specifically, the `q` argument requires a closure, such as `sim_logistic`, for defining the probability of catching specific age groups, trawl dimensions are defined in the `trawl_dim` argument, and set, length and age sampling effort are defined using the `set_den`, `lengths_cap` and `ages_cap` arguments, respectively.
-Like `sim_abundance` and `sim_distribution`, custom closures can be supplied to `sim_survey` to impose alternate parametric curves for catchability at age.
+Also note that the `min_sets` argument imposes a minimum of number of sets to conduct per strata, regardless of its allocation given strat area and set density.
+This argument imposes a useful constraint for generating data to be analyzed using design-based approaches that require more than one value for the calculation of a mean.
+
+Like `sim_abundance` and `sim_distribution`, custom closures can be supplied to `sim_survey` to impose alternate parametric curves for catchability at age (i.e.
+a closure including an equation for a dome can be constructed and used in lieu of `sim_logistic` to impose a dome-shaped catchability).
 Multiple simulations of the same survey can be run using the `n_sims` argument, however, requesting large numbers of simulations can be computationally demanding depending on the processing capacity available.
@@
 Again, this function retains all the objects listed in the output of `sim_distribution` and adds data.tables that detail the set locations (`setdet`) and sampling details (`samp`).
-Catchability corrected abundance matrices, named `I` and `I_at_length`, are also produced and added to the output; these matrices are useful for comparing the true abundance available to the survey to abundance estimates obtained using design-based or model-based analyses of the simulated survey data.
+Catchability corrected abundance matrices (abundance at age matrix multiplied by survey catchability), named `I` and `I_at_length`, are also produced and added to the output; these matrices are useful for comparing the true abundance available to the survey to abundance estimates obtained using design-based or model-based analyses of the simulated survey data.
 Specific surveys can be explored using the `plot_survey` function, which uses plotly [@sievert2018] and crosstalk [@cheng2016] in the background to link the bubble plot of aggregate set catch to the histogram of lengths and ages sampled to facilitate explorations of set-specific catches (Figure \@ref(fig:survey)).
@@
 ```{r fan1, out.height = "250px", fig.cap = "Fan chart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{\\mathrm{sets}}$.
-The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+The thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%.
 This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
@@
 
-```{r fan2, out.height = "600px", fig.cap = "Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{\\mathrm{sets}}$, and length sampling protocol, $M_{\\mathrm{lengths}}$.
-The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+```{r fan2, out.height = "600px", fig.cap = "Fan chart of stratified estimates of abundance at length from year seven of the simulation from surveys with different set densities, $D_{\\mathrm{sets}}$, and the maximum number of length samples, $M_{\\mathrm{lengths}}$.
+The thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%.
 This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
@@
 
-```{r fan3, out.height = "600px", fig.cap = "Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{\\mathrm{sets}}$, and length sampling protocol, $M_{\\mathrm{lengths}}$.
+```{r fan3, out.height = "600px", fig.cap = "Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{\\mathrm{sets}}$, and the maximum number of length samples, $M_{\\mathrm{lengths}}$.
 Number of ages sampled per length group, $M_{\\mathrm{ages}}$, was 10 in all scenarios.
-The thick black line indicates the true trend in the total population available to the survey and the color gradient represents a range of probability envelopes from 10% to 90%.
+The thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%.
 This plot is a facet of plots produced by `plot_total_strat_fan` when supplied results from `test_surveys`."}
@@
 Records are ranked by lowest to highest RMSE score.
-Within each plot, a color ramp is applied from lowest to highest value as an additional visual aid.
+Within each plot, a yellow to purple color gradient is applied from lowest to highest value as an additional visual aid.
 Note that the exponent format of the axes defaults to SI unit symbols (e.g.

### 2019-12-17 R/sim_dist.R

@@ -37,8 +37,8 @@
 #'                    length as ages
 #' @param phi_year    Defines autocorrelation through years. Can be one value or a vector of the same
 #'                    length as years
-#' @param group_ages  Make space-age-year variance equal across these ages
-#' @param group_years Make space-age-year variance equal across these years
+#' @param group_ages  Make space-age-year noise equal across these ages
+#' @param group_years Make space-age-year noise equal across these years
 #' @param model       String indicating either "exponential" or "matern" as the correlation function
 #'
 #' @export

### 2019-12-18 analysis/paper/rebuttal_letter.Rmd
Address minor comments from Reviewer 2

@@
 
-*Given a suggestion from Reviewer #2, we have restructured the paper to have two core sections: "Model structure" and "Using SimSurvey".
+*Given a suggestion from Reviewer #2, we have restructured the paper to have two core sections: "Model structure" and "Using __`SimSurvey`__".
 We have included a blanket statement under the "Model structure" section noting that users can circumvent specific components of the framework.*
@@
 Some users may find this parameterization insufficient for their species and we hope they will implement their own closure to use in the `sim_distribution` function to better simulate the effect.
-In addition to our blanket statement under the "Model structure" section, we have added a more specific statement under the "Using SimSurvey" section stating that alternate formulations can be used by supplying alternate closures to the core functions.*
+In addition to our blanket statement under the "Model structure" section, we have added a more specific statement under the "Using __`SimSurvey`__" section stating that alternate formulations can be used by supplying alternate closures to the core functions.*
 
@@
 
-*TODO*
+*This is an excellent suggestion! We have re-structured our manuscript accordingly and feel that this structure will be much easier for a reader to follow.*
 
@@
 
-*TODO*
+*We have now noted in the manuscript that a plus group is not modeled as the number of ages can easily be extended to include groups with zero fish.
+This choice simplifies the simulation, including the setting of initial numbers which is done via exponential decay.
+Further, the lack of a plus group is inconsequential for survey based estimates of abundance at age.*
 
@@
 
-*TODO*
+*We have improved our explanation of a closure at the beginning of the Using __`SimSurvey`__ section*
 
@@
 
-*We have made the suggested change.*
+*This was actually correct and, given an improved description of a closure, the logic behind this line of code should be clearer now.*
 
@@
 
-*TODO*
+*Again, we hope that our improved description of a closure will clarify what we mean by this.*
 
@@
 
-*CHECK*
-
-*We have added the following text: Also note that the key functions in the SimSurvey package have been set-up to be pipe [13] friendly, i.e.
-pipes are an operator to forward the values into a function or call expression.*
-
-*Object b is the same as Object a but the variance in the size of clusters has been changed.
-We believe that this is clear from the text on L228-231 and the code on L235-243 but defer to the editor’s judgement on this matter.*
+*We have clarified how a pipe works, noting that it forwards values from one function call to the next function call, and we now state that the output from the two examples provided (nested approach vs.
+pipe approach) are functionally the same though the approach is slightly different.*
 

### 2019-12-18 analysis/paper/sim_survey_paper.Rmd
Address minor comments from Reviewer 2

@@
 
-cat("1. Adjust font size of code strings  \n")
+cat("1. Adjust font size of code strings (title and headers, and convert links to sections with code to code font [Using SimSurvey and test_surveys])  \n")
 cat("2. Decrease font size of tables  \n")
@@
 The steps taken to simulate surveys of spatial, age-structure populations are outlined below.
-First we outline the equations underlying the package in the *[Model structure]* section and then move into a vignette of the package under the *[Using **`SimSurvey`**]* section.
+First we outline the equations underlying the package in the **[Model structure]** section and then move into a vignette of the package under the **[Using **`SimSurvey`**]** section.
 While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study.
-Output from default function calls are therefore relevant to the case study and these results are described and discussed in [*S1 Appendix*][S1 Appendix: Case study].
+Output from default function calls are therefore relevant to the case study and these results are described and discussed in **[S1 Appendix][S1 Appendix: Case study]**.
 
@@
 Though there is a degree of flexibility in each of these steps, keep in mind that users can circumvent specific components by applying user defined equations, inputs and/or analyses.
-Details on how to use the package and circumvent some aspects of its structure are outlined in the *[Using **`SimSurvey`**]* section.
+Details on how to use the package and circumvent some aspects of its structure are outlined in the **[Using **`SimSurvey`**]** section.
 
@@
 total mortality can be made to be more similar for fish that are closer together in age and/or time).
+Note that a plus group is not modeled as the number of ages can easily be extended to include groups with zero fish.
 Abundance at age is then converted to abundance at length using the original von Bertalanffy growth curve [@von1938]:
@@
 The rate at which point-to-point spatial correlation decays with distance is controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
-The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see [*S2 Appendix*][S2 Appendix: Age-year-space covariance] for a more detailed description of the space-age-year covariance structure).
+The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see **[S2 Appendix][S2 Appendix: Age-year-space covariance]** for a more detailed description of the space-age-year covariance structure).
 In short, this formulation allows control of depth preferences, the level of spatial aggregation and the degree of age and year specific clustering.
@@
 
-Once catches are simulated, lengths of the fish sampled by each set are simulated using the von Bertalanffy growth equation found above in the *[Simulate abundance]* section.
+Once catches are simulated, lengths of the fish sampled by each set are simulated using the von Bertalanffy growth equation found above in the **[Simulate abundance]** section.
 Sub-sampling is then conducted whereby a subset of fish are sampled for length measurements and a subset of this subset are sampled for age determination.
@@
 @Thorson2015], design-based approaches, such as stratified analyses, are often used.
-Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in [*S3 Appendix*][S3 Appendix: Stratified analysis equations]] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
+Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in **[S3 Appendix][S3 Appendix: Stratified analysis equations]**] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
 Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be scaled up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key.
@@
 
-The equations behind the functions listed in Table \@ref(tab:fun) are detailed in the *[Model structure]* section.
-Note that several of the core equations are implemented using closures, which are functions that contain data and return functions [@wickham2014].
-This was done to avoid the repeated specifications of key arguments, such as ages and years, and also provide an option for advanced R users to supply custom closures with alternate equations.
+The equations behind the functions listed in Table \@ref(tab:fun) are detailed in the **[Model structure]** section.
+Note that several of the core equations are implemented using "closures", which are functions that contain data and return functions [@wickham2014].
+For example, `sim_R` returns a function that holds the supplied parameter values and requires a sequence of years to be supplied.
+
+```{r, echo = TRUE, eval = FALSE}
+R_fun <- sim_R(log_mean = log(500), log_sd = 0.5)
+R_vec <- R_fun(years = 1:100)
+```
+
+Here, the `R_vec` object holds 100 years of simulated recruitment values and each run of the `R_fun` function will result in different simulated values using the internal formulation and the parameters supplied to `sim_R`.
+The other closures included in the package operate in a similar way in that parameter inputs are supplied to the closure and the functions returned by the closure requires inputs such as ages and/or years.
+This was done to avoid the repeated specifications of key arguments, such as ages and years.
+Moreover, this approach provides an option for advanced R users inspect and modify the closures implemented in the package to supply custom closures with alternate equations.
+Also note that each of the closures implemented in the package includes a plot argument such that quick visuals can be obtained using a line of code like this: `sim_R(log_mean = log(500), log_sd = 0.5, plot = TRUE)(years = 1:100)`).
 
@@
 
-Abundance at age and length is simulated using the `sim_abundance` function and a default function call is described in Table \@ref(tab:abundance) along with associated symbols from the equations outlined above.
-This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions known as "closures", such as `sim_R`, `sim_Z` and `sim_vonB`, for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`), respectively.
-Closures are functions that contain data and return functions [@wickham2014].
-For example, `sim_R(log_mean = log(500), log_sd = 0.5)` returns a function that holds the supplied parameter values and requires a sequence of years to be supplied.
-(Note that each of the above-mentioned closures include a plot argument such that quick visuals can be obtained using a line of code like this: `sim_R(log_mean = log(500), log_sd = 0.5, plot = TRUE)(years = 1:100)`).
+Abundance at age and length is simulated using the `sim_abundance` function and a default function call is described in Table \@ref(tab:abundance) along with associated symbols from the equations outlined in the **[Simulate abundance]** section.
+This function has a simple structure and requires the specification of a series of `ages` and `years` along with a series of closures, such as `sim_R`, `sim_Z` and `sim_vonB`, for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`), respectively.
 Overall, the function provides a simple tool for simulating a range of dynamic age-structured populations.
@@
 
-The above equations are used in the `make_grid`, `sim_ays_covar` and `sim_parabola` functions, and these functions are used within `sim_distribution` to distribute a population simulated using `sim_abundance` throughout a grid (Table \@ref(tab:distribution)).
+The equations outlined in the *[Simulate spatial distribution]* section are used in the `make_grid`, `sim_ays_covar` and `sim_parabola` functions, and these functions are used within `sim_distribution` to distribute a population simulated using `sim_abundance` throughout a grid (Table \@ref(tab:distribution)).
 The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`.
@@
 Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete.
-Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe [`%>%`; @bache2014] friendly.
+Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe [`%>%`; @bache2014] friendly such that values from one function call are forwarded to the next function (i.e.
+output from the two calls below are functionally the same though the approach is slightly different).
 
@@
 Keep in mind that low numbers of `n_sims` and high numbers of `n_loops` will be less demanding on RAM, but may be slower, especially if the work is spread across few `cores`.
-Because most of the default settings of the functions match the case study settings, the code below will replicate the results from our case study (see [*S1 Appendix*][S1 Appendix: Case study] for more detail).
+Because most of the default settings of the functions match the case study settings, the code below will replicate the results from our case study (see **[S1 Appendix][S1 Appendix: Case study]** for more detail).
 The `expand_surveys` function sets up a series of `r nrow(expand_surveys())` surveys to test (i.e.
@@
 
-The case study described in [*S1 Appendix*][S1 Appendix: Case study] provides one example of how **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
+The case study described in **[S1 Appendix][S1 Appendix: Case study]** provides one example of how **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
 There are multiple layers to the sampling design of such surveys and the end results are linked with the way such data are analyzed.
@@
 One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012].
-Results from the case study (see [*S1 Appendix*][S1 Appendix: Case study]) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
+Results from the case study (see **[S1 Appendix][S1 Appendix: Case study]**) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
 @Berg2012].
@@
 The code below displays the core defaults will replicate the results of this case study.
-Results produced here are the same as those produced and displayed in a series of figures in the *[Testing survey protocol]* section of the paper.
+Results produced here are the same as those produced and displayed in a series of figures in the **[`test_surveys`]** section of the paper.
 

### 2019-12-19 analysis/paper/rebuttal_letter.Rmd
Draft Parameterisation section (suggested by reviewer 2)

@@
 
-*TODO*
+*This is yet another very helpful suggestion and, now that we have included a Parameterisation section, it is easy to see the value of such a section to prospective users of the package.
+However, we should note that we deviated from the recommendation to a degree as we did not produce a table that summaries how to specify the parameters, rather we wrote a section that outlines recommended steps for setting up a custom simulation.
+We started drafting a table, however, it quickly became apparent that it would be too big and cumbersome and somewhat redundant with the core tables that describe the arguments and parameters.
+We hope that the narrative/outline we included will serve the practical purpose the reviewer had in mind.*
 

### 2019-12-19 analysis/paper/sim_survey_paper.Rmd
Draft Parameterisation section (suggested by reviewer 2)

@@
 
-cat("1. Adjust font size of code strings (title and headers, and convert links to sections with code to code font [Using SimSurvey and test_surveys])  \n")
+cat("1. Adjust font size of code strings (title and headers, and convert links to sections with code to code font [Using SimSurvey, test_surveys])  \n")
 cat("2. Decrease font size of tables  \n")
@@
 
+# Parameterisation
+
+The default parameter values used in this package were chosen to emulate  survey data from a specific cod population (see **[S1 Appendix][S1 Appendix: Case study]** for details) and, as such, these values ought to provide a good place to start for another population of cod.
+In this section we outline a series of recommended steps to take to tailor the default settings to suit another system.
+Throughout this outline, we also provide some context on how default parameter values were chosen as this information may help guide a users' choices.
+
+## 1. Define survey area and protocol
+
+Parameterising the survey design should be the easiest place to start as the area and protocol should be clearly defined.
+The survey area can be defined using `make_grid` or it can be manually constructed.
+For the case study, `make_grid` arguments were modified to create a grid with similar dimensions to the stratified design actually applied to the case study area (i.e.
+total area, shelf depth, number of strata, etc.).
+Also note that the resolution (`res` argument) of the grid was defined to be similar to the resolution at which sets are allocated in the survey.
+A square grid was chosen in lieu of the actual survey grid to create an area that avoids potential issues associated with land as a barrier while being sufficiently complex to test survey designs.
+When setting up a grid, be careful not to drastically increase the dimensions (`x_range` or `y_range` arguments) or resolution (`res` argument) as simulating the spatial noise over a large field can be computationally demanding.
+
+With a survey grid defined, the settings of `sim_survey` can be modified to match the protocol of a users' survey of interest.
+For the case study, the trawl dimensions (`trawl_dim`) were modified to match the dimensions of a standard trawl conducted in the region.
+Likewise, area is used to define the number of sets to allocate to a strata and the value used by default for `set_den` matches the protocol, as does the `min_sets` setting (i.e.
+at least two sets are conducted in each strata to facilitate stratified analyses).
+Finally, length and age sub-sampling settings were modified to match the protocol for the survey of cod in the case study area.
+The most difficult parameter to define in the `sim_survey` function is `q`.
+A logistic curve is used by default to define these unknown values and the `sim_logistic` parameters were tweaked to approximate the catchability curve supplied to the survey-based model that was used to assess the case study population [@ings2019].
+A similar tactic can be applied to other cases with estimates of survey catchability.
+
+## 2. Define abundance and growth parameters
+
+Abundance and growth parameters are more difficult to define than the survey protocol as the true values are unknown.
+Nevertheless, spatially-aggregated age-structured models are frequently applied to assess fish populations and estimates from such models can be used to define `sim_abundance` argument settings.
+Alternatively, key values from these models (e.g.
+Z and N matrices) can be used instead of simulated values.
+Default parameters used in `sim_abundance` were largely inspired by estimates from the survey-based model used to assess the stock [@ings2019].
+Based on this work, average total mortality was roughly centered around 0.5 ($\mu_Z$; argument `log_mean` in `sim_Z` closure) and we allowed it to make similar changes in magnitude through time and across ages by setting the standard deviation of log total mortality to 0.2 ($\sigma_Z$; argument `log_sd` in `sim_Z` closure) and the correlations to 0.9 and 0.5 in the age and year dimensions ($\varphi_{\delta, \mathrm{age}}$ and $\varphi_{\delta, \mathrm{year}}$; `phi_age` and `phi_year`), respectively.
+Likewise, `sim_R` values were modified to generate similar patterns in recruitment as those estimated by the model.
+Ages groups to simulate across ranged from age 1 to 20 ($a$; `ages` argument) to capture the tails of the population, and 20 years were simulated ($y$; `years` argument) to capture some dynamics.
+Finally, von Bertalanffy growth parameters were roughly based on estimates from cod in the region [@Cadigan2016b].
+While the choices made are obviously imperfect and ad hoc, it is not necessary to generate a perfect representation of the population as the goal here is to generate a dynamic population to survey.
+What is key is that we work with a known population from which we can evaluate the performance of specific survey designs and analyses.
+
+## 3. Define spatial parameters
+
+Ideally this step would be no more difficult than step 2 as the inputs required could come from a spatially-explicit age-structured model fit to real data.
+Such estimates, however, are rare and we can confirm that no such model has been developed and fit to data from our case study population.
+Unless a user has access to such estimates, parameter estimates will need to manually defined using visual comparisons of real and simulated data.
+A user will have a first-cut simulation of some survey data by following steps 1 and 2, and the plotting functions provided with the package will provide a means of visually exploring these data.
+For comparison purposes, actual survey data will need to be processed and visual patterns in the spatial distribution of each age will need to be inspected.
+It was in this step that we noticed that ages 1 to 4 tend to occupy different areas while ages 5+ occupy similar space for the cod population we focused on.
+This observation motivated the coupling of the spatial noise of ages 5+ (`group_ages` argument in the `sim_ays_covar` closure), which forces this component of the population to occupy the same space, and the moderate value of 0.5 for the correlation in the spatial distribution across ages ($\varphi_{\xi, \mathrm{age}}$; `phi_age` argument in the `sim_ays_covar` closure).
+Year-to-year plots of the real data also revealed considerable inertia in the space occupied by each age group; this observation motivated a fairly high value of 0.9 for the correlation in the spatial distribution across years ($\varphi_{\xi, \mathrm{year}}$; `phi_year` argument in the `sim_ays_covar` closure).
+Finally, the range of the spatial correlation ($r$; `range` argument in the `sim_ays_covar` closure) was iteratively modified until the simulated population exhibited a similar level of patchiness as the real data, and the standard deviation of the age-year-space process ($\sigma_{\xi}$; `sd` argument in the `sim_ays_covar` closure) was increased to generate set catches with tails similar to those observed in the real data (cod catches in the area tend to be zero-inflated and heavy-tailed).
+Finally, visual inspection of the real data revealed that the largest catches tended to be centered on sets conducted in 200 m of water; a parabolic relationship, defined using the `sim_parabola` closure, with a mean of 200 m and standard deviation of 70 m approximated this pattern.
+
 # Research opportunities

### 2019-12-19 analysis/imitate_cod_data.R

@@ -45,9 +45,9 @@ library(plotly)
 ## REAL DATA
 
 ## Load survey data compiled for Rstrap
-load("analysis/rv_data/converted_set_details_2018-02-26.Rdata")
-load("analysis/rv_data/converted_length-frequency_data_2018-02-26.Rdata")
-load("analysis/rv_data/age-growth_data_2018-02-26.Rdata")
+load("analysis/rv_data/converted_set_details_2019-10-03.Rdata")
+load("analysis/rv_data/converted_length-frequency_data_2019-10-03.Rdata")
+load("analysis/rv_data/age-growth_data_2019-10-03.Rdata")
 
 ## Subset data to cod
 con.setdet <- con.setdet[(con.setdet$rec == 5 | (con.setdet$rec == 6 & con.setdet$spec == 438)), ]

### 2019-12-29 analysis/paper/sim_survey_paper.Rmd
Address Greg's comments/edits on the revision

@@
 The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of built-in or user-defined sampling protocols to reproduce the population parameters of the known population.
-**`SimSurvey`** also includes a function for estimating the stratified mean and variance of the simulated data.
+**`SimSurvey`** also includes a function for estimating the stratified mean and variance of the population from the simulated survey data.
 **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring.
@@
 The framework follows four general steps: 1) simulate a spatially-aggregated age-structured population; 2) distribute the population throughout a spatial grid, imposing correlation across space, time and age; 3) sample the population using random sampling; and, 4) obtain population estimates using a design-based analysis.
-Though there is a degree of flexibility in each of these steps, keep in mind that users can circumvent specific components by applying user defined equations, inputs and/or analyses.
-Details on how to use the package and circumvent some aspects of its structure are outlined in the **[Using **`SimSurvey`**]** section.
+Though there is a degree of flexibility in each of these steps, users can circumvent specific components by applying user defined equations, inputs and/or analyses.
+Details on how to use the package and, if desired, circumvent some aspects of its structure are outlined in the **[Using **`SimSurvey`**]** section.
 
@@
 Though some typical relationships have yet to be implemented (e.g.
-stock-recruitment), this formulation facilitates the simulation of age-structured populations that are dynamic enough to test an array of survey designs.
+stock-recruitment), this formulation facilitates the simulation of age-structured populations that are dynamic enough to assess survey performance across a range of abundance levels across years and ages.
 
@@
 Rather than developing a full spatially-explicit model, population and spatial dynamics are modeled as independent processes for simplicity.
-The complexities of spatial population dynamics - such as larval dispersal, spatial differences in growth and meta-population connectivity - are therefor not accounted for and, as such, the model is a substantial simplification of reality.
+The complexities of spatial population dynamics - such as larval dispersal, spatial differences in growth and population connectivity - are not explicitly accounted for and, as such, the model is a necessary simplification of reality.
 Despite this limitation, the approach taken facilitates the simulation of spatial, age-structured populations with sufficient complexity for testing the efficacy of various survey designs.
@@
 Parameter estimates from  spatially-aggregated age-structured models, which are commonly used in stock assessments, can therefore be used to simulate a population using the cohort model and the resultant abundance at age values can be distributed across a spatial grid.
-Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across longitude only, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
-The grid can be divided into two hierarchical levels.
+Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across one spatial axis, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
+We use depth as our main stratification variable, but note that any other appropriate stratification variables could be used.
+The grid can be divided into two hierarchical levels, management divisions and habitat-based survey strata.
 For demonstration purposes, we envision these levels as part of a stratified-random survey within international fishery divisions, i.e., $H_{\mathrm{strat}}$ depth-based strata within $H_{\mathrm{div}}$ divisions (e.g.
-NAFO or ICES divisions).
+NAFO or ICES divisions, or any other geographically bounded area).
 The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space.
-This noise term is normalized to sum to 1 to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model:
+This noise term is scaled to sum to 1 to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model:
 
@@
 The final step in the simulation is to sample the simulated population over the age-year-space array generated.
-The sampling is random or stratified random, emulating real-world surveys conducted by many research institutions around the world.
+The sampling is random or stratified random, emulating surveys conducted by many research institutions around the world.
 The area of each strata $A_{\mathrm{strat}}$ is calculated and this is used to define the number of sampling stations $H_{\mathrm{sets}}$, hereafter referred to as sets, allocated to one or more strata under a particular set density, $D_{\mathrm{sets}}$ (i.e.
@@
 Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$.
-In cases where there are multiple sets in one cell, the population in that cell is split across the sets.
+In cases where there are multiple sets in one cell, the population in that cell is divided across the sets.
 While this means that numbers caught in an isolated simulation cannot exceed the numbers in the population, keep in mind that the survey, no matter how intense, is assumed to have no impact on the population from one year to the next.
@@
 
-The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtual fish populations with correlation across space, time and age (`sim_abundance`, `sim_distribution`), 2) simulating surveys with a range of sampling strategies and intensities (`sim_survey`), and 3) estimating the stratified mean and variance of simulated survey data (`test_surveys`; Table \@ref(tab:fun)).
+The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtual fish populations with correlation across space, time and age (`sim_abundance`, `sim_distribution`), 2) simulating surveys with a range of sampling strategies and intensities (`sim_survey`), and 3) estimating the stratified mean and variance of simulated survey data (`run_strat`; Table \@ref(tab:fun)).
 **`SimSurvey`** relies heavily on functions from the **`data.table`** [@dowle2017], **`raster`** [@hijmans2016] and **`plotly`** [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively.
@@
 This was done to avoid the repeated specifications of key arguments, such as ages and years.
-Moreover, this approach provides an option for advanced R users inspect and modify the closures implemented in the package to supply custom closures with alternate equations.
+Moreover, this approach provides an option for advanced R users to inspect and modify the closures implemented in the package to supply custom closures with alternate equations.
 Also note that each of the closures implemented in the package includes a plot argument such that quick visuals can be obtained using a line of code like this: `sim_R(log_mean = log(500), log_sd = 0.5, plot = TRUE)(years = 1:100)`).
@@
 | `+                   n_div = 1,`                        | Number of divisions                                          | $H_{\mathrm{div}}$                      |
-| `+                   strat_splits = 2,`                 | Number of times to horizontally split strat                  | &nbsp;                                  |
+| `+                   strat_splits = 2,`                 | Number of strata within each depth class                     | &nbsp;                                  |
 | `+                   strat_breaks = seq(0, 1000, 40)),` | Series of depth breaks for defining strata                   | &nbsp;                                  |
@@
 
-The equations outlined in the *[Simulate spatial distribution]* section are used in the `make_grid`, `sim_ays_covar` and `sim_parabola` functions, and these functions are used within `sim_distribution` to distribute a population simulated using `sim_abundance` throughout a grid (Table \@ref(tab:distribution)).
+The equations outlined in the **[Simulate spatial distribution]** section are used in the `make_grid`, `sim_ays_covar` and `sim_parabola` functions, and these functions are used within `sim_distribution` to distribute a population simulated using `sim_abundance` throughout a grid (Table \@ref(tab:distribution)).
 The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`.
@@
 
-```{r grid, out.height = "320px", fig.cap = "Plots produced by `plot_grid` when supplied a raster object produced by `make_grid` a) using default settings, b) settings that increase the number of times depth strata are horizontally split, and c) settings that produce a more linear depth gradient and increase the number of divisions.
+```{r grid, out.height = "320px", fig.cap = "Plots produced by `plot_grid` when supplied a raster object produced by `make_grid` a) using default settings, b) settings that increase the number of times depth strata are split, and c) settings that produce a more linear depth gradient and increase the number of divisions.
 In these plots, the yellow to purple color gradient represents depth, the thick grey lines delineate divisions and thin white lines delineate strata."}
@@
 In addition to supplying objects produced by `sim_abundance` and `make_grid`, the `sim_distribution` function requires two closures that describe the age-year-space covariance and the relationship with depth.
-Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by tweaking a few parameters in these closures (Figure \@ref(fig:distribution)).
+Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by adjusting a few parameters in these closures (Figure \@ref(fig:distribution)).
 Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e.
@@
 | `+                   x0 = 3),`      | Midpoint of logistic curve of catchability (age)                        | $x_0$                   |
-| `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (length, width; km) - i.e. area covered by a trawl     | $A_{\mathrm{trawl}}$    |
+| `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (distance towed, trawl width; km) - i.e. area trawled  | $A_{\mathrm{trawl}}$    |
 | `+  min_sets = 2`                   | Minimum number of sets to conduct per strata                            | &nbsp;                  |
@@
 Specifically, the `q` argument requires a closure, such as `sim_logistic`, for defining the probability of catching specific age groups, trawl dimensions are defined in the `trawl_dim` argument, and set, length and age sampling effort are defined using the `set_den`, `lengths_cap` and `ages_cap` arguments, respectively.
-Also note that the `min_sets` argument imposes a minimum of number of sets to conduct per strata, regardless of its allocation given strat area and set density.
+Also note that the `min_sets` argument imposes a minimum of number of sets to conduct per strata, regardless of its allocation given strata area and set density.
 This argument imposes a useful constraint for generating data to be analyzed using design-based approaches that require more than one value for the calculation of a mean.
@@
 
-The default parameter values used in this package were chosen to emulate  survey data from a specific cod population (see **[S1 Appendix][S1 Appendix: Case study]** for details) and, as such, these values ought to provide a good place to start for another population of cod.
-In this section we outline a series of recommended steps to take to tailor the default settings to suit another system.
-Throughout this outline, we also provide some context on how default parameter values were chosen as this information may help guide a users' choices.
+The default parameter values used in this package were chosen to emulate survey data from a specific cod population  (see **[S1 Appendix][S1 Appendix: Case study]** for details) and, as such, these values ought to provide a good place to start for another population of cod or similar groundfish.
+In this section we outline a series of recommended steps to take in order to tailor the default settings to suit other systems.
+Throughout this outline, we also provide some context on how default parameter values were chosen as this information may help guide a users’ choices.
 

### 2019-12-29 analysis/paper/rebuttal_letter.Rmd
Address Greg's edits/comments on the rebuttal letter

@@
 ---
-output: word_document
+output:
+  md_document: default
+  word_document: default
 ---
@@
 Regular, Gregory J.
-Robertson, Bob Rogers, and Keith P.
-Lewis.
+Robertson, Keith P.
+Lewis, Jonathan Babyn, Brian Healey and Fran Mowbray.
 In accordance to one of the reviewers' suggestions, the paper has been renamed **"`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"**.
@@
 Below are the reviewers' comments with a brief accounting of our response to each concern or suggestion.
-
-Lines numbers refer to the revised manuscript.
 
@@
 
-*We hope we have addressed each of these main comments by 1) clarifying several equations, 2) adding more details and justifications, and 3) modifying naming conventions and clarifying extensibility.
+*We hope we have addressed each of these main comments by a) clarifying several equations, b) adding more details and justifications, and c) modifying naming conventions and clarifying extensibility.
 See replies below for more details.*
@@
 
-*We have noted that both built-in and user-defined strategies can be utilised.*
+*We have noted that both built-in and user-defined sampling strategies can be utilised.*
 
@@
 *Good catch.
-The answer to the question is no, bias correction was not applied.
-This is a matter of naming convention that we struggled with.
-The intent was to operate in log space for these equations and, as such, the simulations were set-up to receive log means.
-When developing the functions, however, we waffled on whether to use `log_mean` or `mean` as the argument name and we landed on `mean` as we figured this would be a more relatable value for users to supply.
-Kudos to the reviewer for noticing that we failed to consider bias-correction.
-To be more explicit with regards to what we "mean", we have renamed the arguments to `log_mean`.*
+The answer to the question is no, bias correction was not applied, but nor was it needed, as all demographic parameters operated in log space.
+The issue is a matter of naming convention that we admittedly struggled with.
+When developing the functions, we vacillated on whether to use `log_mean` or `mean` as the argument name and we landed on mean as we believed this would be a more relatable value for users to supply.
+To be more explicit with regards to what value should be supplied, we have renamed the arguments to `log_mean`.*
 
@@
 *We have prefaced this section with the caveat that population and spatial dynamics are modeled as independent processes and we note the pros and cons of this approach.
-In short: the con is that the approach is a simplification of reality as it does not explicitly account for dynamics such as larval dispersal, spatial differences in growth and meta-population connectivity; the pro is that this simplification limits the number of unknown parameters that need to be specified while facilitating the simulation of a sufficiently complex population for testing the efficacy of various survey designs.*
+In short: the con is that the approach is a simplification of reality as it does not explicitly account for dynamics such as larval dispersal, spatial differences in growth and population connectivity; the pro is that this simplification limits the number of unknown parameters that need to be specified while facilitating the simulation of a sufficiently complex population for testing the efficacy of various survey designs.*
 
@@
 
-*Fair point, however, we focused on "depth" for two reasons: 1) it is an important covariate for many species (as the reviewer noted), and 2) most surveys that we are aware of are depth-stratified (also noted by the reviewer below).
-We therefore retained the "depth" naming as it ought to satisfy most use cases.
-Users can satisfy other, less common, cases by supplying user-specified grids and/or spatial populations.*
+*A good point, this variable could be used for any habitat gradient over which stratification is applied.
+However, we retain the term “depth” for three reasons: 1) it is an important covariate for many species (as the reviewer noted), 2) most surveys that we are aware of are depth-stratified (also noted by the reviewer below) and 3) it is a concise variable name (as opposed to something like “habitat_covar”).
+We therefore retained the “depth” naming as it ought to satisfy most use cases.
+We now note directly in the text that any covariate could be supplied and used as ‘depth’*
 
@@
 *Given a suggestion from Reviewer #2, we have restructured the paper to have two core sections: "Model structure" and "Using __`SimSurvey`__".
-We have included a blanket statement under the "Model structure" section noting that users can circumvent specific components of the framework.*
+We have included a blanket statement under the "Model structure" section noting that users can circumvent specific components of the __`SimSurvey`__ framework and supply their own objects, including spatial structure of the survey grid.*
 
@@
 *Good idea!
-We actually pursued this idea in an earlier iteration of `make_grid`, however, we abandoned the option because it was difficult to automate; the random field created random problems with the depth-stratification (e.g.
-one cell strata).*
+We actually pursued this idea in an earlier iteration of `make_grid`, however, we abandoned the option because it was difficult to automate because the random field sometimes created small strata (one cell) or strata that were split in space.
+We would certainly welcome more development integrating other packages with __`SimSurvey`__*
 
@@
 
-*We have added more detail on the structure of the divisions and strata.*
+*We have added more detail on the structure of the divisions and strata.
+In short, divisions are pre-defined areas, generally within which a fisheries management framework is applied.
+Age-length keys are specified at the level of the division, so for populations spanning large habitat gradients (in temperature for example) in which age-length relationships change along that gradient, multiple divisions can be supplied.
+Strata, on the other hand, are defined by an important habitat gradient.
+As structured in __`SimSurvey`__, strata are nested in divisions.
+*
 
@@
 
-*See reply to point 2. above.*
+*We agree, see reply to point 2. above, and note that any habitat variable could be used for stratification.*
 
@@
 
-*We have made the suggested change.*
+*We have made the correction.*
 
@@
 
-*We have clarified how number of sets per strata is calculated.*
+*We have clarified how number of sets per strata is calculated; it is done within the `sim_survey` function using the `set_den` argument, with further adjustment possible with the `min_sets` argument.*
 
@@
 
-*We are not sure what the reviewer would like to have implemented here. Is the suggestion to impose a minimum number of ages to collect across all length groups?*
+*We are not sure what the reviewer would like to have implemented here.
+Is the suggestion to impose a minimum number of ages to collect across all length groups?
+In practice, an `ages_min` constraint is already applied for length bins where the total number of fish caught is lower than `ages_cap`.
+A specific `ages_min` constraint cannot really be applied, as if there are too few fish of a certain size caught in any one survey, there is no means of getting more fish of that size.*
 
@@
 We have added this missing detail to our manuscript.
-We also added a note that the survey is assumed to have no impact on the population from one year to the next.*
+We also added a note that the survey is assumed to have no impact on the population from one year to the next, and users planning to conduct surveys at levels that will affect the overall population will need to use caution in interpreting their results.*
 
@@
 
-*Honestly, we do not recall what we mean by "depending on the number of fish caught".
-Perhaps we added those words to cover off cases where no fish are caught.
-Whatever the case, we have removed the statement to minimize confusion.
+*Our wording "depending on the number of fish caught"  was unfortunately vague and, in hindsight, is not adding anything to the paper.
+Therefore, we have removed the statement to minimize confusion.
 We have also clarified the sub-sampling sequence.
@@
 age-specific clustering tends to result in sets with high intraclass correlation).
-We are open to learning more about other processes that may contribute to correlated samples.*
+Correlation in the sampling procedure could be a future development for __`SimSurvey`__, but we feel at this time the package is sufficiently complex and we did not want to overwhelm users with too many options.*
 
@@
 
-*We have included an example that ought to clarify how a custom closure can be supplied.*
+*We have included an example in the second paragraph of the "Using __`SimSurvey`__" section that we hope will clarify how a user can supply a custom closure.*
 
@@
 
-*We agree with the Reviewer and retitled the ms "`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks".*
+*We agree with the Reviewer and retitled the ms "`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks".
+Currently the package provides options for stratified random sampling proportional to surface area, but other survey designs could added in future versions of __`SimSurvey`__.*
 
@@
 *This is yet another very helpful suggestion and, now that we have included a Parameterisation section, it is easy to see the value of such a section to prospective users of the package.
-However, we should note that we deviated from the recommendation to a degree as we did not produce a table that summaries how to specify the parameters, rather we wrote a section that outlines recommended steps for setting up a custom simulation.
+Instead of a table that summarizes how to specify the parameters, we wrote an entire section that outlines recommended steps for setting up a their own simulation.
 We started drafting a table, however, it quickly became apparent that it would be too big and cumbersome and somewhat redundant with the core tables that describe the arguments and parameters.
@@
 
-*We have now noted in the manuscript that a plus group is not modeled as the number of ages can easily be extended to include groups with zero fish.
+*We have now noted in the manuscript that a plus group is not modeled explicitly as the number of ages can easily be extended to include groups with zero fish.
 This choice simplifies the simulation, including the setting of initial numbers which is done via exponential decay.
@@
 
-*This was actually correct and, given an improved description of a closure, the logic behind this line of code should be clearer now.*
+*We have provided an improved description of a closure; the logic behind this line of code should be clearer now*
 

### 2019-12-29 analysis/paper/rebuttal_letter.md
Address Greg's edits/comments on the rebuttal letter

--- 2019-12-29 analysis/paper/rebuttal_letter.Rmd
@@
----
-output:
-  md_document: default
-  word_document: default
----
-
 Daniel E. Duplisea, PhD  
@@
 
-`r format(Sys.time(), '%Y-%m-%d')`
+2019-12-29
 
@@
 
-Thank you for considering a revision of the manuscript PONE-D-19-26904,  **"`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"** by Paul M.
-Regular, Gregory J.
-Robertson, Keith P.
-Lewis, Jonathan Babyn, Brian Healey and Fran Mowbray.
-In accordance to one of the reviewers' suggestions, the paper has been renamed **"`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"**.
+Thank you for considering a revision of the manuscript PONE-D-19-26904,
+**“`SimSurvey`: an `R` package to optimize the design and analysis of
+fisheries surveys by simulating spatially-correlated fish stocks”** by
+Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, Jonathan Babyn,
+Brian Healey and Fran Mowbray. In accordance to one of the reviewers’
+suggestions, the paper has been renamed **“`SimSurvey`: an `R` package
+for comparing the design and analysis of fisheries surveys by simulating
+spatially-correlated fish stocks”**.
 
-We thank the Reviewers for their thoughtful and thorough comments which will clarify and improve the manuscript.
-We agree with many of the comments and have made every effort to incorporate these into the manuscript.
-Below are the reviewers' comments with a brief accounting of our response to each concern or suggestion.
+We thank the Reviewers for their thoughtful and thorough comments which
+will clarify and improve the manuscript. We agree with many of the
+comments and have made every effort to incorporate these into the
+manuscript. Below are the reviewers’ comments with a brief accounting of
+our response to each concern or suggestion.
 
-We submit this revised manuscript for your consideration and look forward to your decision.
+We submit this revised manuscript for your consideration and look
+forward to your decision.
 
@@
 A1C 5X1, Canada  
-E-mail: Paul.Regular@dfo-mpo.gc.ca  
+E-mail:
+<a href="mailto:Paul.Regular@dfo-mpo.gc.ca" class="email">Paul.Regular@dfo-mpo.gc.ca</a>  
 Phone: (709) 772-2067
 
-----------------------------------------------------------------------------------------------------
+------------------------------------------------------------------------
 
-Reviewer #1:
+Reviewer \#1:
 
-This paper describes the R package “SimSurvey” which is a package to test the design and analysis of fisheries independent surveys.
+This paper describes the R package “SimSurvey” which is a package to
+test the design and analysis of fisheries independent surveys.
 
-I found the paper very interesting and the package of great potential value and interest to many fishery scientists.
-However, I have several comments that I would like the authors to address first before I recommend acceptance of this paper.
-My main comments relate to:
+I found the paper very interesting and the package of great potential
+value and interest to many fishery scientists. However, I have several
+comments that I would like the authors to address first before I
+recommend acceptance of this paper. My main comments relate to:
 
-1.
-Verify some of the equations and model description used in the model.
-I think that some of them lack description to fully understand what was done and few others are misleading, even incorrect.
-The detail can be found below but in general, they are linked to the problem of bias correction for log-normal distribution, re-scaling”, moving from abundance at age to abundance at length.
+1.  Verify some of the equations and model description used in the
+    model. I think that some of them lack description to fully
+    understand what was done and few others are misleading, even
+    incorrect. The detail can be found below but in general, they are
+    linked to the problem of bias correction for log-normal
+    distribution, re-scaling”, moving from abundance at age to abundance
+    at length.
 
-2. Add additional explanation to justify the choices. E.g. why population dynamics is modeled independently of the spatial distribution. Is this reasonable, limitation, etc.
+2.  Add additional explanation to justify the choices. E.g. why
+    population dynamics is modeled independently of the spatial
+    distribution. Is this reasonable, limitation, etc.
 
-3. Few suggestions (add-ons, renaming) to increase the generality of the paper (without requiring too much work)
+3.  Few suggestions (add-ons, renaming) to increase the generality of
+    the paper (without requiring too much work)
 
-*We hope we have addressed each of these main comments by a) clarifying several equations, b) adding more details and justifications, and c) modifying naming conventions and clarifying extensibility.
-See replies below for more details.*
+*We hope we have addressed each of these main comments by a) clarifying
+several equations, b) adding more details and justifications, and c)
+modifying naming conventions and clarifying extensibility. See replies
+below for more details.*
 
@@
 
-What about the flexibility to include new sampling strategies (user defined one) or including new estimation approaches
+What about the flexibility to include new sampling strategies (user
+defined one) or including new estimation approaches
 
-*We have noted that both built-in and user-defined sampling strategies can be utilised.*
+*We have noted that both built-in and user-defined sampling strategies
+can be utilised.*
 
@@
 
-L69-73: I think it would be good to add the reference to the specific function name listed in Table2.
-This way, I would be more explicit and avoid possible confusion.
-For example, in my case, when I first read the section header “simulate abundance”, I was expecting to see the way you dealt with the spatial distribution of abundance.
+L69-73: I think it would be good to add the reference to the specific
+function name listed in Table2. This way, I would be more explicit and
+avoid possible confusion. For example, in my case, when I first read the
+section header “simulate abundance”, I was expecting to see the way you
+dealt with the spatial distribution of abundance.
 
@@
 
-L94&95&104: I just want to make sure that the bias correction has been applied to these two equations.
-If logN (or logZ) is normally distribution with mean mu and sigma, then the N (or Z) is lognormally distributed with mean exp(mu+sigma^2/2).
-Therefore, when you write “mean” in table 3 for R and Z, I hope that you modified in your actual equation for log(Z) to:
-log(Z) ~ normal(log(mean)-sigma^2/2, sigma)
-In other word, the mean(R) is not simply exp(meanR_log_scale). exp(meanR_log_scale) is the median of the distribution of R.
+L94&95&104: I just want to make sure that the bias correction has been
+applied to these two equations. If logN (or logZ) is normally
+distribution with mean mu and sigma, then the N (or Z) is lognormally
+distributed with mean exp(mu+sigma^2/2). Therefore, when you write
+“mean” in table 3 for R and Z, I hope that you modified in your actual
+equation for log(Z) to: log(Z) ~ normal(log(mean)-sigma^2/2, sigma) In
+other word, the mean(R) is not simply exp(meanR\_log\_scale).
+exp(meanR\_log\_scale) is the median of the distribution of R.
 
-*Good catch.
-The answer to the question is no, bias correction was not applied, but nor was it needed, as all demographic parameters operated in log space.
-The issue is a matter of naming convention that we admittedly struggled with.
-When developing the functions, we vacillated on whether to use `log_mean` or `mean` as the argument name and we landed on mean as we believed this would be a more relatable value for users to supply.
-To be more explicit with regards to what value should be supplied, we have renamed the arguments to `log_mean`.*
+*Good catch. The answer to the question is no, bias correction was not
+applied, but nor was it needed, as all demographic parameters operated
+in log space. The issue is a matter of naming convention that we
+admittedly struggled with. When developing the functions, we vacillated
+on whether to use `log_mean` or `mean` as the argument name and we
+landed on mean as we believed this would be a more relatable value for
+users to supply. To be more explicit with regards to what value should
+be supplied, we have renamed the arguments to `log_mean`.*
 
-L107: there is not enough description to understand how you converted the abundance at age to abundance at length.
-What were the choice of length bin groups?
-Every 5 cm, 10cm?
-Such information is also missing in the tables.
-Please add that.
-But then you have to calculate some cumulative distribution (until length l) for the length at age distribution (using the VB growth equation) and do some subtractions to be able to allocate the abundance at age to each length group.
-See equation A.1.14 in here for example. https://www.nwfsc.noaa.gov/news/events/program_reviews/documents/C.2_Methot_Wetzel_SSTechnicalDescription.pdf
+L107: there is not enough description to understand how you converted
+the abundance at age to abundance at length. What were the choice of
+length bin groups? Every 5 cm, 10cm? Such information is also missing in
+the tables. Please add that. But then you have to calculate some
+cumulative distribution (until length l) for the length at age
+distribution (using the VB growth equation) and do some subtractions to
+be able to allocate the abundance at age to each length group. See
+equation A.1.14 in here for example.
+<a href="https://www.nwfsc.noaa.gov/news/events/program_reviews/documents/C.2_Methot_Wetzel_SSTechnicalDescription.pdf" class="uri">https://www.nwfsc.noaa.gov/news/events/program_reviews/documents/C.2_Methot_Wetzel_SSTechnicalDescription.pdf</a>
 
-*Another good catch as we did not provide enough information to replicate the approach.
-We have added more detail on the calculations behind the sim_vonB function; turns out our approach was similar to the one described in the exemplar the reviewer provided.*
+*Another good catch as we did not provide enough information to
+replicate the approach. We have added more detail on the calculations
+behind the sim\_vonB function; turns out our approach was similar to the
+one described in the exemplar the reviewer provided.*
 
@@
 
-1.
-First of all, I think it is important to state that you assume that the population dynamic model and the spatial distribution is independent i.e.
-you do not have a spatially-explicit population dynamics model.
-And talk about what it means (is it realistic, limitation, etc).
+1.  First of all, I think it is important to state that you assume that
+    the population dynamic model and the spatial distribution is
+    independent i.e. you do not have a spatially-explicit population
+    dynamics model. And talk about what it means (is it realistic,
+    limitation, etc).
 
-*We have prefaced this section with the caveat that population and spatial dynamics are modeled as independent processes and we note the pros and cons of this approach.
-In short: the con is that the approach is a simplification of reality as it does not explicitly account for dynamics such as larval dispersal, spatial differences in growth and population connectivity; the pro is that this simplification limits the number of unknown parameters that need to be specified while facilitating the simulation of a sufficiently complex population for testing the efficacy of various survey designs.*
+*We have prefaced this section with the caveat that population and
+spatial dynamics are modeled as independent processes and we note the
+pros and cons of this approach. In short: the con is that the approach
+is a simplification of reality as it does not explicitly account for
+dynamics such as larval dispersal, spatial differences in growth and
+population connectivity; the pro is that this simplification limits the
+number of unknown parameters that need to be specified while
+facilitating the simulation of a sufficiently complex population for
+testing the efficacy of various survey designs.*
 
-2.
-L161.
-I think it would be better to rename “depth” as “main covariate influencing the species distribution”.
-And “depth” happens to be one of them for many species, but there are cases when this is not the case.
+1.  L161. I think it would be better to rename “depth” as “main
+    covariate influencing the species distribution”. And “depth” happens
+    to be one of them for many species, but there are cases when this is
+    not the case.
 
-*A good point, this variable could be used for any habitat gradient over which stratification is applied.
-However, we retain the term “depth” for three reasons: 1) it is an important covariate for many species (as the reviewer noted), 2) most surveys that we are aware of are depth-stratified (also noted by the reviewer below) and 3) it is a concise variable name (as opposed to something like “habitat_covar”).
-We therefore retained the “depth” naming as it ought to satisfy most use cases.
-We now note directly in the text that any covariate could be supplied and used as ‘depth’*
+*A good point, this variable could be used for any habitat gradient over
+which stratification is applied. However, we retain the term “depth” for
+three reasons: 1) it is an important covariate for many species (as the
+reviewer noted), 2) most surveys that we are aware of are
+depth-stratified (also noted by the reviewer below) and 3) it is a
+concise variable name (as opposed to something like “habitat\_covar”).
+We therefore retained the “depth” naming as it ought to satisfy most use
+cases. We now note directly in the text that any covariate could be
+supplied and used as ‘depth’*
 
-3. L161. I think it would be good to say upfront that there are two ways of defining the spatial structure. Using the function in the package or user defined.
+1.  L161. I think it would be good to say upfront that there are two
+    ways of defining the spatial structure. Using the function in the
+    package or user defined.
 
-*Given a suggestion from Reviewer #2, we have restructured the paper to have two core sections: "Model structure" and "Using __`SimSurvey`__".
-We have included a blanket statement under the "Model structure" section noting that users can circumvent specific components of the __`SimSurvey`__ framework and supply their own objects, including spatial structure of the survey grid.*
+*Given a suggestion from Reviewer \#2, we have restructured the paper to
+have two core sections: “Model structure” and “Using **`SimSurvey`**”.
+We have included a blanket statement under the “Model structure” section
+noting that users can circumvent specific components of the
+**`SimSurvey`** framework and supply their own objects, including
+spatial structure of the survey grid.*
 
-4.
-L161: I think another possibility is to generate a random field by using the package “Randomfields” for example.
-This way you can generate a map where “depth” is patchily distributed (maybe more like an island type case study)
+1.  L161: I think another possibility is to generate a random field by
+    using the package “Randomfields” for example. This way you can
+    generate a map where “depth” is patchily distributed (maybe more
+    like an island type case study)
 
-*Good idea!
-We actually pursued this idea in an earlier iteration of `make_grid`, however, we abandoned the option because it was difficult to automate because the random field sometimes created small strata (one cell) or strata that were split in space.
-We would certainly welcome more development integrating other packages with __`SimSurvey`__*
+*Good idea! We actually pursued this idea in an earlier iteration of
+`make_grid`, however, we abandoned the option because it was difficult
+to automate because the random field sometimes created small strata (one
+cell) or strata that were split in space. We would certainly welcome
+more development integrating other packages with **`SimSurvey`***
 
-5.
-L163-164: please add more description about the division or strata.
-How can we set it up and what do you specifically mean by division and strata?
-Is one nested within the other, or not necessarily?
-Your examples are based on Atlantic Ocean and people in other regions might not be familiar with how these divisions are created.
+1.  L163-164: please add more description about the division or strata.
+    How can we set it up and what do you specifically mean by division
+    and strata? Is one nested within the other, or not necessarily? Your
+    examples are based on Atlantic Ocean and people in other regions
+    might not be familiar with how these divisions are created.
 
 *We have added more detail on the structure of the divisions and strata.
-In short, divisions are pre-defined areas, generally within which a fisheries management framework is applied.
-Age-length keys are specified at the level of the division, so for populations spanning large habitat gradients (in temperature for example) in which age-length relationships change along that gradient, multiple divisions can be supplied.
-Strata, on the other hand, are defined by an important habitat gradient.
-As structured in __`SimSurvey`__, strata are nested in divisions.
-*
+In short, divisions are pre-defined areas, generally within which a
+fisheries management framework is applied. Age-length keys are specified
+at the level of the division, so for populations spanning large habitat
+gradients (in temperature for example) in which age-length relationships
+change along that gradient, multiple divisions can be supplied. Strata,
+on the other hand, are defined by an important habitat gradient. As
+structured in **`SimSurvey`**, strata are nested in divisions. *
 
-6.
-L163: Why only focus on “depth-based strata”?
-I think it would be good to allow the user to choose their own stratification approach.
-It could be depth based as you did (which happens most often in surveys) but it could technically be any other thing (user supplied).
-This allows more flexibility.
+1.  L163: Why only focus on “depth-based strata”? I think it would be
+    good to allow the user to choose their own stratification approach.
+    It could be depth based as you did (which happens most often in
+    surveys) but it could technically be any other thing (user
+    supplied). This allows more flexibility.
 
-*We agree, see reply to point 2. above, and note that any habitat variable could be used for stratification.*
+*We agree, see reply to point 2. above, and note that any habitat
+variable could be used for stratification.*
 
-7.
-L167: the equation is misleading and I am not sure it is right based on what is written in the text.
-You mentioned later on L178: that you “re-scaled” so that the total number of fish in a specific year and age across space is equal to the single number from the population dynamics model.
-If so, the re-scaling should be done in the identify (natural) scale, not in log scale.
-In log scale, even if ,, sums to zero across space, the sum in the identity scale won’t match.
-This is often refer to as “bias correction” for the log normal models.
-And you should ideally show how the rescaling was done in terms of the equations too.
+1.  L167: the equation is misleading and I am not sure it is right based
+    on what is written in the text. You mentioned later on L178: that
+    you “re-scaled” so that the total number of fish in a specific year
+    and age across space is equal to the single number from the
+    population dynamics model. If so, the re-scaling should be done in
+    the identify (natural) scale, not in log scale. In log scale, even
+    if ,, sums to zero across space, the sum in the identity scale won’t
+    match. This is often refer to as “bias correction” for the log
+    normal models. And you should ideally show how the rescaling was
+    done in terms of the equations too.
 
-*Right, the equation presented was not an accurate reflection of the calculations. We have revised the equation to explicitly show how the values were normalized to sum to 1.*
+*Right, the equation presented was not an accurate reflection of the
+calculations. We have revised the equation to explicitly show how the
+values were normalized to sum to 1.*
 
-8. L167: this “depth preference” function is very simplistic and gives only very “smooth” symmetric distribution. More often, fish have a skewed depth preference: often right-skewed.
+1.  L167: this “depth preference” function is very simplistic and gives
+    only very “smooth” symmetric distribution. More often, fish have a
+    skewed depth preference: often right-skewed.
 
-*True.
-Some users may find this parameterization insufficient for their species and we hope they will implement their own closure to use in the `sim_distribution` function to better simulate the effect.
-In addition to our blanket statement under the "Model structure" section, we have added a more specific statement under the "Using __`SimSurvey`__" section stating that alternate formulations can be used by supplying alternate closures to the core functions.*
+*True. Some users may find this parameterization insufficient for their
+species and we hope they will implement their own closure to use in the
+`sim_distribution` function to better simulate the effect. In addition
+to our blanket statement under the “Model structure” section, we have
+added a more specific statement under the “Using **`SimSurvey`**”
+section stating that alternate formulations can be used by supplying
+alternate closures to the core functions.*
 
-9. L173-174: I think it might be worth adding, in simpler terms, the meaning of the spatial smoothing and scaling parameters.
+1.  L173-174: I think it might be worth adding, in simpler terms, the
+    meaning of the spatial smoothing and scaling parameters.
 
-*We have prefaced that sentence with "The rate at which point-to-point spatial correlation decays with distance is controlled by...".*
+*We have prefaced that sentence with “The rate at which point-to-point
+spatial correlation decays with distance is controlled by…”.*
 
-10.
-L178: it is another question of scaling.
-How did you exactly do the scaling?
-In the identity scale?
-By dividing my the sum of the effects?
-I am asking this because depending on how you did the rescaling, your correlation structure in space and age might have been affected and is not the same as the one specified in L172.
-Did you verify that?
+1.  L178: it is another question of scaling. How did you exactly do the
+    scaling? In the identity scale? By dividing my the sum of the
+    effects? I am asking this because depending on how you did the
+    rescaling, your correlation structure in space and age might have
+    been affected and is not the same as the one specified in L172. Did
+    you verify that?
 
@@
 
-Table 4. “group_ages”. Ok but how is the variance controlled for the other age classes?
+Table 4. “group\_ages”. Ok but how is the variance controlled for the
+other age classes?
 
-*"Variance" was a poor word choice. We have replaced it with "noise" as it is the simulated noise that we fix across multiple age groups.*
+*“Variance” was a poor word choice. We have replaced it with “noise” as
+it is the simulated noise that we fix across multiple age groups.*
 
-L189: “user supplied”.
-This is a good feature.
-However, I think it is important to mention here that user have to make sure that they use the correct projection method to ensure that each grid is of the same dimension.
+L189: “user supplied”. This is a good feature. However, I think it is
+important to mention here that user have to make sure that they use the
+correct projection method to ensure that each grid is of the same
+dimension.
 
@@
 
-L221: there is not “group_years” argument in Table 4.
+L221: there is not “group\_years” argument in Table 4.
 
@@
 
-L237: “this function”. I think it would be better to replace “this function” with “sim_distribution” as you do not mention the word “sim_distribution” in the sentence above this.
+L237: “this function”. I think it would be better to replace “this
+function” with “sim\_distribution” as you do not mention the word
+“sim\_distribution” in the sentence above this.
 
@@
 
-L158-251: In general, I think re-organising this section using sub-section headers could be useful. Just to guide the readers
+L158-251: In general, I think re-organising this section using
+sub-section headers could be useful. Just to guide the readers
 
-*This is an excellent suggestion. By following a suggestion by Reviewer #2 to re-organize the paper into two core sections we have added more headers to help guide the readers.*
+*This is an excellent suggestion. By following a suggestion by Reviewer
+\#2 to re-organize the paper into two core sections we have added more
+headers to help guide the readers.*
 
-L254: you say that sampling is stratified random but SRS is also an option based on Table5. Please correct.
+L254: you say that sampling is stratified random but SRS is also an
+option based on Table5. Please correct.
 
@@
 
-L257: what does this mean? Does this control the number of set but how is this calculated?
+L257: what does this mean? Does this control the number of set but how
+is this calculated?
 
@@
 
-L257: I do not see how you control for the total number of set in the survey? How do you control it?
+L257: I do not see how you control for the total number of set in the
+survey? How do you control it?
 
-*We have clarified how number of sets per strata is calculated; it is done within the `sim_survey` function using the `set_den` argument, with further adjustment possible with the `min_sets` argument.*
+*We have clarified how number of sets per strata is calculated; it is
+done within the `sim_survey` function using the `set_den` argument, with
+further adjustment possible with the `min_sets` argument.*
 
-L261: I think you should mention here that you can also force the sample size (as seen on table 5).
-Moreover, in table 5, it would be good to set-up a “ages_min” for the minimum number of ages to sample […]” so that it gives the ability to fix the sample size if needed by writing the same value for “ages_min” and “ages_cap”.
+L261: I think you should mention here that you can also force the sample
+size (as seen on table 5). Moreover, in table 5, it would be good to
+set-up a “ages\_min” for the minimum number of ages to sample \[…\]” so
+that it gives the ability to fix the sample size if needed by writing
+the same value for “ages\_min” and “ages\_cap”.
 
 *We are not sure what the reviewer would like to have implemented here.
-Is the suggestion to impose a minimum number of ages to collect across all length groups?
-In practice, an `ages_min` constraint is already applied for length bins where the total number of fish caught is lower than `ages_cap`.
-A specific `ages_min` constraint cannot really be applied, as if there are too few fish of a certain size caught in any one survey, there is no means of getting more fish of that size.*
+Is the suggestion to impose a minimum number of ages to collect across
+all length groups? In practice, an `ages_min` constraint is already
+applied for length bins where the total number of fish caught is lower
+than `ages_cap`. A specific `ages_min` constraint cannot really be
+applied, as if there are too few fish of a certain size caught in any
+one survey, there is no means of getting more fish of that size.*
 
-L261: How are you making sure that the number of sampled fish for that specific cell, age and year won’t be above the total number of fish in that cell, age, and year?
-The probability value could be close to one and if you fish in a few a time, then you are at risk.
-Especially because your population dynamics model is not spatially explicit and is completely independent of the distribution function itself i.e.
-you can technically fish out all the fish in an area but it will be populated back the year after the way you implemented in this study… Maybe you need to put a condition (or just a note) for general users to make sure that this probability value is much below 1?
+L261: How are you making sure that the number of sampled fish for that
+specific cell, age and year won’t be above the total number of fish in
+that cell, age, and year? The probability value could be close to one
+and if you fish in a few a time, then you are at risk. Especially
+because your population dynamics model is not spatially explicit and is
+completely independent of the distribution function itself i.e. you can
+technically fish out all the fish in an area but it will be populated
+back the year after the way you implemented in this study… Maybe you
+need to put a condition (or just a note) for general users to make sure
+that this probability value is much below 1?
 
-*The sampling is implemented such that the number of fish sampled in a cell cannot exceed the number of fish in a cell because the population is split across sets in cases where more than one set is conducted in a cell.
-We have added this missing detail to our manuscript.
-We also added a note that the survey is assumed to have no impact on the population from one year to the next, and users planning to conduct surveys at levels that will affect the overall population will need to use caution in interpreting their results.*
+*The sampling is implemented such that the number of fish sampled in a
+cell cannot exceed the number of fish in a cell because the population
+is split across sets in cases where more than one set is conducted in a
+cell. We have added this missing detail to our manuscript. We also added
+a note that the survey is assumed to have no impact on the population
+from one year to the next, and users planning to conduct surveys at
+levels that will affect the overall population will need to use caution
+in interpreting their results.*
 
-L267: I recommend to clarify something here.
-1.
-Depending on the number of fish caught?
-What do you mean?
-What is the rule you used?
-2.
-The way you coded, sample by age is first decided, then the corresponding length is calculated, then age-subsample is determined.
-In reality, length sample is taken in the field, then age sub-samples are taken.
-While similar, I do not think it always equal.
-Especially, when you start including some correlation structure in the sampling.
-By the way, did you consider including some correlation structure in the sampling process to make more realistic?
+L267: I recommend to clarify something here. 1. Depending on the number
+of fish caught? What do you mean? What is the rule you used? 2. The way
+you coded, sample by age is first decided, then the corresponding length
+is calculated, then age-subsample is determined. In reality, length
+sample is taken in the field, then age sub-samples are taken. While
+similar, I do not think it always equal. Especially, when you start
+including some correlation structure in the sampling. By the way, did
+you consider including some correlation structure in the sampling
+process to make more realistic?
 
-*Our wording "depending on the number of fish caught"  was unfortunately vague and, in hindsight, is not adding anything to the paper.
-Therefore, we have removed the statement to minimize confusion.
-We have also clarified the sub-sampling sequence.
-Finally, we have yet to consider including correlation structure in the sampling process as we went about imposing correlation via the spatial correlation of age groups (i.e.
-age-specific clustering tends to result in sets with high intraclass correlation).
-Correlation in the sampling procedure could be a future development for __`SimSurvey`__, but we feel at this time the package is sufficiently complex and we did not want to overwhelm users with too many options.*
+*Our wording “depending on the number of fish caught” was unfortunately
+vague and, in hindsight, is not adding anything to the paper. Therefore,
+we have removed the statement to minimize confusion. We have also
+clarified the sub-sampling sequence. Finally, we have yet to consider
+including correlation structure in the sampling process as we went about
+imposing correlation via the spatial correlation of age groups
+(i.e. age-specific clustering tends to result in sets with high
+intraclass correlation). Correlation in the sampling procedure could be
+a future development for **`SimSurvey`**, but we feel at this time the
+package is sufficiently complex and we did not want to overwhelm users
+with too many options.*
 
@@
 
-L275: Table5: “age_sammpling” should be “age_sampling”
+L275: Table5: “age\_sammpling” should be “age\_sampling”
 
@@
 
-L275: “min_sets” you have not described it yet and what is it? You have sample from all cells? If no, this is not realistic.
+L275: “min\_sets” you have not described it yet and what is it? You have
+sample from all cells? If no, this is not realistic.
 
-*We have clarified the meaning and utility of the `min_sets` argument (i.e.
-a small strata may be allocated only one set under a low set density scenario; this argument overrides the allocation and imposes the `min_sets` if it is greater than the allocation).*
+*We have clarified the meaning and utility of the `min_sets` argument
+(i.e. a small strata may be allocated only one set under a low set
+density scenario; this argument overrides the allocation and imposes the
+`min_sets` if it is greater than the allocation).*
 
@@
 
-L285-286: Could you be more specific on how custom closures can be supplied and where?
+L285-286: Could you be more specific on how custom closures can be
+supplied and where?
 
-*We have included an example in the second paragraph of the "Using __`SimSurvey`__" section that we hope will clarify how a user can supply a custom closure.*
+*We have included an example in the second paragraph of the “Using
+**`SimSurvey`**” section that we hope will clarify how a user can supply
+a custom closure.*
 
-L306: how are these catchability corrected abundance matrices calculated?
-It is important to write this information somewhere (or write “please refer to the section “Stratified analysis” for further information on the calculation of abundance indices”) or something alike and Appendix S3.
+L306: how are these catchability corrected abundance matrices
+calculated? It is important to write this information somewhere (or
+write “please refer to the section “Stratified analysis” for further
+information on the calculation of abundance indices”) or something alike
+and Appendix S3.
 
@@
 
-L336: I think it would be good to say that other methods exist and people can use it in this package (maybe)?
+L336: I think it would be good to say that other methods exist and
+people can use it in this package (maybe)?
 
-*Good point, however, we think this is covered by referencing a paper that describes a geostatistical R package and we also note that other options can be used under the "Research opportunities" section.*
+*Good point, however, we think this is covered by referencing a paper
+that describes a geostatistical R package and we also note that other
+options can be used under the “Research opportunities” section.*
 
-L421: color gradient. Even though it is obvious it might be good to say green to purple gradient.
+L421: color gradient. Even though it is obvious it might be good to say
+green to purple gradient.
 
@@
 
-L427: instead of “sampling protocol”, I think it would be more meaning full to say the maximum number of length samples.
+L427: instead of “sampling protocol”, I think it would be more meaning
+full to say the maximum number of length samples.
 
@@
 
-----------------------------------------------------------------------------------------------------
+------------------------------------------------------------------------
 
-Reviewer #2:
+Reviewer \#2:
 
-This manuscript describes an R package called SimSurvey.
-The package includes a set of functions for simulating point-based fisheries survey designs, e.g.
-bottom trawl surveys, for estimating abundance indices.
-It focuses on number of stations and number of fish sampled ignores other constraints such as distance between stations and day-time duration which impose strong constraints on real surveys.
-The functions included allow the user to first simulate age-and length structured population dynamics, distribute individuals randomly in space (assuming a certain correlation structure), carry out a survey and finally calculate abundance indices from the simulated data.
-The package will likely be of interest to researchers wanting to explore the precision achievable with different survey designs.
-My comments regarding the package and the presentation are summarised below.
+This manuscript describes an R package called SimSurvey. The package
+includes a set of functions for simulating point-based fisheries survey
+designs, e.g. bottom trawl surveys, for estimating abundance indices. It
+focuses on number of stations and number of fish sampled ignores other
+constraints such as distance between stations and day-time duration
+which impose strong constraints on real surveys. The functions included
+allow the user to first simulate age-and length structured population
+dynamics, distribute individuals randomly in space (assuming a certain
+correlation structure), carry out a survey and finally calculate
+abundance indices from the simulated data. The package will likely be of
+interest to researchers wanting to explore the precision achievable with
+different survey designs. My comments regarding the package and the
+presentation are summarised below.
 
@@
 
-The title announces a package for optimizing survey designs.
-As far as I can see the package does not allow survey design optimization, neither in terms of defining survey strata nor in terms of number of stations per stratum.
-The strata are defined by the user.
-The only option available for the number of stations is proportional to stratum surface; the user sets the minimum number of stations taken in the smallest stratum.
-It would be useful to be able to specify the total number of stations and test different allocation schemes, such as proportional to surface area (implemented), equal number per stratum, Neyman allocation (accounting for surface area and abundance variability), etc.
-Please consider revising the title (e.g. “compare” and instead of “optimize)” and spell out the available sampling design options.
+The title announces a package for optimizing survey designs. As far as I
+can see the package does not allow survey design optimization, neither
+in terms of defining survey strata nor in terms of number of stations
+per stratum. The strata are defined by the user. The only option
+available for the number of stations is proportional to stratum surface;
+the user sets the minimum number of stations taken in the smallest
+stratum. It would be useful to be able to specify the total number of
+stations and test different allocation schemes, such as proportional to
+surface area (implemented), equal number per stratum, Neyman allocation
+(accounting for surface area and abundance variability), etc. Please
+consider revising the title (e.g. “compare” and instead of “optimize)”
+and spell out the available sampling design options.
 
-*We agree with the Reviewer and retitled the ms "`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks".
-Currently the package provides options for stratified random sampling proportional to surface area, but other survey designs could added in future versions of __`SimSurvey`__.*
+*We agree with the Reviewer and retitled the ms “`SimSurvey`: an `R`
+package for comparing the design and analysis of fisheries surveys by
+simulating spatially-correlated fish stocks”. Currently the package
+provides options for stratified random sampling proportional to surface
+area, but other survey designs could added in future versions of
+**`SimSurvey`**.*
 
-2. Manuscript structure
+1.  Manuscript structure
 
-The manuscript might be easier to follow if the manuscript was restructured: 1) Model description, 2) Using SimSurvey.
-The later section would then group all example code which could again be subdivided into running simulations and exploring results (plot functions).
+The manuscript might be easier to follow if the manuscript was
+restructured: 1) Model description, 2) Using SimSurvey. The later
+section would then group all example code which could again be
+subdivided into running simulations and exploring results (plot
+functions).
 
-*This is an excellent suggestion! We have re-structured our manuscript accordingly and feel that this structure will be much easier for a reader to follow.*
+*This is an excellent suggestion! We have re-structured our manuscript
+accordingly and feel that this structure will be much easier for a
+reader to follow.*
 
-3. Parameterisation
-To use SimSurvey for a real world problem realistic parameter values are needed.
-The package comes with default values chosen for a particular case study.
-However, no mention is made in the manuscript how to choose appropriate values for the many model parameters to tailor the simulations to a population of interest.
-I suggest the authors add a section on parameterisation and a table summarizing all parameters with a column specifying how to parameterize.
-For example, parameters for population dynamics and growth could probably be taken from the literature (or a stock assessment report).
-However this is not possible for the parameters of the spatial distribution function sim_distribution() such as correlation between ages etc.
-Ideally the package would include a fitting function for estimating these parameters from actual data.
-These input data could come from a pilot survey and include location (lat, long) and numbers by length/age.
+1.  Parameterisation To use SimSurvey for a real world problem realistic
+    parameter values are needed. The package comes with default values
+    chosen for a particular case study. However, no mention is made in
+    the manuscript how to choose appropriate values for the many model
+    parameters to tailor the simulations to a population of interest. I
+    suggest the authors add a section on parameterisation and a table
+    summarizing all parameters with a column specifying how to
+    parameterize. For example, parameters for population dynamics and
+    growth could probably be taken from the literature (or a stock
+    assessment report). However this is not possible for the parameters
+    of the spatial distribution function sim\_distribution() such as
+    correlation between ages etc. Ideally the package would include a
+    fitting function for estimating these parameters from actual data.
+    These input data could come from a pilot survey and include location
+    (lat, long) and numbers by length/age.
 
-*This is yet another very helpful suggestion and, now that we have included a Parameterisation section, it is easy to see the value of such a section to prospective users of the package.
-Instead of a table that summarizes how to specify the parameters, we wrote an entire section that outlines recommended steps for setting up a their own simulation.
-We started drafting a table, however, it quickly became apparent that it would be too big and cumbersome and somewhat redundant with the core tables that describe the arguments and parameters.
-We hope that the narrative/outline we included will serve the practical purpose the reviewer had in mind.*
+*This is yet another very helpful suggestion and, now that we have
+included a Parameterisation section, it is easy to see the value of such
+a section to prospective users of the package. Instead of a table that
+summarizes how to specify the parameters, we wrote an entire section
+that outlines recommended steps for setting up a their own simulation.
+We started drafting a table, however, it quickly became apparent that it
+would be too big and cumbersome and somewhat redundant with the core
+tables that describe the arguments and parameters. We hope that the
+narrative/outline we included will serve the practical purpose the
+reviewer had in mind.*
 
-Minor issues
-- line 93: I assume there is an age plus but this needs to be mentioned. Also, please specify how you set the initial numbers for plus group ().
+Minor issues - line 93: I assume there is an age plus but this needs to
+be mentioned. Also, please specify how you set the initial numbers for
+plus group ().
 
-*We have now noted in the manuscript that a plus group is not modeled explicitly as the number of ages can easily be extended to include groups with zero fish.
-This choice simplifies the simulation, including the setting of initial numbers which is done via exponential decay.
-Further, the lack of a plus group is inconsequential for survey based estimates of abundance at age.*
+*We have now noted in the manuscript that a plus group is not modeled
+explicitly as the number of ages can easily be extended to include
+groups with zero fish. This choice simplifies the simulation, including
+the setting of initial numbers which is done via exponential decay.
+Further, the lack of a plus group is inconsequential for survey based
+estimates of abundance at age.*
 
-- line 121: I don’t understand the explanation of a closure. What do you mean by “return functions”? Do you mean it returns an object with different attributes?
+-   line 121: I don’t understand the explanation of a closure. What do
+    you mean by “return functions”? Do you mean it returns an object
+    with different attributes?
 
-*We have improved our explanation of a closure at the beginning of the Using __`SimSurvey`__ section*
+*We have improved our explanation of a closure at the beginning of the
+Using **`SimSurvey`** section*
 
-- line 125: the number of right and left brackets is unbalanced, please check
+-   line 125: the number of right and left brackets is unbalanced,
+    please check
 
-*We have provided an improved description of a closure; the logic behind this line of code should be clearer now*
+*We have provided an improved description of a closure; the logic behind
+this line of code should be clearer now*
 
-- line 126 “This structure was chosen to avoid the repeated specifications of ages and years”. As far as I can see the example code only specifies years, not ages.
+-   line 126 “This structure was chosen to avoid the repeated
+    specifications of ages and years”. As far as I can see the example
+    code only specifies years, not ages.
 
-*Again, we hope that our improved description of a closure will clarify what we mean by this.*
+*Again, we hope that our improved description of a closure will clarify
+what we mean by this.*
 
-- line 227 Please explain what a pipe is and how it is used.
-In the example I understand that the output of sim_abundance( ) is provided to (piped) into sim_distribution().
-I am unclear what the object b contains.
-Is it the result of sim_distribution()?
+-   line 227 Please explain what a pipe is and how it is used. In the
+    example I understand that the output of sim\_abundance( ) is
+    provided to (piped) into sim\_distribution(). I am unclear what the
+    object b contains. Is it the result of sim\_distribution()?
 
-*We have clarified how a pipe works, noting that it forwards values from one function call to the next function call, and we now state that the output from the two examples provided (nested approach vs.
-pipe approach) are functionally the same though the approach is slightly different.*
+*We have clarified how a pipe works, noting that it forwards values from
+one function call to the next function call, and we now state that the
+output from the two examples provided (nested approach vs. pipe
+approach) are functionally the same though the approach is slightly
+different.*
 

### 2019-12-29 analysis/paper/sim_survey_paper.Rmd
Complete rev1 for PLOS ONE

@@
 output: 
-  bookdown::word_document2:
-    reference_docx: template.docx
   bookdown::html_document2:
@@
     toc_float: true
+  bookdown::word_document2:
+    reference_docx: template.docx
 bibliography: references.bib

### 2020-02-05 analysis/paper/sim_survey_paper.Rmd
Start revision given latest reveiw; add paragraph to intro

--- 2019-12-29 analysis/paper/sim_survey_paper.Rmd
@@
 # Introduction
+
+Good survey design is a critical prerequisite for obtaining accurate and precise estimates of a population of interest.
+The need for sound sampling techniques is as relevant today it was in the 1930's when much of the theory and principles behind the collection of scientific data were laid by statisticians such as R.A.
+Fisher and Jerzy Neyman [@fienberg1996].
+Then and now, we are experiencing a period of increasing human activity and robust methods are required to detect the impacts of our actions on the natural world [@nichols2006].
+Survey quality is especially important for conservation-oriented science as monitoring data are used to identify species at risk of extinction [@martin2007], to assess the efficacy of recovery plans [@campbell2002], and to determine precautionary levels of exploitation [@sutherland2001].
+Though large-scale surveys require careful planning and are costly to implement [@reynolds2011], they are not easily replaced by other approaches.
+For instance, numerous attempts to infer population size using catch per unit effort (CPUE) data from commercial or recreational fisheries have been froth with bias because CPUE and abundance are rarely proportional [@harley2001; @maunder2006].
+As such, fisheries-independent surveys have become the preferred source of information as designs with a solid basis in sampling theory can be imposed.
+
+<!-- consider keeping last sentence above general then suegue in the below paragraph to fisheries surveys somehow... -->
 

### 2020-02-05 analysis/paper/sim_survey_paper.Rmd
Revise first two paragraphs of intro -> new draft of intro

--- 2019-12-29 analysis/paper/sim_survey_paper.Rmd
@@
 
-Fisheries-independent surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies.
-While costly to obtain, this information forms the basis of many stock assessments throughout the world [@Pennington1998] and the quality of this information depends on surveys and analyses that maximize information while minimizing the expense of data collection.
-While simulations provide a platform for exploring solutions to this optimization problem, building the necessary simulation framework is not a trivial task given the multi-stage nature of the typical sampling program and the complexity of the population processes we aim to represent.
-Notably, many fish stocks show size-specific patterns of aggregation, making it difficult to collect sufficient biological sub-samples (e.g.
-lengths and ages) to represent the entire target population [@Pennington1994; @Pennington2002; @Stewart2014].
-This undermines the assumptions of many analyses as sub-samples from such populations tend to exhibit strong positive intracluster correlation [@Aanes2015].
-These complexities may explain why simulations that test the full sampling and analytical pathway of fisheries-independent surveys are rare [but see @puerta2019; @schnute2003].
+Good survey design is a critical prerequisite for obtaining accurate and precise estimates of a population of interest.
+The need for sound sampling techniques is as relevant today as it was in the 1930's when much of the theory and principles behind the collection of scientific data were laid by statisticians such as R.A.
+Fisher and Jerzy Neyman [@fienberg1996].
+Then and now, we are experiencing a period of increasing human activity and robust methods are required to detect the impacts of our actions on the natural world [@nichols2006].
+Survey quality is especially important for conservation-oriented science as monitoring data are used to identify species at risk of extinction [@martin2007], to assess the efficacy of recovery plans [@campbell2002], and to determine precautionary levels of exploitation [@sutherland2001].
+Surveys with a solid basis in sampling theory have therefore become a core component of many ecological monitoring programs throughout the world.
+
+One of the biggest challenges associated with surveys is the costs of planning and implementing effective and efficient sampling designs [@reynolds2011].
+Despite the costs, surveys have become a mainstay in the management of dynamic fish stocks as survey data are not easily replaced by other sources of information [@Pennington1998].
+For instance, large population declines can be overlooked by relying solely on catch and effort data from commercial fisheries [@rose1999].
+Such difficulties have contributed to the growing reliance on fisheries-independent surveys to monitor fish populations.
+Planning or improving such surveys is an important but non-trivial task because of the inherent complexity of both the survey and target population.
+Specifically, fisheries-independent surveys usually involve multiple levels of sampling and fish populations often show size-structured spatial and temporal dynamics [@Nelson2014].
+These features are not easily captured by simple statistical models, which makes it difficult to determine optimal sampling and sub-sampling schemes [@Pennington1994; @Pennington2002; @Stewart2014].
+In such situations it is common for practitioners to resort to simulations as a means of testing the efficacy of various sampling designs (e.g.
+[@thomas2010]), however, the complexities associated with fisheries-independent surveys may explain why simulations that test the full sampling and analytical pathway are rare (but see [@puerta2019; @schnute2003]).
 
@@
 Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and a two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years.
-The rate at which point-to-point spatial correlation decays with distance is controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
+The rate at which point-to-point spatial correlation decays with distance is controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) (here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015]) and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
 The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see **[S2 Appendix][S2 Appendix: Age-year-space covariance]** for a more detailed description of the space-age-year covariance structure).
@@
 
-While there are many model-based options for obtaining an abundance index from survey data [e.g.
-@Thorson2015], design-based approaches, such as stratified analyses, are often used.
+While there are many model-based options for obtaining an abundance index from survey data (e.g.
+[@Thorson2015]), design-based approaches, such as stratified analyses, are often used.
 Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in **[S3 Appendix][S3 Appendix: Stratified analysis equations]**] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
@@
 Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete.
-Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe [`%>%`; @bache2014] friendly such that values from one function call are forwarded to the next function (i.e.
+Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe (`%>%`; [@bache2014]) friendly such that values from one function call are forwarded to the next function (i.e.
 output from the two calls below are functionally the same though the approach is slightly different).
@@
 
-The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formula for stratified-random designs [e.g.
-@Cochran1977].
-Nevertheless, there has long been interest in using model-based approaches to improve abundance estimates [e.g.
-@smith1990; @Berg2014; @Thorson2015].
+The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formula for stratified-random designs (e.g.
+[@Cochran1977]).
+Nevertheless, there has long been interest in using model-based approaches to improve abundance estimates (e.g.
+[@smith1990; @Berg2014; @Thorson2015]).
 **`SimSurvey`** can serve as a convenient tool for simulation testing mean and variance estimates provided by a range of different approaches (design-based analyses, bootstrap estimates, generalized additive models, geostatistical models, etc.).
@@
 One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012].
-Results from the case study (see **[S1 Appendix][S1 Appendix: Case study]**) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem [e.g.
-@Berg2012].
+Results from the case study (see **[S1 Appendix][S1 Appendix: Case study]**) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem (e.g.
+[@Berg2012]).
 
@@
 Ideally, the simulation would also include multiple species with different life-histories and distributions, and also integrate a cost component to assess the trade-offs between information and cost.
-Finally, it would be interesting to test alternate analyses of these data that may account for the spatial structure of the age-length key [e.g.
-@Berg2012].
+Finally, it would be interesting to test alternate analyses of these data that may account for the spatial structure of the age-length key (e.g.
+[@Berg2012]).
 
@@
 
-Standard notation for use in the analysis of stratified-random survey data [modified from @Smith1981]:
+Standard notation for use in the analysis of stratified-random survey data (modified from [@Smith1981]):
 

### 2020-02-06 R/sim_pop.R
Export some real and simulated data for Jin

@@ -65,7 +65,7 @@ sim_R <- function(log_mean = log(30000000), log_sd = 0.5, random_walk = TRUE, pl
 
 #' @export
 #' @rdname sim_R
-sim_Z <- function(log_mean = 0.5, log_sd = 0.2, phi_age = 0.9, phi_year = 0.5, plot = FALSE) {
+sim_Z <- function(log_mean = log(0.5), log_sd = 0.2, phi_age = 0.9, phi_year = 0.5, plot = FALSE) {
   function(ages = NULL, years = NULL) {
 
     na <- length(ages)

### 2020-02-06 analysis/exports_for_Jin.R

@@ -1,0 +1,89 @@
+
+library(SimSurvey)
+library(data.table)
+
+## Simulate cod-like population
+## See "imitate_cod_data.R" file for details on parameter choices
+
+set.seed(438)
+pop <- sim_abundance(ages = 1:20,
+                     years = 1:20,
+                     R = sim_R(log_mean = log(30000000),
+                               log_sd = 0.5,
+                               random_walk = TRUE),
+                     Z = sim_Z(log_mean = log(0.5),
+                               log_sd = 0.2,
+                               phi_age = 0.9,
+                               phi_year = 0.5),
+                     growth = sim_vonB(Linf = 120, L0 = 5, K = 0.1,
+                                       log_sd = 0.1, length_group = 3,
+                                       digits = 0)) %>%
+  sim_distribution(grid = make_grid(x_range = c(-140, 140),
+                                    y_range = c(-140, 140),
+                                    res = c(3.5, 3.5),
+                                    shelf_depth = 200,
+                                    shelf_width = 100,
+                                    depth_range = c(0, 1000),
+                                    n_div = 1,
+                                    strat_breaks = seq(0, 1000, by = 40),
+                                    strat_splits = 2),
+                   ays_covar = sim_ays_covar(sd = 2.8,
+                                             range = 300,
+                                             phi_age = 0.5,
+                                             phi_year = 0.9,
+                                             group_ages = 5:20),
+                   depth_par = sim_parabola(mu = 200,
+                                            sigma = 70))
+
+
+
+## ~ current sampling protocol
+base_case <- sim_survey_parallel(pop, n_sims = 5, n_loops = 5, cores = 6,
+                                 set_den = 2e-03, lengths_cap = 500, ages_cap = 10,
+                                 q = sim_logistic(k = 2, x0 = 3),
+                                 quiet = FALSE)
+
+setdet <- base_case$setdet
+samp <- base_case$samp
+fwrite(setdet, file = "analysis/cod_sim_exports/2020-02-06_for_Jin/sim_cod_3Ps_setdet.csv")
[…]

### 2020-02-06 analysis/paper/sim_survey_paper.Rmd
Start structural changes to paper: modify headers

@@
 
-# Model structure
+# Methods
+
+## Model structure
 
@@
 
-## Simulate abundance
+### Simulate abundance
 
@@
 
-## Simulate spatial distribution
+### Simulate spatial distribution
 
@@
 
-## Simulate survey
+### Simulate survey
 
@@
 
-## Stratified analysis
+### Stratified analysis
 
@@
 
-## Assumptions
+### Assumptions
 
@@
 
-# Using **`SimSurvey`**
+## Core functions
 
@@
 
-## `sim_abundance`
+### `sim_abundance`
 
@@
 
-## `sim_distribution`
+### `sim_distribution`
 
@@
 
-## `sim_survey`
+### `sim_survey`
 
@@
 
-## `run_strat`
+### `run_strat`
 
@@
 
-## `test_surveys`
+### `test_surveys`
 
@@
 
-# Parameterisation
+## Parameterisation
 
@@
 
-## 1. Define survey area and protocol
+### 1. Define survey area and protocol
 
@@
 
-## 2. Define abundance and growth parameters
+### 2. Define abundance and growth parameters
 
@@
 
-## 3. Define spatial parameters
+### 3. Define spatial parameters
 
@@
 
-# Research opportunities
+# Results
+
+CONSIDER INCLUDING AN ABSTRACT OF THE CASE STUDY RESULTS, MOSTLY AS A MEANS OF INTREPRETING THE PLOTS. PREVIOUS ABSTRACT BELOW:
+
+A central question to fish stock assessments is: how many samples are needed to adequately represent the population being monitored?
+For assessments based on trawl surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling (i.e.
+the number of sets, lengths and ages sampled in a survey).
+We therefore need to assess the impact of haul design and length and age sampling on survey accuracy to determine optimal sampling effort.
+In this study we simulated repeated stratified-random samples from a spatially correlated field of simulated fish.
+These samples were then analyzed and estimates of abundance at age were compared against a known truth.
+By varying the sampling protocol, we demonstrate that increasing the number of sets conducted per stratum provides the greatest gains in precision when sampling a clustered population.
+Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent.
+These results indicate that length and age sampling efforts for some species(e.g.
+cod *Gadus morhua*) in trawl surveys conducted by regional fisheries management organizations can be scaled back without significant losses to precision.
+Time saved by reducing said sampling may afford more time to conduct additional sets and/or increase sampling efforts of other species.
+While this simulation is tailored to a specific case, the method may be adapted to different systems and survey procedures.
+
+# Discussion
+
+GENERAL
+
+## Research opportunities
 
@@
 
-## Design or model-based approach
+### Design or model-based approach
 
@@
 
-## Growth analyses
+### Growth analyses
 
@@
 
-## Random or stratified sampling
+### Random or stratified sampling
 
@@
 
-# Future directions
+## Future directions
 
@@
 
-# Summary
+## Summary
 

### 2020-02-06 R/sim_survey.R
Add argument allowing age sampling to be conducted at the strat spatial scale

@@ -113,12 +113,18 @@ sim_sets <- function(sim, n_sims = 1, trawl_dim = c(1.5, 0.02),
 #'                            \code{resample_cells = FALSE} because the number of sets allocated may
 #'                            exceed the number of cells in a strata.
 #' @param lengths_cap         Maximum number of lengths measured per set
-#' @param length_group        Length group for otolith collection
-#' @param ages_cap            If \code{age_sampling = "length stratified"}, this cap represents the
-#'                            maximum number of otoliths to collect per length group
-#'                            per division per year. If \code{age_sampling = "random"}, it is the
-#'                            maximum number of otoliths to collect from measured fish per set.
-#' @param age_sampling        Should age sampling be length "stratified" or "random"?
+#' @param ages_cap            If \code{age_sampling = "stratified"}, this cap represents the maximum
+#'                            number of ages to sample per length group (defined using the \code{age_length_group}
+#'                            argument) per division or strat (defined using the \code{age_space_group} argument)
+#'                            per year. If \code{age_sampling = "random"}, it is the maximum number of ages to sample
+#'                            from measured fish per set.
+#' @param age_sampling        Should age sampling be "stratified" (default) or "random"?
+#' @param age_length_group    Numeric value indicating the size of the length bins for stratified
+#'                            age sampling. Ignored if \code{age_sampling = "random"}.
+#' @param age_space_group     Should age sampling occur at the "division" (default) or "strat" spatial scale?
+#'                            That is, age sampling can be spread across each "division" or "strat"
+#'                            in each year to a maximum number within each length bin (cap is defined using
+#'                            the \code{age_cap} argument). Ignored if \code{age_sampling = "random"}.
 #' @param light               Drop some objects from the output to keep object size low?
 #'
 #' @return A list including rounded population simulation, set locations and details
@@ -138,7 +144,8 @@ sim_sets <- function(sim, n_sims = 1, trawl_dim = c(1.5, 0.02),
 sim_survey <- function(sim, n_sims = 1, q = sim_logistic(), trawl_dim = c(1.5, 0.02),
                        resample_cells = FALSE, binom_error = TRUE,
                        min_sets = 2, set_den = 2 / 1000, lengths_cap = 500,
[…]

### 2020-02-06 analysis/cod_sim.R

@@ -7,10 +7,10 @@ library(SimSurvey)
 set.seed(438)
 pop <- sim_abundance(ages = 1:20,
                      years = 1:20,
-                     R = sim_R(mean = 30000000,
+                     R = sim_R(log_mean = log(30000000),
                                log_sd = 0.5,
                                random_walk = TRUE),
-                     Z = sim_Z(mean = 0.5,
+                     Z = sim_Z(log_mean = log(0.5),
                                log_sd = 0.2,
                                phi_age = 0.9,
                                phi_year = 0.5),
@@ -72,10 +72,10 @@ gc()
 set.seed(438)
 pop <- sim_abundance(ages = 1:20,
                      years = 1:20,
-                     R = sim_R(mean = 30000000,
+                     R = sim_R(log_mean = log(30000000),
                                log_sd = 0.5,
                                random_walk = TRUE),
-                     Z = sim_Z(mean = 0.5,
+                     Z = sim_Z(log_mean = log(0.5),
                                log_sd = 0.2,
                                phi_age = 0.9,
                                phi_year = 0.5),
@@ -123,5 +123,32 @@ setMKLthreads() # turn hyperthreading on again
 # ## visualize results
 # load("analysis/cod_sim_exports/2018-10-28_no_age_clust_test/test_output.RData")
 # vis_sim(sim)
+
+
+
+## Test survey with strat-specific age-length-keys --------------------------------------
+
+library(SimSurvey)
+
+set.seed(438)
+pop <- sim_abundance() %>%
+  sim_distribution()
+
+
+default_survey <- pop %>%
+  sim_survey(n_sims = 10)
+
+alt_survey <- pop %>%
+  sim_survey(n_sims = 10, ages_cap = 5, age_length_group = 5,
+             age_space_group = "strat")
+
+
+
+
+
+
+
+
+
 
 

### 2020-02-08 analysis/paper/sim_survey_paper.Rmd
Revise paper up to the Core functions section (add details on the age sampling and analysis changes)

@@
 The steps taken to simulate surveys of spatial, age-structure populations are outlined below.
-First we outline the equations underlying the package in the **[Model structure]** section and then move into a vignette of the package under the **[Using **`SimSurvey`**]** section.
+Under the **[Methods]** section, we first outline the equations underlying the package (**[Model structure]** section) and then move into a vignette of the package (**[Core functions]** section).
 While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study.
-Output from default function calls are therefore relevant to the case study and these results are described and discussed in **[S1 Appendix][S1 Appendix: Case study]**.
+Output from default function calls are therefore relevant to the case study and some of these results are highlighted in the **[Results]** and **[Discussion]** section; a more detailed description and discussion of these results can be found in **[S1 Appendix][S1 Appendix: Case study]**.
 
@@
 Though there is a degree of flexibility in each of these steps, users can circumvent specific components by applying user defined equations, inputs and/or analyses.
-Details on how to use the package and, if desired, circumvent some aspects of its structure are outlined in the **[Using **`SimSurvey`**]** section.
+Details on how to use the package and, if desired, circumvent some aspects of its structure are outlined in the **[Core functions]** section.
 
@@
 Parameter estimates from  spatially-aggregated age-structured models, which are commonly used in stock assessments, can therefore be used to simulate a population using the cohort model and the resultant abundance at age values can be distributed across a spatial grid.
-Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across one spatial axis, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
+Here, a grid of $s$ cells is generated where each cell has an area of $S$ and depth $d$; depth is defined using a sigmoid curve, applied across one spatial axis, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$.
 We use depth as our main stratification variable, but note that any other appropriate stratification variables could be used.
@@
 Sub-sampling is then conducted whereby a subset of fish are sampled for length measurements and a subset of this subset are sampled for age determination.
-Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division.
+Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division, strata or set.
 Such sub-sampling is common in fisheries-independent surveys as it is costly, impractical and unnecessary to sample every fish captured.
@@
 [@Thorson2015]), design-based approaches, such as stratified analyses, are often used.
-Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in **[S3 Appendix][S3 Appendix: Stratified analysis equations]**] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$).
-Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be scaled up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key.
+Here we apply formula presented in Smith and Somerton [-@Smith1981] (equations are replicated in **[S3 Appendix][S3 Appendix: Stratified analysis equations]**) to calculate year $y$ and simulation $j$ specific stratified estimates of total abundance ($\hat{I}_{y,j}$), abundance at length ($\hat{I}_{l,y,j}$) and abundance at age ($\hat{I}_{a,y,j}$).
+Note that estimates of total abundance are based on total numbers caught at each set, $n_i$, while abundance at length requires the sub-sampled length frequencies at each set, $m_{l, i}$, to be scaled up using set-specific ratios of the number of fish measured, $m_{i}$, to numbers caught, $n_i$:
+
+$$n_{l, i} = m_{l, i} \times \frac{n_i}{m_i}$$
+
+Likewise, age frequencies need to be calculated to obtain stratified estimates of abundance at age.
+This is done by constructing an age-length key, which is the proportion of fish in each length bin that fall into specific ages classes, and applying this key to the bumped up length frequencies, $n_{l,i}$, to approximate age frequencies, $n_{a,i}$, that is:
+
+$$n_{a,i} = \sum_{l = 1}^{L} n_{l,i} \times \frac{m^{\prime}_{a,l,g}}{m^{\prime}_{l,g}}$$
+
+Here, the prime symbol $^{\prime}$ indicates that these values are tertiary sampling units as they represent fish sub-sampled for age determination from those sub-sampled for length measurements.
+Within this sub-sample, $m^{\prime}_{l,g}$ represents the number of fish within length bin $l$ and from area $g$, and $m^{\prime}_{a,l,g}$ represents the number of fish within age-class $a$, length bin $l$ and area $g$.
+The area $g$ can be division, strata or set specific.
+Though age-length keys are typically constructed and applied over large spatial scales (e.g.
+division), calculating proportions at finer scales may better account for the hierarchical sample designs [@Aanes2015].
+
 We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:
 
-$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})^2}{N_a N_y N_k}}$$
-
-Where $N_a$, $N_y$, and $N_k$ are the number of ages, years and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e.
+$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{A} \sum_{y = 1}^{Y} \sum_{j = 1}^{J} (\hat{I}_{a,y,j} - I_{a,y})^2}{A \times Y \times J}}$$
+
+Where $A$, $Y$, and $J$ are the number of ages, years and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e.
 catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$).

### 2020-02-10 analysis/paper/sim_survey_paper.Rmd
Add details on the age sampling and analysis changes to the Core functions section

@@
 Sub-sampling is then conducted whereby a subset of fish are sampled for length measurements and a subset of this subset are sampled for age determination.
-Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division, strata or set.
+Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division, strata or set $s_{\mathrm{group}}$.
 Such sub-sampling is common in fisheries-independent surveys as it is costly, impractical and unnecessary to sample every fish captured.
@@
 Though age-length keys are typically constructed and applied over large spatial scales (e.g.
-division), calculating proportions at finer scales may better account for the hierarchical sample designs [@Aanes2015].
+division), Aanes and Vølstad [@Aanes2015] recommends calculating proportions at finer scales to better account for hierarchical sample designs.
 
@@
 | `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_L$                          |
-| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{group}^{N}$                     |
+| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{\mathrm{group}}^{N}$            |
 
@@
 | **Function call**                   | **Description**                                                             | **Symbol**              |
-|:----------------------------------- |:----------------------------------------------------------------------- | :---------------------- |
+|:----------------------------------- |:--------------------------------------------------------------------------- | :---------------------- |
 | `sim_survey(`                       | &nbsp;                                                                      | &nbsp;                  |
@@
 | `+  lengths_cap = 500,`             | Maximum number of lengths to collect / set                                  | $M_{\mathrm{lengths}}$  |
-| `+  length_group = 1,`              | Length group bin size for age sampling (cm)                             | $l_{\mathrm{group}}$    |
-| `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / division              | $M_{\mathrm{ages}}$     |
+| `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / spatial group ^1^         | $M_{\mathrm{ages}}$     |
 | `+  age_sampling = "stratified")`   | Controls whether age sampling is length `"stratified"` or `"random"`        | &nbsp;                  |
+| `+  age_length_group = 1)`          | Length group bin size for stratified age sampling (cm)                      | $l_{\mathrm{group}}$    |
+| `+  age_space_group = "division")`  | Spatial scale of stratified age sampling (`"division"`, `"strat"`, `"set"`) | $s_{\mathrm{group}}$    |
+
+^1^ Length group and spatial group are defined using the `age_length_group` and `age_space_group` arguments, respectively.
+These arguments are ignored if `age_sampling` is set to `"random"` and the value supplied to `ages_cap` represents the maximum number of ages to sample per set.
 
@@
 Stratified estimates of abundance are obtained by supplying the output from `sim_survey` to the `run_strat` function.
-RMSE of the stratified estimates can then be calculated using the `strat_error` function.
+There are only four arguments in the `run_strat` function: `length_group`, `alk_scale`, `strat_data_fun` and `strat_means_fun`.
+The `length_group` argument defines the size of the length frequency bins for both abundance at length calculations and age-length key construction; this argument is set to `"inherit"`, by default, to utilize the `length_group` value defined inside the closure supplied to the `growth` argument in `sim_abundance`.
+The `alk_scale` argument defines the scale at which to construct and apply age-length keys (`"division"` [default], `"strat"` or `"set"`).
+RMSE of the stratified estimates from `run_strat` can then be calculated using the `strat_error` function.
 Results and error of a stratified analysis of one survey over a population are obtained using the following code (using default values):

### 2020-02-12 analysis/paper/sim_survey_paper.Rmd
Start to work in a short results/discussion section for the case study into the body of the paper

@@
 
-# Methods
-
-## Model structure
+# Model structure
 
@@
 
-### Simulate abundance
+## Simulate abundance
 
@@
 
-### Simulate spatial distribution
+## Simulate spatial distribution
 
@@
 
-### Simulate survey
+## Simulate survey
 
@@
 
-### Stratified analysis
+## Stratified analysis
 
@@
 
-### Assumptions
+## Assumptions
 
@@
 
-## Core functions
+# Core functions
 
@@
 
-### `sim_abundance`
+## `sim_abundance`
 
@@
 
-### `sim_distribution`
+## `sim_distribution`
 
@@
 
-### `sim_survey`
+## `sim_survey`
 
@@
 
-### `run_strat`
+## `run_strat`
 
@@
 
-### `test_surveys`
+## `test_surveys`
 
@@
 True values of abundance available to the survey are overlaid on the series of probability envelopes.
-These plots help visually assess the level of precision and bias from a specific set of survey protocol.
+These plots help visually assess the level of precision and bias from a specific set of survey protocol; ideally, the probability envelopes will be tightly centered around the true values.
 The three lines of code below will produce interactive fan charts for stratified estimates of total abundance, abundance at length and abundance at age, respectively (e.g.
@@
 
-```{r fan3, out.height = "600px", fig.cap = "Fan chart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{\\mathrm{sets}}$, and the maximum number of length samples, $M_{\\mathrm{lengths}}$.
+```{r fan3, out.height = "600px", fig.cap = "Fan chart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{\\mathrm{sets}}$, and the maximum number of length samples, $M_{\\mathrm{lengths}}$.
 Number of ages sampled per length group, $M_{\\mathrm{ages}}$, was 10 in all scenarios.
@@
 The relative performance of the surveys tested can be compared using `plot_survey_rank` and `plot_error_surface`.
-The `plot_survey_rank` function produces a divergent dot plot of the results which ranks the surveys by RMSE.
+The `plot_survey_rank` function produces a divergent dot plot of the results which ranks the surveys by RMSE, where lower values indicate sampling strategies that minimize bias and maximize precision.
 Using the `which_strat` argument, the plot can be focused on total, length or age based stratified results (Figure \@ref(fig:rank)).
@@
 
-## Parameterisation
+# Interpretation
+
+Results shown in the **[Core functions]** section demonstrates one use case of the **`SimSurvey`** package that is focused on evaluating the efficacy of increasing the sampling effort of a stratified random survey of a cod population (see **[S1 Appendix][S1 Appendix: Case study]** for details).
+These results largely align with expectations from sampling theory that 1) design-based estimators for stratified random surveys are unbiased, and 2) precision is increased by increasing the number of primary sampling units [@Cochran1977].
+Specifically, estimates of total abundance and abundance at length are centered around true values and their probability envelopes tighten as set density increases (Figures \@ref(fig:fan1), \@ref(fig:fan2)).
+Case study results also echo the growing body of literature which concludes that extra sub-sampling is an ineffective means of improving estimates relative to sampling more locations [@Pennington1994; @Pennington2002; @Stewart2014; @Bogstad1995; @Coggins2013; @Zhang2013].
+This is exemplified by the relatively large drops in RMSE when set density is increased compared to when sub-sampling effort is increased (Figures \@ref(fig:rank), \@ref(fig:rmse-surface)).
+Moreover, it appears that it is more advantageous to measure fewer total fish at more locations than measuring many fish at fewer locations (Figure \@ref(fig:rank)).
+Again, this result was expected because the size-structured spatial clustering imposed by the simulation causes set-to-set variation to be greater than within-set variation in characteristics such as length and age.
+This structure was imposed because fish caught together tend to have similar characteristics and it is well known that this reduces the effective sample size [@Pennington1994].
+Though ineffective, increased sub-sampling effort should never lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)).
+Bias, in particular, appears to be a problem with stratified estimates of abundance at age under certain scenarios (Figure \@ref(fig:fan3)); in contrast, there appears to be little bias in the stratified estimates of abundance at length (Figure \@ref(fig:fan2)).
+This result was unexpected as the design-based analysis was presumed to be unbiased.
+The contrast between the length and age based analysis, however, indicates that the problem lies with the intervening age-length-key.
+
+Age-length-key...large scale typical...danger, assumptions...segue to solution...
+
+# Parameterisation
 
@@
 
-### 1. Define survey area and protocol
+## 1. Define survey area and protocol
 
@@
 
-### 2. Define abundance and growth parameters
+## 2. Define abundance and growth parameters
 
@@
 
-### 3. Define spatial parameters
+## 3. Define spatial parameters
 
@@
 Finally, visual inspection of the real data revealed that the largest catches tended to be centered on sets conducted in 200 m of water; a parabolic relationship, defined using the `sim_parabola` closure, with a mean of 200 m and standard deviation of 70 m approximated this pattern.
-
-# Results
-
-CONSIDER INCLUDING AN ABSTRACT OF THE CASE STUDY RESULTS, MOSTLY AS A MEANS OF INTREPRETING THE PLOTS. PREVIOUS ABSTRACT BELOW:
-
-A central question to fish stock assessments is: how many samples are needed to adequately represent the population being monitored?
-For assessments based on trawl surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling (i.e.
-the number of sets, lengths and ages sampled in a survey).
-We therefore need to assess the impact of haul design and length and age sampling on survey accuracy to determine optimal sampling effort.
-In this study we simulated repeated stratified-random samples from a spatially correlated field of simulated fish.
-These samples were then analyzed and estimates of abundance at age were compared against a known truth.
-By varying the sampling protocol, we demonstrate that increasing the number of sets conducted per stratum provides the greatest gains in precision when sampling a clustered population.
-Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent.
-These results indicate that length and age sampling efforts for some species(e.g.
-cod *Gadus morhua*) in trawl surveys conducted by regional fisheries management organizations can be scaled back without significant losses to precision.
-Time saved by reducing said sampling may afford more time to conduct additional sets and/or increase sampling efforts of other species.
-While this simulation is tailored to a specific case, the method may be adapted to different systems and survey procedures.
 

### 2020-02-12 analysis/cod_sim.R

@@ -149,7 +149,7 @@ vis_sim(alt_survey)
 ## Test alternate survey with strat specific age sampling and age-length-keys
 surveys <- expand_surveys(set_den = c(0.5, 1, 2, 5, 10) / 1000,
                           lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
-                          ages_cap = c(2, 5, 10, 20, 50))
+                          ages_cap = c(1, 2, 5, 10, 20))
 sim <- test_surveys(pop,
                     surveys = surveys,
                     keep_details = 1,
@@ -168,4 +168,41 @@ vis_sim(sim)
 
 plot_age_strat_fan(sim)
 
+plot_survey_rank(sim, which_strat = "length")
+
+
+
+
+## 3 cm bins would be more comprable...
+
+set.seed(438)
+pop <- sim_abundance() %>%
+  sim_distribution()
+
+
+## Test alternate survey with strat specific age sampling and age-length-keys
+surveys <- expand_surveys(set_den = 2 / 1000,
+                          lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
+                          ages_cap = c(1, 2, 3, 5, 10))
+sim <- test_surveys(pop,
+                    surveys = surveys,
+                    keep_details = 1,
+                    n_sims = 5,
+                    n_loops = 200,
+                    cores = 7,
+                    export_dir = "analysis/cod_sim_exports/2020-02-10_strat_alk",
+                    age_length_group = 3,
+                    age_space_group = "set",
+                    alk_scale = "set")
+# sim <- resume_test(export_dir = "analysis/cod_sim_exports/2020-02-10_strat_alk")
+
+## TODO: perhaps run sim_survey_parallel and get fan plots working for that ouptut?
+
+vis_sim(sim)
+
+sim$surveys
+
+plot_age_strat_fan(sim, surveys = 5, ages = 1:10, years = 7)
+
+plot_survey_rank(sim, which_strat = "length")
 

### 2020-02-12 analysis/cod_sim.R
Tidy up code used to generate set_alk simulation results

@@ -122,63 +122,13 @@ sim <- test_surveys(pop,
 
 
 
-## Test survey with strat-specific age-length-keys --------------------------------------
+## Test survey with set-specific age-length-keys --------------------------------------
 
 library(SimSurvey)
 
 set.seed(438)
-pop <- sim_abundance(growth = sim_vonB(length_group = 5)) %>%
-  sim_distribution()
-
-
-default_survey <- pop %>%
-  sim_survey(n_sims = 5)
-
-sum(default_survey$samp$aged)
-
-alt_survey <- pop %>%
-  sim_survey(n_sims = 5, lengths_cap = 100, ages_cap = 5,
-             age_length_group = 5, age_space_group = "set") %>%
-  run_strat(alk_scale = "set")
-
-sum(alt_survey$samp$aged)
-
-vis_sim(alt_survey)
-
-
-## Test alternate survey with strat specific age sampling and age-length-keys
-surveys <- expand_surveys(set_den = c(0.5, 1, 2, 5, 10) / 1000,
-                          lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
-                          ages_cap = c(1, 2, 5, 10, 20))
-sim <- test_surveys(pop,
-                    surveys = surveys,
-                    keep_details = 1,
-                    n_sims = 5,
-                    n_loops = 200,
-                    cores = 7,
-                    export_dir = "analysis/cod_sim_exports/2020-02-06_strat_alk",
-                    age_length_group = 5,
-                    age_space_group = "set",
-                    alk_scale = "set")
-# sim <- resume_test(export_dir = "analysis/cod_sim_exports/2020-02-06_strat_alk")
-
-## TODO: perhaps run sim_survey_parallel and get fan plots working for that ouptut?
-
-vis_sim(sim)
-
-plot_age_strat_fan(sim)
-
-plot_survey_rank(sim, which_strat = "length")
-
-
-
-
-## 3 cm bins would be more comprable...
-
-set.seed(438)
 pop <- sim_abundance() %>%
   sim_distribution()
-
 
 ## Test alternate survey with strat specific age sampling and age-length-keys
 surveys <- expand_surveys(set_den = 2 / 1000,
@@ -190,11 +140,11 @@ sim <- test_surveys(pop,
                     n_sims = 5,
                     n_loops = 200,
                     cores = 7,
-                    export_dir = "analysis/cod_sim_exports/2020-02-10_strat_alk",
[…]

### 2020-02-12 analysis/paper/sim_survey_paper.Rmd
Keep working on the text in the interpretation section and add a plot from new simulation results

@@
 The `alk_scale` argument defines the scale at which to construct and apply age-length keys (`"division"` [default], `"strat"` or `"set"`).
+Finally, `strat_data_fun` and `strat_means_fun` allow users to supply custom functions for processing the simulated data and calculating stratified means, respectively; the built in functions `strat_data` and `strat_means` are supplied by default.
 RMSE of the stratified estimates from `run_strat` can then be calculated using the `strat_error` function.
@@
 To this, `strat_error` adds data.tables ending with `_strat_error` or `_strat_error_stats`.
-The `_strat_error` objects simply contain stratified estimates of abundance (column named `I_hat`) with corresponding true values of abundance available to the survey (column named `I`) and the `strat_error_stats` data.frame includes metrics of mean absolute error (`MAE`), mean-squared error (`MSE`) and root-mean-squared error (`RMSE`).
+The `_strat_error` objects simply contain stratified estimates of abundance (column named `I_hat`) with corresponding true values of abundance available to the survey (column named `I`) and the `strat_error_stats` data.frame includes metrics of mean error (`ME`), mean absolute error (`MAE`), mean-squared error (`MSE`) and root-mean-squared error (`RMSE`).
 
@@
 all possible combinations of the `set_den`, `lengths_cap` and `ages_cap` vectors) and the `test_surveys` function will run 1000 simulations of each survey and compare stratified estimates of abundance to the true abundance available to the survey.
+While `test_surveys` is set-up for testing key sampling effort settings (`set_den`, `lengths_cap` and `ages_cap`), other options can be assessed using independent calls of `test_surveys` as it accepts several arguments from `sim_survey` (`q`, `trawl_dim`, `min_sets`, `age_sampling`, `age_length_group` and `age_space_group`) and `run_strat` (`length_group` and `alk_scale`).
 
@@
 This structure was imposed because fish caught together tend to have similar characteristics and it is well known that this reduces the effective sample size [@Pennington1994].
-Though ineffective, increased sub-sampling effort should never lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)).
+Increases to the number of sub-samples, however, should not lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)).
 Bias, in particular, appears to be a problem with stratified estimates of abundance at age under certain scenarios (Figure \@ref(fig:fan3)); in contrast, there appears to be little bias in the stratified estimates of abundance at length (Figure \@ref(fig:fan2)).
 This result was unexpected as the design-based analysis was presumed to be unbiased.
-The contrast between the length and age based analysis, however, indicates that the problem lies with the intervening age-length-key.
+The contrast between the length and age based analysis indicates that the problem lies with the intervening age-length-key.
+
+Age-length keys have long been used in conjunction with length frequency tables to estimate the age distribution of fish populations over large scales [@fridriksson1934].
+The default option applied in the package is to construct and apply the key at the division level, as this is the standard protocol for the actual survey the case study is based on.
+There is, however, a potential cost to the spatial scale of the key.
+Namely, it is unlikely that one age-length key is representative for the whole region because the probability of being a specific age given length varies in space [@Berg2012].
+Because different age groups sometimes occur in different places, the translation of lengths to ages may be biased by the samples used to generate the age-length key.
+In short, multi-stage sampling of a clustered population violates the assumption that the age-length key is generated from a simple random sample [@Aanes2015].
+Bias introduced by this assumption can be resolved by explicitly accounting for hierarchical sampling using designed-based estimators [@Aanes2015].
+For the case study, the estimator will need to account for cluster sampling in a stratified random survey.
+Following recommendations in Aanes and Vølstad [@Aanes2015], we used **`SimSurvey`** to 1) conduct a surveys with concurrent length and age sampling at every set, 2) construct and apply age-length keys on a set-by-set basis, and 3) weight the resultant age frequencies at each set using stratified random estimators.
+This test was accomplished using this code:
+
+```{r, echo = TRUE, eval = FALSE}
+
+alt_surveys <- expand_surveys(set_den = 2 / 1000,
+                              lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
+                              ages_cap = c(1, 2, 3, 5, 10))
+
+alt_tests <- test_surveys(pop, surveys = alt_surveys,
+                          n_sims = 5, n_loops = 200, cores = 3,
+                          age_length_group = 3,
+                          age_space_group = "set", 
+                          alk_scale = "set")
+
+```
+
+```{r alt-data, eval = replot | is_html}
+
+load("analysis/cod_sim_exports/2020-02-10_set_alk/test_output.RData")
+alt_tests <- sim
+
+```
+
+This is a minor modification of the code used in the **[`test_surveys`]** section whereby the same simulated population was used (`pop` object) but set density scenarios were reduced to one option, for simplicity, and age sampling protocol and stratified analysis options were changed.
+That is, 1) `age_length_group` and `age_space_group` were changed to simulate a survey that conducts concurrent length and age sampling at every set (e.g.
+collect age samples from a maximum of 1 fish per 3 cm length group per set), 2) `alk_scale` was changed to `"set"` to construct and apply age-length keys on a set-by-set basis, and 3) the resultant age frequencies at each set are weighted accordingly using the design-based estimators built into `run_strat`.
+Again, results are visualized using `plot_age_strat_fan` and `plot_error_surface`:
+
+```{r, echo = TRUE, eval = FALSE}
+
+plot_age_strat_fan(alt_tests, years = 1:20, ages = 1:10)
+plot_error_surface(alt_tests, plot_by = "rule")
+
+```
+
+```{r, eval = replot | is_html}
+
+a <- plot_age_strat_fan(alt_tests, surveys = 19, ages = 1:10, years = 7) %>% 
+  layout(yaxis = list(range = c(0, 51000000))) %>% 
+  add_labels(x = -0.15, y = 1.05, text = "a)")
+
+sim <- alt_tests
+totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught),
+                                 n_measured = mean(n_measured), n_aged = mean(n_aged)),
+                          by = "survey"]
+errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
+d <- merge(errors, totals, by = "survey")
+
+x <- sort(unique(d$ages_cap))
+y <- sort(unique(d$lengths_cap))
+
+scene <- list(
+  xaxis = list(title = "Ages cap"),
+  yaxis = list(title = "Lengths cap"),
+  zaxis = list(title = "RMSE"),
+  camera = list(eye = list(x = 1.5, y = 1.5, z = 1.5))
+)
+
+z <- stats::xtabs(RMSE ~ ages_cap + lengths_cap, data = d, subset = NULL)
+
+b <- plot_ly(x = x, y = y) %>% 
+  add_surface(z = t(z)) %>% 
+  layout(scene = scene) %>% 
+  hide_guides() %>% 
+  add_labels(x = 0.05, y = 1.05,  text = "b)")
+
+p <- subplot(a, b, shareX = FALSE, shareY = FALSE, titleX = TRUE, titleY = TRUE) %>% 
+  layout(xaxis1 = list(domain = list(x = c(0, 0.5), y = c(0, 1))),
+         scene = list(domain = list(x = c(0.5, 1), y = c(0, 1))),
+         autosize = is_html)
+
+if (replot) export_plot(p, file = "analysis/paper/figures/alt_tests_plots.png", width = 800, height = 390)
+
+```
+
+```{r alt-tests, out.height = "600px", fig.cap = "Key results from an alternative design-based sampling and analysis procedure that accounts for multi-stage cluster sampling of ages.
+Specifically, a) is a fan chart of stratified estimates of abundance at age from year seven of the simulation from a survey with a set density ($D_{\\mathrm{sets}}$) of 0.002 sets / km^2^, maximum number of length samples ($M_{\\mathrm{lengths}}$) of 100, and maximum number of age samples per 3 cm length group per set ($M_{\\mathrm{ages}}$) of 3 (the thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%); and b) is a surface plot of RMSE from an array of surveys with different sub-sampling effort, where the x-axes represent the maximum number of lengths to measure per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per 3 cm length group per set ($M_{\\mathrm{ages}}$)."}
+
+if (is_html) {
+  p
+} else {
+  knitr::include_graphics("figures/alt_tests_plots.png")
+}
+
+```
+
+These estimators require concurrent length and age sampling at every primary sampling unit
+
+This bias is perhaps compounded by the length-stratified sampling of ages such that the ages sampled may be skewed towards sets with the most catch, especially under scenarios where length sampling effort is high.
+If this is the case, then the representativeness of the age-length key to the whole population could be diminished by excessive length sampling.
 

### 2020-02-12 analysis/paper/sim_survey_paper.Rmd
Wrap up draft interpretation section and add a sentence to the abstract

@@
 Fish populations often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective surveys.
-Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision and increasing the risk of bias) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics).
+Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics).
 For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of fish captured to measure biological characteristics (e.g.
@@
 **`SimSurvey`** also includes a function for estimating the stratified mean and variance of the population from the simulated survey data.
-**`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring.
+We demonstrate the use of this package using a case study and show that the it can reveal unexpected sources of bias and be used to explore design-based solutions to such problems.
+In summary, **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring.
 Various statistical approaches can then be applied to the results to test the efficacy of different analytical approaches.
@@
 While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study.
-Output from default function calls are therefore relevant to the case study and some of these results are highlighted in the **[Results]** and **[Discussion]** section; a more detailed description and discussion of these results can be found in **[S1 Appendix][S1 Appendix: Case study]**.
+Output from default function calls are therefore relevant to the case study and some of these results are highlighted in the **[Interpretation]** section; a more detailed description and discussion of these results can be found in **[S1 Appendix][S1 Appendix: Case study]**.
 
@@
 This structure was imposed because fish caught together tend to have similar characteristics and it is well known that this reduces the effective sample size [@Pennington1994].
-Increases to the number of sub-samples, however, should not lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)).
+Increases sub-sampling effort, however, should not lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)).
 Bias, in particular, appears to be a problem with stratified estimates of abundance at age under certain scenarios (Figure \@ref(fig:fan3)); in contrast, there appears to be little bias in the stratified estimates of abundance at length (Figure \@ref(fig:fan2)).
-This result was unexpected as the design-based analysis was presumed to be unbiased.
+This result was unexpected as the analysis was presumed to be unbiased.
 The contrast between the length and age based analysis indicates that the problem lies with the intervening age-length-key.
@@
 Age-length keys have long been used in conjunction with length frequency tables to estimate the age distribution of fish populations over large scales [@fridriksson1934].
-The default option applied in the package is to construct and apply the key at the division level, as this is the standard protocol for the actual survey the case study is based on.
+By default, `sim_survey` spreads length-stratified age sampling across the division and `run_strat` constructs and applies age-length keys at the division scale.
+This default was chosen because it is standard protocol for the actual survey the case study is based on.
 There is, however, a potential cost to the spatial scale of the key.
@@
 In short, multi-stage sampling of a clustered population violates the assumption that the age-length key is generated from a simple random sample [@Aanes2015].
-Bias introduced by this assumption can be resolved by explicitly accounting for hierarchical sampling using designed-based estimators [@Aanes2015].
-For the case study, the estimator will need to account for cluster sampling in a stratified random survey.
+Bias introduced by this assumption can be resolved by explicitly accounting for the hierarchical sampling by using designed-based estimators [@Aanes2015].
+For the case study, a proper design-based estimator will need to account for cluster sampling in a stratified random survey.
 Following recommendations in Aanes and Vølstad [@Aanes2015], we used **`SimSurvey`** to 1) conduct a surveys with concurrent length and age sampling at every set, 2) construct and apply age-length keys on a set-by-set basis, and 3) weight the resultant age frequencies at each set using stratified random estimators.
@@
 
-```{r alt-tests, out.height = "600px", fig.cap = "Key results from an alternative design-based sampling and analysis procedure that accounts for multi-stage cluster sampling of ages.
+```{r alt, out.height = "600px", fig.cap = "Key results from an alternative design-based sampling and analysis procedure that accounts for multi-stage cluster sampling of ages.
 Specifically, a) is a fan chart of stratified estimates of abundance at age from year seven of the simulation from a survey with a set density ($D_{\\mathrm{sets}}$) of 0.002 sets / km^2^, maximum number of length samples ($M_{\\mathrm{lengths}}$) of 100, and maximum number of age samples per 3 cm length group per set ($M_{\\mathrm{ages}}$) of 3 (the thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%); and b) is a surface plot of RMSE from an array of surveys with different sub-sampling effort, where the x-axes represent the maximum number of lengths to measure per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per 3 cm length group per set ($M_{\\mathrm{ages}}$)."}
@@
 
-These estimators require concurrent length and age sampling at every primary sampling unit
-
-This bias is perhaps compounded by the length-stratified sampling of ages such that the ages sampled may be skewed towards sets with the most catch, especially under scenarios where length sampling effort is high.
-If this is the case, then the representativeness of the age-length key to the whole population could be diminished by excessive length sampling.
-
-Age-length-key...large scale typical...danger, assumptions...segue to solution...
+Unlike the default approach to collecting and analyzing age samples, results from this test conform to expectations in that the estimator appears to be unbiased and, ignoring the costs associated with unnecessary sampling effort, there appears to be no negative consequences of increasing sub-sampling effort (Figure \@ref(fig:alt)).
+This is an important result with real-world implications for the survey the case study is based on.
+This example shows how **`SimSurvey`** can be used to reveal problems and explore solutions to the sampling designs of fisheries-independent surveys.
 

### 2020-02-13 analysis/paper/sim_survey_paper.Rmd

@@
 ```{r alt, out.height = "600px", fig.cap = "Key results from an alternative design-based sampling and analysis procedure that accounts for multi-stage cluster sampling of ages.
-Specifically, a) is a fan chart of stratified estimates of abundance at age from year seven of the simulation from a survey with a set density ($D_{\\mathrm{sets}}$) of 0.002 sets / km^2^, maximum number of length samples ($M_{\\mathrm{lengths}}$) of 100, and maximum number of age samples per 3 cm length group per set ($M_{\\mathrm{ages}}$) of 3 (the thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%); and b) is a surface plot of RMSE from an array of surveys with different sub-sampling effort, where the x-axes represent the maximum number of lengths to measure per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per 3 cm length group per set ($M_{\\mathrm{ages}}$)."}
+Specifically, a) is a fan chart of stratified estimates of abundance at age from year seven of the simulation from a survey with a set density ($D_{\\mathrm{sets}}$) of 0.002 sets / km^2^, maximum number of length samples ($M_{\\mathrm{lengths}}$) of 100, and maximum number of age samples per 3 cm length group per set ($M_{\\mathrm{ages}}$) of 3 (the thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%); and b) is a surface plot of RMSE from an array of surveys with set density ($D_{\\mathrm{sets}}$) fixed to 0.002 sets / km^2^ but different sub-sampling effort, where the x-axes represent the maximum number of lengths to measure per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per 3 cm length group per set ($M_{\\mathrm{ages}}$)."}
 
@@
 Here we evaluate the efficacy of alternate sampling protocols by assessing deviation of stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
-Although assessing model-based analyses was an option, we focused on the design-based stratified analysis for simplicity and because of its widespread use.
-The code below displays the core defaults will replicate the results of this case study.
-Results produced here are the same as those produced and displayed in a series of figures in the **[`test_surveys`]** section of the paper.
+We also evaluate designed-based estimators recommended by Aanes and Vølstad [@Aanes2015] that account for cluster sampling of ages.
+Although assessing model-based analyses was an option, we focused on the design-based analyses for simplicity and because of its widespread use.
+The code below displays the core defaults and will replicate the results of this case study.
+The last few lines of code will also replicate a simulation configured to test an alternate age sampling and analysis approach.
+Results produced here are the same as those produced and displayed in a series of figures in the **[`test_surveys`]** and **[Interpretation]** sections of the paper.
 
@@
 
+alt_surveys <- test_surveys(
+  distribution,
+  surveys = expand_surveys(set_den = 2 / 1000,
+                           lengths_cap = c(5, 10, 20, 50, 100, 500, 1000),
+                           ages_cap = c(1, 2, 3, 5, 10)),
+  n_sims = 5, n_loops = 200, cores = 3,
+  q = sim_logistic(k = 2, x0 = 3),
+  age_length_group = 3,
+  age_space_group = "set", 
+  alk_scale = "set"
+)
+
 ```
@@
 
+Bias in stratified estimates of abundance at age appears to disappear when length-stratified age sampling is conducted and age-length keys are applied on a set-by-set basis (Figure \@ref(fig:alt)a).
+Under this scenario, there also appears to be a plateau, rather than a valley, in RMSE values as length and age sampling effort is increased (Figure \@ref(fig:alt)b).
+Like the default approach (division-wide age sampling and age-length key), RMSE declines when length sampling effort is increased from around 5 to 100 measurement per set, however, beyond 100 there is little change in the RMSE values (Figure \@ref(fig:alt)b).
+In contrast to the length sampling axis, there is relatively little change in the age sampling axis; a small decline in RMSE is apparent as age sampling per length group per set is decreased from 1 to 3, with little change above this level of effort (Figure \@ref(fig:alt)b).
+
 ## Discussion
@@
 Ages are generally determined from length-stratified sub-samples of the catch and raw proportions of age-at-length are used to assign ages to fish in specific length groups.
-Age-length keys are usually constructed at larger spatial scales because there are rarely enough samples to construct a key at a finer spatial scale.
+Age sampling is often spread across large spatial scales and age-length-keys are usually applied at this scale (i.e.
+age sampling may not occur at every set).
 Here we construct one age-length key for the division, as is done for the analysis of 3Ps cod.
@@
 If this is the case, then the representativeness of the age-length key to the whole population could be diminished by excessive length sampling.
-
-The negative consequences of excessive length sampling on abundance at age estimates are most apparent under the low set density scenarios.
-The precision of the estimates are also the poorest when fewer sets are conducted.
-Sets, however, are the most costly sample to obtain and, as such, the set densities of fisheries-independent surveys tend to be on the lower end of the scenarios presented in this paper.
-The set density of the multi-species survey conducted by Fisheries and Oceans Canada in the Newfoundland region, for instance, ranges between 0.001 and 0.002 sets / km^2^.
-Length sampling effort, in contrast, tends to be on the higher end of the scenarios presented here because lengths are relatively easy to obtain.
-Again, using Fisheries and Oceans Canada sampling protocol from the Newfoundland region as an example, it is not uncommon for length sampling effort to be capped at 500 measurements per set.
-The results presented in this paper paradoxically suggest that better abundance at age estimates are obtained by lowering the length sampling cap.
-Results also indicate that reductions to length sampling effort would not significantly impact abundance at length estimates.
-Nonetheless, it is hard to be prescriptive because this simulation is far from a perfect reflection of reality and it focuses on one case study.
-Further research is required on the consequences of some of the assumptions of this simulation as well as the interaction between age-specific clustering, length-stratified sampling and the age-length key.
-Ideally, the simulation would also include multiple species with different life-histories and distributions, and also integrate a cost component to assess the trade-offs between information and cost.
-Finally, it would be interesting to test alternate analyses of these data that may account for the spatial structure of the age-length key (e.g.
-[@Berg2012]).
+In short, the primary cause of the problem is that the classic age-length key method assumes that fish are drawn from a simple random sample; this assumption, however, is rarely satisfied in practice [@Aanes2015].
+
+While there are potential model-based solutions to bias introduced by large scale age-length keys (e.g.
+spatial age-length keys [@Berg2012]), a design-based solution is to utilize estimators that account for multi-stage cluster sampling designs [@Aanes2015].
+This approach was tested here and, indeed, it appears to resolve the bias in the stratified estimates of abundance at age.
+This implies that inference from the actual survey of 3Ps cod can be improved by a simple adjustment of the sampling design and analytical approach.
+In the context of stock assessment, and unbiased estimator is certainly the preferred option because under-estimates may lead to unnecessary cuts to a fishery and over-estimates may contribute to ill-advised quota increases and potentially avoidable population declines.
 
@@
 This is largely because fish caught together tend to be more similar than those in the general population [@Pennington1994].
+Such cluster sampling can also introduce bias if the hierarchical sampling design is not explicitly accounted for [@Aanes2015].
 Therefore, if the goal of a trawl survey is to maximize information, it is likely better to stop collecting correlated samples and, instead, focus efforts the next set and/or other species.

### 2020-02-14 analysis/paper/sim_survey_paper.Rmd
Draft starting paragraph of the discussion

@@
 RMSE was also calculated for abundance at length estimates, where the above formula is indexed by length groups $l$, and total abundance, which lacks a group index of $a$ or $l$.
-
-## Assumptions
-
-Like any model, this simulation is a simplification of a much more complex reality.
-For instance, the population is assumed to aggregate by age-class and be uniformly distributed within a cell, instead fish may aggregate by length and form finer-scale clusters.
-The survey is also an instantaneous snapshot of the population, meaning that the population is assumed to be in the same location from the beginning to the end of the survey.
-Also, fish are aged at random within length bins and ages are estimated without error.
-Finally, area trawled is assumed to be perfectly standard.
-These assumptions, plus a range of others, will surely under-represent the natural variability of fish populations and survey protocol.
-Nevertheless, the **`SimSurvey`** package provides a relatively complex and flexible operating model for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
 
@@
 For the case study, a proper design-based estimator will need to account for cluster sampling in a stratified random survey.
-Following recommendations in Aanes and Vølstad [@Aanes2015], we used **`SimSurvey`** to 1) conduct a surveys with concurrent length and age sampling at every set, 2) construct and apply age-length keys on a set-by-set basis, and 3) weight the resultant age frequencies at each set using stratified random estimators.
+Following recommendations in Aanes and Vølstad [-@Aanes2015], we used **`SimSurvey`** to 1) conduct a surveys with concurrent length and age sampling at every set, 2) construct and apply age-length keys on a set-by-set basis, and 3) weight the resultant age frequencies at each set using stratified random estimators.
 This test was accomplished using this code:
@@
 
-GENERAL
+Though somewhat narrow in scope, results from our case study demonstrate why it is important to evaluate the design and analysis of complex surveys.
+**`SimSurvey`** settings for our case study were configured to reflect the key properties of an actual survey of a cod population.
+In short, the population was simulated to exhibit age-specific clustering and these clusters were sampled by multi-staged stratified random surveys where age sub-samples are from length sub-samples from fish caught in a set.
+This type of sampling is common in fisheries research, however, the clustered nature of the samples are rarely taken into consideration [@Nelson2014].
+Instead, it is commonly assumed that individuals are simple random samples from the target populations even though such sampling is virtually impossible to accomplish in practice.
+The assumption of simple random sampling implies that the characteristics of fish caught at the same location are independent and, if estimators that assume independence are used, this can introduce bias and underestimate variance [@Nelson2014; @Aanes2015].
+Moreover, this assumption may contribute to the perception that more sub-samples will lead to better population estimates.
+Increasing sub-sampling effort, however, has been shown to be an ineffective means of improving estimates because samples at the same location are usually correlated; the best way to increase the number of independent samples is to sample more locations [@Pennington1994; @Pennington2002; @Stewart2014; @Bogstad1995; @Coggins2013; @Zhang2013].
+Results from the case study reiterate these points and indicate that there is room for improvements to the actual survey the case study was based on.
+Specifically, it appears that valuable human resources may be wasted by collecting excessive samples of correlated metrics, and abundance at age estimates may be biased by the assumption that age samples are from a simple random sample.
+We have used **`SimSurvey`** to identify solutions to these problems and, given current results, we suggest that time can be saved by decreasing sub-sampling effort and bias can be reduced by applying a design-based solution to the age sampling and analysis that accounts for the cluster sampling.
 
@@
 
-The case study described in **[S1 Appendix][S1 Appendix: Case study]** provides one example of how **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
+The case study used here provides one example of how **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
 There are multiple layers to the sampling design of such surveys and the end results are linked with the way such data are analyzed.
@@
 
+## Assumptions
+
+Like any model, this simulation is a simplification of a much more complex reality.
+For instance, the population is assumed to aggregate by age-class and be uniformly distributed within a cell, instead fish may aggregate by length and form finer-scale clusters.
+The survey is also an instantaneous snapshot of the population, meaning that the population is assumed to be in the same location from the beginning to the end of the survey.
+Also, fish are aged at random within length bins and ages are estimated without error.
+Finally, area trawled is assumed to be perfectly standard.
+These assumptions, plus a range of others, will surely under-represent the natural variability of fish populations and survey protocol.
+Nevertheless, the **`SimSurvey`** package provides a relatively complex and flexible operating model for simulating stratified-random survey data from a population that varies across age, year and space dimensions.
+
 ## Summary
 
-The **`SimSurvey`** package serves as a tool for simulating stratified random surveys of dynamic populations that vary across ages, time and space.
+The **`SimSurvey`** package serves as a tool for simulating sample surveys of dynamic populations that vary across ages, time and space.
 The core of the simulation is based on the widely used cohort equation and, even though the processes that define recruitment and total mortality are simple, a wide range of stock dynamics can be simulated by changing a few parameters.
@@
 The next necessary steps to generating data similar to actual observations is to conduct a survey.
-In this package we implement a function, `sim_survey`, that conducts a stratified random survey of the population.
+In this package we implement a function, `sim_survey`, that conducts a survey of the population.
 The sampling process is governed by the area covered by the trawl as well as age-specific catchability.

### 2020-02-14 analysis/paper/sim_survey_paper.Rmd
Tweak abstract and introduction

@@
 **`SimSurvey`** also includes a function for estimating the stratified mean and variance of the population from the simulated survey data.
-We demonstrate the use of this package using a case study and show that the it can reveal unexpected sources of bias and be used to explore design-based solutions to such problems.
+We demonstrate the use of this package using a case study and show that it can reveal unexpected sources of bias and be used to explore design-based solutions to such problems.
 In summary, **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring.
@@
 These features are not easily captured by simple statistical models, which makes it difficult to determine optimal sampling and sub-sampling schemes [@Pennington1994; @Pennington2002; @Stewart2014].
-In such situations it is common for practitioners to resort to simulations as a means of testing the efficacy of various sampling designs (e.g.
-[@thomas2010]), however, the complexities associated with fisheries-independent surveys may explain why simulations that test the full sampling and analytical pathway are rare (but see [@puerta2019; @schnute2003]).
-
-Here we document **`SimSurvey`**, an `R` package designed to simplify and facilitate realistic simulations of fisheries-independent trawl surveys.
+In such situations it is common for practitioners to resort to simulations to test various sampling and analytical pathways (e.g.
+[@thomas2010]), however, such tests are rare in fisheries research (but see [@schnute2003; @puerta2019]), perhaps because of the complexities associated with fisheries-independent surveys.
+
+Here we document **`SimSurvey`**, an `R` package designed to facilitate realistic simulations of fisheries-independent trawl surveys.
 In short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time.
@@
 The steps taken to simulate surveys of spatial, age-structure populations are outlined below.
-Under the **[Methods]** section, we first outline the equations underlying the package (**[Model structure]** section) and then move into a vignette of the package (**[Core functions]** section).
-While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study.
-Output from default function calls are therefore relevant to the case study and some of these results are highlighted in the **[Interpretation]** section; a more detailed description and discussion of these results can be found in **[S1 Appendix][S1 Appendix: Case study]**.
+We first outline the equations underlying the package (**[Model structure]** section) and then move into a vignette (**[Core functions]** section).
+The core functions of the package are largely demonstrated using default settings, and these settings are based on a case study (see for **[S1 Appendix][S1 Appendix: Case study]** details).
+Several of the results from the case study are described and discussed in the **[Interpretation]** section as they highlight the utility of the package.
+Finally, we provide some guidance on how to modify default settings to suit specific needs (**[Parameterisation]** section) and then move onto discuss the potential uses and future directions of the package (**[Discussion]** section).
 

### 2020-02-14 R/run_strat.R
Add warning for cases where age sampling scale does not match age-length key scale

@@ -44,6 +44,16 @@ strat_data <- function(sim, length_group = 3, alk_scale = "division") {
   }
   samp <- merge(setdet[, cols, with = FALSE], samp, by = "set")
 
+  ## Produce warnings if age sampling scale does not match scale requested for age-length key
+  ag_check <- setdet[n_measured > 0, list(n_aged = sum(n_aged)), by = c("sim", "year", "strat")]
+  if (alk_scale == "strat" && any(ag_check$n_aged == 0)) {
+    warning("Age-length keys cannot be constructed at the strat level because ages were not sampled at every strat. This discrepancy may introduce bias. Consider running sim_survey with age_space_group = 'strat'")
+  }
+  ag_check <- setdet[n_measured > 0, list(n_aged = sum(n_aged)), by = c("sim", "year", "set")]
+  if (alk_scale == "set" && any(ag_check$n_aged == 0)) {
+    warning("Age-length keys cannot be constructed at the set level because ages were not sampled at every set. This discrepancy may introduce bias. Consider running sim_survey with age_space_group = 'set'")
+  }
+
   ## Construct length-frequency table
   lf <- samp
   lf[, n := .N, by = "set"]

### 2020-02-14 analysis/paper/sim_survey_paper.Rmd
Other small tweaks to the paper

@@
 Likewise, age frequencies need to be calculated to obtain stratified estimates of abundance at age.
-This is done by constructing an age-length key, which is the proportion of fish in each length bin that fall into specific ages classes, and applying this key to the bumped up length frequencies, $n_{l,i}$, to approximate age frequencies, $n_{a,i}$, that is:
+This is done by constructing an age-length key, which is the proportion of fish in each length bin that fall into specific ages classes.
+Once these proportions are calculated, they applied to the bumped up length frequencies, $n_{l,i}$, to approximate age frequencies, $n_{a,i}$, that is:
 
@@
 Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by adjusting a few parameters in these closures (Figure \@ref(fig:distribution)).
+Though we only provide one option for simulating a spatial noise (see **[S2 Appendix][S2 Appendix: Age-year-space covariance]** for details), advance users could construct closures that leverage simulation models provided by packages such as **`RandomFields`** [@schlather2020] or **`INLA`** [@rue2009].
 Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e.
@@
 This structure was imposed because fish caught together tend to have similar characteristics and it is well known that this reduces the effective sample size [@Pennington1994].
-Increases sub-sampling effort, however, should not lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)).
+Increasing sub-sampling effort, however, should not lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)).
 Bias, in particular, appears to be a problem with stratified estimates of abundance at age under certain scenarios (Figure \@ref(fig:fan3)); in contrast, there appears to be little bias in the stratified estimates of abundance at length (Figure \@ref(fig:fan2)).
@@
 There is, however, a potential cost to the spatial scale of the key.
-Namely, it is unlikely that one age-length key is representative for the whole region because the probability of being a specific age given length varies in space [@Berg2012].
+Namely, it is unlikely that one age-length key is representative for the whole region as the probability of being a specific age given length varies in space [@Berg2012].
 Because different age groups sometimes occur in different places, the translation of lengths to ages may be biased by the samples used to generate the age-length key.
@@
 
-Unlike the default approach to collecting and analyzing age samples, results from this test conform to expectations in that the estimator appears to be unbiased and, ignoring the costs associated with unnecessary sampling effort, there appears to be no negative consequences of increasing sub-sampling effort (Figure \@ref(fig:alt)).
-This is an important result with real-world implications for the survey the case study is based on.
+Unlike the default approach to collecting and analyzing age samples, results from this test appear to be unbiased and additional sub-sampling effort beyond a certain point appears to have little to no effect on the estimates (Figure \@ref(fig:alt))).
+This is an important result with real-world implications as they imply that 1) inference from the actual survey can be improved by altering the sampling and analysis of ages and, 2) valuable survey time can be saved by cutting back on sub-sampling effort.
 This example shows how **`SimSurvey`** can be used to reveal problems and explore solutions to the sampling designs of fisheries-independent surveys.
@@
 Though somewhat narrow in scope, results from our case study demonstrate why it is important to evaluate the design and analysis of complex surveys.
-**`SimSurvey`** settings for our case study were configured to reflect the key properties of an actual survey of a cod population.
-In short, the population was simulated to exhibit age-specific clustering and these clusters were sampled by multi-staged stratified random surveys where age sub-samples are from length sub-samples from fish caught in a set.
+Specifically, we tested the design and analysis of a multi-staged stratified random survey of a cod population.
 This type of sampling is common in fisheries research, however, the clustered nature of the samples are rarely taken into consideration [@Nelson2014].
-Instead, it is commonly assumed that individuals are simple random samples from the target populations even though such sampling is virtually impossible to accomplish in practice.
+Instead, it is commonly assumed that individuals are randomly drawn from the target populations even though such sampling is virtually impossible to accomplish in practice.
 The assumption of simple random sampling implies that the characteristics of fish caught at the same location are independent and, if estimators that assume independence are used, this can introduce bias and underestimate variance [@Nelson2014; @Aanes2015].
@@
 Specifically, it appears that valuable human resources may be wasted by collecting excessive samples of correlated metrics, and abundance at age estimates may be biased by the assumption that age samples are from a simple random sample.
-We have used **`SimSurvey`** to identify solutions to these problems and, given current results, we suggest that time can be saved by decreasing sub-sampling effort and bias can be reduced by applying a design-based solution to the age sampling and analysis that accounts for the cluster sampling.
+We have used **`SimSurvey`** to identify solutions to these problems and, given current results, we suggest that time can be saved by decreasing sub-sampling effort and bias can be reduced by using an approach that account for the hierarchical design of the survey.
 
@@
 
-The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formula for stratified-random designs (e.g.
+The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formula (e.g.
 [@Cochran1977]).
@@
 the probability a fish is a specific age given length) and age frequencies are obtained by applying this key to length-frequencies obtained via more expansive random sampling.
-One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012].
+One age-length key is typically assumed to be representative of the whole stock area, however, spatial variability in the relationship may introduce bias in abundance-at-age estimates [@Berg2012; @Aanes2015].
 Results from the case study (see **[S1 Appendix][S1 Appendix: Case study]**) reiterate this point and **`SimSurvey`** may serve as a platform for testing potential model-based solutions to this problem (e.g.
@@
 We will therefore endeavor to add functions for simulating multi-species surveys.
+Another limitation of the package is that we have yet to implement alternatives to random or stratified random survey designs (e.g.
+systematic sampling); expanding these options would allow for a more comprehensive evaluations of various designs.
 Finally, it would be useful to add an option for testing the consequences of surveys with partial coverage of a population as survey coverage is a frequent concern in stock assessment.
@@
 We also thank Dave Cote and Joanne Morgan for providing constructive comments on a previous version of this manuscript.
-Finally, the package and the manuscript were greatly improved by feedback from two anonymous reviewers.
+Finally, the package and the manuscript were greatly improved by feedback from editor Daniel Duplisea and two anonymous reviewers.
 This work was supported by the NSERC visiting-fellow program and Fisheries and Oceans Canada.

### 2020-02-17 analysis/paper/rebuttal_letter.Rmd
Start drafting response letter

--- 2019-12-18 analysis/paper/rebuttal_letter.Rmd
@@
 ---
-output: word_document
+output:
+  word_document: default
+  md_document: default
 ---
@@
 
-Thank you for considering a revision of the manuscript PONE-D-19-26904,  **"`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"** by Paul M.
+Thank you for considering another revision of manuscript PONE-D-19-26904R1,  **"`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"** by Paul M.
 Regular, Gregory J.
-Robertson, Bob Rogers, and Keith P.
-Lewis.
-In accordance to one of the reviewers' suggestions, the paper has been renamed **"`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"**.
-
-We thank the Reviewers for their thoughtful and thorough comments which will clarify and improve the manuscript.
-We agree with many of the comments and have made every effort to incorporate these into the manuscript.
-Below are the reviewers' comments with a brief accounting of our response to each concern or suggestion.
-
-Lines numbers refer to the revised manuscript.
+Robertson, Keith P.
+Lewis, Jonathan Babyn, Brian Healey and Fran Mowbray.
+We are also grateful for your detailed suggestions and we have made every effort to do justice to the changes you recommend.
+Most importantly, we hope we have added sufficient content to the core of the paper elevate the manuscript from solely a software manual to a primary scientific publication.
+Though the how-to approach remains, we now see that describing some of the case study results in the core of the manuscript makes it more interesting and it adds another tangible reason for prospective users to learn how to use the package.
+Please see below for more details on the changes we made in response to your suggestions.
 
@@
 
-Reviewer #1:
+Additional Editor Comments (if provided):
 
-This paper describes the R package “SimSurvey” which is a package to test the design and analysis of fisheries independent surveys.
+Dear Paul, you have made lots of excellent changes that I think help with the use of the package and responded to the many specific comments of reviewers which has no doubt corrected many technical issues and clarifications.
+The one thing I would say that you have not really done is get at the deeper research merits of this work beyond introducing a new piece of software.
+I think paper, as it stands, lacks a larger context and content which is important for the primary publication.
+My diagnosis for why this is is that the manuscript does not conform very well to more typical scientific reports (Intro, M&M, Result, Discussion) which can make it difficult for readers to find the larger scientific merits of the work.
+It is useful of course for those who already understand the merits of this kind of work but this work is for primary publication and it needs to appeal more to the former than the latter group.
+There is a very "how-to" feel to it (e.g.
+line 64 "In this section") which I think detracts from getting at the larger purpose of the work.
 
-I found the paper very interesting and the package of great potential value and interest to many fishery scientists.
-However, I have several comments that I would like the authors to address first before I recommend acceptance of this paper.
-My main comments relate to:
+I would really like you to address this issue of moving it from a software manual to a primary scientific publication.
+I do not think it should involve that much work but there will be some restructuring of sections as well as places to put in content and bring out conclusions.
+Here are my suggestions for this:
 
-1.
-Verify some of the equations and model description used in the model.
-I think that some of them lack description to fully understand what was done and few others are misleading, even incorrect.
-The detail can be found below but in general, they are linked to the problem of bias correction for log-normal distribution, re-scaling”, moving from abundance at age to abundance at length.
-
-2. Add additional explanation to justify the choices. E.g. why population dynamics is modeled independently of the spatial distribution. Is this reasonable, limitation, etc.
-
-3. Few suggestions (add-ons, renaming) to increase the generality of the paper (without requiring too much work)
-
-*We hope we have addressed each of these main comments by 1) clarifying several equations, 2) adding more details and justifications, and 3) modifying naming conventions and clarifying extensibility.
-See replies below for more details.*
-
-Detailed comments:
-
-Abstract:
-
-What about the flexibility to include new sampling strategies (user defined one) or including new estimation approaches
-
-*We have noted that both built-in and user-defined strategies can be utilised.*
+Try to follow a more traditional paper structure.
+This will help readers and it likely will also make it clearer for you on how you can inject content into the paper to move it beyond the software manual approach:
 
@@
 
-L43: quality of
+- You need to talk about wider issues and examples.
+e.g.
+examples of when poor survey design meant that scientific questions could not be properly addressed when good survey design meant they were.
+Examples of when good survey design allowed researchers to address needs that were unanticipated at the time it was designed.
+i.e.
+you need to build a better case (not just cost) of why survey design is really important and it is most powerful to do this with examples.
+You should try to bring in ideas related to ecosystem and climate changes and being able to track communities.
+Perhaps bring in species at risk ideas and tracking decline, you could bring in ideas related to MSC certification.
+These are just examples of specifics but you get the idea: the Introduction needs to have more general information outlining in both a broad and specific sense why we should be concerned about this.
 
-*We have made the suggested change.*
+- Keep in mind an educated reader who may not be in fisheries but is interested in why anyone should care about this or could, for example, be interested in surveying say caribou or songbirds or something outside of fisheries but where many of the motivation and concepts may be similar and they are doing a more general literature search before designing their own survey.
 
-It would be good to add few references here e.g. L50,
+*Thank you for thinking beyond a fisheries biologist audience, as this is the audience we have been targeting from the onset.
+In the case that this is short-sighted, we have modified our introduction, as suggested, to pitch the concept to a broader audience.
+The reason we did not think about other biologists is because fisheries biologist tend to have access to age-based data while others (e.g.
+seabird and caribou biologists) do not.
+Alas, we agree that there could be biologists outside the field of fisheries that have such data or stage-based data who may find the package useful; if nothing else, the general framework may be useful to adapt to other marine or terrestrial situations.*
 
-*We have added references as suggested.*
+Methods:
 
-Method:
+- This will start at your "Model Structure" section. You should put a higher level heading just before that called Methods.
 
-L69-73: I think it would be good to add the reference to the specific function name listed in Table2.
-This way, I would be more explicit and avoid possible confusion.
-For example, in my case, when I first read the section header “simulate abundance”, I was expecting to see the way you dealt with the spatial distribution of abundance.
+- I can see that it is hard to separate your Methods from Results.
+Your results are really in the section "Using SimSurvey" I would not be opposed to putting this as a separate Results section but I leave it to you to decide.
+Essentially what you have is a case study as an example of how it works and therefore specific results are less interesting than how you got there with the package.
+You might title it something like "Results: running a SimSurvey simulation".
+This section also has a lot of content without a lot of explanation.
+for example, you have several large and complicated figures in a row.
+You should discuss not only what figures are meant to offer in the package but also what they mean.
+So for example on line 473 you refer to three figures (5,6,7) but you offer little interpretation of those figures just why you can make them which is another example of the limitations of the how-to manual approach.
+So you need to think of it as a case study and help someone decide the implications of their survey design.
 
-*We have made the suggested change.*
+*We really struggled to accommodate a traditional Methods and Results section as any structure we attempted betrayed the definition of these sections.
+----DESCRIBE WHAT WE DID HERE----.
+Below we discuss a couple of failed attempts:*
 
-L94&95&104: I just want to make sure that the bias correction has been applied to these two equations.
-If logN (or logZ) is normally distribution with mean mu and sigma, then the N (or Z) is lognormally distributed with mean exp(mu+sigma^2/2).
-Therefore, when you write “mean” in table 3 for R and Z, I hope that you modified in your actual equation for log(Z) to:
-log(Z) ~ normal(log(mean)-sigma^2/2, sigma)
-In other word, the mean(R) is not simply exp(meanR_log_scale). exp(meanR_log_scale) is the median of the distribution of R.
+*1) We tried to start the methods section with the "Model Structure" section, as suggested, and shunt the figures to a "Results" section; this, however, broke the flow and left a gap in the methods section where we show the methods (functions) for running a simulation but do not show the plotting methods.*
 
-*Good catch.
-The answer to the question is no, bias correction was not applied.
-This is a matter of naming convention that we struggled with.
-The intent was to operate in log space for these equations and, as such, the simulations were set-up to receive log means.
-When developing the functions, however, we waffled on whether to use `log_mean` or `mean` as the argument name and we landed on `mean` as we figured this would be a more relatable value for users to supply.
-Kudos to the reviewer for noticing that we failed to consider bias-correction.
-To be more explicit with regards to what we "mean", we have renamed the arguments to `log_mean`.*
+*2) We tried to keep the methods section as is (i.e.
+describe equations and functions) and describe some of the key case study results in the results section, but it was difficult to find a logical placement for the "Parameterisation" section and it did not seem right to segue to a solution to a problem identified in the case study in a traditional results section.
+In contrast, it seemed fine to describe and discuss the case study results demonstrated through the "Core functions" section and then jump to a solution to a problem in an independent section.*
 
-L107: there is not enough description to understand how you converted the abundance at age to abundance at length.
-What were the choice of length bin groups?
-Every 5 cm, 10cm?
-Such information is also missing in the tables.
-Please add that.
-But then you have to calculate some cumulative distribution (until length l) for the length at age distribution (using the VB growth equation) and do some subtractions to be able to allocate the abundance at age to each length group.
-See equation A.1.14 in here for example. https://www.nwfsc.noaa.gov/news/events/program_reviews/documents/C.2_Methot_Wetzel_SSTechnicalDescription.pdf
+Results:
 
-*Another good catch as we did not provide enough information to replicate the approach.
-We have added more detail on the calculations behind the sim_vonB function; turns out our approach was similar to the one described in the exemplar the reviewer provided.*
+- see comment above
 
-“Simulate spatial distribution”:
+Discussion:
 
-I have a few comments on this section:
+- This starts just before "Research Opportunities".
+You can keep these sections but there should be more preamble before jumping right into research opportunities.
+I suggest a general paragraph(s) that segue into your subsections of "research opportunities", "future research".
+In the Discussion you should address again some of the broader issues from the Introduction, e.g.
+how could spatial (depth) distribution changes anticipated under climate change for some population be tackled by survey design exploration now so that we can continue to track these changing populations 20 years down the road and do not lose the signal.
+How can this software help with that?
 
-1.
-First of all, I think it is important to state that you assume that the population dynamic model and the spatial distribution is independent i.e.
-you do not have a spatially-explicit population dynamics model.
-And talk about what it means (is it realistic, limitation, etc).
+- I think something that can be useful for readers is to outline the steps in a thought process a researcher might undertake when setting up a survey (perhaps a separate subsection) and then how one might go about a SimSurvey run for this.
+It also gives you a good opportunity to discuss your multispecies ideas:
+    - What is your current problem/needs
+    - What are the constraints (HR and $ perspective)
+    - What are the constraints from a biological, physical perspective
+    - Resources available
+    - What are your anticipated future needs (i.e. if climate change is going to make cod go deeper, will you be able to track it in 20 years?)
+    - How could you consider some or all of these with SimSurvey
 
-*We have prefaced this section with the caveat that population and spatial dynamics are modeled as independent processes and we note the pros and cons of this approach.
-In short: the con is that the approach is a simplification of reality as it does not explicitly account for dynamics such as larval dispersal, spatial differences in growth and meta-population connectivity; the pro is that this simplification limits the number of unknown parameters that need to be specified while facilitating the simulation of a sufficiently complex population for testing the efficacy of various survey designs.*
+- Your "Assumptions" section should come down into the Discussion
 
-2.
-L161.
-I think it would be better to rename “depth” as “main covariate influencing the species distribution”.
-And “depth” happens to be one of them for many species, but there are cases when this is not the case.
-
-*Fair point, however, we focused on "depth" for two reasons: 1) it is an important covariate for many species (as the reviewer noted), and 2) most surveys that we are aware of are depth-stratified (also noted by the reviewer below).
-We therefore retained the "depth" naming as it ought to satisfy most use cases.
-Users can satisfy other, less common, cases by supplying user-specified grids and/or spatial populations.*
-
-3. L161. I think it would be good to say upfront that there are two ways of defining the spatial structure. Using the function in the package or user defined.
-
-*Given a suggestion from Reviewer #2, we have restructured the paper to have two core sections: "Model structure" and "Using __`SimSurvey`__".
-We have included a blanket statement under the "Model structure" section noting that users can circumvent specific components of the framework.*
-
-4.
-L161: I think another possibility is to generate a random field by using the package “Randomfields” for example.
-This way you can generate a map where “depth” is patchily distributed (maybe more like an island type case study)
-
-*Good idea!
-We actually pursued this idea in an earlier iteration of `make_grid`, however, we abandoned the option because it was difficult to automate; the random field created random problems with the depth-stratification (e.g.
-one cell strata).*
-
-5.
-L163-164: please add more description about the division or strata.
-How can we set it up and what do you specifically mean by division and strata?
-Is one nested within the other, or not necessarily?
-Your examples are based on Atlantic Ocean and people in other regions might not be familiar with how these divisions are created.
-
-*We have added more detail on the structure of the divisions and strata.*
-
-6.
-L163: Why only focus on “depth-based strata”?
-I think it would be good to allow the user to choose their own stratification approach.
-It could be depth based as you did (which happens most often in surveys) but it could technically be any other thing (user supplied).
-This allows more flexibility.
-
-*See reply to point 2. above.*
-
-7.
-L167: the equation is misleading and I am not sure it is right based on what is written in the text.
-You mentioned later on L178: that you “re-scaled” so that the total number of fish in a specific year and age across space is equal to the single number from the population dynamics model.
-If so, the re-scaling should be done in the identify (natural) scale, not in log scale.
-In log scale, even if ,, sums to zero across space, the sum in the identity scale won’t match.
-This is often refer to as “bias correction” for the log normal models.
-And you should ideally show how the rescaling was done in terms of the equations too.
-
-*Right, the equation presented was not an accurate reflection of the calculations. We have revised the equation to explicitly show how the values were normalized to sum to 1.*
-
-8. L167: this “depth preference” function is very simplistic and gives only very “smooth” symmetric distribution. More often, fish have a skewed depth preference: often right-skewed.
-
-*True.
-Some users may find this parameterization insufficient for their species and we hope they will implement their own closure to use in the `sim_distribution` function to better simulate the effect.
-In addition to our blanket statement under the "Model structure" section, we have added a more specific statement under the "Using __`SimSurvey`__" section stating that alternate formulations can be used by supplying alternate closures to the core functions.*
-
-9. L173-174: I think it might be worth adding, in simpler terms, the meaning of the spatial smoothing and scaling parameters.
-
-*We have prefaced that sentence with "The rate at which point-to-point spatial correlation decays with distance is controlled by...".*
-
-10.
-L178: it is another question of scaling.
-How did you exactly do the scaling?
-In the identity scale?
-By dividing my the sum of the effects?
-I am asking this because depending on how you did the rescaling, your correlation structure in space and age might have been affected and is not the same as the one specified in L172.
-Did you verify that?
-
-*See reply to point 7. above*
-
-Table 4. “group_ages”. Ok but how is the variance controlled for the other age classes?
-
-*"Variance" was a poor word choice. We have replaced it with "noise" as it is the simulated noise that we fix across multiple age groups.*
-
-L189: “user supplied”.
-This is a good feature.
-However, I think it is important to mention here that user have to make sure that they use the correct projection method to ensure that each grid is of the same dimension.
-
-*We have made the suggested change.*
-
-L216: reference to figure 3?
-
-*We have made the suggested change.*
-
-L221: there is not “group_years” argument in Table 4.
-
-*We have added it to the table.*
-
-L237: “this function”. I think it would be better to replace “this function” with “sim_distribution” as you do not mention the word “sim_distribution” in the sentence above this.
-
-*We have made the suggested change.*
-
-L158-251: In general, I think re-organising this section using sub-section headers could be useful. Just to guide the readers
-
-*This is an excellent suggestion. By following a suggestion by Reviewer #2 to re-organize the paper into two core sections we have added more headers to help guide the readers.*
-
-L254: you say that sampling is stratified random but SRS is also an option based on Table5. Please correct.
-
-*We have made the suggested change.*
-
-L257: what does this mean? Does this control the number of set but how is this calculated?
-
-*We have clarified how number of sets per strata is calculated.*
-
-L257: I do not see how you control for the total number of set in the survey? How do you control it?
-
-*We have clarified how number of sets per strata is calculated.*
-
-L261: I think you should mention here that you can also force the sample size (as seen on table 5).
-Moreover, in table 5, it would be good to set-up a “ages_min” for the minimum number of ages to sample […]” so that it gives the ability to fix the sample size if needed by writing the same value for “ages_min” and “ages_cap”.
-
-*We are not sure what the reviewer would like to have implemented here. Is the suggestion to impose a minimum number of ages to collect across all length groups?*
-
-L261: How are you making sure that the number of sampled fish for that specific cell, age and year won’t be above the total number of fish in that cell, age, and year?
-The probability value could be close to one and if you fish in a few a time, then you are at risk.
-Especially because your population dynamics model is not spatially explicit and is completely independent of the distribution function itself i.e.
-you can technically fish out all the fish in an area but it will be populated back the year after the way you implemented in this study… Maybe you need to put a condition (or just a note) for general users to make sure that this probability value is much below 1?
-
-*The sampling is implemented such that the number of fish sampled in a cell cannot exceed the number of fish in a cell because the population is split across sets in cases where more than one set is conducted in a cell.
-We have added this missing detail to our manuscript.
-We also added a note that the survey is assumed to have no impact on the population from one year to the next.*
-
-L267: I recommend to clarify something here.
-1.
-Depending on the number of fish caught?
-What do you mean?
-What is the rule you used?
-2.
-The way you coded, sample by age is first decided, then the corresponding length is calculated, then age-subsample is determined.
-In reality, length sample is taken in the field, then age sub-samples are taken.
-While similar, I do not think it always equal.
-Especially, when you start including some correlation structure in the sampling.
-By the way, did you consider including some correlation structure in the sampling process to make more realistic?
-
-*Honestly, we do not recall what we mean by "depending on the number of fish caught".
-Perhaps we added those words to cover off cases where no fish are caught.
-Whatever the case, we have removed the statement to minimize confusion.
-We have also clarified the sub-sampling sequence.
-Finally, we have yet to consider including correlation structure in the sampling process as we went about imposing correlation via the spatial correlation of age groups (i.e.
-age-specific clustering tends to result in sets with high intraclass correlation).
-We are open to learning more about other processes that may contribute to correlated samples.*
-
-L275 Table 1 on should be table 5
-
-*We have changed the page numbers accordingly.*
-
-L275: Table5: “age_sammpling” should be “age_sampling”
-
-*We have made the suggested change.*
-
-L275: “min_sets” you have not described it yet and what is it? You have sample from all cells? If no, this is not realistic.
-
-*We have clarified the meaning and utility of the `min_sets` argument (i.e.
-a small strata may be allocated only one set under a low set density scenario; this argument overrides the allocation and imposes the `min_sets` if it is greater than the allocation).*
-
-L279: Table5 not Table1?
-
-*We have changed the numbers accordingly.*
-
-L285-286: Could you be more specific on how custom closures can be supplied and where?
-
-*We have included an example that ought to clarify how a custom closure can be supplied.*
-
-L306: how are these catchability corrected abundance matrices calculated?
-It is important to write this information somewhere (or write “please refer to the section “Stratified analysis” for further information on the calculation of abundance indices”) or something alike and Appendix S3.
-
-*We have clarified how this was calculated.*
-
-L336: I think it would be good to say that other methods exist and people can use it in this package (maybe)?
-
-*Good point, however, we think this is covered by referencing a paper that describes a geostatistical R package and we also note that other options can be used under the "Research opportunities" section.*
-
-L421: color gradient. Even though it is obvious it might be good to say green to purple gradient.
-
-*We have made the suggested change.*
-
-L427: instead of “sampling protocol”, I think it would be more meaning full to say the maximum number of length samples.
-
-*We have made the suggested change.*
-
-L452: say that the color ramps from yellow to purple
-
-*We have made the suggested change.*
-
-S1 appendix: missing figure in S1
-
-*We have included the figure*
-
-----------------------------------------------------------------------------------------------------
-
-Reviewer #2:
-
-This manuscript describes an R package called SimSurvey.
-The package includes a set of functions for simulating point-based fisheries survey designs, e.g.
-bottom trawl surveys, for estimating abundance indices.
-It focuses on number of stations and number of fish sampled ignores other constraints such as distance between stations and day-time duration which impose strong constraints on real surveys.
-The functions included allow the user to first simulate age-and length structured population dynamics, distribute individuals randomly in space (assuming a certain correlation structure), carry out a survey and finally calculate abundance indices from the simulated data.
-The package will likely be of interest to researchers wanting to explore the precision achievable with different survey designs.
-My comments regarding the package and the presentation are summarised below.
-
-1. Optimization
-
-The title announces a package for optimizing survey designs.
-As far as I can see the package does not allow survey design optimization, neither in terms of defining survey strata nor in terms of number of stations per stratum.
-The strata are defined by the user.
-The only option available for the number of stations is proportional to stratum surface; the user sets the minimum number of stations taken in the smallest stratum.
-It would be useful to be able to specify the total number of stations and test different allocation schemes, such as proportional to surface area (implemented), equal number per stratum, Neyman allocation (accounting for surface area and abundance variability), etc.
-Please consider revising the title (e.g. “compare” and instead of “optimize)” and spell out the available sampling design options.
-
-*We agree with the Reviewer and retitled the ms "`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks".*
-
-2. Manuscript structure
-
-The manuscript might be easier to follow if the manuscript was restructured: 1) Model description, 2) Using SimSurvey.
-The later section would then group all example code which could again be subdivided into running simulations and exploring results (plot functions).
-
-*This is an excellent suggestion! We have re-structured our manuscript accordingly and feel that this structure will be much easier for a reader to follow.*
-
-3. Parameterisation
-To use SimSurvey for a real world problem realistic parameter values are needed.
-The package comes with default values chosen for a particular case study.
-However, no mention is made in the manuscript how to choose appropriate values for the many model parameters to tailor the simulations to a population of interest.
-I suggest the authors add a section on parameterisation and a table summarizing all parameters with a column specifying how to parameterize.
-For example, parameters for population dynamics and growth could probably be taken from the literature (or a stock assessment report).
-However this is not possible for the parameters of the spatial distribution function sim_distribution() such as correlation between ages etc.
-Ideally the package would include a fitting function for estimating these parameters from actual data.
-These input data could come from a pilot survey and include location (lat, long) and numbers by length/age.
-
-*TODO*
-
-Minor issues
-- line 93: I assume there is an age plus but this needs to be mentioned. Also, please specify how you set the initial numbers for plus group ().
-
-*We have now noted in the manuscript that a plus group is not modeled as the number of ages can easily be extended to include groups with zero fish.
-This choice simplifies the simulation, including the setting of initial numbers which is done via exponential decay.
-Further, the lack of a plus group is inconsequential for survey based estimates of abundance at age.*
-
-- line 121: I don’t understand the explanation of a closure. What do you mean by “return functions”? Do you mean it returns an object with different attributes?
-
-*We have improved our explanation of a closure at the beginning of the Using __`SimSurvey`__ section*
-
-- line 125: the number of right and left brackets is unbalanced, please check
-
-*This was actually correct and, given an improved description of a closure, the logic behind this line of code should be clearer now.*
-
-- line 126 “This structure was chosen to avoid the repeated specifications of ages and years”. As far as I can see the example code only specifies years, not ages.
-
-*Again, we hope that our improved description of a closure will clarify what we mean by this.*
-
-- line 227 Please explain what a pipe is and how it is used.
-In the example I understand that the output of sim_abundance( ) is provided to (piped) into sim_distribution().
-I am unclear what the object b contains.
-Is it the result of sim_distribution()?
-
-*We have clarified how a pipe works, noting that it forwards values from one function call to the next function call, and we now state that the output from the two examples provided (nested approach vs.
-pipe approach) are functionally the same though the approach is slightly different.*
-
-- There is no table 1, please revise table numbering.
-
-*We have made the suggested change.*
+- "Future Directions" should say something about randomfields re: Reviewer 1.
+Even if you just outline that you have considered it.
+You might also try to say something about optimisation of design which Reviewer 2 mentioned.

### 2020-02-18 analysis/paper/rebuttal_letter.Rmd
Complete draft of responses to the editor's suggestions

@@
 
-Additional Editor Comments (if provided):
+Additional Editor Comments:
 
@@
 *Thank you for thinking beyond a fisheries biologist audience, as this is the audience we have been targeting from the onset.
-In the case that this is short-sighted, we have modified our introduction, as suggested, to pitch the concept to a broader audience.
+In case this is short-sighted, we have modified our introduction, as suggested, to pitch the concept to a broader audience.
 The reason we did not think about other biologists is because fisheries biologist tend to have access to age-based data while others (e.g.
@@
 
-*We really struggled to accommodate a traditional Methods and Results section as any structure we attempted betrayed the definition of these sections.
-----DESCRIBE WHAT WE DID HERE----.
-Below we discuss a couple of failed attempts:*
+*We really struggled to accommodate a traditional Methods and Results section because any structure we attempted defied the definition of these sections.
+To maintain a logical flow that describes how and why to use the package, we landed on a mix of methods and case study results/discussion through the body of the paper.
+We first describe the "Model structure", then the "Core functions", then we describe and discuss the case study results in the "Interpretation" section (this section also includes a solution to bias generated by the standard sampling/analysis approach) and, finally, we move to the "Parameterisation" section where we provide some guidance on how to modify default settings to suit specific needs.
+The how to content still remains with this structure, however, it has been augmented with more case study results (and one new result) which should bring out the real-world implications of the package, plus it provides some meaning to several of the complex figures shown through the paper.*
 
-*1) We tried to start the methods section with the "Model Structure" section, as suggested, and shunt the figures to a "Results" section; this, however, broke the flow and left a gap in the methods section where we show the methods (functions) for running a simulation but do not show the plotting methods.*
+*Below we discuss a couple of failed attempts to force the content into Methods and Results sections:*
+
+*1) We tried to start the methods section with "Model Structure", as suggested, and shunt the figures to a "Results" section; this, however, broke the flow and left a gap in the methods section where we demonstrate the methods (functions) for running a simulation but not the plotting methods.*
 
 *2) We tried to keep the methods section as is (i.e.
-describe equations and functions) and describe some of the key case study results in the results section, but it was difficult to find a logical placement for the "Parameterisation" section and it did not seem right to segue to a solution to a problem identified in the case study in a traditional results section.
-In contrast, it seemed fine to describe and discuss the case study results demonstrated through the "Core functions" section and then jump to a solution to a problem in an independent section.*
+describe equations and functions) and describe some of the key case study results in the results section, but it was difficult to find a logical placement for the "Parameterisation" section (it seemed awkward to keep it in between the "Core functions" section and the "Results" section).
+It also seemed awkward to segue from a description of the case study results to a discussion that leads to a potential solution to a problem in a traditional "Results" section.*
+
+*3) Earlier iterations of the paper were focused entirely on the case study, the format was traditional and the package was a side-note in the methods section.
+It became apparent, however, that the package was more interesting and generally applicable than the case study results.
+We then began re-working the paper towards a vignette style and we used a couple papers as a guide on how to document a new R package (e.g.
+https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0092725).*
 
@@
 - see comment above
+
+*Ditto*
 
@@
 
+*In contrast to a Methods or Results section, a Discussion section was easier to accommodate.
+Here we lean on the case study results to reiterate the importance of planing a survey or testing an existing survey.
+We then use this preamble to segue into the "Research opportunities" section.*
+
 - I think something that can be useful for readers is to outline the steps in a thought process a researcher might undertake when setting up a survey (perhaps a separate subsection) and then how one might go about a SimSurvey run for this.
@@
 
+*As much as we would like to address each of these points and work towards a survey design handbook, it would be hard to be as cohesive and comprehensive as Cochran (1977; "Sampling techniques") or Sutherland (2006; "Ecological census techniques: a handbook").
+Moreover, we have yet to implement features that could help users assess some of these trade-offs (i.e.
+cost-benefit analysis).
+We are hopeful that these are questions that future versions of SimSurvey may be able to help address as more features are added to elevate it from a purely statistical toolbox to something that also considers the human element.*
+
 - Your "Assumptions" section should come down into the Discussion
+
+*Agreed, done.*
 
@@
 You might also try to say something about optimisation of design which Reviewer 2 mentioned.
+
+*Good point.
+We have noted in under the "sim_distribution" section that a custom closure can be created that uses randomfields to simulate spatial noise.
+We have also noted in the "Future directions" section that we hope to implement alternatives to random or stratified random designs to allow for more comprehensive evaluations of various designs.*

### 2020-02-18 analysis/paper/response_letter.Rmd
Tweak draft response letter

--- 2020-02-18 analysis/paper/rebuttal_letter.Rmd
@@
 We are also grateful for your detailed suggestions and we have made every effort to do justice to the changes you recommend.
-Most importantly, we hope we have added sufficient content to the core of the paper elevate the manuscript from solely a software manual to a primary scientific publication.
+Most importantly, we hope we have added sufficient content to the core of the paper elevate the manuscript from a software manual to a primary scientific publication.
 Though the how-to approach remains, we now see that describing some of the case study results in the core of the manuscript makes it more interesting and it adds another tangible reason for prospective users to learn how to use the package.
@@
 *Thank you for thinking beyond a fisheries biologist audience, as this is the audience we have been targeting from the onset.
-In case this is short-sighted, we have modified our introduction, as suggested, to pitch the concept to a broader audience.
-The reason we did not think about other biologists is because fisheries biologist tend to have access to age-based data while others (e.g.
+In hindsight we see that this could be short-sighted so we have modified our introduction, as suggested, to pitch the concept to a broader audience.
+The reason we did not think beyond fisheries biologist is because they tend to have access to age-based data while others (e.g.
 seabird and caribou biologists) do not.
-Alas, we agree that there could be biologists outside the field of fisheries that have such data or stage-based data who may find the package useful; if nothing else, the general framework may be useful to adapt to other marine or terrestrial situations.*
+Alas, there are probably exceptions to our perceptions; if nothing else, the general framework we present may be useful to adapt to other marine or terrestrial situations.*
 
@@
 
-*We really struggled to accommodate a traditional Methods and Results section because any structure we attempted defied the definition of these sections.
+*We really struggled to accommodate traditional Methods and Results sections because all of our re-structuring attempts defied the definitions of these sections.
 To maintain a logical flow that describes how and why to use the package, we landed on a mix of methods and case study results/discussion through the body of the paper.
 We first describe the "Model structure", then the "Core functions", then we describe and discuss the case study results in the "Interpretation" section (this section also includes a solution to bias generated by the standard sampling/analysis approach) and, finally, we move to the "Parameterisation" section where we provide some guidance on how to modify default settings to suit specific needs.
-The how to content still remains with this structure, however, it has been augmented with more case study results (and one new result) which should bring out the real-world implications of the package, plus it provides some meaning to several of the complex figures shown through the paper.*
+The how-to content still remains with this structure, however, it has been augmented with more case study results (and one new result) which should bring out the real-world implications of the package, plus it provides some meaning to several of the complex figures shown through the paper.*
 
@@
 
-*1) We tried to start the methods section with "Model Structure", as suggested, and shunt the figures to a "Results" section; this, however, broke the flow and left a gap in the methods section where we demonstrate the methods (functions) for running a simulation but not the plotting methods.*
+*1) We tried to start the methods section with "Model Structure", as suggested, and shunt the figures to a "Results" section; this, however, broke the flow and left a gap in the methods section which held demonstrations of the core simulation functions but not the plotting functions.*
 
-*2) We tried to keep the methods section as is (i.e.
-describe equations and functions) and describe some of the key case study results in the results section, but it was difficult to find a logical placement for the "Parameterisation" section (it seemed awkward to keep it in between the "Core functions" section and the "Results" section).
-It also seemed awkward to segue from a description of the case study results to a discussion that leads to a potential solution to a problem in a traditional "Results" section.*
+*2) We tried to keep the methods section as is and describe some of the case study results in the results section, but it was difficult to find a logical placement for the "Parameterisation" section (i.e.
+it seemed awkward to keep it in between the "Core functions" section and the "Results" section).
+With this structure it was also hard to find a sensible location for describing and discussing a solution to the age-length key problem.*
 
@@
 It became apparent, however, that the package was more interesting and generally applicable than the case study results.
-We then began re-working the paper towards a vignette style and we used a couple papers as a guide on how to document a new R package (e.g.
+We then began re-working the paper into a how-to format and, while moving away from a conventional format, we looked to several papers as a guide on how to document a R package  (e.g.
 https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0092725).*
@@
 *Good point.
-We have noted in under the "sim_distribution" section that a custom closure can be created that uses randomfields to simulate spatial noise.
+We have noted in the "sim_distribution" section that a custom closure can be created that uses randomfields to simulate spatial noise.
 We have also noted in the "Future directions" section that we hope to implement alternatives to random or stratified random designs to allow for more comprehensive evaluations of various designs.*

### 2020-02-18 analysis/paper/response_letter.md
Tweak draft response letter

--- 2020-02-17 analysis/paper/rebuttal_letter.Rmd
@@
----
-output:
-  word_document: default
-  md_document: default
----
-
 Daniel E. Duplisea, PhD  
@@
 
-`r format(Sys.time(), '%Y-%m-%d')`
+2020-02-17
 
@@
 
-Thank you for considering another revision of manuscript PONE-D-19-26904R1,  **"`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"** by Paul M.
-Regular, Gregory J.
-Robertson, Keith P.
-Lewis, Jonathan Babyn, Brian Healey and Fran Mowbray.
-We are also grateful for your detailed suggestions and we have made every effort to do justice to the changes you recommend.
-Most importantly, we hope we have added sufficient content to the core of the paper elevate the manuscript from solely a software manual to a primary scientific publication.
-Though the how-to approach remains, we now see that describing some of the case study results in the core of the manuscript makes it more interesting and it adds another tangible reason for prospective users to learn how to use the package.
-Please see below for more details on the changes we made in response to your suggestions.
+Thank you for considering another revision of manuscript
+PONE-D-19-26904R1, **“`SimSurvey`: an `R` package for comparing the
+design and analysis of fisheries surveys by simulating
+spatially-correlated fish stocks”** by Paul M. Regular, Gregory J.
+Robertson, Keith P. Lewis, Jonathan Babyn, Brian Healey and Fran
+Mowbray. We are also greatful for your detailed suggestions and we have
+made every effort to do justice to the changes you reccomend. Most
+importantly, we hope we have added sufficient content to the core of the
+paper elevate the manuscript from soley a software manual to a primary
+scientific publication. Though the how-to approach remains, we now see
+that describing some of the case study results in the core of the
+manuscript makes it more interesting and it adds another tangible reason
+for prospective users to learn how to use the package. Please see below
+for more details on the changes we made in response to your suggestions.
 
-We submit this revised manuscript for your consideration and look forward to your decision.
+We submit this revised manuscript for your consideration and look
+forward to your decision.
 
@@
 A1C 5X1, Canada  
-E-mail: Paul.Regular@dfo-mpo.gc.ca  
+E-mail:
+<a href="mailto:Paul.Regular@dfo-mpo.gc.ca" class="email">Paul.Regular@dfo-mpo.gc.ca</a>  
 Phone: (709) 772-2067
 
-----------------------------------------------------------------------------------------------------
+------------------------------------------------------------------------
 
@@
 
-Dear Paul, you have made lots of excellent changes that I think help with the use of the package and responded to the many specific comments of reviewers which has no doubt corrected many technical issues and clarifications.
-The one thing I would say that you have not really done is get at the deeper research merits of this work beyond introducing a new piece of software.
-I think paper, as it stands, lacks a larger context and content which is important for the primary publication.
-My diagnosis for why this is is that the manuscript does not conform very well to more typical scientific reports (Intro, M&M, Result, Discussion) which can make it difficult for readers to find the larger scientific merits of the work.
-It is useful of course for those who already understand the merits of this kind of work but this work is for primary publication and it needs to appeal more to the former than the latter group.
-There is a very "how-to" feel to it (e.g.
-line 64 "In this section") which I think detracts from getting at the larger purpose of the work.
+Dear Paul, you have made lots of excellent changes that I think help
+with the use of the package and responded to the many specific comments
+of reviewers which has no doubt corrected many technical issues and
+clarifications. The one thing I would say that you have not really done
+is get at the deeper research merits of this work beyond introducing a
+new piece of software. I think paper, as it stands, lacks a larger
+context and content which is important for the primary publication. My
+diagnosis for why this is is that the manuscript does not conform very
+well to more typical scientific reports (Intro, M&M, Result, Discussion)
+which can make it difficult for readers to find the larger scientific
+merits of the work. It is useful of course for those who already
+understand the merits of this kind of work but this work is for primary
+publication and it needs to appeal more to the former than the latter
+group. There is a very “how-to” feel to it (e.g. line 64 “In this
+section”) which I think detracts from getting at the larger purpose of
+the work.
 
-I would really like you to address this issue of moving it from a software manual to a primary scientific publication.
-I do not think it should involve that much work but there will be some restructuring of sections as well as places to put in content and bring out conclusions.
+I would really like you to address this issue of moving it from a
+software manual to a primary scientific publication. I do not think it
+should involve that much work but there will be some restructuring of
+sections as well as places to put in content and bring out conclusions.
 Here are my suggestions for this:
 
-Try to follow a more traditional paper structure.
-This will help readers and it likely will also make it clearer for you on how you can inject content into the paper to move it beyond the software manual approach:
+Try to follow a more traditional paper structure. This will help readers
+and it likely will also make it clearer for you on how you can inject
+content into the paper to move it beyond the software manual approach:
 
@@
 
-- You need to talk about wider issues and examples.
-e.g.
-examples of when poor survey design meant that scientific questions could not be properly addressed when good survey design meant they were.
-Examples of when good survey design allowed researchers to address needs that were unanticipated at the time it was designed.
-i.e.
-you need to build a better case (not just cost) of why survey design is really important and it is most powerful to do this with examples.
-You should try to bring in ideas related to ecosystem and climate changes and being able to track communities.
-Perhaps bring in species at risk ideas and tracking decline, you could bring in ideas related to MSC certification.
-These are just examples of specifics but you get the idea: the Introduction needs to have more general information outlining in both a broad and specific sense why we should be concerned about this.
+-   You need to talk about wider issues and examples. e.g. examples of
+    when poor survey design meant that scientific questions could not be
+    properly addressed when good survey design meant they were. Examples
+    of when good survey design allowed researchers to address needs that
+    were unanticipated at the time it was designed. i.e. you need to
+    build a better case (not just cost) of why survey design is really
+    important and it is most powerful to do this with examples. You
+    should try to bring in ideas related to ecosystem and climate
+    changes and being able to track communities. Perhaps bring in
+    species at risk ideas and tracking decline, you could bring in ideas
+    related to MSC certification. These are just examples of specifics
+    but you get the idea: the Introduction needs to have more general
+    information outlining in both a broad and specific sense why we
+    should be concerned about this.
 
-- Keep in mind an educated reader who may not be in fisheries but is interested in why anyone should care about this or could, for example, be interested in surveying say caribou or songbirds or something outside of fisheries but where many of the motivation and concepts may be similar and they are doing a more general literature search before designing their own survey.
-
-*Thank you for thinking beyond a fisheries biologist audience, as this is the audience we have been targeting from the onset.
-In the case that this is short-sighted, we have modified our introduction, as suggested, to pitch the concept to a broader audience.
-The reason we did not think about other biologists is because fisheries biologist tend to have access to age-based data while others (e.g.
-seabird and caribou biologists) do not.
-Alas, we agree that there could be biologists outside the field of fisheries that have such data or stage-based data who may find the package useful; if nothing else, the general framework may be useful to adapt to other marine or terrestrial situations.*
+-   Keep in mind an educated reader who may not be in fisheries but is
+    interested in why anyone should care about this or could, for
+    example, be interested in surveying say caribou or songbirds or
+    something outside of fisheries but where many of the motivation and
+    concepts may be similar and they are doing a more general literature
+    search before designing their own survey.
 
@@
 
-- This will start at your "Model Structure" section. You should put a higher level heading just before that called Methods.
+-   This will start at your “Model Structure” section. You should put a
+    higher level heading just before that called Methods.
 
 -   I can see that it is hard to separate your Methods from Results.
-Your results are really in the section "Using SimSurvey" I would not be opposed to putting this as a separate Results section but I leave it to you to decide.
-Essentially what you have is a case study as an example of how it works and therefore specific results are less interesting than how you got there with the package.
-You might title it something like "Results: running a SimSurvey simulation".
-This section also has a lot of content without a lot of explanation.
-for example, you have several large and complicated figures in a row.
-You should discuss not only what figures are meant to offer in the package but also what they mean.
-So for example on line 473 you refer to three figures (5,6,7) but you offer little interpretation of those figures just why you can make them which is another example of the limitations of the how-to manual approach.
-So you need to think of it as a case study and help someone decide the implications of their survey design.
-
-*We really struggled to accommodate a traditional Methods and Results section as any structure we attempted betrayed the definition of these sections.
-----DESCRIBE WHAT WE DID HERE----.
-Below we discuss a couple of failed attempts:*
-
-*1) We tried to start the methods section with the "Model Structure" section, as suggested, and shunt the figures to a "Results" section; this, however, broke the flow and left a gap in the methods section where we show the methods (functions) for running a simulation but do not show the plotting methods.*
-
-*2) We tried to keep the methods section as is (i.e.
-describe equations and functions) and describe some of the key case study results in the results section, but it was difficult to find a logical placement for the "Parameterisation" section and it did not seem right to segue to a solution to a problem identified in the case study in a traditional results section.
-In contrast, it seemed fine to describe and discuss the case study results demonstrated through the "Core functions" section and then jump to a solution to a problem in an independent section.*
+    Your results are really in the section “Using SimSurvey” I would not
+    be opposed to putting this as a separate Results section but I leave
+    it to you to decide. Essentially what you have is a case study as an
+    example of how it works and therefore specific results are less
+    interesting than how you got there with the package. You might title
+    it something like “Results: running a SimSurvey simulation”. This
+    section also has a lot of content without a lot of explanation. for
+    example, you have several large and complicated figures in a row.
+    You should discuss not only what figures are meant to offer in the
+    package but also what they mean. So for example on line 473 you
+    refer to three figures (5,6,7) but you offer little interpretation
+    of those figures just why you can make them which is another example
+    of the limitations of the how-to manual approach. So you need to
+    think of it as a case study and help someone decide the implications
+    of their survey design.
 
@@
 
-- This starts just before "Research Opportunities".
-You can keep these sections but there should be more preamble before jumping right into research opportunities.
-I suggest a general paragraph(s) that segue into your subsections of "research opportunities", "future research".
-In the Discussion you should address again some of the broader issues from the Introduction, e.g.
-how could spatial (depth) distribution changes anticipated under climate change for some population be tackled by survey design exploration now so that we can continue to track these changing populations 20 years down the road and do not lose the signal.
-How can this software help with that?
+-   This starts just before “Research Opportunities”. You can keep these
+    sections but there should be more preamble before jumping right into
+    research opportunities. I suggest a general paragraph(s) that segue
+    into your subsections of “research opportunities”, “future
+    research”. In the Discussion you should address again some of the
+    broader issues from the Introduction, e.g. how could spatial (depth)
+    distribution changes anticipated under climate change for some
+    population be tackled by survey design exploration now so that we
+    can continue to track these changing populations 20 years down the
+    road and do not lose the signal. How can this software help with
+    that?
 
-- I think something that can be useful for readers is to outline the steps in a thought process a researcher might undertake when setting up a survey (perhaps a separate subsection) and then how one might go about a SimSurvey run for this.
-It also gives you a good opportunity to discuss your multispecies ideas:
+-   I think something that can be useful for readers is to outline the
+    steps in a thought process a researcher might undertake when setting
+    up a survey (perhaps a separate subsection) and then how one might
+    go about a SimSurvey run for this. It also gives you a good
+    opportunity to discuss your multispecies ideas:
     -   What is your current problem/needs
@@
     -   Resources available
-    - What are your anticipated future needs (i.e. if climate change is going to make cod go deeper, will you be able to track it in 20 years?)
+    -   What are your anticipated future needs (i.e. if climate change
+        is going to make cod go deeper, will you be able to track it in
+        20 years?)
     -   How could you consider some or all of these with SimSurvey
+-   Your “Assumptions” section should come down into the Discussion
 
-- Your "Assumptions" section should come down into the Discussion
-
-- "Future Directions" should say something about randomfields re: Reviewer 1.
-Even if you just outline that you have considered it.
-You might also try to say something about optimisation of design which Reviewer 2 mentioned.
+-   “Future Directions” should say something about randomfields re:
+    Reviewer 1. Even if you just outline that you have considered it.
+    You might also try to say something about optimisation of design
+    which Reviewer 2 mentioned.

### 2020-02-19 analysis/paper/response_letter.Rmd
Integrate Greg and Keith's edits to the manuscript

@@
 We are also grateful for your detailed suggestions and we have made every effort to do justice to the changes you recommend.
-Most importantly, we hope we have added sufficient content to the core of the paper elevate the manuscript from a software manual to a primary scientific publication.
+Most importantly, we hope we have added sufficient content to the core of the paper to elevate the manuscript from a software manual to a primary scientific publication.
 Though the how-to approach remains, we now see that describing some of the case study results in the core of the manuscript makes it more interesting and it adds another tangible reason for prospective users to learn how to use the package.

### 2020-02-19 analysis/paper/sim_survey_paper.Rmd
Integrate Greg and Keith's edits to the manuscript

@@
 ---
-title: "`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
+title: "`SimSurvey`: an `R` package for comparing the design and analysis of surveys by simulating spatially-correlated populations"
 output: 
@@
 
-Fish populations often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective surveys.
+Populations often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective surveys.
 Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics).
-For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of fish captured to measure biological characteristics (e.g.
+For many large-scale monitoring programs, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of individuals to measure biological characteristics (e.g.
 lengths or ages).
 Population estimates are also affected by the pathway taken to analyze such data.
-Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for fisheries-independent survey data.
-Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations.
+Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for multi-stage survey data.
+Here we introduce the R package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed populations.
 The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of built-in or user-defined sampling protocols to reproduce the population parameters of the known population.
@@
 We demonstrate the use of this package using a case study and show that it can reveal unexpected sources of bias and be used to explore design-based solutions to such problems.
-In summary, **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring.
+In summary, **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks and other populations that show complex structuring.
 Various statistical approaches can then be applied to the results to test the efficacy of different analytical approaches.
@@
 Good survey design is a critical prerequisite for obtaining accurate and precise estimates of a population of interest.
-The need for sound sampling techniques is as relevant today as it was in the 1930's when much of the theory and principles behind the collection of scientific data were laid by statisticians such as R.A.
+The need for sound sampling techniques is as relevant today as it was in the 1930's when much of the theory and principles behind the collection of scientific data were developed by statisticians such as R.A.
 Fisher and Jerzy Neyman [@fienberg1996].
@@
 
-One of the biggest challenges associated with surveys is the costs of planning and implementing effective and efficient sampling designs [@reynolds2011].
-Despite the costs, surveys have become a mainstay in the management of dynamic fish stocks as survey data are not easily replaced by other sources of information [@Pennington1998].
+One of the biggest challenges associated with surveys is the cost of planning and implementing effective and efficient sampling designs [@reynolds2011].
+Despite their costs, surveys have become a mainstay in the management of dynamic fish stocks as survey data are not easily replaced by other sources of information [@Pennington1998].
 For instance, large population declines can be overlooked by relying solely on catch and effort data from commercial fisheries [@rose1999].
@@
 In such situations it is common for practitioners to resort to simulations to test various sampling and analytical pathways (e.g.
-[@thomas2010]), however, such tests are rare in fisheries research (but see [@schnute2003; @puerta2019]), perhaps because of the complexities associated with fisheries-independent surveys.
+[@thomas2010]).
+However, such tests are rare in fisheries research (but see [@schnute2003; @puerta2019]), perhaps because of the complexities associated with fisheries-independent surveys.
 
@@
 The package has two main components: the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated population distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.
+Although constructed to assess the design of fisheries-independent trawl surveys, this package can be used to model any population that shows complex spatial and age structure.
 
 This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al.
-[-@puerta2019], however, efforts were focused on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways.
+[-@puerta2019], however, we focused our efforts on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways.
 The steps taken to simulate surveys of spatial, age-structure populations are outlined below.
-We first outline the equations underlying the package (**[Model structure]** section) and then move into a vignette (**[Core functions]** section).
-The core functions of the package are largely demonstrated using default settings, and these settings are based on a case study (see for **[S1 Appendix][S1 Appendix: Case study]** details).
-Several of the results from the case study are described and discussed in the **[Interpretation]** section as they highlight the utility of the package.
-Finally, we provide some guidance on how to modify default settings to suit specific needs (**[Parameterisation]** section) and then move onto discuss the potential uses and future directions of the package (**[Discussion]** section).
+We first outline the equations underlying the package (**[Model structure]** section) and then we demonstrate how to use its core functions (**[Core functions]** section).
+The core functions of the package are largely demonstrated using default settings, and these settings are based on a case study (see for **[S1 Appendix][S1 Appendix: Case study]** details on the case study, and see **[S2 Appendix][S2 Appendix: Parameterisation]** for guidance on how to modify default settings to suit specific needs).
+Several of the results from the case study are described and discussed in the **[Interpretation]** section as they highlight one use case of the package.
+Finally, we discuss the broader research opportunities and future directions of the package (**[Discussion]** section).
 
@@
 The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space.
-This noise term is scaled to sum to 1 to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model:
-
-$$\begin{aligned}N_{a,y,s} &= N_{a,y} \frac{\eta_{a,y,s}}{\sum_{s = 1}^{N_s} \eta_{a,y}} \\ \eta_{a,y,s} &= \frac{(d_s - \mu_{d}) ^ 2}{2 \sigma_d^2} + \xi_{a,y,s} \end{aligned}$$
+This noise term, $\eta_{a,y,s}$, is scaled to sum to 1 to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model:
+
+$$\begin{aligned}N_{a,y,s} &= N_{a,y} \frac{\eta_{a,y,s}}{\sum_{s = 1}^{N_s} \eta_{a,y,s}} \\ \eta_{a,y,s} &= \frac{(d_s - \mu_{d}) ^ 2}{2 \sigma_d^2} + \xi_{a,y,s} \end{aligned}$$
 
@@
 The rate at which point-to-point spatial correlation decays with distance is controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) (here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015]) and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively.
-The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see **[S2 Appendix][S2 Appendix: Age-year-space covariance]** for a more detailed description of the space-age-year covariance structure).
+The overall variance of the spatial process is controlled by $\sigma_{\xi}^2$ (see **[S3 Appendix][S3 Appendix: Age-year-space covariance]** for a more detailed description of the space-age-year covariance structure).
 In short, this formulation allows control of depth preferences, the level of spatial aggregation and the degree of age and year specific clustering.
@@
 [@Thorson2015]), design-based approaches, such as stratified analyses, are often used.
-Here we apply formula presented in Smith and Somerton [-@Smith1981] (equations are replicated in **[S3 Appendix][S3 Appendix: Stratified analysis equations]**) to calculate year $y$ and simulation $j$ specific stratified estimates of total abundance ($\hat{I}_{y,j}$), abundance at length ($\hat{I}_{l,y,j}$) and abundance at age ($\hat{I}_{a,y,j}$).
+Here we apply formulae presented in Smith and Somerton [-@Smith1981] (equations are replicated in **[S4 Appendix][S4 Appendix: Stratified analysis equations]**) to calculate year $y$ and simulation $j$ specific stratified estimates of total abundance ($\hat{I}_{y,j}$), abundance at length ($\hat{I}_{l,y,j}$) and abundance at age ($\hat{I}_{a,y,j}$).
 Note that estimates of total abundance are based on total numbers caught at each set, $n_i$, while abundance at length requires the sub-sampled length frequencies at each set, $m_{l, i}$, to be scaled up using set-specific ratios of the number of fish measured, $m_{i}$, to numbers caught, $n_i$:
@@
 | **`sim_abundance`**  | Simulate a basic age-structured population dynamics model |
-| `sim_R`, `sim_Z`, `sim_N0`, `sim_vonB` | Closures, to use inside `sim_abundance`, for simulating recruitment, total mortality, initial abundance and growth, respectively | 
+| `sim_R`, `sim_Z`, `sim_N0`, `sim_vonB` | Closures, to use inside `sim_abundance`, for simulating recruitment, total mortality, initial abundance, and growth, respectively | 
 | **`sim_distribution`** | Simulate spatial and temporal distribution of an age-structured population |
@@
 However, for simplicity, we use `make_grid` to construct a square grid for a default run of `sim_distribution`.
-Below we generate and plot (Figure \@ref(fig:grid)) a default grid, another grid with the number of strata increased by increasing the number of `strat_splits`, and another with four divisions and a linear depth gradient (the sigmoid curve is forced to be linear when `shelf_width` is set to zero).
+Below we generate and plot (Figure \@ref(fig:grid)) a default grid, another grid with the number of strata increased by increasing the number of `strat_splits`, and another with the number of divisions increased using `n_div` and a linear depth gradient (the sigmoid curve is forced to be linear when `shelf_width` is set to zero).
 
@@
 
-```{r grid, out.height = "320px", fig.cap = "Plots produced by `plot_grid` when supplied a raster object produced by `make_grid` a) using default settings, b) settings that increase the number of times depth strata are split, and c) settings that produce a more linear depth gradient and increase the number of divisions.
+```{r grid, out.height = "320px", fig.cap = "Plots produced by `plot_grid` when supplied a raster object produced by `make_grid` a) using default settings, b) settings that increase the number of times depth strata are split (`strat_splits = 3`), and c) settings that produce a more linear depth gradient (`shelf_width = 0`) and increase the number of divisions (`n_div = 4`).
 In these plots, the yellow to purple color gradient represents depth, the thick grey lines delineate divisions and thin white lines delineate strata."}
@@
 Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by adjusting a few parameters in these closures (Figure \@ref(fig:distribution)).
-Though we only provide one option for simulating a spatial noise (see **[S2 Appendix][S2 Appendix: Age-year-space covariance]** for details), advance users could construct closures that leverage simulation models provided by packages such as **`RandomFields`** [@schlather2020] or **`INLA`** [@rue2009].
+Though we only provide one option for simulating a spatial noise (see **[S3 Appendix][S3 Appendix: Age-year-space covariance]** for details), advanced users could construct closures that leverage simulation models provided by packages such as **`RandomFields`** [@schlather2020] or **`INLA`** [@rue2009].
 Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e.
@@
 a <- sim_distribution(sim = sim_abundance(), # nested approach
-                      ays_covar = sim_ays_covar(range = 300,
+                      ays_covar = sim_ays_covar(range = 300, # clustered
                                                 phi_year = 0.9,
@@
 b <- sim_abundance() %>%  # pipe approach
-  sim_distribution(ays_covar = sim_ays_covar(range = 2000,
+  sim_distribution(ays_covar = sim_ays_covar(range = 2000, # diffuse
                                              phi_year = 0.2,
@@
 A default function call is described in Table \@ref(tab:survey).
-The `sim_survey` function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and definitions of the design of the survey.
+The `sim_survey` function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and survey protocol settings.
 Specifically, the `q` argument requires a closure, such as `sim_logistic`, for defining the probability of catching specific age groups, trawl dimensions are defined in the `trawl_dim` argument, and set, length and age sampling effort are defined using the `set_den`, `lengths_cap` and `ages_cap` arguments, respectively.
@@
 Objects produced by `sim_survey` (set and sampling details) and `run_strat` (full stratified analysis results) are not retained to minimize object size.
+
 Like other core functions, some convenience functions are included in **`SimSurvey`** for creating interactive plots of the results from `test_surveys`.
@@
 
-```{r rank, out.height = "400px", fig.cap = "Divergent dot plot of the precision and accuracy (RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{\\mathrm{sets}}$] and length measurements [$N_{\\mathrm{measured}}$]), under various sampling protocols (set density [$D_{\\mathrm{sets}}$] and maximum number of lengths measured per set [$M_{\\mathrm{lengths}}$]).
+```{r rank, out.height = "400px", fig.cap = "Divergent dot plot of the precision and accuracy (root-mean-squared error; RMSE) of length based stratified estimates of abundance, and total sampling effort (number of sets [$N_{\\mathrm{sets}}$] and length measurements [$N_{\\mathrm{measured}}$]), under various sampling protocols (set density [$D_{\\mathrm{sets}}$] and maximum number of lengths measured per set [$M_{\\mathrm{lengths}}$]).
 Records are ranked by lowest to highest RMSE score.
@@
 ```{r}
-cap <- "Surface plots of RMSE from an array of surveys with different sampling protocol.
+cap <- "Surface plots of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol.
 Panels represent surveys with different set densities ($D_{\\mathrm{sets}}$), x-axes represent the maximum sampling effort of lengths per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per length group ($M_{\\mathrm{ages}}$).
@@
 
-Results shown in the **[Core functions]** section demonstrates one use case of the **`SimSurvey`** package that is focused on evaluating the efficacy of increasing the sampling effort of a stratified random survey of a cod population (see **[S1 Appendix][S1 Appendix: Case study]** for details).
+Results shown in the **[Core functions]** section demonstrates one use of the **`SimSurvey`** package that is focused on evaluating the efficacy of increasing the sampling effort of a stratified random survey of a cod population (see **[S1 Appendix][S1 Appendix: Case study]** for details).
 These results largely align with expectations from sampling theory that 1) design-based estimators for stratified random surveys are unbiased, and 2) precision is increased by increasing the number of primary sampling units [@Cochran1977].
@@
 Bias, in particular, appears to be a problem with stratified estimates of abundance at age under certain scenarios (Figure \@ref(fig:fan3)); in contrast, there appears to be little bias in the stratified estimates of abundance at length (Figure \@ref(fig:fan2)).
-This result was unexpected as the analysis was presumed to be unbiased.
+This result was unexpected as these estimates of abundance at age were presumed to be unbiased.
 The contrast between the length and age based analysis indicates that the problem lies with the intervening age-length-key.
@@
 
-```{r alt, out.height = "600px", fig.cap = "Key results from an alternative design-based sampling and analysis procedure that accounts for multi-stage cluster sampling of ages.
-Specifically, a) is a fan chart of stratified estimates of abundance at age from year seven of the simulation from a survey with a set density ($D_{\\mathrm{sets}}$) of 0.002 sets / km^2^, maximum number of length samples ($M_{\\mathrm{lengths}}$) of 100, and maximum number of age samples per 3 cm length group per set ($M_{\\mathrm{ages}}$) of 3 (the thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%); and b) is a surface plot of RMSE from an array of surveys with set density ($D_{\\mathrm{sets}}$) fixed to 0.002 sets / km^2^ but different sub-sampling effort, where the x-axes represent the maximum number of lengths to measure per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per 3 cm length group per set ($M_{\\mathrm{ages}}$)."}
+```{r alt, out.height = "600px", fig.cap = "Main results from an alternative design-based sampling and analysis procedure that accounts for multi-stage cluster sampling of ages.
+Specifically, a) is a fan chart of stratified estimates of abundance at age from year seven of the simulation from a survey with a set density ($D_{\\mathrm{sets}}$) of 0.002 sets / km^2^, maximum number of length samples ($M_{\\mathrm{lengths}}$) of 100, and maximum number of age samples per 3 cm length group per set ($M_{\\mathrm{ages}}$) of 3 (the thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%); and b) is a surface plot of root-mean-squared error (RMSE) from an array of surveys with set density ($D_{\\mathrm{sets}}$) fixed to 0.002 sets / km^2^ but different sub-sampling effort, where the x-axes represent the maximum number of lengths to measure per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per 3 cm length group per set ($M_{\\mathrm{ages}}$)."}
 
@@
 This is an important result with real-world implications as they imply that 1) inference from the actual survey can be improved by altering the sampling and analysis of ages and, 2) valuable survey time can be saved by cutting back on sub-sampling effort.
-This example shows how **`SimSurvey`** can be used to reveal problems and explore solutions to the sampling designs of fisheries-independent surveys.
-
-# Parameterisation
-
-The default parameter values used in this package were chosen to emulate survey data from a specific cod population  (see **[S1 Appendix][S1 Appendix: Case study]** for details) and, as such, these values ought to provide a good place to start for another population of cod or similar groundfish.
-In this section we outline a series of recommended steps to take in order to tailor the default settings to suit other systems.
-Throughout this outline, we also provide some context on how default parameter values were chosen as this information may help guide a users’ choices.
-
-## 1. Define survey area and protocol
-
-Parameterising the survey design should be the easiest place to start as the area and protocol should be clearly defined.
-The survey area can be defined using `make_grid` or it can be manually constructed.
-For the case study, `make_grid` arguments were modified to create a grid with similar dimensions to the stratified design actually applied to the case study area (i.e.
-total area, shelf depth, number of strata, etc.).
-Also note that the resolution (`res` argument) of the grid was defined to be similar to the resolution at which sets are allocated in the survey.
-A square grid was chosen in lieu of the actual survey grid to create an area that avoids potential issues associated with land as a barrier while being sufficiently complex to test survey designs.
-When setting up a grid, be careful not to drastically increase the dimensions (`x_range` or `y_range` arguments) or resolution (`res` argument) as simulating the spatial noise over a large field can be computationally demanding.
-
-With a survey grid defined, the settings of `sim_survey` can be modified to match the protocol of a users' survey of interest.
-For the case study, the trawl dimensions (`trawl_dim`) were modified to match the dimensions of a standard trawl conducted in the region.
-Likewise, area is used to define the number of sets to allocate to a strata and the value used by default for `set_den` matches the protocol, as does the `min_sets` setting (i.e.
-at least two sets are conducted in each strata to facilitate stratified analyses).
-Finally, length and age sub-sampling settings were modified to match the protocol for the survey of cod in the case study area.
-The most difficult parameter to define in the `sim_survey` function is `q`.
-A logistic curve is used by default to define these unknown values and the `sim_logistic` parameters were tweaked to approximate the catchability curve supplied to the survey-based model that was used to assess the case study population [@ings2019].
-A similar tactic can be applied to other cases with estimates of survey catchability.
-
-## 2. Define abundance and growth parameters
-
-Abundance and growth parameters are more difficult to define than the survey protocol as the true values are unknown.
-Nevertheless, spatially-aggregated age-structured models are frequently applied to assess fish populations and estimates from such models can be used to define `sim_abundance` argument settings.
-Alternatively, key values from these models (e.g.
-Z and N matrices) can be used instead of simulated values.
-Default parameters used in `sim_abundance` were largely inspired by estimates from the survey-based model used to assess the stock [@ings2019].
-Based on this work, average total mortality was roughly centered around 0.5 ($\mu_Z$; argument `log_mean` in `sim_Z` closure) and we allowed it to make similar changes in magnitude through time and across ages by setting the standard deviation of log total mortality to 0.2 ($\sigma_Z$; argument `log_sd` in `sim_Z` closure) and the correlations to 0.9 and 0.5 in the age and year dimensions ($\varphi_{\delta, \mathrm{age}}$ and $\varphi_{\delta, \mathrm{year}}$; `phi_age` and `phi_year`), respectively.
-Likewise, `sim_R` values were modified to generate similar patterns in recruitment as those estimated by the model.
-Ages groups to simulate across ranged from age 1 to 20 ($a$; `ages` argument) to capture the tails of the population, and 20 years were simulated ($y$; `years` argument) to capture some dynamics.
-Finally, von Bertalanffy growth parameters were roughly based on estimates from cod in the region [@Cadigan2016b].
-While the choices made are obviously imperfect and ad hoc, it is not necessary to generate a perfect representation of the population as the goal here is to generate a dynamic population to survey.
-What is key is that we work with a known population from which we can evaluate the performance of specific survey designs and analyses.
-
-## 3. Define spatial parameters
-
-Ideally this step would be no more difficult than step 2 as the inputs required could come from a spatially-explicit age-structured model fit to real data.
-Such estimates, however, are rare and we can confirm that no such model has been developed and fit to data from our case study population.
-Unless a user has access to such estimates, parameter estimates will need to manually defined using visual comparisons of real and simulated data.
-A user will have a first-cut simulation of some survey data by following steps 1 and 2, and the plotting functions provided with the package will provide a means of visually exploring these data.
-For comparison purposes, actual survey data will need to be processed and visual patterns in the spatial distribution of each age will need to be inspected.
-It was in this step that we noticed that ages 1 to 4 tend to occupy different areas while ages 5+ occupy similar space for the cod population we focused on.
-This observation motivated the coupling of the spatial noise of ages 5+ (`group_ages` argument in the `sim_ays_covar` closure), which forces this component of the population to occupy the same space, and the moderate value of 0.5 for the correlation in the spatial distribution across ages ($\varphi_{\xi, \mathrm{age}}$; `phi_age` argument in the `sim_ays_covar` closure).
-Year-to-year plots of the real data also revealed considerable inertia in the space occupied by each age group; this observation motivated a fairly high value of 0.9 for the correlation in the spatial distribution across years ($\varphi_{\xi, \mathrm{year}}$; `phi_year` argument in the `sim_ays_covar` closure).
-Finally, the range of the spatial correlation ($r$; `range` argument in the `sim_ays_covar` closure) was iteratively modified until the simulated population exhibited a similar level of patchiness as the real data, and the standard deviation of the age-year-space process ($\sigma_{\xi}$; `sd` argument in the `sim_ays_covar` closure) was increased to generate set catches with tails similar to those observed in the real data (cod catches in the area tend to be zero-inflated and heavy-tailed).
-Finally, visual inspection of the real data revealed that the largest catches tended to be centered on sets conducted in 200 m of water; a parabolic relationship, defined using the `sim_parabola` closure, with a mean of 200 m and standard deviation of 70 m approximated this pattern.
+This example demonstrates how **`SimSurvey`** can be used to reveal problems and explore solutions to the sampling designs of fisheries-independent surveys.
 
@@
 The case study used here provides one example of how **`SimSurvey`** can be used to simulation test the design of fisheries-independent trawl surveys.
-There are multiple layers to the sampling design of such surveys and the end results are linked with the way such data are analyzed.
+Default settings can, of course, be modified to emulate data from an array of different surveys of different populations (see **[S2 Appendix][S2 Appendix: Parameterisation]** for guidance on how to modify default settings to suit specific needs).
+The package therefore facilitates the exploration of a range of research questions.
 Below we outline some examples where the **`SimSurvey`** package may aid future research efforts.
@@
 
-The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formula (e.g.
+The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formulae (e.g.
 [@Cochran1977]).
@@
 Together, two functions (`sim_abundance` and `sim_distribution`) are capable of simulating a wide range of populations with different life histories, depth associations and spatial properties.
-The next necessary steps to generating data similar to actual observations is to conduct a survey.
+The next step to generating data similar to actual observations is to conduct a survey.
 In this package we implement a function, `sim_survey`, that conducts a survey of the population.
@@
 As a tangible first step, we focused our efforts on emulating data from the survey of cod in NAFO Subdivision 3Ps since it is one of the most dynamic and heavily sampled stocks in the region.
-Function settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data; these values were set as the defaults for the main functions in the package.
+Function settings were iteratively modified until it was difficult to distinguish the simulated data from real survey data; these values were set as the defaults for the main functions in the package (see **[S2 Appendix][S2 Appendix: Parameterisation]** for more details on how these defaults were chosen and for guidance on how to modify these settings to suit specific needs).
 Here we evaluate the efficacy of alternate sampling protocols by assessing deviation of stratified estimates of abundance [@Smith1981] from the true abundance available to the survey.
@@
 
-Bias in stratified estimates of abundance at age appears to disappear when length-stratified age sampling is conducted and age-length keys are applied on a set-by-set basis (Figure \@ref(fig:alt)a).
+Bias in stratified estimates of abundance at age is minimal when length-stratified age sampling is conducted and age-length keys are applied on a set-by-set basis (Figure \@ref(fig:alt)a).
 Under this scenario, there also appears to be a plateau, rather than a valley, in RMSE values as length and age sampling effort is increased (Figure \@ref(fig:alt)b).
@@
 
-# S2 Appendix: Age-year-space covariance
+# S2 Appendix: Parameterisation
+
+The default parameter values used in this package were chosen to emulate survey data from a specific cod population  (see **[S1 Appendix][S1 Appendix: Case study]** for details) and, as such, these values ought to provide a reasonable place to start for another population of cod or similar groundfish.
+In this section we outline a series of recommended steps to take in order to tailor the default settings to suit other systems.
+Throughout this outline, we also provide some context on how default parameter values were chosen as this information may help guide a users’ choices.
+
+## 1. Define survey area and protocol
+
+Parameterising the survey design should be the easiest place to start as the area and protocol should be clearly defined.
+The survey area can be defined using `make_grid` or it can be manually constructed.
+For the case study, `make_grid` arguments were modified to create a grid with similar dimensions to the stratified design actually applied to the case study area (i.e.
+total area, shelf depth, number of strata, etc.).
+Also note that the resolution (`res` argument) of the grid was defined to be similar to the resolution at which sets are allocated in the survey.
+A square grid was chosen in lieu of the actual survey grid to create an area that avoids potential issues associated with land as a barrier while being sufficiently complex to test survey designs.
+When setting up a grid, be careful not to drastically increase the dimensions (`x_range` or `y_range` arguments) or resolution (`res` argument) as simulating the spatial noise over a large field can be computationally demanding.
+
+With a survey grid defined, the settings of `sim_survey` can be modified to match the protocol of a users' survey of interest.
+For the case study, the trawl dimensions (`trawl_dim`) were modified to match the dimensions of a standard trawl conducted in the region.
+Likewise, area is used to define the number of sets to allocate to a strata and the value used by default for `set_den` matches the protocol, as does the `min_sets` setting (i.e.
+at least two sets are conducted in each strata to facilitate stratified analyses).
+Finally, length and age sub-sampling settings were modified to match the protocol for the survey of cod in the case study area.
+The most difficult parameter to define in the `sim_survey` function is `q`.
+A logistic curve is used by default to define these unknown values and the `sim_logistic` parameters were tweaked to approximate the catchability curve supplied to the survey-based model that was used to assess the case study population [@ings2019].
+A similar tactic can be applied to other cases with estimates of survey catchability.
+
+## 2. Define abundance and growth parameters
+
+Abundance and growth parameters are more difficult to define than the survey protocol as the true values are unknown.
+Nevertheless, spatially-aggregated age-structured models are frequently applied to assess fish populations and estimates from such models can be used to define `sim_abundance` argument settings.
+Alternatively, key values from these models (e.g.
+Z and N matrices) can be used instead of simulated values.
+Default parameters used in `sim_abundance` were largely inspired by estimates from the survey-based model used to assess the cod stock the case study was based on [@ings2019].
+Based on this work, average total mortality was roughly centered around 0.5 ($\mu_Z$; argument `log_mean` in `sim_Z` closure) and we allowed it to make similar changes in magnitude through time and across ages by setting the standard deviation of log total mortality to 0.2 ($\sigma_Z$; argument `log_sd` in `sim_Z` closure) and the correlations to 0.9 and 0.5 in the age and year dimensions ($\varphi_{\delta, \mathrm{age}}$ and $\varphi_{\delta, \mathrm{year}}$; `phi_age` and `phi_year`), respectively.
+Likewise, `sim_R` values were modified to generate similar patterns in recruitment as those estimated by the model.
+We simulated data for ages 1 to 20 ($a$; `ages` argument) to capture the tails of the population, and 20 years were simulated ($y$; `years` argument) to capture some dynamics.
+Finally, von Bertalanffy growth parameters were roughly based on estimates from cod in the region [@Cadigan2016b].
+While the choices made are obviously imperfect and ad hoc, it is not necessary to generate a perfect representation of the population as the goal here is to generate a dynamic population to survey.
+What is key is that a known population is simulated from which the performance of specific survey designs and analyses can be evaluated.
+
+## 3. Define spatial parameters
+
+Ideally this step would be no more difficult than step 2 as the inputs required could come from a spatially-explicit age-structured model fit to real data.
+Such estimates, however, are rare and we can confirm that no such model has been developed and fit to data from our case study population.
+Unless a user has access to such estimates, parameter estimates will need to be manually defined using visual comparisons of real and simulated data.
+A user will have an initial simulation of some survey data by following steps 1 and 2, and the plotting functions provided with the package will provide a means of visually exploring these data.
+For comparison purposes, actual survey data will need to be processed and visual patterns in the spatial distribution of each age will need to be inspected.
+It was in this step that we noticed that ages 1 to 4 tend to occupy different areas while ages 5+ occupy similar space for the cod population we focused on.
+Observing these patterns in real data motivated the coupling of the spatial noise of ages 5+ (`group_ages` argument in the `sim_ays_covar` closure), which forces this component of the population to occupy the same space, and the moderate value of 0.5 for the correlation in the spatial distribution across ages ($\varphi_{\xi, \mathrm{age}}$; `phi_age` argument in the `sim_ays_covar` closure).
+Year-to-year plots of the real data also revealed considerable inertia in the space occupied by each age group; this observation motivated a fairly high value of 0.9 for the correlation in the spatial distribution across years ($\varphi_{\xi, \mathrm{year}}$; `phi_year` argument in the `sim_ays_covar` closure).
+Finally, the range of the spatial correlation ($r$; `range` argument in the `sim_ays_covar` closure) was iteratively modified until the simulated population exhibited a similar level of patchiness as the real data, and the standard deviation of the age-year-space process ($\sigma_{\xi}$; `sd` argument in the `sim_ays_covar` closure) was increased to generate set catches with tails similar to those observed in the real data (cod catches in the area tend to be zero-inflated and heavy-tailed).
+Finally, visual inspection of the real data revealed that the largest catches tended to be centered on sets conducted in 200 m of water; a parabolic relationship, defined using the `sim_parabola` closure, with a mean of 200 m and standard deviation of 70 m approximated this pattern.
+
+# S3 Appendix: Age-year-space covariance
 
@@
 
-# S3 Appendix: Stratified analysis equations
+# S4 Appendix: Stratified analysis equations
 

### 2020-02-20 analysis/paper/sim_survey_paper.Rmd
Tweak abstract (bring back some signs that the package is fish-focused)

@@
 Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics).
-For many large-scale monitoring programs, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of individuals to measure biological characteristics (e.g.
-lengths or ages).
-Population estimates are also affected by the pathway taken to analyze such data.
+These issues can be difficult to identify and avoid in sample surveys of fish populations as they tend to be costly and comprised of multiple levels of sampling.
+Population estimates are therefore affected by each level of sampling as well as the pathway taken to analyze such data.
 Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for multi-stage survey data.

### 2020-02-20 analysis/paper/response_letter.Rmd
Revise response letter given Greg and Keith's feedback

--- 2020-02-18 analysis/paper/response_letter.Rmd
@@
 We are also grateful for your detailed suggestions and we have made every effort to do justice to the changes you recommend.
-Most importantly, we hope we have added sufficient content to the core of the paper elevate the manuscript from a software manual to a primary scientific publication.
-Though the how-to approach remains, we now see that describing some of the case study results in the core of the manuscript makes it more interesting and it adds another tangible reason for prospective users to learn how to use the package.
+Most importantly, we hope we have added sufficient content to the core of the paper to alleviate concerns that the manuscript is simply a software manual and further elevate the paper to be a suitable primary scientific publication.
+Though the how-to approach remains, we now see that describing some of the case study results in the core of the manuscript makes it more interesting and it adds another tangible reason for prospective users to ask survey design questions relevant to the their study system by learning how to use the package.
 Please see below for more details on the changes we made in response to your suggestions.
@@
 *Thank you for thinking beyond a fisheries biologist audience, as this is the audience we have been targeting from the onset.
-In hindsight we see that this could be short-sighted so we have modified our introduction, as suggested, to pitch the concept to a broader audience.
-The reason we did not think beyond fisheries biologist is because they tend to have access to age-based data while others (e.g.
-seabird and caribou biologists) do not.
-Alas, there are probably exceptions to our perceptions; if nothing else, the general framework we present may be useful to adapt to other marine or terrestrial situations.*
+In hindsight we see that was short-sighted so we have modified our introduction, as suggested, to pitch the concept to a broader audience.
+The reason we did not think beyond fisheries biology is because we tend to have access to extensive age-based data while others (e.g.
+songbird and caribou biologists) do not.
+However, there are certainly other marine or terrestrial situations that could use our approach, the general framework we now present should be useful to adapt our work to systems we have not even considered.*
+
+*Specifically, we have added an additional paragraph at the start of the paper discussing the importance of good survey design at a very general level.
+We also have removed references to fish and fisheries from some places in the abstract, to show our work can easily be generalized to other systems.*
 
@@
 *We really struggled to accommodate traditional Methods and Results sections because all of our re-structuring attempts defied the definitions of these sections.
-To maintain a logical flow that describes how and why to use the package, we landed on a mix of methods and case study results/discussion through the body of the paper.
-We first describe the "Model structure", then the "Core functions", then we describe and discuss the case study results in the "Interpretation" section (this section also includes a solution to bias generated by the standard sampling/analysis approach) and, finally, we move to the "Parameterisation" section where we provide some guidance on how to modify default settings to suit specific needs.
-The how-to content still remains with this structure, however, it has been augmented with more case study results (and one new result) which should bring out the real-world implications of the package, plus it provides some meaning to several of the complex figures shown through the paper.*
+To maintain a logical flow that describes the how and why to use the package, we landed on a mix of methods and case study results/discussion through the body of the paper.
+We first describe the "Model structure", then the "Core functions", and then we describe and discuss the case study results in the "Interpretation" section.*
 
-*Below we discuss a couple of failed attempts to force the content into Methods and Results sections:*
+*The new "Interpretation" section is essentially a shortened version of the appendix on the case study, however, this iteration of the manuscript includes one new result.
+In the process of revising the paper and reading more literature, we stumbled upon a potential design-based solution to the bias introduced by the division-level age-length key (default approach).
+The design-based fix was easily implemented using the package and, after running the necessary simulations, we discovered that the resultant estimates appear to be unbiased.
+We feel this is an interesting and useful addition to the paper as it shows that the package can be used to identify issues and explore solutions.*
+
+*We should also note that we have moved the "Parameterisation" section to an appendix.
+This was the section where we provide some guidance on how to modify default settings to suit specific needs.
+Following the addition of the "Interpretation" section, it became clear that this section no longer fits with the logical flow of the paper.*
+
+*Overall, we feel that the structural changes we have made to the paper have minimized the how-to content, and the increased focus on the case study should bring more meaning to the complex figures shown through the paper and highlight the real-world implications of the package.*
+
+*Below we discuss a number of our failed attempts to place the content into Methods and Results sections:*
 
@@
 
-*2) We tried to keep the methods section as is and describe some of the case study results in the results section, but it was difficult to find a logical placement for the "Parameterisation" section (i.e.
-it seemed awkward to keep it in between the "Core functions" section and the "Results" section).
-With this structure it was also hard to find a sensible location for describing and discussing a solution to the age-length key problem.*
+*2) We tried to keep the methods section as is and describe some of the case study results in the results section, however it seemed awkward to describe and discuss a solution to the age-length key problem in a results section.*
 
@@
 
-*Ditto*
+*Please see reply to Methods comment.*
 
@@
 
-*In contrast to a Methods or Results section, a Discussion section was easier to accommodate.
-Here we lean on the case study results to reiterate the importance of planing a survey or testing an existing survey.
-We then use this preamble to segue into the "Research opportunities" section.*
+*In contrast to a Methods or Results section, a Discussion section was easier to accommodate and we have provided one.
+Here we lean on the case study results to reiterate the importance of planning a survey or testing an existing survey.
+We then use this preamble to segue into the "Research opportunities" section as suggested.*
 
@@
 cost-benefit analysis).
-We are hopeful that these are questions that future versions of SimSurvey may be able to help address as more features are added to elevate it from a purely statistical toolbox to something that also considers the human element.*
+We are hopeful that these are questions that future versions of SimSurvey may be able to help address as more features are added to elevate it from a purely statistical toolbox to something that also considers the operational constraints of delivering a survey.*
 
@@
 
-*Agreed, done.*
+*Agreed, we have made the suggested change.*
 

### 2020-02-20 analysis/paper/sim_survey_paper.Rmd
Revise response letter given Greg and Keith's feedback

@@
 
-```{r alt, out.height = "600px", fig.cap = "Main results from an alternative design-based sampling and analysis procedure that accounts for multi-stage cluster sampling of ages.
+```{r alt, out.height = "400px", fig.cap = "Main results from an alternative design-based sampling and analysis procedure that accounts for multi-stage cluster sampling of ages.
 Specifically, a) is a fan chart of stratified estimates of abundance at age from year seven of the simulation from a survey with a set density ($D_{\\mathrm{sets}}$) of 0.002 sets / km^2^, maximum number of length samples ($M_{\\mathrm{lengths}}$) of 100, and maximum number of age samples per 3 cm length group per set ($M_{\\mathrm{ages}}$) of 3 (the thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%); and b) is a surface plot of root-mean-squared error (RMSE) from an array of surveys with set density ($D_{\\mathrm{sets}}$) fixed to 0.002 sets / km^2^ but different sub-sampling effort, where the x-axes represent the maximum number of lengths to measure per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per 3 cm length group per set ($M_{\\mathrm{ages}}$)."}

### 2020-02-20 analysis/paper/sim_survey_paper.Rmd
Compile files for submission R2

@@
 Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for multi-stage survey data.
-Here we introduce the R package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed populations.
+Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed populations.
 The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of built-in or user-defined sampling protocols to reproduce the population parameters of the known population.

### 2020-04-09 analysis/paper/response_letter.Rmd
Start response to latest round of peer-review

--- 2020-02-18 analysis/paper/response_letter.Rmd
@@
 
-Thank you for considering another revision of manuscript PONE-D-19-26904R1,  **"`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"** by Paul M.
+Thank you for considering another revision of manuscript PONE-D-19-26904R2,  **"`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"** by Paul M.
 Regular, Gregory J.
@@
 Lewis, Jonathan Babyn, Brian Healey and Fran Mowbray.
-We are also grateful for your detailed suggestions and we have made every effort to do justice to the changes you recommend.
-Most importantly, we hope we have added sufficient content to the core of the paper elevate the manuscript from a software manual to a primary scientific publication.
-Though the how-to approach remains, we now see that describing some of the case study results in the core of the manuscript makes it more interesting and it adds another tangible reason for prospective users to learn how to use the package.
-Please see below for more details on the changes we made in response to your suggestions.
+We have carefully considered the latest round of suggestions and below we outline how we have addressed each point.
 
@@
 
-Additional Editor Comments:
+Reviewer #1: This is my second time reviewing this manuscript and the authors did a great job addressing people's comments.
+Great job for that.
+The manuscript is now much clearer and read well, I think.
 
-Dear Paul, you have made lots of excellent changes that I think help with the use of the package and responded to the many specific comments of reviewers which has no doubt corrected many technical issues and clarifications.
-The one thing I would say that you have not really done is get at the deeper research merits of this work beyond introducing a new piece of software.
-I think paper, as it stands, lacks a larger context and content which is important for the primary publication.
-My diagnosis for why this is is that the manuscript does not conform very well to more typical scientific reports (Intro, M&M, Result, Discussion) which can make it difficult for readers to find the larger scientific merits of the work.
-It is useful of course for those who already understand the merits of this kind of work but this work is for primary publication and it needs to appeal more to the former than the latter group.
-There is a very "how-to" feel to it (e.g.
-line 64 "In this section") which I think detracts from getting at the larger purpose of the work.
+**We thank the reviewer for the kind words.**
 
-I would really like you to address this issue of moving it from a software manual to a primary scientific publication.
-I do not think it should involve that much work but there will be some restructuring of sections as well as places to put in content and bring out conclusions.
-Here are my suggestions for this:
+This time, I only have a few minor comments:  
+L19 : «there is a limited number of»
 
-Try to follow a more traditional paper structure.
-This will help readers and it likely will also make it clearer for you on how you can inject content into the paper to move it beyond the software manual approach:
+**We have removed the 'the' that should not have been in that sentence.**
 
-Introduction:
+L103: Maybe it would be a good idea to justify why you are converting to abundance at length.
 
-- You need to talk about wider issues and examples.
-e.g.
-examples of when poor survey design meant that scientific questions could not be properly addressed when good survey design meant they were.
-Examples of when good survey design allowed researchers to address needs that were unanticipated at the time it was designed.
-i.e.
-you need to build a better case (not just cost) of why survey design is really important and it is most powerful to do this with examples.
-You should try to bring in ideas related to ecosystem and climate changes and being able to track communities.
-Perhaps bring in species at risk ideas and tracking decline, you could bring in ideas related to MSC certification.
-These are just examples of specifics but you get the idea: the Introduction needs to have more general information outlining in both a broad and specific sense why we should be concerned about this.
+**Agreed, we have started a new paragraph regarding abundance at length which begins: "In practice, abundance at age is often inferred from length data as it is easier to collect.
+Abundance at length is therefore simulated from abundance at age using the original von Bertalanffy growth curve"**
 
-- Keep in mind an educated reader who may not be in fisheries but is interested in why anyone should care about this or could, for example, be interested in surveying say caribou or songbirds or something outside of fisheries but where many of the motivation and concepts may be similar and they are doing a more general literature search before designing their own survey.
+L240-242: I think it could be a good idea to show how to exactly do this (maybe obvious to advanced people but not to the general audience).
+This could be a simple reference to an example on the github repo.
 
-*Thank you for thinking beyond a fisheries biologist audience, as this is the audience we have been targeting from the onset.
-In hindsight we see that this could be short-sighted so we have modified our introduction, as suggested, to pitch the concept to a broader audience.
-The reason we did not think beyond fisheries biologist is because they tend to have access to age-based data while others (e.g.
-seabird and caribou biologists) do not.
-Alas, there are probably exceptions to our perceptions; if nothing else, the general framework we present may be useful to adapt to other marine or terrestrial situations.*
+**...**
 
-Methods:
+Table 2: If bias correction was not implemented, I would suggest to change the labeling in the table here and elsewhere (if needed) and replace the “mean” by “median” (if the variable is transformed back to the original scale).
+If not, I would make clear that you talk about mean in log scale.
 
-- This will start at your "Model Structure" section. You should put a higher level heading just before that called Methods.
+L326: I did not see how one can use INLA or RandomFields in Appendix S3. This is a similar comment as the one for line 240-242.
 
-- I can see that it is hard to separate your Methods from Results.
-Your results are really in the section "Using SimSurvey" I would not be opposed to putting this as a separate Results section but I leave it to you to decide.
-Essentially what you have is a case study as an example of how it works and therefore specific results are less interesting than how you got there with the package.
-You might title it something like "Results: running a SimSurvey simulation".
-This section also has a lot of content without a lot of explanation.
-for example, you have several large and complicated figures in a row.
-You should discuss not only what figures are meant to offer in the package but also what they mean.
-So for example on line 473 you refer to three figures (5,6,7) but you offer little interpretation of those figures just why you can make them which is another example of the limitations of the how-to manual approach.
-So you need to think of it as a case study and help someone decide the implications of their survey design.
-
-*We really struggled to accommodate traditional Methods and Results sections because all of our re-structuring attempts defied the definitions of these sections.
-To maintain a logical flow that describes how and why to use the package, we landed on a mix of methods and case study results/discussion through the body of the paper.
-We first describe the "Model structure", then the "Core functions", then we describe and discuss the case study results in the "Interpretation" section (this section also includes a solution to bias generated by the standard sampling/analysis approach) and, finally, we move to the "Parameterisation" section where we provide some guidance on how to modify default settings to suit specific needs.
-The how-to content still remains with this structure, however, it has been augmented with more case study results (and one new result) which should bring out the real-world implications of the package, plus it provides some meaning to several of the complex figures shown through the paper.*
-
-*Below we discuss a couple of failed attempts to force the content into Methods and Results sections:*
-
-*1) We tried to start the methods section with "Model Structure", as suggested, and shunt the figures to a "Results" section; this, however, broke the flow and left a gap in the methods section which held demonstrations of the core simulation functions but not the plotting functions.*
-
-*2) We tried to keep the methods section as is and describe some of the case study results in the results section, but it was difficult to find a logical placement for the "Parameterisation" section (i.e.
-it seemed awkward to keep it in between the "Core functions" section and the "Results" section).
-With this structure it was also hard to find a sensible location for describing and discussing a solution to the age-length key problem.*
-
-*3) Earlier iterations of the paper were focused entirely on the case study, the format was traditional and the package was a side-note in the methods section.
-It became apparent, however, that the package was more interesting and generally applicable than the case study results.
-We then began re-working the paper into a how-to format and, while moving away from a conventional format, we looked to several papers as a guide on how to document a R package  (e.g.
-https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0092725).*
-
-Results:
-
-- see comment above
-
-*Ditto*
-
-Discussion:
-
-- This starts just before "Research Opportunities".
-You can keep these sections but there should be more preamble before jumping right into research opportunities.
-I suggest a general paragraph(s) that segue into your subsections of "research opportunities", "future research".
-In the Discussion you should address again some of the broader issues from the Introduction, e.g.
-how could spatial (depth) distribution changes anticipated under climate change for some population be tackled by survey design exploration now so that we can continue to track these changing populations 20 years down the road and do not lose the signal.
-How can this software help with that?
-
-*In contrast to a Methods or Results section, a Discussion section was easier to accommodate.
-Here we lean on the case study results to reiterate the importance of planing a survey or testing an existing survey.
-We then use this preamble to segue into the "Research opportunities" section.*
-
-- I think something that can be useful for readers is to outline the steps in a thought process a researcher might undertake when setting up a survey (perhaps a separate subsection) and then how one might go about a SimSurvey run for this.
-It also gives you a good opportunity to discuss your multispecies ideas:
-    - What is your current problem/needs
-    - What are the constraints (HR and $ perspective)
-    - What are the constraints from a biological, physical perspective
-    - Resources available
-    - What are your anticipated future needs (i.e. if climate change is going to make cod go deeper, will you be able to track it in 20 years?)
-    - How could you consider some or all of these with SimSurvey
-
-*As much as we would like to address each of these points and work towards a survey design handbook, it would be hard to be as cohesive and comprehensive as Cochran (1977; "Sampling techniques") or Sutherland (2006; "Ecological census techniques: a handbook").
-Moreover, we have yet to implement features that could help users assess some of these trade-offs (i.e.
-cost-benefit analysis).
-We are hopeful that these are questions that future versions of SimSurvey may be able to help address as more features are added to elevate it from a purely statistical toolbox to something that also considers the human element.*
-
-- Your "Assumptions" section should come down into the Discussion
-
-*Agreed, done.*
-
-- "Future Directions" should say something about randomfields re: Reviewer 1.
-Even if you just outline that you have considered it.
-You might also try to say something about optimisation of design which Reviewer 2 mentioned.
-
-*Good point.
-We have noted in the "sim_distribution" section that a custom closure can be created that uses randomfields to simulate spatial noise.
-We have also noted in the "Future directions" section that we hope to implement alternatives to random or stratified random designs to allow for more comprehensive evaluations of various designs.*
+L481-490: the processing time is surprisingly long… it is not an issue in itself but this might refrain some people to using this package… I think some code optimization could help (e.g.
+vectorize, use matrix operation as much as possible, etc)

### 2020-04-09 analysis/paper/sim_survey_paper.Rmd
Start response to latest round of peer-review

@@
 Population estimates are therefore affected by each level of sampling as well as the pathway taken to analyze such data.
-Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for multi-stage survey data.
+Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for multi-stage survey data.
 Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed populations.
@@
 Note that a plus group is not modeled as the number of ages can easily be extended to include groups with zero fish.
-Abundance at age is then converted to abundance at length using the original von Bertalanffy growth curve [@von1938]:
+
+In practice, abundance at age is often inferred from length data as it is easier to collect.
+Abundance at length is therefore simulated from abundance at age using the original von Bertalanffy growth curve [@von1938]:
 
@@
 
+Overall, this formulation facilitates the simulation of a dynamic length and age structured population.
 Though some typical relationships have yet to be implemented (e.g.
-stock-recruitment), this formulation facilitates the simulation of age-structured populations that are dynamic enough to assess survey performance across a range of abundance levels across years and ages.
+stock-recruitment), sufficient information can be simulated to assess survey performance across a range of abundance levels across years, lengths and ages.
 
@@
 Moreover, this approach provides an option for advanced R users to inspect and modify the closures implemented in the package to supply custom closures with alternate equations.
+...
+.
 Also note that each of the closures implemented in the package includes a plot argument such that quick visuals can be obtained using a line of code like this: `sim_R(log_mean = log(500), log_sd = 0.5, plot = TRUE)(years = 1:100)`).

### 2020-04-15 analysis/paper/response_letter.Rmd
Complete draft of response letter and revised paper

@@
 
-**We thank the reviewer for the kind words.**
+*We thank the reviewer for the kind words.*
 
@@
 
-**We have removed the 'the' that should not have been in that sentence.**
+*We have removed the 'the' that should not have been in that sentence.*
 
@@
 
-**Agreed, we have started a new paragraph regarding abundance at length which begins: "In practice, abundance at age is often inferred from length data as it is easier to collect.
-Abundance at length is therefore simulated from abundance at age using the original von Bertalanffy growth curve"**
+*Agreed, we have started a new paragraph regarding abundance at length which begins: "In practice, abundance at age is often inferred from length data as it is easier to collect.
+Abundance at length is therefore simulated from abundance at age using the original von Bertalanffy growth curve"*
 
@@
 
-**...**
+*Indeed, this is a helpful idea and we have now created a short vignette on creating closures and this can be accessed via the pkgdown site (https://paulregular.github.io/SimSurvey/articles/custom_closures.html).
+A link to this vignette has been added to the paper.*
 
@@
 
+*Good catch, we have added log after mean to ensure that readers know the functions are operating in log scale.*
+
 L326: I did not see how one can use INLA or RandomFields in Appendix S3. This is a similar comment as the one for line 240-242.
+
+*We have added the following in parentheses: see the code behind `sim_ays_covar_sped` [https://github.com/PaulRegular/SimSurvey/blob/master/R/sim_dist_spde.R] for an example of how the `sim_ays_covar` closure was modified to apply a Stochastic Partial Differential Equation approach using the `INLA` package.*
 
@@
 vectorize, use matrix operation as much as possible, etc)
+
+*We have stretched our programming skills to the limit to reduce the processing time as much as possible.
+Of course, said programming skills are largely self-taught, so there is surely room for improvement.
+While developing the package, we repeatedly profiled our code and vectorized as much as we could, leaning heavily on the `data.table` package.
+We have also applied parallel processing in places in an attempt to speed up the code.
+Another step to further optimize the code would be to translate many of the core functions to C++, however this is on the wish list and well outside the scope of the project at this point in time.
+We are hopeful that we have produced enough here to catalyze collaborations that help extend and optimize the package.*

### 2020-04-15 analysis/paper/sim_survey_paper.Rmd
Expand README and add some structure to pkgdown reference page
Complete draft of response letter and revised paper

@@
 library(magrittr)
+library(here)
 Sys.setenv("MAPBOX_TOKEN" = "pk.eyJ1IjoicGF1bHJlZ3VsYXIiLCJhIjoiY2pueGIxejMxMGM0MjNqbjR1aXdwYm92bSJ9.R-t1SCa0VHKf4N-PQmYI5Q")
@@
 This was done to avoid the repeated specifications of key arguments, such as ages and years.
-Moreover, this approach provides an option for advanced R users to inspect and modify the closures implemented in the package to supply custom closures with alternate equations.
-...
-.
+Moreover, this approach provides an option for advanced R users to inspect and modify the closures implemented in the package to supply custom closures with alternate equations (see https://paulregular.github.io/SimSurvey/articles/custom_closures.html for a short vignette on creating custom closures).
 Also note that each of the closures implemented in the package includes a plot argument such that quick visuals can be obtained using a line of code like this: `sim_R(log_mean = log(500), log_sd = 0.5, plot = TRUE)(years = 1:100)`).
@@
 | `+  years = 1:20,`                          | Years                                                      | $y$                                 |
-| `+  R = sim_R(log_mean = log(30000000),`    | Mean recruitment^1^                                        | $\mu_r$                             |
-| `+            log_sd = 0.5),`               | Standard deviation of log-recruitment                      | $\sigma_r$                          |
-| `+  Z = sim_Z(log_mean = log(0.5),`         | Mean total mortality^2^                                    | $\mu_Z$                             |
-| `+            log_sd = 0.2,`                | Standard deviation of total mortality (log)                | $\sigma_Z$                          |
+| `+  R = sim_R(log_mean = log(30000000),`    | Mean log recruitment^1^                                    | $\mu_r$                             |
+| `+            log_sd = 0.5),`               | Standard deviation of log recruitment                      | $\sigma_r$                          |
+| `+  Z = sim_Z(log_mean = log(0.5),`         | Mean log total mortality ^2^                               | $\mu_Z$                             |
+| `+            log_sd = 0.2,`                | Standard deviation of log total mortality                  | $\sigma_Z$                          |
 | `+            phi_age = 0.9,`               | Correlation across ages in error around total mortality    | $\varphi_{\delta, \mathrm{age}}$    |
@@
 } else {
-  knitr::include_graphics("figures/plot_surface.png")
+  knitr::include_graphics(here("analysis", "paper", "figures", "plot_surface.png"))
 }
@@
 } else {
-  knitr::include_graphics("figures/plot_grid.png")
+  knitr::include_graphics(here("analysis", "paper", "figures", "plot_grid.png"))
 }
@@
 Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by adjusting a few parameters in these closures (Figure \@ref(fig:distribution)).
-Though we only provide one option for simulating a spatial noise (see **[S3 Appendix][S3 Appendix: Age-year-space covariance]** for details), advanced users could construct closures that leverage simulation models provided by packages such as **`RandomFields`** [@schlather2020] or **`INLA`** [@rue2009].
+While here we only describe one option for simulating a spatial noise (see **[S3 Appendix][S3 Appendix: Age-year-space covariance]** for details), custom colsures can be used that leverage simulation models provided by packages such as **`RandomFields`** [@schlather2020] or **`INLA`** [@rue2009] (see the code behind `sim_ays_covar_sped` [https://github.com/PaulRegular/SimSurvey/blob/master/R/sim_dist_spde.R] for an example of how the `sim_ays_covar` closure was modified to apply a Stochastic Partial Differential Equation approach using the **`INLA`** package).
 Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e.
@@
 } else {
-  knitr::include_graphics("figures/plot_distribution.png")
+  knitr::include_graphics(here("analysis", "paper", "figures", "plot_distribution.png"))
 }
@@
 } else {
-  knitr::include_graphics("figures/plot_survey.png")
+  knitr::include_graphics(here("analysis", "paper", "figures", "plot_survey.png"))
 }
@@
 } else {
-  knitr::include_graphics("figures/plot_total_strat_fan.png")
+  knitr::include_graphics(here("analysis", "paper", "figures", "plot_total_strat_fan.png"))
 }
@@
 } else {
-  knitr::include_graphics("figures/plot_length_strat_fan.png")
+  knitr::include_graphics(here("analysis", "paper", "figures", "plot_length_strat_fan.png"))
 }
@@
 } else {
-  knitr::include_graphics("figures/plot_age_strat_fan.png")
+  knitr::include_graphics(here("analysis", "paper", "figures", "plot_age_strat_fan.png"))
 }
@@
 } else {
-  knitr::include_graphics("figures/plot_survey_rank.png")
+  knitr::include_graphics(here("analysis", "paper", "figures", "plot_survey_rank.png"))
 }
@@
 } else {
-  knitr::include_graphics("figures/plot_error_surface.png")
+  knitr::include_graphics(here("analysis", "paper", "figures", "plot_error_surface.png"))
 }
@@
 } else {
-  knitr::include_graphics("figures/alt_tests_plots.png")
+  knitr::include_graphics(here("analysis", "paper", "figures", "alt_tests_plots.png"))
 }
@@
 
-knitr::include_graphics("figures/real_sim_comp.png")
+knitr::include_graphics(here("analysis", "paper", "figures", "real_sim_comp.png"))
 

### 2020-04-15 analysis/paper/sim_survey_paper.Rmd
Prep rev3 submission files

@@
 Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by adjusting a few parameters in these closures (Figure \@ref(fig:distribution)).
-While here we only describe one option for simulating a spatial noise (see **[S3 Appendix][S3 Appendix: Age-year-space covariance]** for details), custom colsures can be used that leverage simulation models provided by packages such as **`RandomFields`** [@schlather2020] or **`INLA`** [@rue2009] (see the code behind `sim_ays_covar_sped` [https://github.com/PaulRegular/SimSurvey/blob/master/R/sim_dist_spde.R] for an example of how the `sim_ays_covar` closure was modified to apply a Stochastic Partial Differential Equation approach using the **`INLA`** package).
+While here we only describe one option for simulating a spatial noise (see **[S3 Appendix][S3 Appendix: Age-year-space covariance]** for details), custom closures can be used that leverage simulation models provided by packages such as **`RandomFields`** [@schlather2020] or **`INLA`** [@rue2009] (see the code behind `sim_ays_covar_sped` [https://github.com/PaulRegular/SimSurvey/blob/master/R/sim_dist_spde.R] for an example of how the `sim_ays_covar` closure was modified to apply a Stochastic Partial Differential Equation approach using the **`INLA`** package).
 Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e.

### 2020-04-16 analysis/paper/response_letter.Rmd
Minor tweaks to rev3 files given feedback from co-authors

@@
 We have also applied parallel processing in places in an attempt to speed up the code.
-Another step to further optimize the code would be to translate many of the core functions to C++, however this is on the wish list and well outside the scope of the project at this point in time.
+In other parts of the package, we learned that there is little we can do to reduce the processing time (e.g.
+Cholesky decomposition is a factor limiting the speed of the `sim_distribution` function).
+Another step to further optimize the code would be to translate some of the core functions to C++, however this is on the wish list and well outside the scope of the project at this point in time.
 We are hopeful that we have produced enough here to catalyze collaborations that help extend and optimize the package.*

### 2020-04-16 analysis/paper/sim_survey_paper.Rmd
Minor tweaks to rev3 files given feedback from co-authors

@@
 Likewise, age frequencies need to be calculated to obtain stratified estimates of abundance at age.
-This is done by constructing an age-length key, which is the proportion of fish in each length bin that fall into specific ages classes.
-Once these proportions are calculated, they applied to the bumped up length frequencies, $n_{l,i}$, to approximate age frequencies, $n_{a,i}$, that is:
+This is done by constructing an age-length key, which is the proportion of fish in each length bin that fall into specific age classes.
+Once these proportions are calculated, they are applied to the bumped up length frequencies, $n_{l,i}$, to approximate age frequencies, $n_{a,i}$, that is:
 
@@
 The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`.
-If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure structure and correct projection is used.
+If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure and correct projection is used.
 The package includes a manually constructed survey grid of NAFO Subdivision 3Ps off the southern coast of Newfoundland (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid.
@@
 | `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / spatial group ^1^         | $M_{\mathrm{ages}}$     |
-| `+  age_sampling = "stratified")`   | Controls whether age sampling is length `"stratified"` or `"random"`        | &nbsp;                  |
-| `+  age_length_group = 1)`          | Length group bin size for stratified age sampling (cm)                      | $l_{\mathrm{group}}$    |
+| `+  age_sampling = "stratified",`   | Controls whether age sampling is length `"stratified"` or `"random"`        | &nbsp;                  |
+| `+  age_length_group = 1,`          | Length group bin size for stratified age sampling (cm)                      | $l_{\mathrm{group}}$    |
 | `+  age_space_group = "division")`  | Spatial scale of stratified age sampling (`"division"`, `"strat"`, `"set"`) | $s_{\mathrm{group}}$    |
@@
 Specifically, estimates of total abundance and abundance at length are centered around true values and their probability envelopes tighten as set density increases (Figures \@ref(fig:fan1), \@ref(fig:fan2)).
-Case study results also echo the growing body of literature which concludes that extra sub-sampling is an ineffective means of improving estimates relative to sampling more locations [@Pennington1994; @Pennington2002; @Stewart2014; @Bogstad1995; @Coggins2013; @Zhang2013].
+Our case study results also echo the growing body of literature which concludes that extra sub-sampling is an ineffective means of improving estimates relative to sampling more locations [@Pennington1994; @Pennington2002; @Stewart2014; @Bogstad1995; @Coggins2013; @Zhang2013].
 This is exemplified by the relatively large drops in RMSE when set density is increased compared to when sub-sampling effort is increased (Figures \@ref(fig:rank), \@ref(fig:rmse-surface)).
@@
 For the case study, a proper design-based estimator will need to account for cluster sampling in a stratified random survey.
-Following recommendations in Aanes and Vølstad [-@Aanes2015], we used **`SimSurvey`** to 1) conduct a surveys with concurrent length and age sampling at every set, 2) construct and apply age-length keys on a set-by-set basis, and 3) weight the resultant age frequencies at each set using stratified random estimators.
+Following recommendations in Aanes and Vølstad [-@Aanes2015], we used **`SimSurvey`** to 1) conduct a survey with concurrent length and age sampling at every set, 2) construct and apply age-length keys on a set-by-set basis, and 3) weight the resultant age frequencies at each set using stratified random estimators.
 This test was accomplished using this code:
@@
 
-This is a minor modification of the code used in the **[`test_surveys`]** section whereby the same simulated population was used (`pop` object) but set density scenarios were reduced to one option, for simplicity, and age sampling protocol and stratified analysis options were changed.
+This is a minor modification of the code used in the **[`test_surveys`]** section whereby the same simulated population was used (`pop` object) but set density scenarios were reduced to one option for simplicity, and age sampling protocol and stratified analysis options were changed.
 That is, 1) `age_length_group` and `age_space_group` were changed to simulate a survey that conducts concurrent length and age sampling at every set (e.g.

### 2020-04-16 R/sim_dist_spde.R

@@ -15,6 +15,7 @@
 #' @param model 'spde' for the SPDE approach, 'barrier' for barrier approach
 #'
 #' @return Q a sparse precision matrix of type dgTMatrix
+#' @keywords internal
 #'
 .Q <- function(mesh,barrier.triangles,range=50,range_fraction=0.2,
                sigma_u = 1,model="spde"){

### 2020-04-16 analysis/paper/sim_survey_paper.Rmd
One more tweak before approving re-submission to PLOS ONE

@@
 
-Unlike the default approach to collecting and analyzing age samples, results from this test appear to be unbiased and additional sub-sampling effort beyond a certain point appears to have little to no effect on the estimates (Figure \@ref(fig:alt))).
+Unlike the default approach to collecting and analyzing age samples, results from this test appear to be unbiased and additional sub-sampling effort beyond a certain point appears to have little to no effect on the estimates (Figure \@ref(fig:alt)).
 This is an important result with real-world implications as they imply that 1) inference from the actual survey can be improved by altering the sampling and analysis of ages and, 2) valuable survey time can be saved by cutting back on sub-sampling effort.

### 2020-07-30 R/vis_sim.R
Use base approach for installing previous version of plotly when running vis_fit

@@ -47,7 +47,8 @@ vis_sim <- function(sim, ...) {
     ans <- readline(prompt = "are you sure you want to revert to plotly version 4.8? ")
     if (substr(ans, 1, 1) == "y") {
       utils::remove.packages("plotly")
-      devtools::install_version("plotly", version = "4.8.0", repos = "http://cran.us.r-project.org")
+      pkg_url <- "https://cran.r-project.org/src/contrib/Archive/plotly/plotly_4.8.0.tar.gz"
+      utils::install.packages(pkg_url, repos = NULL, type = "source")
     } else {
       stop("plotly version 4.8 is required.")
     }

### 2020-07-30 R/data.R
Fix several spelling mistakes

@@ -42,14 +42,14 @@
 #'  @format A list containing the R-INLA survey mesh, the set of triangles in the barrier and the barrier polygons for plotting
 #'
 #' An example of a mesh containing barrier information for use with
-#' sim_ays_covar_spde. Also derived from global adminstrative boundaries
+#' sim_ays_covar_spde. Also derived from global administrative boundaries
 #' data (http://gadm.org). Details on creation provided in the data-raw
 #' folder of this package in the survey_mesh.R file. Includes the set
 #' of barrier triangles needed to use the barrier approach, barrier
 #' polygons for plotting and the set of triangles in the barrier.
 "survey_mesh"
 
-#'Lite sample survey meshs and related items
+#' Lite sample survey mesh and related items
 #'
 #' @format A list containing the same items as survey_mesh, but with fewer nodes to save on computational time
 #'

### 2020-07-30 R/run_strat.R

@@ -113,7 +113,7 @@ strat_data <- function(sim, length_group = 3, alk_scale = "division") {
 #' Calculate stratified means, variances and confidence intervals across groups
 #'
 #' @details Function was mainly created for use in the \code{\link{run_strat}} function.
-#' It first calculates strat-level statistics and then the larger-scale stistics like total abundance
+#' It first calculates strat-level statistics and then the larger-scale statistics like total abundance
 #'
 #' @param data            Expects data.table with all grouping variables in stacked format (must include
 #'                        strat_area and tow_area for scaling values)

### 2020-07-30 R/sim_pop.R

@@ -6,7 +6,7 @@
 #' Given parameters, it generates N0, R and Z values.
 #'
 #' @param log_mean One mean value or a vector of means, in log scale, of length equal to years for \code{sim_R} or a matrix of means with
-#' rows equaling the number of ages and colums equaling the number of years for \code{sim_Z}.
+#' rows equaling the number of ages and columns equaling the number of years for \code{sim_Z}.
 #' @param random_walk Simulate recruitment as a random walk?
 #' @param log_sd Standard deviation of the variable in the log scale.
 #' @param phi_age Autoregressive parameter for the age dimension.
@@ -133,7 +133,7 @@ sim_N0 <- function(N0 = "exp", plot = FALSE) {
 #'
 #' @description Helper function for converting lengths to length groups
 #' (Note: this isn't a general function; the output midpoints defining the
-#' groups alligns with DFO specific method/labeling)
+#' groups aligns with DFO specific method/labeling)
 #'
 #' @param length       Interval from \code{\link[base]{findInterval}}
 #' @param group        Length group used to cut the length data
@@ -254,7 +254,7 @@ convert_N <- function(N_at_age = NULL, lak = NULL) {
 #'                 \code{\link{sim_N0}}, for generating starting abundance vector.
 #' @param growth   Closure, such as \code{\link{sim_vonB}}, for simulating length given age.
 #'                 The function is used here to generate a abundance-at-age matrix and
-#'                 it is carried foward for later use in \code{\link{sim_survey}} to simulate
+#'                 it is carried forward for later use in \code{\link{sim_survey}} to simulate
 #'                 lengths from survey catch at age.
 #'
 #' @return A \code{list} of length 9:

### 2020-07-30 R/sim_survey.R

@@ -257,7 +257,7 @@ sim_survey <- function(sim, n_sims = 1, q = sim_logistic(), trawl_dim = c(1.5, 0
 #' Simulate stratified random surveys using parallel computation
 #'
 #' This function is a wrapper for \code{\link{sim_survey}} except it allows for
-#' many more total itterations to be run than \code{\link{sim_survey}} before running
+#' many more total iterations to be run than \code{\link{sim_survey}} before running
 #' into RAM limitations. Unlike \code{\link{test_surveys}}, this function retains
 #' the full details of the survey and it may therefore be more useful for testing
 #' alternate approaches to a stratified analysis for obtaining survey indices.

### 2020-07-30 R/test_surveys.R

@@ -47,7 +47,7 @@ expand_surveys <- function(set_den = c(0.5, 1, 2, 5, 10) / 1000,
 #' @inherit                 run_strat
 #' @inheritDotParams        sim_survey -sim -n_sims -set_den -lengths_cap -ages_cap -light
 #'
-#' @details Depening on the settings, \code{test_surveys} may take a long time to run.
+#' @details Depending on the settings, \code{test_surveys} may take a long time to run.
 #' The \code{resume_test} function is for resuming partial runs of \code{test_surveys}.
 #' Note that progress bar time estimates will be biased here by previous completions.
 #' \code{test_loop} is a helper function used in both \code{test_surveys} and

### 2021-12-13 tutorial/tutorial.Rmd
Start making learnr tutorial

---
title: "SimSurvey demo"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(SimSurvey)
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

### Fisheries-independent surveys

<div style="float: left; width: 50%;">

- Conducted by many RFMOs
- Becoming increasingly important in stock assessment
- Associated with difficult questions:
    - How many fish are in the sea?
    - Where are they located?
    - How should we sample the population?
    - How should we analyze these data?

</div>

<div style="float: right; width: 50%;">

<img src="https://pbs.twimg.com/media/DDRLGEmXUAAXhGf?format=jpg&name=large" width="400"/>
<br>
&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">Canadian Coast Guard | Twitter</a>
</font>

</div>

## Core functions

### `sim_abundance`

```{r sim-abundance, exercise = TRUE, exercise.eval = FALSE}
set.seed(438)
long <- sim_abundance(ages = 1:20,
                      R = sim_R(log_mean = log(3e+07)),
                      Z = sim_Z(log_mean = log(0.2)))
short <- sim_abundance(ages = 1:6,
                       R = sim_R(log_mean = log(1e+10)),
                       Z = sim_Z(log_mean = log(0.8)))
```

### 2021-12-14 tutorial/tutorial.Rmd
Introduce sim_abundance

@@
 
-<img src="https://pbs.twimg.com/media/DDRLGEmXUAAXhGf?format=jpg&name=large" width="400"/>
+<img src="https://pbs.twimg.com/media/DDRLGEmXUAAXhGf?format=jpg&name=large" width="350"/>
 <br>
@@
 
-## Core functions
+<br>
 
-### `sim_abundance`
+<div style="float: left; width: 100%;">
 
-```{r sim-abundance, exercise = TRUE, exercise.eval = FALSE}
+### How can [SimSurvey](https://github.com/PaulRegular/SimSurvey) help?
+
+- A sandbox for simulating of fisheries-independent trawl surveys
+  - **Survey:** random or stratified-random
+  - **Population:** age-structured and spatially-correlated
+- Facilitates tests of the design and analysis of survey data
+  - How many sets, lengths, ages are enough?
+  - Should the analysis be design or model based?
+  - Can models be used to stitch surveys or fill gaps?
+
+</div>
+
+## Simulation steps
+
+1. **Simulate abundance** - `sim_abundance()`
+    - Common cohort model
+
+2. **Simulate spatial aggregation** - `sim_distribution()`
+    - Includes depth associations and noise correlated across space, years and ages
+
+3. **Simulate survey data** - `sim_survey()`
+    - Sample population using stratified random survey
+    - These data can be analyzed using `run_strat()` (design-based analysis), or using a package like [sdmTMB](https://github.com/pbs-assess/sdmTMB)
+
+## Simulate abundance
+
+### `sim_abundance()`
+
+- Abundance at age and length is simulated using `sim_abundance()`
+- Key arguments:
+  - `ages`, `years` - supply vectors of ages and years to simulate across
+  - `R`, `Z`, `growth` - supply closures such as `sim_R()`, `sim_Z()`, and `sim_vonB()`
+
+### Closure
+
+- Functions that retain argument values and return functions
+
+```{r closure, exercise = TRUE, exercise.eval = FALSE}
+R_fun <- sim_R(log_mean = log(500), log_sd = 0.5)
+R_fun
+```
+
+### Usage
+
+- Closures minimize the repeated specification of arguments such as ages and years
+  - Specify once in `sim_abundance` and the values are used by the functions returned by the closures
+
+```{r usage, exercise = TRUE, exercise.eval = FALSE}
+pop <- sim_abundance(ages = 1:6, years = 1:4,
+                     R = sim_R(log_mean = log(1000)),
+                     Z = sim_Z(log_mean = log(0.8)))
+round(pop$N)
+```
+
+### Tinker and plot
+
+- For instance, a relatively short and long lived species are simulated below
+- Results can be visualized using `plot_surface()`
+
+```{r tinker, exercise = TRUE, exercise.eval = FALSE}
 set.seed(438)
@@
                        Z = sim_Z(log_mean = log(0.8)))
+plot_surface(short, mat = "N")
 ```
+
+## Simulate spatial aggregation
+
+### `sim_distribution()`

### 2021-12-15 tutorial/tutorial.Rmd
Introduce the rest of the core functions in the tutorial and start sdmTMB section

@@
 library(SimSurvey)
-knitr::opts_chunk$set(echo = FALSE)
+library(dplyr)
+library(sdmTMB)
+library(plotly)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
+                      out.width = "100%", out.height = "450px")
 ```
@@
 - Key arguments:
-  - `ages`, `years` - supply vectors of ages and years to simulate across
-  - `R`, `Z`, `growth` - supply closures such as `sim_R()`, `sim_Z()`, and `sim_vonB()`
+  - `ages`, `years` -- supply vectors of ages and years to simulate across
+  - `R`, `Z`, `growth` -- supply closures such as `sim_R()`, `sim_Z()`, and `sim_vonB()`
 
@@
 ### `sim_distribution()`
+
+- `sim_distribution()` is used to distribute a population simulated using `sim_abundance()` throughout a grid
+- Key arguments:
+  - `sim` -- A list with abundance details like produced by `sim_abundance()`
+  - `grid` -- A raster object defining the survey grid, like the one produced by `make_grid()`
+  - `ays_covar` -- Closure for simulating age-year-space covariance, like `sim_ays_covar()`
+  - `depth_par` -- Closure for defining relationship between abundance and depth, like `sim_parabola()`
+
+### Make a grid
+
+- `make_grid()` sets up a depth stratified survey grid
+  - Returns a raster object with four layers: `depth`, `cell`, `division`, and `strat`
+  - Number of strata are affected by the number of divisions (`n_div`), horizontal splits (`strat_splits`), the depth gradient (`shelf_depth`, `shelf_width`, `depth_range`), and depth breaks (`strat_breaks`)
+
+```{r grid, exercise = TRUE, exercise.eval = FALSE}
+g <- make_grid(res = c(10, 19), n_div = 4, strat_splits = 1, 
+               shelf_depth = 500, shelf_width = 0, depth_range = c(0, 1000))
+plot_grid(g)
+```
+
+### Populating the grid
+
+- The degree of spatial clustering across ages and years can be controled using `sim_ays_covar()` by adjusting the decorrelation range (`range`), and the degree of correlation across age (`phi_year`) and year (`phi_age`) dimensions
+- A parabolic depth preference can be controled using `sim_parabola()` where `mu` defines centran tendency and `sigma` the spread
+
+```{r populate-grid, exercise = TRUE, exercise.eval = FALSE}
+pop <- sim_abundance(ages = 1:10, years = 1:10) %>%
+           sim_distribution(grid = make_grid(res = c(10, 10)),
+                            ays_covar = sim_ays_covar(range = 300,
+                                                      phi_age = 0.8,
+                                                      phi_year = 0.1),
+                            depth_par = sim_parabola(mu = 200,
+                                                     sigma = 50))
+plot_distribution(pop, ages = 1:3, years = 1:3, type = "heatmap")
+```
+
+## Simulate survey data
+
+### `sim_survey()`
+
+- Samples from a survey are generated using `sim_survey()`
+- Key arguments:
+  - `n_sims` -- Number of surveys to simulate over the same population
+  - `q` -- Closure, such as `sim_logistic()`, for simulating catchability at age
+  - `trawl_dim` -- Trawl width and distance
+  - `set_den`, `lengths_cap`, and `ages_cap` -- Sampling protocol for sets, length measurement, and age sampling
+
+### Collect samples
+
+- Conduct a survey with a relatively low and high set density:
+
+```{r survey, exercise = TRUE, exercise.eval = FALSE}
+set.seed(438)
+pop <- sim_abundance(ages = 1:10, years = 1:5) %>%
+  sim_distribution(grid = make_grid(res = c(10, 10)))
+
+a <- pop %>%
+  sim_survey(set_den = 1 / 1000)
+b <- pop %>% 
+  sim_survey(set_den = 5 / 1000)
+
+plot_survey(a, which_year = 2, which_sim = 1)
+```
+
+## Analyze
+
+### Design or model-based? That is the question
+
+- Simulated data returned from `sim_survey()` can be analyzed using 
+  - a design-based approach (e.g., stratified analysis; built-in `run_strat()` function), or
+  - a model-based approach (e.g., geostatistical model; [sdmTMB](https://github.com/pbs-assess/sdmTMB))
+
+### Design-based -- stratified analysis
+
+- `run_strat()` runs a basic stratified analysis and `strat_error()` computes error statistics
+
+```{r strat, exercise = TRUE, exercise.eval = FALSE}
+sim <- sim_abundance(ages = 1:5, years = 1:5,
+              R = sim_R(log_mean = log(1e+7)),
+              growth = sim_vonB(length_group = 1)) %>%
+  sim_distribution(grid = make_grid(res = c(10, 10)),
+                   ays_covar = sim_ays_covar(sd = 1)) %>%
+  sim_survey(n_sims = 1, q = sim_logistic(k = 2, x0 = 3)) %>%
+  run_strat() %>%
+  strat_error()
+
+sim$total_strat_error_stats
+sim$length_strat_error_stats
+sim$age_strat_error_stats
+```
+
+### Model-based -- [sdmTMB](https://github.com/pbs-assess/sdmTMB)
+
+- sdmTMB is an R package for fast and flexible fitting of spatiotemporal species distribution GLMMs with [TMB](https://github.com/kaskr/adcomp)
+
+- Modeling steps:
+
+  1. **Make an R-INLA "mesh" representing spatial "knots"** - `make_mesh()`
+  2. **Fit a spatiotemporal GLMM** - `sdmTMB()`
+      - Implements a geostatistical model using TMB, R-INLA, and Gaussian Markov random fields
+      - Function is set up similarly to `glm()`, `lmer()`, or `glmmTMB()`
+  3. **Inspect the model** - `print()`, `tidy()`, `predict()`, `residuals()`, `sdmTMB_cv()`, `sdmTMB_sim()`, `AIC()`
+  4. **Predict on a grid that represents the survey domain** - `predict()`
+  5. **Calculate the survey index** - `get_index()`
+      - Use `get_index()` to sum the expected density or abundance across the grid and calculate standard errors
+
+### `sdmTMB::sdmTMB()`
+
+```{r sdmTMB, exercise = TRUE, exercise.eval = TRUE}
+
+## Simulate population
+set.seed(271)
+pop <- sim_abundance(ages = seq(1, 10), years = seq(1, 20)) %>%
+  sim_distribution(grid = make_grid(res = c(10, 10), depth_range = c(10, 500)),
+                   ays_covar = sim_ays_covar(phi_age = 0.8, phi_year = 0.2),
+                   depth_par = sim_parabola(mu = 200, sigma = 30))
+
+## Simulate survey data
+survey <- sim_survey(pop, n_sims = 1) %>% run_strat()
+
+## Add coordinates to set details
+xy <- as_tibble(survey$grid_xy)
+dat <- as_tibble(survey$setdet) %>%
+  select(x, y, set, year, N = n, tow_area) %>%
+  left_join(., xy, by = c("x", "y")) %>%
+  mutate(offset = log(tow_area))
+
+## Make mesh and fit geostatistical model to survey data
+mesh <- sdmTMB::make_mesh(dat, xy_cols = c("x", "y"), cutoff = 20)
+fit <- sdmTMB(N ~ 0 + as.factor(year) + offset,
+              data = dat,
+              family = nbinom2(link = "log"), spde = mesh,
+              include_spatial = TRUE, time = "year")
+
+## Expand grid data for predictions
+grid_dat <- as_tibble(select(pop$grid_xy, x, y, depth)) %>% distinct()
+grid_dat <- purrr::map_dfr(sort(unique(dat$year)), ~ bind_cols(grid_dat, year = .))
+grid_dat$offset <- mean(dat$offset)
+
+## Predict across full grid and get index
+tow_area <- survey$setdet$tow_area[1] # cell and tow area is consistent in the simulation
+cell_area <- survey$setdet$cell_area[1]
+pred <- predict(fit, newdata = grid_dat, return_tmb_object = TRUE, area = cell_area / tow_area)
+index <- get_index(pred)
+
+## Combine indices
+true_abund <- tibble(year = unique(dat$year), N = as.numeric(colSums(survey$I))) %>% mutate(type = "True")
+strat_abund <- tibble::as_tibble(survey$total_strat) %>%
+  mutate(N = total, type = "Design-based") %>%
+  select(year, N, type)
+index <- index %>%
+  mutate(type = "Model-based", N = est) %>%
+  bind_rows(strat_abund) %>%
+  bind_rows(true_abund)
+
+## Visualize result
+index %>% 
+  plot_ly(x = ~year, color = ~type, legendgroup = ~type) %>% 
+  add_ribbons(ymin = ~lwr, ymax = ~upr, line = list(width = 0), showlegend = FALSE) %>% 
+  add_lines(y = ~N)
+
+```

### 2021-12-15 tutorial/tutorial.Rmd
Wrap-up complete draft of tutorial

@@
 
+- Fit a geostatistical using [sdmTMB](https://github.com/pbs-assess/sdmTMB) using survey data simulated using [SimSurvey](https://github.com/PaulRegular/SimSurvey)
+
 ```{r sdmTMB, exercise = TRUE, exercise.eval = TRUE}
@@
 ## Simulate population
-set.seed(271)
-pop <- sim_abundance(ages = seq(1, 10), years = seq(1, 20)) %>%
+set.seed(17)
+pop <- sim_abundance(ages = seq(1, 10), years = seq(1, 15)) %>%
   sim_distribution(grid = make_grid(res = c(10, 10), depth_range = c(10, 500)),
@@
 ## Expand grid data for predictions
-grid_dat <- as_tibble(select(pop$grid_xy, x, y, depth)) %>% distinct()
+grid_dat <- as_tibble(select(survey$grid_xy, x, y, depth)) %>% distinct()
 grid_dat <- purrr::map_dfr(sort(unique(dat$year)), ~ bind_cols(grid_dat, year = .))
@@
 ```
+
+## What if?
+
+- With these tools we can test whether a model can fill in holes in a survey caused by problems (e.g., vessel issues) or changing protocol (e.g., restricted areas)
+- In the example below, strata 5 and 6 are excluded following year 10 and the analyses run in the previous section are re-run using the affected survey data
+
+```{r what-if, exercise = TRUE, exercise.eval = TRUE}
+
+## Simulate population
+set.seed(17)
+pop <- sim_abundance(ages = seq(1, 10), years = seq(1, 15)) %>%
+  sim_distribution(grid = make_grid(res = c(10, 10), depth_range = c(10, 500)),
+                   ays_covar = sim_ays_covar(phi_age = 0.8, phi_year = 0.2),
+                   depth_par = sim_parabola(mu = 200, sigma = 30))
+
+## Simulate survey data
+survey <- sim_survey(pop, n_sims = 1)
+
+## DROP DATA FROM STRATA 5 AND 6 FOLLOWING YEAR 10
+survey$setdet <- filter(survey$setdet, !(year > 10 & strat %in% 5:6))
+
+## Run stratified analysis with missing areas
+survey <- survey %>% run_strat()
+
+## Add coordinates to set details
+xy <- as_tibble(survey$grid_xy)
+dat <- as_tibble(survey$setdet) %>%
+  select(x, y, set, year, N = n, tow_area) %>%
+  left_join(., xy, by = c("x", "y")) %>%
+  mutate(offset = log(tow_area))
+
+## Make mesh and fit geostatistical model to survey data
+mesh <- sdmTMB::make_mesh(dat, xy_cols = c("x", "y"), cutoff = 20)
+fit <- sdmTMB(N ~ 0 + as.factor(year) + offset,
+              data = dat,
+              family = nbinom2(link = "log"), spde = mesh,
+              include_spatial = TRUE, time = "year")
+
+## Expand grid data for predictions
+grid_dat <- as_tibble(select(survey$grid_xy, x, y, depth)) %>% distinct()
+grid_dat <- purrr::map_dfr(sort(unique(dat$year)), ~ bind_cols(grid_dat, year = .))
+grid_dat$offset <- mean(dat$offset)
+
+## Predict across full grid and get index
+tow_area <- survey$setdet$tow_area[1] # cell and tow area is consistent in the simulation
+cell_area <- survey$setdet$cell_area[1]
+pred <- predict(fit, newdata = grid_dat, return_tmb_object = TRUE, area = cell_area / tow_area)
+index <- get_index(pred)
+
+## Combine indices
+true_abund <- tibble(year = unique(dat$year), N = as.numeric(colSums(survey$I))) %>% mutate(type = "True")
+strat_abund <- tibble::as_tibble(survey$total_strat) %>%
+  mutate(N = total, type = "Design-based") %>%
+  select(year, N, type)
+index <- index %>%
+  mutate(type = "Model-based", N = est) %>%
+  bind_rows(strat_abund) %>%
+  bind_rows(true_abund)
+
+## Visualize result
+index %>% 
+  plot_ly(x = ~year, color = ~type, legendgroup = ~type) %>% 
+  add_ribbons(ymin = ~lwr, ymax = ~upr, line = list(width = 0), showlegend = FALSE) %>% 
+  add_lines(y = ~N)
+
+```
+
+## Conclusion
+
+- Surveys are expensive and complex and we want to make the most of the data we obtain
+- [SimSurvey](https://github.com/PaulRegular/SimSurvey) offers a means of simulating some of the complexities
+- Design-based analyses are elegant and efficient and work best when everything is standard
+- Problem is that surveys change and this is where model-based approaches may be more powerful
+- [sdmTMB](https://github.com/pbs-assess/sdmTMB) facilitates fast fitting of powerful geostatistical models

### 2022-01-25 R/sim_dist_spde.R
Suggest rgdal since it is required by INLA

@@ -212,10 +212,18 @@ sim_ays_covar_spde <- function(sd = 2.8,range = 300,model = "barrier",phi_age = 
 #'
 #' @export
 make_mesh <- function(grid = make_grid(),max.edge=50,bound.outer=150,cutoff=10,offset=c(max.edge,bound.outer),...){
+
+    for (pkg in c("rgdal", "INLA")) {
+        if (!requireNamespace(pkg, quietly = TRUE)) {
+            stop(paste(pkg, "is needed for make_mesh to work. Please install it."), call. = FALSE)
+        }
+    }
+
     gridPoints <- raster::rasterToPoints(grid)
     locs <- as.matrix(gridPoints[,1:2])
     mesh <- INLA::inla.mesh.2d(locs,offset=offset,max.edge=max.edge,cutoff=cutoff,...)
     mesh
+
 }
 
 

### 2022-01-25 R/plot_sim.R
Fix bug in plotting scripts; plotly returns an error if supplied xtabs class data

@@ -133,6 +133,7 @@ plot_distribution <- function(sim, ages = sim$ages, years = sim$years,
     z <- split_d[[i]]
     attr(z, "class") <- NULL # plotly didn't like the xtabs attributes
     attr(z, "call") <- NULL
+    dimnames(z) <- NULL
     p <- p %>% add_trace(type = type,
                          z = z,
                          visible = i == 1,
@@ -519,6 +520,9 @@ plot_error_surface <- function(sim, plot_by = "rule") {
     vis[[i]] <- c(FALSE, TRUE, TRUE)
     if (plot_by == "rule") {
       z <- stats::xtabs(RMSE ~ ages_cap + lengths_cap, data = split_d[[i]], subset = NULL)
+      dimnames(z) <- NULL # plotly didn't like the xtabs attributes
+      attr(z, "class") <- NULL
+      attr(z, "call") <- NULL
       p <- p %>% add_surface(z = t(z),
                              cmin = min(d$RMSE), cmax = max(d$RMSE),
                              showscale = i == 1,

### 2022-01-25 R/vis_sim.R
Define environment for rendering vis_sim rmarkdown file

@@ -41,18 +41,11 @@ vis_sim <- function(sim, ...) {
       stop(paste(p, "is needed for vis_fit to work. Please install it."), call. = FALSE)
     }
   }
-  if (utils::packageVersion("plotly") > "4.8") {
-    stop('Plots will not render correctly in the dashboard using a version of plotly > 4.8. Plotly 4.8.0 can be instaled using the following line: install.packages("https://cran.r-project.org/src/contrib/Archive/plotly/plotly_4.8.0.tar.gz", repos = NULL, type = "source")')
-  }
 
-  ## save this object to load in vis_sim.Rmd
-  ## (I cannot get rmarkdown::run to work on objects supplied to a function)
-  save(sim, file = file.path(tempdir(), "sim.rda"))
-
-  ## keep most objects created in the rmd file local (i.e. not in the global)
-  local({
-    rmarkdown::run(file = "inst/rmd/vis_sim.Rmd", ...)
-  })
+  rmd_env <- new.env()
+  rmd_env$sim <- sim
+  rmarkdown::run(file = "inst/rmd/vis_sim.Rmd",
+                 render_args = list(envir = rmd_env), ...)
 
 }
 

### 2022-01-26 R/sim_survey.R
Switch to a split-apply-combine approach using base R functions to randomly sample cells (sets); data.table option not working in R devel version

@@ -87,12 +87,22 @@ sim_sets <- function(sim, n_sims = 1, trawl_dim = c(1.5, 0.02),
   cells <- cells[i, ]
   cells$sim <- s
 
-  browser()
+  ## Simulate sets; randomly sample row id by group using split-apply-combine approach
+  ind <- seq(nrow(cells))
+  grp <- paste0(cells$sim, "-", cells$year, "-", cells$strat)
+  siz <- cells$strat_sets
+  split_ind <- split(ind, grp)
+  split_siz <- split(siz, grp)
+  samp_ind <- lapply(seq_along(split_ind), function(i) {
+    sample(split_ind[[i]], size = split_siz[[i]], replace = resample_cells)
+  })
+  samp_ind <- unlist(samp_ind)
 
-  ## Simulate sets; randomly sample data.table row by group
-  ind <- cells[, .I[sample(.N, size = strat_sets, replace = resample_cells)],
-               by = c("sim", "year", "strat")][[4]]
-  sets <- cells[ind, ]
+  # ## Alternative data.table approach; not working under R development version 2022-01-24
+  # samp_ind <- cells[, .I[sample(.N, size = strat_sets, replace = resample_cells)],
+  #                   by = c("sim", "year", "strat")][[4]]
+
+  sets <- cells[samp_ind, ]
   sets[, cell_sets := .N, by = c("sim", "year", "cell")] # useful for identifying cells with more than one set (when resample_units = TRUE)
   sets$set <- seq(nrow(sets))
   sets

### 2022-01-26 R/sim_survey.R
Revert to improved data.table approach for sampling sets; discovered that sample.int requires length(size) == 1

@@ -87,22 +87,10 @@ sim_sets <- function(sim, n_sims = 1, trawl_dim = c(1.5, 0.02),
   cells <- cells[i, ]
   cells$sim <- s
 
-  ## Simulate sets; randomly sample row id by group using split-apply-combine approach
-  ind <- seq(nrow(cells))
-  grp <- paste0(cells$sim, "-", cells$year, "-", cells$strat)
-  siz <- cells$strat_sets
-  split_ind <- split(ind, grp)
-  split_siz <- split(siz, grp)
-  samp_ind <- lapply(seq_along(split_ind), function(i) {
-    sample(split_ind[[i]], size = split_siz[[i]], replace = resample_cells)
-  })
-  samp_ind <- unlist(samp_ind)
-
-  # ## Alternative data.table approach; not working under R development version 2022-01-24
-  # samp_ind <- cells[, .I[sample(.N, size = strat_sets, replace = resample_cells)],
-  #                   by = c("sim", "year", "strat")][[4]]
-
-  sets <- cells[samp_ind, ]
+  ## Simulate sets; randomly sample row id by group
+  ind <- cells[, .I[sample(.N, size = unique(strat_sets), replace = resample_cells)],
+                    by = c("sim", "year", "strat")][[4]]
+  sets <- cells[ind, ]
   sets[, cell_sets := .N, by = c("sim", "year", "cell")] # useful for identifying cells with more than one set (when resample_units = TRUE)
   sets$set <- seq(nrow(sets))
   sets

### 2022-01-26 R/zzz.R
Drop zzz.R as it is not really needed

@@ -1,15 +1,0 @@
-
-# pkgdown::build_site(examples = FALSE)
-# devtools::check(env_vars = c(NOT_CRAN = "true", R_CHECK_DONTTEST_EXAMPLES = "false"))
-# devtools::check_win_devel(email = "Paul.Regular@dfo-mpo.gc.ca")
-# cran_check <- rhub::check_for_cran(email = "paul.regular@gmail.com", env_vars = c(`_R_CHECK_DONTTEST_EXAMPLES_` = "false", `_R_CHECK_FORCE_SUGGESTS_` = "false", `_R_CHECK_CRAN_INCOMING_USE_ASPELL_` = "true"))
-# # cran_check <- rhub::get_check(c("SimSurvey_0.1.3.tar.gz-5c5c6c6d39094fb69979eff456c8af95", "SimSurvey_0.1.3.tar.gz-d8ce1e5cb7c54a5599a472a18022dfc3", "SimSurvey_0.1.3.tar.gz-a2249a8410f947d590bfe474b61e6f98"))
-# cran_check$cran_summary()
-# devtools::release()
-
-release_questions <- function() {
-  c(
-    "Did you update the pkgdown site?",
-    "Did you update cran-coments?"
-  )
-}

### 2022-01-26 tutorial/tutorial.Rmd
Improve index predictions from sdmTMB in tutorial

@@
 grid_dat <- purrr::map_dfr(sort(unique(dat$year)), ~ bind_cols(grid_dat, year = .))
-grid_dat$offset <- mean(dat$offset)
+grid_dat$offset <- 0
 
 ## Predict across full grid and get index
-tow_area <- survey$setdet$tow_area[1] # cell and tow area is consistent in the simulation
 cell_area <- survey$setdet$cell_area[1]
-pred <- predict(fit, newdata = grid_dat, return_tmb_object = TRUE, area = cell_area / tow_area)
-index <- get_index(pred)
+pred <- predict(fit, newdata = grid_dat, sims = 1000)
+index <- get_index_sims(pred, area = rep(cell_area, nrow(pred)))
 
@@
   add_ribbons(ymin = ~lwr, ymax = ~upr, line = list(width = 0), showlegend = FALSE) %>% 
-  add_lines(y = ~N)
+  add_lines(y = ~N, line = list(dash = ~ifelse(type == "True", "solid", "dot")))
 
@@
 
-## DROP DATA FROM STRATA 5 AND 6 FOLLOWING YEAR 10
+## DROP DATA FROM STRATA 5, 6 FOLLOWING YEAR 10
 survey$setdet <- filter(survey$setdet, !(year > 10 & strat %in% 5:6))
@@
   add_ribbons(ymin = ~lwr, ymax = ~upr, line = list(width = 0), showlegend = FALSE) %>% 
-  add_lines(y = ~N)
+  add_lines(y = ~N, line = list(dash = ~ifelse(type == "True", "solid", "dot"))) %>% 
+  add_segments(x = 10, xend = 10, y = 0, yend = max(index$upr, na.rm = TRUE) * 1.05, 
+               inherit = FALSE, showlegend = FALSE, hoverinfo = "none", color = I("black"),
+               size = I(0.5), linetype = I(2))
 

### 2025-01-31 tutorial/tutorial.Rmd
Minor tweak of tutorial

@@
 mesh <- sdmTMB::make_mesh(dat, xy_cols = c("x", "y"), cutoff = 20)
-fit <- sdmTMB(N ~ 0 + as.factor(year) + offset,
+fit <- sdmTMB(N ~ 0 + as.factor(year),
+              offset = dat$offset,
+              mesh = mesh,
               data = dat,
-              family = nbinom2(link = "log"), spde = mesh,
-              include_spatial = TRUE, time = "year")
+              family = nbinom2(link = "log"), 
+              time = "year")
 
@@
 mesh <- sdmTMB::make_mesh(dat, xy_cols = c("x", "y"), cutoff = 20)
-fit <- sdmTMB(N ~ 0 + as.factor(year) + offset,
+fit <- sdmTMB(N ~ 0 + as.factor(year),
+              offset = dat$offset,
+              mesh = mesh,
               data = dat,
-              family = nbinom2(link = "log"), spde = mesh,
-              include_spatial = TRUE, time = "year")
+              family = nbinom2(link = "log"),
+              time = "year")
 
